### 2018-02-19 manuscript.Rnw
added knitr file for manuscript

% run this from the command line to compile: Rscript -e "library(knitr); knit('./manuscript.Rnw')"
\documentclass{article}

\usepackage[margin=1in]{geometry}

\input{newphyt_style}

\begin{document}

\title{Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns}

\author{Joel H. Nitta}
%\author[2]{James E. Watkins, Jr.}
%\author[3]{Ravahere Taputuarai}
%\author[1]{Charles C. Davis}

%\affil[1]{Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, Cambridge, MA 02138, U.S.A.}
%\affil[2]{Department of Biology, Colgate University, 13 Oak Drive, Hamilton, New York 13346, U.S.A.}
%\affil[3]{Association Te Rau Ati Ati a Tau a Hiti Noa Tu, B.P. 11553, Mahina, Tahiti, French Polynesia}

\maketitle

% setup knitr
<<setup, echo=FALSE>>=
library(knitr)
# set global chunk options
opts_chunk$set(echo=FALSE)
opts_chunk$set(message=FALSE)
opts_chunk$set(warning=FALSE)
# set working directory
setwd(here::here())
@

% run analyses silently so results don't print to screen
<<run_analyses, results="hide">>=
source(file = "bin/phylosig.R")
source(file = "bin/microclimate.R")
source(file = "bin/binaryPGLMM.R")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Summary %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

% Set left justification (better conversion to Word than default full justification)
\raggedright

% Set indent to about 3 characters throughout
\setlength{\parindent}{3ex}

\section*{Summary}

\begin{itemize}
\item The appearance of angiosperm-dominated forests in the Cretaceous and early Cenozoic had a profound effect on terrestrial biota by creating novel ecological niches. The diversity of modern fern lineages are thought to be a response to angiosperm dominance, radiating both in the canopy and on the forest floor. Despite the significance of epiphytic growth for fern evolution, however, few studies have investigated traits related to epiphytism in ferns while accounting for phylogeny.
\item We investigated evolutionary patterns in traits related to epiphytic growth at the species and community level using a phylogeny including representatives of multiple epiphytic fern radiations on the island of Moorea, French Polynesia. 
\item We observed phylogenetic conservatism in multiple traits related to epiphytic growth in fern sporophytes and gametophytes. Epiphytes had significantly shorter stipes, smaller rhizomes, and a higher frequency of non-cordate gametophytes while accounting for phylogeny. Epiphytic communities were clustered both phylogenetically and functionally, consistent with the hypothesis that the canopy acts as a filter due to its low humidity. 
\item Our results provide an important phylogenetic perspective on the role of functional traits in ferns, and pave the way for more detailed studies examining how niche space is filled within terrestrial and epiphytic habitats.
\end{itemize}

\textbf{Key words:} community ecology, functional traits, gametophyte, Moorea, Society Islands, Tahiti.

%%%%%%%%%%%%%%%%%%%%%%%%%%% Tables %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage

<<setup_tables>>=
library(xtable)
# load latex functions
source("bin/latex_functions.R")
@

<<print_phylosig, results="asis">>=
phylosig.results[sapply(phylosig.results, class) == "numeric" ] <- sapply(phylosig.results[sapply(phylosig.results, class) == "numeric" ], round, 3)
phylosig.results <- embolden_p(phylosig.results, col.select=c("k.pval","lambda.pval", "prob_random", "prob_brownian"))

print(xtable(phylosig.results,
             label="tab:phylosig",
             caption="Results of phylogenetic signal analysis for growth habit and related traits. Values of $\\lambda$ or $\\K >= 1$ indicate that traits are evolving according to Brownian Motion (BM); values of $\\lambda$ or \\K near 0 indicate random distribution of traits. For the \\D statistic, values of 0 or less indicate evolution according to BM; values near 1 indicate random trait distribution."),
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      sanitize.text.function = function (x) {x},
      include.rownames=TRUE,
      caption.placement="top")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%% Figures %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage

\begin{figure}[ht]
  \includegraphics[scale=1]{microclimate.pdf}
  \caption{Climate}
  \label{fig:climate}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%% Appendix Tables %%%%%%%%%%%%%%%%%%%%%%
% Add S to table number and reset counter
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

<<climate_ancova, results="asis">>=
ancova.results[sapply(ancova.results, class) == "numeric" ] <- sapply(ancova.results[sapply(ancova.results, class) == "numeric" ], round, 3)
ancova.results <- embolden_p(ancova.results, col.select=c("pval.intercept","pval.slope"))
print(xtable(ancova.results,
             label="apptab:climate_model",
             caption="Results of analysis of covariance (ANCOVA) for microclimatic variables between epiphytic and terrestrial plots along an elevational gradient on Moorea, French Polynesia. Linear models were constructed for each response variable with growth habit as a fixed effect and elevation as a co-factor. Differences in intercept between growth habits was tested with models of the form $y \\sim growth habit + elevation$. Differences in slope between growth habits were tested with models of the form $y \\sim growth habit \\times elevation$. Models were compared using the `aov' function in R."),
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      sanitize.text.function = function (x) {x},
      include.rownames=TRUE,
      caption.placement="top")
@

<<climate_slopes, results="asis">>=
slopes.results[sapply(slopes.results, class) == "numeric" ] <- sapply(slopes.results[sapply(slopes.results, class) == "numeric" ], round, 3)
print(xtable(slopes.results,
             caption=NULL),
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      include.rownames=TRUE)
@

<<binaryPGLMM, results="asis">>=
bpglmm.results <- as.data.frame(sapply(bpglmm.results, round, 3))
bpglmm.results <- embolden_p(bpglmm.results, col.select=c("sigmap","pvalue"))
print(xtable(bpglmm.results,
             label="apptab:PGLMM_gameto",
             caption="Phylogenetic generalized linear mixed models (PGLMMs) for binary (gametophyte) traits traits related to epiphytic growth. The response of each trait was modeled against growth habit using a PGLMM with phylogeny as a random effect. $\\sigma^2$ is a measure of phylogenetic signal calculated from the variance in the PGLMM. Significance of $\\sigma^2$ calculated using an approximate likelihood ratio test between a model with the observed value of $\\sigma^2$ against a null model with $\\sigma^2 = 0$. Traits significantly correlated with growth habit in bold."),
      sanitize.text.function = function (x) {x},
      sanitize.colnames.function = custom_cols,
      include.rownames=TRUE,
      caption.placement="top")
@

\end{document}

### 2018-02-20 manuscript.Rnw
Mostly finalize table formatting
Add function to latex_functions to add part label to tables. Add some packages to manuscript preamble, body text, and Table 1 (static text). Get tables formatted basically how I want them: caption is on top. First sentence of caption should be short title, followed by description (to be converted to footnote in final MS).

@@
 
+% Set margins
 \usepackage[margin=1in]{geometry}
 
+% Set line spacing to 1.5
+\usepackage{setspace}
+\onehalfspace
+
+% Enable bibliography
+\usepackage{natbib}
+
+% Set unit and abbreviations formatting
 \input{newphyt_style}
@@
 
-% setup knitr
 <<setup, echo=FALSE>>=
+# setup knitr
 library(knitr)
-# set global chunk options
 opts_chunk$set(echo=FALSE)
@@
 opts_chunk$set(warning=FALSE)
-# set working directory
-setwd(here::here())
 @
 
-% run analyses silently so results don't print to screen
 <<run_analyses, results="hide">>=
-source(file = "bin/phylosig.R")
+# set working directory, run all analyses
+setwd(here::here())
+source(file = "bin/cont_phylosig.R")
+source(file = "bin/binary_phylosig.R")
 source(file = "bin/microclimate.R")
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
+
+\section*{Introduction}
+
+The widespread appearance of angiosperm-dominated forests in the late Cretaceous and early Cenozoic had an immense effect on terrestrial ecosystems \citep{Niklas1983, Kenrick1997a,Lloyd2008}.
+The heterogeneous niche space of the angiosperm canopy comprising stratified light and humidity levels and a range of substrates for growth became rapidly filled by a variety of epiphytic plants, including mosses \citep{Fiz-Palacios2011}, leafy liverworts \citep{Feldberg2014}, orchids \citep{Givnish2015b}, bromeliads \citep{Givnish2014a}, and ferns \citep{Schneider2004d}.
+Diversification rates in some of these groups may have further increased during the Paleocene / Eocene thermal maximum, a period of globally warmer and wetter climates that may have enabled movement into the canopy \citep{Schuettpelz2009,Testo2016a}.
+The transition to epiphytic growth has been particularly important for fern diversity, with ca.
+29\% of all species exhibiting this growth form relative to ca.
+10\% across all vascular plants \citep{Kress1986}.
+Major transitions to epiphytism occurred at least five times during the evolution of ferns, each followed by diversification during the late Cretaceous or early Cenozoic \citep{Schuettpelz2009,WatkinsJr.2012}.
+Recent studies \citep{Sundue2015,Testo2016a,Lehtonen2017} have failed to find evidence of rapid diversification characteristic of adaptive radiations associated with epiphytic growth in ferns, but have confirmed gradually increasing rates of diversification in Polypodiaceae \citep{Sundue2015} and generally support the significance of epiphytic growth as a key innovation for extant fern diversity (\citealt{Testo2016a,WatkinsJr.2012}; but see \citealt{Lehtonen2017}).
+
+Despite the ecological opportunities afforded by life in the canopy, including escape from disturbance \citep{Page2002} and abundant light for photosynthesis \citep{Benzing2000}, epiphytes must contend with greater abiotic stress in the form of increased evaporation and more extreme oscillations in temperature relative to terrestrial plants \citep{Zotz2001,Lowman2012}.
+Associated with this change, epiphytic ferns have evolved various morphological adaptations that minimize water loss from their leaves, including shorter stipes \citep{WatkinsJr.2010}, thicker leaves \citep{Kluge2007,WatkinsJr.2007b}, and protection with scales or hairs \citep{WatkinsJr.2006b}.
+Some lineages have specialized adaptations to epiphytic growth, such as humus-collection (\eg, \textit{Asplenium}; \citealt{Benzing1990}, \textit{Drynaria}; \citealt{Janssen2005}), vegetative desiccation tolerance (\eg, Hymenophyllaceae; \citealt{Shreve1911}, \textit{Pleopeltis}; \citealt{Pessin1925}), CAM photosynthesis \citep[\eg, \textit{Pyrrosia},][]{Wong1976}, and symbiotic relationships with ants (\eg, \textit{Antrophyum}; \citealt{WatkinsJr.2008}, \textit{Lecanopteris}; \citealt{Gay1993a}).
+Many of these same adaptations are found in epiphytic seed plants.
+However, the evolution of epiphytisim in ferns was much more complex given that ferns rely on a free-living gametophyte.
+The fern gametophyte is anatomically simple, and differs from the sporophyte markedly in anatomy, morphology, and physiology.
+Such differences create a system of “dual ecology” where selective pressures acting on one stage are very different than those acting on another.
+For example, desiccation tolerance, a key trait for non-vascular plants living in water-limited environments, is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is quite rare in sporophytes \citep{WatkinsJr.2007}.
+Despite the evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no studies to our knowledge have analyzed the evolution of such traits across the fern evolutionary tree including both life stages and representatives of all five Cretaceous epiphytic radiations \sensu \cite{Schuettpelz2009}.
+
+The majority of studies investigating adaptations to epiphytism have utilized a univariate, cross-species approach, analyzing traits individually between terrestrial and epiphytic species \citep[\eg,][]{WatkinsJr.2007b,WatkinsJr.2010}.
+However, given the evolutionary history of ferns with multiple radiations within the epiphytic niche \citep{Schuettpelz2009}, it is likely that traits related to epiphytism carry at least some degree of phylogenetic signal (\ie, the tendency for closely related species to resemble each other).
+Therefore, a comparative approach incorporating phylogenetic relationships (and if possible, branch lengths) is appropriate \citep{Felsenstein1985}.
+However, such approaches have only recently been applied to studies of epiphytism in ferns, and at relatively limited phylogenetic scales \citep[\eg][]{Zhang2014b,Sundue2015}.
+
+Furthermore, it should be possible to gain greater insight into the selective strength exerted by the epiphytic habitat by utilizing a community-based approach.
+If the relatively harsh abiotic conditions of the canopy act as a filter, we would expect to find lower functional richness at a community level, which should also result in decreased phylogenetic diversity under a scenario of phylogenetic trait conservatism \citep{Webb2002}.
+However, the relative strength of this environmental filter may also depend on overall abiotic conditions; for example, relative humidity and temperature are expected to vary not only between epiphytic and terrestrial environments, but also with elevation \citep{Cardelus2006}.
+Elevation is a strong predictor of fern species richness, with maximal richness tending to occur at mid-elevations, especially in the tropics \citep{Kessler2011a}.
+Functional diversity is also correlated with species richness (as few species tend to have exactly the same functional traits; \citealt{Schleuter2010}).
+Thus, environmental variation along elevational gradients is expected to impact the functional, phylogenetic, and species diversity of epiphytic and terrestrial fern communities.
+
+Here, we investigate the evolution and role in community assembly of functional traits related to epiphytic growth in fern gametophytes and sporophytes on Moorea, French Polynesia.
+Moorea is an isolated tropical oceanic island with a high phylogenetic diversity of ferns distributed over a steep elevational gradient, making it ideal for a community phylogenetic analysis of functional traits.
+Specifically, we test the following hypotheses related to epiphytic growth in ferns:
+
+\begin{enumerate}
+\item The epiphytic environment experiences greater extremes in temperature and humidity than the terrestrial environment, thus acting as a filter on the traits of plants that grow there.
+\item Epiphytic ferns have converged upon similar morphologies multiple times in both sporophyte and gametophyte phases.
+\item Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities due to environmental filtering, and the strength of this filtering varies with changes in environment.
+\end{enumerate}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% Acknowledgements %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section*{Acknowledgements}
+
+We thank members of the Davis Lab, N.
+Michele Holbrook, and Jonathan Losos for helpful discussion and comments on drafts.
+Suzanne Vinette, Saad Amer, Tristan Wang, and Tohei Theophilus helped with community surveys and measurement of traits.
+Staff at the University of California, Berkeley Richard B.
+Gump South Pacific Research Station, Moorea, French Polynesia, in particular Valentine Brotherson, Hinano Teavai-Murphy, Frank Murphy, and Neil Davies, provided support with obtaining research and collection permits, laboratory space, and logistics.
+Funding provided in part by National Science Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to JHN and CCD), Setup Funds from Harvard University to CCD, the American Society of Plant Taxonomists (Research Grant for Graduate Students to JHN), the Garden Club of America (Award in Tropical Botany to JHN), the Harvard University Herbaria (Fernald Fieldwork Fellowship to JHN), the Society of Systematic Biologists (Graduate Student Research Award to JHN), and the Systematics Association (Systematics Research Fund to JHN).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% Author Contribution Statement %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Author Contributions}
+
+J.H.N. conceived and carried out the study. R.T. collected data. J.H.N. performed the analyses. J.H.N., J.E.W., and C.C.D. wrote the manuscript.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% set style
+\bibliographystyle{newphyt}
+
+% change the bibliography title
+\renewcommand\refname{References}
+
+% insert bibliography
+\bibliography{references}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%% Tables %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
 
@@
 library(xtable)
-# load latex functions
 source("bin/latex_functions.R")
@@
 
-<<print_phylosig, results="asis">>=
-phylosig.results[sapply(phylosig.results, class) == "numeric" ] <- sapply(phylosig.results[sapply(phylosig.results, class) == "numeric" ], round, 3)
-phylosig.results <- embolden_p(phylosig.results, col.select=c("k.pval","lambda.pval", "prob_random", "prob_brownian"))
+% Table 1
+\clearpage
 
-print(xtable(phylosig.results,
-             label="tab:phylosig",
-             caption="Results of phylogenetic signal analysis for growth habit and related traits.
+\begin{table}[ht]
+\centering
+\caption{Traits used in this study. Noncordate gametophyte morphotypes include ribbon, strap and filamentous.}
+\begin{tabular}{ p{0.3\textwidth} p{0.2\textwidth} p{0.4\textwidth} }
+\hline
+ & Data type (unit) & Ecological Significance \\
+\hline
+Sporophyte Traits & & \\
+\hspace{3ex}Stipe Length  & Continuous (\si{\centi\meter})  & Shorter stipes compensate for low conductivity \citep{WatkinsJr.2010} \\
+\hspace{3ex}Frond Length  & Continuous (\si{\centi\meter})  & Smaller leaf size reduces evapotranspiration \citep{Vogel1968} \\
+\hspace{3ex}Frond Width  & Continuous (\si{\centi\meter})  & Smaller leaf size reduces evapotranspiration \citep{Vogel1968} \\
+\hspace{3ex}Rhizome Diameter  & Continuous (\si{\centi\meter})  & Scales with overall plant size (Creese et al. 2011) \\
+\hspace{3ex}Frond Dissection  & Ordinal  & Simple laminae reduce evapotranspiration (Kluge and Kessler 2007) \\
+\hspace{3ex}Pinna number & Integer & Scales with overall plant size (Creese et al. 2011) \\
+\hspace{3ex}Specific Leaf Area & Continuous (\si{\square\meter\per\kilo\gram}) & Species in more stressful environments invest in thicker leaves \citep{Wright2004a} \\
+Gametophyte Traits & & \\
+\hspace{3ex}Morphotype & Binary (cordate vs. noncordate) & Noncordate morphology reduces drying rates by holding external water in crevices and folds \citep{WatkinsJr.2007,Pittermann2013} \\
+\hspace{3ex}Gemmae & Binary (present/absent) & Enable asexual reproduction, thereby overcoming local extinction due to drying \citep{Farrar2008} \\
+\hspace{3ex}Glands & Binary (present/absent) & Lipophilic exudates produced by glands may reduce transpiration rates \citep{Wollenweber2000} \\
+\hspace{3ex}Hairs & Binary (present/absent) & Hairs reduce drying rates by holding external water and increasing boundary layer thickness \citep{WatkinsJr.2007} \\
+\hline
+\end{tabular}
+\label{table:traits}
+\end{table}
+
+% Table 2
+\clearpage
+
+<<cont_phylosig, results="asis">>=
+cont_phylosig.results <- round_df(cont_phylosig.results)
+cont_phylosig.results <- embolden_p(cont_phylosig.results, col.select=c("k.pval","lambda.pval"))
+part_label <- add_part_label(cont_phylosig.results, "(a)")
+
+print(xtable(cont_phylosig.results,
+             label="table:phylosig",
+             caption="Phylogenetic signal in traits.
+(a) Continuous traits and (b) binary traits.
 Values of $\\lambda$ or $\\K >= 1$ indicate that traits are evolving according to Brownian Motion (BM); values of $\\lambda$ or \\K near 0 indicate random distribution of traits.
@@
       include.rownames=TRUE,
+      add.to.row = part_label,
       caption.placement="top")
+@
+
+<<binary_phylosig, results="asis">>=
+binary_phylosig.results <- round_df(binary_phylosig.results)
+binary_phylosig.results <- embolden_p(binary_phylosig.results, col.select=c("prob_random","prob_browian"))
+part_label <- add_part_label(binary_phylosig.results, "(b)")
+
+print(xtable(binary_phylosig.results,
+             caption=NULL),
+      sanitize.colnames.function = custom_cols,
+      sanitize.rownames.function = custom_rows,
+      sanitize.text.function = function (x) {x},
+      include.rownames=TRUE,      
+      add.to.row = part_label)
 @
@@
 
+% Table S1
+\clearpage
 <<climate_ancova, results="asis">>=
-ancova.results[sapply(ancova.results, class) == "numeric" ] <- sapply(ancova.results[sapply(ancova.results, class) == "numeric" ], round, 3)
+ancova.results <- round_df(ancova.results)
 ancova.results <- embolden_p(ancova.results, col.select=c("pval.intercept","pval.slope"))
+part_label <- add_part_label(ancova.results , "(a)")
+
 print(xtable(ancova.results,
-             label="apptab:climate_model",
+             label="supp_table:climate_model",
              caption="Results of analysis of covariance (ANCOVA) for microclimatic variables between epiphytic and terrestrial plots along an elevational gradient on Moorea, French Polynesia.
+(a) Model statistics and (b) coefficients.
 Linear models were constructed for each response variable with growth habit as a fixed effect and elevation as a co-factor.
@@
       include.rownames=TRUE,
+      add.to.row = part_label,
       caption.placement="top")
@@
 <<climate_slopes, results="asis">>=
-slopes.results[sapply(slopes.results, class) == "numeric" ] <- sapply(slopes.results[sapply(slopes.results, class) == "numeric" ], round, 3)
+slopes.results <- round_df(slopes.results)
+part_label <- add_part_label(slopes.results , "(b)")
+
 print(xtable(slopes.results,
@@
       sanitize.rownames.function = custom_rows,
+      add.to.row = part_label,
       include.rownames=TRUE)
@@
 
+% Table S2
+\clearpage
 <<binaryPGLMM, results="asis">>=
-bpglmm.results <- as.data.frame(sapply(bpglmm.results, round, 3))
+bpglmm.results <- round_df(bpglmm.results)
 bpglmm.results <- embolden_p(bpglmm.results, col.select=c("sigmap","pvalue"))
 print(xtable(bpglmm.results,
-             label="apptab:PGLMM_gameto",
+             label="supp_table:PGLMM_gameto",
              caption="Phylogenetic generalized linear mixed models (PGLMMs) for binary (gametophyte) traits traits related to epiphytic growth.

### 2018-02-20 bin/latex_functions.R

@@ -2,6 +2,21 @@
 # various functions to assist with latex formatting when compiling 
 # manuscript.tex and manuscript.pdf from manuscript.Rnw
 
+# add_part_label
+# add a table part label on top of a table (as an extra row)
+# df is dataframe to be printed using xtable
+# labeltext is text to be printed
+add_part_label <- function (df, labeltext) {
+
+  part_label <- list ()
+  part_label$pos <- list()
+  part_label$pos[[1]] <- c(-1) # position label at top of table
+  part_label$command <- c(paste0("\\multicolumn{", ncol(df)+1, "}{l}{", labeltext, "} \\\\ \n"))
+  
+  return(part_label)
+  
+}
+
 # round_df
 # round the numbers in a dataframe to the specified number of digits
 round_df <- function (df, digits=3) {
@@ -60,7 +75,7 @@ custom_cols <- function (cols) {
 custom_rows <- function (rows) {
   
   # traits
-  rows <- gsub ("habit", "Growth habit", rows)
+  # continuous
   rows <- gsub ("stipe", "Stipe length", rows)
   rows <- gsub ("^length$", "Frond length", rows)
   rows <- gsub ("width", "Frond width", rows)
@@ -68,10 +83,13 @@ custom_rows <- function (rows) {
   rows <- gsub ("pinna", "Pinna number", rows)
   rows <- gsub ("sla", "Specific leaf area", rows)
   rows <- gsub ("rhizome", "Rhizome \\\\diameter", rows)
+  # binary
+  rows <- gsub ("habit", "Growth habit", rows)
+  rows <- gsub ("epiphytic", "Epiphytic growth", rows)
   rows <- gsub ("gemmae", "Gemmae", rows)
   rows <- gsub ("glands", "Glands", rows)
   rows <- gsub ("hairs", "Hairs", rows)
-  rows <- gsub ("morph_binary", "Morphotype", rows)
+  rows <- gsub ("cordate_morph", "Cordate morphotype", rows)
   
   # microclimate
   rows <- gsub ("max_temp", "Max. temperature", rows)
@@ -85,7 +103,6 @@ custom_rows <- function (rows) {
   
   return(rows)
 }
-
 
 # embolden_p
 # makes p-values less than chosen significance level bold in results table

### 2018-03-01 manuscript.Rnw
Fixed typo in table caption
Separate out SI into its own knitr file

@@
 <<run_analyses, results="hide">>=
-# set working directory, run all analyses
+# set working directory, run all analyses, save results objects
 setwd(here::here())
@@
 source(file = "bin/binaryPGLMM.R")
+source(file = "bin/trait_PCA.R")
+save(list = ls(all = TRUE), file= "results.RData")
 @
@@
 Values of $\\lambda$ or $\\K >= 1$ indicate that traits are evolving according to Brownian Motion (BM); values of $\\lambda$ or \\K near 0 indicate random distribution of traits.
-For the \\D statistic, values of 0 or less indicate evolution according to BM; values near 1 indicate random trait distribution."),
+For \\D, values of 0 (or less) indicate evolution according to (or more conserved than) BM; values near 1 indicate random trait distribution.
+\\pval\\textsubscript{Rnd}, result of testing that observed \\D is significantly different from 1; \\pval\\textsubscript{BM}, result of testing that observed \\D is significantly different from 0."),
       sanitize.colnames.function = custom_cols,
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%% Appendix Tables %%%%%%%%%%%%%%%%%%%%%%
-% Add S to table number and reset counter
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{table}{0}
-
-% Table S1
-\clearpage
-<<climate_ancova, results="asis">>=
-ancova.results <- round_df(ancova.results)
-ancova.results <- embolden_p(ancova.results, col.select=c("pval.intercept","pval.slope"))
-part_label <- add_part_label(ancova.results , "(a)")
-
-print(xtable(ancova.results,
-             label="supp_table:climate_model",
-             caption="Results of analysis of covariance (ANCOVA) for microclimatic variables between epiphytic and terrestrial plots along an elevational gradient on Moorea, French Polynesia.
-(a) Model statistics and (b) coefficients.
-Linear models were constructed for each response variable with growth habit as a fixed effect and elevation as a co-factor.
-Differences in intercept between growth habits was tested with models of the form $y \\sim growth habit + elevation$.
-Differences in slope between growth habits were tested with models of the form $y \\sim growth habit \\times elevation$.
-Models were compared using the `aov' function in R."),
-      sanitize.colnames.function = custom_cols,
-      sanitize.rownames.function = custom_rows,
-      sanitize.text.function = function (x) {x},
-      include.rownames=TRUE,
-      add.to.row = part_label,
-      caption.placement="top")
-@
-
-<<climate_slopes, results="asis">>=
-slopes.results <- round_df(slopes.results)
-part_label <- add_part_label(slopes.results , "(b)")
-
-print(xtable(slopes.results,
-             caption=NULL),
-      sanitize.colnames.function = custom_cols,
-      sanitize.rownames.function = custom_rows,
-      add.to.row = part_label,
-      include.rownames=TRUE)
-@
-
-% Table S2
-\clearpage
-<<binaryPGLMM, results="asis">>=
-bpglmm.results <- round_df(bpglmm.results)
-bpglmm.results <- embolden_p(bpglmm.results, col.select=c("sigmap","pvalue"))
-print(xtable(bpglmm.results,
-             label="supp_table:PGLMM_gameto",
-             caption="Phylogenetic generalized linear mixed models (PGLMMs) for binary (gametophyte) traits traits related to epiphytic growth.
-The response of each trait was modeled against growth habit using a PGLMM with phylogeny as a random effect.
-$\\sigma^2$ is a measure of phylogenetic signal calculated from the variance in the PGLMM.
-Significance of $\\sigma^2$ calculated using an approximate likelihood ratio test between a model with the observed value of $\\sigma^2$ against a null model with $\\sigma^2 = 0$.
-Traits significantly correlated with growth habit in bold."),
-      sanitize.text.function = function (x) {x},
-      sanitize.colnames.function = custom_cols,
-      include.rownames=TRUE,
-      caption.placement="top")
-@
-
 \end{document}

### 2018-03-01 supporting_information.Rnw
Separate out SI into its own knitr file

\documentclass{article}
\date{}

% Set margins
\usepackage[margin=1in]{geometry}

% Set line spacing to 1.5
\usepackage{setspace}
\onehalfspace

% Enable bibliography
\usepackage{natbib}

% Set unit and abbreviations formatting
\input{newphyt_style}

\begin{document}

<<setup, echo=FALSE>>=
# setup knitr
library(knitr)
opts_chunk$set(echo=FALSE)
opts_chunk$set(message=FALSE)
opts_chunk$set(warning=FALSE)
@

<<setup_tables>>=
library(xtable)
source("bin/latex_functions.R")
# load results objects produced by running manuscript.Rnw
load("results.RData")
@

\noindent \textbf{\textit{New Phytologist} Supporting Information} \\
\\
\noindent Article title: Life in the canopy: phylogenetic filtering and morphological canalization in epiphytic ferns \\
\noindent Authors: Joel H. Nitta, James E. Watkins, Jr., Ravahere Taputuarai and Charles C. Davis \\
\noindent Article acceptance date: someday soon \\
\\
\noindent The following Supporting Information is available for this article: \\

\noindent \textbf{Methods S1} We performed alternative tests for correlation of growth habit with functional traits using phylogenetic generalized linear mixed models (PGLMMs). Models were constructed with each trait as the response variable in turn dependent on growth habit, with phylogeny as a random effect. For binary (gametophyte) traits we used the \function{binaryGLMM} function in the \package{ape} package \citep{Paradis2004}. For quantitative (sporophyte) traits, we used the \function{MCMCglmm} function in the R package of the same name \citep{Hadfield2010}. Priors were set to an inverse-Gamma distribution with shape and scale parameters equal to \num{0.01} \citep{Hadfield2010a}. The MCMC algorithm was run for \num{5000000} iterations with a burnin of \num{1000} and thinning every \num{500} iterations.

%%%%%%%%%%%%%%%%%%%%%%%%%%% Tables %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Add S to table number and reset counter
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

% Table S1 (supp_table:PGLMM_sporo)
<<PGLMM, results="asis">>=
# read in results from running PGLMM on cluster, clean up dataframe
pglmm.results <- read.csv("data/mcmcglmm_habit_all_results.csv")
pglmm.results <- pglmm.results[pglmm.results$Effect == "habit", ]
pglmm.results$Effect <- NULL
pglmm.results$Model <- gsub(" ~ habit", "", pglmm.results$Model)
pglmm.results$Model <- tolower(pglmm.results$Model)
rownames(pglmm.results) <- pglmm.results$Model
pglmm.results$Model <- NULL

# add latex formatting
pglmm.results <- round_df(pglmm.results)
pglmm.results <- embolden_p(pglmm.results, col.select="P.value")

print(xtable(pglmm.results,
             label="supp_table:PGLMM_sporo",
             caption="Phylogenetic generalized linear mixed models (PGLMMs) for quantitative (sporophyte) traits related to epiphytic growth. The response of each trait was modeled against growth habit using a PGLMM with phylogeny as a random effect. \\pval for traits significantly correlated with growth habit in bold."),
      sanitize.text.function = function (x) {x},
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      include.rownames=TRUE,
      caption.placement="top")

@

% Table S2
\clearpage
<<binaryPGLMM, results="asis">>=
bpglmm.results <- round_df(bpglmm.results)
bpglmm.results <- embolden_p(bpglmm.results, col.select=c("sigmap","pvalue"))

print(xtable(bpglmm.results,
             label="supp_table:PGLMM_gameto",
             caption="Phylogenetic generalized linear mixed models (PGLMMs) for binary (gametophyte) traits traits related to epiphytic growth. The response of each trait was modeled against growth habit using a PGLMM with phylogeny as a random effect. $\\sigma^2$ is a measure of phylogenetic signal calculated from the variance in the PGLMM. Significance of $\\sigma^2$ calculated using an approximate likelihood ratio test between a model with the observed value of $\\sigma^2$ against a null model with $\\sigma^2 = 0$. Statistically significant values in bold."),
      sanitize.text.function = function (x) {x},
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      include.rownames=TRUE,
      caption.placement="top")
@

% Table S3
\clearpage
<<climate_ancova, results="asis">>=
ancova.results <- round_df(ancova.results)
ancova.results <- embolden_p(ancova.results, col.select=c("pval.intercept","pval.slope"))
part_label <- add_part_label(ancova.results , "(a)")

print(xtable(ancova.results,
             label="supp_table:climate_model",
             caption="Results of analysis of covariance (ANCOVA) for microclimatic variables between epiphytic and terrestrial plots along an elevational gradient on Moorea, French Polynesia. (a) Model statistics and (b) estimates. Linear models were constructed for each response variable with growth habit as a fixed effect and elevation as a co-factor. Differences in intercept between growth habits was tested with models of the form $y \\sim growth habit + elevation$. Differences in slope between growth habits were tested with models of the form $y \\sim growth habit \\times elevation$. Models were compared using the `aov' function in R."),
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      sanitize.text.function = function (x) {x},
      include.rownames=TRUE,
      add.to.row = part_label,
      caption.placement="top")
@

<<climate_slopes, results="asis">>=
slopes.results <- round_df(slopes.results)
part_label <- add_part_label(slopes.results , "(b)")

print(xtable(slopes.results,
             caption=NULL),
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      add.to.row = part_label,
      include.rownames=TRUE)
@

% Table S4
\clearpage
<<trait_pca, results="asis">>=
PCA.results <- round_df(PCA.results)
PCA.results <- embolden_top_values(PCA.results, row.select = c("dissection", "length", "pinna", "rhizome", "sla", "stipe", "width"))

print(xtable(PCA.results,
             label="supp_table:PCA",
             caption="PC loadings for principal components analyses (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia. For each PC, values of the three traits with the greatest loadings (absolute values) shown in bold."),
      sanitize.text.function = function (x) {x},
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      include.rownames=TRUE,
      caption.placement="top")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% set style
\bibliographystyle{newphyt}

% change the bibliography title
\renewcommand\refname{References}

% insert bibliography
\bibliography{references}

\end{document}

### 2018-03-02 manuscript.Rnw
Clean up sporo_PIC
Add Table 3 to manuscript

@@
 setwd(here::here())
+source(file = "bin/microclimate.R")
+source(file = "bin/trait_PCA.R")
 source(file = "bin/cont_phylosig.R")
 source(file = "bin/binary_phylosig.R")
-source(file = "bin/microclimate.R")
+source(file = "bin/sporo_PIC.R")
 source(file = "bin/binaryPGLMM.R")
-source(file = "bin/trait_PCA.R")
 save(list = ls(all = TRUE), file= "results.RData")
@@
 
+% Table 3
+\clearpage
+
+<<sporo_PIC, results="asis">>=
+PIC.results <- round_df(PIC.results)
+PIC.results <- embolden_p(PIC.results, col.select=c("pval"))
+
+print(xtable(PIC.results,
+             label="table:PIC_sporo",
+             caption="Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth.
+All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades.
+Significant differences between epiphytes and terrestrial clades in bold."),
+      sanitize.colnames.function = custom_cols,
+      sanitize.rownames.function = custom_rows,
+      sanitize.text.function = function (x) {x},
+      include.rownames=TRUE,
+      caption.placement="top")
+@
 %%%%%%%%%%%%%%%%%%%%%%%%%%% Figures %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-03-02 bin/latex_functions.R

@@ -19,10 +19,14 @@ add_part_label <- function (df, labeltext) {
 
 # round_df
 # round the numbers in a dataframe to the specified number of digits
-round_df <- function (df, digits=3) {
+round_df <- function (df, digits=3, col.select = NULL) {
 
-  df[sapply(df, class) == "numeric" ] <- sapply(df[sapply(df, class) == "numeric" ], round, digits)
-
+  if (length(col.select) == 0) {
+    df[sapply(df, class) == "numeric" ] <- sapply(df[sapply(df, class) == "numeric" ], round, digits)
+  }
+  
+  df[,col.select] <- sapply(df[,col.select], round, digits)
+    
   return(df)
 }
 
@@ -43,6 +47,13 @@ custom_cols <- function (cols) {
   cols <- gsub ("^D$", "\\\\D", cols)
   cols <- gsub ("^prob_random$", "\\\\pval\\\\textsubscript{rnd}", cols)
   cols <- gsub ("^prob_brownian$", "\\\\pval\\\\textsubscript{BM}", cols)
+  
+  # continuous pglmmm columns
+  cols <- gsub ("^Parameter.estimate$", "Estimate", cols)
+  cols <- gsub ("^Lower.95..CI$", "Lower 95\\\\% CI", cols)
+  cols <- gsub ("^Upper.95..CI$", "Upper 95\\\\% CI", cols)
+  cols <- gsub ("^Effective.sample.size$", "Effective sample size", cols)
+  cols <- gsub ("^P.value$", "\\\\pval", cols)
   
   # binary pglmm columns
   cols <- gsub ("^sigmasq$", "$\\\\sigma^2$", cols)
@@ -65,6 +76,18 @@ custom_cols <- function (cols) {
   cols <- gsub ("inter\\.epi", "Intercept (epiphytic)", cols)
   cols <- gsub ("slope\\.ter", "Slope (terrestrial)", cols)
   cols <- gsub ("inter\\.ter", "Intercept (terrestrial)", cols)
+  
+  # trait PCA
+  cols <- gsub ("Std_PCA_Dim\\.1", "Standard PC1", cols)
+  cols <- gsub ("Std_PCA_Dim\\.2", "Standard PC2", cols)
+  cols <- gsub ("Phy_PCA_Dim\\.1", "Phylogenetic PC1", cols)
+  cols <- gsub ("Phy_PCA_Dim\\.2", "Phylogenetic PC2", cols)
+  
[…]

### 2018-03-02 bin/sporo_PIC.R

@@ -1,34 +1,33 @@
-# library(vegan)
-# library(picante)
-#library(ade4)
+# sporoPIC
+
+# Calculate phylogenetically independent contrasts for quantitative (sporophyte) traits 
+# between terrestrial and epiphytic species using brunch function in caper
+
+# load packages
+library(tidyverse)
 library(caper)
-library(xlsx)
-library(phytools)
+library(mooreaferns)
 
-setwd("/Users/joelnitta/R/moorea/")
-source("bin/my_funcs.R")
+# set working directory
+setwd(here::here())
 
-# Use brunch to run PIC against binary state (epiphytic or not)
-
-# set input files
-tree_file <- set_input_files()[[2]]
+####################
+### prepare data ###
+####################
 
 # load tree
-phy <- read.tree(tree_file)
+phy <- mooreaferns::fern_tree
 
 # load traits
-source("bin/traits/get_traits.R")
+traits <- mooreaferns::fern_traits
 
-# keep only quantitative traits with decent sampling
-sporo_traits <-c("habit", "stipe", "length", "width", "rhizome", "dissection", "pinna", "SLA")
+# reset levels for growth habit so epiphytic trait values are subtracted from terrestrial
+traits$habit <- factor(traits$habit, levels = c("terrestrial", "epiphytic"))
 
-# keep only quantitative traits with decent sampling
-traits <- traits[,sporo_traits]
-
-# make habit a factor
-traits$habit[traits$habit == 0] <- "terrestrial"
-traits$habit[traits$habit == 1] <- "epiphytic"
-traits$habit <- as.factor(traits$habit)
+# keep only species, habit, and quantitative traits to test
+traits.test <- c("stipe", "length", "width", "rhizome", "dissection", "pinna", "sla")
+traits <- traits[,c("species", "habit", traits.test)]
+rownames(traits) <- traits$species
 
 # trim to only species with trait data
 phy <- drop.tip(phy, phy$tip.label[!(phy$tip.label %in% rownames(traits))])
@@ -36,55 +35,19 @@ phy <- drop.tip(phy, phy$tip.label[!(phy$tip.label %in% rownames(traits))])
 # get traits in same order as tips
 traits <- traits[phy$tip.label,]
 
-########################################################
-### phylogenetic ANOVA using phylANOVA from phytools ###
[…]

### 2018-03-03 readme.Rmd
Add readme

---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns

This repository contains code needed to run the analyses and generate the manuscript:

Nitta, J.H., Watkins, J.E., Taputuarai, R., and Davis, C.C. Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns. Submitted.

### Dependencies
- R (including knitr and various packages called by scripts)  
- LaTeX (optional, see below)  
- RStudio (optional, see below)
- latex2rtf (optional, see below)

Data (and a few functions) are contained in a separate package. This needs to be installed as follows:

```r
devtools::install_github("joelnitta/mooreaferns", dependencies=TRUE)
```

### Instructions
Clone or [download](https://github.com/joelnitta/canopylife/archive/master.zip) and unzip this repository, and navigate to it.

Analyses were conducted in R, and each can be run separately from `bin/`.

To run all analyses, output figures, and compile the manuscript as a pdf, the following options are availble depending on whether/how extensively LaTeX is installed on your machine:

#### - From the terminal
This method should work with a full LaTeX distribution.

1. Open the terminal and execute this command to run all analyses and output (i.e., "knit") the tex file:

```
Rscript -e "library(knitr); knit('./manuscript.Rnw')"
```

2. Next, typeset the pdf from the tex file:

```
xelatex manuscript.tex
```

The second step may fail depending on how Latex has been installed. In this case, try one of the methods below.

#### - Overleaf
This method does not require LaTeX to be installed at all.

1. Execute the first command in "From the terminal" above.
2. Upload all the files in this directory to an account on <https://www.overleaf.com> and typeset the pdf there. XeLaTex should be specified as the LaTeX Engine under Settings.

#### - RStudio
This method should work with a minimal LaTeX distribution if the packages used in the tex file are installed.

1. In RStudio Preferences:Sweve, set "Typeset LaTeX into PDF" to XeLaTex, if it isn't already.
2. Open `manuscript.Rnw` in RStudio and press "Compile PDF".

#### - Tables only
To convert the pdf to MS Word for submission of the manuscript, the [copy-and-paste method](https://joelnitta.com/blog/getting-latex-into-word) was used. This works Pretty Darn Well, except for tables, which get garbled during pasting. To convert tables to MS Word (i.e., rich text) format, latex2rtf works better. However, latex2rtf is limited in the LaTeX packages it can use. A simple knitr file using only bare-bones LaTeX packages is available to output a tex file containing only tables, which can then be processed by latex2rtf. After knitting manuscript.Rnw, run the following:

```
Rscript -e "library(knitr); knit('./tables.Rnw')"
latex2rtf tables.tex
```
### Cleanup
Run this script from the terminal to remove intermediate files generated during compiling the pdf. This can't be undone, so use with care!

```
bash cleanup.sh
```

### 2018-03-03 readme.Rmd
Add instructions for compiling SI

@@
 ### Dependencies
+
 - R (including knitr and various packages called by scripts)  
+- RStudio (optional, see below)
 - LaTeX (optional, see below)  
-- RStudio (optional, see below)
 - latex2rtf (optional, see below)
@@
 
-### Instructions
+### Getting started
+
 Clone or [download](https://github.com/joelnitta/canopylife/archive/master.zip) and unzip this repository, and navigate to it.
 
-Analyses were conducted in R, and each can be run separately from `bin/`.
+All analyses were conducted in R, and each can be run separately from `bin/`.
+
+### Compiling the manuscript
 
@@
 #### - From the terminal
+
 This method should work with a full LaTeX distribution.
@@
 #### - Overleaf
+
 This method does not require LaTeX to be installed at all.
@@
 #### - RStudio
+
 This method should work with a minimal LaTeX distribution if the packages used in the tex file are installed.
@@
 
-#### - Tables only
+### Compiling Supporting Information
+
+After compiling the manuscript using one of the methods above and before running `cleanup.sh` (so that `results.Rdata` is still present), compile the SI in the same way but replace `manuscript` with `supporting_information`.
+
+### Compiling tables only
+
 To convert the pdf to MS Word for submission of the manuscript, the [copy-and-paste method](https://joelnitta.com/blog/getting-latex-into-word) was used.
@@
 A simple knitr file using only bare-bones LaTeX packages is available to output a tex file containing only tables, which can then be processed by latex2rtf.
-After knitting manuscript.Rnw, run the following:
+
+After knitting manuscript.Rnw and before running `cleanup.sh`, run the following:
 
@@
 Rscript -e "library(knitr); knit('./tables.Rnw')"
+latex tables.tex
+bibtex tables.aux
+latex tables.tex
 latex2rtf tables.tex
 ```
+
 ### Cleanup
+
 Run this script from the terminal to remove intermediate files generated during compiling the pdf. This can't be undone, so use with care!

### 2018-03-03 cleanup.sh
add cleanup.sh

@@ -1,0 +1,19 @@
+#!bin/bash
+
+# deletes intermediate files. 
+# these can't be recovered, so use with care!
+
+# various intermediate latex files
+rm -f *.aux
+rm -f *.bbl
+rm -f *.blg
+rm -f *.dvi
+rm -f *.log
+
+# R output
+rm -f results.RData
+rm -f PCA.pdf
+rm -f microclimate.pdf
+
+# etc
+rm -r readme.html

### 2018-03-03 tables.Rnw
Add tables.Rnw to output only tables

--- 2018-03-01 supporting_information.Rnw
@@
 \documentclass{article}
-\date{}
 
@@
 \usepackage[margin=1in]{geometry}
-
-% Set line spacing to 1.5
-\usepackage{setspace}
-\onehalfspace
 
@@
 % Set unit and abbreviations formatting
-\input{newphyt_style}
+\input{table_style}
 
@@
 library(knitr)
+opts_knit$set(concordance=FALSE)
 opts_chunk$set(echo=FALSE)
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%% Tables %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 <<setup_tables>>=
 library(xtable)
+setwd(here::here())
 source("bin/latex_functions.R")
@@
 
-\noindent \textbf{\textit{New Phytologist} Supporting Information} \\
-\\
-\noindent Article title: Life in the canopy: phylogenetic filtering and morphological canalization in epiphytic ferns \\
-\noindent Authors: Joel H. Nitta, James E. Watkins, Jr., Ravahere Taputuarai and Charles C. Davis \\
-\noindent Article acceptance date: someday soon \\
-\\
-\noindent The following Supporting Information is available for this article: \\
+% Table 1
+\clearpage
 
-\noindent \textbf{Methods S1} We performed alternative tests for correlation of growth habit with functional traits using phylogenetic generalized linear mixed models (PGLMMs).
-Models were constructed with each trait as the response variable in turn dependent on growth habit, with phylogeny as a random effect.
-For binary (gametophyte) traits we used the \function{binaryGLMM} function in the \package{ape} package \citep{Paradis2004}.
-For quantitative (sporophyte) traits, we used the \function{MCMCglmm} function in the R package of the same name \citep{Hadfield2010}.
-Priors were set to an inverse-Gamma distribution with shape and scale parameters equal to \num{0.01} \citep{Hadfield2010a}.
-The MCMC algorithm was run for \num{5000000} iterations with a burnin of \num{1000} and thinning every \num{500} iterations.
+\begin{table}[ht]
+\centering
+\caption{Traits used in this study. Noncordate gametophyte morphotypes include ribbon, strap and filamentous.}
+\begin{tabular}{ p{0.3\textwidth} p{0.2\textwidth} p{0.4\textwidth} }
+\hline
+ & Data type (unit) & Ecological Significance \\
+\hline
+Sporophyte Traits & & \\
+\hspace{3ex}Stipe Length  & Continuous (cm)  & Shorter stipes compensate for low conductivity \citep{WatkinsJr.2010} \\
+\hspace{3ex}Frond Length  & Continuous (cm)  & Smaller leaf size reduces evapotranspiration \citep{Vogel1968} \\
+\hspace{3ex}Frond Width  & Continuous (cm)  & Smaller leaf size reduces evapotranspiration \citep{Vogel1968} \\
+\hspace{3ex}Rhizome Diameter  & Continuous (cm)  & Scales with overall plant size (Creese et al. 2011) \\
+\hspace{3ex}Frond Dissection  & Ordinal  & Simple laminae reduce evapotranspiration (Kluge and Kessler 2007) \\
+\hspace{3ex}Pinna number & Integer & Scales with overall plant size (Creese et al. 2011) \\
+\hspace{3ex}Specific Leaf Area & Continuous (m\textsuperscript{2} kg\textsuperscript{-1}) & Species in more stressful environments invest in thicker leaves \citep{Wright2004a} \\
+Gametophyte Traits & & \\
+\hspace{3ex}Morphotype & Binary (cordate vs. noncordate) & Noncordate morphology reduces drying rates by holding external water in crevices and folds \citep{WatkinsJr.2007,Pittermann2013} \\
+\hspace{3ex}Gemmae & Binary (present/absent) & Enable asexual reproduction, thereby overcoming local extinction due to drying \citep{Farrar2008} \\
+\hspace{3ex}Glands & Binary (present/absent) & Lipophilic exudates produced by glands may reduce transpiration rates \citep{Wollenweber2000} \\
+\hspace{3ex}Hairs & Binary (present/absent) & Hairs reduce drying rates by holding external water and increasing boundary layer thickness \citep{WatkinsJr.2007} \\
+\hline
+\end{tabular}
+\label{table:traits}
+\end{table}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%% Tables %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Table 2
+\clearpage
 
+<<cont_phylosig, results="asis">>=
+cont_phylosig.results <- round_df(cont_phylosig.results)
+cont_phylosig.results <- embolden_p(cont_phylosig.results, col.select=c("k.pval","lambda.pval"))
+part_label <- add_part_label(cont_phylosig.results, "(a)")
+
+print(xtable(cont_phylosig.results,
+             label="table:phylosig",
+             caption="Phylogenetic signal in traits.
+(a) Continuous traits and (b) binary traits.
+Values of $\\lambda$ or $\\K >= 1$ indicate that traits are evolving according to Brownian Motion (BM); values of $\\lambda$ or \\K near 0 indicate random distribution of traits.
+For \\D, values of 0 (or less) indicate evolution according to (or more conserved than) BM; values near 1 indicate random trait distribution.
+\\pval\\textsubscript{Rnd}, result of testing that observed \\D is significantly different from 1; \\pval\\textsubscript{BM}, result of testing that observed \\D is significantly different from 0."),
+      sanitize.colnames.function = custom_cols,
+      sanitize.rownames.function = custom_rows,
+      sanitize.text.function = function (x) {x},
+      include.rownames=TRUE,
+      add.to.row = part_label,
+      caption.placement="top")
+@
+
+<<binary_phylosig, results="asis">>=
+binary_phylosig.results <- round_df(binary_phylosig.results)
+binary_phylosig.results <- embolden_p(binary_phylosig.results, col.select=c("prob_random","prob_browian"))
+part_label <- add_part_label(binary_phylosig.results, "(b)")
+
+print(xtable(binary_phylosig.results,
+             caption=NULL),
+      sanitize.colnames.function = custom_cols,
+      sanitize.rownames.function = custom_rows,
+      sanitize.text.function = function (x) {x},
+      include.rownames=TRUE,      
+      add.to.row = part_label)
+@
+
+% Table 3
+\clearpage
+
+<<sporo_PIC, results="asis">>=
+PIC.results <- round_df(PIC.results)
+PIC.results <- embolden_p(PIC.results, col.select=c("pval"))
+
+print(xtable(PIC.results,
+             label="table:PIC_sporo",
+             caption="Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth.
+All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades.
+Significant differences between epiphytes and terrestrial clades in bold."),
+      sanitize.colnames.function = custom_cols,
+      sanitize.rownames.function = custom_rows,
+      sanitize.text.function = function (x) {x},
+      include.rownames=TRUE,
+      caption.placement="top")
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% Appendix Tables %%%%%%%%%%%%%%%%%%%%%%
 % Add S to table number and reset counter
@@
 
-% Table S1 (supp_table:PGLMM_sporo)
+% Table S1
+\clearpage
+
 <<PGLMM, results="asis">>=
@@
 \clearpage
+
 <<binaryPGLMM, results="asis">>=
@@
 \clearpage
+
 <<climate_ancova, results="asis">>=
@@
 \clearpage
+
 <<trait_pca, results="asis">>=
@@
 @
-%%%%%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% set style
+%%%%%%%%%%%%%%%%%%%%%%%%%%% Refs in Tables %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \bibliographystyle{newphyt} 
-
-% change the bibliography title
-\renewcommand\refname{References}
-
-% insert bibliography
 \bibliography{references}

### 2018-03-03 manuscript.Rnw
Add Pagel's correlated evo test to manuscript
Also, change font to TNR in MS and SI

@@
-% run this from the command line to compile: Rscript -e "library(knitr); knit('./manuscript.Rnw')"
 \documentclass{article}
@@
 \usepackage[margin=1in]{geometry}
+
+% Use Times New Roman font (better conversion of special characters in Word)
+\usepackage{fontspec}
+\setmainfont{Times New Roman}
 
@@
 
-\begin{document}
-
 \title{Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns}
@@
 %\affil[3]{Association Te Rau Ati Ati a Tau a Hiti Noa Tu, B.P. 11553, Mahina, Tahiti, French Polynesia}
+
+\begin{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title Page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 source(file = "bin/sporo_PIC.R")
+source(file = "bin/correlated_evo.R")
 source(file = "bin/binaryPGLMM.R")
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Summary %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \clearpage
@@
 @
+
+% Table 4
+\clearpage
+
+<<pagel_test, results="asis">>=
+pagel.results <- round_df(pagel.results)
+pagel.results <- embolden_p(pagel.results, col.select=c("pval"))
+
+print(xtable(pagel.results,
+             label="table:pagel_gameto",
+             caption="Pagel's \\citeyearpar{Pagel1994} test of correlated evolution between binary traits.
+Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
+terrestrial).
+Models with significantly different log likelihoods (LL) in bold; lower LL for the dependent model indicates correlation with growth habit."),
+      sanitize.colnames.function = custom_cols,
+      sanitize.rownames.function = custom_rows,
+      sanitize.text.function = function (x) {x},
+      include.rownames=TRUE,
+      caption.placement="top")
+@
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%% Figures %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-03-03 bin/correlated_evo.R

@@ -1,37 +1,50 @@
+# correlated_evo
+
 # run Pagel's (1994) test of correlated evolution on two binary traits using fitPagel in phytools
 
-library(xlsx)
+# load packages
 library(phytools)
+library(mooreaferns)
 
-setwd("/Users/joelnitta/R/moorea/")
-source("bin/my_funcs.R")
-
-######################
-### data wranglin' ###
-######################
-
-### get raw data
-# set input files
-tree_file <- set_input_files()[[2]]
+#####################
+### prepare data  ###
+#####################
 
 # load tree
-phy <- read.tree(tree_file)
+phy <- mooreaferns::fern_tree
 
-# load untransformed traits
-traits <- get.my.traits(log.trans = FALSE, scale.traits = FALSE)
+# load un-transformed species traits
+traits <- mooreaferns::fern_traits
+rownames(traits) <- traits$species
 
-# load distribution data (widespread or not)
-# distribution data is in "results.df"
-source("bin/alpha_diversity/elevational range table.R")
+# ### get raw data
+# # set input files
+# tree_file <- set_input_files()[[2]]
+# 
+# # load tree
+# phy <- read.tree(tree_file)
+# 
+# # load untransformed traits
+# traits <- get.my.traits(log.trans = FALSE, scale.traits = FALSE)
 
-### merge distribution data with traits
-# first make all not widespread
-traits$distribution <- "notwidespread"
-# get list of widespread species
-widespread_species <- rownames(results.df[results.df$widespread == 1,])
-# make those in list widespread
-traits$distribution[rownames(traits) %in% widespread_species] <- "widespread"
-rm(results.df)
+
+# #***** add widespread observations to data folder
+# 
+# # load distribution data (widespread or not)
+# # distribution data is in "results.df"
+# source("bin/alpha_diversity/elevational range table.R")
+# 
+# ### merge distribution data with traits
+# # first make all not widespread
+# traits$distribution <- "notwidespread"
+# # get list of widespread species
+# widespread_species <- rownames(results.df[results.df$widespread == 1,])
+# # make those in list widespread
+# traits$distribution[rownames(traits) %in% widespread_species] <- "widespread"
+# rm(results.df)
+
+
+
 
 ### match up traits and tree
[…]

### 2018-03-03 bin/latex_functions.R

@@ -35,6 +35,7 @@ round_df <- function (df, digits=3, col.select = NULL) {
 # this function is to be used as input to sanitize.colnames.function of print.xtable
 custom_cols <- function (cols) {
   
+  # Table 2
   # continuous phylogenetic signal
   cols <- gsub ("^lambda$", "$\\\\lambda$", cols)
   cols <- gsub ("^lambda\\.pval$", "$\\\\pval_\\\\lambda$", cols)
@@ -48,6 +49,20 @@ custom_cols <- function (cols) {
   cols <- gsub ("^prob_random$", "\\\\pval\\\\textsubscript{rnd}", cols)
   cols <- gsub ("^prob_brownian$", "\\\\pval\\\\textsubscript{BM}", cols)
   
+  # Table 3
+  # quantitative PICs
+  cols <- gsub ("^num_contrasts$", "Number of contrasts", cols)
+  cols <- gsub ("^num_pos_con$", "Positive contrasts", cols)
+  cols <- gsub ("^tval$", "\\\\tval", cols)
+  cols <- gsub ("^pval$", "\\\\pval", cols)
+  
+  # Table 4
+  # Pagel's correlated evolution
+  cols <- gsub ("^logL_indep$", "LL (independent model)", cols)
+  cols <- gsub ("^logL_dep$", "LL (dependent model)", cols)
+  cols <- gsub ("^likelihood_ratio$", "Likelihood ratio", cols)
+  
+  # Table S1
   # continuous pglmmm columns
   cols <- gsub ("^Parameter.estimate$", "Estimate", cols)
   cols <- gsub ("^Lower.95..CI$", "Lower 95\\\\% CI", cols)
@@ -55,6 +70,7 @@ custom_cols <- function (cols) {
   cols <- gsub ("^Effective.sample.size$", "Effective sample size", cols)
   cols <- gsub ("^P.value$", "\\\\pval", cols)
   
+  # Table S2
   # binary pglmm columns
   cols <- gsub ("^sigmasq$", "$\\\\sigma^2$", cols)
   cols <- gsub ("^sigmap$", "\\\\pval($\\\\sigma^2=0$)", cols)
@@ -63,6 +79,7 @@ custom_cols <- function (cols) {
   cols <- gsub ("^zscore$", "\\\\zval-score", cols)
   cols <- gsub ("^pvalue$", "\\\\pval", cols)
   
+  # Table S3
   # microclimate ancova columns
[…]

### 2018-03-03 cleanup.sh
Fix typo in cleanup.sh

@@ -16,4 +16,4 @@ rm -f PCA.pdf
 rm -f microclimate.pdf
 
 # etc
-rm -r readme.html
+rm -f readme.html

### 2018-03-12 bin/microclimate.R
Fix syncing between microclimate and microclimate_means
Add italics to R2 in microclimate plots

@@ -21,6 +21,9 @@ source("bin/microclimate_means.R")
 ### run ANCOVA ###
 ##################
 
+# use only sites on Moorea
+moorea_sites <- mooreaferns::sites[grep("Aorai", mooreaferns::sites$site, invert=TRUE), ]
+
 # add latitude, longitude, and elevation to microclimate means
 daily_values <- merge(daily_values, moorea_sites, by="site", all.x=TRUE)
 grand_means <- merge(grand_means, moorea_sites, by="site", all.x=TRUE)
@@ -29,9 +32,9 @@ grand_means <- merge(grand_means, moorea_sites, by="site", all.x=TRUE)
 #anyNA(daily_values)
 #anyNA(grand_means)
 
-# make sure growth habit is factor
-daily_values$habit <- as.factor(daily_values$habit)
-grand_means$habit <- as.factor(grand_means$habit)
+# convert growth habit to factor
+daily_values$habit <- factor(daily_values$habit, levels=c("terrestrial", "epiphytic"))
+grand_means$habit <- factor(grand_means$habit, levels=c("terrestrial", "epiphytic"))
 
 # set response variables
 resp_vars <- c("max_temp", "mean_temp", "min_temp", "sd_temp", "max_RH", "mean_RH", "min_RH", "sd_RH")
@@ -70,8 +73,6 @@ for (i in 1:length(resp_vars)) {
 ancova.results <- as.data.frame(dplyr::bind_rows(parameters))
 rownames(ancova.results) <- resp_vars
 
-# write.csv(ancova.results, file="climate_ancova.csv")
-
 ###############################################
 #### Get slopes and intercepts for plotting ###
 ###############################################
@@ -109,26 +110,34 @@ rownames(slopes.results) <- resp_vars
 
 # function to make scatterplot with lines and R2 values from linear model
 make_scatter_plot <- function (input.plot.data, model_data, indep_var, dep_var, ylabel) {
+  
+  # extract formatted r squared value for this plot
+  r.squared <- paste("italic(R)^2 ==", round(model_data[dep_var,"r.squared"], 2), sep=" " )    
+  
   p <- ggplot (data = input.plot.data, aes_string(x = indep_var, y = dep_var, color = "habit")) +
     geom_abline(intercept = model_data[dep_var,"inter.ter"], slope = model_data[dep_var,"slope.ter"], color=cols[1]) +
[…]

### 2018-03-12 bin/microclimate_means.R

@@ -8,7 +8,7 @@ library(mooreaferns) # moorea_climate
 setwd(here::here())
 
 # calculate daily maximum, mean, minimum, and SD of temperature and RH by site
-daily_values <- plyr::ddply(moorea_climate, plyr::.(site, date), summarize,
+daily_values <- plyr::ddply(moorea_climate, plyr::.(site, date), plyr::summarize,
                             max_temp = max(temp),
                             mean_temp = mean(temp),
                             min_temp = min(temp),
@@ -19,7 +19,7 @@ daily_values <- plyr::ddply(moorea_climate, plyr::.(site, date), summarize,
                             sd_RH = sd(RH))
 
 # calculate grand means of daily values for temperature and RH by site
-grand_means <- plyr::ddply(daily_values, "site", summarize,
+grand_means <- plyr::ddply(daily_values, "site", plyr::summarize,
                            max_temp = mean(max_temp),
                            mean_temp = mean(mean_temp),
                            min_temp = mean(min_temp),

### 2018-03-12 manuscript.Rnw
Add functional diversity to manuscript
And rest of figures

@@
 source(file = "bin/binaryPGLMM.R")
+source(file = "bin/functional_diversity.R")
 save(list = ls(all = TRUE), file= "results.RData")
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%% Figure Captions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
+
+\noindent\fig \ref{fig:traits}.
+A) Sporophyte traits.
+All sporophyte traits were quantitative, including frond length and width, stipe length, rhizome dia., number of pinna pairs (in this example, nine) and degree of lamina dissection (in this example, once-pinnate).
+Specific leaf area (ratio of area to mass of leaf lamina) not depicted.
+B) Gametophyte traits.
+Morphotype was defined as a binary trait, either cordate or non-cordate.
+Examples of non-cordate morphotypes include ribbon (thallus elongate and two-dimensional) and filamentous (thallus single lines of cells).
+Other gametophyte traits (gemmae, hairs, and glands) were scored as present or absent.
+Arrows point out instances of each binary trait.
+Scalebars 1 mm except for glands, which is 0.1 mm.
+Photographs by J.
+H.
+Nitta.
+% include blank line between each caption
+\vspace{\baselineskip}
+
+\noindent\fig \ref{fig:climate}.
+Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter on Moorea, French Polynesia.
+Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from July 7, 2013 to July 5, 2014, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).
+`Epiphytic' dataloggers were placed at ca.
+2 m on trees, and `terrestrial' dataloggers placed at ground level.
+Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown.
+Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval \textless \num{.05}
+\vspace{\baselineskip}
+
+\noindent\fig \ref{fig:PCA}. Principal components analysis (PCA) of traits related to epiphytic growth
+in ferns from Moorea, French Polynesia. A–B) PC loadings. C–D) species scores. A
+and C are standard PCA, B and \D are phylogenetic PCA. Epiphytes in green, terrestrial
+species in brown. Quantitative traits only, including only species with no missing
+observations for any traits ($\n = 104$ \si{\species}). Abbreviations for traits as follows: Di, degree
+of frond dissection; Le, frond length; Pi, number of pinna pairs; Rh, rhizome diameter;
+SL, specific leaf area; St, stipe length; Wi, frond width.
+\vspace{\baselineskip}
+
+\noindent\fig \ref{fig:traits_tree}. Time-calibrated phylogenetic tree of ferns from Moorea, French
+Polynesia with growth habit and associated traits mapped on the tips. Relative value
+of quantitative (sporophyte) traits in greyscale: low values are lighter, high values are
+darker. States of qualitative (gametophyte) states and growth habit indicated by colors in
+key. Slashes indicate missing data or non-applicable trait states. Trends for three leaf size
+traits (stipe length, frond length, and frond with) were quantitatively very similar, so of
+these, we only present stipe length. Species missing data for six or more traits not shown.
+For a summary of traits, see Table 3.1. Abbreviations for clades of interest are as follows.
+Major Cretaceous epiphytic radiations \sensu \cite{Schuettpelz2009} in green: H,
+hymenophylloids; T, trichomanoids, V, vittarioids; A, asplenioids; E, elaphoglossoids; P,
+polygrammoids (trichomanoids diversified prior to the Cretaceous but are included as an
+important extant epiphytic clade). Other major clades: EI, eupolypods I; EII, eupolypods
+II.
+\vspace{\baselineskip}
+
+\noindent\fig \ref{fig:CWM}. Mean values of selected functional morphological traits of epiphytic and
+terrestrial ferns on Moorea, French Polynesia. Epiphytic values in green, terrestrial
+values in brown. A) Community-weighted mean (CWM) values for each trait plotted
+against either minimum relative humidity ($\n = 13$ communities each for epiphytes and
+terrestrial species) or elevation ($\n = 17$ communities each). Trendlines indicate
+significant relationships determined by a linear model at $\pval<0.05$. B) Mean values for all
+species by growth habit for each trait. Numbers of species with observations available
+differed for each trait and are shown at bottom of each barplot. Asterisks indicate
+statistical significance of one-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+Trends for three leaf size traits (stipe length, frond length, and frond with) were
+quantitatively very similar, so of these we only present stipe length.
+\vspace{\baselineskip}
+
+\noindent\fig \ref{fig:FDIV}. Functional and phylogenetic diversity of epiphytic and terrestrial fern
+communities on Moorea, French Polynesia. Epiphytic communities in green, terrestrial
+communities in brown. Response variable abbreviations as follows: MPD, standard effect
+size (SES) of mean phylogenetic distance; MNTD, SES of mean nearest taxon distance;
+FRic, functional richness; FEve, functional evenness; FDiv, functional diversity. A)
+Response variables plotted against either minimum relative humidity ($\n = 13$
+communities each for epiphytes and terrestrial species) or elevation ($\n = 17$ communities
+each). Trendlines indicate significant relationships determined by a linear model at $\pval<0.05$. B) Median values (box plots) of response variables by growth habit (epiphytic or
+terrestrial) across all communities (epiphytic communities, $\n = 17$; terrestrial
+communities, $\n = 17$). Medians shown with bold lines, lower and upper hinges
+correspond to first and third quartiles, and whiskers extend to values within 1.5 $\times$ the
+interquartile range. Asterisks indicate statistical significance of one-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%% Figures %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\begin{figure}[ht]
+  \includegraphics[scale=1]{traits_examples.pdf}
+  \caption{Traits}
+  \label{fig:traits}
+\end{figure}
+
+\clearpage
 \begin{figure}[ht]
@@
 
+\clearpage
+\begin{figure}[ht]
+  \includegraphics[scale=1]{PCA.pdf}
+  \caption{Principal Components Analysis}
+  \label{fig:PCA}
+\end{figure}
+
+\clearpage
+\begin{figure}[ht]
+  \includegraphics[scale=1]{traits_on_tree.pdf}
+  \caption{Traits on tree}
+  \label{fig:traits_tree}
+\end{figure}
+
+\clearpage
+\begin{figure}[ht]
+  \includegraphics[scale=.75]{cwm_plots.pdf}
+  \caption{Community-weighted means}
+  \label{fig:CWM}
+\end{figure}
+
+\clearpage
+\begin{figure}[ht]
+  \includegraphics[scale=.75]{fdiv_plots.pdf}
+  \caption{Functional Diversity}
+  \label{fig:FDIV}
+\end{figure}
+
 \end{document}

### 2018-03-12 manuscript.Rnw
Fix figure captions for func div

@@
 
-\noindent\fig \ref{fig:CWM}. Mean values of selected functional morphological traits of epiphytic and
-terrestrial ferns on Moorea, French Polynesia. Epiphytic values in green, terrestrial
-values in brown. A) Community-weighted mean (CWM) values for each trait plotted
+\noindent\fig \ref{fig:CWM}.
+Change in selected functional traits of ferns on Moorea, French Polynesia across environmental gradients and between growth habit types (epiphytic or terrestrial).
+Epiphytic values in green, terrestrial
+values in brown. Scatterplots show community-weighted mean (CWM) values for each trait plotted
 against either minimum relative humidity ($\n = 13$ communities each for epiphytes and
 terrestrial species) or elevation ($\n = 17$ communities each). Trendlines indicate
-significant relationships determined by a linear model at $\pval<0.05$. B) Mean values for all
-species by growth habit for each trait. Numbers of species with observations available
-differed for each trait and are shown at bottom of each barplot. Asterisks indicate
-statistical significance of one-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+significant relationships determined by a linear model at $\pval<0.05$. Barplots show mean values for all
+species by growth habit (epiphytic or
+terrestrial) for each trait. Error bars show standard error. Asterisks indicate
+statistical significance of two-sided \tval-test for differences between epiphytic and terrestrial species; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
 Trends for three leaf size traits (stipe length, frond length, and frond with) were
@@
 size (SES) of mean phylogenetic distance; MNTD, SES of mean nearest taxon distance;
-FRic, functional richness; FEve, functional evenness; FDiv, functional diversity. A)
-Response variables plotted against either minimum relative humidity ($\n = 13$
+FRic, functional richness; FEve, functional evenness; FDiv, functional diversity. Scatterplots show
+response variables plotted against either minimum relative humidity ($\n = 13$
 communities each for epiphytes and terrestrial species) or elevation ($\n = 17$ communities
-each). Trendlines indicate significant relationships determined by a linear model at $\pval<0.05$. B) Median values (box plots) of response variables by growth habit (epiphytic or
+each). Trendlines indicate significant relationships determined by a linear model at $\pval<0.05$. Boxplots show median values (bold lines) of response variables by growth habit (epiphytic or
 terrestrial) across all communities (epiphytic communities, $\n = 17$; terrestrial
-communities, $\n = 17$). Medians shown with bold lines, lower and upper hinges
+communities, $\n = 17$). Lower and upper hinges
 correspond to first and third quartiles, and whiskers extend to values within 1.5 $\times$ the

### 2018-03-18 manuscript.Rnw
Add methods, start on results
Add get_results function to extract rounded result value from dataframe, start using Sexpr to insert values into text. Add supplemental table captions at end so they get referenced properly (even though supplemental info is a separated rnw file).

@@
 \documentclass{article}
+
+%%% Notes:
+% Formatted for New Phytologist.
+% Should be compiled with XeLaTeX (supports unicode chars).
+% This version is designed to be easy to convert to .doc by copying the pdf text and pasting into Word.
+
+% Don't print date
+\date{}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title Page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\maketitle
-
 <<setup, echo=FALSE>>=
@@
 @
+
+<<get_result_func, echo=FALSE>>=
+# simple function to get rounded results with less typing
+get_result <- function (results_table, col_name, row_name, signif=FALSE, digits=3) {
+  if (signif == FALSE) {
+    value <- round(results_table[row_name, col_name], digits=3)
+  } else if (signif == TRUE) {
+    value <- signif(results_table[row_name, col_name], digits=digits)
+  }
+  return(value)
+}
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title Page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% For conversion to word, remove page numbers
+\pagenumbering{gobble}
+
+\maketitle
 
@@
 
-The widespread appearance of angiosperm-dominated forests in the late Cretaceous and early Cenozoic had an immense effect on terrestrial ecosystems \citep{Niklas1983, Kenrick1997a,Lloyd2008}.
+The appearance of angiosperm-dominated forests in the late Cretaceous and early Cenozoic had an immense effect on terrestrial ecosystems \citep{Niklas1983, Kenrick1997a,Lloyd2008}.
 The heterogeneous niche space of the angiosperm canopy comprising stratified light and humidity levels and a range of substrates for growth became rapidly filled by a variety of epiphytic plants, including mosses \citep{Fiz-Palacios2011}, leafy liverworts \citep{Feldberg2014}, orchids \citep{Givnish2015b}, bromeliads \citep{Givnish2014a}, and ferns \citep{Schneider2004d}.
-Diversification rates in some of these groups may have further increased during the Paleocene / Eocene thermal maximum, a period of globally warmer and wetter climates that may have enabled movement into the canopy \citep{Schuettpelz2009,Testo2016a}.
+Diversification rates in some of these groups may have further increased during the Paleocene / Eocene thermal maximum, a period of warmer and wetter climates that may have enabled movement into the canopy \citep{Schuettpelz2009,Testo2016a}.
 The transition to epiphytic growth has been particularly important for fern diversity, with ca.
@@
 
-Despite the ecological opportunities afforded by life in the canopy, including escape from disturbance \citep{Page2002} and abundant light for photosynthesis \citep{Benzing2000}, epiphytes must contend with greater abiotic stress in the form of increased evaporation and more extreme oscillations in temperature relative to terrestrial plants \citep{Zotz2001,Lowman2012}.
+Despite the ecological opportunities afforded by life in the canopy, including escape from disturbance \citep{Page2002} and abundant light for photosynthesis \citep{Benzing2000}, epiphytes must contend with greater stress in the form of increased evaporation and extreme oscillations in temperature relative to terrestrial plants \citep{Zotz2001,Lowman2012}.
 Associated with this change, epiphytic ferns have evolved various morphological adaptations that minimize water loss from their leaves, including shorter stipes \citep{WatkinsJr.2010}, thicker leaves \citep{Kluge2007,WatkinsJr.2007b}, and protection with scales or hairs \citep{WatkinsJr.2006b}.
-Some lineages have specialized adaptations to epiphytic growth, such as humus-collection (\eg, \textit{Asplenium}; \citealt{Benzing1990}, \textit{Drynaria}; \citealt{Janssen2005}), vegetative desiccation tolerance (\eg, Hymenophyllaceae; \citealt{Shreve1911}, \textit{Pleopeltis}; \citealt{Pessin1925}), CAM photosynthesis \citep[\eg, \textit{Pyrrosia},][]{Wong1976}, and symbiotic relationships with ants (\eg, \textit{Antrophyum}; \citealt{WatkinsJr.2008}, \textit{Lecanopteris}; \citealt{Gay1993a}).
-Many of these same adaptations are found in epiphytic seed plants.
-However, the evolution of epiphytisim in ferns was much more complex given that ferns rely on a free-living gametophyte.
-The fern gametophyte is anatomically simple, and differs from the sporophyte markedly in anatomy, morphology, and physiology.
-Such differences create a system of “dual ecology” where selective pressures acting on one stage are very different than those acting on another.
-For example, desiccation tolerance, a key trait for non-vascular plants living in water-limited environments, is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is quite rare in sporophytes \citep{WatkinsJr.2007}.
-Despite the evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no studies to our knowledge have analyzed the evolution of such traits across the fern evolutionary tree including both life stages and representatives of all five Cretaceous epiphytic radiations \sensu \cite{Schuettpelz2009}.
+Some lineages have specialized adaptations to epiphytic growth, such as humus-collection (\eg, \textit{Asplenium}; \citealt{Benzing1990}, \textit{Drynaria}; \citealt{Janssen2005}), vegetative desiccation tolerance (DT) (\eg, Hymenophyllaceae; \citealt{Shreve1911}, \textit{Pleopeltis}; \citealt{Pessin1925}), CAM photosynthesis \citep[\eg, \textit{Pyrrosia},][]{Wong1976}, and symbiotic relationships with ants (\eg, \textit{Antrophyum}; \citealt{WatkinsJr.2008}, \textit{Lecanopteris}; \citealt{Gay1993a}).
+Many of these same adaptations are found in epiphytic seed plants \citep{Benzing1990}.
+However, the evolution of epiphytism in ferns was more complex given that ferns rely on a free-living gametophyte.
+The fern gametophyte is anatomically simple, and differs from the sporophyte in anatomy, morphology, and physiology.
+Such differences create a system of `dual ecology' where selective pressures acting on one stage are different than those acting on another.
+For example, DT is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is rare in sporophytes \citep{WatkinsJr.2007}.
+Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no studies to our knowledge have analyzed the evolution of such traits across the fern phylogeny including both life stages and representatives of all five Cretaceous epiphytic radiations \sensu \cite{Schuettpelz2009}.
 
-The majority of studies investigating adaptations to epiphytism have utilized a univariate, cross-species approach, analyzing traits individually between terrestrial and epiphytic species \citep[\eg,][]{WatkinsJr.2007b,WatkinsJr.2010}.
-However, given the evolutionary history of ferns with multiple radiations within the epiphytic niche \citep{Schuettpelz2009}, it is likely that traits related to epiphytism carry at least some degree of phylogenetic signal (\ie, the tendency for closely related species to resemble each other).
+Most studies investigating adaptations to epiphytism have utilized a univariate, cross-species approach, analyzing traits individually between terrestrial and epiphytic species \citep[\eg,][]{WatkinsJr.2007b,WatkinsJr.2010}.
+However, given multiple radiations into the epiphytic niche \citep{Schuettpelz2009}, it is likely that traits related to epiphytism carry at least some degree of phylogenetic signal (\ie, the tendency for closely related species to resemble each other).
 Therefore, a comparative approach incorporating phylogenetic relationships (and if possible, branch lengths) is appropriate \citep{Felsenstein1985}.
@@
 
-Furthermore, it should be possible to gain greater insight into the selective strength exerted by the epiphytic habitat by utilizing a community-based approach.
-If the relatively harsh abiotic conditions of the canopy act as a filter, we would expect to find lower functional richness at a community level, which should also result in decreased phylogenetic diversity under a scenario of phylogenetic trait conservatism \citep{Webb2002}.
-However, the relative strength of this environmental filter may also depend on overall abiotic conditions; for example, relative humidity and temperature are expected to vary not only between epiphytic and terrestrial environments, but also with elevation \citep{Cardelus2006}.
-Elevation is a strong predictor of fern species richness, with maximal richness tending to occur at mid-elevations, especially in the tropics \citep{Kessler2011a}.
-Functional diversity is also correlated with species richness (as few species tend to have exactly the same functional traits; \citealt{Schleuter2010}).
+It should be possible to gain greater insight into the selective strength exerted by the epiphytic habitat by utilizing a community-based approach.
+If the harsh abiotic conditions of the canopy act as a filter, we  expect to find lower functional richness at a community level, which should result in decreased phylogenetic diversity under a scenario of phylogenetic trait conservatism \citep{Webb2002}.
+However, the relative strength of this environmental filter may depend on overall abiotic conditions; for example, relative humidity (RH) and temperature also vary with elevation \citep{Cardelus2006}.
+Elevation is a strong predictor of fern species richness, with maximal richness occurring at mid-elevations, especially in the tropics \citep{Kessler2011a}.
+Functional diversity is also correlated with species richness (few species tend to have exactly the same functional traits; \citealt{Schleuter2010}).
 Thus, environmental variation along elevational gradients is expected to impact the functional, phylogenetic, and species diversity of epiphytic and terrestrial fern communities.
@@
 \end{enumerate}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% Methods %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Materials and Methods}
+
+\subsection*{Study Site}
+
+Moorea, French Polynesia (\ang{17;30;}S, \ang{149;50;}W) is a small (\SI{135}{\kilo\meter\squared}) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
+This distance acts as a strong barrier to dispersal \citep{Carson1995,Dassler2001}, which combined with the young age of the island (\circa \SI{1.5}{\ma}; \citealt{Duncan1976}) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders \sensu \citealt{PteridophytePhylogenyGroupI2016}).
+Ferns of Moorea include representatives of all five of the Cretaceous epiphytic radiations described by \cite{Schuettpelz2009} (\ie, polygrammoids, elaphoglossoids, asplenioids, vittarioids, and filmy ferns) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
+
+\subsection*{Community Survey}
+
+A fern community survey was carried out on Moorea comprising 17, \SI{10}{\meter} $\times$ \SI{10}{\meter} plots spanning an elevational gradient from \SIrange{20}{2000}{\meter}, with plots placed \circa every \SI{200}{\meter} \citep{Nitta2017}.
+Most plots were located along trails leading to the three main peaks on the island, Mt.
+Rotui (\SI{899}{\meter}), Mt.
+Mouaputa (\SI{890}{\meter}), and Mt.
+Tohiea (\SI{1206}{\meter}).
+Each plot was divided into 25, \SI{2}{\meter} $\times$ \SI{2}{\meter} subplots.
+Presence/absence of all sporophytes in each subplot was scored and summed to produce an abundance ranking for each species per plot from zero (not observed) to 25 (observed in all subplots).
+Epiphytic ferns were sampled by hand to \circa \SI{2}{\meter} on tree trunks or confirmed visually from the ground.
+Species were categorized as epiphytic or terrestrial based on field observations (see Trait Selection and Measurements below).
+The list of species in each plot was split into epiphytic and terrestrial groups and treated separately during data analysis, hereafter referred to as `epiphytic' and `terrestrial' communities.
+Vouchers were deposited at GH and UC.
+
+\subsection*{Environmental Survey}
+
+Temperature and RH of epiphytic and terrestrial habitats were measured using Hobo Pro v2 dataloggers with the RS3 Solar Radiation Shield (Onset Corporation, USA).
+A pair of dataloggers was installed for each plot, one mounted at \circa \SI{2}{\meter} on a tree (`epiphytic datalogger'), and one at ground level (`terrestrial datalogger'), except for the Mt.
+Tohiea \SI{1200}{\meter} plot, which only had an epiphytic datalogger.
+Temperature and RH were logged once every \SI{15}{\minute} during two survey periods: a preliminary survey from July 18, 2012 to July 7, 2013 at six plots, and a final survey from July 7, 2013 to July 5, 2014 at 15 plots.
+Two plots were not included in the environmental survey.
+We used the 2013–-2014 data for the analysis, with one exception: the terrestrial datalogger at the Mt.
+Tohiea \SI{393}{\meter} plot malfunctioned and was missing data for a substantial part of the final survey period (March 12--July 5, 2014), so we used data for the same dates from the preliminary survey period for this datalogger instead.
+Three other dataloggers also failed early during the final survey period, but lacked data from the preliminary period, so were excluded from analysis.
+Other days missing data for any datalogger during the final survey period were excluded from the dataset.
+The final dataset included 201 days of data for 13 epiphytic dataloggers and 13 terrestrial dataloggers ($\n = 26$).
+
+\subsection*{Trait Selection and Measurements}
+
+We selected several traits relevant to epiphytic growth (\tab \ref{table:traits} and \fig \ref{fig:traits}).
+We preferred morphological traits over physiological traits for our study because data for morphological traits are easier to gather at the scale of our study (\circa \SI{130}{\species} distributed over an elevational gradient from \SIrange{200}{2000}{\meter}).
+In addition, plant morphology and physiology are closely linked, and physiological traits such as rates of photosynthesis and water use efficiency are known to be correlated with morphological traits such as stipe length in ferns \citep{WatkinsJr.2010}.
+Studying the relationships between ecology and morphological traits is useful from a paleobotanical and climate change perspective, as trends in morphological trait variation in the fossil record can be used to gain insight into past climatic conditions \citep{Chaloner1997,McElwain2007a,McElwain2007b,Royer2008}.
+
+Morphological traits of sporophytes were measured from herbarium specimens for most species, or obtained from the literature when plant material was unavailable.
+Measurements were made directly on plants in the field for a few species with large fronds.
+A single leaf or rhizome per individual was measured on one to three separately collected individuals per species, and mean values were calculated for each species.
+Although trait values can vary within fern species with elevation \citep{Kessler2007a}, our study includes a broad phylogenetic sampling, so we assume intraspecific variation is outweighed by interspecies differences.
+To measure specific leaf area (SLA; \si{\square\meter\per\kilo\gram}), \circa 10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} \diameter punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
+For species with \textless \SI{2}{\milli\meter} between primary veins, leaf fragments excluding primary veins were obtained by dissection, dried at \SI{60}{\celsius} for \SI{24}{\hour}, weighed, and scanned.
+ImageJ \citep{Abramoff2004} software was used to measure area per fragment, and SLA calculated in the same way as leaf punches.
+
+Morphological traits of gametophytes were all qualitative (either categorical or binary; \tab \ref{table:traits}).
+Trait states were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding \citep{Nitta2017}, grown from spores of known species in the lab, or taken from the literature.
+For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie, heart-shaped) or non-cordate.
+Non-cordate morphotypes include elongate (strap or ribbon \sensu \citealt{Farrar2008}) and filamentous forms (\fig \ref{fig:traits}B).
+Gametophytes were cultivated on Bold's media \citep{Bold1957} supplemented with Nitch's micronutrients \citep{Nitsch1951}.
+
+Growth habit was coded as a binary trait (epiphytic or terrestrial) based on field observations.
+Some species do not fall clearly into either category (\eg, hemi-epiphytes; \citealt{Benzing1990,Dubuisson2003,Nitta2009}).
+We used connection to the soil as the criterion for defining binary growth habit; thus, epipetric plants were treated as epiphytic, and hemi-epiphytes and climbing plants were treated as terrestrial.
+For the purposes of this study, we treated growth habit as a fixed trait and used these categories to define epiphytic and terrestrial communities.
+Other studies have demonstrated that exceptional epiphytic growth of terrestrial fern species (and vice-versa) is  rare \citep{Cardelus2006,Kluge2006,WatkinsJr.2009}, and we do not believe that such rare exceptions would affect our analysis.
+
+\subsection*{Statistical Analysis}
+
+We calculated mean daily maximum, mean, minimum, and standard variation of temperature and RH from the raw climate data, then tested for differences in these summary statistics between epiphytic and terrestrial dataloggers using ANCOVA \citep{Scheffers2013}.
+We used daily mean values of temperature and RH each in turn as the response variable, and constructed linear models with growth habit as the factor and elevation as the covariate, and with their interaction to test for differences in response of climatic variables to elevation between epiphytic and terrestrial dataloggers.
+We also calculated the number of drying events at various minimum RH values, as well as daily mean values for water potential, but found that these were both highly correlated with mean RH and did not analyze them further (data not shown).
+
+For comparative analyses, we used the dated phylogenetic tree of \cite{Nitta2017}, which includes all species from Moorea and some from the neighboring island of Tahiti.
+We trimmed the tree to species occurring only on Moorea ($\n = 130$ \si{\species}) using the \function{drop.tip} function of the \package{ape} package \citep{Paradis2004} in R \citep{RCoreTeam2015}.
+We tested for phylogenetic signal in quantitative traits using two metrics that relate the observed trait values to those expected under a model of trait evolution following Brownian motion (BM), Blomberg's \K \citep{Blomberg2003} and Pagel's $\lambda$ \citep{Pagel1999}, with the \function{phylosig} function in the \package{phytools} R package \citep{Revell2012}.
+$\lambda$ is a scaling parameter that ranges from 0 (traits evolving randomly, as on a star phylogeny) to 1 (traits evolving along branches according to Brownian motion).
+\K describes the ratio between the amount of observed variance in traits vs.
+the amount of variance expected under a BM model of trait evolution: when $\K = 1$, traits are evolving according to BM; when $\K > 1$, traits are more conserved than expected under BM; when $\K < 1$, traits have less phylogenetic signal than expected under BM \citep{Blomberg2003}.
+We tested for significance in \K by comparing observed values of \K against a null distribution of \num{1000} trees with trait values shuffled randomly across the tips, and in $\lambda$ by conducting a likelihood ratio test comparing the log likelihoods of the observed value of $\lambda$ vs.
+$\lambda = 0$ (no phylogenetic signal).
+We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' \D \citep{Fritz2010}, which is designed to test for signal in binary traits.
+Values of \D range from 0 under a model of trait evolution by BM, and 1 for random distribution of traits; negative values (greater trait conservation than BM) and values greater than 1 (trait overdispersion) are possible.
+We analyzed \D using the \function{phylo.d} function in the \package{caper} R package \citep{Orme2013}, which also conducts a significance test by comparing the observed value of \D with distributions of simulated values produced under one of two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM threshold model.
+The probability of obtaining the observed value of \D under either model is then calculated.
+
+We used a multivariate approach to investigate differences in traits between epiphytes and terrestrial species.
+We used only quantitative traits, and included only species that had no missing data for any traits ($\n = 104$ \si{\species}).
+We first log-transformed traits as necessary to attain normality, then scaled each trait by its range.
+We subjected the transformed, scaled traits to a Principal Components Analysis (PCA) to reduce dimensionality.
+We conducted standard and phylogenetically corrected PCA, using the \function{PCA} and \function{phyl.pca} functions in the \package{FactoMineR} \citep{Le2008} and \package{phytools} R packages, respectively.
+The position of each species was plotted in the multi-dimensional trait space using the first two PC axes, then color-coded according to growth habit (epiphytic vs.
+terrestrial).
+
+We detected significant phylogenetic signal in most traits, so we tested for differences between traits of epiphytic vs.
+terrestrial species using methods that can test for significant differences between traits associated with a binary character while accounting for phylogeny.
+For quantitative traits, we used the \function{brunch} function in the \package{caper} package \citep{Orme2013}, which calculates phylogenetically independent contrasts in a quantitative character between alternative states of a binary trait.
+For qualitative (binary) traits, we used Pagel's test of correlated evolution \citep{Pagel1994} as implemented with the \function{fitPagel} function in the \package{phytools} package \citep{Revell2012}.
+This method compares the likelihood between two alternative models for a pair of binary traits.
+The first model tests the null hypothesis that rates of evolution of the two traits are independent.
+The second model tests the alternative hypothesis that the rates of evolution of one trait depends on the other.
+The likelihoods of obtaining the observed data are calculated for each model, then compared using a log likelihood test.
+A significantly better fit of the dependent model indicates that evolution of the two traits is correlated.
+To verify the effect of analysis method on our results, we performed alternative tests for correlation of growth habit with functional traits using phylogenetic generalized linear mixed models (PGLMMs; see Methods S1 and \tabs \ref{supp_table:PGLMM_sporo}, \ref{supp_table:PGLMM_gameto}).
+
+We analyzed community phylogenetic diversity using the Mean Phylogenetic Distance (MPD) and Mean Nearest Taxon Distance (MNTD) indices of \citep{Webb2000}.
+MPD is the mean phylogenetic distance of each species in a community to all other species in the same community; MNTD is the mean phylogenetic distance of each species in a community to its sister species in the same community.
+Communities with low MPD or MNTD are `phylogenetically clustered' (co-occurring species are closely related to each other), whereas those with high MPD or MNTD are `phylogenetically overdispersed' (co-occurring species are distantly related to each other).
+MPD therefore provides a measure of phylogenetic diversity across the whole tree, while MNTD is more sensitive to distances towards the tips.
+Alone, MPD and MNTD are not very informative; to assess their statistical significance, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{1000} randomized communities (equivalent to the Net Relatedness Index and Nearest Taxon Index, respectively, of \citealt{Webb2000} with the sign reversed).
+Observed values above 95\% or less than 5\% of the null distribution were considered significant.
+Selection of the randomization procedure can strongly bias outcomes \citep{Gotelli2006,Ulrich2012}.
+Our null hypothesis is that fern spores are easily capable of dispersal between terrestrial and epiphytic habitats, and across the relatively short distances separating our plots.
+We therefore constructed null communities by drawing species randomly from a regional pool including terrestrial and epiphytic taxa.
+The regional pool included all known species of ferns from Moorea, and all species from the neighboring island of Tahiti with genetic data available ($\n = 146$ \si{\species} total; \citealt{Nitta2017}).
+We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} \citep{Kembel2010} to calculate community phylogenetic diversity metrics.
+
+We analyzed community functional trait diversity using the \function{dbFD} function in the \package{FD} R package \citep{Laliberte2011}.
+This function takes as input values any number of traits for species in a given community, plots them in a multidimensional trait space using PCA, and calculates several summary statistics.
+Functional richness (FRic) is equivalent to convex hull volume containing all observed trait values plotted in multidimensional trait space \citep{Cornwell2006}, functional evenness (FEve) is the average distance to the center of all the traits, and functional dispersion (FDis) is the mean distance between traits \citep{Villeger2008}.
+We chose these three metrics because they are statistically independent and can each provide different information about the structure of community trait diversity \citep{Savage2012}.
+Functional richness does not use abundance data, but functional evenness and divergence can be weighted by abundance, so we preformed analyses for these metrics with abundance weighting to correct for the effect of rare species with unusual morphologies.
+Selection, transformation, and scaling of traits was performed similarly as for the PCA, except that the method used to calculate distance \citep{Gower1971} can accommodate missing data, so we included some species with observations missing for some traits ($\n = 126$ \si{\species} total).
+We calculated each metric separately for epiphytic and terrestrial communities, and compared results between the two.
+Under a scenario of environmental filtering (as hypothesized for epiphytic communities), we expect relatively lower values for the three metrics of functional diversity \citep{Savage2012}.
+We also investigated relationships between functional trait diversity and environment with linear models by testing for dependence of each functional trait diversity metric against elevation and daily minimum RH.
+We chose these independent variables because we found that overall climate tended to change with elevation, and to test if minimum RH specifically had a filtering effect on diversity along the gradient.
+
+Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% Results %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section*{Results}
+
+\subsection*{Environmental Survey}
+
+There were significant interaction effects between growth habit and elevation on all microclimatic variables except for minimum temperature (\tab \ref{supp_table:climate_model}).
+Slope and intercept for variables related to temperature were very similar between epiphytic and terrestrial dataloggers, but differed more strongly for those related to humidity (\tab \ref{supp_table:climate_model}, \fig \ref{fig:climate}).
+Minimum RH increased \Sexpr{get_result(slopes.results, "slope.epi", "min_RH", TRUE, 2)*100}\% per \SI{100}{\meter} elevation for epiphytic dataloggers, but was nearly flat (\Sexpr{get_result(slopes.results, "slope.ter", "min_RH", TRUE,2)*100}\% per \SI{100}{\meter}) for terrestrial dataloggers (linear model, $\rval^2 = \Sexpr{get_result(slopes.results, "r.squared", "min_RH")}$, $\pval < 0.001$).
+The Mt.
+Rotui \SI{830}{\meter} plot was a clear outlier, with greater mean temperature and lower minimum RH relative to other plots at similar elevation (outliers in  \fig \ref{fig:climate} at \circa \SI{800}{\meter}).
+Mt.
+Rotui is a solitary peak located between Cook's Bay and Oponohu Bay.
+It is more isolated with greater exposure than the other two mountains where the rest of the plots were established, which are part of a mountain complex on the interior of the island (J.
+Nitta, pers.
+obs.).
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%% SI Captions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Add S to table number and reset counter
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{table}{0}
+\clearpage
+
+% Table S1
+\begin{table}[ht]
+\caption{Phylogenetic generalized linear mixed models (PGLMMs) for quantitative (sporophyte) traits related to epiphytic growth.
+The response of each trait was modeled against growth habit using a PGLMM with phylogeny as a random effect.
+\pval for traits significantly correlated with growth habit in bold.}
+\label{supp_table:PGLMM_sporo}
+\end{table}
+
+% Table S2
+\begin{table}[ht]
+\caption{Phylogenetic generalized linear mixed models (PGLMMs) for binary (gametophyte) traits traits related to epiphytic growth.
+The response of each trait was modeled against growth habit using a PGLMM with phylogeny as a random effect.
+$\sigma^2$ is a measure of phylogenetic signal calculated from the variance in the PGLMM.
+Significance of $\sigma^2$ calculated using an approximate likelihood ratio test between a model with the observed value of $\sigma^2$ against a null model with $\sigma^2 = 0$.
+Traits significantly correlated with growth habit in bold.}
+\label{supp_table:PGLMM_gameto}
+\end{table}
+
+% Table S3
+\begin{table}[ht]
+\caption{Results of analysis of covariance (ANCOVA) for microclimatic variables between epiphytic and terrestrial dataloggers along an elevational gradient on Moorea, French Polynesia.
+Linear models were constructed for each response variable with growth habit as a fixed effect and elevation as a co-factor.
+Differences in intercept between growth habits was tested with models of the form $y \sim growth habit + elevation$.
+Differences in slope between growth habits were tested with models of the form $y \sim growth habit \times elevation$.
+Models were compared using the \function{aov} function in R.}
+\label{supp_table:climate_model}
+\end{table}
+
+% Table S4
+\begin{table}[ht]
+\caption{PC loadings for principal components analyses (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
+For each PC, values of the three traits with the greatest loadings (absolute values) shown in bold.}
+\label{supp_table:PCA}
+\end{table}
+
+% Table S5 
+\begin{table}[ht]
+\caption{List of all ferns from Moorea, French Polynesia with observed trait values and growth habit.
+For continuous trait values, means $\pm$ \stdev shown.
+For growth habit, E = epiphytic, T = terrestrial.
+For frond dissection, 1 = simple, 2 = pinnatifid or pinnatisect, 3 = 1-pinnate, 4 = 1- pinnate-pinnatifid, 5 = 2-pinnate, 6 = 2-pinnate-pinnatifid, 7 = 3- pinnate, 8 = 3-pinnate-pinnatifid, 9 = \textgreater 3-pinnate-pinnatifid, 10 = binpinnatifid or tripinnatifid.}
+\label{supp_table:traitvals}
+\end{table}
+
 \end{document}

### 2018-03-19 manuscript.Rnw
Paste in rest of results and discussion
Still not done going through and changing text results values to Sexpr to grab actual results values.

@@
 obs.).
+
+\subsection*{Principal Components Analysis of Trait Data}
+
+Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (\fig \ref{fig:PCA}C, D).
+The first two PC axes explained \Sexpr{get_result(PCA.results, "Std_PCA_Dim.2", "cumulative_variance", digits=3, signif=TRUE)}\% and \Sexpr{get_result(PCA.results, "Phy_PCA_Dim.2", "cumulative_variance", digits=3, signif=TRUE)}\% of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
+In both analyses, stipe length and frond width were loaded on PC1, and SLA, degree of frond dissection, and frond length were loaded on PC2 (\fig \ref{fig:PCA}A, B, \tab \ref{supp_table:PCA}).
+Results for number of pinna pairs differed between the two analyses, tending toward PC2 in the standard PCA and PC1 in the phylogenetic PCA (\tab \ref{supp_table:PCA}).
+
+\subsection*{Phylogenetic Signal}
+
+Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tab \ref{table:phylosig}, \fig \ref{fig:traits_tree}), and for quantitative traits, different results were obtained for $\lambda$ and K.
+When measured with $\lambda$, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of $\lambda$ close to 1; only number of pinna pairs had $\lambda$ close to zero (\tab \ref{table:phylosig}a).
+However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of $\K < 1$ (\tab \ref{table:phylosig}a).
+All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, $\D = \Sexpr{get_result(binary_phylosig.results, "D", "hairs", digits=2, signif=TRUE)}$) or more conserved (gemmmae $\D =  \Sexpr{get_result(binary_phylosig.results, "D", "gemmae", digits=2, signif=TRUE)}$; glands, $\D =  \Sexpr{get_result(binary_phylosig.results, "D", "glands", digits=2, signif=TRUE)}$; morphotype, $\D = \Sexpr{get_result(binary_phylosig.results, "D", "cordate_morph", digits=2, signif=TRUE)}$) than expected under a scenario of evolution by BM (\tab \ref{table:phylosig}b).
+Phylogenetic generalized linear mixed models (PGLMMs) also indicated significant signal in all gametophyte traits (\tab \ref{supp_table:PGLMM_gameto}).
+Growth habit showed phylogenetic signal nearly identical to that expected under BM ($\D = \Sexpr{get_result(binary_phylosig.results, "D", "epiphytic", digits=2, signif=TRUE)}$; \tab \ref{table:phylosig}b).
+
+\subsection*{Correlation of Traits with Growth Habit}
+
+Most sporophyte traits, including all those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (\tab \ref{table:PIC_sporo}).
+Epiphytic species tend to have smaller values for all traits (\tab \ref{table:PIC_sporo}; contrasts were made by subtracting epiphytic from terrestrial values).
+Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, $\pval = \Sexpr{get_result(PIC.results, "pval", "stipe", digits=2, signif=TRUE)}$), as well as frond width ($\pval = \Sexpr{get_result(PIC.results, "pval", "width", digits=2, signif=TRUE)}$) and degree of laminae dissection ($\pval = \Sexpr{get_result(PIC.results, "pval", "dissection", digits=2, signif=TRUE)}$).
+For the gametophyte traits, only morphotype (non-cordate vs.
+cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, $\pval = 0.0011$; \tab \ref{table:pagel_gameto}).
+Quantitatively similar results were obtained using PGLMMs, except that no significant relationship was observed between rhizome diameter and growth habit, whereas SLA did show a significant relationship with growth habit (\tabs \ref{supp_table:PGLMM_sporo}, \ref{supp_table:PGLMM_gameto}).
+
+\subsection*{Functional and Phylogenetic Diversity}
+
+On a single-trait basis, distribution of community-weighted mean values (CWMs) showed significant correlation with elevation for some traits, but no significant relationship was detected for any traits with minimum RH (\fig \ref{fig:CWM}A).
+SLA decreased with elevation for epiphytic (linear model, $\rval^2 = 0.6787$, $\pval <0.001$) and terrestrial (linear model, $\rval^2 = 0.8118$, $\pval <0.001$) communities.
+Degree of lamina dissection decreased with elevation in epiphytes (linear model, $\rval^2 = 0.3634$, $\pval = 0.0104$), but not terrestrial communities.
+Stipe length showed a very weak, but significant, increase with elevation for epiphytic communities (linear model, $\rval^2 = 0.2721$, $\pval = 0.0317$).
+Rhizome diameter decreased with elevation for terrestrial communities (linear model, $\rval^2 = 0.238$, $\pval = 0.047$), but increased slightly in epiphytes (linear model, $\rval^2 = 0.3525$, $\pval = 0.0120$).
+
+We observed a total of 104 species across all plots, including 48 epiphytic and 56 terrestrial species.
+Mean species richness was $11.12 \pm 4.6$ \si{\species} per plot for epiphytes and $12.35 \pm 5.0$ \si{\species} per plot for terrestrial species (not significantly different; one sided \tval-test, $\pval = 0.23$; \fig \ref{fig:FDIV}A).
+Species richness increased linearly with elevation and minimum RH for epiphytes, but not terrestrial communities (\fig \ref{fig:FDIV}B).
+Trait richness (FRic) was lower in epiphytic vs.
+terrestrial communities overall (one sided \tval-test, $\pval < 0.001$), whereas functional evenness (FEve) and diversity (FDiv) did not differ between the two.
+FEve decreased with elevation, but not minimum RH, in terrestrial communities (\fig \ref{fig:FDIV}A).
+Epiphytic communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, $\rval^2 = 0.5836$, $\pval < 0.001$) and plots with greater minimum RH (linear model, $\rval^2 = 0.6626$, $\pval = 0.009$).
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (one sided \tval-test, $\pval < 0.001$) but not MPD ($\pval = 0.30$; \fig \ref{fig:FDIV}B).
+No significant trend was detected in phylogenetic community structure with environment in either terrestrial or epiphytic communities (\fig \ref{fig:FDIV}A).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% Discussion %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section*{Discussion}
+
+Here, we show for the first time using a broadly sampled phylogenetic dataset that morphological traits related to epiphytic growth are strongly evolutionarily conserved in ferns.
+The canalized morphology of epiphytic ferns suggests that this habitat strongly filters for a specialized growth form.
+We find that communities of epiphytic ferns are more clustered functionally and phylogenetically relative to terrestrial species, again supporting the role of environmental filtering in determining epiphytic fern community structure.
+Our results are similar between gametophyte and sporophyte traits, suggesting that morphology of each life stage is important for epiphytic growth.
+
+\subsection*{Effect of Phylogeny on Analysis of Epiphytic Traits}
+
+Although several other studies have addressed the relationship of functional traits with growth habit in ferns \citep[\eg, ][]{Hietz1998,Kluge2007,WatkinsJr.2010}, few have accounted for evolutionary history.
+We find that growth habit shows phylogenetic signal, as do several of the traits associated with it.
+In one of the few previous studies to analyze phylogenetic signal in functional traits of ferns, \citet{Zhang2014b} found that only one of 16 traits showed significant phylogenetic signal.
+The only trait in common between the current study and that of \citet{Zhang2014b} is SLA, which did not show phylogenetic signal in either study.
+Given the high degree of phylogenetic signal we observe in most traits and that at least some of the traits between our two studies are likely to be correlated, it is surprising that \citet{Zhang2014b} did not detect phylogenetic signal in more traits.
+However, \citet{Zhang2014b} only used \K, which tended to indicate less phylogenetic signal than $\lambda$ in our study.
+In simulation studies, $\lambda$ has lower error (both Type I and Type II) than \K, which can be inconsistent unless the strength of BM is great \citep{Munkemuller2012}.
+Our results indicate the importance of including a phylogenetic component, and including multiple measures of phylogenetic signal, when analyzing traits related to epiphytic growth in ferns.
+
+\subsection*{Sporophyte Traits}
+
+The sporophyte traits that differed most strongly between epiphytic and terrestrial taxa were related to overall size, including stipe and frond length, frond width, and rhizome diameter.
+Stipe length in particular differed strongly between epiphytes and terrestrial species (\fig\ref{fig:CWM}B, \tab\ref{table:PIC_sporo}).
+Stipe length may be a critical trait controlling whole leaf hydraulic conductance; since ferns have resistive stipes in general, species growing in water-limited habitats are expected to decrease stipe length to minimize resistance \citep{WatkinsJr.2010}.
+Our results support previous non-phylogenetically controlled studies that found a prevalence of short stipes in epiphytic ferns \citep{WatkinsJr.2010,Creese2011}.
+However, given that many of these size-related traits scale together in ferns \citep{Arcand2008,Creese2011}, careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
+
+Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (\tab\ref{table:PIC_sporo}, \fig\ref{fig:CWM}B).
+SLA is an important part of the leaf economic spectrum, which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
+slow-growing species with low SLA, high nutrient content, and long life spans \citep{Wright2004a}.
+We hypothesized that epiphytes would have lower SLA, given their more extreme growing conditions.
+Indeed, the species with the lowest SLA values were epiphytes (\eg, \textit{Selliguea plantaginea} Brack., \SI{8.7}{\square\meter\per\kilo\gram}; \textit{Davallia anderssonii} Mett.
+ex Kuhn, \SI{8.43}{\square\meter\per\kilo\gram}), but this relationship was not significant across the dataset (\tab\ref{table:PIC_sporo}, \fig\ref{fig:CWM}).
+The tradeoffs posited in the leaf economics spectrum have been most thoroughly documented in global studies including the widest possible range of taxa and environments, and do not always hold up in local studies \citep{Petter2016}.
+For example, \citet{Zhu2016} did not find a significant relationship between SLA and leaf life span in a survey of 16 ferns across a disturbance (light) gradient in China.
+It is possible that other small-scale effects (\eg, microclimate) obscure this pattern in local studies such as ours and that of \citet{Zhu2016}.
+This is supported by the strong, parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (\fig\ref{fig:CWM}A).
+
+\subsection*{Gametophyte Traits}
+
+Only gametophyte morphotype was found to have a significant association with epiphytic growth.
+Morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived ($< \SI{1}{\year}$), cordate gametophytes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate (\eg, elongate or filamentous) gametophytes that persist over multiple growing seasons and produce sporophytes more slowly \citep{WatkinsJr.2007a,Farrar2008}.
+The complex, three-dimensional structure of non-cordate gametophytes slows their drying rates, and may be an adaptation to drought-prone epiphytic habitats \citep{WatkinsJr.2007,Pittermann2013}, and promote out-crossing \citep{Farrar2008}.
+Our study is the first to our knowledge to demonstrate that the correlation between non-cordate morphology and epiphytism in ferns is significant while controlling for phylogeny.
+\cite{Sundue2015} found that non-cordate morphology is significantly correlated with increased diversification rates within Polypodiaceae, consistent with the hypothesis that this trait promoted diversification within the epiphytic niche.
+
+We did not find that gemmae production was linked with epiphytism.
+Gemmae are asexual propagules that allow gametophytes to persist over multiple growing seasons and attain large population sizes \citep{Farrar2008}.
+Gemmae production should be a useful trait in an epiphytic context because it may allow gametophytes to increase their surface area and hedge against population loss due to extreme drying events \citep{Farrar2008}.
+Gemmae can also function sexually by producing antherida in the presence of antheridiogen secreted by other gametophytes \citep{Emigh1977}, which may confer a selective advantage in the epiphytic environment by enabling outcrossing \citep{Dassler2001}.
+It is possible that some, but not all, epiphytic ferns rely on such strategies, and that they are not under similarly strong selection as other traits such as stipe length and gametophyte morphotype.
+Interestingly, when we classified `widespread' species as those with gametophytes occurring more than \SI{200}{\meter} above or below the minimum or maximum elevational range of conspecific sporophytes (data not shown), we found that gemmae production is correlated with widespread growth (log likelihood dependent model = \num{-54.5}, log likelihood independent model = \num{-63.6}, $\pval = 0.001$, Pagel's test of correlated evolution).
+Thus, it seems that gemmae do function in long-term gametophyte persistence, but may not be required for epiphytic growth.
+
+Presence of hairs and glands has strong phylogenetic signal (\tabs\ref{table:phylosig}, \ref{supp_table:PGLMM_gameto}), and these traits are correlated with certain taxonomic groups \citep{Stokey1951,Nayar1971}.
+Although hairs have been hypothesized to reduce rates of drying by increasing the boundary layer and holding external water \citep{WatkinsJr.2007}, we did not detect a correlation between hair production and epiphytic growth.
+Similarly, production of glands does not appear to be correlated with epiphytic growth.
+Wax-secreting glands are prevalent in both sporophytes and gametophytes of many notholaenid ferns, a xeric-adapted clade, but their functional significance has not been established \citep{Johnson2012}.
+Further tests are needed to determine if such glands serve a functional role in preventing water loss in desert ferns and epiphytes.
+
+\subsection*{Environmental Filtering}
+
+We observed the predicted trend of decreased trait richness (FRic) in epiphytic relative to terrestrial communities overall, and a significant positive correlation of FRic with humidity in epiphytes (\ie, decreased FRic at drier, low elevation plots and increased FRic at more humid, high elevation plots; \fig\ref{fig:FDIV}A).
+However, we did not observe similar patterns in the other two measures of functional diversity, evenness (FEve) and divergence (FDiv) (\fig\ref{fig:FDIV}B).
+Indeed, FEve decreased with elevation (but not minimum RH) in terrestrial communities (\fig\ref{fig:FDIV}A).
+Overall, the lack of a clear pattern in functional evenness and divergence suggests that within each trait pool, there is no significant difference in trait spacing between epiphytes and terrestrial species.
+This could be maintained by niche partitioning within epiphytic and terrestrial habitats.
+Epiphytic habitats are highly heterogeneous, and microclimatic conditions vary with height along the host and distance from the trunk \citep{Hietz1995}.
+The vertical stratification of epiphytic species into different zones along height gradients in host trees is well documented \citep{Zotz2007}.
+In a study comparing terrestrial and epiphytic ferns along an elevational gradient from \SIrange{30}{3000}{\meter} in Costa Rica, \citet{WatkinsJr.2006a} found that epiphytes tended to sort into low-trunk and canopy zones, with only 18\% found in both, and no species that overlapped between canopy and terrestrial habitats.
+It is possible that epiphytic growth per se limits overall morphological variation in epiphytic ferns, but that species with different traits are then sorted into different niches.
+
+In addition to their decreased functional diversity, we also found that epiphytic communities are more phylogenetically clustered relative to terrestrial communities overall (\fig\ref{fig:FDIV}B).
+However, we did not detect significant trends in phylogenetic community structure of epiphytic and terrestrial communities along the elevational gradient (\fig\ref{fig:FDIV}A).
+It is possible that the scale of the current study (from \SIrange{200}{1200}{\meter}) prevented the detection of such trends.
+When the plots in this study were analyzed together with additional plots from Tahiti reaching a maximum altitude of just over \SI{2000}{\meter}, \citet{Nitta2017}  detected increasing phylogenetic clustering in fern sporophyte communities (epiphytes and terrestrial species together) with elevation.
+\cite{Kluge2011} also compared phylogenetic and trait diversity between epiphytic and terrestrial ferns over an elevational gradient from \SIrange{100}{3400}{\meter} in Costa Rica, and found trait overdispersion at mid-elevations for epiphytes but no trends in phylogenetic diversity for either group.
+They accounted this to evolutionary plasticity in traits, but did not quantify the degree of phylogenetic signal in the traits they studied.
+
+\subsection*{Conclusion and Future Directions}
+
+We have demonstrated that growth habit itself, and several traits associated with it, show significant phylogenetic signal in ferns.
+Furthermore, we find that epiphytes have restricted morphologies both on a per-species basis and at the community level, and that epiphytic communities are phylogenetically clustered relative to terrestrial ones.
+Finally, we observed drier conditions in the canopy relative to terrestrial habitats.
+Taken together, our results support a scenario of environmental filtering acting on conserved traits to limit functional and phylogenetic diversity in epiphytic ferns.
+The generality of this pattern could be further tested by investigating the functional traits of other groups with large epiphytic radiations such as orchids and bromeliads.
+
+Recent studies \citep[\eg,][]{Sundue2015,Testo2016a,Lehtonen2017} have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by \cite{Schuettpelz2009} by failing to detect increased rates of diversification associated with epiphytic growth.
+Epiphytism itself is likely a complex trait requiring the evolution of multiple associated traits, and the timing and order of the evolution of such traits may be related to the complex diversification dynamics observed in these studies.
+Our study identifies several conserved morphological traits (\eg, stipe length, frond dissection, gametophyte morphotype) that may belong to this suite of epiphytic traits.
+We hope future studies will build on our results by analyzing these traits across much more densely sampled phylogenies \citep[\eg][]{Testo2016a} to dissect how each may be related to diversification and growth habit.
+Additionally, future studies should focus on determining in more detail how niche space is partitioned within epiphytic vs.
+terrestrial habitats, which is clearly a major dividing line for ferns.
 

### 2018-03-21 manuscript.Rnw
Change hard-coded results values to Sexpr for almost all results and discussion
Add mean_trait_values to analyses. Still need to fix sla for Selliguea plantaginea

@@
 source(file = "bin/functional_diversity.R")
+source(file = "bin/mean_trait_values.R") # this must be last to keep sla.grand.mean
 save(list = ls(all = TRUE), file= "results.RData")
@@
 
-<<get_result_func, echo=FALSE>>=
+<<get_result_func>>=
 # simple function to get rounded results with less typing
-get_result <- function (results_table, col_name, row_name, signif=FALSE, digits=3) {
+library(dplyr) # use to filter df prior to get_result in some cases
+get_result <- function (results_table, col_name, row_name=1, signif=FALSE, digits=3) {
   if (signif == FALSE) {
@@
 }
+
 @
@@
 For the gametophyte traits, only morphotype (non-cordate vs.
-cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, $\pval = 0.0011$; \tab \ref{table:pagel_gameto}).
+cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, $\pval = \Sexpr{get_result(pagel.results, "pval", "morph_binary", digits=2, signif=TRUE)}$; \tab \ref{table:pagel_gameto}).
 Quantitatively similar results were obtained using PGLMMs, except that no significant relationship was observed between rhizome diameter and growth habit, whereas SLA did show a significant relationship with growth habit (\tabs \ref{supp_table:PGLMM_sporo}, \ref{supp_table:PGLMM_gameto}).
@@
 
+<<calculate_richness>>=
+### calculate number of species observed in plots total and by growth habit
+# use only sites on Moorea
+sites <- mooreaferns::sites[grep("Aorai", mooreaferns::sites$site, invert=TRUE), ]
+# load untransformed traits
+traits <- mooreaferns::fern_traits
+# load community data
+comm <- mooreaferns::sporocomm
+# trim community data to only moorea sites
+comm <- comm[rownames(comm) %in% sites$site, ]
+# keep only species with at least one observation
+comm <- comm[, colSums(comm) > 0]
+# trim trait data to only species with at least one observation
+traits <- traits[traits$species %in% colnames(comm), ]
+@
+
 On a single-trait basis, distribution of community-weighted mean values (CWMs) showed significant correlation with elevation for some traits, but no significant relationship was detected for any traits with minimum RH (\fig \ref{fig:CWM}A).
-SLA decreased with elevation for epiphytic (linear model, $\rval^2 = 0.6787$, $\pval <0.001$) and terrestrial (linear model, $\rval^2 = 0.8118$, $\pval <0.001$) communities.
-Degree of lamina dissection decreased with elevation in epiphytes (linear model, $\rval^2 = 0.3634$, $\pval = 0.0104$), but not terrestrial communities.
-Stipe length showed a very weak, but significant, increase with elevation for epiphytic communities (linear model, $\rval^2 = 0.2721$, $\pval = 0.0317$).
-Rhizome diameter decreased with elevation for terrestrial communities (linear model, $\rval^2 = 0.238$, $\pval = 0.047$), but increased slightly in epiphytes (linear model, $\rval^2 = 0.3525$, $\pval = 0.0120$).
+SLA decreased with elevation for epiphytic (linear model, $\rval^2 = \Sexpr{get_result(filter(univar.el.results, variable=="sla", habit=="epiphytic"), "r.squared.plain")}$, $\pval = \Sexpr{get_result(filter(univar.el.results, variable=="sla", habit=="epiphytic"), "p", signif=TRUE)}$) and terrestrial (linear model, $\rval^2 = \Sexpr{get_result(filter(univar.el.results, variable=="sla", habit=="terrestrial"), "r.squared.plain")}$, $\pval = \Sexpr{get_result(filter(univar.el.results, variable=="sla", habit=="terrestrial"), "p", signif=TRUE)}$) communities.
+Degree of lamina dissection decreased with elevation in epiphytes (linear model, $\rval^2 = \Sexpr{get_result(filter(univar.el.results, variable=="dissection", habit=="epiphytic"), "r.squared.plain")}$, $\pval = \Sexpr{get_result(filter(univar.el.results, variable=="dissection", habit=="epiphytic"), "p", signif=TRUE)}$), but not terrestrial communities.
+Stipe length showed a small, but significant, increase with elevation for epiphytic communities (linear model, $\rval^2 = \Sexpr{get_result(filter(univar.el.results, variable=="stipe", habit=="epiphytic"), "r.squared.plain")}$, $\pval = \Sexpr{get_result(filter(univar.el.results, variable=="stipe", habit=="epiphytic"), "p", signif=TRUE)}$).
+Rhizome diameter decreased with elevation for terrestrial communities (linear model, $\rval^2 = \Sexpr{get_result(filter(univar.el.results, variable=="rhizome", habit=="terrestrial"), "r.squared.plain")}$, $\pval = \Sexpr{get_result(filter(univar.el.results, variable=="rhizome", habit=="terrestrial"), "p", signif=TRUE)}$), but increased slightly in epiphytes (linear model, $\rval^2 = \Sexpr{get_result(filter(univar.el.results, variable=="rhizome", habit=="epiphytic"), "r.squared.plain")}$, $\pval = \Sexpr{get_result(filter(univar.el.results, variable=="rhizome", habit=="epiphytic"), "p", signif=TRUE)}$).
 
-We observed a total of 104 species across all plots, including 48 epiphytic and 56 terrestrial species.
-Mean species richness was $11.12 \pm 4.6$ \si{\species} per plot for epiphytes and $12.35 \pm 5.0$ \si{\species} per plot for terrestrial species (not significantly different; one sided \tval-test, $\pval = 0.23$; \fig \ref{fig:FDIV}A).
+We observed a total of \Sexpr{ncol(comm)} species across all plots, including \Sexpr{length(traits$habit[traits$habit == "epiphytic"])} epiphytic and \Sexpr{length(traits$habit[traits$habit == "terrestrial"])} terrestrial species.
+Mean species richness was $\Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="epiphytic"), "mean", signif=TRUE)} \pm \stdev \Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="epiphytic"), "sd", signif=TRUE)}$ \si{\species} per plot for epiphytes and $\Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="terrestrial"), "mean", signif=TRUE)} \pm \stdev \Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="terrestrial"), "sd", signif=TRUE)}$ \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, $\pval = \Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="terrestrial"), "pval", signif=TRUE)}$; \fig \ref{fig:FDIV}A).
 Species richness increased linearly with elevation and minimum RH for epiphytes, but not terrestrial communities (\fig \ref{fig:FDIV}B).
 Trait richness (FRic) was lower in epiphytic vs.
-terrestrial communities overall (one sided \tval-test, $\pval < 0.001$), whereas functional evenness (FEve) and diversity (FDiv) did not differ between the two.
+terrestrial communities (two-sided \tval-test, $\pval = \Sexpr{get_result(filter(multvar.t.results, resp_var=="FRic", habit=="terrestrial"), "pval", signif=TRUE)}$), whereas functional diversity (FDiv) was lower in terrestrial communities (two-sided \tval-test, $\pval = \Sexpr{get_result(filter(multvar.t.results, resp_var=="FDiv", habit=="terrestrial"), "pval", signif=TRUE)}$), and functional evenness (FEve) did not differ between the two.
 FEve decreased with elevation, but not minimum RH, in terrestrial communities (\fig \ref{fig:FDIV}A).
-Epiphytic communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, $\rval^2 = 0.5836$, $\pval < 0.001$) and plots with greater minimum RH (linear model, $\rval^2 = 0.6626$, $\pval = 0.009$).
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (one sided \tval-test, $\pval < 0.001$) but not MPD ($\pval = 0.30$; \fig \ref{fig:FDIV}B).
+Epiphytic communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, $\rval^2 = \Sexpr{get_result(filter(multvar.el.results, habit=="epiphytic", variable=="FRic"), "r.squared.plain", signif=TRUE)}$, $\pval = \Sexpr{get_result(filter(multvar.el.results, habit=="epiphytic", variable=="FRic"), "p", signif=TRUE, digits=1)}$) and plots with greater minimum RH (linear model, $\rval^2 = \Sexpr{get_result(filter(multvar.env.results, habit=="epiphytic", variable=="FRic"), "r.squared.plain")}$, $\pval = \Sexpr{get_result(filter(multvar.env.results, habit=="epiphytic", variable=="FRic"), "p")}$).
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (two-sided \tval-test, $\pval = \Sexpr{get_result(filter(multvar.t.results, habit=="terrestrial", resp_var=="mntd.obs.z"), "pval", signif=TRUE)}$) but not MPD ($\pval = \Sexpr{get_result(filter(multvar.t.results, habit=="terrestrial", resp_var=="mpd.obs.z"), "pval", signif=TRUE, digits=2)}$; \fig \ref{fig:FDIV}B).
 No significant trend was detected in phylogenetic community structure with environment in either terrestrial or epiphytic communities (\fig \ref{fig:FDIV}A).
@@
 We hypothesized that epiphytes would have lower SLA, given their more extreme growing conditions.
-Indeed, the species with the lowest SLA values were epiphytes (\eg, \textit{Selliguea plantaginea} Brack., \SI{8.7}{\square\meter\per\kilo\gram}; \textit{Davallia anderssonii} Mett.
-ex Kuhn, \SI{8.43}{\square\meter\per\kilo\gram}), but this relationship was not significant across the dataset (\tab\ref{table:PIC_sporo}, \fig\ref{fig:CWM}).
+Indeed, the species with the lowest SLA values were epiphytes (\eg, \textit{Selliguea plantaginea} Brack., \SI{\Sexpr{as.numeric(get_result(filter(sla.grand.mean, species=="Selliguea_plantaginea"), "mean", signif=TRUE))}}{\square\meter\per\kilo\gram}; \textit{Ctenopterella blechnoides} (Grev.) Parris, \SI{\Sexpr{as.numeric(get_result(filter(sla.grand.mean, species=="Ctenopterella_blechnoides"), "mean", signif=TRUE))}}{\square\meter\per\kilo\gram}), but this relationship was not significant across the dataset (\tab\ref{table:PIC_sporo}, \fig\ref{fig:CWM}).
 The tradeoffs posited in the leaf economics spectrum have been most thoroughly documented in global studies including the widest possible range of taxa and environments, and do not always hold up in local studies \citep{Petter2016}.
@@
 
-We observed the predicted trend of decreased trait richness (FRic) in epiphytic relative to terrestrial communities overall, and a significant positive correlation of FRic with humidity in epiphytes (\ie, decreased FRic at drier, low elevation plots and increased FRic at more humid, high elevation plots; \fig\ref{fig:FDIV}A).
-However, we did not observe similar patterns in the other two measures of functional diversity, evenness (FEve) and divergence (FDiv) (\fig\ref{fig:FDIV}B).
-Indeed, FEve decreased with elevation (but not minimum RH) in terrestrial communities (\fig\ref{fig:FDIV}A).
-Overall, the lack of a clear pattern in functional evenness and divergence suggests that within each trait pool, there is no significant difference in trait spacing between epiphytes and terrestrial species.
+Results for functional trait richness (FRic) supported our hypotheses: we observed lower FRic in epiphytic relative to terrestrial communities overall, and a significant positive correlation of FRic with humidity in epiphytes (\ie, decreased FRic at drier, low elevation plots and increased FRic at more humid, high elevation plots; \fig\ref{fig:FDIV}A).
+However, results for the other two measures of functional diversity, evenness (FEve) and divergence (FDiv) (\fig\ref{fig:FDIV}B) did not match our predictions.
+Rather, FEve decreased with elevation (but not minimum RH) in terrestrial communities, and FDiv was lower in terrestrial communities overall while showing no trend along the gradient (\fig\ref{fig:FDIV}A).
+The lack of a clear pattern in functional evenness and divergence suggests that within each trait pool, there is no significant difference in trait spacing between epiphytes and terrestrial species.
 This could be maintained by niche partitioning within epiphytic and terrestrial habitats.

### 2018-03-21 bin/mean_trait_values.R

@@ -1,0 +1,102 @@
+library(tidyverse)
+library(mooreaferns)
+
+setwd(here::here())
+
+###########
+### SLA ###
+###########
+
+# load raw measurements, including multiple measurements per individual
+sla.raw <- mooreaferns::sla.raw
+
+# Nitta 2543 Ptisana salicina is in ERROR, should be Nitta 2544.
+# Already have measurements for 2544 P. salicina, so it must have been measured twice.
+# Exclude 2544 P. salicina
+sla.raw <- sla.raw[!(sla.raw$specimen=="Nitta_2544" & sla.raw$species=="Ptisana_salicina"), ]
+
+# calculate mean SLA for each individual
+sla.mean <- 
+  sla.raw %>%
+  group_by(specimen) %>%
+  summarize(sla = mean(sla) , sd = sd(sla), n = n()) %>%
+  left_join(unique(select(sla.raw, species, specimen)))
+
+# calculate grand mean for each species based on individual means
+sla.grand.mean <- 
+  sla.mean %>%
+  group_by(species) %>%
+  summarize(mean = mean(sla) , sd = sd(sla), n = n())
+  
+sla.grand.mean$trait <- "sla"
+
+####################
+### other traits ###
+####################
+
+# load raw measurements, in long format
+# includes only one measurment per individual but multiple individuals per species
+morph.long <- 
+  mooreaferns::morph.raw %>%
+  as_tibble %>%
+  select (-source, -specimen) %>%
+  gather(key = trait, value = measurement, -species) %>%
+  na.omit()
+
+# calculate means
+morph.mean <- 
+  morph.long %>%
+  group_by(species, trait) %>%
+  summarise(mean = mean(measurement) , sd = sd(measurement), n = n())
+rm(morph.long)
+
+# merge with sla grand means
+morph.mean <- bind_rows(morph.mean, sla.grand.mean)
+
+# split means into numeric and integer (only number of pinna pairs) measurements
+morph.numeric <- filter(morph.mean, trait == "frond.length" | trait == "lamina.width" | trait == "rhizome.dia" | trait == "stipe.length" | trait == "sla") 
+morph.int <- filter(morph.mean, trait == "pinna.pairs")
+
+# set number of signif digits differently for each
+# numeric measurements made with normal ruler in cm, so should have to 2 digits after 0
+# need to use formatC to keep trailing zeros
[…]

### 2018-03-21 bin/functional_diversity.R
Fi typo in univar.el.results

@@ -651,7 +651,7 @@ right_plots <- plot_scatter_dataset (resp_vars, compare.df, "el", xlabel="Elevat
 
 # also output linear model results in dataframe for MS
 univar.env.results <- plot_scatter_dataset (resp_vars, compare.env, "min_RH", xlabel="Minimum RH (%)", ylabels=my.ylabels)[[2]]
-univar.el.results <- plot_scatter_dataset (resp_vars, compare.env, "min_RH", xlabel="Minimum RH (%)", ylabels=my.ylabels)[[2]]
+univar.el.results <- plot_scatter_dataset (resp_vars, compare.df, "el", xlabel="Elevation (m)", ylabels=my.ylabels)[[2]]
 
 ### make various adjustments to plots
 

### 2018-03-23 manuscript.Rnw
Fixes to mean_trait_values
Move sourcing of mean_trait_values to top of analyses list, add gameto_source to mean_trait_values, fix caption in supporting information Table S5

@@
 setwd(here::here())
+source(file = "bin/mean_trait_values.R")
 source(file = "bin/microclimate.R")
@@
 source(file = "bin/functional_diversity.R")
-source(file = "bin/mean_trait_values.R") # this must be last to keep sla.grand.mean
 save(list = ls(all = TRUE), file= "results.RData")
+@
+
+<<calculate_richness>>=
+### calculate number of species observed in plots total and by growth habit
+# use only sites on Moorea
+sites <- mooreaferns::sites[grep("Aorai", mooreaferns::sites$site, invert=TRUE), ]
+# load untransformed traits
+traits <- mooreaferns::fern_traits
+# load community data
+comm <- mooreaferns::sporocomm
+# trim community data to only moorea sites
+comm <- comm[rownames(comm) %in% sites$site, ]
+# keep only species with at least one observation
+comm <- comm[, colSums(comm) > 0]
+# trim trait data to only species with at least one observation
+traits <- traits[traits$species %in% colnames(comm), ]
 @
@@
 \item The appearance of angiosperm-dominated forests in the Cretaceous and early Cenozoic had a profound effect on terrestrial biota by creating novel ecological niches.
-The diversity of modern fern lineages are thought to be a response to angiosperm dominance, radiating both in the canopy and on the forest floor.
+The diversity of modern fern lineages is thought to be a response to angiosperm dominance, radiating both in the canopy and on the forest floor.
 Despite the significance of epiphytic growth for fern evolution, however, few studies have investigated traits related to epiphytism in ferns while accounting for phylogeny.
@@
 
-<<calculate_richness>>=
-### calculate number of species observed in plots total and by growth habit
-# use only sites on Moorea
-sites <- mooreaferns::sites[grep("Aorai", mooreaferns::sites$site, invert=TRUE), ]
-# load untransformed traits
-traits <- mooreaferns::fern_traits
-# load community data
-comm <- mooreaferns::sporocomm
-# trim community data to only moorea sites
-comm <- comm[rownames(comm) %in% sites$site, ]
-# keep only species with at least one observation
-comm <- comm[, colSums(comm) > 0]
-# trim trait data to only species with at least one observation
-traits <- traits[traits$species %in% colnames(comm), ]
-@
-
 On a single-trait basis, distribution of community-weighted mean values (CWMs) showed significant correlation with elevation for some traits, but no significant relationship was detected for any traits with minimum RH (\fig \ref{fig:CWM}A).
@@
 
-The sporophyte traits that differed most strongly between epiphytic and terrestrial taxa were related to overall size, including stipe and frond length, frond width, and rhizome diameter.
+The sporophyte traits that differed most strongly between epiphytic and terrestrial taxa were related to plant size, including stipe and frond length, frond width, and rhizome diameter.
 Stipe length in particular differed strongly between epiphytes and terrestrial species (\fig\ref{fig:CWM}B, \tab\ref{table:PIC_sporo}).
@@
 We hypothesized that epiphytes would have lower SLA, given their more extreme growing conditions.
-Indeed, the species with the lowest SLA values were epiphytes (\eg, \textit{Selliguea plantaginea} Brack., \SI{\Sexpr{as.numeric(get_result(filter(sla.grand.mean, species=="Selliguea_plantaginea"), "mean", signif=TRUE))}}{\square\meter\per\kilo\gram}; \textit{Ctenopterella blechnoides} (Grev.) Parris, \SI{\Sexpr{as.numeric(get_result(filter(sla.grand.mean, species=="Ctenopterella_blechnoides"), "mean", signif=TRUE))}}{\square\meter\per\kilo\gram}), but this relationship was not significant across the dataset (\tab\ref{table:PIC_sporo}, \fig\ref{fig:CWM}).
+Indeed, the species with the lowest SLA values were epiphytes (\eg, \textit{Ctenopterella blechnoides} (Grev.) Parris, \SI{\Sexpr{as.numeric(get_result(filter(sla.grand.mean.results, species=="Ctenopterella_blechnoides"), "mean", signif=TRUE))}} {\square\meter\per\kilo\gram}; \textit{Selliguea plantaginea} Brack., \SI{\Sexpr{as.numeric(get_result(filter(sla.grand.mean.results, species=="Selliguea_plantaginea"), "mean", signif=TRUE))}} {\square\meter\per\kilo\gram}), but this relationship was not significant across the dataset (\tab\ref{table:PIC_sporo}, \fig\ref{fig:CWM}).
 The tradeoffs posited in the leaf economics spectrum have been most thoroughly documented in global studies including the widest possible range of taxa and environments, and do not always hold up in local studies \citep{Petter2016}.
@@
 \begin{table}[ht]
-\caption{List of all ferns from Moorea, French Polynesia with observed trait values and growth habit.
+\caption{Trait values and growth habit of ferns from Moorea, French Polynesia.
 For continuous trait values, means $\pm$ \stdev shown.

### 2018-03-23 supporting_information.Rnw
Add Pagel's correlated evo test to manuscript
Also, change font to TNR in MS and SI
Fixes to mean_trait_values
Move sourcing of mean_trait_values to top of analyses list, add gameto_source to mean_trait_values, fix caption in supporting information Table S5

@@
 @
+
+% Table S5 
+\begin{table}[ht]
+\caption{Trait values and growth habit of ferns from Moorea, French Polynesia.
+For continuous trait values, means $\pm$ \stdev shown.
+Growth habit was coded as a binary character (terrestrial or epiphytic).
+In case of intermediate growth habits, connection to the soil was used as the criterion for selecting the character state; \eg, epipetric plants were treated as epiphytic, and hemi-epiphytes and climbing plants were treated as terrestrial.
+For frond dissection, 1 = simple, 2 = pinnatifid or pinnatisect, 3 = 1-pinnate, 4 = 1- pinnate-pinnatifid, 5 = 2-pinnate, 6 = 2-pinnate-pinnatifid, 7 = 3-pinnate, 8 = 3-pinnate-pinnatifid, 9 = \textgreater 3-pinnate-pinnatifid, 10 = binpinnatifid or tripinnatifid.
+Morphotype and presence/absence of gemmae, glands, and hairs are traits of gametophytes; other traits are from sporophytes.
+For sporophyte data sources, "M" indicates measurements made on specimens collected for this study; 1 = \cite{NationalMuseumofNatureandScience2008}, 2= \cite{Copeland1932}, 3= \cite{Palmer2003}, 4 = \cite{RoyalBotanicGardensandDomainTrust}, 5 = \cite{Brownsey1987}}.
+For gametophyte data sources, "L" indicates observations of gametophytes cultured from spores in the lab; "F" indicates observations of gametophytes in the field and identified to species using DNA barcodes \citep{Nitta2017}; "T" indicates characters scored according to taxonomy (genus or family); 6 = \cite{Nayar1971}; 7 = \cite{Lloyd1980}; 8 = \cite{Zhang2008}; 9 = \cite{Bierhorst1967}; 10 = \cite{Martin2006}; 11 = \cite{Tigerschiold1989}; 12 = \cite{Tigerschiold1990}; 13 = \cite{Atkinson1975}.
+\label{supp_table:traitvals}
+\end{table}
 %%%%%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-03-23 bin/mean_trait_values.R

@@ -1,5 +1,7 @@
-library(tidyverse)
+#library(tidyverse)
 library(mooreaferns)
+library(dplyr)
+library(tidyr)
 
 setwd(here::here())
 
@@ -10,9 +12,9 @@ setwd(here::here())
 # load raw measurements, including multiple measurements per individual
 sla.raw <- mooreaferns::sla.raw
 
-# Nitta 2543 Ptisana salicina is in ERROR, should be Nitta 2544.
+# Nitta 2543 Ptisana salicina is in error, it should be Nitta 2544.
 # Already have measurements for 2544 P. salicina, so it must have been measured twice.
-# Exclude 2544 P. salicina
+# So exclude 2544 P. salicina
 sla.raw <- sla.raw[!(sla.raw$specimen=="Nitta_2544" & sla.raw$species=="Ptisana_salicina"), ]
 
 # calculate mean SLA for each individual
@@ -23,12 +25,12 @@ sla.mean <-
   left_join(unique(select(sla.raw, species, specimen)))
 
 # calculate grand mean for each species based on individual means
-sla.grand.mean <- 
+sla.grand.mean.results <- 
   sla.mean %>%
   group_by(species) %>%
   summarize(mean = mean(sla) , sd = sd(sla), n = n())
   
-sla.grand.mean$trait <- "sla"
+sla.grand.mean.results$trait <- "sla"
 
 ####################
 ### other traits ###
@@ -51,13 +53,13 @@ morph.mean <-
 rm(morph.long)
 
 # merge with sla grand means
-morph.mean <- bind_rows(morph.mean, sla.grand.mean)
+morph.mean <- bind_rows(morph.mean, sla.grand.mean.results)
 
 # split means into numeric and integer (only number of pinna pairs) measurements
 morph.numeric <- filter(morph.mean, trait == "frond.length" | trait == "lamina.width" | trait == "rhizome.dia" | trait == "stipe.length" | trait == "sla") 
 morph.int <- filter(morph.mean, trait == "pinna.pairs")
 
-# set number of signif digits differently for each
+# set number of signif digits differently for numeric vs integer measurements
 # numeric measurements made with normal ruler in cm, so should have to 2 digits after 0
 # need to use formatC to keep trailing zeros
 # see https://kmyu.wordpress.com/2011/01/11/formatting-numbers-for-printing-in-r-rounding-and-trailing-zeroes/
@@ -86,17 +88,51 @@ morph.mean <-
 ###################################
 
 morph.results <- 
[…]

### 2019-06-20 bin/functional_diversity.R
Export RDS to presentation

@@ -515,6 +515,9 @@ multvar.t.results <- make_barplot_means (compare.df, resp_vars, t.test.out)
 # generate boxplots as list
 boxplots <- plot_box_dataset (compare.df, resp_vars, multvar.t.results)
 
+# save RDS for exporting to presentation
+saveRDS(boxplots, "tmp/multvar_boxplots.RDS")
+
 # modifications for boxplots
 # strip y axis titles and scale, since these are same across all plots in a row
  for (i in 1:length(boxplots)) {
@@ -630,6 +633,9 @@ univar.t.results <- make_barplot_means (traits.test, resp_vars, t.test.out)
 
 ### make barplots
 barplots <- plot_bar_dataset(univar.t.results, resp_vars)
+
+# save RDS for exporting to presentation
+saveRDS(barplots, "tmp/univar_barplots.RDS")
 
 # modifications for barplots
 # strip y axis titles (but keep scale, because different from part A)

### 2019-06-20 manuscript.Rnw
Update RNW

@@
 library(dplyr) # use to filter df prior to get_result in some cases
-get_result <- function (results_table, col_name, row_name=1, signif=FALSE, digits=3) {
+get_result <- function (results_table, col_name, row_name=1, signif=FALSE, digits=2) {
   if (signif == FALSE) {
-    value <- round(results_table[row_name, col_name], digits=3)
+    value <- round(results_table[row_name, col_name], digits=digits)
   } else if (signif == TRUE) {
@@
 Slope and intercept for variables related to temperature were very similar between epiphytic and terrestrial dataloggers, but differed more strongly for those related to humidity (\tab \ref{supp_table:climate_model}, \fig \ref{fig:climate}).
-Minimum RH increased \Sexpr{get_result(slopes.results, "slope.epi", "min_RH", TRUE, 2)*100}\% per \SI{100}{\meter} elevation for epiphytic dataloggers, but was nearly flat (\Sexpr{get_result(slopes.results, "slope.ter", "min_RH", TRUE,2)*100}\% per \SI{100}{\meter}) for terrestrial dataloggers (linear model, $\rval^2 = \Sexpr{get_result(slopes.results, "r.squared", "min_RH")}$, $\pval < 0.001$).
+Minimum RH increased \Sexpr{get_result(slopes.results, "slope.epi", "min_RH", TRUE)*100}\% per \SI{100}{\meter} elevation for epiphytic dataloggers, but was nearly flat (\Sexpr{get_result(slopes.results, "slope.ter", "min_RH", TRUE)*100}\% per \SI{100}{\meter}) for terrestrial dataloggers (linear model, $\rval^2 = \Sexpr{get_result(slopes.results, "r.squared", "min_RH")}$, $\pval = \Sexpr{get_result(slopes.results, "pval", "min_RH", TRUE)}$).
 The Mt.
@@
 Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (\fig \ref{fig:PCA}C, D).
-The first two PC axes explained \Sexpr{get_result(PCA.results, "Std_PCA_Dim.2", "cumulative_variance", digits=3, signif=TRUE)}\% and \Sexpr{get_result(PCA.results, "Phy_PCA_Dim.2", "cumulative_variance", digits=3, signif=TRUE)}\% of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
+The first two PC axes explained \Sexpr{get_result(PCA.results, "Std_PCA_Dim.2", "cumulative_variance")}\% and \Sexpr{get_result(PCA.results, "Phy_PCA_Dim.2", "cumulative_variance")}\% of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
 In both analyses, stipe length and frond width were loaded on PC1, and SLA, degree of frond dissection, and frond length were loaded on PC2 (\fig \ref{fig:PCA}A, B, \tab \ref{supp_table:PCA}).
@@
 However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of $\K < 1$ (\tab \ref{table:phylosig}a).
-All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, $\D = \Sexpr{get_result(binary_phylosig.results, "D", "hairs", digits=2, signif=TRUE)}$) or more conserved (gemmmae $\D =  \Sexpr{get_result(binary_phylosig.results, "D", "gemmae", digits=2, signif=TRUE)}$; glands, $\D =  \Sexpr{get_result(binary_phylosig.results, "D", "glands", digits=2, signif=TRUE)}$; morphotype, $\D = \Sexpr{get_result(binary_phylosig.results, "D", "cordate_morph", digits=2, signif=TRUE)}$) than expected under a scenario of evolution by BM (\tab \ref{table:phylosig}b).
+All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, $\D = \Sexpr{get_result(binary_phylosig.results, "D", "hairs")}$) or more conserved (gemmmae $\D =  \Sexpr{get_result(binary_phylosig.results, "D", "gemmae")}$; glands, $\D =  \Sexpr{get_result(binary_phylosig.results, "D", "glands")}$; morphotype, $\D = \Sexpr{get_result(binary_phylosig.results, "D", "cordate_morph")}$) than expected under a scenario of evolution by BM (\tab \ref{table:phylosig}b).
 Phylogenetic generalized linear mixed models (PGLMMs) also indicated significant signal in all gametophyte traits (\tab \ref{supp_table:PGLMM_gameto}).
-Growth habit showed phylogenetic signal nearly identical to that expected under BM ($\D = \Sexpr{get_result(binary_phylosig.results, "D", "epiphytic", digits=2, signif=TRUE)}$; \tab \ref{table:phylosig}b).
+Growth habit showed phylogenetic signal nearly identical to that expected under BM ($\D = \Sexpr{get_result(binary_phylosig.results, "D", "epiphytic")}$; \tab \ref{table:phylosig}b).
 
@@
 Epiphytic species tend to have smaller values for all traits (\tab \ref{table:PIC_sporo}; contrasts were made by subtracting epiphytic from terrestrial values).
-Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, $\pval = \Sexpr{get_result(PIC.results, "pval", "stipe", digits=2, signif=TRUE)}$), as well as frond width ($\pval = \Sexpr{get_result(PIC.results, "pval", "width", digits=2, signif=TRUE)}$) and degree of laminae dissection ($\pval = \Sexpr{get_result(PIC.results, "pval", "dissection", digits=2, signif=TRUE)}$).
+Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, $\pval = \Sexpr{get_result(PIC.results, "pval", "stipe", TRUE)}$), as well as frond width ($\pval = \Sexpr{get_result(PIC.results, "pval", "width", TRUE)}$) and degree of laminae dissection ($\pval = \Sexpr{get_result(PIC.results, "pval", "dissection", TRUE)}$).
 For the gametophyte traits, only morphotype (non-cordate vs.
-cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, $\pval = \Sexpr{get_result(pagel.results, "pval", "morph_binary", digits=2, signif=TRUE)}$; \tab \ref{table:pagel_gameto}).
+cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, $\pval = \Sexpr{get_result(pagel.results, "pval", "morph_binary", TRUE)}$; \tab \ref{table:pagel_gameto}).
 Quantitatively similar results were obtained using PGLMMs, except that no significant relationship was observed between rhizome diameter and growth habit, whereas SLA did show a significant relationship with growth habit (\tabs \ref{supp_table:PGLMM_sporo}, \ref{supp_table:PGLMM_gameto}).
@@
 We observed a total of \Sexpr{ncol(comm)} species across all plots, including \Sexpr{length(traits$habit[traits$habit == "epiphytic"])} epiphytic and \Sexpr{length(traits$habit[traits$habit == "terrestrial"])} terrestrial species.
-Mean species richness was $\Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="epiphytic"), "mean", signif=TRUE)} \pm \stdev \Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="epiphytic"), "sd", signif=TRUE)}$ \si{\species} per plot for epiphytes and $\Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="terrestrial"), "mean", signif=TRUE)} \pm \stdev \Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="terrestrial"), "sd", signif=TRUE)}$ \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, $\pval = \Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="terrestrial"), "pval", signif=TRUE)}$; \fig \ref{fig:FDIV}A).
+Mean species richness was $\Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="epiphytic"), "mean", signif=TRUE)} \pm \Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="epiphytic"), "sd", signif=TRUE)}$ (\stdev) \si{\species}per plot for epiphytes and $\Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="terrestrial"), "mean", signif=TRUE)} \pm \Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="terrestrial"), "sd", signif=TRUE)}$ (\stdev) \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, $\pval = \Sexpr{get_result(filter(multvar.t.results, resp_var=="ntaxa", habit=="terrestrial"), "pval", signif=TRUE)}$; \fig \ref{fig:FDIV}A).
 Species richness increased linearly with elevation and minimum RH for epiphytes, but not terrestrial communities (\fig \ref{fig:FDIV}B).
@@
 FEve decreased with elevation, but not minimum RH, in terrestrial communities (\fig \ref{fig:FDIV}A).
-Epiphytic communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, $\rval^2 = \Sexpr{get_result(filter(multvar.el.results, habit=="epiphytic", variable=="FRic"), "r.squared.plain", signif=TRUE)}$, $\pval = \Sexpr{get_result(filter(multvar.el.results, habit=="epiphytic", variable=="FRic"), "p", signif=TRUE, digits=1)}$) and plots with greater minimum RH (linear model, $\rval^2 = \Sexpr{get_result(filter(multvar.env.results, habit=="epiphytic", variable=="FRic"), "r.squared.plain")}$, $\pval = \Sexpr{get_result(filter(multvar.env.results, habit=="epiphytic", variable=="FRic"), "p")}$).
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (two-sided \tval-test, $\pval = \Sexpr{get_result(filter(multvar.t.results, habit=="terrestrial", resp_var=="mntd.obs.z"), "pval", signif=TRUE)}$) but not MPD ($\pval = \Sexpr{get_result(filter(multvar.t.results, habit=="terrestrial", resp_var=="mpd.obs.z"), "pval", signif=TRUE, digits=2)}$; \fig \ref{fig:FDIV}B).
+Epiphytic communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, $\rval^2 = \Sexpr{get_result(filter(multvar.el.results, habit=="epiphytic", variable=="FRic"), "r.squared.plain")}$, $\pval = \Sexpr{get_result(filter(multvar.el.results, habit=="epiphytic", variable=="FRic"), "p", signif=TRUE)}$) and plots with greater minimum RH (linear model, $\rval^2 = \Sexpr{get_result(filter(multvar.env.results, habit=="epiphytic", variable=="FRic"), "r.squared.plain")}$, $\pval = \Sexpr{get_result(filter(multvar.env.results, habit=="epiphytic", variable=="FRic"), "p")}$).
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (two-sided \tval-test, $\pval = \Sexpr{get_result(filter(multvar.t.results, habit=="terrestrial", resp_var=="mntd.obs.z"), "pval", signif=TRUE)}$) but not MPD ($\pval = \Sexpr{get_result(filter(multvar.t.results, habit=="terrestrial", resp_var=="mpd.obs.z"), "pval", signif=TRUE)}$; \fig \ref{fig:FDIV}B).
 No significant trend was detected in phylogenetic community structure with environment in either terrestrial or epiphytic communities (\fig \ref{fig:FDIV}A).
@@
 \begin{table}[ht]
-\caption{Trait values and growth habit of ferns from Moorea, French Polynesia.
-For continuous trait values, means $\pm$ \stdev shown.
-For growth habit, E = epiphytic, T = terrestrial.
-For frond dissection, 1 = simple, 2 = pinnatifid or pinnatisect, 3 = 1-pinnate, 4 = 1- pinnate-pinnatifid, 5 = 2-pinnate, 6 = 2-pinnate-pinnatifid, 7 = 3- pinnate, 8 = 3-pinnate-pinnatifid, 9 = \textgreater 3-pinnate-pinnatifid, 10 = binpinnatifid or tripinnatifid.}
+\caption{Trait values and growth habit of ferns from Moorea, French Polynesia.}
 \label{supp_table:traitvals}

### 2019-06-20 ms/manuscript.Rmd
Start on new manuscript using Rmd instead of Rnw

---
title: "Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns"
bibliography: "references_clean.bib"
output: word_document
csl: "new-phytologist.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Joel H. Nitta^1,4^, James E. Watkins, Jr.^2^, Ravahere Taputuarai^3^, and Charles C. Davis^1^

^1^Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, Cambridge, MA 02138, U.S.A.

^2^Department of Biology, Colgate University, 13 Oak Drive, Hamilton, New York 13346, U.S.A.

^3^Association Te Rau Ati Ati a Tau a Hiti Noa Tu, B.P. 11553, Mahina, Tahiti, French Polynesia

^4^Current affiliation: Department of Botany, National Museum of Natural History, Smithsonian Institute, Washington, D.C., 20013, U.S.A.

## Summary

- The appearance of angiosperm-dominated forests in the Cretaceous and early Cenozoic had a profound effect on terrestrial biota by creating novel ecological niches. The diversity of modern fern lineages is thought to be a response to angiosperm dominance, radiating both in the canopy and on the forest floor. Despite the significance of epiphytic growth for fern evolution, however, few studies have investigated traits related to epiphytism in ferns while accounting for phylogeny.

- We investigated evolutionary patterns in traits related to epiphytic growth at the species and community level using a phylogeny including representatives of multiple epiphytic fern radiations on the island of Moorea, French Polynesia.

- We observed phylogenetic conservatism in multiple traits related to epiphytic growth in fern sporophytes and gametophytes. Epiphytes had significantly shorter stipes, smaller rhizomes, and a higher frequency of non-cordate gametophytes while accounting for phylogeny. Epiphytic communities were clustered both phylogenetically and functionally, consistent with the hypothesis that the canopy acts as a filter due to its low humidity.

**Key words:** community ecology, functional traits, gametophyte, Moorea, Society Islands, Tahiti.Our results provide an important phylogenetic perspective on the role of functional traits in ferns, and pave the way for more detailed studies examining how niche space is filled within terrestrial and epiphytic habitats.

## Introduction

The appearance of angiosperm-dominated forests in the late Cretaceous and early Cenozoic had an immense effect on terrestrial ecosystems [@Niklas1983; @Kenrick1997a; @Lloyd2008]. The heterogeneous niche space of the angiosperm canopy comprising stratified light and humidity levels and a range of substrates for growth became rapidly filled by a variety of epiphytic plants, including mosses [@Fiz-Palacios2011], leafy liverworts [@Feldberg2014], orchids [@Givnish2015b], bromeliads [@Givnish2014a], and ferns [@Schneider2004d]. Diversification rates in some of these groups may have further increased during the Paleocene / Eocene thermal maximum, a period of warmer and wetter climates that may have enabled movement into the canopy [@Schuettpelz2009; @Testo2016a]. The transition to epiphytic growth has been particularly important for fern diversity, with ca. 29% of all species exhibiting this growth form relative to ca. 10% across all vascular plants [@Kress1986]. Major transitions to epiphytism occurred at least five times during the evolution of ferns, each followed by diversification during the late Cretaceous or early Cenozoic [@Schuettpelz2009; @WatkinsJr.2012]. Recent studies [@Sundue2015; Testo2016a; @Lehtonen2017] have failed to find evidence of rapid diversification characteristic of adaptive radiations associated with epiphytic growth in ferns, but have confirmed gradually increasing rates of diversification in Polypodiaceae [@Sundue2015] and generally support the significance of epiphytic growth as a key innovation for extant fern diversity (@Testo2016a; @WatkinsJr.2012; but see @Lehtonen2017).

Despite the ecological opportunities afforded by life in the canopy, including escape from disturbance [@Page2002] and abundant light for photosynthesis [@Benzing2000], epiphytes must contend with greater stress in the form of increased evaporation and extreme oscillations in temperature relative to terrestrial plants [@Zotz2001; Lowman2012]. Associated with this change, epiphytic ferns have evolved various morphological adaptations that minimize water loss from their leaves, including shorter stipes [@WatkinsJr.2010], thicker leaves [@Kluge2007; WatkinsJr.2007b], and protection with scales or hairs [@WatkinsJr.2006b]. Some lineages have specialized adaptations to epiphytic growth, such as humus-collection (e.g., *Asplenium*; @Benzing1990, *Drynaria*; @Janssen2005), vegetative desiccation tolerance (DT) (e.g., Hymenophyllaceae; @Shreve1911, *Pleopeltis*; @Pessin1925), CAM photosynthesis (e.g., *Pyrrosia*; @Wong1976), and symbiotic relationships with ants (e.g., *Antrophyum*; @WatkinsJr.2008, *Lecanopteris*; @Gay1993a). Many of these same adaptations are found in epiphytic seed plants [@Benzing1990]. However, the evolution of epiphytism in ferns was more complex given that ferns rely on a free-living gametophyte. The fern gametophyte is anatomically simple, and differs from the sporophyte in anatomy, morphology, and physiology. Such differences create a system of 'dual ecology' where selective pressures acting on one stage are different than those acting on another. For example, DT is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is rare in sporophytes [@WatkinsJr.2007]. Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no studies to our knowledge have analyzed the evolution of such traits across the fern phylogeny including both life stages and representatives of all five Cretaceous epiphytic radiations *sensu* [@Schuettpelz2009].

Most studies investigating adaptations to epiphytism have utilized a univariate, cross-species approach, analyzing traits individually between terrestrial and epiphytic species [e.g., @WatkinsJr.2007b; @WatkinsJr.2010]. However, given multiple radiations into the epiphytic niche [@Schuettpelz2009], it is likely that traits related to epiphytism carry at least some degree of phylogenetic signal (i.e., the tendency for closely related species to resemble each other). Therefore, a comparative approach incorporating phylogenetic relationships (and if possible, branch lengths) is appropriate [@Felsenstein1985]. However, such approaches have only recently been applied to studies of epiphytism in ferns, and at relatively limited phylogenetic scales [e.g., @Zhang2014b; @Sundue2015].

It should be possible to gain greater insight into the selective strength exerted by the epiphytic habitat by utilizing a community-based approach. If the harsh abiotic conditions of the canopy act as a filter, we  expect to find lower functional richness at a community level, which should result in decreased phylogenetic diversity under a scenario of phylogenetic trait conservatism [@Webb2002]. However, the relative strength of this environmental filter may depend on overall abiotic conditions; for example, relative humidity (RH) and temperature also vary with elevation [@Cardelus2006]. Elevation is a strong predictor of fern species richness, with maximal richness occurring at mid-elevations, especially in the tropics [@Kessler2011a]. Functional diversity is also correlated with species richness (few species tend to have exactly the same functional traits; @Schleuter2010). Thus, environmental variation along elevational gradients is expected to impact the functional, phylogenetic, and species diversity of epiphytic and terrestrial fern communities.

Here, we investigate the evolution and role in community assembly of functional traits related to epiphytic growth in fern gametophytes and sporophytes on Moorea, French Polynesia. Moorea is an isolated tropical oceanic island with a high phylogenetic diversity of ferns distributed over a steep elevational gradient, making it ideal for a community phylogenetic analysis of functional traits. Specifically, we test the following hypotheses related to epiphytic growth in ferns:

1. The epiphytic environment experiences greater extremes in temperature and humidity than the terrestrial environment, thus acting as a filter on the traits of plants that grow there.

2. Epiphytic ferns have converged upon similar morphologies multiple times in both sporophyte and gametophyte phases.

3. Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities due to environmental filtering, and the strength of this filtering varies with changes in environment.

### 2019-06-24 ms/manuscript.Rmd
Add figures

@@
 title: "Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns"
-bibliography: "references_clean.bib"
-output: word_document
-csl: "new-phytologist.csl"
+bibliography: "references.bib"
+mainfont: Roboto-Regular.ttf
+fontsize: 12pt
+mainfontoptions: 
+- BoldFont=Roboto-Bold.ttf
+- ItalicFont=Roboto-Italic.ttf
+- BoldItalicFont=Roboto-BoldItalic.ttf
+output: 
+  bookdown::pdf_document2:
+    citation_package: none
+    toc: no
+    number_sections: no
+    fig_caption: yes
+    keep_tex: no
+    latex_engine: xelatex
+    pandoc_args: [ "--csl", "new-phytologist.csl"]
+  bookdown::word_document2:
+    toc: no
+    fig_caption: yes
+    pandoc_args: [ "--csl", "new-phytologist.csl"]
+editor_options: 
+  chunk_output_type: console
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
+library(tidyverse)
+library(drake)
+library(patchwork)
 ```
@@
 Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities due to environmental filtering, and the strength of this filtering varies with changes in environment.
+
+## Materials and Methods
+
+## Results
+
+## Discussion
+
+## Acknowledgements
+
+## Figures
+
+```{r climate-plot, fig.height = 4, fig.width = 6, fig.caption = "climate"}
+loadd(climate_plot)
+climate_plot
+```
+
+```{r pca-plot, fig.height = 6, fig.width = 6, fig.caption = "trait PCA"}
+loadd(pca_plot)
+pca_plot
+```
+
+```{r tree, fig.height = 8, fig.width = 6, fig.caption = "traits on tree"}
+loadd(traits_with_tree)
+traits_with_tree
+```
+
+```{r combined-comm-div-plots, fig.height = 8, fig.width = 6, fig.caption = "community diversity metrics"}
+loadd(combined_comm_div_plots)
+combined_comm_div_plots
+```
+
+```{r combined-cwm-div-plots, fig.height = 8, fig.width = 6, fig.caption = "community-weighted means"}
+loadd(combined_cwm_plots)
+combined_cwm_plots
+```
+
+## Tables
+
+## References

### 2019-07-11 ms/manuscript.Rmd
Update manuscript Rmd
- Add font files
- Add latex style file for New Phytologist
- Add clean bib file

--- 2019-06-20 ms/manuscript.Rmd
@@
 title: "Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns"
-bibliography: "references_clean.bib"
-output: word_document
-csl: "new-phytologist.csl"
+bibliography: "references.bib"
+mainfont: Times New Roman.ttf
+fontsize: 12pt
+mainfontoptions: 
+- BoldFont=Times New Roman Bold.ttf
+- ItalicFont=Times New Roman Italic.ttf
+- BoldItalicFont=Times New Roman Bold Italic.ttf
+output: 
+  bookdown::pdf_document2:
+    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
+    toc: no
+    number_sections: no
+    fig_caption: yes
+    keep_tex: no
+    latex_engine: xelatex
+    pandoc_args: [ "--csl", "new-phytologist.csl"]
+    includes:
+      in_header: newphyt.sty
+editor_options: 
+  chunk_output_type: console
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
+library(tidyverse)
+library(drake)
+library(patchwork)
+library(kableExtra)
+library(here)
+library(magrittr)
+library(scales)
 ```
+
+<!-- FOR CONVERSION TO MS WORD ONLY: remove page numbers -->
+\pagenumbering{gobble}
 
@@
 ^4^Current affiliation: Department of Botany, National Museum of Natural History, Smithsonian Institute, Washington, D.C., 20013, U.S.A.
+
+\clearpage
+
+<!-- Set left justification (better conversion to Word than default full justification) -->
+\raggedright
+
+<!-- Set indent to about 3 characters throughout -->
+\setlength{\parindent}{3ex}
 
@@
 
-**Key words:** community ecology, functional traits, gametophyte, Moorea, Society Islands, Tahiti.Our results provide an important phylogenetic perspective on the role of functional traits in ferns, and pave the way for more detailed studies examining how niche space is filled within terrestrial and epiphytic habitats.
+- Our results provide an important phylogenetic perspective on the role of functional traits in ferns, and pave the way for more detailed studies examining how niche space is filled within terrestrial and epiphytic habitats.
+
+**Key words:** community ecology, functional traits, gametophyte, Moorea, Society Islands, Tahiti.
+
+\clearpage
 
@@
 Major transitions to epiphytism occurred at least five times during the evolution of ferns, each followed by diversification during the late Cretaceous or early Cenozoic [@Schuettpelz2009; @WatkinsJr.2012].
-Recent studies [@Sundue2015; Testo2016a; @Lehtonen2017] have failed to find evidence of rapid diversification characteristic of adaptive radiations associated with epiphytic growth in ferns, but have confirmed gradually increasing rates of diversification in Polypodiaceae [@Sundue2015] and generally support the significance of epiphytic growth as a key innovation for extant fern diversity (@Testo2016a; @WatkinsJr.2012; but see @Lehtonen2017).
+Recent studies [@Sundue2015; @Testo2016a; @Lehtonen2017] have failed to find evidence of rapid diversification characteristic of adaptive radiations associated with epiphytic growth in ferns, but have confirmed gradually increasing rates of diversification in Polypodiaceae [@Sundue2015] and generally support the significance of epiphytic growth as a key innovation for extant fern diversity (@Testo2016a; @WatkinsJr.2012; but see @Lehtonen2017).
 
-Despite the ecological opportunities afforded by life in the canopy, including escape from disturbance [@Page2002] and abundant light for photosynthesis [@Benzing2000], epiphytes must contend with greater stress in the form of increased evaporation and extreme oscillations in temperature relative to terrestrial plants [@Zotz2001; Lowman2012].
-Associated with this change, epiphytic ferns have evolved various morphological adaptations that minimize water loss from their leaves, including shorter stipes [@WatkinsJr.2010], thicker leaves [@Kluge2007; WatkinsJr.2007b], and protection with scales or hairs [@WatkinsJr.2006b].
+Despite the ecological opportunities afforded by life in the canopy, including escape from disturbance [@Page2002] and abundant light for photosynthesis [@Benzing2000], epiphytes must contend with greater stress in the form of increased evaporation and extreme oscillations in temperature relative to terrestrial plants [@Zotz2001; @Lowman2012].
+Associated with this change, epiphytic ferns have evolved various morphological adaptations that minimize water loss from their leaves, including shorter stipes [@WatkinsJr.2010], thicker leaves [@Kluge2007; @WatkinsJr.2007b], and protection with scales or hairs [@WatkinsJr.2006b].
 Some lineages have specialized adaptations to epiphytic growth, such as humus-collection (e.g., *Asplenium*; @Benzing1990, *Drynaria*; @Janssen2005), vegetative desiccation tolerance (DT) (e.g., Hymenophyllaceae; @Shreve1911, *Pleopeltis*; @Pessin1925), CAM photosynthesis (e.g., *Pyrrosia*; @Wong1976), and symbiotic relationships with ants (e.g., *Antrophyum*; @WatkinsJr.2008, *Lecanopteris*; @Gay1993a).
@@
 Such differences create a system of 'dual ecology' where selective pressures acting on one stage are different than those acting on another.
-For example, DT is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is rare in sporophytes [@WatkinsJr.2007].
+For example, DT is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is rare in sporophytes [@WatkinsJr.2007c].
 Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no studies to our knowledge have analyzed the evolution of such traits across the fern phylogeny including both life stages and representatives of all five Cretaceous epiphytic radiations *sensu* [@Schuettpelz2009].
@@
 Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities due to environmental filtering, and the strength of this filtering varies with changes in environment.
+
+## Materials and Methods
+
+### Study Site
+
+Moorea, French Polynesia (\ang{17;30;}S, \ang{149;50;}W) is a small (\SI{135}{\kilo\meter\squared}) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
+This distance acts as a strong barrier to dispersal [@Carson1995;@Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders \sensu @PteridophytePhylogenyGroupI2016).
+Ferns of Moorea include representatives of all five of the Cretaceous epiphytic radiations described by @Schuettpelz2009 (\ie, polygrammoids, elaphoglossoids, asplenioids, vittarioids, and filmy ferns) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
+
+### Community Survey
+
+A fern community survey was carried out on Moorea comprising 17, \SI{10}{\meter} $\times$ \SI{10}{\meter} plots spanning an elevational gradient from \SIrange{20}{2000}{\meter}, with plots placed \circa every \SI{200}{\meter} @[Nitta2017].
+Most plots were located along trails leading to the three main peaks on the island, Mt.
+Rotui (\SI{899}{\meter}), Mt.
+Mouaputa (\SI{890}{\meter}), and Mt.
+Tohiea (\SI{1206}{\meter}).
+Each plot was divided into 25, \SI{2}{\meter} $\times$ \SI{2}{\meter} subplots.
+Presence/absence of all sporophytes in each subplot was scored and summed to produce an abundance ranking for each species per plot from zero (not observed) to 25 (observed in all subplots).
+Epiphytic ferns were sampled by hand to \circa \SI{2}{\meter} on tree trunks or confirmed visually from the ground.
+Species were categorized as epiphytic or terrestrial based on field observations (see Trait Selection and Measurements below).
+The list of species in each plot was split into epiphytic and terrestrial groups and treated separately during data analysis, hereafter referred to as 'epiphytic' and 'terrestrial' communities.
+Vouchers were deposited at GH and UC.
+
+### Environmental Survey
+
+Temperature and RH of epiphytic and terrestrial habitats were measured using Hobo Pro v2 dataloggers with the RS3 Solar Radiation Shield (Onset Corporation, USA).
+A pair of dataloggers was installed for each plot, one mounted at \circa \SI{2}{\meter} on a tree ('epiphytic datalogger'), and one at ground level ('terrestrial datalogger'), except for the Mt.
+Tohiea \SI{1200}{\meter} plot, which only had an epiphytic datalogger.
+Temperature and RH were logged once every \SI{15}{\minute} during two survey periods: a preliminary survey from July 18, 2012 to July 7, 2013 at six plots, and a final survey from July 7, 2013 to July 5, 2014 at 15 plots.
+Two plots were not included in the environmental survey.
+We used the 2013–-2014 data for the analysis, with one exception: the terrestrial datalogger at the Mt.
+Tohiea \SI{393}{\meter} plot malfunctioned and was missing data for a substantial part of the final survey period (March 12--July 5, 2014), so we used data for the same dates from the preliminary survey period for this datalogger instead.
+Three other dataloggers also failed early during the final survey period, but lacked data from the preliminary period, so were excluded from analysis.
+Other days missing data for any datalogger during the final survey period were excluded from the dataset.
+The final dataset included 201 days of data for 13 epiphytic dataloggers and 13 terrestrial dataloggers ($\n = 26$).
+
+### Trait Selection and Measurements
+
+We selected several traits relevant to epiphytic growth (Table \@ref(tab:traits-used) and Figure \@ref(fig:traits-used-image)).
+We preferred morphological traits over physiological traits for our study because data for morphological traits are easier to gather at the scale of our study (\circa \SI{130}{\species} distributed over an elevational gradient from \SIrange{200}{2000}{\meter}).
+In addition, plant morphology and physiology are closely linked, and physiological traits such as rates of photosynthesis and water use efficiency are known to be correlated with morphological traits such as stipe length in ferns [@WatkinsJr.2010].
+Studying the relationships between ecology and morphological traits is useful from a paleobotanical and climate change perspective, as trends in morphological trait variation in the fossil record can be used to gain insight into past climatic conditions [@Chaloner1997; @McElwain2007a; @McElwain2007b; @Royer2008].
+
+Morphological traits of sporophytes were measured from herbarium specimens for most species, or obtained from the literature when plant material was unavailable.
+Measurements were made directly on plants in the field for a few species with large fronds.
+A single leaf or rhizome per individual was measured on one to three separately collected individuals per species, and mean values were calculated for each species.
+Although trait values can vary within fern species with elevation [@Kessler2007a], our study includes a broad phylogenetic sampling, so we assume intraspecific variation is outweighed by interspecies differences.
+To measure specific leaf area (SLA; \si{\square\meter\per\kilo\gram}), \circa 10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} \diameter punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
+For species with \textless \SI{2}{\milli\meter} between primary veins, leaf fragments excluding primary veins were obtained by dissection, dried at \SI{60}{\celsius} for \SI{24}{\hour}, weighed, and scanned.
+ImageJ [@Abramoff2004] software was used to measure area per fragment, and SLA calculated in the same way as leaf punches.
+
+Morphological traits of gametophytes were all qualitative (either categorical or binary; ref table 1).
+Trait states were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017], grown from spores of known species in the lab, or taken from the literature.
+For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie, heart-shaped) or non-cordate.
+Non-cordate morphotypes include elongate (strap or ribbon \sensu @Farrar2008) and filamentous forms (ref traits figure).
+Gametophytes were cultivated on Bold's media [@Bold1957] supplemented with Nitch's micronutrients [@Nitsch1951].
+
+Growth habit was coded as a binary trait (epiphytic or terrestrial) based on field observations.
+Some species do not fall clearly into either category (\eg, hemi-epiphytes; [@Benzing1990; @Dubuisson2003; @Nitta2009]).
+We used connection to the soil as the criterion for defining binary growth habit; thus, epipetric plants were treated as epiphytic, and hemi-epiphytes and climbing plants were treated as terrestrial.
+For the purposes of this study, we treated growth habit as a fixed trait and used these categories to define epiphytic and terrestrial communities.
+Other studies have demonstrated that exceptional epiphytic growth of terrestrial fern species (and vice-versa) is rare [@Cardelus2006; @Kluge2006; @WatkinsJr.2009], and we do not believe that such rare exceptions would affect our analysis.
+
+### Statistical analysis
+
+We calculated mean daily maximum, mean, minimum, and standard variation of temperature and RH from the raw climate data, then tested for differences in these summary statistics between epiphytic and terrestrial dataloggers using ANCOVA [@Scheffers2013].
+We used daily mean values of temperature and RH each in turn as the response variable, and constructed linear models with growth habit as the factor and elevation as the covariate, and with their interaction to test for differences in response of climatic variables to elevation between epiphytic and terrestrial dataloggers.
+We also calculated the number of drying events at various minimum RH values, as well as daily mean values for water potential, but found that these were both highly correlated with mean RH and did not analyze them further (data not shown).
+
+For comparative analyses, we used the dated phylogenetic tree of [@Nitta2017], which includes all species from Moorea and some from the neighboring island of Tahiti.
+We trimmed the tree to species occurring only on Moorea ($\n = 130$ \si{\species}) using the \function{drop.tip} function of the \package{ape} package [@Paradis2004] in R [@RCoreTeam2015].
+We tested for phylogenetic signal in quantitative traits using two metrics that relate the observed trait values to those expected under a model of trait evolution following Brownian motion (BM), Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
+$\lambda$ is a scaling parameter that ranges from 0 (traits evolving randomly, as on a star phylogeny) to 1 (traits evolving along branches according to Brownian motion).
+\K describes the ratio between the amount of observed variance in traits vs.
+the amount of variance expected under a BM model of trait evolution: when $\K = 1$, traits are evolving according to BM; when $\K > 1$, traits are more conserved than expected under BM; when $\K < 1$, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
+We tested for significance in \K by comparing observed values of \K against a null distribution of \num{1000} trees with trait values shuffled randomly across the tips, and in $\lambda$ by conducting a likelihood ratio test comparing the log likelihoods of the observed value of $\lambda$ vs.
+$\lambda = 0$ (no phylogenetic signal).
+We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' *D* [@Fritz2010], which is designed to test for signal in binary traits.
+Values of \D range from 0 under a model of trait evolution by BM, and 1 for random distribution of traits; negative values (greater trait conservation than BM) and values greater than 1 (trait overdispersion) are possible.
+We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], which also conducts a significance test by comparing the observed value of \D with distributions of simulated values produced under one of two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM threshold model.
+The probability of obtaining the observed value of \D under either model is then calculated.
+
+We used a multivariate approach to investigate differences in traits between epiphytes and terrestrial species.
+We used only quantitative traits, and included only species that had no missing data for any traits ($\n = 104$ \si{\species}).
+We first log-transformed traits as necessary to attain normality, then scaled each trait by its range.
+We subjected the transformed, scaled traits to a Principal Components Analysis (PCA) to reduce dimensionality.
+We conducted standard and phylogenetically corrected PCA, using the \function{PCA} and \function{phyl.pca} functions in the \package{FactoMineR} [@Le2008] and \package{phytools} R packages, respectively.
+The position of each species was plotted in the multi-dimensional trait space using the first two PC axes, then color-coded according to growth habit (epiphytic vs.
+terrestrial).
+
+We detected significant phylogenetic signal in most traits, so we tested for differences between traits of epiphytic vs.
+terrestrial species using methods that can test for significant differences between traits associated with a binary character while accounting for phylogeny.
+For quantitative traits, we used the \function{brunch} function in the \package{caper} package [@Orme2013], which calculates phylogenetically independent contrasts in a quantitative character between alternative states of a binary trait.
+For qualitative (binary) traits, we used Pagel's test of correlated evolution [@Pagel1994] as implemented with the \function{fitPagel} function in the \package{phytools} package [@Revell2012].
+This method compares the likelihood between two alternative models for a pair of binary traits.
+The first model tests the null hypothesis that rates of evolution of the two traits are independent.
+The second model tests the alternative hypothesis that the rates of evolution of one trait depends on the other.
+The likelihoods of obtaining the observed data are calculated for each model, then compared using a log likelihood test.
+A significantly better fit of the dependent model indicates that evolution of the two traits is correlated.
+To verify the effect of analysis method on our results, we performed alternative tests for correlation of growth habit with functional traits using phylogenetic generalized linear mixed models (PGLMMs; see Methods S1 and Tables S1 and S2).
+
+We analyzed community phylogenetic diversity using the Mean Phylogenetic Distance (MPD) and Mean Nearest Taxon Distance (MNTD) indices of [@Webb2000].
+MPD is the mean phylogenetic distance of each species in a community to all other species in the same community; MNTD is the mean phylogenetic distance of each species in a community to its sister species in the same community.
+Communities with low MPD or MNTD are 'phylogenetically clustered' (co-occurring species are closely related to each other), whereas those with high MPD or MNTD are 'phylogenetically overdispersed' (co-occurring species are distantly related to each other).
+MPD therefore provides a measure of phylogenetic diversity across the whole tree, while MNTD is more sensitive to distances towards the tips.
+Alone, MPD and MNTD are not very informative; to assess their statistical significance, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{1000} randomized communities (equivalent to the Net Relatedness Index and Nearest Taxon Index, respectively, of @Webb2000 with the sign reversed).
+Observed values above 95\% or less than 5\% of the null distribution were considered significant.
+Selection of the randomization procedure can strongly bias outcomes [@Gotelli2006;@Ulrich2012].
+Our null hypothesis is that fern spores are easily capable of dispersal between terrestrial and epiphytic habitats, and across the relatively short distances separating our plots.
+We therefore constructed null communities by drawing species randomly from a regional pool including terrestrial and epiphytic taxa.
+The regional pool included all known species of ferns from Moorea, and all species from the neighboring island of Tahiti with genetic data available ($\n = 146$ \si{\species} total; [@Nitta2017]).
+We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate community phylogenetic diversity metrics.
+
+We analyzed community functional trait diversity using the \function{dbFD} function in the \package{FD} R package [@Laliberte2011].
+This function takes as input values any number of traits for species in a given community, plots them in a multidimensional trait space using PCA, and calculates several summary statistics.
+Functional richness (FRic) is equivalent to convex hull volume containing all observed trait values plotted in multidimensional trait space [@Cornwell2006], functional evenness (FEve) is the average distance to the center of all the traits, and functional dispersion (FDis) is the mean distance between traits [@Villeger2008].
+We chose these three metrics because they are statistically independent and can each provide different information about the structure of community trait diversity [@Savage2012].
+Functional richness does not use abundance data, but functional evenness and divergence can be weighted by abundance, so we preformed analyses for these metrics with abundance weighting to correct for the effect of rare species with unusual morphologies.
+Selection, transformation, and scaling of traits was performed similarly as for the PCA, except that the method used to calculate distance [@Gower1971] can accommodate missing data, so we included some species with observations missing for some traits ($\n = 126$ \si{\species} total).
+We calculated each metric separately for epiphytic and terrestrial communities, and compared results between the two.
+Under a scenario of environmental filtering (as hypothesized for epiphytic communities), we expect relatively lower values for the three metrics of functional diversity [@Savage2012].
+We also investigated relationships between functional trait diversity and environment with linear models by testing for dependence of each functional trait diversity metric against elevation and daily minimum RH.
+We chose these independent variables because we found that overall climate tended to change with elevation, and to test if minimum RH specifically had a filtering effect on diversity along the gradient.
+
+Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife.
+
+## Results
+
+### Environmental Survey
+
+```{r env-survey}
+# loadd(climate_model_summaries)
+# loadd(climate_models_parameter_summaries)
+# 
+# # Extract the slope of the minimum RH for epiphytes
+# # (change in min. % rel humidity per meter).
+# min_RH_epi_slope <- climate_models_parameter_summaries %>%
+#   filter(var == "min_RH", term == "el") %>%
+#   pull(estimate)
+# 
+# # Extract the slope of the minimum RH for terrestrial species.
+# # In the results table, terrestrial slope is the difference
+# # between the estimate for the elevation term and the
+# # interaction term (add though because it's already negative).
+# min_RH_ter_slope <- 
+#   magrittr::add(
+#     min_RH_epi_slope,
+#     climate_models_parameter_summaries %>%
+#       filter(var == "min_RH", term == "el:habitterrestrial") %>%
+#       pull(estimate)
+#   )
+
+# There were significant interaction effects between growth habit and elevation on all microclimatic variables except for minimum temperature (Table S1).
+Slope and intercept for variables related to temperature were very similar between epiphytic and terrestrial dataloggers, but differed more strongly for those related to humidity (Table S1, Figure \@ref(fig:climate).
+Minimum RH increased `r min_RH_epi_slope %>% signif(2) %>% multiply_by(100) %>% percent()` per \SI{100}{\meter} elevation for epiphytic dataloggers, but only `r min_RH_ter_slope %>% signif(2) %>% multiply_by(100) %>% percent()` for terrestrial dataloggers (linear model, $\rval^2 = \Sexpr{get_result(slopes.results, "r.squared", "min_RH")}$, $\pval = \Sexpr{get_result(slopes.results, "pval", "min_RH", TRUE)}$).
+The Mt.
+Rotui \SI{830}{\meter} plot was a clear outlier, with greater mean temperature and lower minimum RH relative to other plots at similar elevation (outliers in  \fig \ref{fig:climate} at \circa \SI{800}{\meter}).
+Mt.
+Rotui is a solitary peak located between Cook's Bay and Oponohu Bay.
+It is more isolated with greater exposure than the other two mountains where the rest of the plots were established, which are part of a mountain complex on the interior of the island (J.
+Nitta, pers.
+obs.).
+
+```
+
+## Discussion
+
+Here, we show for the first time using a broadly sampled phylogenetic dataset that morphological traits related to epiphytic growth are strongly evolutionarily conserved in ferns.
+The canalized morphology of epiphytic ferns suggests that this habitat strongly filters for a specialized growth form.
+We find that communities of epiphytic ferns are more clustered functionally and phylogenetically relative to terrestrial species, again supporting the role of environmental filtering in determining epiphytic fern community structure.
+Our results are similar between gametophyte and sporophyte traits, suggesting that morphology of each life stage is important for epiphytic growth.
+
+### Effect of Phylogeny on Analysis of Epiphytic Traits
+
+Although several other studies have addressed the relationship of functional traits with growth habit in ferns [e.g., @Hietz1998; @Kluge2007; @WatkinsJr.2010], few have accounted for evolutionary history.
+We find that growth habit shows phylogenetic signal, as do several of the traits associated with it.
+In one of the few previous studies to analyze phylogenetic signal in functional traits of ferns, @Zhang2014b found that only one of 16 traits showed significant phylogenetic signal.
+The only trait in common between the current study and that of @Zhang2014b is SLA, which did not show phylogenetic signal in either study.
+Given the high degree of phylogenetic signal we observe in most traits and that at least some of the traits between our two studies are likely to be correlated, it is surprising that @Zhang2014b did not detect phylogenetic signal in more traits.
+However, @Zhang2014b only used \K, which tended to indicate less phylogenetic signal than $\lambda$ in our study.
+In simulation studies, $\lambda$ has lower error (both Type I and Type II) than \K, which can be inconsistent unless the strength of BM is great [@Munkemuller2012].
+Our results indicate the importance of including a phylogenetic component, and including multiple measures of phylogenetic signal, when analyzing traits related to epiphytic growth in ferns.
+
+### Sporophyte Traits
+
+The sporophyte traits that differed most strongly between epiphytic and terrestrial taxa were related to plant size, including stipe and frond length, frond width, and rhizome diameter.
+Stipe length in particular differed strongly between epiphytes and terrestrial species (Figure \@ref(fig:cwm-div), TABLE PIC SPOROS).
+Stipe length may be a critical trait controlling whole leaf hydraulic conductance; since ferns have resistive stipes in general, species growing in water-limited habitats are expected to decrease stipe length to minimize resistance [@WatkinsJr.2010].
+Our results support previous non-phylogenetically controlled studies that found a prevalence of short stipes in epiphytic ferns [@WatkinsJr.2010; Creese2011].
+However, given that many of these size-related traits scale together in ferns [@Arcand2008; Creese2011], careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
+
+```{r sla-numbers}
+loadd(fern_traits)
+
+# First smallest SLA: Humata_anderssonii
+smallest_sla_1 <- fern_traits %>% arrange(sla) %>% slice(1) %>% pull(sla) %>% signif(3)
+assert_that(fern_traits %>% arrange(sla) %>% slice(1) %>% pull(species) == "Humata_anderssonii")
+
+# Second smallest SLA: Ctenopterella_blechnoides
+smallest_sla_2 <- fern_traits %>% arrange(sla) %>% slice(2) %>% pull(sla) %>% signif(3)
+assert_that(fern_traits %>% arrange(sla) %>% slice(2) %>% pull(species) == "Ctenopterella_blechnoides")
+
+```
+
+Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (Figure \@ref(fig:cwm-div), TABLE PIC SPOROS).
+SLA is an important part of the leaf economic spectrum, which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
+slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a].
+We hypothesized that epiphytes would have lower SLA, given their more extreme growing conditions.
+Indeed, the species with the lowest SLA values were epiphytes (\eg, *Davallia anderssonii* Mett.
+ex Kuhn, \SI{`r smallest_sla_1`} {\square\meter\per\kilo\gram}; *Ctenopterella blechnoides* (Grev.) Parris, \SI{`r smallest_sla_2`} {\square\meter\per\kilo\gram}), but this relationship was not significant across the dataset (Figure \@ref(fig:cwm-div), TABLE PIC SPOROS).
+The tradeoffs posited in the leaf economics spectrum have been most thoroughly documented in global studies including the widest possible range of taxa and environments, and do not always hold up in local studies [@Petter2016].
+For example, @Zhu2016 did not find a significant relationship between SLA and leaf life span in a survey of 16 ferns across a disturbance (light) gradient in China.
+It is possible that other small-scale effects (\eg, microclimate) obscure this pattern in local studies such as ours and that of @Zhu2016.
+This is supported by the strong, parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (Figure \@ref(fig:cwm-div)).
+
+### Gametophyte Traits
+
+Only gametophyte morphotype was found to have a significant association with epiphytic growth.
+Morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived ($< \SI{1}{\year}$), cordate gametophytes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate (\eg, elongate or filamentous) gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; Farrar2008].
+The complex, three-dimensional structure of non-cordate gametophytes slows their drying rates, and may be an adaptation to drought-prone epiphytic habitats [@WatkinsJr.2007c; @Pittermann2013], and promote out-crossing [@Farrar2008].
+Our study is the first to our knowledge to demonstrate that the correlation between non-cordate morphology and epiphytism in ferns is significant while controlling for phylogeny.
+@Sundue2015 found that non-cordate morphology is significantly correlated with increased diversification rates within Polypodiaceae, consistent with the hypothesis that this trait promoted diversification within the epiphytic niche.
+
+We did not find that gemmae production was linked with epiphytism.
+Gemmae are asexual propagules that allow gametophytes to persist over multiple growing seasons and attain large population sizes [@Farrar2008].
+Gemmae production should be a useful trait in an epiphytic context because it may allow gametophytes to increase their surface area and hedge against population loss due to extreme drying events [@Farrar2008].
+Gemmae can also function sexually by producing antherida in the presence of antheridiogen secreted by other gametophytes [@Emigh1977], which may confer a selective advantage in the epiphytic environment by enabling outcrossing [@Dassler2001].
+It is possible that some, but not all, epiphytic ferns rely on such strategies, and that they are not under similarly strong selection as other traits such as stipe length and gametophyte morphotype.
+Interestingly, when we classified 'widespread' species as those with gametophytes occurring more than \SI{200}{\meter} above or below the minimum or maximum elevational range of conspecific sporophytes (ADD TO SI METHODS), we found that gemmae production is correlated with widespread growth (log likelihood dependent model = \num{-54.5}, log likelihood independent model = \num{-63.6}, $\pval = 0.001$, Pagel's test of correlated evolution).
+Thus, it seems that gemmae do function in long-term gametophyte persistence, but may not be required for epiphytic growth.
+
+Presence of hairs and glands has strong phylogenetic signal (REF TABLE phylosig, REF TABLE supp pglmm gameto), and these traits are correlated with certain taxonomic groups [@Stokey1951; @Nayar1971].
+Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], we did not detect a correlation between hair production and epiphytic growth.
+Similarly, production of glands does not appear to be correlated with epiphytic growth.
+Wax-secreting glands are prevalent in both sporophytes and gametophytes of many notholaenid ferns, a xeric-adapted clade, but their functional significance has not been established [@Johnson2012].
+Further tests are needed to determine if such glands serve a functional role in preventing water loss in desert ferns and epiphytes.
+
+### Environmental Filtering
+
+Results for functional trait richness (FRic) supported our hypotheses: we observed lower FRic in epiphytic relative to terrestrial communities overall, and a significant positive correlation of FRic with humidity in epiphytes (\ie, decreased FRic at drier, low elevation plots and increased FRic at more humid, high elevation plots; Figure \@ref(fig:comm-div)).
+However, results for the other two measures of functional diversity, evenness (FEve) and divergence (FDiv) did not match our predictions.
+Rather, FEve decreased with elevation (but not minimum RH) in terrestrial communities, and FDiv was lower in terrestrial communities overall while showing no trend along the gradient  (Figure \@ref(fig:comm-div)).
+The lack of a clear pattern in functional evenness and divergence suggests that within each trait pool, there is no significant difference in trait spacing between epiphytes and terrestrial species.
+This could be maintained by niche partitioning within epiphytic and terrestrial habitats.
+Epiphytic habitats are highly heterogeneous, and microclimatic conditions vary with height along the host and distance from the trunk [@Hietz1995].
+The vertical stratification of epiphytic species into different zones along height gradients in host trees is well documented [@Zotz2007].
+In a study comparing terrestrial and epiphytic ferns along an elevational gradient from \SIrange{30}{3000}{\meter} in Costa Rica, @WatkinsJr.2006a found that epiphytes tended to sort into low-trunk and canopy zones, with only 18\% found in both, and no species that overlapped between canopy and terrestrial habitats.
+It is possible that epiphytic growth per se limits overall morphological variation in epiphytic ferns, but that species with different traits are then sorted into different niches.
+
+In addition to their decreased functional diversity, we also found that epiphytic communities are more phylogenetically clustered relative to terrestrial communities overall (Figure \@ref(fig:comm-div)).
+However, we did not detect significant trends in phylogenetic community structure of epiphytic and terrestrial communities along the elevational gradient.
+It is possible that the scale of the current study (from \SIrange{200}{1200}{\meter}) prevented the detection of such trends.
+When the plots in this study were analyzed together with additional plots from Tahiti reaching a maximum altitude of just over \SI{2000}{\meter}, @Nitta2017 detected increasing phylogenetic clustering in fern sporophyte communities (epiphytes and terrestrial species together) with elevation.
+@Kluge2011 also compared phylogenetic and trait diversity between epiphytic and terrestrial ferns over an elevational gradient from \SIrange{100}{3400}{\meter} in Costa Rica, and found trait overdispersion at mid-elevations for epiphytes but no trends in phylogenetic diversity for either group.
+They accounted this to evolutionary plasticity in traits, but did not quantify the degree of phylogenetic signal in the traits they studied.
+
+### Conclusion and Future Directions
+
+We have demonstrated that growth habit, and several traits associated with it, show significant phylogenetic signal in ferns.
+Furthermore, we find that epiphytes have restricted morphologies both on a per-species basis and at the community level, and that epiphytic communities are phylogenetically clustered relative to terrestrial ones.
+Finally, we observed drier conditions in the canopy relative to terrestrial habitats.
+Taken together, our results support a scenario of environmental filtering acting on conserved traits to limit functional and phylogenetic diversity in epiphytic ferns.
+The generality of this pattern could be further tested by investigating the functional traits of other groups with large epiphytic radiations such as orchids and bromeliads.
+
+Recent studies [e.g., @Sundue2015; @Testo2016a; @Lehtonen2017] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
+Epiphytism itself is likely a complex trait requiring the evolution of multiple associated traits, and the timing and order of the evolution of such traits may be related to the complex diversification dynamics observed in these studies.
+Our study identifies several conserved morphological traits (\eg, stipe length, frond dissection, gametophyte morphotype) that may belong to this suite of epiphytic traits.
+We hope future studies will build on our results by analyzing these traits across much more densely sampled phylogenies [e.g., @Testo2016a] to dissect how each may be related to diversification and growth habit.
+Additionally, future studies should focus on determining in more detail how niche space is partitioned within epiphytic vs.
+terrestrial habitats, which is clearly a major dividing line for ferns.
+
+## Acknowledgements
+
+We thank members of the Davis Lab, N.
+Michele Holbrook, and Jonathan Losos for helpful discussion and comments on drafts.
+Suzanne Vinette, Saad Amer, Tristan Wang, and Tohei Theophilus helped with community surveys and measurement of traits.
+Staff at the University of California, Berkeley Richard B.
+Gump South Pacific Research Station, Moorea, French Polynesia, in particular Valentine Brotherson, Hinano Teavai-Murphy, Frank Murphy, and Neil Davies, provided support with obtaining research and collection permits, laboratory space, and logistics.
+Funding provided in part by National Science Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to JHN and CCD), Setup Funds from Harvard University to CCD, the American Society of Plant Taxonomists (Research Grant for Graduate Students to JHN), the Garden Club of America (Award in Tropical Botany to JHN), the Harvard University Herbaria (Fernald Fieldwork Fellowship to JHN), the Society of Systematic Biologists (Graduate Student Research Award to JHN), and the Systematics Association (Systematics Research Fund to JHN).
+
+## Author Contribution
+
+J.H.N. conceived and carried out the study. R.T. collected data. J.H.N. performed the analyses. J.H.N., J.E.W., and C.C.D. wrote the manuscript.
+
+## References
+
+## Figure Legends
+
+Figure \@ref(fig:traits-used-image): Examples of traits used in this study.
+(**a**) Sporophyte traits.
+All sporophyte traits were quantitative, including frond length and width, stipe length, rhizome dia., number of pinna pairs (in this example, nine) and degree of lamina dissection (in this example, once-pinnate).
+Specific leaf area (ratio of area to mass of leaf lamina) not depicted.
+(**b**) Gametophyte traits.
+Morphotype was defined as a binary trait, either cordate or non-cordate.
+Examples of non-cordate morphotypes include ribbon (thallus elongate and two-dimensional) and filamentous (thallus single lines of cells).
+Other gametophyte traits (gemmae, hairs, and glands) were scored as present or absent.
+Arrows point out instances of each binary trait.
+Scalebars 1 mm except for glands, which is 0.1 mm.
+Photographs by J.
+H.
+Nitta.
+
+Figure \@ref(fig:climate): Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter on Moorea, French Polynesia.
+Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from July 7, 2013 to July 5, 2014, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).
+'Epiphytic' dataloggers were placed at ca.
+2 m on trees, and 'terrestrial' dataloggers placed at ground level.
+Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown.
+Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval \textless \num{.05}
+
+Figure \@ref(fig:pca): Principal components analysis (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
+**(a, b)** PC loadings.
+**(c, d)** species scores.
+**(a)** and **(c)** are standard PCA, **(b)** and **(d)** are phylogenetic PCA.
+Epiphytes in green, terrestrial species in brown.
+Quantitative traits only, including only species with no missing observations for any traits ($\n = 104$ \si{\species}).
+
+Figure \@ref(fig:tree): Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
+Relative value of quantitative (sporophyte) traits in heat gradient: low values are lighter, high values are darker.
+States of qualitative (gametophyte) states and growth habit indicated by colors in key.
+Grey indicates missing data or non-applicable trait states.
+Trends for three leaf size traits (stipe length, frond length, and frond with) were quantitatively very similar, so of these, we only present stipe length.
+Species missing data for six or more traits not shown.
+For a summary of traits, see Table \@ref(tab:traits-used).
+Major Cretaceous epiphytic radiations \sensu @Schuettpelz2009 labeled on tree in green: H, hymenophylloids; T, trichomanoids, V, vittarioids; A, asplenioids; E, elaphoglossoids; P, polygrammoids (trichomanoids diversified prior to the Cretaceous but are included as an important extant epiphytic clade).
+
+Figure \@ref(fig:cwm-div): Change in selected functional traits of ferns on Moorea, French Polynesia across environmental gradients and between growth habit types (epiphytic or terrestrial).
+Epiphytic values in green, terrestrial values in brown.
+Scatterplots show community-weighted mean (CWM) values for each trait plotted against either minimum relative humidity ($\n = 13$ communities each for epiphytes and terrestrial species) or elevation ($\n = 17$ communities each).
+Trendlines indicate significant relationships determined by a linear model at $\pval<0.05$.
+Boxplots show median values (bold lines) by growth habit (epiphytic or terrestrial) across all communities (epiphytic communities, $\n = 17$; terrestrial communities, $\n = 17$).
+Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 $\times$ the interquartile range.
+Asterisks indicate statistical significance of two-sided \tval-test for differences between epiphytic and terrestrial species; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+Trends for three leaf size traits (stipe length, frond length, and frond with) were quantitatively very similar, so of these we only present stipe length.
+
+Figure \@ref(fig:comm-div): Functional and phylogenetic diversity of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
+Epiphytic communities in green, terrestrial communities in brown.
+Response variable abbreviations as follows: MPD, standard effect size (SES) of mean phylogenetic distance; MNTD, SES of mean nearest taxon distance; FRic, functional richness; FEve, functional evenness; FDiv, functional diversity.
+Scatterplots show response variables plotted against either minimum relative humidity ($\n = 13$ communities each for epiphytes and terrestrial species) or elevation ($\n = 17$ communities each).
+Trendlines indicate significant relationships determined by a linear model at $\pval<0.05$.
+Boxplots show median values (bold lines) of response variables by growth habit (epiphytic or terrestrial) across all communities (epiphytic communities, $\n = 17$; terrestrial communities, $\n = 17$).
+Hinges and whiskers as in Figure \@ref(fig:cwm-div).
+Asterisks indicate statistical significance of one-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+
+## Tables
+
+```{r traits-used}
+traits_used <- read_csv(here("data/table_1.csv"))
+
+# super_pattern <- "([[:digit:]].*$)"
+# super_text <- str_extract(traits_used$ecological_significance, super_pattern) 
+# super_text <- paste0("\\\\textsuperscript{", super_text, "}")
+
+traits_used %>%
+  # mutate(ecological_significance = str_replace_all(ecological_significance, super_pattern, super_text)) %>%
+  column_to_rownames("trait") %>%
+  knitr::kable(caption = "Traits used", format = "latex", booktabs = TRUE, escape = TRUE) %>%
+  kable_styling(full_width = FALSE) %>%
+  pack_rows("Sporophyte traits", 1, 6) %>%
+  pack_rows("Gametophyte traits", 7, 11) %>%
+  column_spec(3, width = "15em")
+
+# 
+# traits_used %>%
+#   gt %>%
+#   fmt_markdown(columns = everything()) %>%
+#   gt(rowname_col = "trait") %>%
+#   tab_row_group(
+#     group = "Sporophyte traits",
+#     rows = matches("Stipe|Frond|Rhizome|Pinna|Leaf")
+#   ) %>%
+#   tab_row_group(
+#     group = "Gametophyte traits",
+#     rows = matches("Morphotype|Gemmae|Glands|Hairs")
+#   ) %>%
+#   cols_label(
+#     data_type = "Data type (unit)",
+#     ecological_significance = "Ecological Significance"
+#   ) %>%
+#   knitr::kable(caption = "Traits used", booktabs = TRUE)
+
+```
+
+```{r, results='asis', echo = FALSE, eval=(knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex')}
+# Add pagebreak if rendering to pdf
+cat('\\pagebreak')
+```
+
+## Figures
+
+```{r traits-used-image, fig.width=7, fig.height=7, fig.cap = "Examples of traits used in this study."}
+knitr::include_graphics(here("images/Figure_3_1_traits_lowres.pdf"))
+```
+
+```{r climate, fig.height = 4, fig.width = 6, fig.cap = "Microclimatic variables."}
+loadd(climate_plot)
+climate_plot
+```
+
+```{r pca, fig.height = 6, fig.width = 6, fig.cap = "Trait PCA."}
+loadd(pca_plot)
+pca_plot
+```
+
+```{r tree, fig.height = 8, fig.width = 6, fig.cap = "Traits on tree."}
+loadd(traits_with_tree)
+traits_with_tree
+```
+
+```{r cwm-div, fig.height = 8, fig.width = 6, fig.cap = "Community-weighted means."}
+loadd(combined_cwm_plots)
+combined_cwm_plots
+```
+
+```{r comm-div, fig.height = 8, fig.width = 6, fig.cap = "Community diversity metrics."}
+loadd(combined_comm_div_plots)
+combined_comm_div_plots
+```
+
+## Tables
+
+## References

### 2019-07-11 code/setup.R
Fix typo

@@ -27,4 +27,4 @@ conflict_prefer("gather", "tidyr")
 conflict_prefer("standard_theme", "jntools")
 conflict_prefer("map", "purrr")
 conflict_prefer("select", "dplyr")
-conflict_prefer("ggsave", "ggplot")
+conflict_prefer("ggsave", "ggplot2")

### 2019-07-14 code/functions.R
Update plan and functions
- Add strict dataset excluding gametos with traits scored by taxonomy

- Separate model parameter summaries and overall model summaries

@@ -282,11 +282,21 @@ extract_model_fits <- function (supported_models) {
     unnest
 }
 
-# Extract model summaries
-extract_model_summaries <- function (supported_models) {
+# Extract model parameter summaries (slope, intersect, etc)
+extract_model_parameter_summaries <- function (supported_models) {
   supported_models %>%
     mutate(
       summary = map(model, tidy)
+    ) %>%
+    select(var, model_type, summary) %>%
+    unnest
+}
+
+# Extract model summaries (pvalue, rsquared of model)
+extract_model_summaries <- function (supported_models) {
+  supported_models %>%
+    mutate(
+      summary = map(model, glance)
     ) %>%
     select(var, model_type, summary) %>%
     unnest

### 2019-07-14 code/plan.R

@@ -17,6 +17,10 @@ plan <- drake_plan(
     # make sure binary traits are coded as character
     mutate_at(vars(glands, hairs, gemmae), as.character),
   
+  # - Also make "strict" trait dataset by excluding species with 
+  # gameto traits only known from taxonomy
+  fern_traits_strict = filter(fern_traits, gameto_source != "T"),
+  
   # Site data
   moorea_sites = mooreaferns::sites %>%
     as_tibble %>%
@@ -70,6 +74,11 @@ plan <- drake_plan(
     supported_models = climate_models
   ),
   
+  # Get summaries of supported model parameters.
+  climate_model_parameter_summaries = extract_model_parameter_summaries(
+    supported_models = climate_models
+  ),
+  
   # Get summaries of supported models.
   climate_model_summaries = extract_model_summaries(
     supported_models = climate_models
@@ -97,10 +106,24 @@ plan <- drake_plan(
     phy = phy
   ),
   
+  # - also run for "strict" trait dataset
+  phylosig_binary_traits_strict = analyze_binary_phylosig(
+    traits = fern_traits_strict,
+    phy = phy
+  ),
+  
   # Correlated evolution ----
   correlated_evo_test = list(
     as.list(c("morphotype", "glands", "hairs", "gemmae")),
     traits = list(fern_traits),
+    phy = list(phy)
+  ) %>%
+    pmap_dfr(test_corr_evo),
+  
+  # - also run for "strict" trait dataset
+  correlated_evo_test_strict = list(
+    as.list(c("morphotype", "glands", "hairs", "gemmae")),
+    traits = list(fern_traits_strict),
     phy = list(phy)
   ) %>%
     pmap_dfr(test_corr_evo),
@@ -119,6 +142,7 @@ plan <- drake_plan(
   ),
   
   # Trait community diversity ----
+  # (includes quantitative, i.e., sporophyte, traits only)
   func_div_epi = analyze_fd_by_habit(
     traits = fern_traits,
     comm = comm,
@@ -134,6 +158,7 @@ plan <- drake_plan(
   func_div = bind_rows(func_div_epi, func_div_ter),
   
   # Models ----
+  # (includes quantitative, i.e., sporophyte, traits only)
   
   # Merge all diversity metrics.
   diversity_data = list(comm_struc, func_div, grand_mean_climate, moorea_sites) %>% 
[…]

### 2019-07-14 make.R
Add set.seed

@@ -19,5 +19,8 @@ pkgconfig::set_config("drake::strings_in_dots" = "literals")
 source("code/functions.R")
 source("code/plan.R")
 
+# Set seed for reproducibility.
+set.seed(1954)
+
 # Make plan.
 make(plan)

### 2019-07-14 ms/manuscript.Rmd
Update MS Rmd

@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
-library(tidyverse)
 library(drake)
 library(patchwork)
+library(knitr)
 library(kableExtra)
@@
 library(scales)
+library(tidyverse)
+library(conflicted)
+
+conflict_prefer("expand", "tidyr")
+conflict_prefer("select", "dplyr")
+conflict_prefer("filter", "dplyr")
+
+# Function for rounding with trailing zeros
+round_t <- function (x, digits) round(x, digits) %>% sprintf(glue::glue("%.{digits}f"), .)
+
+# Function for formatting p-values: round to 3 digits,
+# or just say "< 0.001"
+format_pval <- function (x) {
+  case_when(
+    x < 0.001 ~ "< 0.001",
+    TRUE ~ round(x, 3) %>% as.character()
+  )
+}
+
 ```
@@
 
-We selected several traits relevant to epiphytic growth (Table \@ref(tab:traits-used) and Figure \@ref(fig:traits-used-image)).
+We selected several traits relevant to epiphytic growth (\tab \@ref(tab:traits-used) and \fig \@ref(fig:traits-used-image)).
 We preferred morphological traits over physiological traits for our study because data for morphological traits are easier to gather at the scale of our study (\circa \SI{130}{\species} distributed over an elevational gradient from \SIrange{200}{2000}{\meter}).
@@
 
-Morphological traits of gametophytes were all qualitative (either categorical or binary; ref table 1).
+Morphological traits of gametophytes were all qualitative (either categorical or binary; \tab \@ref(tab:traits-used)).
 Trait states were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017], grown from spores of known species in the lab, or taken from the literature.
 For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie, heart-shaped) or non-cordate.
-Non-cordate morphotypes include elongate (strap or ribbon \sensu @Farrar2008) and filamentous forms (ref traits figure).
+Non-cordate morphotypes include elongate (strap or ribbon \sensu @Farrar2008) and filamentous forms (\fig \@ref(fig:traits-used-image)).
 Gametophytes were cultivated on Bold's media [@Bold1957] supplemented with Nitch's micronutrients [@Nitsch1951].
@@
 We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' *D* [@Fritz2010], which is designed to test for signal in binary traits.
-Values of \D range from 0 under a model of trait evolution by BM, and 1 for random distribution of traits; negative values (greater trait conservation than BM) and values greater than 1 (trait overdispersion) are possible.
+Values of \D range from 0 under a model of trait evolution by BM, to 1 for random distribution of traits; negative values (greater trait conservation than BM) and values greater than 1 (trait overdispersion) are possible.
 We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], which also conducts a significance test by comparing the observed value of \D with distributions of simulated values produced under one of two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM threshold model.
@@
 A significantly better fit of the dependent model indicates that evolution of the two traits is correlated.
-To verify the effect of analysis method on our results, we performed alternative tests for correlation of growth habit with functional traits using phylogenetic generalized linear mixed models (PGLMMs; see Methods S1 and Tables S1 and S2).
+To verify the effect of analysis method on our results, we performed alternative tests for correlation of growth habit with functional traits using phylogenetic generalized linear mixed models (PGLMMs; see Methods S1 and \tabs S1 and S2).
 
@@
 
+### Principal Components Analysis of Trait Data
+
+```{r pca-numbers}
+# Extract summary stats for mention in text
+loadd(pca_results)
+
+# Cumulative amount of variance explained by first two axes, standard PCA
+pca_std_cum_var <-
+  pca_results$variance %>% 
+  filter(analysis_type == "standard", variance_type == "Cumulative Proportion") %>%
+  pull(PC2) %>%
+  signif(2)
+
+# Cumulative amount of variance explained by first two axes, phylogenetic PCA
+pca_phy_cum_var <-
+  pca_results$variance %>% 
+  filter(analysis_type == "phylogenetic", variance_type == "Cumulative Proportion") %>%
+  pull(PC2) %>%
+  signif(2)
+```
+
+<!-- CHECK PCA CODE: missing dissection -->
+
+Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (\fig \@ref(fig:pca)(c, d)).
+The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
+The trait loadings were similar between the two analyses, except that most switched axes; for example, SLA was loaded on PC1 in the standard analysis but PC2 in the phylogenetic analysis (REF SUPP TABLE pca).
+
+### Phylogenetic Signal
+
+```{r physig-numbers}
+loadd(phylosig_binary_traits)
+habit_d <- phylosig_binary_traits %>% filter(trait == "habit") %>% pull(D) %>% round_t(2)
+hairs_d <- phylosig_binary_traits %>% filter(trait == "hairs") %>% pull(D) %>% round_t(2)
+gemmae_d <- phylosig_binary_traits %>% filter(trait == "gemmae") %>% pull(D) %>% round_t(2)
+glands_d <- phylosig_binary_traits %>% filter(trait == "glands") %>% pull(D) %>% round_t(2)
+morphotype_d <- phylosig_binary_traits %>% filter(trait == "morphotype") %>% pull(D) %>% round_t(2)
+```
+
+Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs \@ref(tab:phylosig-cont), \@ref(tab:phylosig-binary), \fig \@ref(fig:tree)), and for quantitative traits, different results were obtained for $\lambda$ and K.
+When measured with $\lambda$, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of $\lambda$ close to 1; only number of pinna pairs had $\lambda$ close to zero (\tab \@ref(tab:phylosig-cont)).
+However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of $\K < 1$ (\tab \@ref(tab:phylosig-cont)).
+All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmmae \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under a scenario of evolution by BM (\tab \@ref(tab:phylosig-binary)).
+Phylogenetic generalized linear mixed models (PGLMMs) also indicated significant signal in all gametophyte traits (REF SUPP TABLE).
+Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; \tab \@ref(tab:phylosig-binary)).
+
+### Correlation of Traits with Growth Habit
+
+```{r corr-numbers}
+loadd(pic_results)
+stipe_pval <- pic_results %>% filter(trait == "stipe") %>% pull(p.value) %>% signif(2)
+width_pval <- pic_results %>% filter(trait == "width") %>% pull(p.value) %>% signif(2)
+dissection_pval <- pic_results %>% filter(trait == "dissection") %>% pull(p.value) %>% signif(2)
+
+loadd(correlated_evo_test)
+morphotype_pval <- correlated_evo_test %>% filter(trait == "morphotype") %>% pull(pval) %>% signif(2)
+
+```
+
+Most sporophyte traits, including all those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (\tab \@ref(tab:pics)).
+Epiphytic species tend to have smaller values for all traits (\tab \@ref(tab:pics); contrasts were made by subtracting epiphytic from terrestrial values).
+Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval = `r stipe_pval`), as well as frond width (\pval = `r width_pval`) and degree of laminae dissection (\pval = `r dissection_pval`).
+For the gametophyte traits, only morphotype (non-cordate vs.
+cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval = `r morphotype_pval`; \tab \@ref(tab:corr-evo)).
+Quantitatively similar results were obtained using PGLMMs, except that no significant relationship was observed between rhizome diameter and growth habit, whereas SLA did show a significant relationship with growth habit (REF SUPP TABLE).
+
 ## Discussion
@@
 The sporophyte traits that differed most strongly between epiphytic and terrestrial taxa were related to plant size, including stipe and frond length, frond width, and rhizome diameter.
-Stipe length in particular differed strongly between epiphytes and terrestrial species (Figure \@ref(fig:cwm-div), TABLE PIC SPOROS).
+Stipe length in particular differed strongly between epiphytes and terrestrial species (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
 Stipe length may be a critical trait controlling whole leaf hydraulic conductance; since ferns have resistive stipes in general, species growing in water-limited habitats are expected to decrease stipe length to minimize resistance [@WatkinsJr.2010].
@@
 
-Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (Figure \@ref(fig:cwm-div), TABLE PIC SPOROS).
-SLA is an important part of the leaf economic spectrum, which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
+Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics).
+Specific leaf area is an important part of the leaf economic spectrum, which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
 slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a].
@@
 Indeed, the species with the lowest SLA values were epiphytes (\eg, *Davallia anderssonii* Mett.
-ex Kuhn, \SI{`r smallest_sla_1`} {\square\meter\per\kilo\gram}; *Ctenopterella blechnoides* (Grev.) Parris, \SI{`r smallest_sla_2`} {\square\meter\per\kilo\gram}), but this relationship was not significant across the dataset (Figure \@ref(fig:cwm-div), TABLE PIC SPOROS).
+ex Kuhn, \SI{`r smallest_sla_1`} {\square\meter\per\kilo\gram}; *Ctenopterella blechnoides* (Grev.) Parris, \SI{`r smallest_sla_2`} {\square\meter\per\kilo\gram}), but this relationship was not significant across the dataset (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
 The tradeoffs posited in the leaf economics spectrum have been most thoroughly documented in global studies including the widest possible range of taxa and environments, and do not always hold up in local studies [@Petter2016].
@@
 It is possible that other small-scale effects (\eg, microclimate) obscure this pattern in local studies such as ours and that of @Zhu2016.
-This is supported by the strong, parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (Figure \@ref(fig:cwm-div)).
+This is supported by the strong, parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (\fig \@ref(fig:cwm-div)).
 
@@
 Only gametophyte morphotype was found to have a significant association with epiphytic growth.
-Morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived ($< \SI{1}{\year}$), cordate gametophytes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate (\eg, elongate or filamentous) gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; Farrar2008].
+Morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived ($< \SI{1}{\year}$), cordate gametophytes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate (\eg, elongate or filamentous) gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; @Farrar2008].
 The complex, three-dimensional structure of non-cordate gametophytes slows their drying rates, and may be an adaptation to drought-prone epiphytic habitats [@WatkinsJr.2007c; @Pittermann2013], and promote out-crossing [@Farrar2008].
@@
 
-Presence of hairs and glands has strong phylogenetic signal (REF TABLE phylosig, REF TABLE supp pglmm gameto), and these traits are correlated with certain taxonomic groups [@Stokey1951; @Nayar1971].
+Presence of hairs and glands has strong phylogenetic signal (\tabs \@ref(tab:phylosig-binary), REF TABLE supp pglmm gameto), and these traits are correlated with certain taxonomic groups [@Stokey1951; @Nayar1971].
 Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], we did not detect a correlation between hair production and epiphytic growth.
@@
 
-Results for functional trait richness (FRic) supported our hypotheses: we observed lower FRic in epiphytic relative to terrestrial communities overall, and a significant positive correlation of FRic with humidity in epiphytes (\ie, decreased FRic at drier, low elevation plots and increased FRic at more humid, high elevation plots; Figure \@ref(fig:comm-div)).
+Results for functional trait richness (FRic) supported our hypotheses: we observed lower FRic in epiphytic relative to terrestrial communities overall, and a significant positive correlation of FRic with humidity in epiphytes (\ie, decreased FRic at drier, low elevation plots and increased FRic at more humid, high elevation plots; \fig \@ref(fig:comm-div)).
 However, results for the other two measures of functional diversity, evenness (FEve) and divergence (FDiv) did not match our predictions.
-Rather, FEve decreased with elevation (but not minimum RH) in terrestrial communities, and FDiv was lower in terrestrial communities overall while showing no trend along the gradient  (Figure \@ref(fig:comm-div)).
+Rather, FEve decreased with elevation (but not minimum RH) in terrestrial communities, and FDiv was lower in terrestrial communities overall while showing no trend along the gradient  (\fig \@ref(fig:comm-div)).
 The lack of a clear pattern in functional evenness and divergence suggests that within each trait pool, there is no significant difference in trait spacing between epiphytes and terrestrial species.
@@
 
-In addition to their decreased functional diversity, we also found that epiphytic communities are more phylogenetically clustered relative to terrestrial communities overall (Figure \@ref(fig:comm-div)).
+In addition to their decreased functional diversity, we also found that epiphytic communities are more phylogenetically clustered relative to terrestrial communities overall (\fig \@ref(fig:comm-div)).
 However, we did not detect significant trends in phylogenetic community structure of epiphytic and terrestrial communities along the elevational gradient.
@@
 
-Figure \@ref(fig:traits-used-image): Examples of traits used in this study.
+**\fig \@ref(fig:traits-used-image)**: Examples of traits used in this study.
 (**a**) Sporophyte traits.
@@
 
-Figure \@ref(fig:climate): Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter on Moorea, French Polynesia.
+**\fig \@ref(fig:climate)**: Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter  on Moorea, French Polynesia.
 Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from July 7, 2013 to July 5, 2014, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).
@@
 
-Figure \@ref(fig:pca): Principal components analysis (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
+**\fig \@ref(fig:pca)**: Principal components analysis (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
 **(a, b)** PC loadings.
@@
 
-Figure \@ref(fig:tree): Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
+**\fig \@ref(fig:tree)**: Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
 Relative value of quantitative (sporophyte) traits in heat gradient: low values are lighter, high values are darker.
@@
 Species missing data for six or more traits not shown.
-For a summary of traits, see Table \@ref(tab:traits-used).
+For a summary of traits, see \tab \@ref(tab:traits-used).
 Major Cretaceous epiphytic radiations \sensu @Schuettpelz2009 labeled on tree in green: H, hymenophylloids; T, trichomanoids, V, vittarioids; A, asplenioids; E, elaphoglossoids; P, polygrammoids (trichomanoids diversified prior to the Cretaceous but are included as an important extant epiphytic clade).
 
-Figure \@ref(fig:cwm-div): Change in selected functional traits of ferns on Moorea, French Polynesia across environmental gradients and between growth habit types (epiphytic or terrestrial).
+**\fig \@ref(fig:cwm-div)**: Change in selected functional traits of ferns on Moorea, French Polynesia across environmental gradients and between growth habit types (epiphytic or terrestrial).
 Epiphytic values in green, terrestrial values in brown.
@@
 
-Figure \@ref(fig:comm-div): Functional and phylogenetic diversity of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
+**\fig \@ref(fig:comm-div)**: Functional and phylogenetic diversity of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
 Epiphytic communities in green, terrestrial communities in brown.
@@
 Boxplots show median values (bold lines) of response variables by growth habit (epiphytic or terrestrial) across all communities (epiphytic communities, $\n = 17$; terrestrial communities, $\n = 17$).
-Hinges and whiskers as in Figure \@ref(fig:cwm-div).
+Hinges and whiskers as in \fig \@ref(fig:cwm-div).
 Asterisks indicate statistical significance of one-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+
+\pagebreak
 
@@
   kable_styling(full_width = FALSE) %>%
-  pack_rows("Sporophyte traits", 1, 6) %>%
-  pack_rows("Gametophyte traits", 7, 11) %>%
+  pack_rows("Sporophyte traits", 1, 7) %>%
+  pack_rows("Gametophyte traits", 8, 11) %>%
   column_spec(3, width = "15em")
@@
 
-```{r, results='asis', echo = FALSE, eval=(knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex')}
-# Add pagebreak if rendering to pdf
-cat('\\pagebreak')
+\pagebreak
+
+```{r phylosig-cont}
+loadd(phylosig_cont_traits)
+phylosig_cont_traits %>%
+  mutate_at(vars(kval, lambda), ~round_t(., 2)) %>%
+  mutate(lambda_pval = format_pval(lambda_pval)) %>%
+  column_to_rownames("trait") %>%
+  kable(caption = "Phylogenetic signal in continuous traits", format = "latex", booktabs = TRUE)
 ```
+
+\pagebreak
+
+```{r phylosig-binary}
+loadd(phylosig_binary_traits)
+phylosig_binary_traits %>%
+  mutate(D = round_t(D, 2)) %>%
+  column_to_rownames("trait") %>%
+  kable(caption = "Phylogenetic signal in binary traits", format = "latex", booktabs = TRUE)
+```
+
+\pagebreak
+
+```{r pics}
+loadd(pic_results)
+
+pic_results %>%
+  mutate_at(vars(estimate, std.error, statistic), ~round_t(., 2)) %>%
+  mutate(p.value = format_pval(p.value)) %>%
+  select(trait, n_contrasts, n_pos_con, statistic, p.value) %>%
+  column_to_rownames("trait") %>%
+  kable(caption = "Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth", format = "latex", booktabs = TRUE)
+```
+
+All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades.
+
+\pagebreak
+
+```{r corr-evo}
+
+loadd(correlated_evo_test)
+
+correlated_evo_test %>%
+  mutate(pval = format_pval(pval)) %>%
+  mutate_if(is.numeric, ~round_t(., 1)) %>%
+  column_to_rownames("trait") %>%
+  kable(caption = "Pagel's (1994) test of correlated evolution between binary traits", format = "latex", booktabs = TRUE)
+```
+
+Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
+terrestrial).
+Lower log likelihood for the dependent model indicates correlation with growth habit.
+
+\pagebreak
 
@@
 
+\pagebreak
+
 ```{r climate, fig.height = 4, fig.width = 6, fig.cap = "Microclimatic variables."}
@@
 ```
+
+\pagebreak
 
@@
 
+\pagebreak
+
 ```{r tree, fig.height = 8, fig.width = 6, fig.cap = "Traits on tree."}
@@
 
+\pagebreak
+
 ```{r comm-div, fig.height = 8, fig.width = 6, fig.cap = "Community diversity metrics."}
@@
 
-## Tables
+\pagebreak
 

### 2019-07-18 code/functions.R
Extract R2 and pvalues from models

@@ -925,7 +925,7 @@ analyze_fd_by_habit <- function (traits, comm, habit_type = c("epiphytic", "terr
 # Models ----
 
 #' Run a linear model on terrestrial and epiphytic species
-#' separately.
+#' separately and tidy the output parameters.
 #'
 #' @param data Input data. Must include column for "habit"
 #' (growth habit, either epiphtyic or terrestrial).
@@ -941,7 +941,7 @@ analyze_fd_by_habit <- function (traits, comm, habit_type = c("epiphytic", "terr
 #' rep("epiphtyic", 0.5*nrow(mtcars)))
 #' )
 #' run_lm_by_habit(mtcars_hab, "mpg", "wt")
-run_lm_by_habit <- function (data, indep_var, resp_var) {
+tidy_lm_by_habit <- function (data, indep_var, resp_var) {
   
   resp_var_enq <- enquo(resp_var)
   indep_var_enq <- enquo(indep_var)
@@ -961,6 +961,44 @@ run_lm_by_habit <- function (data, indep_var, resp_var) {
       TRUE ~ term
     ))
 }
+
+#' Run a linear model on terrestrial and epiphytic species
+#' separately and extract overall statistics (p value and
+#' r-squared).
+#'
+#' @param data Input data. Must include column for "habit"
+#' (growth habit, either epiphtyic or terrestrial).
+#' @param indep_var Independent variable in data
+#' @param resp_var Response variable in data
+#'
+#' @return Dataframe; model summary
+#'
+#' @examples
+#' # A rather ridiculous example
+#' mtcars_hab <- mutate(mtcars, habit = c(
+#' rep("terrestrial", 0.5*nrow(mtcars)), 
+#' rep("epiphtyic", 0.5*nrow(mtcars)))
+#' )
+#' glance_lm_by_habit(mtcars_hab, "mpg", "wt")
+glance_lm_by_habit <- function (data, indep_var, resp_var) {
+  
+  resp_var_enq <- enquo(resp_var)
+  indep_var_enq <- enquo(indep_var)
+  
+  data %>%
+    select(y = !!resp_var_enq, x = !!indep_var_enq, habit) %>%
+    nest(-habit) %>%
+    mutate(
+      resp_var = resp_var,
+      indep_var = indep_var,
+      model = map(data, ~lm(y ~ x, data = .)),
+      summ = map(model, glance)
+    ) %>%
+    select(-model, -data) %>%
+    unnest
+}
+
+
 
[…]

### 2019-07-18 code/plan.R

@@ -175,12 +175,24 @@ plan <- drake_plan(
   
   # Run linear models on each combination of variables for
   # terrestrial and epiphytic communities separately.
+  
+  # - Extract model parameter summaries (pval, intersect 
+  # for each part of model)
+  model_parameter_summaries = map2_df(
+    .x = args$indep_vars, 
+    .y = args$resp_vars, 
+    ~ tidy_lm_by_habit(diversity_data, indep_var = .x, resp_var = .y)
+  ),
+
+  # - Extract overall model summaries (pval, rsquared
+  # for overall model)
   model_summaries = map2_df(
     .x = args$indep_vars, 
     .y = args$resp_vars, 
-    ~ run_lm_by_habit(diversity_data, indep_var = .x, resp_var = .y)
+    ~ glance_lm_by_habit(diversity_data, indep_var = .x, resp_var = .y)
   ),
   
+  # Extract predicted model fits
   model_fits = map2_df(
     .x = args$indep_vars, 
     .y = args$resp_vars, 
@@ -226,7 +238,7 @@ plan <- drake_plan(
     .y = args$resp_vars, 
     ~ make_scatterplot_by_habit(
       model_fits, 
-      model_summaries, 
+      model_parameter_summaries, 
       x_var = .x, 
       y_var = .y,
       habit_colors = habit_colors)

### 2019-07-18 ms/manuscript.Rmd
Hide results when asserting
Update manuscript

@@
 }
+
+# Load analysis results
+loadd(climate_plot)
+loadd(combined_comm_div_plots)
+loadd(combined_cwm_plots)
+loadd(comm)
+loadd(correlated_evo_test)
+loadd(correlated_evo_test)
+loadd(diversity_data)
+loadd(fern_traits)
+loadd(model_summaries)
+loadd(pca_plot)
+loadd(pca_results)
+loadd(phylosig_binary_traits)
+loadd(phylosig_binary_traits)
+loadd(phylosig_cont_traits)
+loadd(pic_results)
+loadd(t_test_results)
+loadd(traits_with_tree)
 
@@
 Slope and intercept for variables related to temperature were very similar between epiphytic and terrestrial dataloggers, but differed more strongly for those related to humidity (Table S1, Figure \@ref(fig:climate).
-Minimum RH increased `r min_RH_epi_slope %>% signif(2) %>% multiply_by(100) %>% percent()` per \SI{100}{\meter} elevation for epiphytic dataloggers, but only `r min_RH_ter_slope %>% signif(2) %>% multiply_by(100) %>% percent()` for terrestrial dataloggers (linear model, $\rval^2 = \Sexpr{get_result(slopes.results, "r.squared", "min_RH")}$, $\pval = \Sexpr{get_result(slopes.results, "pval", "min_RH", TRUE)}$).
+Minimum RH increased `r min_RH_epi_slope %>% signif(2) %>% multiply_by(100) %>% percent()` per \SI{100}{\meter} elevation for epiphytic dataloggers, but only `r min_RH_ter_slope %>% signif(2) %>% multiply_by(100) %>% percent()` for terrestrial dataloggers (linear model, $\rval^2^ = \Sexpr{get_result(slopes.results, "r.squared", "min_RH")}$, $\pval = \Sexpr{get_result(slopes.results, "pval", "min_RH", TRUE)}$).
 The Mt.
@@
 # Extract summary stats for mention in text
-loadd(pca_results)
 
@@
 ```{r physig-numbers}
-loadd(phylosig_binary_traits)
 habit_d <- phylosig_binary_traits %>% filter(trait == "habit") %>% pull(D) %>% round_t(2)
@@
 ```{r corr-numbers}
-loadd(pic_results)
 stipe_pval <- pic_results %>% filter(trait == "stipe") %>% pull(p.value) %>% signif(2)
@@
 dissection_pval <- pic_results %>% filter(trait == "dissection") %>% pull(p.value) %>% signif(2)
-
-loadd(correlated_evo_test)
 morphotype_pval <- correlated_evo_test %>% filter(trait == "morphotype") %>% pull(pval) %>% signif(2)
-
 ```
@@
 Quantitatively similar results were obtained using PGLMMs, except that no significant relationship was observed between rhizome diameter and growth habit, whereas SLA did show a significant relationship with growth habit (REF SUPP TABLE).
+
+### Functional and Phylogenetic Diversity
+
+```{r comm-div-numbers}
+# Format model summaries for printing
+model_summ_print <-
+  model_summaries %>%
+  mutate(r = round_t(r.squared, 2), 
+         p = format_pval(p.value))
+
+# Format t-test results for printing
+t_test_results_print <-
+  t_test_results %>%
+  mutate(p = format_pval(p.value))
+
+# Define simple function for extracting value of interest from
+# diversity model results
+get_mod_result <- function(resp_var_select, habit_select, 
+                           indep_var_select, value) {
+  model_summ_print %>% 
+    filter(
+      resp_var == resp_var_select, 
+      habit == habit_select, 
+      indep_var == indep_var_select) %>% 
+    pull(!!enquo(value))
+}
+
+# Make tibble of abundances
+abundances <-
+  comm %>%
+  mutate(abundance = 
+           comm %>%
+           select_if(is.numeric) %>%
+           rowSums()
+  ) %>%
+  select(species, abundance)
+
+# Join this with trait data
+traits_with_abun <- left_join(
+  fern_traits, abundances
+)
+
+# Get total number of epiphytic and terrestrial species observed in all plots
+n_epi <- traits_with_abun %>% filter(abundance > 0, habit == "epiphytic") %>% 
+  pull(species) %>% n_distinct
+n_ter <- traits_with_abun %>% filter(abundance > 0, habit == "terrestrial") %>% 
+  pull(species) %>% n_distinct
+
+# Calculate mean richness by growth habit
+mean_rich <-
+  diversity_data %>%
+  group_by(habit) %>%
+  summarize(
+    mean_rich = mean(ntaxa) %>% round_t(1),
+    sd_rich = sd(ntaxa) %>% round_t(1)
+  ) %>%
+  mutate(stat = glue("{mean_rich} ± {sd_rich}"))
+```
+
+On a single-trait basis, distribution of community-weighted mean values (CWMs) showed significant correlation with elevation for some traits, but no significant relationship was detected for any traits with minimum RH (\fig \@ref(fig:cwm-div)).
+Specific leaf area decreased with elevation for epiphytic (linear model, \rval^2^ = `r get_mod_result("sla", "epiphytic", "el", "r")`, \pval = `r get_mod_result("sla", "epiphytic", "el", "p")`) and terrestrial (linear model, \rval^2^ = `r get_mod_result("sla", "terrestrial", "el", "r")`, \pval = `r get_mod_result("sla", "terrestrial", "el", "p")`) communities.
+Degree of lamina dissection decreased with elevation in epiphytes (linear model, \rval^2^ = `r get_mod_result("dissection", "epiphytic", "el", "r")`, \pval = `r get_mod_result("dissection", "epiphytic", "el", "p")`), but not terrestrial communities.
+Stipe length showed a small, but significant, increase with elevation for epiphytic communities (linear model, \rval^2^ = `r get_mod_result("stipe", "epiphytic", "el", "r")`, \pval = `r get_mod_result("stipe", "epiphytic", "el", "p")`).
+Rhizome diameter decreased with elevation for terrestrial communities (linear model, \rval^2^ = `r get_mod_result("rhizome", "terrestrial", "el", "r")`, \pval = `r get_mod_result("rhizome", "terrestrial", "el", "p")`), but increased slightly in epiphytes (linear model, \rval^2^ = `r get_mod_result("rhizome", "epiphytic", "el", "r")`, \pval = `r get_mod_result("rhizome", "epiphytic", "el", "p")`).
+
+We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial species.
+Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \si{\species}per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval = `r t_test_results_print %>% filter(resp_var == "ntaxa") %>% pull(p)`; \fig \@ref(fig:comm-div)).
+Species richness increased linearly with elevation and minimum RH for epiphytes, but not terrestrial communities (\fig \@ref(fig:comm-div)).
+Trait richness (FRic) was lower in epiphytic vs.
+terrestrial communities (two-sided \tval-test, \pval = `r t_test_results_print %>% filter(resp_var == "FRic") %>% pull(p)`), whereas functional diversity (FDiv) was lower in terrestrial communities (two-sided \tval-test, \pval = `r t_test_results_print %>% filter(resp_var == "FDiv") %>% pull(p)`), and functional evenness (FEve) did not differ between the two.
+FEve decreased with elevation, but not minimum RH, in terrestrial communities (\fig \@ref(fig:comm-div)).
+Epiphytic communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, \rval^2^ = `r get_mod_result("FRic", "epiphytic", "el", "r")`, \pval = `r get_mod_result("FRic", "epiphytic", "el", "p")`) and plots with greater minimum RH (linear model, \rval^2^ = `r get_mod_result("FRic", "epiphytic", "min_RH", "r")`, \pval = `r get_mod_result("FRic", "epiphytic", "min_RH", "p")`).
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (two-sided \tval-test, \pval = `r t_test_results_print %>% filter(resp_var == "mntd.obs.z") %>% pull(p)`) but not MPD (\pval = `r t_test_results_print %>% filter(resp_var == "mpd.obs.z") %>% pull(p)`; \fig \@ref(fig:comm-div)).
+No significant trend was detected in phylogenetic community structure with environment in either terrestrial or epiphytic communities (\fig \@ref(fig:comm-div)).
 
@@
 
-```{r sla-numbers}
-loadd(fern_traits)
-
+```{r sla-numbers, results = "hide"}
 # First smallest SLA: Humata_anderssonii
@@
 assert_that(fern_traits %>% arrange(sla) %>% slice(2) %>% pull(species) == "Ctenopterella_blechnoides")
-
 ```
@@
 ```{r phylosig-cont}
-loadd(phylosig_cont_traits)
 phylosig_cont_traits %>%
@@
 ```{r phylosig-binary}
-loadd(phylosig_binary_traits)
 phylosig_binary_traits %>%
@@
 ```{r pics}
-loadd(pic_results)
-
 pic_results %>%
@@
 ```{r corr-evo}
-
-loadd(correlated_evo_test)
-
 correlated_evo_test %>%
@@
 ```{r climate, fig.height = 4, fig.width = 6, fig.cap = "Microclimatic variables."}
-loadd(climate_plot)
 climate_plot
@@
 ```{r pca, fig.height = 6, fig.width = 6, fig.cap = "Trait PCA."}
-loadd(pca_plot)
 pca_plot
@@
 ```{r tree, fig.height = 8, fig.width = 6, fig.cap = "Traits on tree."}
-loadd(traits_with_tree)
 traits_with_tree
@@
 ```{r cwm-div, fig.height = 8, fig.width = 6, fig.cap = "Community-weighted means."}
-loadd(combined_cwm_plots)
 combined_cwm_plots
@@
 ```{r comm-div, fig.height = 8, fig.width = 6, fig.cap = "Community diversity metrics."}
-loadd(combined_comm_div_plots)
 combined_comm_div_plots

### 2019-07-18 ms/manuscript.Rmd
Use pearson's correlation to subset climatic vars and AIC to choose climate models

@@
 
-We calculated mean daily maximum, mean, minimum, and standard variation of temperature and RH from the raw climate data, then tested for differences in these summary statistics between epiphytic and terrestrial dataloggers using ANCOVA [@Scheffers2013].
-We used daily mean values of temperature and RH each in turn as the response variable, and constructed linear models with growth habit as the factor and elevation as the covariate, and with their interaction to test for differences in response of climatic variables to elevation between epiphytic and terrestrial dataloggers.
-We also calculated the number of drying events at various minimum RH values, as well as daily mean values for water potential, but found that these were both highly correlated with mean RH and did not analyze them further (data not shown).
+We calculated mean daily mean, minimum, and standard variation of temperature and RH for each site from the raw climate data, then grand means from the daily means.
+We analyzed correlation between the grand means with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (hereafter, "climatic variables").
+We constructed models to test for differences in response of climatic variables to elevation between epiphytic and terrestrial dataloggers.
+For each retained climatic variable, we constructed three linear models: one with growth habit (epiphytic vs.
+terrestrial) only, one with elevation only, and one with the interaction of elevation and growth habit.
+We compared the fit of the models using the Akaike Information Criteron (AIC) [@Akaike1973], and kept the model with the smallest AIC for each climatic variable.
+A single outlier site located on an exposed south-facing slope with unusually high temperature and low humidity (Mt.
+Rotui, 830 m) site was excluded from analysis.
 

### 2019-07-18 code/functions.R

@@ -206,7 +206,56 @@ get_grand_means <- function(daily_means) {
     )
 }
 
-#' Make climate models
+#' Check for correlation between climate variables
+#' and retain non-correlated ones
+#'
+#' @param climate_data Grand means of climate variables
+#' by site
+#'
+#' @return Grand means of climate variables subsetted
+#' to only those that are correlated less than 0.9
+#' 
+select_climate_vars <- function (climate_data) {
+  
+  # Calculate total correlation for all climatic variables
+  climate_corr_all <-
+    climate_data %>%
+    filter(is_outlier == "no") %>%
+    select_if(is.numeric) %>%
+    corrr::correlate()
+  
+  # Mean and min temp are correlated, but not with SD.
+  temp_corr <-
+    climate_corr_all %>%
+    corrr::focus(contains("temp"), mirror = TRUE)
+  
+  # (mean and min RH) and (min and SD RH) are correlated.
+  rh_corr <-
+    climate_corr_all %>%
+    corrr::focus(contains("RH"), mirror = TRUE)
+  
+  # None of the temp vars are correlated with Rel. Hum.
+  temp_rh_corr <-
+    climate_corr_all %>%
+    corrr::focus(contains("RH"))
+  
+  # Drop variables correlated > 0.9 and check what the results look like
+  climate_corr_select <-
+    climate_corr_all %>%
+    corrr::focus(-max_temp, -min_temp, -min_RH, mirror = TRUE)
+  
+  # Double-check that we've dropped any correlations > 0.9
+  # and keep these as the final variables for analysis.
+  selected_climate_vars <-
+  climate_corr_select %>%
+    assertr::assert(function (x) x < 0.9 | is.na(x), -starts_with("rowname")) %>%
+    pull(rowname)
+  
+  climate_data %>%
+    select(site, habit, selected_climate_vars, is_outlier)
+}
+
+#' Make climate models and choose the best one using AIC
 #'
 #' @param climate_data 
 #' @param site_data 
@@ -240,36 +289,21 @@ make_climate_models <- function (climate_data, site_data) {
     select(-data) %>%
     gather(model_type, model, -var)
   
-  # Run analysis of variance on each interaction model and filter to only
-  # the significant parts of the model. Use this to determine which of the
-  # possible set of models is best.
-  aov_results <-
-    data_for_modeling %>%
-    # Construct a linear model for each variable, and pull out summaries
-    mutate(
[…]

### 2019-07-18 code/plan.R

@@ -61,6 +61,10 @@ plan <- drake_plan(
   
   # Climate analysis ----
   
+  # Subset climate variables to only those with correlation
+  # coefficients less than 0.9
+  grand_mean_climate_select = select_climate_vars(grand_mean_climate),
+  
   # Make full set of climate models including
   # each climate var by elevation, growth habit, and
   # their interaction, then choose the best model for each.

### 2019-07-18 code/setup.R

@@ -20,6 +20,7 @@ library(ggtree)
 library(magrittr)
 library(viridis)
 library(cowplot)
+library(corrr)
 library(tidyverse)
 
 conflict_prefer("filter", "dplyr")

### 2019-07-18 ms/manuscript.Rmd
Use mean instead of min RH
- Move round_t and format_pval to functions from Rmd
- Add function for making single climate plots

@@
 conflict_prefer("filter", "dplyr")
-
-# Function for rounding with trailing zeros
-round_t <- function (x, digits) round(x, digits) %>% sprintf(glue::glue("%.{digits}f"), .)
-
-# Function for formatting p-values: round to 3 digits,
-# or just say "< 0.001"
-format_pval <- function (x) {
-  case_when(
-    x < 0.001 ~ "< 0.001",
-    TRUE ~ round(x, 3) %>% as.character()
-  )
-}
 
@@
 
+Mean and standard deviation of temperature and relative humidity were retained after excluding variables with correlation coefficients >0.9.
+
 ### Principal Components Analysis of Trait Data
@@
 
+<!-- NEED TO UPDATE TO REFLECT mean_RH instead of min_RH -->
 On a single-trait basis, distribution of community-weighted mean values (CWMs) showed significant correlation with elevation for some traits, but no significant relationship was detected for any traits with minimum RH (\fig \@ref(fig:cwm-div)).
@@
 FEve decreased with elevation, but not minimum RH, in terrestrial communities (\fig \@ref(fig:comm-div)).
-Epiphytic communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, \rval^2^ = `r get_mod_result("FRic", "epiphytic", "el", "r")`, \pval = `r get_mod_result("FRic", "epiphytic", "el", "p")`) and plots with greater minimum RH (linear model, \rval^2^ = `r get_mod_result("FRic", "epiphytic", "min_RH", "r")`, \pval = `r get_mod_result("FRic", "epiphytic", "min_RH", "p")`).
+Epiphytic communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, \rval^2^ = `r get_mod_result("FRic", "epiphytic", "el", "r")`, \pval = `r get_mod_result("FRic", "epiphytic", "el", "p")`) and plots with greater minimum RH (linear model, \rval^2^ = `r get_mod_result("FRic", "epiphytic", "mean_RH", "r")`, \pval = `r get_mod_result("FRic", "epiphytic", "mean_RH", "p")`).
 Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (two-sided \tval-test, \pval = `r t_test_results_print %>% filter(resp_var == "mntd.obs.z") %>% pull(p)`) but not MPD (\pval = `r t_test_results_print %>% filter(resp_var == "mpd.obs.z") %>% pull(p)`; \fig \@ref(fig:comm-div)).

### 2019-07-18 code/functions.R

@@ -160,11 +160,9 @@ get_daily_means <- function (climate_raw) {
   climate_raw %>%
     group_by(site, date, habit) %>%
     summarize(
-      max_temp = max(temp),
       mean_temp = mean(temp),
       min_temp = min(temp),
       sd_temp = sd(temp),
-      max_RH = max(RH),
       mean_RH = mean(RH),
       min_RH = min(RH),
       sd_RH = sd(RH)
@@ -188,11 +186,9 @@ get_grand_means <- function(daily_means) {
   daily_means %>%
     group_by(site, habit) %>%
     summarize(
-      max_temp = mean(max_temp),
       mean_temp = mean(mean_temp),
       min_temp = mean(min_temp),
       sd_temp = mean(sd_temp),
-      max_RH = mean(max_RH),
       mean_RH = mean(mean_RH),
       min_RH = mean(min_RH),
       sd_RH = mean(sd_RH)
@@ -242,7 +238,7 @@ select_climate_vars <- function (climate_data) {
   # Drop variables correlated > 0.9 and check what the results look like
   climate_corr_select <-
     climate_corr_all %>%
-    corrr::focus(-max_temp, -min_temp, -min_RH, mirror = TRUE)
+    corrr::focus(-min_temp, -min_RH, mirror = TRUE)
   
   # Double-check that we've dropped any correlations > 0.9
   # and keep these as the final variables for analysis.
@@ -271,7 +267,7 @@ make_climate_models <- function (climate_data, site_data) {
     # REMOVE OUTLIER: Mt Rotui exposed slope
     filter(is_outlier == "no") %>%
     # Reformat data into nested set by variable (mean and SD of temp and RH)
-    select(site, habit, el, mean_temp, sd_temp, min_RH, sd_RH) %>%
+    select(-lat, -long, -is_outlier) %>%
     gather(var, value, -site, -habit, -el) %>%
     nest(-var)
   
@@ -1186,23 +1182,99 @@ match_comm_and_tree <- function (comm, phy, return = c("comm", "tree")) {
   
 }
 
+# Function for rounding with trailing zeros
+round_t <- function (x, digits) {
+  round(x, digits) %>% sprintf(glue::glue("%.{digits}f"), .)
+}
+
+# Function for formatting p-values: round to 3 digits,
+# or just say "< 0.001"
+format_pval <- function (x, equals_sign = FALSE) {
+  case_when(
+    x < 0.001 ~ "< 0.001",
[…]

### 2019-07-18 code/plan.R

@@ -67,9 +67,10 @@ plan <- drake_plan(
   
   # Make full set of climate models including
   # each climate var by elevation, growth habit, and
-  # their interaction, then choose the best model for each.
+  # their interaction, then choose the best model for each
+  # (doesn't include single outlier site)
   climate_models = make_climate_models(
-    climate_data = grand_mean_climate, 
+    climate_data = grand_mean_climate_select, 
     site_data = moorea_sites
   ),
   
@@ -165,7 +166,7 @@ plan <- drake_plan(
   # (includes quantitative, i.e., sporophyte, traits only)
   
   # Merge all diversity metrics.
-  diversity_data = list(comm_struc, func_div, grand_mean_climate, moorea_sites) %>% 
+  diversity_data = list(comm_struc, func_div, grand_mean_climate_select, moorea_sites) %>% 
     reduce(left_join) %>%
     mutate(habit = fct_relevel(habit, c("terrestrial", "epiphytic"))),
   
@@ -174,7 +175,7 @@ plan <- drake_plan(
     resp_vars = c(
       "ntaxa", "mpd.obs.z", "mntd.obs.z", "FDiv", "FEve", "FRic",
       "stipe", "length", "width", "dissection", "pinna", "sla", "rhizome"), 
-    indep_vars = c("el", "min_RH")
+    indep_vars = c("el", "mean_RH")
   )),
   
   # Run linear models on each combination of variables for
@@ -215,8 +216,8 @@ plan <- drake_plan(
     set_names(levels(moorea_climate_raw$habit)),
   
   # Make climate plot.
-  climate_plot = make_climate_plot(
-    climate_data = grand_mean_climate,
+  climate_plot = make_four_part_climate_plot(
+    climate_data = grand_mean_climate_select,
     site_data = moorea_sites,
     fits = climate_model_fits, 
     summaries = climate_model_summaries, 

### 2019-07-22 ms/manuscript.Rmd
Make major changes to climate analysis:
- Add VPD

- Square-root transform some climate variables

- Select non-correlated climate variables and diversity metrics

- Use full-subset GAMs to analyze contribution of each climate variable to each diversity metric

@@
 The final dataset included 201 days of data for 13 epiphytic dataloggers and 13 terrestrial dataloggers ($\n = 26$).
+We calculated vapor pressure deficit (VPD) from relative humidity and temperature as implemented in the \package{plantecophys} R package [@Duursma2015].
 
@@
 
-We calculated mean daily mean, minimum, and standard variation of temperature and RH for each site from the raw climate data, then grand means from the daily means.
+We calculated mean daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each site from the raw climate data, then grand means from the daily means.
 We analyzed correlation between the grand means with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (hereafter, "climatic variables").
-We constructed models to test for differences in response of climatic variables to elevation between epiphytic and terrestrial dataloggers.
-For each retained climatic variable, we constructed three linear models: one with growth habit (epiphytic vs.
-terrestrial) only, one with elevation only, and one with the interaction of elevation and growth habit.
-We compared the fit of the models using the Akaike Information Criteron (AIC) [@Akaike1973], and kept the model with the smallest AIC for each climatic variable.
+We tested for differences between epiphytic and terrestrial dataloggers while accounting for elevation with ANCOVA, using growth habit as the categorical variable and elevation as the covariate for each climatic response variable.
+For visualizing the results, we selected the linear model with the lowest Akaike Information Criterion (AIC) out of the set of models including interaction effects and the effect of each independent variable alone.
 A single outlier site located on an exposed south-facing slope with unusually high temperature and low humidity (Mt.
@@
 Selection, transformation, and scaling of traits was performed similarly as for the PCA, except that the method used to calculate distance [@Gower1971] can accommodate missing data, so we included some species with observations missing for some traits ($\n = 126$ \si{\species} total).
-We calculated each metric separately for epiphytic and terrestrial communities, and compared results between the two.
-Under a scenario of environmental filtering (as hypothesized for epiphytic communities), we expect relatively lower values for the three metrics of functional diversity [@Savage2012].
-We also investigated relationships between functional trait diversity and environment with linear models by testing for dependence of each functional trait diversity metric against elevation and daily minimum RH.
-We chose these independent variables because we found that overall climate tended to change with elevation, and to test if minimum RH specifically had a filtering effect on diversity along the gradient.
+
+Prior to testing for differences in diversity metrics, we first checked for correlation between metrics with Pearson's correlation coefficient, and retained a set of metrics with absolute correlation coefficients < 0.9.
+We then tested for differences between epiphytic and terrestrial communities with a paired t-test.
+We also conducted a more thorough analysis simultaneously testing the effects of environment and growth habit on diversity metrics using a full-subsets multiple regression approach with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
+For each diversity metric, we built a set of models (Gaussian family, with site as a random effect) taking into account all pairwise interactions of explanatory variables, and ranked these by corrected AIC (AICc) to find the models that best explain the data.
+Our dataset included `r nrow(diversity_data) - nrow(grand_mean_climate_select)` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
+Since we found that all environmental variables are related to elevation (see Results), we also constructed linear models each diversity metric by growth habit with elevation as the covariate.
 

### 2019-07-22 code/functions.R

@@ -146,7 +146,8 @@ reformat_raw_climate <- function (moorea_climate_raw) {
         str_detect(site,  "ter") ~ "terrestrial"
       ),
       habit = fct_relevel(habit, c("epiphytic", "terrestrial")),
-      site = str_remove_all(site, "_epi|_ter")
+      site = str_remove_all(site, "_epi|_ter"),
+      vpd = plantecophys::RHtoVPD(RH, temp)
     )
 }
 
@@ -159,13 +160,14 @@ reformat_raw_climate <- function (moorea_climate_raw) {
 get_daily_means <- function (climate_raw) {
   climate_raw %>%
     group_by(site, date, habit) %>%
-    summarize(
-      mean_temp = mean(temp),
-      min_temp = min(temp),
-      sd_temp = sd(temp),
-      mean_RH = mean(RH),
-      min_RH = min(RH),
-      sd_RH = sd(RH)
+    summarize_at(
+      vars(temp, RH, vpd),
+      list(
+        max = ~ max(., na.rm = TRUE),
+        mean = ~ mean(., na.rm = TRUE),
+        min = ~ min(., na.rm = TRUE),
+        sd = ~ sd(., na.rm = TRUE)
+      )
     ) %>%
     ungroup() %>%
     mutate(
@@ -185,14 +187,7 @@ get_daily_means <- function (climate_raw) {
 get_grand_means <- function(daily_means) {
   daily_means %>%
     group_by(site, habit) %>%
-    summarize(
-      mean_temp = mean(mean_temp),
-      min_temp = mean(min_temp),
-      sd_temp = mean(sd_temp),
-      mean_RH = mean(mean_RH),
-      min_RH = mean(min_RH),
-      sd_RH = mean(sd_RH)
-    ) %>%
+    summarize_if(is.numeric, ~ mean(., na.rm = TRUE)) %>%
     ungroup() %>%
     mutate(
       is_outlier = case_when(
@@ -216,11 +211,11 @@ select_climate_vars <- function (climate_data) {
   # Calculate total correlation for all climatic variables
   climate_corr_all <-
     climate_data %>%
-    filter(is_outlier == "no") %>%
+    # filter(is_outlier == "no") %>%
     select_if(is.numeric) %>%
     corrr::correlate()
   
-  # Mean and min temp are correlated, but not with SD.
+  # Mean and min temp are correlated
   temp_corr <-
     climate_corr_all %>%
[…]

### 2019-07-22 code/plan.R

@@ -21,13 +21,14 @@ plan <- drake_plan(
   # gameto traits only known from taxonomy
   fern_traits_strict = filter(fern_traits, gameto_source != "T"),
   
-  # Site data
+  # Site data with elevation
   moorea_sites = mooreaferns::sites %>%
     as_tibble %>%
     filter(str_detect(site, "Aorai", negate = TRUE)),
   
   # Climate data
   # - Raw climate data (rel. humidity and temperature every 15 min.)
+  # - Also calculate vapor pressure deficit (VPD) from temp and RH
   moorea_climate_raw = mooreaferns::moorea_climate %>% 
     as_tibble %>% 
     reformat_raw_climate,
@@ -62,32 +63,60 @@ plan <- drake_plan(
   # Climate analysis ----
   
   # Subset climate variables to only those with correlation
-  # coefficients less than 0.9
-  grand_mean_climate_select = select_climate_vars(grand_mean_climate),
+  # coefficients less than 0.9,
+  # and add elevation.
+  climate_select = select_climate_vars(grand_mean_climate) %>%
+    left_join(select(moorea_sites, site, el)),
   
-  # Make full set of climate models including
+  # Make three versions of climate data for further analysis:
+  # - Square-root transform selected variables
+  # so they have a more normal distribution.
+  climate_trans = climate_select %>%
+    mutate_at(vars(temp_sd, vpd_mean, vpd_min, vpd_sd), sqrt),
+  
+  # - Remove the outliers, but don't transform.
+  climate_no_outlier = climate_select %>%
+    filter(is_outlier == "no"),
+  
+  # - Remove outliers and transform.
+  climate_trans_no_outlier = climate_trans %>%
+    filter(is_outlier == "no"),
+  
+  # - Make vector of climate variables for modeling
+  climate_vars = climate_select %>%
+    select(-site, -habit, -el, -is_outlier) %>%
+    colnames() %>%
+    rlang::set_names(),
+  
+  # Run ANCOVA
+  climate_ancova_results = run_ancova(
+    data = climate_trans_no_outlier, 
+    resp_vars = climate_vars),
+  
+  #### Make models for plotting climate ###
+  
+  # - Make full set of climate models including
   # each climate var by elevation, growth habit, and
-  # their interaction, then choose the best model for each
-  # (doesn't include single outlier site)
-  climate_models = make_climate_models(
[…]

### 2019-07-22 code/setup.R

@@ -21,6 +21,8 @@ library(magrittr)
 library(viridis)
 library(cowplot)
 library(corrr)
+library(plantecophys)
+library(FSSgam)
 library(tidyverse)
 
 conflict_prefer("filter", "dplyr")

### 2019-07-23 ms/manuscript.Rmd
Add SI

@@
       in_header: newphyt.sty
+header-includes: 
+  \usepackage{float}
+  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
 editor_options: 
@@
 # Load analysis results
+loadd(best_fit_div_models)
+loadd(climate_ancova_results)
 loadd(climate_plot)
+loadd(climate_select)
 loadd(combined_comm_div_plots)
@@
 loadd(correlated_evo_test)
-loadd(correlated_evo_test)
-loadd(diversity_data)
+loadd(div_el_model_summaries)
+loadd(div_t_test_results)
+loadd(div_metrics_select)
 loadd(fern_traits)
-loadd(model_summaries)
+loadd(importance_heatmap)
 loadd(pca_plot)
@@
 loadd(phylosig_binary_traits)
-loadd(phylosig_binary_traits)
 loadd(phylosig_cont_traits)
 loadd(pic_results)
-loadd(t_test_results)
 loadd(traits_with_tree)
-
 ```
@@
 
-A fern community survey was carried out on Moorea comprising 17, \SI{10}{\meter} $\times$ \SI{10}{\meter} plots spanning an elevational gradient from \SIrange{20}{2000}{\meter}, with plots placed \circa every \SI{200}{\meter} @[Nitta2017].
+A fern community survey was carried out on Moorea comprising 17, \SI{10}{\meter} $\times$ \SI{10}{\meter} plots spanning an elevational gradient from \SIrange{20}{2000}{\meter}, with plots placed \circa every \SI{200}{\meter} [@Nitta2017].
 Most plots were located along trails leading to the three main peaks on the island, Mt.
@@
 Morphological traits of gametophytes were all qualitative (either categorical or binary; \tab \@ref(tab:traits-used)).
-Trait states were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017], grown from spores of known species in the lab, or taken from the literature.
+Traits were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017] or grown from spores of known species in the lab, taken from the literature, or inferred from taxonomic group (i.e., genus) if documented as a stable taxonomic character.
 For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie, heart-shaped) or non-cordate.
@@
 
+Trait values and data sources are summarized in Table S1.
+
 ### Statistical analysis
 
-We calculated mean daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each site from the raw climate data, then grand means from the daily means.
-We analyzed correlation between the grand means with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (hereafter, "climatic variables").
-We tested for differences between epiphytic and terrestrial dataloggers while accounting for elevation with ANCOVA, using growth habit as the categorical variable and elevation as the covariate for each climatic response variable.
-For visualizing the results, we selected the linear model with the lowest Akaike Information Criterion (AIC) out of the set of models including interaction effects and the effect of each independent variable alone.
-A single outlier site located on an exposed south-facing slope with unusually high temperature and low humidity (Mt.
-Rotui, 830 m) site was excluded from analysis.
+We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each site from the raw climate data, then overall means from the daily values.
+We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (hereafter, "climatic variables").
+Variables relating to VPD, which have no meaning if less than zero, and some other variables that had skewed distributions, were square-root transformed prior to fitting models.
+We tested for differences between epiphytic and terrestrial dataloggers while accounting for elevation with ANCOVA, using growth habit as the categorical variable and elevation as the covariate.
+We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone and their interactions, then selecting the one with the lowest Akaike Information Criterion (AIC).
+A single outlier site with unusually high temperature and low humidity (Mt.
+Rotui, 830 m) site was excluded from ANCOVA and linear models.
+Mt.
+Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established, which are part of a mountain complex on the interior of the island.
 
@@
 We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], which also conducts a significance test by comparing the observed value of \D with distributions of simulated values produced under one of two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM threshold model.
-The probability of obtaining the observed value of \D under either model is then calculated.
+To account for possible circularity in measuring the phylogenetic signal of traits that were scored based on taxonomy, we conducted the same analyses on a dataset excluding species that had been scored this way ("strict" dataset).
 
@@
 We conducted standard and phylogenetically corrected PCA, using the \function{PCA} and \function{phyl.pca} functions in the \package{FactoMineR} [@Le2008] and \package{phytools} R packages, respectively.
-The position of each species was plotted in the multi-dimensional trait space using the first two PC axes, then color-coded according to growth habit (epiphytic vs.
-terrestrial).
 
@@
 A significantly better fit of the dependent model indicates that evolution of the two traits is correlated.
-To verify the effect of analysis method on our results, we performed alternative tests for correlation of growth habit with functional traits using phylogenetic generalized linear mixed models (PGLMMs; see Methods S1 and \tabs S1 and S2).
+The same analyses were also done on the "strict" dataset.
 
@@
 Prior to testing for differences in diversity metrics, we first checked for correlation between metrics with Pearson's correlation coefficient, and retained a set of metrics with absolute correlation coefficients < 0.9.
-We then tested for differences between epiphytic and terrestrial communities with a paired t-test.
-We also conducted a more thorough analysis simultaneously testing the effects of environment and growth habit on diversity metrics using a full-subsets multiple regression approach with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
-For each diversity metric, we built a set of models (Gaussian family, with site as a random effect) taking into account all pairwise interactions of explanatory variables, and ranked these by corrected AIC (AICc) to find the models that best explain the data.
-Our dataset included `r nrow(diversity_data) - nrow(grand_mean_climate_select)` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
-Since we found that all environmental variables are related to elevation (see Results), we also constructed linear models each diversity metric by growth habit with elevation as the covariate.
+We tested for differences between epiphytic and terrestrial communities with a paired \tval-test.
+We simultaneously tested the effects of environment and growth habit on diversity metrics using a full-subsets multiple regression approach with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
+For each diversity metric, we built a set of models (Gaussian family, with site as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by corrected AIC (AICc) to find the models that best explain the data.
+Our dataset included `r nrow(div_metrics_select) - nrow(climate_select)` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
+Since we found that all climatic variables are related to elevation (see Results), we also constructed linear models for each diversity metric using same model selection procedure as in the climate analysis.
 
@@
 
-```{r env-survey}
-# loadd(climate_model_summaries)
-# loadd(climate_models_parameter_summaries)
-# 
-# # Extract the slope of the minimum RH for epiphytes
-# # (change in min. % rel humidity per meter).
-# min_RH_epi_slope <- climate_models_parameter_summaries %>%
-#   filter(var == "min_RH", term == "el") %>%
-#   pull(estimate)
-# 
-# # Extract the slope of the minimum RH for terrestrial species.
-# # In the results table, terrestrial slope is the difference
-# # between the estimate for the elevation term and the
-# # interaction term (add though because it's already negative).
-# min_RH_ter_slope <-
-#   magrittr::add(
-#     min_RH_epi_slope,
-#     climate_models_parameter_summaries %>%
-#       filter(var == "min_RH", term == "el:habitterrestrial") %>%
-#       pull(estimate)
-# )
-
-# There were significant interaction effects between growth habit and elevation on all microclimatic variables except for minimum temperature (Table S1).
-Slope and intercept for variables related to temperature were very similar between epiphytic and terrestrial dataloggers, but differed more strongly for those related to humidity (Table S1, Figure \@ref(fig:climate).
-Minimum RH increased `r min_RH_epi_slope %>% signif(2) %>% multiply_by(100) %>% percent()` per \SI{100}{\meter} elevation for epiphytic dataloggers, but only `r min_RH_ter_slope %>% signif(2) %>% multiply_by(100) %>% percent()` for terrestrial dataloggers (linear model, $\rval^2^ = \Sexpr{get_result(slopes.results, "r.squared", "min_RH")}$, $\pval = \Sexpr{get_result(slopes.results, "pval", "min_RH", TRUE)}$).
-The Mt.
-Rotui \SI{830}{\meter} plot was a clear outlier, with greater mean temperature and lower minimum RH relative to other plots at similar elevation (outliers in  \fig \ref{fig:climate} at \circa \SI{800}{\meter}).
-Mt.
-Rotui is a solitary peak located between Cook's Bay and Oponohu Bay.
-It is more isolated with greater exposure than the other two mountains where the rest of the plots were established, which are part of a mountain complex on the interior of the island (J.
-Nitta, pers.
-obs.).
-
-```
-
-Mean and standard deviation of temperature and relative humidity were retained after excluding variables with correlation coefficients >0.9.
+Mean, maximum, and standard deviation (SD) of temperature and mean, minimum, and SD of vapor pressure deficit (VPD) were retained after excluding variables with correlation coefficients > 0.9.
+Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (\tab S2; \fig \ref{fig:climate}).
+Variables related to VPD differed by growth habit and elevation.
+Mean and maximum or minimum temperature and VPD all decreased with elevation, whereas SD of temperature increased, but SD of VPD decreased with elevation.
+Epiphytic dataloggers had higher values than terrestrial ones for variables related to VPD.
 
@@
 The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
-The trait loadings were similar between the two analyses, except that most switched axes; for example, SLA was loaded on PC1 in the standard analysis but PC2 in the phylogenetic analysis (REF SUPP TABLE pca).
+The trait loadings were similar between the two analyses, except that most switched axes; for example, SLA was loaded on PC1 in the standard analysis but PC2 in the phylogenetic analysis.
 
@@
 All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmmae \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under a scenario of evolution by BM (\tab \@ref(tab:phylosig-binary)).
-Phylogenetic generalized linear mixed models (PGLMMs) also indicated significant signal in all gametophyte traits (REF SUPP TABLE).
 Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; \tab \@ref(tab:phylosig-binary)).
+Similar results were obtained using the "strict" dataset (\tab S3).
 
@@
 cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval = `r morphotype_pval`; \tab \@ref(tab:corr-evo)).
-Quantitatively similar results were obtained using PGLMMs, except that no significant relationship was observed between rhizome diameter and growth habit, whereas SLA did show a significant relationship with growth habit (REF SUPP TABLE).
+Similar results were obtained using the "strict" dataset (\tab S4).
 
@@
 model_summ_print <-
-  model_summaries %>%
+  div_el_model_summaries %>%
   mutate(r = round_t(r.squared, 2), 
@@
 # Format t-test results for printing
-t_test_results_print <-
-  t_test_results %>%
+t_test_print <-
+  div_t_test_results %>%
   mutate(p = format_pval(p.value))
@@
 # diversity model results
-get_mod_result <- function(resp_var_select, habit_select, 
-                           indep_var_select, value) {
-  model_summ_print %>% 
-    filter(
-      resp_var == resp_var_select, 
-      habit == habit_select, 
-      indep_var == indep_var_select) %>% 
+get_mod_result <- function(data, resp_var_select, value) {
+  data %>% 
+    filter(var == resp_var_select) %>% 
     pull(!!enquo(value))
@@
 mean_rich <-
-  diversity_data %>%
+  div_metrics_select %>%
   group_by(habit) %>%
@@
 
-<!-- NEED TO UPDATE TO REFLECT mean_RH instead of min_RH -->
-On a single-trait basis, distribution of community-weighted mean values (CWMs) showed significant correlation with elevation for some traits, but no significant relationship was detected for any traits with minimum RH (\fig \@ref(fig:cwm-div)).
-Specific leaf area decreased with elevation for epiphytic (linear model, \rval^2^ = `r get_mod_result("sla", "epiphytic", "el", "r")`, \pval = `r get_mod_result("sla", "epiphytic", "el", "p")`) and terrestrial (linear model, \rval^2^ = `r get_mod_result("sla", "terrestrial", "el", "r")`, \pval = `r get_mod_result("sla", "terrestrial", "el", "p")`) communities.
-Degree of lamina dissection decreased with elevation in epiphytes (linear model, \rval^2^ = `r get_mod_result("dissection", "epiphytic", "el", "r")`, \pval = `r get_mod_result("dissection", "epiphytic", "el", "p")`), but not terrestrial communities.
-Stipe length showed a small, but significant, increase with elevation for epiphytic communities (linear model, \rval^2^ = `r get_mod_result("stipe", "epiphytic", "el", "r")`, \pval = `r get_mod_result("stipe", "epiphytic", "el", "p")`).
-Rhizome diameter decreased with elevation for terrestrial communities (linear model, \rval^2^ = `r get_mod_result("rhizome", "terrestrial", "el", "r")`, \pval = `r get_mod_result("rhizome", "terrestrial", "el", "p")`), but increased slightly in epiphytes (linear model, \rval^2^ = `r get_mod_result("rhizome", "epiphytic", "el", "r")`, \pval = `r get_mod_result("rhizome", "epiphytic", "el", "p")`).
+In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (\fig \ref{fig:heatmap}).
+Most of the the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (\tab S5).
+Community-weighted means of frond length, frond width, stipe length, and rhizome diameter were correlated (Pearson's correlation coefficient > 0.9), so of these we only analyzed stipe length.
+
+Distribution of community-weighted mean values (CWMs) differed between epiphytic and terrestrial communities overall (pinna number, stipe length) or co-varied with elevation (dissection, SLA) (\fig \@ref(fig:cwm-div)).
+Specific leaf area decreased with elevation, and was lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval = `r get_mod_result(model_summ_print, "sla", "p")`).
+Degree of lamina dissection decreased with elevation in epiphytes, but slightly increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval = `r get_mod_result(model_summ_print, "dissection", "p")`).
+Stipe length and pinna number showed no significant variation with elevation, but were much lower overall for epiphytes relative to terrestrial communities (\tval-test, both \pval = `r get_mod_result(t_test_print, "stipe", "p")`).
 
 We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial species.
-Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \si{\species}per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval = `r t_test_results_print %>% filter(resp_var == "ntaxa") %>% pull(p)`; \fig \@ref(fig:comm-div)).
-Species richness increased linearly with elevation and minimum RH for epiphytes, but not terrestrial communities (\fig \@ref(fig:comm-div)).
+Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \si{\species}per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "ntaxa", "p")`; \fig \@ref(fig:comm-div)).
+Species richness increased  with elevation for both epiphytes and terrestrial communities (\fig \@ref(fig:comm-div)).
 Trait richness (FRic) was lower in epiphytic vs.
-terrestrial communities (two-sided \tval-test, \pval = `r t_test_results_print %>% filter(resp_var == "FRic") %>% pull(p)`), whereas functional diversity (FDiv) was lower in terrestrial communities (two-sided \tval-test, \pval = `r t_test_results_print %>% filter(resp_var == "FDiv") %>% pull(p)`), and functional evenness (FEve) did not differ between the two.
-FEve decreased with elevation, but not minimum RH, in terrestrial communities (\fig \@ref(fig:comm-div)).
-Epiphytic communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, \rval^2^ = `r get_mod_result("FRic", "epiphytic", "el", "r")`, \pval = `r get_mod_result("FRic", "epiphytic", "el", "p")`) and plots with greater minimum RH (linear model, \rval^2^ = `r get_mod_result("FRic", "epiphytic", "mean_RH", "r")`, \pval = `r get_mod_result("FRic", "epiphytic", "mean_RH", "p")`).
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (two-sided \tval-test, \pval = `r t_test_results_print %>% filter(resp_var == "mntd.obs.z") %>% pull(p)`) but not MPD (\pval = `r t_test_results_print %>% filter(resp_var == "mpd.obs.z") %>% pull(p)`; \fig \@ref(fig:comm-div)).
-No significant trend was detected in phylogenetic community structure with environment in either terrestrial or epiphytic communities (\fig \@ref(fig:comm-div)).
+terrestrial communities (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "FRic", "p")`), whereas functional diversity (FDiv) was lower in terrestrial communities (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "FDiv", "p")`), and functional evenness (FEve) did not differ between the two.
+Both epiphytic and terrestrial communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "FRic", "r")`, \pval = `r get_mod_result(model_summ_print, "FRic", "p")`).
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (two-sided \tval-test, \pval = `get_mod_result(t_test_print, "mntd.obs.z", "p")`) but not MPD (\pval = `r get_mod_result(t_test_print, "mpd.obs.z", "p")`; \fig \@ref(fig:comm-div)).
+No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (\fig \@ref(fig:comm-div)).
 
@@
 
-The sporophyte traits that differed most strongly between epiphytic and terrestrial taxa were related to plant size, including stipe and frond length, frond width, and rhizome diameter.
-Stipe length in particular differed strongly between epiphytes and terrestrial species (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
+The sporophyte traits that differed most strongly between epiphytic and terrestrial taxa were related to plant size, including stipe and frond length, frond width, and rhizome diameter (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
 Stipe length may be a critical trait controlling whole leaf hydraulic conductance; since ferns have resistive stipes in general, species growing in water-limited habitats are expected to decrease stipe length to minimize resistance [@WatkinsJr.2010].
 Our results support previous non-phylogenetically controlled studies that found a prevalence of short stipes in epiphytic ferns [@WatkinsJr.2010; Creese2011].
-However, given that many of these size-related traits scale together in ferns [@Arcand2008; Creese2011], careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
+However, given that these were all correlated with each other (Pearson's correlation coefficient > 0.9), and similar traits have been shown to scale together previously [@Arcand2008; Creese2011], careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
 
@@
 
-Presence of hairs and glands has strong phylogenetic signal (\tabs \@ref(tab:phylosig-binary), REF TABLE supp pglmm gameto), and these traits are correlated with certain taxonomic groups [@Stokey1951; @Nayar1971].
+Presence of hairs and glands has strong phylogenetic signal (\tab \@ref(tab:phylosig-binary)), and these traits are correlated with certain taxonomic groups [@Stokey1951; @Nayar1971].
 Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], we did not detect a correlation between hair production and epiphytic growth.
@@
 
-Results for functional trait richness (FRic) supported our hypotheses: we observed lower FRic in epiphytic relative to terrestrial communities overall, and a significant positive correlation of FRic with humidity in epiphytes (\ie, decreased FRic at drier, low elevation plots and increased FRic at more humid, high elevation plots; \fig \@ref(fig:comm-div)).
+Growth habit and temperature are clearly important factors in most of the community diversity metrics we studied, as shown by the full-subsets model analysis (\fig \ref{fig:heatmap}).
+This is also supported by the linear elevation models (\figs \ref{fig:cwm-div}, \ref{fig:comm-div}), given that temperature shows a strong linear relationship with elevation (\fig \ref{fig:climate}).
+It is somewhat puzzling that VPD did not contribute significantly to any models, as we hypothesized that this would be important for epiphytes that are frequently subjected to water stress.
+One reason could be due to conditions at our study site: periods of prolonged rain result in the humidity sensor tip becoming wet, and it does not function normally again (i.e., shows 100% RH) until it dries out.
+The raw data show several extended periods of 100% RH in terrestrial dataloggers at higher elevations.
+Despite these possible inaccuracies however, epiphytic habitats are clearly drier than terrestrial (\fig \ref{fig:climate}).
+
+Results for functional trait richness (FRic) supported our hypotheses: we observed much lower FRic in epiphytic relative to terrestrial communities overall, and a significant positive correlation of FRic with elevation (\fig \@ref(fig:comm-div)).
 However, results for the other two measures of functional diversity, evenness (FEve) and divergence (FDiv) did not match our predictions.
-Rather, FEve decreased with elevation (but not minimum RH) in terrestrial communities, and FDiv was lower in terrestrial communities overall while showing no trend along the gradient  (\fig \@ref(fig:comm-div)).
-The lack of a clear pattern in functional evenness and divergence suggests that within each trait pool, there is no significant difference in trait spacing between epiphytes and terrestrial species.
+Rather, FDiv was lower in terrestrial communities overall but did not show any trend along the gradient, while FEve did not show any significant differences at all (\fig \@ref(fig:comm-div)).
+Functional divergence is similar to richness, but takes into account abundances; so the FDiv results may be driven by a small number of abundant species.
+The lack of a clear pattern in FEve suggests that within each trait pool, there is no significant difference in trait spacing between epiphytes and terrestrial species.
 This could be maintained by niche partitioning within epiphytic and terrestrial habitats.
@@
 
-In addition to their decreased functional diversity, we also found that epiphytic communities are more phylogenetically clustered relative to terrestrial communities overall (\fig \@ref(fig:comm-div)).
+In addition to their decreased functional diversity, we also found that epiphytic communities are more phylogenetically clustered relative to terrestrial communities when measured with MNTD, but not MPD (\fig \@ref(fig:comm-div)).
+This makes sense given that most epiphytes have radiated within five major clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely-related species from two or more epiphytic clades.
 However, we did not detect significant trends in phylogenetic community structure of epiphytic and terrestrial communities along the elevational gradient.
@@
 J.H.N. conceived and carried out the study. R.T. collected data. J.H.N. performed the analyses. J.H.N., J.E.W., and C.C.D. wrote the manuscript.
-
-## References
 
@@
 
-\pagebreak
+\clearpage
 
 ## Tables
+
+\clearpage
 
@@
 #   knitr::kable(caption = "Traits used", booktabs = TRUE)
-
 ```
 
-\pagebreak
+\clearpage
 
@@
 
-\pagebreak
+\clearpage
 
@@
   mutate(D = round_t(D, 2)) %>%
+  mutate(trait = factor(
+    trait, 
+    levels = c("habit", "gemmae", "glands", "hairs","morphotype"))) %>%
+  mutate(trait = lvls_revalue(
+    trait,
+    new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
+  arrange(trait) %>%
+  select(trait, `Number of presences` = num_present, `Number of absences` = num_absent,
+         P_rnd = prob_random, P_bm = prob_brownian) %>%
 column_to_rownames("trait") %>%
@@
   column_to_rownames("trait") %>%
-  kable(caption = "Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth", format = "latex", booktabs = TRUE)
+  kable(caption = "Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth", format = "latex", booktabs = TRUE) %>%
+  footnote(general = "All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades.", general_title = "", threeparttable = TRUE)
 ```
 
-All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades.
-
-\pagebreak
+\clearpage
 
@@
   column_to_rownames("trait") %>%
-  kable(caption = "Pagel's (1994) test of correlated evolution between binary traits", format = "latex", booktabs = TRUE)
+  kable(caption = "Pagel's (1994) test of correlated evolution between binary traits", format = "latex", booktabs = TRUE) %>%
+  footnote(general = "Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
+terrestrial).
+Lower log likelihood for the dependent model indicates correlation with growth habit.", general_title = "", threeparttable = TRUE)
 ```
 
-Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
-terrestrial).
-Lower log likelihood for the dependent model indicates correlation with growth habit.
-
-\pagebreak
+\clearpage
 
@@
 
-\pagebreak
+\clearpage
 
-```{r cwm-div, fig.height = 8, fig.width = 6, fig.cap = "Community-weighted means."}
+```{r heatmap, fig.height = 8, fig.width = 4, fig.cap = "Variable importance scores."}
+importance_heatmap
+```
+
+\clearpage
+
+```{r cwm-div, fig.height = 8, fig.width = 4, fig.cap = "Community-weighted means."}
 combined_cwm_plots
@@
 
-\pagebreak
+\clearpage
 
-```{r comm-div, fig.height = 8, fig.width = 6, fig.cap = "Community diversity metrics."}
+```{r comm-div, fig.height = 8, fig.width = 4, fig.cap = "Community diversity metrics."}
 combined_comm_div_plots

### 2019-07-23 si/SI.rmd
Add SI

---
title: ""
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)

# Load analysis results
loadd(trait_data_for_si)
loadd(climate_ancova_results)
loadd(phylosig_binary_traits_strict)
loadd(correlated_evo_test_strict)
loadd(best_fit_div_models)
```

***New Phytologist*** **Supporting Information**

Article title: Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns

Authors: Joel H. Nitta, James E. Watkins, Jr., Ravahere Taputuarai, and Charles C. Davis

Article acceptance date:

The following Supporting Information is available for this article:

**Table S1** Trait values of ferns on Moorea, French Polynesia. Abbreviations: SLA, specific leaf area. For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen, etc.

```{r}
trait_data_for_si %>%
  write_csv(here::here("si/Table_S1.csv"))
```

**Table S2** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate. Abbreviations: d.f., degrees of freedom; MS, mean squares.

```{r climate-ancova}
# Format and write out table as CSV file
climate_ancova_results %>% 
  mutate(
    p.value = signif(p.value, 2) %>% format.pval(., digits = 1),
    meansq = round_t(meansq, 2),
    term = case_when(
      term == "habit" ~ "Growth habit",
      term == "el" ~ "Elevation",
      term == "habit:el" ~ "Growth habit × elevation",
      TRUE ~ term
    )
  ) %>%
  select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value) %>%
  write_csv(here::here("si/Table_S2.csv"))

```

**Table S3** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy. Abbreviations: P_rnd, probability of obtaining the observed value of *D* under random trait evolution; P_bm, probability of obtaining the observed value of *D* under simulation of trait evolution by Brownian motion.

```{r phylosig-binary-strict}
# Format and write out table as CSV file
phylosig_binary_traits_strict %>%
  mutate(D = round_t(D, 2)) %>%
  mutate(trait = factor(
    trait, 
    levels = c("habit", "gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
  arrange(trait) %>%
  select(Trait = trait, `Number of presences` = num_present, `Number of absences` = num_absent, D,
         P_rnd = prob_random, P_bm = prob_brownian) %>%
  write_csv(here::here("si/Table_S3.csv"))
```

**Table S4** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy. Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs. terrestrial). Lower log likelihood for the dependent model indicates correlation with growth habit. Abbreviations: LL, log likelihood.

```{r corr-evo-strict}
# Format and write out table as CSV file
correlated_evo_test_strict %>%
  mutate(trait = factor(
    trait, 
    levels = c("gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Gemmae", "Glands", "Hairs", "Morphotype"))) %>%
  arrange(trait) %>%
  select(
    Trait = trait,
    `LL (independent model)` = logL_indep,
    `LL (dependent model)` = logL_dep,
    `Likelihood ratio` = likelihood_ratio,
    `P` = pval
  ) %>%
  write_csv(here::here("si/Table_S4.csv"))
```

**Table S5** Best-fitting models explaining community diversity metrics by climate and growth habit. Models within delta AICc 3 of the best-fitting model shown. Abbreviations: FDiv, functional diversity; FEve, functional evenness; FRic, functional richness; MPD, mean phylogenetic distance; MNTD, mean nearest taxon distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion. Values of MPD and MNTD are the standard effect sizes calculated from comparison of observed values with null distribution (see Methods). Trait names refer to community-weighted mean values (see Methods).

```{r best-fit-mods}
# Format and write out table as CSV file
best_fit_div_models %>%
  mutate(resp_var = factor(
        resp_var, 
        levels = c("ntaxa", "mpd.obs.z", "mntd.obs.z", 
                   "FDiv", "FEve", "FRic",
                   "dissection", "sla", "stipe", "pinna"))) %>%
  mutate(
    resp_var = lvls_revalue(
        resp_var, 
        new_levels = c("Richness", "MPD", "MNTD", 
                       "FDiv", "FEve", "FRic",
                       "Dissection", "SLA", "Stipe length", "Pinna no.")) 
  ) %>%
  arrange(resp_var, delta_AICc) %>%
  mutate(
    modname = modname %>%
      str_replace_all("\\+", " \\+ ") %>%
      str_replace_all("\\.by\\.", " × ") %>%
      str_replace_all("habit", "Growth habit") %>%
      str_replace_all("temp_mean", "Mean temp.") %>%
      str_replace_all("temp_sd", "SD temp.") %>%
      str_replace_all("temp_max", "Max. temp.") %>%
      str_replace_all("vpd_mean", "Mean VPD") %>%
      str_replace_all("vpd_sd", "SD VPD") %>%
      str_replace_all("vpd_min", "Min. VPD")
  ) %>% 
  select(
    `Metric` = resp_var,
    `Model` = modname,
    AICc,
    `R2` = r2,
    delta_AICc
  ) %>%
  write_csv(here::here("si/Table_S5.csv"))
```

### 2019-07-23 code/plan.R

@@ -333,6 +333,11 @@ plan <- drake_plan(
   # Write out MS ----
   ms = rmarkdown::render(
     knitr_in("ms/manuscript.Rmd"),
+    quiet = TRUE),
+  
+  # Write out supplemental information
+  si = rmarkdown::render(
+    knitr_in("si/SI.Rmd"),
     quiet = TRUE)
   
 )

### 2019-07-23 code/setup.R

@@ -19,6 +19,7 @@ library(picante)
 library(ggtree)
 library(magrittr)
 library(viridis)
+library(scico)
 library(cowplot)
 library(corrr)
 library(plantecophys)

### 2019-07-28 ms/manuscript.Rmd
Add check of model residuals with Moran's I

@@
 We tested for differences between epiphytic and terrestrial dataloggers while accounting for elevation with ANCOVA, using growth habit as the categorical variable and elevation as the covariate.
-We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone and their interactions, then selecting the one with the lowest Akaike Information Criterion (AIC).
+We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone and their interactions, then selecting the one with the lowest corrected Akaike Information Criterion (AIC~c~) [@Akaike1973].
 A single outlier site with unusually high temperature and low humidity (Mt.
@@
 Mt.
-Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established, which are part of a mountain complex on the interior of the island.
+Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established, which are part of a mountain complex on the interior of the island (J.
+Nitta, pers.
+obs.).
 
@@
 
-Prior to testing for differences in diversity metrics, we first checked for correlation between metrics with Pearson's correlation coefficient, and retained a set of metrics with absolute correlation coefficients < 0.9.
-We tested for differences between epiphytic and terrestrial communities with a paired \tval-test.
-We simultaneously tested the effects of environment and growth habit on diversity metrics using a full-subsets multiple regression approach with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
-For each diversity metric, we built a set of models (Gaussian family, with site as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by corrected AIC (AICc) to find the models that best explain the data.
+We checked for correlation between diveristy metrics with Pearson's correlation coefficient, and retained a set of metrics with absolute correlation coefficients < 0.9.
+We tested for differences in diversity metrics between epiphytic and terrestrial communities with a paired \tval-test.
+We simultaneously tested the effects of climate and growth habit on diversity metrics using a full-subsets multiple regression approach with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
+For each diversity metric, we built a set of models (Gaussian family, with site as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
 Our dataset included `r nrow(div_metrics_select) - nrow(climate_select)` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
-Since we found that all climatic variables are related to elevation (see Results), we also constructed linear models for each diversity metric using same model selection procedure as in the climate analysis.
+Since we found that all climatic variables are related to elevation (see Results), we also constructed linear models for each diversity metric using the same model selection procedure as in the climate analysis.
+We checked for spatial autocorrelation in the model residuals with Moran's *I* [@Moran1950] as implemented in the \package{spdep} R package [@Bivand2013].
 

### 2019-07-28 code/functions.R

@@ -1131,6 +1131,50 @@ get_best_fss_mods <- function(fssgam_results) {
   
 }
 
+# Spatial autocorrelation ----
+
+#' Check for spatial autocorrelation in residuals of models
+#' selected using full subsets analysis with Moran's I 
+#'
+#' @param model_set List of models; results of running full subsets analysis
+#' as a loop over response variables
+#' @param resp_var Response variable to select from models
+#' @param best_mod Name of best model
+#' @param site_data Dataframe with names and gps points (long, lat) 
+#' of sites.
+#'
+#' @return Nested tibble with results of running Moran's I test
+#' for spatial autocorrelation in residuals
+#' 
+check_moran_fss <- function (model_set, resp_var, best_mod, site_data) {
+  
+  # resp_var, best_mod must be character or will select wrong model (if factor)!
+  resp_var <- as.character(resp_var)
+  best_mod <- as.character(best_mod)
+  
+  # Extract best-fitting model residuals from fss results, and add long-lats.
+  model_fits <- model_set[[resp_var]][["success.models"]][[best_mod]] %>%
+    broom::augment() %>%
+    dplyr::left_join(site_data)
+  
+  # Convert to spatial dataframe
+  sp::coordinates(model_fits) <- c("long", "lat")
+  sp::proj4string(model_fits) <- sp::CRS("+proj=longlat +datum=WGS84 +no_defs")
+  
+  # Make distance matrix
+  dist_mat <- 1/as.matrix(dist(sp::coordinates(model_fits)))
+  diag(dist_mat) <- 0
+  # Some sites have the exact same GPS points; repalce Inf values for these with 0
+  dist_mat[is.infinite(dist_mat)] <- 0
+  
+  # Conduct Moran's test on the residuals.
+  # This will tell us if there is any remaining unexplained variation
+  # in the model fit that is due to spatial autocorrelation
+  # spdep::moran.test(model_fits$.resid, spdep::mat2listw(dist_mat)) %>% broom::tidy()
+  spdep::moran.mc(model_fits$.resid, spdep::mat2listw(dist_mat), 10000) %>% broom::tidy()
+}
+
+
 # Misc ----
 
 #' Match trait data and tree

### 2019-07-28 code/plan.R

@@ -67,7 +67,7 @@ plan <- drake_plan(
   # coefficients less than 0.9,
   # and add elevation.
   climate_select = select_climate_vars(grand_mean_climate) %>%
-    left_join(select(moorea_sites, site, el)),
+    left_join(moorea_sites),
   
   # Make three versions of climate data for further analysis:
   # - Square-root transform selected variables
@@ -85,7 +85,7 @@ plan <- drake_plan(
   
   # - Make vector of climate variables for modeling
   climate_vars = climate_select %>%
-    select(-site, -habit, -el, -is_outlier) %>%
+    select(-site, -habit, -el, -is_outlier, -lat, -long) %>%
     colnames() %>%
     rlang::set_names(),
   
@@ -207,7 +207,7 @@ plan <- drake_plan(
   # - Subset diversity metrics to only those with correlation
   # coefficients less than 0.9, and add elevation and outlier status.
   div_metrics_select = select_div_metrics(div_metrics_all) %>%
-    left_join(select(moorea_sites, site, el)) %>%
+    left_join(moorea_sites) %>%
     mutate(
       is_outlier = case_when(
         site == "Rotui_800m_slope" ~ "yes",
@@ -223,7 +223,7 @@ plan <- drake_plan(
   
   # - Make named vector of response variables to test
   resp_vars = div_metrics_select %>%
-    select(-site, -habit, -el, -is_outlier) %>% 
+    select(-site, -habit, -el, -is_outlier, -long, -lat) %>% 
     colnames %>% rlang::set_names(),
   
   # - Run full-subsets model analysis.
@@ -240,6 +240,23 @@ plan <- drake_plan(
   
   # - Extract table of best-fit models.
   best_fit_div_models = get_best_fss_mods(fss_div_results),
+  
+  # - Test for spatial autocorrelation in residuals with Moran's I.
+  best_fit_div_models_moran = mutate(
+      best_fit_div_models,
+      moran_test = map2(
+        .x = resp_var, .y = modname,
+        ~ check_moran_fss(
+          model_set = fss_div_results,
+          resp_var = .x,
+          best_mod = .y,
+          site_data = moorea_sites
+        )
+      )
+    ) %>%
+    unnest() %>%
+    select(resp_var, modname, AICc, r2, delta_AICc, 
+           morans_I = statistic, morans_I_pval = p.value),
   
   ### Run linear models of diversity metrics by elevation ###
   

### 2019-07-28 ms/manuscript.Rmd
Update MS

@@
 ex Kuhn, \SI{`r smallest_sla_1`} {\square\meter\per\kilo\gram}; *Ctenopterella blechnoides* (Grev.) Parris, \SI{`r smallest_sla_2`} {\square\meter\per\kilo\gram}), but this relationship was not significant across the dataset (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
-The tradeoffs posited in the leaf economics spectrum have been most thoroughly documented in global studies including the widest possible range of taxa and environments, and do not always hold up in local studies [@Petter2016].
+The tradeoffs posited in the leaf economics spectrum have been most thoroughly documented in global studies including the widest possible range of taxa and environments, and do not always hold up in local studies [@Petter2016; @Messier2017].
 For example, @Zhu2016 did not find a significant relationship between SLA and leaf life span in a survey of 16 ferns across a disturbance (light) gradient in China.

### 2019-08-01 code/functions.R
Don't subset tree for mpd, mntd

@@ -1042,23 +1042,19 @@ analyze_comm_struc_by_habit <- function (comm, phy, traits) {
   
   ### Prepare data ###
   
-  # Keep only species in community with trait data
-  comm <- comm %>%
-    filter(species %in% traits$species)
-  
   # Add growth habit to community data
   comm <- left_join(
     comm,
     select(traits, species, habit)
     ) %>%
+    # Make sure habit isn't missing for any species
   assert(not_na, habit)
   
-  # check for mismatches/missing species between phy and comm
-  phy <- match_comm_and_tree(comm, phy, "tree")
-  comm <- match_comm_and_tree(comm, phy, "comm")
-  
-  # Make sure that worked
-  assert_that(isTRUE(all.equal(comm$species, phy$tip.label)))
+  # Make sure all species are in the phylogeny
+  # (but don't subset phylogeny to only species in Moorea
+  # communities - we want to use the species pool including
+  # both ferns of Moorea and Tahiti fro mpd, mntd).
+  assert_that(all(comm$species %in% phy$tip.label))
   
   # Split communities into epiphytic / terrestrial taxa, 
   # but keep together in a single dataframe so we can
@@ -1111,7 +1107,7 @@ analyze_comm_struc_by_habit <- function (comm, phy, traits) {
     cophenetic(phy), 
     null.model = "phylogeny.pool", 
     abundance.weighted = TRUE, 
-    runs = 999, iterations = 1000) %>%
+    runs = 999) %>%
     rownames_to_column("site") %>%
     as_tibble
   
@@ -1121,7 +1117,7 @@ analyze_comm_struc_by_habit <- function (comm, phy, traits) {
     cophenetic(phy), 
     null.model = "phylogeny.pool", 
     abundance.weighted = TRUE, 
-    runs = 999, iterations = 1000) %>%
+    runs = 999) %>%
     rownames_to_column("site") %>%
     as_tibble
   

### 2019-08-01 ms/manuscript.Rmd
Update MS
- Check numbers using code

@@
 loadd(correlated_evo_test)
+loadd(daily_mean_climate)
 loadd(div_el_model_summaries)
@@
 loadd(pic_results)
+loadd(species_list)
 loadd(traits_with_tree)
@@
 The list of species in each plot was split into epiphytic and terrestrial groups and treated separately during data analysis, hereafter referred to as 'epiphytic' and 'terrestrial' communities.
-Vouchers were deposited at GH and UC.
+Vouchers were deposited at GH and UC (abbreviations follow Index Herbariorum [@Thiers2019]).
 
 ### Environmental Survey
+
+```{r climate-numbers}
+# Count total number of days in climate data
+total_days <- daily_mean_climate %>% pull(date) %>% n_distinct()
+
+# Check start and end dates
+# - start July 7, 2013
+assert_that(daily_mean_climate %>% pull(date) %>% min == "2013-07-07")
+# - end July 5, 2014
+assert_that(daily_mean_climate %>% pull(date) %>% max == "2014-07-05")
+
+# Count total number of sites in climate data
+total_sites <-
+daily_mean_climate %>%
+  select(site, habit) %>%
+  unique() %>%
+  count(habit)
+```
 
@@
 Other days missing data for any datalogger during the final survey period were excluded from the dataset.
-The final dataset included 201 days of data for 13 epiphytic dataloggers and 13 terrestrial dataloggers ($\n = 26$).
+The final dataset included `r total_days` days of data for `r total_sites %>% filter(habit == "epiphytic") %>% pull(n)` epiphytic dataloggers and `r total_sites %>% filter(habit == "terrestrial") %>% pull(n)` terrestrial dataloggers.
 We calculated vapor pressure deficit (VPD) from relative humidity and temperature as implemented in the \package{plantecophys} R package [@Duursma2015].
@@
 For comparative analyses, we used the dated phylogenetic tree of [@Nitta2017], which includes all species from Moorea and some from the neighboring island of Tahiti.
-We trimmed the tree to species occurring only on Moorea ($\n = 130$ \si{\species}) using the \function{drop.tip} function of the \package{ape} package [@Paradis2004] in R [@RCoreTeam2015].
+We trimmed the tree to species occurring only on Moorea (*n* = `r length(species_list)` \si{\species}) using the \function{drop.tip} function of the \package{ape} package [@Paradis2004] in R [@RCoreTeam2015].
 We tested for phylogenetic signal in quantitative traits using two metrics that relate the observed trait values to those expected under a model of trait evolution following Brownian motion (BM), Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].

### 2019-08-01 code/functions.R
Move vector of selected traits outside of function

@@ -683,7 +683,8 @@ transform_traits <- function (traits,
 #'
 #' @param traits Dataframe of pre-processed traits
 #' @param phy Phylogeny
-#' @param analysis 
+#' @param cont_traits Vector of names of continuous traits
+#' to include in the analysis 
 #'
 #' @return List of dataframes including
 #'   species_locs: Positions of species along PC axes
@@ -694,24 +695,23 @@ transform_traits <- function (traits,
 #' "standard" for standard PCA (not using phylogeny), 
 #' or "phylo" for PCA including phylogeny.
 #' 
-run_trait_PCA <- function (traits, phy) {
+run_trait_PCA <- function (traits, phy, 
+                           cont_traits = c("sla", "stipe", "length", "width", 
+                                           "rhizome", "pinna", "dissection")) {
   
   ### Data wrangling ###
   
-  # Make vector of continuous traits to include in PCA
-  cont_vars <- c(
-    "sla", "stipe", "length", "width", "rhizome", "pinna"
-  )
-  
   # Prepare trait data
   traits <- traits %>%
+    # Make sure all selected continuous traits are in the trait data
+    verify(all(cont_traits %in% colnames(.))) %>%
     # Log-transform and scale
     transform_traits() %>%
     # Subset to continuous traits
-    select(species, habit, sla, stipe, length, width, rhizome, pinna) %>%
+    select(species, habit, cont_traits) %>%
     # Keep only completely sampled species
     filter(complete.cases(.)) %>%
-    # Keep only species in phylogeny
+    # Keep only species in phylogeny, in phylogenetic order
     match_traits_and_tree(traits = ., phy = phy, "traits") 
   
   # Trim to only species with trait data
@@ -721,7 +721,7 @@ run_trait_PCA <- function (traits, phy) {
   assert_that(isTRUE(all.equal(traits$species, phy$tip.label)))
   
   # Both standard and phylo PCA expect data frames with row names
-  traits_df_for_pca <- select(traits, species, cont_vars) %>%
+  traits_df_for_pca <- select(traits, species, cont_traits) %>%
     as.data.frame
   rownames(traits_df_for_pca) <- traits_df_for_pca$species
   traits_df_for_pca$species <- NULL

### 2019-08-01 ms/manuscript.Rmd
Update MS

@@
 
-```{r climate-numbers}
+```{r climate-numbers, results="hide"}
 # Count total number of days in climate data
@@
 The final dataset included `r total_days` days of data for `r total_sites %>% filter(habit == "epiphytic") %>% pull(n)` epiphytic dataloggers and `r total_sites %>% filter(habit == "terrestrial") %>% pull(n)` terrestrial dataloggers.
-We calculated vapor pressure deficit (VPD) from relative humidity and temperature as implemented in the \package{plantecophys} R package [@Duursma2015].
+We calculated vapor pressure deficit (VPD) from relative humidity and temperature using the \package{plantecophys} R package [@Duursma2015].
 
@@
 
+```{r check-num-traits}
+# Check the number of species included in trait PCA
+cont_traits = c("sla", "stipe", "length", "width", 
+                "rhizome", "pinna", "dissection")
+
+quant_traits_nsp <-
+fern_traits %>%
+  # Subset to continuous traits
+  select(species, habit, cont_traits) %>%
+  # Keep only completely sampled species
+  filter(complete.cases(.)) %>%
+  nrow
+```
+
 We used a multivariate approach to investigate differences in traits between epiphytes and terrestrial species.
-We used only quantitative traits, and included only species that had no missing data for any traits ($\n = 104$ \si{\species}).
+We used only quantitative traits, and included only species that had no missing data for any traits (*\n* = `r quant_traits_nsp` \si{\species}).
 We first log-transformed traits as necessary to attain normality, then scaled each trait by its range.
@@
 Functional richness does not use abundance data, but functional evenness and divergence can be weighted by abundance, so we preformed analyses for these metrics with abundance weighting to correct for the effect of rare species with unusual morphologies.
-Selection, transformation, and scaling of traits was performed similarly as for the PCA, except that the method used to calculate distance [@Gower1971] can accommodate missing data, so we included some species with observations missing for some traits ($\n = 126$ \si{\species} total).
+Selection, transformation, and scaling of traits was performed similarly as for the PCA, except that the method used to calculate distance [@Gower1971] can accommodate missing data, so we included species with observations missing for some traits.
 
@@
 ```
-
-<!-- CHECK PCA CODE: missing dissection -->
 
@@
 Both epiphytic and terrestrial communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "FRic", "r")`, \pval = `r get_mod_result(model_summ_print, "FRic", "p")`).
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (two-sided \tval-test, \pval = `get_mod_result(t_test_print, "mntd.obs.z", "p")`) but not MPD (\pval = `r get_mod_result(t_test_print, "mpd.obs.z", "p")`; \fig \@ref(fig:comm-div)).
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "mntd.obs.z", "p")`) but not MPD (\pval = `r get_mod_result(t_test_print, "mpd.obs.z", "p")`; \fig \@ref(fig:comm-div)).
 No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (\fig \@ref(fig:comm-div)).

### 2019-08-05 ms/manuscript.Rmd
Update MS to use mpd, mntd instead of FRic etc

@@
 ---
-title: "Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns"
+title: "Constrained, yet diverse: Phylogenetic and morphological filtering in epiphytic ferns"
 bibliography: "references.bib"
@@
 
-We analyzed community phylogenetic diversity using the Mean Phylogenetic Distance (MPD) and Mean Nearest Taxon Distance (MNTD) indices of [@Webb2000].
-MPD is the mean phylogenetic distance of each species in a community to all other species in the same community; MNTD is the mean phylogenetic distance of each species in a community to its sister species in the same community.
-Communities with low MPD or MNTD are 'phylogenetically clustered' (co-occurring species are closely related to each other), whereas those with high MPD or MNTD are 'phylogenetically overdispersed' (co-occurring species are distantly related to each other).
-MPD therefore provides a measure of phylogenetic diversity across the whole tree, while MNTD is more sensitive to distances towards the tips.
-Alone, MPD and MNTD are not very informative; to assess their statistical significance, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{1000} randomized communities (equivalent to the Net Relatedness Index and Nearest Taxon Index, respectively, of @Webb2000 with the sign reversed).
-Observed values above 95\% or less than 5\% of the null distribution were considered significant.
+We used distance-based methods to analyze phylogenetic and functional diversity.
+Mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) describe the mean distance between all species in a community (MPD) or the mean distance between the closest pairs of species in a community (MNTD) [@Webb2000].
+These metrics were originally formulated for phylogenetic distances (thus "phylogenetic" in the name), but can be applied in the same way to trait-based distances, enabling a direct comparison between the two [@Sessa2018].
+For functional diversity, we used Gower's dissimilarity coeffecient [@Gower1971] because it can use continuous, binary, and qualitative traits, and accepts missing data.
+We log-transformed traits as needed to account for skew, excluded frond width and length (both frond size traits highly correlated with stipe length), and weighted sporophyte and gametophyte traits equally before calculating the distance matrix using the \function{gowdis} function in the R pacakge \package{FD}.
+
+Raw values of MPD and MNTD are statistically related to species richness [@Swenson2014]; to obtain metrics that are uncorrelated with species richness, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{999} randomized communities (equivalent to the Net Relatedness Index and Nearest Taxon Index, respectively, of @Webb2000 with the sign reversed).
 Selection of the randomization procedure can strongly bias outcomes [@Gotelli2006;@Ulrich2012].
 Our null hypothesis is that fern spores are easily capable of dispersal between terrestrial and epiphytic habitats, and across the relatively short distances separating our plots.
-We therefore constructed null communities by drawing species randomly from a regional pool including terrestrial and epiphytic taxa.
-The regional pool included all known species of ferns from Moorea, and all species from the neighboring island of Tahiti with genetic data available ($\n = 146$ \si{\species} total; [@Nitta2017]).
-We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate community phylogenetic diversity metrics.
+We therefore constructed null communities using the "independent swap" algorithm of @Gotelli2000 that maintains species frequency and abundance while randomizing species identities.
+We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available (`r Ntip(phy)` \si{\species} total; [@Nitta2017]) for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct`) for functional null communities.
+We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate phylogenetic and functional diversity metrics.
 
-We analyzed community functional trait diversity using the \function{dbFD} function in the \package{FD} R package [@Laliberte2011].
-This function takes as input values any number of traits for species in a given community, plots them in a multidimensional trait space using PCA, and calculates several summary statistics.
-Functional richness (FRic) is equivalent to convex hull volume containing all observed trait values plotted in multidimensional trait space [@Cornwell2006], functional evenness (FEve) is the average distance to the center of all the traits, and functional dispersion (FDis) is the mean distance between traits [@Villeger2008].
-We chose these three metrics because they are statistically independent and can each provide different information about the structure of community trait diversity [@Savage2012].
-Functional richness does not use abundance data, but functional evenness and divergence can be weighted by abundance, so we preformed analyses for these metrics with abundance weighting to correct for the effect of rare species with unusual morphologies.
-Selection, transformation, and scaling of traits was performed similarly as for the PCA, except that the method used to calculate distance [@Gower1971] can accommodate missing data, so we included species with observations missing for some traits.
-
-We checked for correlation between diveristy metrics with Pearson's correlation coefficient, and retained a set of metrics with absolute correlation coefficients < 0.9.
 We tested for differences in diversity metrics between epiphytic and terrestrial communities with a paired \tval-test.

### 2019-08-05 ms/manuscript.Rmd
Update MS
- remove references to FD metrics

- add discussion of unexpectedly diverse epiphytic comms

@@
 - We observed phylogenetic conservatism in multiple traits related to epiphytic growth in fern sporophytes and gametophytes.
-Epiphytes had significantly shorter stipes, smaller rhizomes, and a higher frequency of non-cordate gametophytes while accounting for phylogeny.
-Epiphytic communities were clustered both phylogenetically and functionally, consistent with the hypothesis that the canopy acts as a filter due to its low humidity.
+Epiphytes had significantly shorter fronds, smaller rhizomes, and a higher frequency of non-cordate gametophytes.
+Epiphytic communities were clustered phylogenetically but diverse functionally, indicating that the canopy acts as a filter due to its low humidity to limit size while simultaneously enabling diversification into finely divided niches.
 
@@
 Most of the the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (\tab S5).
-Community-weighted means of frond length, frond width, stipe length, and rhizome diameter were correlated (Pearson's correlation coefficient > 0.9), so of these we only analyzed stipe length.
+Community-weighted means of frond length, frond width, stipe length, and rhizome diameter were correlated (Pearson's correlation coefficient > 0.9), so of these we only show stipe length.
 
-Distribution of community-weighted mean values (CWMs) differed between epiphytic and terrestrial communities overall (pinna number, stipe length) or co-varied with elevation (dissection, SLA) (\fig \@ref(fig:cwm-div)).
+Community-weighted mean values (CWMs) were smaller for epiphytic than terrestial communities for pinna number, and size-related traits (frond length, frond width, stipe length, and rhizome diameter) (\tval-test, all \pval < 0.05; \fig \@ref(fig:cwm-div)).
+Standard deviation of community-weighted trait values was larger for size-related traits in terrestrial communities, but smaller for other traits such as frond dissection and pinna number.
 Specific leaf area decreased with elevation, and was lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval = `r get_mod_result(model_summ_print, "sla", "p")`).
 Degree of lamina dissection decreased with elevation in epiphytes, but slightly increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval = `r get_mod_result(model_summ_print, "dissection", "p")`).
-Stipe length and pinna number showed no significant variation with elevation, but were much lower overall for epiphytes relative to terrestrial communities (\tval-test, both \pval = `r get_mod_result(t_test_print, "stipe", "p")`).
+Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (\fig \@ref(fig:cwm-div)).
 
@@
 Species richness increased  with elevation for both epiphytes and terrestrial communities (\fig \@ref(fig:comm-div)).
-Trait richness (FRic) was lower in epiphytic vs.
-terrestrial communities (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "FRic", "p")`), whereas functional diversity (FDiv) was lower in terrestrial communities (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "FDiv", "p")`), and functional evenness (FEve) did not differ between the two.
-Both epiphytic and terrestrial communities showed a trend of increasing trait richness along the environmental gradient, with greater FRic at high elevations (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "FRic", "r")`, \pval = `r get_mod_result(model_summ_print, "FRic", "p")`).
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "mntd.obs.z", "p")`) but not MPD (\pval = `r get_mod_result(t_test_print, "mpd.obs.z", "p")`; \fig \@ref(fig:comm-div)).
+
+Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "mpd.obs.z.func", "p")` and `r get_mod_result(t_test_print, "mntd.obs.z.func", "p")`, respectively).
+Functional diversity increased increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but hot MNTD~func~.
+
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "mntd.obs.z", "p")`) but not MPD~phy~ (\pval = `r get_mod_result(t_test_print, "mpd.obs.z", "p")`; \fig \@ref(fig:comm-div)).
 No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (\fig \@ref(fig:comm-div)).
@@
 Here, we show for the first time using a broadly sampled phylogenetic dataset that morphological traits related to epiphytic growth are strongly evolutionarily conserved in ferns.
-The canalized morphology of epiphytic ferns suggests that this habitat strongly filters for a specialized growth form.
-We find that communities of epiphytic ferns are more clustered functionally and phylogenetically relative to terrestrial species, again supporting the role of environmental filtering in determining epiphytic fern community structure.
+We find that epiphtyic ferns are constrained by size, having consistently smaller fronds and rhizomes than terrerstial species.
+However, contrary to our hypothesis, epiphytic fern communities appear to be more diverse than terrestrial communities.
+This suggests that environmental filtering plays a strong role in limiting plant size, while allowing for diversification across other trait axes in epiphytes.
 Our results are similar between gametophyte and sporophyte traits, suggesting that morphology of each life stage is important for epiphytic growth.
@@
 
-The sporophyte traits that differed most strongly between epiphytic and terrestrial taxa were related to plant size, including stipe and frond length, frond width, and rhizome diameter (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
+Epiphytes were markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
 Stipe length may be a critical trait controlling whole leaf hydraulic conductance; since ferns have resistive stipes in general, species growing in water-limited habitats are expected to decrease stipe length to minimize resistance [@WatkinsJr.2010].
@@
 
+In contrast to their clearly smaller size, epiphytes were remarkably diverse in other traits.
+Epiphytes tended to have less dissected fronds with fewer pinna than terrestrial species (\tab \@ref(tab:pics)), but the range of values for these traits was quite wide at the community level (\fig \@ref(fig:cwm-div).
 Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics).
@@
 We hypothesized that epiphytes would have lower SLA, given their more extreme growing conditions.
-Indeed, the species with the lowest SLA values were epiphytes (\eg, *Davallia anderssonii* Mett.
-ex Kuhn, \SI{`r smallest_sla_1`} {\square\meter\per\kilo\gram}; *Ctenopterella blechnoides* (Grev.) Parris, \SI{`r smallest_sla_2`} {\square\meter\per\kilo\gram}), but this relationship was not significant across the dataset (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
 The tradeoffs posited in the leaf economics spectrum have been most thoroughly documented in global studies including the widest possible range of taxa and environments, and do not always hold up in local studies [@Petter2016; @Messier2017].
 For example, @Zhu2016 did not find a significant relationship between SLA and leaf life span in a survey of 16 ferns across a disturbance (light) gradient in China.
-It is possible that other small-scale effects (\eg, microclimate) obscure this pattern in local studies such as ours and that of @Zhu2016.
-This is supported by the strong, parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (\fig \@ref(fig:cwm-div)).
+Specific leaf area was the only response variable for which we detected spatial autocorrelation in the residuals (REF SUPP DATA).
+This, and the strong, parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (\fig \@ref(fig:cwm-div)), indicates that another unsampled environmental variable may be driving differences in SLA along the gradient.
 
@@
 
-Results for functional trait richness (FRic) supported our hypotheses: we observed much lower FRic in epiphytic relative to terrestrial communities overall, and a significant positive correlation of FRic with elevation (\fig \@ref(fig:comm-div)).
-However, results for the other two measures of functional diversity, evenness (FEve) and divergence (FDiv) did not match our predictions.
-Rather, FDiv was lower in terrestrial communities overall but did not show any trend along the gradient, while FEve did not show any significant differences at all (\fig \@ref(fig:comm-div)).
-Functional divergence is similar to richness, but takes into account abundances; so the FDiv results may be driven by a small number of abundant species.
-The lack of a clear pattern in FEve suggests that within each trait pool, there is no significant difference in trait spacing between epiphytes and terrestrial species.
-This could be maintained by niche partitioning within epiphytic and terrestrial habitats.
-Epiphytic habitats are highly heterogeneous, and microclimatic conditions vary with height along the host and distance from the trunk [@Hietz1995].
-The vertical stratification of epiphytic species into different zones along height gradients in host trees is well documented [@Zotz2007].
-In a study comparing terrestrial and epiphytic ferns along an elevational gradient from \SIrange{30}{3000}{\meter} in Costa Rica, @WatkinsJr.2006a found that epiphytes tended to sort into low-trunk and canopy zones, with only 18\% found in both, and no species that overlapped between canopy and terrestrial habitats.
-It is possible that epiphytic growth per se limits overall morphological variation in epiphytic ferns, but that species with different traits are then sorted into different niches.
-
-In addition to their decreased functional diversity, we also found that epiphytic communities are more phylogenetically clustered relative to terrestrial communities when measured with MNTD, but not MPD (\fig \@ref(fig:comm-div)).
-This makes sense given that most epiphytes have radiated within five major clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely-related species from two or more epiphytic clades.
+We obtained partial support for our hypothesis that epiphytic fern communities would be more phylogenetically clustered than terrestrial ones; this was observed for MNTD, but not MPD (\fig \@ref(fig:comm-div)).
+This makes sense given that most epiphytes have radiated within five major clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely-related species from two or more epiphytic clades, but has a similar overall level of phylogentic diversity as terrestrial communities.
 However, we did not detect significant trends in phylogenetic community structure of epiphytic and terrestrial communities along the elevational gradient.
@@
 
+Results for functional community diversity contradicted our hypotheses; epiphytic communities were more functionally diverse than terrestrial as measured by both MPD and MNTD (\fig \@ref(fig:comm-div)).
+However, this result is highly dependent on the traits used to measure functional diversity; when the same analyses were done using only size-related traits (scaled, untransformed stipe length, frond length and width, and rhizome diameter), the exact opposite result was observed (i.e., epiphytic communities are mores clustered; Supp.
+Info).
+Futhermore, terrestrial species show a much wider range of values for size traits (\fig \@ref(fig:cwm-div)).
+This suggests that epiphytic ferns are subject to extreme environmental filtering for size-related traits, while simultaneously diversifying along other trait axes.
+This could be maintained by niche partitioning within epiphytic habitats.
+Epiphytic habitats are highly heterogeneous, and microclimatic conditions vary with height along the host and distance from the trunk [@Hietz1995].
+The vertical stratification of epiphytic species into different zones along height gradients in host trees is well documented [@Zotz2007].
+In a study comparing terrestrial and epiphytic ferns along an elevational gradient from \SIrange{30}{3000}{\meter} in Costa Rica, @WatkinsJr.2006a found that epiphytes tended to sort into low-trunk and canopy zones, with only 18\% found in both.
+It seems that the canopy habitat limits overall size in epiphytic ferns, but that species with different traits then sort into different niches.
+The goal of our study was to quantify differences between epiphytic and terrestrial fern communities; future research should focus on bringing comparative phylogenetic and community ecological methods to bear on the large body of work characterizing differences within epiphytic microhabitats.
+
 ### Conclusion and Future Directions
@@
 We have demonstrated that growth habit, and several traits associated with it, show significant phylogenetic signal in ferns.
-Furthermore, we find that epiphytes have restricted morphologies both on a per-species basis and at the community level, and that epiphytic communities are phylogenetically clustered relative to terrestrial ones.
+Furthermore, we find that epiphytes are constrained by size while being remarkably diverse in other traits.
 Finally, we observed drier conditions in the canopy relative to terrestrial habitats.
-Taken together, our results support a scenario of environmental filtering acting on conserved traits to limit functional and phylogenetic diversity in epiphytic ferns.
+Taken together, our results support a scenario of environmental filtering acting to limit size and phylogenetic diversity in epiphytic ferns, while simulataneously enabling diversification along other trait axes.
 The generality of this pattern could be further tested by investigating the functional traits of other groups with large epiphytic radiations such as orchids and bromeliads.
@@
 
-```{r cwm-div, fig.height = 8, fig.width = 4, fig.cap = "Community-weighted means."}
+```{r cwm-div, fig.height = 4, fig.width = 8, fig.cap = "Community-weighted means."}
 combined_cwm_plots

### 2019-08-06 code/functions.R
Fix typo

@@ -3088,7 +3088,7 @@ test_corr_evo_range_gemmae <- function(community_matrix_path, phy, traits, moore
   
   # Add growth habit
   range_comp <- inner_join(
-    fern_traits, range_comp
+    traits, range_comp
   ) %>%
     select(species, gameto_beyond_sporo, gemmae) %>%
     mutate(gemmae = case_when(

### 2019-08-06 si/SI.rmd
Add Fig S1, correlation plot

@@
 loadd(best_fit_div_models)
+loadd(func_comm_struc_combined)
+
 ```
@@
 ```
+
+**Fig.
+S1** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input: **a**, standard effect size of mean phylogenetic distance (MPD~func~) **b**, standard effect size of mean nearest taxon distance (MNTD~func~).
+For details on calculations of MPD~func~ and MNTD~func~, see Methods.
+Correlations (values in each tile) calculated with Pearson's correlation coefficient using the R package "corrr".
+All trait sets were scaled to standard deviation of each trait.
+The following trait sets were used: "all" (all traits), "sporo" (sporophyte traits; frond length and width, stipe length, rhizome width, pinna number, SLA, degree of lamina dissection), "gameto" (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), "size" (sporophyte size-related traits; frond length and width, stipe length, rhizome width), "other" (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), "reduced" (all traits but excluding frond length and width, which were highly correlated with stipe length and rhizome diameter).
+For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), both log-transformed and untransformed sets were included.
+The log-transformed "reduced" dataset was used for the main results.
+
+```{r fig-s1}
+library(corrr)
+
+# Tweak names for easier viewing
+func_comm_struc_combined_renamed <-
+func_comm_struc_combined %>%
+  mutate(dataset = str_remove_all(dataset, "func_comm_struc_test_fern_traits_")) %>%
+  mutate(dataset = str_remove_all(dataset, "\\.pool")) %>%
+  mutate(dataset = str_replace_all(dataset, "all_reduced", "reduced")) %>%
+  mutate(dataset = str_remove_all(dataset, "scaled_")) %>%
+  mutate(dataset = str_replace_all(dataset, "log_", "log-trans_"))
+
+# Don't need log-transformed sets that don't include log-transformed traits
+func_comm_struc_combined_renamed <- 
+  func_comm_struc_combined_renamed %>%
+  filter(
+    !dataset %in% c("log-trans_gameto", "log-trans_other")
+  )
+
+# Convert to wide format, run corrr: MPD
+mpd_corr <-
+  func_comm_struc_combined_renamed %>%
+  select(dataset, site, habit, mpd.obs.z) %>%
+  spread(dataset, mpd.obs.z) %>%
+  select(-site, -habit) %>%
+  correlate %>% 
+  rearrange(absolute = FALSE) # arrange so most correlated variables are closest
+
+# Modified from https://github.com/tidymodels/corrr/issues/35
+order <- mpd_corr %>% 
+  select(-rowname) %>% 
+  colnames()
+
+mpd_corr_plot <-
+mpd_corr %>% 
+  stretch(na.rm = TRUE) %>%
+  mutate_at(c("x", "y"), forcats::fct_relevel, ... = order) %>% 
+  ggplot(aes(x, y, fill = r)) +
+  geom_tile() +
+  geom_text(aes(label = as.character(fashion(r))), color = "white", size = 3) +
+  scale_fill_scico(palette = "roma", breaks = c(-1, 1)) +
+  labs(x = NULL, y = NULL, title = "a") +
+  theme_minimal() +
+  theme(
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_blank(),
+    title = element_text(face = "bold"),
+    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
+  )
+
+# Convert to wide format, run corrr: MNTD
+mntd_corr <-
+  func_comm_struc_combined_renamed %>%
+  select(dataset, site, habit, mntd.obs.z) %>%
+  spread(dataset, mntd.obs.z) %>%
+  select(-site, -habit) %>%
+  correlate %>% 
+  rearrange(absolute = FALSE) # arrange so most correlated variables are closest
+
+# Modified from https://github.com/tidymodels/corrr/issues/35
+# order <- mntd_corr %>% 
+#   select(-rowname) %>% 
+#   colnames() %>% 
+#   rev()
+
+mntd_corr_plot <-
+  mntd_corr %>% 
+  stretch(na.rm = TRUE) %>%
+  mutate_at(c("x", "y"), forcats::fct_relevel, ... = order) %>% 
+  ggplot(aes(x, y, fill = r)) +
+  geom_tile() +
+  geom_text(aes(label = as.character(fashion(r))), color = "white", size = 3) +
+  scale_fill_scico(palette = "roma", breaks = c(-1, 1)) +
+  labs(x = NULL, y = NULL, title = "b") +
+  theme_minimal() +
+  theme(
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_blank(),
+    title = element_text(face = "bold"),
+    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
+  )
+
+fig_s1 <- mpd_corr_plot + mntd_corr_plot + plot_layout(nrow = 2, ncol = 1)
+
+ggsave(plot = fig_s1, filename = here("si/Fig_S1.pdf"), height = 8, width = 4, units = "in")
+```

### 2019-08-06 si/SI.rmd
For func div SI, use sporo size or other, but not both

@@
 All trait sets were scaled to standard deviation of each trait.
-The following trait sets were used: "all" (all traits), "sporo" (sporophyte traits; frond length and width, stipe length, rhizome width, pinna number, SLA, degree of lamina dissection), "gameto" (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), "size" (sporophyte size-related traits; frond length and width, stipe length, rhizome width), "other" (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), "reduced" (all traits but excluding frond length and width, which were highly correlated with stipe length and rhizome diameter).
+The following trait sets were used: "all" (all traits), "size" (sporophyte traits related to size; frond length and width, stipe length, rhizome width, pinna number, SLA, degree of lamina dissection), "other" (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), "gameto" (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), "reduced" (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter).
 For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), both log-transformed and untransformed sets were included.
@@
 
+# Make plot 
 # Modified from https://github.com/tidymodels/corrr/issues/35
@@
   correlate %>% 
-  rearrange(absolute = FALSE) # arrange so most correlated variables are closest
-
-# Modified from https://github.com/tidymodels/corrr/issues/35
-# order <- mntd_corr %>% 
-#   select(-rowname) %>% 
-#   colnames() %>% 
-#   rev()
+  rearrange(absolute = FALSE)
 
@@
   stretch(na.rm = TRUE) %>%
-  mutate_at(c("x", "y"), forcats::fct_relevel, ... = order) %>% 
+  mutate_at(c("x", "y"), forcats::fct_relevel, ... = order) %>% # Use same order as MPD
   ggplot(aes(x, y, fill = r)) +

### 2019-08-06 code/plan.R

@@ -3,7 +3,7 @@ plan <- drake_plan(
   
   # Data processing ----
   
-  ### Download pre-processed data ###
+  ### Pre-processed data ###
   # Download and unzip data from Nitta et al. 2017 Ecol. Monographs from Dryad
   nitta_2017_data = download_and_unzip_nitta_2017(
     dl_path = "data/nitta_2017_data_and_scripts.zip",
@@ -114,7 +114,7 @@ plan <- drake_plan(
   ### PPGI taxonomy ###
   ppgi = read_csv(file_in("data/ppgi_taxonomy.csv")),
   
-  ###  Format trait data for SI ###
+  ### Format trait data for SI ###
   # Include SD and n for all quantitative (continuous) data
   trait_data_for_si = process_trait_data_for_si(
     sla_raw, morph_cont_raw, morph_qual_raw
@@ -158,8 +158,7 @@ plan <- drake_plan(
   # - Make full set of climate models including
   # each climate var by elevation, growth habit, and
   # their interaction, then choose the best model for each.
-  # Exclude single outlier site, but use untransformed
-  # data so the fits can be plotted.
+  # Exclude single outlier site.
   climate_models = choose_habit_elevation_models(
     data = climate_trans_no_outlier,
     resp_vars = climate_vars),
@@ -260,16 +259,11 @@ plan <- drake_plan(
           "sla", "pinna", "dissection",
           "morphotype", "glands", "hairs", "gemmae"),
   
-  # - traits related to sporophyte size
+  # - sporophyte traits related to size
   size = c("stipe", "length", "width", "rhizome"),
   
-  # - traits other than sporophyte size
-  other = c("sla", "pinna", "dissection",
-            "morphotype", "glands", "hairs", "gemmae"),
-  
-  # - sporophyte traits
-  sporo = c("stipe", "length", "width", "rhizome",
-            "sla", "pinna", "dissection"),
+  # - sporophyte traits other than size
+  other = c("sla", "pinna", "dissection"),
   
   # - gametophyte traits
   gameto = c("morphotype", "glands", "hairs", "gemmae"),
@@ -292,7 +286,7 @@ plan <- drake_plan(
     ),
     transform = cross(
       trait_data = c(fern_traits_scaled, fern_traits_log_scaled),
-      traits_select_list = c(all, size, other, sporo, gameto, reduced),
[…]

### 2019-08-06 ms/manuscript.Rmd
Update MS, SI

@@
   chunk_output_type: console
+nocite: | 
+  @Vogel1968, @Wollenweber2000
 ---
@@
 loadd(correlated_evo_test)
+loadd(correlated_evo_test_gemmae)
 loadd(daily_mean_climate)
@@
 loadd(div_t_test_results)
-loadd(div_metrics_select)
+loadd(div_metrics_all)
 loadd(fern_traits)
@@
 loadd(pca_results)
+loadd(phy)
 loadd(phylosig_binary_traits)
@@
 \clearpage
-
+<!-- extra text-->
 ## Introduction
@@
 
-Growth habit was coded as a binary trait (epiphytic or terrestrial) based on field observations.
+Growth habit was coded as a binary trait (epiphytic or terrestrial) based on our field observations.
 Some species do not fall clearly into either category (\eg, hemi-epiphytes; [@Benzing1990; @Dubuisson2003; @Nitta2009]).
@@
 
-We used a multivariate approach to investigate differences in traits between epiphytes and terrestrial species.
-We used only quantitative traits, and included only species that had no missing data for any traits (*\n* = `r quant_traits_nsp` \si{\species}).
-We first log-transformed traits as necessary to attain normality, then scaled each trait by its range.
-We subjected the transformed, scaled traits to a Principal Components Analysis (PCA) to reduce dimensionality.
-We conducted standard and phylogenetically corrected PCA, using the \function{PCA} and \function{phyl.pca} functions in the \package{FactoMineR} [@Le2008] and \package{phytools} R packages, respectively.
-
 We detected significant phylogenetic signal in most traits, so we tested for differences between traits of epiphytic vs.
@@
 
+We investigated differences in multivariate trait space between epiphytes and terrestrial species using Principal Components Analysis (PCA).
+We used only quantitative traits, and included only species that had no missing data for any traits (*n* = `r quant_traits_nsp` \si{\species}).
+We first log-transformed traits as necessary to reduce skew, then scaled each trait by its standard deviation.
+We conducted standard and phylogenetically corrected PCA, using the \function{PCA} and \function{phyl.pca} functions in the \package{FactoMineR} [@Le2008] and \package{phytools} [@Revell2012] R packages, respectively.
+
 We used distance-based methods to analyze phylogenetic and functional diversity.
@@
 For functional diversity, we used Gower's dissimilarity coeffecient [@Gower1971] because it can use continuous, binary, and qualitative traits, and accepts missing data.
-We log-transformed traits as needed to account for skew, excluded frond width and length (both frond size traits highly correlated with stipe length), and weighted sporophyte and gametophyte traits equally before calculating the distance matrix using the \function{gowdis} function in the R pacakge \package{FD}.
+We log-transformed traits as needed to account for skew, scaled each trait by its standard deviation, and excluded frond width and length (both frond size traits highly correlated with stipe length) before calculating the distance matrix using the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
+To test for the effect of different trait selection schemes, we compared the results of running the same analyses for various sets of traits with and without log-transformation (Supporting Information).
 
@@
 We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate phylogenetic and functional diversity metrics.
+Hereafter, "MPD~phy~" and "MNTD~phy", and "MPD~func~" and "MNTD~func~", will be used to refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.
 
@@
 For each diversity metric, we built a set of models (Gaussian family, with site as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
-Our dataset included `r nrow(div_metrics_select) - nrow(climate_select)` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
+Our dataset included `r nrow(div_metrics_all) - nrow(climate_select)` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
 Since we found that all climatic variables are related to elevation (see Results), we also constructed linear models for each diversity metric using the same model selection procedure as in the climate analysis.
@@
 
-Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife.
+Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife. Data have been deposited in Dryad (\*\*\* ADD DOI \*\*\*).
 
@@
 The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
-The trait loadings were similar between the two analyses, except that most switched axes; for example, SLA was loaded on PC1 in the standard analysis but PC2 in the phylogenetic analysis.
+In the standard PCA, size-related traits tended to load on PC1, with SLA, pinna, and dissection on PC2.
+In the phylogenetic PCA, SLA was strongly loaded on PC1, while dissection, pinna, and other size-related traits were mostly loaded on PC2.
 
@@
 mean_rich <-
-  div_metrics_select %>%
+  div_metrics_all %>%
   group_by(habit) %>%
@@
 Species richness increased  with elevation for both epiphytes and terrestrial communities (\fig \@ref(fig:comm-div)).
-
 Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "mpd.obs.z.func", "p")` and `r get_mod_result(t_test_print, "mntd.obs.z.func", "p")`, respectively).
-Functional diversity increased increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but hot MNTD~func~.
-
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "mntd.obs.z", "p")`) but not MPD~phy~ (\pval = `r get_mod_result(t_test_print, "mpd.obs.z", "p")`; \fig \@ref(fig:comm-div)).
+Functional diversity increased increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval = `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; \fig \@ref(fig:comm-div)).
 No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (\fig \@ref(fig:comm-div)).
@@
 Stipe length may be a critical trait controlling whole leaf hydraulic conductance; since ferns have resistive stipes in general, species growing in water-limited habitats are expected to decrease stipe length to minimize resistance [@WatkinsJr.2010].
-Our results support previous non-phylogenetically controlled studies that found a prevalence of short stipes in epiphytic ferns [@WatkinsJr.2010; Creese2011].
-However, given that these were all correlated with each other (Pearson's correlation coefficient > 0.9), and similar traits have been shown to scale together previously [@Arcand2008; Creese2011], careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
+Our results support previous non-phylogenetically controlled studies that found a prevalence of short stipes in epiphytic ferns [@WatkinsJr.2010; @Creese2011].
+However, given that these were all correlated with each other (Pearson's correlation coefficient > 0.9), and similar traits have been shown to scale together previously [@Arcand2008; @Creese2011], careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
 
@@
 In contrast to their clearly smaller size, epiphytes were remarkably diverse in other traits.
-Epiphytes tended to have less dissected fronds with fewer pinna than terrestrial species (\tab \@ref(tab:pics)), but the range of values for these traits was quite wide at the community level (\fig \@ref(fig:cwm-div).
+Epiphytes tended to have less dissected fronds with fewer pinna than terrestrial species (\tab \@ref(tab:pics)), but the range of values for these traits was quite wide at the community level (\fig \@ref(fig:cwm-div)).
 Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics).
@@
 It is possible that some, but not all, epiphytic ferns rely on such strategies, and that they are not under similarly strong selection as other traits such as stipe length and gametophyte morphotype.
-Interestingly, when we classified 'widespread' species as those with gametophytes occurring more than \SI{200}{\meter} above or below the minimum or maximum elevational range of conspecific sporophytes (ADD TO SI METHODS), we found that gemmae production is correlated with widespread growth (log likelihood dependent model = \num{-54.5}, log likelihood independent model = \num{-63.6}, $\pval = 0.001$, Pagel's test of correlated evolution).
-Thus, it seems that gemmae do function in long-term gametophyte persistence, but may not be required for epiphytic growth.
+Interestingly, when we classified 'widespread' species as those with gametophytes occurring above or below the minimum or maximum elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with widespread growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution).
+Thus, it seems that gemmae may function in long-term gametophyte persistence, but are not required for epiphytic growth.
 
@@
 Results for functional community diversity contradicted our hypotheses; epiphytic communities were more functionally diverse than terrestrial as measured by both MPD and MNTD (\fig \@ref(fig:comm-div)).
-However, this result is highly dependent on the traits used to measure functional diversity; when the same analyses were done using only size-related traits (scaled, untransformed stipe length, frond length and width, and rhizome diameter), the exact opposite result was observed (i.e., epiphytic communities are mores clustered; Supp.
-Info).
-Futhermore, terrestrial species show a much wider range of values for size traits (\fig \@ref(fig:cwm-div)).
+However, this result is highly dependent on the traits used to measure functional diversity; when the same analyses were done using only size-related traits (scaled, untransformed stipe length, frond length and width, and rhizome diameter), the opposite result was observed (i.e., epiphytic communities are more clustered; \fig S1).
+Futhermore, terrestrial species show a much wider range of values for size traits (\fig \@ref(fig:cwm-div)), and neither functional group appears to be restricted to a smaller area in trait space (\fig \@ref(fig:pca)).
 This suggests that epiphytic ferns are subject to extreme environmental filtering for size-related traits, while simultaneously diversifying along other trait axes.
@@
 
-**\fig \@ref(fig:comm-div)**: Functional and phylogenetic diversity of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
+**\fig \@ref(fig:comm-div)**: Functional and phylogenetic diversity of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
 Epiphytic communities in green, terrestrial communities in brown.
-Response variable abbreviations as follows: MPD, standard effect size (SES) of mean phylogenetic distance; MNTD, SES of mean nearest taxon distance; FRic, functional richness; FEve, functional evenness; FDiv, functional diversity.
-Scatterplots show response variables plotted against either minimum relative humidity ($\n = 13$ communities each for epiphytes and terrestrial species) or elevation ($\n = 17$ communities each).
+Response variable abbreviations as follows: MPD~phy~, standard effect size (SES) of mean phylogenetic distance; MNTD~phy~, SES of mean nearest taxon distance; MPD~func~, SES of mean functional distance; MNTD~func~, SES of mean nearest functional distance.
 Trendlines indicate significant relationships determined by a linear model at $\pval<0.05$.
-Boxplots show median values (bold lines) of response variables by growth habit (epiphytic or terrestrial) across all communities (epiphytic communities, $\n = 17$; terrestrial communities, $\n = 17$).
+Boxplots show median values (bold lines) of response variables by growth habit (epiphytic or terrestrial).
 Hinges and whiskers as in \fig \@ref(fig:cwm-div).
 Asterisks indicate statistical significance of one-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
-
-\clearpage
-
-## Tables
 
@@
 
-# super_pattern <- "([[:digit:]].*$)"
-# super_text <- str_extract(traits_used$ecological_significance, super_pattern) 
-# super_text <- paste0("\\\\textsuperscript{", super_text, "}")
-
 traits_used %>% 
-  # mutate(ecological_significance = str_replace_all(ecological_significance, super_pattern, super_text)) %>%
+  rename(`Data type` = data_type, `Functional significance` = ecological_significance) %>%
   column_to_rownames("trait") %>%
-  knitr::kable(caption = "Traits used", format = "latex", booktabs = TRUE, escape = TRUE) %>%
+  knitr::kable(caption = "Traits used in this study", format = "latex", booktabs = TRUE, escape = TRUE) %>%
   kable_styling(full_width = FALSE) %>%
@@
   pack_rows("Gametophyte traits", 8, 11) %>%
-  column_spec(3, width = "15em")
+  column_spec(3, width = "15em") %>%
+  footnote(general = "References: 1, Watkins \\\\textit{et al.} (2010); 2, Vogel (1968); 3, Creese \\\\textit{et al.} (2011); 4, Kluge \\\\& Kessler (2007); 5, Wright \\\\textit{et al.} (2004); 6, Watkins \\\\textit{et al.} (2007b); 7, Pittermann \\\\textit{et al.} (2013); 8, Farrar \\\\textit{et al.} (2008); 9, Wollenweber \\\\& Schneider (2000).
+Noncordate gametophyte morphotypes include ribbon, strap and filamentous.", general_title = "", threeparttable = TRUE, escape = FALSE)
 
@@
 
-```{r heatmap, fig.height = 8, fig.width = 4, fig.cap = "Variable importance scores."}
+```{r heatmap, fig.height = 4, fig.width = 6, fig.cap = "Variable importance scores."}
 importance_heatmap

### 2019-08-06 si/SI.rmd
Update MS, SI

@@
 Models within delta AICc 3 of the best-fitting model shown.
-Abbreviations: FDiv, functional diversity; FEve, functional evenness; FRic, functional richness; MPD, mean phylogenetic distance; MNTD, mean nearest taxon distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion.
-Values of MPD and MNTD are the standard effect sizes calculated from comparison of observed values with null distribution (see Methods).
+Abbreviations: MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion.
 Trait names refer to community-weighted mean values (see Methods).
+Frond length and width were highly correlated with rhizome diameter and not shown.
+For details on calculations of MPD and MNTD, see Methods.
 
@@
 best_fit_div_models %>%
+  # Drop frond length and width (highly correlated with rhizome dia.)
+  filter(resp_var %in% c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
+                   "mpd.obs.z.func", "mntd.obs.z.func",
+                   "dissection", "sla", "stipe", "rhizome", "pinna")) %>%
   mutate(resp_var = factor(
         resp_var, 
-        levels = c("ntaxa", "mpd.obs.z", "mntd.obs.z", 
-                   "FDiv", "FEve", "FRic",
-                   "dissection", "sla", "stipe", "pinna"))) %>%
+        levels = c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
+                   "mpd.obs.z.func", "mntd.obs.z.func",
+                   "dissection", "sla", "stipe", "rhizome", "pinna"))) %>%
   mutate(
@@
         resp_var, 
-        new_levels = c("Richness", "MPD", "MNTD", 
-                       "FDiv", "FEve", "FRic",
-                       "Dissection", "SLA", "Stipe length", "Pinna no.")) 
+        new_levels = c("Richness", "MPDphy", "MNTDphy", 
+                       "MPDfunc", "MNTDfunc",
+                       "Dissection", "SLA", "Stipe length", "Rhizome diam.", "Pinna no.")) 
   ) %>%
@@
       str_replace_all("\\+", " \\+ ") %>%
-      str_replace_all("\\.by\\.", " × ") %>%
+      str_replace_all("\\.by\\.", " x ") %>%
       str_replace_all("habit", "Growth habit") %>%

### 2019-08-06 ms/manuscript.Rmd
Update conclusion to reflect Testo et al 2018

@@
 
-Here, we show for the first time using a broadly sampled phylogenetic dataset that morphological traits related to epiphytic growth are strongly evolutionarily conserved in ferns.
+Here, we show for the first time using a broadly sampled phylogenetic dataset that multiple morphological traits related to epiphytic growth are evolutionarily conserved in ferns.
 We find that epiphtyic ferns are constrained by size, having consistently smaller fronds and rhizomes than terrerstial species.
@@
 This suggests that environmental filtering plays a strong role in limiting plant size, while allowing for diversification across other trait axes in epiphytes.
-Our results are similar between gametophyte and sporophyte traits, suggesting that morphology of each life stage is important for epiphytic growth.
 
@@
 Epiphytes were markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
+Our results at the local scale are corroborated by @Testo2018a, who also observed abrupt size reduction associated with epiphytic growth in a globally sampled fern phylogeny.
 Stipe length may be a critical trait controlling whole leaf hydraulic conductance; since ferns have resistive stipes in general, species growing in water-limited habitats are expected to decrease stipe length to minimize resistance [@WatkinsJr.2010].
@@
 Epiphytes tended to have less dissected fronds with fewer pinna than terrestrial species (\tab \@ref(tab:pics)), but the range of values for these traits was quite wide at the community level (\fig \@ref(fig:cwm-div)).
-Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics).
+Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
 Specific leaf area is an important part of the leaf economic spectrum, which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
@@
 For example, @Zhu2016 did not find a significant relationship between SLA and leaf life span in a survey of 16 ferns across a disturbance (light) gradient in China.
-Specific leaf area was the only response variable for which we detected spatial autocorrelation in the residuals (REF SUPP DATA).
+Specific leaf area was the only response variable for which we detected spatial autocorrelation in the residuals (\tab S5).
 This, and the strong, parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (\fig \@ref(fig:cwm-div)), indicates that another unsampled environmental variable may be driving differences in SLA along the gradient.
@@
 Results for functional community diversity contradicted our hypotheses; epiphytic communities were more functionally diverse than terrestrial as measured by both MPD and MNTD (\fig \@ref(fig:comm-div)).
-However, this result is highly dependent on the traits used to measure functional diversity; when the same analyses were done using only size-related traits (scaled, untransformed stipe length, frond length and width, and rhizome diameter), the opposite result was observed (i.e., epiphytic communities are more clustered; \fig S1).
+However, this result is highly dependent on the traits used to measure functional diversity; when the same analyses were done using only size-related traits (scaled, untransformed stipe length, frond length and width, and rhizome diameter), the opposite result was observed (\fig S1).
 Futhermore, terrestrial species show a much wider range of values for size traits (\fig \@ref(fig:cwm-div)), and neither functional group appears to be restricted to a smaller area in trait space (\fig \@ref(fig:pca)).
@@
 Taken together, our results support a scenario of environmental filtering acting to limit size and phylogenetic diversity in epiphytic ferns, while simulataneously enabling diversification along other trait axes.
+This is consistent with a much more broadly sampled study finding that epiphytic and terrestrial ferns do not differ in rates of body size evolution [@Testo2018a].
 The generality of this pattern could be further tested by investigating the functional traits of other groups with large epiphytic radiations such as orchids and bromeliads.
 
-Recent studies [e.g., @Sundue2015; @Testo2016a; @Lehtonen2017] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
+Recent studies [e.g., @Sundue2015; @Testo2016a; @Lehtonen2017; @Testo2018a] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
 Epiphytism itself is likely a complex trait requiring the evolution of multiple associated traits, and the timing and order of the evolution of such traits may be related to the complex diversification dynamics observed in these studies.
-Our study identifies several conserved morphological traits (\eg, stipe length, frond dissection, gametophyte morphotype) that may belong to this suite of epiphytic traits.
-We hope future studies will build on our results by analyzing these traits across much more densely sampled phylogenies [e.g., @Testo2016a] to dissect how each may be related to diversification and growth habit.
+Small size seems to be a prerequisite for epiphytic growth, as indicated both by our study and @Testo2018a.
+Here, we identify several traits in addition to plant size (\eg, pinna number, frond dissection, gametophyte morphotype) that may also belong to the suite of epiphytic traits.
+We hope future studies will build on our results by analyzing traits involved in epiphytic growth other than body size across densely sampled phylogenies to dissect how each may relate to diversification and growth habit.
 Additionally, future studies should focus on determining in more detail how niche space is partitioned within epiphytic vs.

### 2019-08-06 ms/manuscript.Rmd
Update MS
- Shorten discussion of SLA, clarify Moran's I was not significant for any metrics

@@
 In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (\fig \ref{fig:heatmap}).
+Community-weighted means of frond length, frond width, stipe length, and rhizome diameter were correlated (Pearson's correlation coefficient > 0.9), so of these we only show stipe length.
 Most of the the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (\tab S5).
-Community-weighted means of frond length, frond width, stipe length, and rhizome diameter were correlated (Pearson's correlation coefficient > 0.9), so of these we only show stipe length.
+Spatial autocorrelation was not detected for any of the diversity metrics (\tab S5).
 
@@
 We hypothesized that epiphytes would have lower SLA, given their more extreme growing conditions.
-The tradeoffs posited in the leaf economics spectrum have been most thoroughly documented in global studies including the widest possible range of taxa and environments, and do not always hold up in local studies [@Petter2016; @Messier2017].
-For example, @Zhu2016 did not find a significant relationship between SLA and leaf life span in a survey of 16 ferns across a disturbance (light) gradient in China.
-Specific leaf area was the only response variable for which we detected spatial autocorrelation in the residuals (\tab S5).
-This, and the strong, parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (\fig \@ref(fig:cwm-div)), indicates that another unsampled environmental variable may be driving differences in SLA along the gradient.
+However, the parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (\fig \@ref(fig:cwm-div)) indicates that SLA may be responding more strongly to temperature in both epiphytes and terrestrial species, while only being weakly modified by humidity.
 

### 2019-08-06 si/SI.rmd
Update MS
- Shorten discussion of SLA, clarify Moran's I was not significant for any metrics

--- 2019-08-06 si/SI.rmd
@@
 loadd(correlated_evo_test_strict)
-loadd(best_fit_div_models)
+loadd(best_fit_div_models_moran)
 loadd(func_comm_struc_combined)
@@
 Models within delta AICc 3 of the best-fitting model shown.
-Abbreviations: FDiv, functional diversity; FEve, functional evenness; FRic, functional richness; MPD, mean phylogenetic distance; MNTD, mean nearest taxon distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion.
-Values of MPD and MNTD are the standard effect sizes calculated from comparison of observed values with null distribution (see Methods).
+Abbreviations: MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, *P* value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's *I*.
 Trait names refer to community-weighted mean values (see Methods).
+Frond length and width were highly correlated with rhizome diameter and not shown.
+For details on calculations of MPD and MNTD, see Methods.
 
@@
 # Format and write out table as CSV file
-best_fit_div_models %>%
+best_fit_div_models_moran %>%
+  # Drop frond length and width (highly correlated with rhizome dia.)
+  filter(resp_var %in% c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
+                   "mpd.obs.z.func", "mntd.obs.z.func",
+                   "dissection", "sla", "stipe", "rhizome", "pinna")) %>%
   mutate(resp_var = factor(
         resp_var, 
-        levels = c("ntaxa", "mpd.obs.z", "mntd.obs.z", 
-                   "FDiv", "FEve", "FRic",
-                   "dissection", "sla", "stipe", "pinna"))) %>%
+        levels = c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
+                   "mpd.obs.z.func", "mntd.obs.z.func",
+                   "dissection", "sla", "stipe", "rhizome", "pinna"))) %>%
   mutate(
@@
     `R2` = r2,
-    delta_AICc
+    delta_AICc,
+    `Moran's I` = morans_I,
+    `P_moran` = morans_I_pval
   ) %>%

### 2019-08-06 ms/manuscript.Rmd
Add t-value to CWMs

@@
 
-**\fig \@ref(fig:cwm-div)**: Change in selected functional traits of ferns on Moorea, French Polynesia across environmental gradients and between growth habit types (epiphytic or terrestrial).
-Epiphytic values in green, terrestrial values in brown.
-Scatterplots show community-weighted mean (CWM) values for each trait plotted against either minimum relative humidity ($\n = 13$ communities each for epiphytes and terrestrial species) or elevation ($\n = 17$ communities each).
-Trendlines indicate significant relationships determined by a linear model at $\pval<0.05$.
-Boxplots show median values (bold lines) by growth habit (epiphytic or terrestrial) across all communities (epiphytic communities, $\n = 17$; terrestrial communities, $\n = 17$).
-Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 $\times$ the interquartile range.
-Asterisks indicate statistical significance of two-sided \tval-test for differences between epiphytic and terrestrial species; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
-Trends for three leaf size traits (stipe length, frond length, and frond with) were quantitatively very similar, so of these we only present stipe length.
+**\fig \@ref(fig:cwm-div)**: Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
+Epiphytic communities in green, terrestrial communities in brown.
+Error bars are community-weighted standard deviation (values below zero not shown).
+Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at $\pval<0.05$.
+\tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at $\pval<0.05$; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
 
@@
 Response variable abbreviations as follows: MPD~phy~, standard effect size (SES) of mean phylogenetic distance; MNTD~phy~, SES of mean nearest taxon distance; MPD~func~, SES of mean functional distance; MNTD~func~, SES of mean nearest functional distance.
-Trendlines indicate significant relationships determined by a linear model at $\pval<0.05$.
-Boxplots show median values (bold lines) of response variables by growth habit (epiphytic or terrestrial).
-Hinges and whiskers as in \fig \@ref(fig:cwm-div).
-Asterisks indicate statistical significance of one-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+Trendlines indicate significant relationships as determined by a linear model at $\pval<0.05$.
+Boxplots show median values (bold lines) by growth habit.
+Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 $\times$ the interquartile range.
+Asterisks indicate statistical significance of two-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
 

### 2019-08-06 code/functions.R

@@ -1226,8 +1226,8 @@ analyze_func_struc_by_habit <- function (
     dist_mat <- FD::gowdis(traits_df, w = weights) 
   } else {
     dist_mat <- FD::gowdis(traits_df) 
-    }
-
+  }
+  
   ### Run community structure analysis ###
   
   # Look at phylo structure of each plot individually, compared to null 
@@ -1984,11 +1984,11 @@ bind_data <- function (..., id_col = "dataset") {
 #'
 #' @return ggplot object
 make_climate_scatterplot <- function (yval, xval = "el", ylab = yval, xlab = "Elevation (m)",
-                              single_line = TRUE,
-                              r_upper = TRUE,
-                              data, 
-                              fits, summaries, 
-                              habit_colors, alpha_val = 0.5) {
+                                      single_line = TRUE,
+                                      r_upper = TRUE,
+                                      data, 
+                                      fits, summaries, 
+                                      habit_colors, alpha_val = 0.5) {
   
   yval_sym <- sym(yval)
   xval_sym <- sym(xval)
@@ -2108,11 +2108,11 @@ make_climate_scatterplot <- function (yval, xval = "el", ylab = yval, xlab = "El
 #'   habit_colors = habit_colors
 #' )
 make_div_scatterplot <- function (yval, xval = "el", ylab = yval, xlab = "Elevation (m)",
-                              single_line = TRUE,
-                              r_upper = TRUE,
-                              data, 
-                              fits, summaries, 
-                              habit_colors) {
+                                  single_line = TRUE,
+                                  r_upper = TRUE,
+                                  data, 
+                                  fits, summaries, 
+                                  habit_colors) {
   
   yval_sym <- sym(yval)
   xval_sym <- sym(xval)
@@ -2436,26 +2436,42 @@ make_cwm_scatterplot <- function (yval, xval = "el", ylab = yval, xlab = "Elevat
                                   r_upper = TRUE,
                                   data, 
                                   fits, summaries, 
+                                  t_test_results,
[…]

### 2019-08-06 code/plan.R

@@ -444,10 +444,15 @@ plan <- drake_plan(
       data = cwm_long,
       fits = div_el_model_fits,
       summaries = div_el_model_summaries,
+      t_test_results = div_t_test_results,
       yval = .x,
       ylab = .y,
       habit_colors = habit_colors)
   ),
+  
+  # Combine community-weighted means scatterplots into final figure.
+  combined_cwm_plots = combine_cwm_plots(
+    scatterplots = cwm_scatterplots),
   
   # Make community diversity scatterplots
   div_scatterplots = map(
@@ -477,11 +482,6 @@ plan <- drake_plan(
   combined_comm_div_plots = combine_comm_div_plots(
     scatterplots = div_scatterplots,
     boxplots = div_boxplots),
-
-  # Combine community-weighted means scatterplots
-  # and boxplots into final figure.
-  combined_cwm_plots = combine_cwm_plots(
-    scatterplots = cwm_scatterplots),
 
   # Make heatmap of importance scores.
   importance_heatmap = make_heatmap(important_div_vars),

### 2019-08-06 ms/manuscript.Rmd
Use lajolla palette

@@
 **\fig \@ref(fig:tree)**: Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
-Relative value of quantitative (sporophyte) traits in heat gradient: low values are lighter, high values are darker.
+Relative value of quantitative (sporophyte) traits shown by heat gradient: low values are lighter, high values are darker.
 States of qualitative (gametophyte) states and growth habit indicated by colors in key.
 Grey indicates missing data or non-applicable trait states.
-Trends for three leaf size traits (stipe length, frond length, and frond with) were quantitatively very similar, so of these, we only present stipe length.
+Values for three leaf size traits (stipe length, frond length, and frond with) were correlated, so of these, we only present stipe length.
 Species missing data for six or more traits not shown.
@@
 Major Cretaceous epiphytic radiations \sensu @Schuettpelz2009 labeled on tree in green: H, hymenophylloids; T, trichomanoids, V, vittarioids; A, asplenioids; E, elaphoglossoids; P, polygrammoids (trichomanoids diversified prior to the Cretaceous but are included as an important extant epiphytic clade).
+Units on timescale for tree in millions of years.
 
@@
   mutate(lambda_pval = format_pval(lambda_pval)) %>%
+  mutate(trait = factor(
+    trait, 
+    levels = c("stipe", "length", "width", "rhizome", 
+               "sla", "pinna", "dissection"))) %>%
+  mutate(trait = lvls_revalue(
+    trait,
+    new_levels = c("Stipe length", "Frond length", "Frond width", "Rhizome diam.", 
+                   "SLA", "Pinna no.", "Frond dissection"))) %>%
+  arrange(trait) %>%
+  rename(K = kval, `P(K)` = k_pval, `P(lambda)` = lambda_pval) %>%
   column_to_rownames("trait") %>%
-  kable(caption = "Phylogenetic signal in continuous traits", format = "latex", booktabs = TRUE)
+  kable(caption = "Phylogenetic signal in continuous traits", format = "latex", booktabs = TRUE) %>%
+  footnote(general = "K, Blomberg's \\\\textit{K} (Blomberg \\\\textit{et al.}, 2003); lambda, Pagel's lambda (Pagel, 1999).
+Values of lambda or \\\\textit{K} greater than 1 indicate traits evolving according to Brownian Motion (BM); values of lambda or K near 0 indicate random distribution of traits.", general_title = "", threeparttable = TRUE, escape = FALSE)
 ```
@@
   select(trait, `Number of presences` = num_present, `Number of absences` = num_absent,
-         P_rnd = prob_random, P_bm = prob_brownian) %>%
+         D, `P(rnd)` = prob_random, `P(bm)` = prob_brownian) %>%
 column_to_rownames("trait") %>%
-  kable(caption = "Phylogenetic signal in binary traits", format = "latex", booktabs = TRUE)
+  kable(caption = "Phylogenetic signal in binary traits", format = "latex", booktabs = TRUE) %>%
+  footnote(general = "D, Fritz and Purvis’ \\\\textit{D} (Fritz \\\\& Purvis, 2010); P(rnd), probability of random distribution of traits; P(bm), probability of traits evolving under Brownian motion (BM).
+Values of \\\\textit{D} 0 or less indicate evolution according to BM; values near 1 indicate random trait distribution.", general_title = "", threeparttable = TRUE, escape = FALSE)
 ```
@@
   select(trait, n_contrasts, n_pos_con, statistic, p.value) %>%
+  mutate(trait = factor(
+    trait, 
+    levels = c("stipe", "length", "width", "rhizome", 
+               "sla", "pinna", "dissection"))) %>%
+  mutate(trait = lvls_revalue(
+    trait,
+    new_levels = c("Stipe length", "Frond length", "Frond width", "Rhizome diam.", 
+                   "SLA", "Pinna no.", "Frond dissection"))) %>%
+  arrange(trait) %>%
+  rename(t = statistic, P = p.value, `Num. constrasts` = n_contrasts, `Num. pos. constrasts` = n_pos_con) %>%
   column_to_rownames("trait") %>%
@@
   mutate_if(is.numeric, ~round_t(., 1)) %>%
+  mutate(trait = factor(
+    trait, 
+    levels = c("gemmae", "glands", "hairs","morphotype"))) %>%
+  mutate(trait = lvls_revalue(
+    trait,
+    new_levels = c("Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
+  arrange(trait) %>%
+  rename(`LogL(independent)` = logL_indep, `LogL(dependent)` = logL_dep, `Likelihood ratio` = likelihood_ratio, P = pval) %>%
   column_to_rownames("trait") %>%
   kable(caption = "Pagel's (1994) test of correlated evolution between binary traits", format = "latex", booktabs = TRUE) %>%
-  footnote(general = "Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
-terrestrial).
-Lower log likelihood for the dependent model indicates correlation with growth habit.", general_title = "", threeparttable = TRUE)
+  footnote(general = "LogL, log-likelihood.
+Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
+terrestrial growth).
+Traits with a significantly lower log likelihood for the dependent model are significantly correlated with growth habit.", general_title = "", threeparttable = TRUE)
 ```

### 2019-08-06 code/functions.R

@@ -2905,8 +2905,8 @@ plot_traits_on_tree <- function (traits, phy, ppgi) {
       position = "top",
       breaks= c("sla", "dissection", "stipe", "rhizome", "pinna"),
       labels= c("SLA", "Dissection", "Stipe length", "Rhizome diam.", "Pinna num.")) +
-    scale_fill_viridis(
-      option = "B",
+    scale_fill_scico(
+      palette = "lajolla",
       breaks = c(0, 1),
       labels = c("Low", "High")) +
     labs(

### 2019-08-06 data_raw/check_diff.R
Add data raw README and change log

@@ -1,0 +1,57 @@
+# Check for differences in morph_qual_traits_2016-09-03_mod.csv from Moorea_ferns_morph_2016-09-03.xlsx
+# I know that I copied morph_qual_traits_2016-09-03_mod.csv from the Moorea_ferns_morph_2016-09-03.xlsx
+# morphology worksheet, but didn't document further changes from there at the time.
+# This script compares them to find where I changed things.
+
+library(tidyverse)
+library(readxl)
+
+# Moorea_ferns_morph_2016-09-03.xlsx is read-only and
+# copied from /Volumes/Transcend/Projects/Moorea CommPhy/Morphology
+traits_raw <- read_excel("data_raw/Moorea_ferns_morph_2016-09-03.xlsx", sheet = "morphology", skip = 11)
+traits_new <- read_csv("data_raw/morph_qual_traits_2016-09-03_mod.csv", na = c("#DIV/0!", "#N/A", ""))
+
+# The order had been rearranged in excel, so sort both by OTU
+traits_raw <- traits_raw %>% arrange(OTU)
+traits_new <- traits_new %>% arrange(OTU)
+
+# Not the same yet, due to differences in the `Key Group` column
+all.equal(traits_raw, traits_new)
+
+# Not using that column any more anyways, so drop it
+traits_raw <- traits_raw %>% select(-`Key Group`)
+traits_new <- traits_new %>% select(-`Key Group`)
+
+# Differences due to included taxa (rows)
+all.equal(traits_raw, traits_new)
+
+# Outgroups have been dropped from the new dataset
+setdiff(traits_new$OTU, traits_raw$OTU)
+setdiff(traits_raw$OTU, traits_new$OTU)
+
+# Drop these and compare again
+traits_raw <- traits_raw %>% filter(traits_raw$OTU %in% traits_new$OTU)
+
+# Most remaining differences are due to coding of NA values
+all.equal(as.data.frame(traits_raw), as.data.frame(traits_new), check.attributes = FALSE)
+
+left_join(
+select(traits_raw, OTU, dissection_raw = dissection),
+select(traits_new, OTU, dissection_new = dissection)
+) %>%
+  mutate_at(vars(dissection_raw, dissection_new), ~replace_na(., "NA")) %>%
+  filter(dissection_raw != dissection_new)
+
+left_join(
[…]

### 2019-08-06 ms/manuscript.Rmd
Fix some spelling errors

@@
 
+<!-- Set left justification (better conversion to Word than default full justification) -->
+\raggedright
+
+<!-- Set indent to about 3 characters throughout -->
+\setlength{\parindent}{3ex}
+
 Joel H. Nitta^1,4^, James E. Watkins, Jr.^2^, Ravahere Taputuarai^3^, and Charles C. Davis^1^
@@
 \clearpage
-
-<!-- Set left justification (better conversion to Word than default full justification) -->
-\raggedright
-
-<!-- Set indent to about 3 characters throughout -->
-\setlength{\parindent}{3ex}
 
@@
 \clearpage
-<!-- extra text-->
+
 ## Introduction
@@
 Tohiea \SI{1200}{\meter} plot, which only had an epiphytic datalogger.
-Temperature and RH were logged once every \SI{15}{\minute} during two survey periods: a preliminary survey from July 18, 2012 to July 7, 2013 at six plots, and a final survey from July 7, 2013 to July 5, 2014 at 15 plots.
+Temperature and RH were logged once every \SI{30}{\minute} during two survey periods: a preliminary survey from July 18, 2012 to July 7, 2013 at six plots, and a final survey from July 7, 2013 to July 5, 2014 at 15 plots.
 Two plots were not included in the environmental survey.
@@
 However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of $\K < 1$ (\tab \@ref(tab:phylosig-cont)).
-All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmmae \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under a scenario of evolution by BM (\tab \@ref(tab:phylosig-binary)).
+All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under a scenario of evolution by BM (\tab \@ref(tab:phylosig-binary)).
 Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; \tab \@ref(tab:phylosig-binary)).
@@
 
-Community-weighted mean values (CWMs) were smaller for epiphytic than terrestial communities for pinna number, and size-related traits (frond length, frond width, stipe length, and rhizome diameter) (\tval-test, all \pval < 0.05; \fig \@ref(fig:cwm-div)).
+Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial communities for pinna number, and size-related traits (frond length, frond width, stipe length, and rhizome diameter) (\tval-test, all \pval < 0.05; \fig \@ref(fig:cwm-div)).
 Standard deviation of community-weighted trait values was larger for size-related traits in terrestrial communities, but smaller for other traits such as frond dissection and pinna number.
@@
 Here, we show for the first time using a broadly sampled phylogenetic dataset that multiple morphological traits related to epiphytic growth are evolutionarily conserved in ferns.
-We find that epiphtyic ferns are constrained by size, having consistently smaller fronds and rhizomes than terrerstial species.
+We find that epiphytic ferns are constrained by size, having consistently smaller fronds and rhizomes than terrerstial species.
 However, contrary to our hypothesis, epiphytic fern communities appear to be more diverse than terrestrial communities.

### 2019-08-08 ms/manuscript.Rmd
Add rich-text tables

@@
 ```{r traits-used}
+# Table 1: traits used in this study
+
+# Read in csv version of the table
 traits_used <- read_csv(here("data/table_1.csv"))
 
+# Write out table in rich text format for submitting
+# MS as word doc
 traits_used %>%
-  rename(`Data type` = data_type, `Functional significance` = ecological_significance) %>%
+  gt %>%
+  fmt_markdown(columns = everything()) %>%
+  gt(rowname_col = "trait") %>%
+  tab_row_group(
+    group = "Sporophyte traits",
+    rows = matches("Stipe|Frond|Rhizome|Pinna|Leaf")
+  ) %>%
+  tab_row_group(
+    group = "Gametophyte traits",
+    rows = matches("Morphotype|Gemmae|Glands|Hairs")
+  ) %>%
+  cols_label(
+    data_type = "Data type (unit)",
+    ecological_significance = "Ecological Significance"
+  ) %>% 
+  as_rtf() %>%
+  write_lines("ms/table1.rtf")
+
+# Include in PDF using kableExtra to convert to latex format
+traits_used %>% 
+  rename(`Data type (unit)` = data_type, `Functional significance` = ecological_significance) %>%
   column_to_rownames("trait") %>%
@@
 Noncordate gametophyte morphotypes include ribbon, strap and filamentous.", general_title = "", threeparttable = TRUE, escape = FALSE)
-
-# 
-# traits_used %>%
-#   gt %>%
-#   fmt_markdown(columns = everything()) %>%
-#   gt(rowname_col = "trait") %>%
-#   tab_row_group(
-#     group = "Sporophyte traits",
-#     rows = matches("Stipe|Frond|Rhizome|Pinna|Leaf")
-#   ) %>%
-#   tab_row_group(
-#     group = "Gametophyte traits",
-#     rows = matches("Morphotype|Gemmae|Glands|Hairs")
-#   ) %>%
-#   cols_label(
-#     data_type = "Data type (unit)",
-#     ecological_significance = "Ecological Significance"
-#   ) %>%
-#   knitr::kable(caption = "Traits used", booktabs = TRUE)
 ```
@@
 ```{r phylosig-cont}
+# Table 2: Phylogenetic signal in continuous traits
+
+# Format the raw results for table presentation
+phylosig_cont_traits_table <-
 phylosig_cont_traits %>%
@@
                    "SLA", "Pinna no.", "Frond dissection"))) %>%
-  arrange(trait) %>%
+  arrange(trait)
+
+# Write out table in rich text format
+phylosig_cont_traits_table %>%
+  gt(rowname_col = "trait") %>%
+  cols_label(
+    kval = md("*K*"),
+    k_pval = md("*P*(*K*)"),
+    lambda_pval = md("*P*(lambda)"),
+    lambda = "lambda"
+  )  %>% 
+  as_rtf() %>%
+  write_lines("ms/table2.rtf")
+
+# Include in PDF in latex format
+phylosig_cont_traits_table %>%
   rename(K = kval, `P(K)` = k_pval, `P(lambda)` = lambda_pval) %>%
@@
 ```{r phylosig-binary}
+# Table 3: Phylogenetic signal in binary traits
+
+# Format the raw results for table presentation
+phylosig_binary_traits_table <-
 phylosig_binary_traits %>%
@@
     new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
-  arrange(trait) %>%
+  arrange(trait)
+
+# Write out table in rich text format
+phylosig_binary_traits_table %>%
+  gt(rowname_col = "trait") %>%
+  cols_label(
+    num_present = "Number of presences",
+    num_absent = "Number of absences",
+    D = md("*D*"),
+    prob_random = md("*P*(rnd)"),
+    prob_brownian = md("*P*(bm)")
+  ) %>% 
+  as_rtf() %>%
+  write_lines("ms/table3.rtf")
+
+# Include in PDF in latex format
+phylosig_binary_traits_table %>%
   select(trait, `Number of presences` = num_present, `Number of absences` = num_absent,
@@
 ```{r pics}
+# Table 4: phylogenetically independent contrasts
+
+# Format the raw results for table presentation
+pic_table <- 
 pic_results %>%
@@
                    "SLA", "Pinna no.", "Frond dissection"))) %>%
-  arrange(trait) %>%
+  arrange(trait)
+
+# Write out table in rich text format
+pic_table %>%
+gt(rowname_col = "trait") %>%
+  cols_label(
+    n_contrasts = "Num. constrasts",
+    n_pos_con = "Num. pos. constrasts",
+    statistic = md("*t*"),
+    p.value = md("*P*"),
+  ) %>% 
+  as_rtf() %>%
+  write_lines("ms/table4.rtf")
+
+# Include in PDF in latex format
+pic_table %>%
   rename(t = statistic, P = p.value, `Num. constrasts` = n_contrasts, `Num. pos. constrasts` = n_pos_con) %>%
@@
 ```{r corr-evo}
+# Table 5: Pagel's test of correlated evolution
+
+# Format the raw results for table presentation
+correlated_evo_test_table <-
 correlated_evo_test %>%
@@
     new_levels = c("Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
-  arrange(trait) %>%
+  arrange(trait)
+
+# Write out table in rich text format
+correlated_evo_test_table %>%
+gt(rowname_col = "trait") %>%
+  cols_label(
+    logL_indep = "LogL(independent)",
+    logL_dep = "LogL(dependent)",
+    likelihood_ratio = "Likelihood ratio",
+    pval = md("*P*"),
+  ) %>% 
+  as_rtf() %>%
+  write_lines("ms/table5.rtf")
+
+# Include in PDF in latex format
+correlated_evo_test_table %>%
   rename(`LogL(independent)` = logL_indep, `LogL(dependent)` = logL_dep, `Likelihood ratio` = likelihood_ratio, P = pval) %>%

### 2019-08-08 ms/manuscript.Rmd
Fix table output location, move references

@@
 library(drake)
+library(gt)
 library(patchwork)
@@
 
-## Figure Legends
+## References
+<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
+<div id="refs"></div>
+
+\clearpage
 
@@
   as_rtf() %>%
-  write_lines("ms/table1.rtf")
+  write_lines(here("ms/table1.rtf"))
 
@@
   as_rtf() %>%
-  write_lines("ms/table2.rtf")
+  write_lines(here("ms/table2.rtf"))
 
@@
   as_rtf() %>%
-  write_lines("ms/table3.rtf")
+  write_lines(here("ms/table3.rtf"))
 
@@
   as_rtf() %>%
-  write_lines("ms/table4.rtf")
+  write_lines(here("ms/table4.rtf"))
 
@@
   as_rtf() %>%
-  write_lines("ms/table5.rtf")
+  write_lines(here("ms/table5.rtf"))
 
@@
 \clearpage
-
-## Figures
 
@@
 ```
-
-\clearpage
-
-## References

### 2019-08-08 ms/manuscript.Rmd
Output figures as pdf to ms/

@@
 climate_plot
+ggsave(plot = climate_plot, filename = here("ms/fig1.pdf"), height = 4, width = 6, units = "in")
 ```
@@
 pca_plot
+ggsave(plot = pca_plot, filename = here("ms/fig2.pdf"), height = 6, width = 6, units = "in")
 ```
@@
 traits_with_tree
+ggsave(plot = traits_with_tree, filename = here("ms/fig3.pdf"), height = 8, width = 6, units = "in")
 ```
@@
 importance_heatmap
+ggsave(plot = importance_heatmap, filename = here("ms/fig4.pdf"), height = 4, width = 6, units = "in")
 ```
@@
 combined_cwm_plots
+ggsave(plot = combined_cwm_plots, filename = here("ms/fig5.pdf"), height = 4, width = 8, units = "in")
 ```
@@
 combined_comm_div_plots
+ggsave(plot = combined_comm_div_plots, filename = here("ms/fig6.pdf"), height = 8, width = 4, units = "in")
 ```

### 2019-08-08 ms/manuscript.Rmd
Add data availability statement
Change figure output to tiff
- as per New Phyt requirements

--- 2019-08-08 ms/manuscript.Rmd
@@
 
-Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife. Data have been deposited in Dryad (\*\*\* ADD DOI \*\*\*).
+### Data accessibility
+
+Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife.
+Raw and processed data are openly available from the Dryad repository at (http://doi.org/[\*\*\* ADD DOI \*\*\*]) [@Nitta2019].
 
@@
 climate_plot
+ggsave(plot = climate_plot, filename = here("ms/fig1.tiff"), height = 4, width = 6, units = "in", dpi = "retina")
 ```
@@
 pca_plot
+ggsave(plot = pca_plot, filename = here("ms/fig2.tiff"), height = 6, width = 6, units = "in", dpi = "retina")
 ```
@@
 traits_with_tree
+ggsave(plot = traits_with_tree, filename = here("ms/fig3.tiff"), height = 8, width = 6, units = "in", dpi = "retina")
 ```
@@
 importance_heatmap
+ggsave(plot = importance_heatmap, filename = here("ms/fig4.tiff"), height = 4, width = 6, units = "in", dpi = "retina")
 ```
@@
 combined_cwm_plots
+ggsave(plot = combined_cwm_plots, filename = here("ms/fig5.tiff"), height = 4, width = 8, units = "in", dpi = "retina")
 ```
@@
 combined_comm_div_plots
+ggsave(plot = combined_comm_div_plots, filename = here("ms/fig6.tiff"), height = 8, width = 4, units = "in", dpi = "retina")
 ```

### 2019-08-08 ms/manuscript.Rmd
Use :: for here::

--- 2019-08-08 ms/manuscript.Rmd
@@
 library(kableExtra)
-library(here)
 library(magrittr)
@@
 # Read in csv version of the table
-traits_used <- read_csv(here("data/table_1.csv"))
+traits_used <- read_csv(here::here("data/table_1.csv"))
 
@@
   as_rtf() %>%
-  write_lines(here("ms/table2.rtf"))
+  write_lines(here::here("ms/table2.rtf"))
 
@@
   as_rtf() %>%
-  write_lines(here("ms/table3.rtf"))
+  write_lines(here::here("ms/table3.rtf"))
 
@@
   as_rtf() %>%
-  write_lines(here("ms/table4.rtf"))
+  write_lines(here::here("ms/table4.rtf"))
 
@@
   as_rtf() %>%
-  write_lines(here("ms/table5.rtf"))
+  write_lines(here::here("ms/table5.rtf"))
 
@@
 ```{r traits-used-image, fig.width=7, fig.height=7, fig.cap = "Examples of traits used in this study."}
-knitr::include_graphics(here("images/Figure_3_1_traits_lowres.pdf"))
+knitr::include_graphics(here::here("images/Figure_3_1_traits_lowres.pdf"))
 ```
@@
 climate_plot
+ggsave(plot = climate_plot, filename = here::here("ms/fig1.tiff"), height = 4, width = 6, units = "in", dpi = "retina")
 ```
@@
 pca_plot
+ggsave(plot = pca_plot, filename = here::here("ms/fig2.tiff"), height = 6, width = 6, units = "in", dpi = "retina")
 ```
@@
 traits_with_tree
+ggsave(plot = traits_with_tree, filename = here::here("ms/fig3.tiff"), height = 8, width = 6, units = "in", dpi = "retina")
 ```
@@
 importance_heatmap
+ggsave(plot = importance_heatmap, filename = here::here("ms/fig4.tiff"), height = 4, width = 6, units = "in", dpi = "retina")
 ```
@@
 combined_cwm_plots
+ggsave(plot = combined_cwm_plots, filename = here::here("ms/fig5.tiff"), height = 4, width = 8, units = "in", dpi = "retina")
 ```
@@
 combined_comm_div_plots
+ggsave(plot = combined_comm_div_plots, filename = here::here("ms/fig6.tiff"), height = 8, width = 4, units = "in", dpi = "retina")
 ```

### 2019-08-08 ms/manuscript.Rmd
Update MS

@@
 ^4^Current affiliation: Department of Botany, National Museum of Natural History, Smithsonian Institute, Washington, D.C., 20013, U.S.A.
+
+Author for correspondence:  
+*Joel H. Nitta*  
+*Tel: +1 6174960515*  
+*Email: joelnitta@gmail.com*
+
+```{r word-count table}
+tibble(
+  words = c(
+    "Total word count (excluding summary, references and legends):_n",
+    "Summary:_n",
+    "Introduction:_n",
+    "Materials and Methods:_n",
+    "Results:_n",
+    "Discussion:_n",
+    "Acknowledgements:_n"),
+  figs_tabs = c(
+    "No. of figures:_6 (all in color)",
+    "No. of tables:_5",
+    "No. of supporting information files:_6 (Fig. S1; Table S1-S5)", "", "", "", "")
+) %>%
+  separate(words, c("words", "count"), sep = "_") %>%
+  separate(figs_tabs, c("figs_tabs", "n"), sep = "_", fill = "right") %>%
+  mutate(n = replace_na("")) %>%
+  gt() %>%
+  cols_label(
+    words = "",
+    count = "",
+    figs_tabs = "",
+    n = "",
+  ) %>% 
+  as_rtf() %>%
+  write_lines(here::here("ms/word_count.rtf"))
+```
 
@@
 We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate phylogenetic and functional diversity metrics.
-Hereafter, "MPD~phy~" and "MNTD~phy", and "MPD~func~" and "MNTD~func~", will be used to refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.
+Hereafter, "MPD~phy~" and "MNTD~phy~", and "MPD~func~" and "MNTD~func~", will be used to refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.
 
@@
 Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife.
-Raw and processed data are openly available from the Dryad repository at (http://doi.org/[\*\*\* ADD DOI \*\*\*]) [@Nitta2019].
+Raw and processed data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED [@Nitta2019].
 

### 2019-08-08 ms/manuscript.Rmd
Separate out references in Table 1 into their own column

@@
     data_type = "Data type (unit)",
-    ecological_significance = "Ecological Significance"
+    ecological_significance = "Ecological Significance",
+    reference = "Reference(s)"
   ) %>% 
   as_rtf() %>%
-  write_lines(here("ms/table1.rtf"))
+  write_lines(here::here("ms/table1.rtf"))
 
@@
 traits_used %>% 
-  rename(`Data type (unit)` = data_type, `Functional significance` = ecological_significance) %>%
+  rename(`Data type (unit)` = data_type, `Functional significance` = ecological_significance, `Reference(s)` = reference) %>%
   column_to_rownames("trait") %>%
@@
   pack_rows("Gametophyte traits", 8, 11) %>%
-  column_spec(3, width = "15em") %>%
-  footnote(general = "References: 1, Watkins \\\\textit{et al.} (2010); 2, Vogel (1968); 3, Creese \\\\textit{et al.} (2011); 4, Kluge \\\\& Kessler (2007); 5, Wright \\\\textit{et al.} (2004); 6, Watkins \\\\textit{et al.} (2007b); 7, Pittermann \\\\textit{et al.} (2013); 8, Farrar \\\\textit{et al.} (2008); 9, Wollenweber \\\\& Schneider (2000).
-Noncordate gametophyte morphotypes include ribbon, strap and filamentous.", general_title = "", threeparttable = TRUE, escape = FALSE)
+  column_spec(2, width = "8em") %>%
+  column_spec(3, width = "12em") %>%
+  column_spec(4, width = "8em") %>%
+  footnote(general = "Noncordate gametophyte morphotypes include ribbon, strap and filamentous.", general_title = "", threeparttable = TRUE, escape = TRUE)
 ```

### 2019-08-08 code/functions.R
Don't show frond with in CWM figure

@@ -2607,14 +2607,14 @@ combine_cwm_plots <- function (scatterplots) {
     mutate(
       resp_var = factor(
         resp_var, 
-        levels = c("stipe", "length", "width", "rhizome", 
+        levels = c("stipe", "length", "rhizome", 
                    "dissection", "pinna", "sla"))
     )
   
   # Remove un-needed plot features.
   # Can't use ggplot `+` with mutate(), etc., so do old-fashioned loop.
   for(i in 1:nrow(plots_df)) {
-    if(plots_df$resp_var[[i]] %in% c("stipe", "length", "width")) {
+    if(plots_df$resp_var[[i]] %in% c("stipe", "length", "rhizome")) {
       plots_df$plot[[i]] <- plots_df$plot[[i]] + 
         theme(
           axis.title.x = element_blank(),
@@ -2627,10 +2627,10 @@ combine_cwm_plots <- function (scatterplots) {
   }
   
   # Combine plots into single output
-  wrap_plots(plots_df$plot, ncol = 4, nrow = 2) & theme(
-    legend.position = "none",
+  wrap_plots(plots_df$plot, ncol = 3, nrow = 2) & theme(
+    legend.position = "none" # ,
     # Tweak margins to remove whitespace between plots
-    plot.margin = margin(t = 0.10, r = 0, b = 0, l = 0.10, unit = "in")
+    # plot.margin = margin(t = 0.10, r = 0, b = 0, l = 0.10, unit = "in")
   )
   
 }

### 2019-08-08 code/plan.R

@@ -436,9 +436,9 @@ plan <- drake_plan(
 
   # Make CWM scatterplots.
   cwm_scatterplots = map2(
-    .x = c("stipe", "length", "width", "rhizome",
+    .x = c("stipe", "length", "rhizome",
     "sla", "pinna", "dissection") %>% set_names(.),
-    .y = c("Stipe length (cm)", "Frond length (cm)", "Frond width (cm)", "Rhizome dia. (cm)",
+    .y = c("Stipe length (cm)", "Frond length (cm)", "Rhizome dia. (cm)",
           "SLA", "Pinna no.", "Frond dissection"),
     ~ make_cwm_scatterplot(
       data = cwm_long,

### 2019-08-08 ms/manuscript.Rmd
Don't show frond with in CWM figure
Use pdf instead of tiff for figs, fix indenting of authors

@@
 
-<!-- Set indent to about 3 characters throughout -->
-\setlength{\parindent}{3ex}
+<!-- Don't indent authors -->
+\setlength{\parindent}{0ex}
 
@@
 *Email: joelnitta@gmail.com*
+
+<!-- Set indent to about 3 characters hereafter -->
+\setlength{\parindent}{3ex}
 
@@
 Epiphytic communities in green, terrestrial communities in brown.
-Error bars are community-weighted standard deviation (values below zero not shown).
+Error bars are each one standard deviation (values below zero not shown).
 Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at $\pval<0.05$.
 \tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at $\pval<0.05$; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+Results for frond width were very similar to frond length and not shown.
 
@@
 
-```{r traits-used-image, fig.width=7, fig.height=7, fig.cap = "Examples of traits used in this study."}
+```{r traits-used-image, fig.height = 7, fig.width = 7, fig.cap = "Examples of traits used in this study."}
 knitr::include_graphics(here::here("images/Figure_3_1_traits_lowres.pdf"))
@@
 
-```{r climate, fig.height = 4, fig.width = 6, fig.cap = "Microclimatic variables."}
+```{r climate, fig.height = 4, fig.width = 7, fig.cap = "Microclimatic variables."}
 climate_plot
-ggsave(plot = climate_plot, filename = here::here("ms/fig1.tiff"), height = 4, width = 6, units = "in", dpi = "retina")
+ggsave(plot = climate_plot, filename = here::here("ms/fig1.pdf"), height = 4, width = 7, units = "in")
 ```
@@
 pca_plot
-ggsave(plot = pca_plot, filename = here::here("ms/fig2.tiff"), height = 6, width = 6, units = "in", dpi = "retina")
+ggsave(plot = pca_plot, filename = here::here("ms/fig2.pdf"), height = 6, width = 6, units = "in")
 ```
@@
 traits_with_tree
-ggsave(plot = traits_with_tree, filename = here::here("ms/fig3.tiff"), height = 8, width = 6, units = "in", dpi = "retina")
+ggsave(plot = traits_with_tree, filename = here::here("ms/fig3.pdf"), height = 8, width = 6, units = "in")
 ```
@@
 importance_heatmap
-ggsave(plot = importance_heatmap, filename = here::here("ms/fig4.tiff"), height = 4, width = 6, units = "in", dpi = "retina")
+ggsave(plot = importance_heatmap, filename = here::here("ms/fig4.pdf"), height = 4, width = 6, units = "in")
 ```
@@
 
-```{r cwm-div, fig.height = 4, fig.width = 8, fig.cap = "Community-weighted means."}
+```{r cwm-div, fig.height = 4, fig.width = 7, fig.cap = "Community-weighted means."}
 combined_cwm_plots
-ggsave(plot = combined_cwm_plots, filename = here::here("ms/fig5.tiff"), height = 4, width = 8, units = "in", dpi = "retina")
+ggsave(plot = combined_cwm_plots, filename = here::here("ms/fig5.pdf"), height = 4, width = 7, units = "in")
 ```
@@
 combined_comm_div_plots
-ggsave(plot = combined_comm_div_plots, filename = here::here("ms/fig6.tiff"), height = 8, width = 4, units = "in", dpi = "retina")
+ggsave(plot = combined_comm_div_plots, filename = here::here("ms/fig6.pdf"), height = 8, width = 4, units = "in")
 ```

### 2019-08-10 ms/manuscript.Rmd
Cut text from statistical analysis section

@@
 
-We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each site from the raw climate data, then overall means from the daily values.
+We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each site, then overall means from the daily values.
 We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (hereafter, "climatic variables").
@@
 Mt.
-Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established, which are part of a mountain complex on the interior of the island (J.
+Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established (J.
 Nitta, pers.
@@
 We tested for phylogenetic signal in quantitative traits using two metrics that relate the observed trait values to those expected under a model of trait evolution following Brownian motion (BM), Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
-$\lambda$ is a scaling parameter that ranges from 0 (traits evolving randomly, as on a star phylogeny) to 1 (traits evolving along branches according to Brownian motion).
+$\lambda$ is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion).
 \K describes the ratio between the amount of observed variance in traits vs.
-the amount of variance expected under a BM model of trait evolution: when $\K = 1$, traits are evolving according to BM; when $\K > 1$, traits are more conserved than expected under BM; when $\K < 1$, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
+the amount of variance expected under BM: when $\K = 1$, traits are evolving according to BM; when $\K > 1$, traits are more conserved than expected under BM; when $\K < 1$, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
 We tested for significance in \K by comparing observed values of \K against a null distribution of \num{1000} trees with trait values shuffled randomly across the tips, and in $\lambda$ by conducting a likelihood ratio test comparing the log likelihoods of the observed value of $\lambda$ vs.
 $\lambda = 0$ (no phylogenetic signal).
-We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' *D* [@Fritz2010], which is designed to test for signal in binary traits.
-Values of \D range from 0 under a model of trait evolution by BM, to 1 for random distribution of traits; negative values (greater trait conservation than BM) and values greater than 1 (trait overdispersion) are possible.
-We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], which also conducts a significance test by comparing the observed value of \D with distributions of simulated values produced under one of two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM threshold model.
+We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' *D* [@Fritz2010].
+Values of \D range from 0 under a model of trait evolution by BM, to 1 for random distribution of traits.
+We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], which also conducts a significance test by comparing the observed \D with distributions of simulated values produced under one of two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM threshold model.
 To account for possible circularity in measuring the phylogenetic signal of traits that were scored based on taxonomy, we conducted the same analyses on a dataset excluding species that had been scored this way ("strict" dataset).
@@
 This method compares the likelihood between two alternative models for a pair of binary traits.
-The first model tests the null hypothesis that rates of evolution of the two traits are independent.
-The second model tests the alternative hypothesis that the rates of evolution of one trait depends on the other.
-The likelihoods of obtaining the observed data are calculated for each model, then compared using a log likelihood test.
-A significantly better fit of the dependent model indicates that evolution of the two traits is correlated.
+In the first model, rates of evolution of the two traits are independent.
+In the second model, the rate of evolution of one trait depends on the other.
+A significantly better fit of the second model indicates that evolution of the two traits is correlated.
 The same analyses were also done on the "strict" dataset.
@@
 
-We used distance-based methods to analyze phylogenetic and functional diversity.
+We used distance-based methods to analyze community phylogenetic and functional diversity.
 Mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) describe the mean distance between all species in a community (MPD) or the mean distance between the closest pairs of species in a community (MNTD) [@Webb2000].
-These metrics were originally formulated for phylogenetic distances (thus "phylogenetic" in the name), but can be applied in the same way to trait-based distances, enabling a direct comparison between the two [@Sessa2018].
-For functional diversity, we used Gower's dissimilarity coeffecient [@Gower1971] because it can use continuous, binary, and qualitative traits, and accepts missing data.
+Although these metrics were originally formulated for phylogenetic distances, they can be applied in the same way to trait-based distances, enabling a direct comparison between the two [@Sessa2018].
+We used Gower's dissimilarity coeffecient [@Gower1971] to calculate functional distances because it can use continuous, binary, and qualitative traits, and accepts missing data.
 We log-transformed traits as needed to account for skew, scaled each trait by its standard deviation, and excluded frond width and length (both frond size traits highly correlated with stipe length) before calculating the distance matrix using the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
@@
 Raw values of MPD and MNTD are statistically related to species richness [@Swenson2014]; to obtain metrics that are uncorrelated with species richness, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{999} randomized communities (equivalent to the Net Relatedness Index and Nearest Taxon Index, respectively, of @Webb2000 with the sign reversed).
-Selection of the randomization procedure can strongly bias outcomes [@Gotelli2006;@Ulrich2012].
-Our null hypothesis is that fern spores are easily capable of dispersal between terrestrial and epiphytic habitats, and across the relatively short distances separating our plots.
-We therefore constructed null communities using the "independent swap" algorithm of @Gotelli2000 that maintains species frequency and abundance while randomizing species identities.
+We constructed null communities using the "independent swap" algorithm of @Gotelli2000 that maintains species frequency and abundance while randomizing species identities.
 We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available (`r Ntip(phy)` \si{\species} total; [@Nitta2017]) for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct`) for functional null communities.
 We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate phylogenetic and functional diversity metrics.
-Hereafter, "MPD~phy~" and "MNTD~phy~", and "MPD~func~" and "MNTD~func~", will be used to refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.
+Hereafter, "MPD~phy~" and "MNTD~phy~", and "MPD~func~" and "MNTD~func~", refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.
 
-We tested for differences in diversity metrics between epiphytic and terrestrial communities with a paired \tval-test.
 We simultaneously tested the effects of climate and growth habit on diversity metrics using a full-subsets multiple regression approach with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].

### 2019-08-10 ms/manuscript.Rmd
Cut text from results and discussion

@@
 The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
-In the standard PCA, size-related traits tended to load on PC1, with SLA, pinna, and dissection on PC2.
+In the standard PCA, size-related traits tended to load on PC1, with SLA, pinna number, and degree of dissection on PC2.
 In the phylogenetic PCA, SLA was strongly loaded on PC1, while dissection, pinna, and other size-related traits were mostly loaded on PC2.
@@
 However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of $\K < 1$ (\tab \@ref(tab:phylosig-cont)).
-All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under a scenario of evolution by BM (\tab \@ref(tab:phylosig-binary)).
+All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under BM (\tab \@ref(tab:phylosig-binary)).
 Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; \tab \@ref(tab:phylosig-binary)).
@@
 In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (\fig \ref{fig:heatmap}).
-Community-weighted means of frond length, frond width, stipe length, and rhizome diameter were correlated (Pearson's correlation coefficient > 0.9), so of these we only show stipe length.
 Most of the the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (\tab S5).
@@
 
-Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial communities for pinna number, and size-related traits (frond length, frond width, stipe length, and rhizome diameter) (\tval-test, all \pval < 0.05; \fig \@ref(fig:cwm-div)).
-Standard deviation of community-weighted trait values was larger for size-related traits in terrestrial communities, but smaller for other traits such as frond dissection and pinna number.
-Specific leaf area decreased with elevation, and was lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval = `r get_mod_result(model_summ_print, "sla", "p")`).
-Degree of lamina dissection decreased with elevation in epiphytes, but slightly increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval = `r get_mod_result(model_summ_print, "dissection", "p")`).
+Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and pinna number (\tval-test, all \pval < 0.05; \fig \@ref(fig:cwm-div)).
+Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval = `r get_mod_result(model_summ_print, "sla", "p")`).
+Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval = `r get_mod_result(model_summ_print, "dissection", "p")`).
 Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (\fig \@ref(fig:cwm-div)).
@@
 However, contrary to our hypothesis, epiphytic fern communities appear to be more diverse than terrestrial communities.
-This suggests that environmental filtering plays a strong role in limiting plant size, while allowing for diversification across other trait axes in epiphytes.
+This suggests that environmental filtering plays a strong role in limiting plant size, while allowing for diversification across other trait axes in epiphytic ferns.
 
@@
 Our results at the local scale are corroborated by @Testo2018a, who also observed abrupt size reduction associated with epiphytic growth in a globally sampled fern phylogeny.
-Stipe length may be a critical trait controlling whole leaf hydraulic conductance; since ferns have resistive stipes in general, species growing in water-limited habitats are expected to decrease stipe length to minimize resistance [@WatkinsJr.2010].
+Stipe length in particular may be a critical trait controlling whole leaf hydraulic conductance; since ferns have resistive stipes in general, species growing in water-limited habitats are expected to decrease stipe length to minimize resistance [@WatkinsJr.2010].
 Our results support previous non-phylogenetically controlled studies that found a prevalence of short stipes in epiphytic ferns [@WatkinsJr.2010; @Creese2011].
-However, given that these were all correlated with each other (Pearson's correlation coefficient > 0.9), and similar traits have been shown to scale together previously [@Arcand2008; @Creese2011], careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
+However, given that these were all correlated with each other in our study, and similar traits have been shown to scale together previously [@Arcand2008; @Creese2011], careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
 
@@
 In contrast to their clearly smaller size, epiphytes were remarkably diverse in other traits.
-Epiphytes tended to have less dissected fronds with fewer pinna than terrestrial species (\tab \@ref(tab:pics)), but the range of values for these traits was quite wide at the community level (\fig \@ref(fig:cwm-div)).
+Epiphytes tended to have less dissected fronds with fewer pinna than terrestrial species (\tab \@ref(tab:pics)), but the range of values for these traits was quite wide at the community level (error bars in \fig \@ref(fig:cwm-div)).
 Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
@@
 slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a].
-We hypothesized that epiphytes would have lower SLA, given their more extreme growing conditions.
+We expected that epiphytes would have lower SLA, given their more extreme growing conditions.
 However, the parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (\fig \@ref(fig:cwm-div)) indicates that SLA may be responding more strongly to temperature in both epiphytes and terrestrial species, while only being weakly modified by humidity.
@@
 Gemmae can also function sexually by producing antherida in the presence of antheridiogen secreted by other gametophytes [@Emigh1977], which may confer a selective advantage in the epiphytic environment by enabling outcrossing [@Dassler2001].
-It is possible that some, but not all, epiphytic ferns rely on such strategies, and that they are not under similarly strong selection as other traits such as stipe length and gametophyte morphotype.
+It is possible that some, but not all, epiphytic ferns rely on such strategies, and that they are not under similarly strong selection as other traits such as leaf size and gametophyte morphotype.
 Interestingly, when we classified 'widespread' species as those with gametophytes occurring above or below the minimum or maximum elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with widespread growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution).
@@
 In a study comparing terrestrial and epiphytic ferns along an elevational gradient from \SIrange{30}{3000}{\meter} in Costa Rica, @WatkinsJr.2006a found that epiphytes tended to sort into low-trunk and canopy zones, with only 18\% found in both.
-It seems that the canopy habitat limits overall size in epiphytic ferns, but that species with different traits then sort into different niches.
 The goal of our study was to quantify differences between epiphytic and terrestrial fern communities; future research should focus on bringing comparative phylogenetic and community ecological methods to bear on the large body of work characterizing differences within epiphytic microhabitats.

### 2019-08-10 ms/manuscript.Rmd
Add caption for heatmap

@@
 
+**\fig \@ref(fig:heatmap)**: Variable importance scores from a full-subsets analysis using general additive models exploring the effect of climate and growth habit (terrestrial vs.
+epiphytic growth) on community diversity of ferns on Moorea, French Polynesia.
+Predictor variable abbreviations: VPD, vapor pressure deficit; SD, standard deviation.
+Response variable abbreviations: MPD~phy~, standard effect size (SES) of mean phylogenetic distance; MNTD~phy~, SES of mean nearest taxon distance; MPD~func~, SES of mean functional distance; MNTD~func~, SES of mean nearest functional distance; SLA, specific leaf area.
+All trait names refer to community-weighted mean values.
+Community-weighted means of frond length, frond width, stipe length, and rhizome diameter were correlated (Pearson's correlation coefficient > 0.9), so of these we only show stipe length.
+
 **\fig \@ref(fig:cwm-div)**: Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.

### 2019-08-11 ms/manuscript.Rmd
Cut text from methods

@@
 We preferred morphological traits over physiological traits for our study because data for morphological traits are easier to gather at the scale of our study (\circa \SI{130}{\species} distributed over an elevational gradient from \SIrange{200}{2000}{\meter}).
-In addition, plant morphology and physiology are closely linked, and physiological traits such as rates of photosynthesis and water use efficiency are known to be correlated with morphological traits such as stipe length in ferns [@WatkinsJr.2010].
-Studying the relationships between ecology and morphological traits is useful from a paleobotanical and climate change perspective, as trends in morphological trait variation in the fossil record can be used to gain insight into past climatic conditions [@Chaloner1997; @McElwain2007a; @McElwain2007b; @Royer2008].
+Plant morphology and physiology are closely linked, and physiological traits such as rates of photosynthesis and water use efficiency are known to be correlated with morphological traits such as stipe length in ferns [@WatkinsJr.2010].
 
@@
 
-For comparative analyses, we used the dated phylogenetic tree of [@Nitta2017], which includes all species from Moorea and some from the neighboring island of Tahiti.
-We trimmed the tree to species occurring only on Moorea (*n* = `r length(species_list)` \si{\species}) using the \function{drop.tip} function of the \package{ape} package [@Paradis2004] in R [@RCoreTeam2015].
-We tested for phylogenetic signal in quantitative traits using two metrics that relate the observed trait values to those expected under a model of trait evolution following Brownian motion (BM), Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
-$\lambda$ is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion).
+For comparative analyses, we used the dated phylogenetic tree of @Nitta2017.
+We tested for phylogenetic signal in quantitative traits using Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
+$\lambda$ is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion [BM]).
 \K describes the ratio between the amount of observed variance in traits vs.
-the amount of variance expected under BM: when $\K = 1$, traits are evolving according to BM; when $\K > 1$, traits are more conserved than expected under BM; when $\K < 1$, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
-We tested for significance in \K by comparing observed values of \K against a null distribution of \num{1000} trees with trait values shuffled randomly across the tips, and in $\lambda$ by conducting a likelihood ratio test comparing the log likelihoods of the observed value of $\lambda$ vs.
+the amount of variance expected under BM: when $\K > 1$, traits are more conserved than expected under BM; when $\K < 1$, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
+We tested for significance by comparing observed \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in $\lambda$ by conducting a likelihood ratio test comparing the log likelihoods of the observed value of $\lambda$ vs.
 $\lambda = 0$ (no phylogenetic signal).
@@
 Values of \D range from 0 under a model of trait evolution by BM, to 1 for random distribution of traits.
-We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], which also conducts a significance test by comparing the observed \D with distributions of simulated values produced under one of two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM threshold model.
+We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], and conducted significance test by comparing observed \D to distributions of simulated values produced under two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM model.
 To account for possible circularity in measuring the phylogenetic signal of traits that were scored based on taxonomy, we conducted the same analyses on a dataset excluding species that had been scored this way ("strict" dataset).
@@
 We detected significant phylogenetic signal in most traits, so we tested for differences between traits of epiphytic vs.
-terrestrial species using methods that can test for significant differences between traits associated with a binary character while accounting for phylogeny.
+terrestrial species using phylogenetic comparative methods.
 For quantitative traits, we used the \function{brunch} function in the \package{caper} package [@Orme2013], which calculates phylogenetically independent contrasts in a quantitative character between alternative states of a binary trait.
 For qualitative (binary) traits, we used Pagel's test of correlated evolution [@Pagel1994] as implemented with the \function{fitPagel} function in the \package{phytools} package [@Revell2012].
-This method compares the likelihood between two alternative models for a pair of binary traits.
+This method compares likelihoods between alternative models for a pair of binary traits.
 In the first model, rates of evolution of the two traits are independent.
@@
 
-We investigated differences in multivariate trait space between epiphytes and terrestrial species using Principal Components Analysis (PCA).
-We used only quantitative traits, and included only species that had no missing data for any traits (*n* = `r quant_traits_nsp` \si{\species}).
+We investigated differences in multivariate trait space between epiphytes and terrestrial species using principal components analysis (PCA).
+We used only quantitative traits, and included only species that had no missing data (*n* = `r quant_traits_nsp` \si{\species}).
 We first log-transformed traits as necessary to reduce skew, then scaled each trait by its standard deviation.
@@
 We used distance-based methods to analyze community phylogenetic and functional diversity.
-Mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) describe the mean distance between all species in a community (MPD) or the mean distance between the closest pairs of species in a community (MNTD) [@Webb2000].
+Mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) describe the mean distance between all species in a community (MPD) or the mean distance between the closest pairs of species in a community (MNTD) [@Webb2000; Webb2002].
 Although these metrics were originally formulated for phylogenetic distances, they can be applied in the same way to trait-based distances, enabling a direct comparison between the two [@Sessa2018].
-We used Gower's dissimilarity coeffecient [@Gower1971] to calculate functional distances because it can use continuous, binary, and qualitative traits, and accepts missing data.
-We log-transformed traits as needed to account for skew, scaled each trait by its standard deviation, and excluded frond width and length (both frond size traits highly correlated with stipe length) before calculating the distance matrix using the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
+We log-transformed traits as needed to account for skew, scaled each trait by its standard deviation, excluded frond width and length (both frond size traits highly correlated with stipe length), then calculated the distance matrix using Gower's dissimilarity coeffecient [@Gower1971] with the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
 To test for the effect of different trait selection schemes, we compared the results of running the same analyses for various sets of traits with and without log-transformation (Supporting Information).
 
-Raw values of MPD and MNTD are statistically related to species richness [@Swenson2014]; to obtain metrics that are uncorrelated with species richness, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{999} randomized communities (equivalent to the Net Relatedness Index and Nearest Taxon Index, respectively, of @Webb2000 with the sign reversed).
+Raw values of MPD and MNTD are statistically related to species richness [@Swenson2014]; to obtain metrics that are uncorrelated with species richness, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{999} randomized communities.
 We constructed null communities using the "independent swap" algorithm of @Gotelli2000 that maintains species frequency and abundance while randomizing species identities.
@@
 
-We simultaneously tested the effects of climate and growth habit on diversity metrics using a full-subsets multiple regression approach with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
+We tested the effects of climate and growth habit on diversity metrics using a full-subsets multiple regression approach with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
 For each diversity metric, we built a set of models (Gaussian family, with site as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
 Our dataset included `r nrow(div_metrics_all) - nrow(climate_select)` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
-Since we found that all climatic variables are related to elevation (see Results), we also constructed linear models for each diversity metric using the same model selection procedure as in the climate analysis.
+Therefore, we also constructed linear models for each diversity metric related to elevation including all sites using the same model selection procedure as in the climate analysis.
 We checked for spatial autocorrelation in the model residuals with Moran's *I* [@Moran1950] as implemented in the \package{spdep} R package [@Bivand2013].

### 2019-08-11 ms/manuscript.Rmd
Update MS
- Add @Tsutsumi2013 for gameto evo
- Add description of CWMs
- Add mention of tree in MPD/MNTD
- Cut text in conclusion, discussion, and methods (null model)

@@
 Many of these same adaptations are found in epiphytic seed plants [@Benzing1990].
-However, the evolution of epiphytism in ferns was more complex given that ferns rely on a free-living gametophyte.
+However, the evolution of epiphytism in ferns was more complex given that ferns rely on a free-living gametophyte [@Tsusumi2013].
 The fern gametophyte is anatomically simple, and differs from the sporophyte in anatomy, morphology, and physiology.
@@
 
+We calculated community-weighted means (CWMs) and standard deviations of quantitative traits by using the overall mean trait value of each species weighted by their abundances in each community.
+
 We used distance-based methods to analyze community phylogenetic and functional diversity.
@@
 Although these metrics were originally formulated for phylogenetic distances, they can be applied in the same way to trait-based distances, enabling a direct comparison between the two [@Sessa2018].
-We log-transformed traits as needed to account for skew, scaled each trait by its standard deviation, excluded frond width and length (both frond size traits highly correlated with stipe length), then calculated the distance matrix using Gower's dissimilarity coeffecient [@Gower1971] with the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
+For phylogenetic distances, we used the tree of @Nitta2017.
+For functional distances, we log-transformed and scaled traits as in the PCA, excluded frond width and length (both frond size traits highly correlated with stipe length), then calculated the distance matrix using Gower's dissimilarity coeffecient [@Gower1971] with the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
 To test for the effect of different trait selection schemes, we compared the results of running the same analyses for various sets of traits with and without log-transformation (Supporting Information).
@@
 Raw values of MPD and MNTD are statistically related to species richness [@Swenson2014]; to obtain metrics that are uncorrelated with species richness, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{999} randomized communities.
-We constructed null communities using the "independent swap" algorithm of @Gotelli2000 that maintains species frequency and abundance while randomizing species identities.
+We constructed null communities using the "independent swap" algorithm [@Gotelli2000].
 We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available (`r Ntip(phy)` \si{\species} total; [@Nitta2017]) for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct`) for functional null communities.
@@
 
-We tested the effects of climate and growth habit on diversity metrics using a full-subsets multiple regression approach with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
+We tested the effects of climate and growth habit on diversity metrics using full-subsets multiple regression with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
 For each diversity metric, we built a set of models (Gaussian family, with site as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
@@
 Results for functional community diversity contradicted our hypotheses; epiphytic communities were more functionally diverse than terrestrial as measured by both MPD and MNTD (\fig \@ref(fig:comm-div)).
-However, this result is highly dependent on the traits used to measure functional diversity; when the same analyses were done using only size-related traits (scaled, untransformed stipe length, frond length and width, and rhizome diameter), the opposite result was observed (\fig S1).
+However, this result is dependent on the traits used to measure functional diversity; when the same analyses were done using only size-related traits (scaled, untransformed stipe length, frond length and width, and rhizome diameter), the opposite result was observed (\fig S1).
 Futhermore, terrestrial species show a much wider range of values for size traits (\fig \@ref(fig:cwm-div)), and neither functional group appears to be restricted to a smaller area in trait space (\fig \@ref(fig:pca)).
-This suggests that epiphytic ferns are subject to extreme environmental filtering for size-related traits, while simultaneously diversifying along other trait axes.
-This could be maintained by niche partitioning within epiphytic habitats.
+This suggests that epiphytic ferns are subject to environmental filtering for size-related traits, while simultaneously diversifying along other trait axes.
+This could be maintained by niche partitioning within epiphytic habitats, which may be more diverse than terrestrial for a given area [@Rojas-Hernandez2018].
 Epiphytic habitats are highly heterogeneous, and microclimatic conditions vary with height along the host and distance from the trunk [@Hietz1995].
-The vertical stratification of epiphytic species into different zones along height gradients in host trees is well documented [@Zotz2007].
-In a study comparing terrestrial and epiphytic ferns along an elevational gradient from \SIrange{30}{3000}{\meter} in Costa Rica, @WatkinsJr.2006a found that epiphytes tended to sort into low-trunk and canopy zones, with only 18\% found in both.
+The vertical stratification of epiphytic species has been well documented in ferns [\eg, @Parra2009; @WatkinsJr.2006a; @Acebey2017] and broader studies [\eg, @Kromer2007; @Zotz2007].
 The goal of our study was to quantify differences between epiphytic and terrestrial fern communities; future research should focus on bringing comparative phylogenetic and community ecological methods to bear on the large body of work characterizing differences within epiphytic microhabitats.
@@
 Taken together, our results support a scenario of environmental filtering acting to limit size and phylogenetic diversity in epiphytic ferns, while simulataneously enabling diversification along other trait axes.
-This is consistent with a much more broadly sampled study finding that epiphytic and terrestrial ferns do not differ in rates of body size evolution [@Testo2018a].
 The generality of this pattern could be further tested by investigating the functional traits of other groups with large epiphytic radiations such as orchids and bromeliads.
@@
 Epiphytism itself is likely a complex trait requiring the evolution of multiple associated traits, and the timing and order of the evolution of such traits may be related to the complex diversification dynamics observed in these studies.
-Small size seems to be a prerequisite for epiphytic growth, as indicated both by our study and @Testo2018a.
+Small size seems to be a prerequisite for epiphytic growth, as indicated both by our results and the global study of @Testo2018a.
 Here, we identify several traits in addition to plant size (\eg, pinna number, frond dissection, gametophyte morphotype) that may also belong to the suite of epiphytic traits.

### 2019-08-14 code/functions.R
Various fixes to tree plot
- Add heatmap scale
- Add "my" to time scale
- Fix order of gametophyte trait levels
- log transform only rhizome and stipe length

@@ -2399,19 +2399,21 @@ plot_traits_on_tree <- function (traits, phy, ppgi) {
     # Convert morphotype to binary trait
     # make binary morph category: 0 is noncordate, 1 is cordate
     mutate(morphotype = case_when(
-      morphotype == "cordate" ~ "1",
-      TRUE ~ "0"
+      morphotype == "cordate" ~ 1,
+      TRUE ~ 0
     )) %>%
     # Only keep species in tree
     match_traits_and_tree(phy = phy, "traits") %>%
     # Transform and rescale traits
-    # SLA is already between 0 and 1, so don't log-transform.
-    # Instead of scaling between min and max range, rescale between 0 and 1.
-    transform_traits(trans_select = c("dissection", "stipe", "length", 
-                                      "width", "rhizome", "pinna"),
-                     scale_traits = FALSE) %>%
+    transform_traits(
+      log_trans = TRUE, 
+      scale_traits = FALSE,
+      # log-transform same traits as for PCA
+      trans_select = c("length", "rhizome", "stipe", "width")
+    ) %>%
+    # Instead of scaling by standard deviation, rescale between 0 and 1.
     mutate_at(vars(dissection, sla, stipe, length, width, rhizome, pinna),
-              ~rescale(., c(0,1)))
+               ~rescale(., c(0,1)))
   
   # For phylogeny, only keep species in traits
   phy <- match_traits_and_tree(traits, phy, "tree")
@@ -2434,7 +2436,7 @@ plot_traits_on_tree <- function (traits, phy, ppgi) {
     pairedcols[c(4,3)], #gemmae: dark and light green
     pairedcols[c(6,5)], #glands: red and pink
     pairedcols[c(8,7)], #hairs: dark and light orange
-    "grey95") %>%
+    "grey50") %>% # missing values: grey50
     set_names(., c(
       "epiphytic", "terrestrial",
       "noncordate", "cordate",
@@ -2476,7 +2478,7 @@ plot_traits_on_tree <- function (traits, phy, ppgi) {
       species = fct_reorder(species, phy_order),
       trait = factor(trait, levels = c("habit", "morphotype", "gemmae", "glands", "hairs")),
       value = factor(value, levels = c("epiphytic", "terrestrial",
-                                       "noncordate", "cordate",
+                                       "cordate", "noncordate",
                                        "gemmae_present", "gemmae_absent",
[…]

### 2019-08-15 code/resolve_names.R
Add script to resolve species names

@@ -1,0 +1,86 @@
+# This script looks up scientific names (specis names with taxonomic author) using
+# a live search of taxonomic databases via the Global Name Resolver 
+# (https://resolver.globalnames.org/). Since results have the possibility
+# of changing, it is run once and the results
+# checked-in to the repo at the time of manuscript submission.
+# The script is left here though for reproducibility.
+
+library(taxize)
+library(janitor)
+library(tidyverse)
+
+# Read in list of accepted species names used in this study,
+# extract the genus and specific epithet
+species_list <- read_csv("data/nitta_2017/species.csv") %>%
+  clean_names() %>%
+  filter(include == 1, tahiti_only == 0) %>%
+  pull(genus_sp)
+
+# Reformat the genus_species list of names into a dataframe
+# with the taxon_code used to refer to that taxon in scripts,
+# and other names at various taxonomic ranks.
+taxa_names <-
+  tibble(taxon_code = species_list) %>%
+  separate(taxon_code, c("genus", "specific_epithet"), by = "_", remove = FALSE) %>%
+  mutate(informal_variety = str_match(specific_epithet, "[0-9]") %>% map_chr(1)) %>%
+  mutate(specific_epithet = str_remove_all(specific_epithet, "[0-9]")) %>%
+  mutate(species = paste(genus, specific_epithet)) %>%
+  mutate(informal_variety = case_when(
+    specific_epithet == "sp" ~ NA_character_,
+    TRUE ~ informal_variety
+  )) %>%
+  mutate(infaspecific_epithet = case_when(
+    species == "Lindsaea repens" ~ "marquesensis",
+    species == "Davallia solida" ~ "solida",
+    species == "Ctenitis sciaphila" ~ "sciaphila",
+    species == "Dicranopteris linearis" ~ "linearis",
+    species == "Deparia petersenii" ~ "congrua",
+    TRUE ~ NA_character_
+  )) %>%
+  mutate(taxon = jntools::paste3(genus, specific_epithet, infaspecific_epithet))
+
+# Use the Global Name Resolver (https://resolver.globalnames.org/) to match the names
+# to either IPNI or TROPICOS databases.
+taxa_names_with_authors <- gnr_resolve(
[…]

### 2019-08-15 ms/manuscript.Rmd
Various fixes to tree plot
- Add heatmap scale
- Add "my" to time scale
- Fix order of gametophyte trait levels
- log transform only rhizome and stipe length
Renumber SI tables after adding a new one

@@
 Moorea, French Polynesia (\ang{17;30;}S, \ang{149;50;}W) is a small (\SI{135}{\kilo\meter\squared}) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
-This distance acts as a strong barrier to dispersal [@Carson1995;@Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders \sensu @PteridophytePhylogenyGroupI2016).
+This distance acts as a strong barrier to dispersal [@Carson1995;@Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders \sensu @PteridophytePhylogenyGroupI2016) (\tab S1).
 Ferns of Moorea include representatives of all five of the Cretaceous epiphytic radiations described by @Schuettpelz2009 (\ie, polygrammoids, elaphoglossoids, asplenioids, vittarioids, and filmy ferns) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
@@
 
-Trait values and data sources are summarized in Table S1.
+Trait values and data sources are summarized in Table S2.
 
@@
 Mean, maximum, and standard deviation (SD) of temperature and mean, minimum, and SD of vapor pressure deficit (VPD) were retained after excluding variables with correlation coefficients > 0.9.
-Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (\tab S2; \fig \ref{fig:climate}).
+Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (\tab S3; \fig \ref{fig:climate}).
 Variables related to VPD differed by growth habit and elevation.
@@
 Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; \tab \@ref(tab:phylosig-binary)).
-Similar results were obtained using the "strict" dataset (\tab S3).
+Similar results were obtained using the "strict" dataset (\tab S4).
 
@@
 cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval = `r morphotype_pval`; \tab \@ref(tab:corr-evo)).
-Similar results were obtained using the "strict" dataset (\tab S4).
+Similar results were obtained using the "strict" dataset (\tab S5).
 
@@
 In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (\fig \ref{fig:heatmap}).
-Most of the the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (\tab S5).
-Spatial autocorrelation was not detected for any of the diversity metrics (\tab S5).
+Most of the the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (\tab S6).
+Spatial autocorrelation was not detected for any of the diversity metrics (\tab S6).
 
@@
 Relative value of quantitative (sporophyte) traits shown by heat gradient: low values are lighter, high values are darker.
+Stipe length and rhizome diameter were log-transformed prior to scaling.
 States of qualitative (gametophyte) states and growth habit indicated by colors in key.

### 2019-08-15 ms/SI.rmd
Move SI from si/ to ms/

--- 2019-08-06 si/SI.rmd
@@
 
-**Table S1** Trait values of ferns on Moorea, French Polynesia.
-Abbreviations: SLA, specific leaf area.
+**Table S1** List of ferns from Moorea, French Polynesia included in this study with taxonomic data.
+
+'taxon_code' is the name used for that taxon in all analyses.
+'informal_variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them.
+'name_source' indicates the source of the scientific name as follows, IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith, 2003.
+
+**Table S2** Trait values of ferns on Moorea, French Polynesia.
+
+SLA, specific leaf area.
 For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen, etc.
@@
 trait_data_for_si %>%
-  write_csv(here::here("si/Table_S1.csv"))
+  write_csv(here::here("si/Table_S2.csv"))
 ```
 
-**Table S2** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs.
-terrestrial) with elevation as a covariate.
-Abbreviations: d.f., degrees of freedom; MS, mean squares.
+**Table S3** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate.
+
+d.f., degrees of freedom; MS, mean squares.
 
@@
   select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value) %>%
-  write_csv(here::here("si/Table_S2.csv"))
+  write_csv(here::here("si/Table_S3.csv"))
 
@@
 
-**Table S3** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
-Abbreviations: P_rnd, probability of obtaining the observed value of *D* under random trait evolution; P_bm, probability of obtaining the observed value of *D* under simulation of trait evolution by Brownian motion.
+**Table S4** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+
+P_rnd, probability of obtaining the observed value of *D* under random trait evolution; P_bm, probability of obtaining the observed value of *D* under simulation of trait evolution by Brownian motion.
 
@@
          P_rnd = prob_random, P_bm = prob_brownian) %>%
-  write_csv(here::here("si/Table_S3.csv"))
+  write_csv(here::here("si/Table_S4.csv"))
 ```
 
-**Table S4** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+**Table S5** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
 Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
@@
   ) %>%
-  write_csv(here::here("si/Table_S4.csv"))
+  write_csv(here::here("si/Table_S5.csv"))
 ```
 
-**Table S5** Best-fitting models explaining community diversity metrics by climate and growth habit.
-Models within delta AICc 3 of the best-fitting model shown.
-Abbreviations: MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, *P* value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's *I*.
+**Table S6** Best-fitting models explaining community diversity metrics by climate and growth habit. Models within delta AICc 3 of the best-fitting model shown.
+
+MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, *P* value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's *I*.
 Trait names refer to community-weighted mean values (see Methods).
@@
   ) %>%
-  write_csv(here::here("si/Table_S5.csv"))
+  write_csv(here::here("si/Table_S6.csv"))
 ```

### 2019-08-15 ms/SI.rmd
Move SI files to ms/ folder

@@
 title: ""
-output: word_document
+bibliography: "si_references.bib"
+mainfont: Times New Roman.ttf
+fontsize: 12pt
+mainfontoptions: 
+- BoldFont=Times New Roman Bold.ttf
+- ItalicFont=Times New Roman Italic.ttf
+- BoldItalicFont=Times New Roman Bold Italic.ttf
+output: 
+  bookdown::pdf_document2:
+    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
+    toc: no
+    number_sections: no
+    fig_caption: yes
+    keep_tex: no
+    latex_engine: xelatex
+    pandoc_args: [ "--csl", "new-phytologist.csl"]
+nocite: | 
+  @NMNS2008, @Copeland1932, @FloraNSW2019, @Brownsey1987, @Murdock2003
 ---
@@
 trait_data_for_si %>%
-  write_csv(here::here("si/Table_S2.csv"))
+  write_csv(here::here("ms/Table_S2.csv"))
 ```
@@
   select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value) %>%
-  write_csv(here::here("si/Table_S3.csv"))
+  write_csv(here::here("ms/Table_S3.csv"))
 
@@
          P_rnd = prob_random, P_bm = prob_brownian) %>%
-  write_csv(here::here("si/Table_S4.csv"))
+  write_csv(here::here("ms/Table_S4.csv"))
 ```
@@
   ) %>%
-  write_csv(here::here("si/Table_S5.csv"))
+  write_csv(here::here("ms/Table_S5.csv"))
 ```
@@
   ) %>%
-  write_csv(here::here("si/Table_S6.csv"))
+  write_csv(here::here("ms/Table_S6.csv"))
 ```
@@
 
-ggsave(plot = fig_s1, filename = here("si/Fig_S1.pdf"), height = 8, width = 4, units = "in")
+ggsave(plot = fig_s1, filename = here::here("ms/Fig_S1.pdf"), height = 8, width = 4, units = "in")
 ```
+
+References

### 2019-08-15 code/plan.R

@@ -493,7 +493,7 @@ plan <- drake_plan(
 
   # SI ----
   si = rmarkdown::render(
-    knitr_in("si/SI.Rmd"),
+    knitr_in("ms/SI.Rmd"),
     quiet = TRUE)
 
 )

### 2019-08-15 code/resolve_names.R

@@ -83,4 +83,4 @@ taxa_names_with_authors <- gnr_resolve(
     TRUE ~ name_source
   ))
 
-write_csv(here("si/Table_S1.csv"))
+write_csv(here("ms/Table_S1.csv"))

### 2019-08-15 ms/manuscript.Rmd
Standardize figure and table names between ms and si
- as Table_1, Fig_1, etc

@@
   as_rtf() %>%
-  write_lines(here::here("ms/table1.rtf"))
+  write_lines(here::here("ms/Table_1.rtf"))
 
@@
   as_rtf() %>%
-  write_lines(here::here("ms/table2.rtf"))
+  write_lines(here::here("ms/Table_2.rtf"))
 
@@
   as_rtf() %>%
-  write_lines(here::here("ms/table3.rtf"))
+  write_lines(here::here("ms/Table_3.rtf"))
 
@@
   as_rtf() %>%
-  write_lines(here::here("ms/table4.rtf"))
+  write_lines(here::here("ms/Table_4.rtf"))
 
@@
   as_rtf() %>%
-  write_lines(here::here("ms/table5.rtf"))
+  write_lines(here::here("ms/Table_5.rtf"))
 
@@
 climate_plot
-ggsave(plot = climate_plot, filename = here::here("ms/fig1.pdf"), height = 4, width = 7, units = "in")
+ggsave(plot = climate_plot, filename = here::here("ms/Fig_1.pdf"), height = 4, width = 7, units = "in")
 ```
@@
 pca_plot
-ggsave(plot = pca_plot, filename = here::here("ms/fig2.pdf"), height = 6, width = 6, units = "in")
+ggsave(plot = pca_plot, filename = here::here("ms/Fig_2.pdf"), height = 6, width = 6, units = "in")
 ```
@@
 traits_with_tree
-ggsave(plot = traits_with_tree, filename = here::here("ms/fig3.pdf"), height = 8, width = 6, units = "in")
+ggsave(plot = traits_with_tree, filename = here::here("ms/Fig_3.pdf"), height = 8, width = 6, units = "in")
 ```
@@
 importance_heatmap
-ggsave(plot = importance_heatmap, filename = here::here("ms/fig4.pdf"), height = 4, width = 6, units = "in")
+ggsave(plot = importance_heatmap, filename = here::here("ms/Fig_4.pdf"), height = 4, width = 6, units = "in")
 ```
@@
 combined_cwm_plots
-ggsave(plot = combined_cwm_plots, filename = here::here("ms/fig5.pdf"), height = 4, width = 7, units = "in")
+ggsave(plot = combined_cwm_plots, filename = here::here("ms/Fig_5.pdf"), height = 4, width = 7, units = "in")
 ```
@@
 combined_comm_div_plots
-ggsave(plot = combined_comm_div_plots, filename = here::here("ms/fig6.pdf"), height = 8, width = 4, units = "in")
+ggsave(plot = combined_comm_div_plots, filename = here::here("ms/Fig_6.pdf"), height = 8, width = 4, units = "in")
 ```

### 2019-08-15 ms/manuscript.Rmd
Fix figure numbers in output

@@
 
-```{r traits-used-image, fig.height = 7, fig.width = 7, fig.cap = "Examples of traits used in this study."}
-knitr::include_graphics(here::here("images/Figure_3_1_traits_lowres.pdf"))
+```{r traits-used-image, fig.height = 9, fig.width = 6, fig.cap = "Examples of traits used in this study."}
+knitr::include_graphics(here::here("ms/Fig_1.pdf"))
 ```
@@
 climate_plot
-ggsave(plot = climate_plot, filename = here::here("ms/Fig_1.pdf"), height = 4, width = 7, units = "in")
+ggsave(plot = climate_plot, filename = here::here("ms/Fig_2.pdf"), height = 4, width = 7, units = "in")
 ```
@@
 pca_plot
-ggsave(plot = pca_plot, filename = here::here("ms/Fig_2.pdf"), height = 6, width = 6, units = "in")
+ggsave(plot = pca_plot, filename = here::here("ms/Fig_3.pdf"), height = 6, width = 6, units = "in")
 ```
@@
 traits_with_tree
-ggsave(plot = traits_with_tree, filename = here::here("ms/Fig_3.pdf"), height = 8, width = 6, units = "in")
+ggsave(plot = traits_with_tree, filename = here::here("ms/Fig_4.pdf"), height = 8, width = 6, units = "in")
 ```
@@
 importance_heatmap
-ggsave(plot = importance_heatmap, filename = here::here("ms/Fig_4.pdf"), height = 4, width = 6, units = "in")
+ggsave(plot = importance_heatmap, filename = here::here("ms/Fig_5.pdf"), height = 4, width = 6, units = "in")
 ```
@@
 combined_cwm_plots
-ggsave(plot = combined_cwm_plots, filename = here::here("ms/Fig_5.pdf"), height = 4, width = 7, units = "in")
+ggsave(plot = combined_cwm_plots, filename = here::here("ms/Fig_6.pdf"), height = 4, width = 7, units = "in")
 ```
@@
 combined_comm_div_plots
-ggsave(plot = combined_comm_div_plots, filename = here::here("ms/Fig_6.pdf"), height = 8, width = 4, units = "in")
+ggsave(plot = combined_comm_div_plots, filename = here::here("ms/Fig_7.pdf"), height = 8, width = 4, units = "in")
 ```

### 2019-08-15 ms/manuscript.Rmd
Add list of SI

@@
   figs_tabs = c(
-    "No. of figures:_6 (all in color)",
+    "No. of figures:_7 (all in color)",
     "No. of tables:_5",
-    "No. of supporting information files:_6 (Fig. S1; Table S1-S5)", "", "", "", "")
+    "No. of supporting information files:_7 (Fig. S1; Table S1-S6)", "", "", "", "")
 ) %>%
@@
 
+## Supporting Information
+
+Additional Supporting Information may be found online in the Supporting Information section at the end of the article:
+
+**Table S1** List of ferns from Moorea, French Polynesia included in this study.
+
+**Table S2** Trait values of ferns on Moorea, French Polynesia.
+
+**Table S3** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate.
+
+**Table S4** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+
+**Table S5** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+
+**Table S6** Best-fitting models explaining community diversity metrics by climate and growth habit.
+
+**Fig. S1** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
+
+\clearpage
+
 **\fig \@ref(fig:traits-used-image)**: Examples of traits used in this study.
-(**a**) Sporophyte traits.
+**(a)** Sporophyte traits.
 All sporophyte traits were quantitative, including frond length and width, stipe length, rhizome dia., number of pinna pairs (in this example, nine) and degree of lamina dissection (in this example, once-pinnate).
 Specific leaf area (ratio of area to mass of leaf lamina) not depicted.
-(**b**) Gametophyte traits.
+**(b)** Gametophyte traits.
 Morphotype was defined as a binary trait, either cordate or non-cordate.

### 2019-08-15 ms/SI.rmd
Add references, details to captions

@@
 nocite: | 
-  @NMNS2008, @Copeland1932, @FloraNSW2019, @Brownsey1987, @Murdock2003
+  @NMNS2008, @Copeland1932, @FloraNSW2019, @Brownsey1987, @Murdock2003, @Nayar1971, @Lloyd1980, @Zhang2008, @Bierhorst1967, @Martin2006, @Tigerschiold1989, @Tigerschiold1990, @Atkinson1975, @Chen2014a
 ---
@@
 
-**Table S1** List of ferns from Moorea, French Polynesia included in this study with taxonomic data.
+**Table S1** List of ferns from Moorea, French Polynesia included in this study.
 
-'taxon_code' is the name used for that taxon in all analyses.
-'informal_variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them.
-'name_source' indicates the source of the scientific name as follows, IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith, 2003.
+'taxon code' is the name used for that taxon in analyses.
+'informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them.
+'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith, 2003.
 
@@
 SLA, specific leaf area.
-For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen, etc.
+
+'taxon code' is the name used for that taxon in analyses (see Table S1).
+For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI.
 
@@
 **Table S5** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+
+LL, log likelihood.
+
 Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
@@
 Lower log likelihood for the dependent model indicates correlation with growth habit.
-Abbreviations: LL, log likelihood.
 
@@
 
-**Table S6** Best-fitting models explaining community diversity metrics by climate and growth habit. Models within delta AICc 3 of the best-fitting model shown.
+**Table S6** Best-fitting models explaining community diversity metrics by climate and growth habit.
 
 MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, *P* value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's *I*.
+
+Models within delta AICc 3 of the best-fitting model shown.
 Trait names refer to community-weighted mean values (see Methods).
@@
 
-**Fig.
-S1** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input: **a**, standard effect size of mean phylogenetic distance (MPD~func~) **b**, standard effect size of mean nearest taxon distance (MNTD~func~).
+**Fig. S1** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
+
+**(a)**, standard effect size of mean phylogenetic distance (MPD~func~) **(b)**, standard effect size of mean nearest taxon distance (MNTD~func~).
 For details on calculations of MPD~func~ and MNTD~func~, see Methods.
@@
 
-References
+**References**

### 2019-08-15 code/resolve_names.R
Fix typo in write_csv (missing object)

@@ -81,6 +81,7 @@ taxa_names_with_authors <- gnr_resolve(
     name_source == "Tropicos - Missouri Botanical Garden" ~ "TROPICOS",
     name_source == "The International Plant Names Index" ~ "IPNI",
     TRUE ~ name_source
-  ))
-
-write_csv(here("ms/Table_S1.csv"))
+  )) %>%
+  rename_all(~str_replace(., "_", " "))
+  
+write_csv(taxa_names_with_authors, here::here("ms/Table_S1.csv"))

### 2019-08-15 code/functions.R
Use "taxon code" as column name instead of "species"

@@ -409,7 +409,7 @@ process_trait_data_for_si <- function (sla_raw, morph_cont_raw, morph_qual_raw) 
   # Add measurment sources and reformat column names/order
   left_join(morph_table, meas_sources) %>%
     select(
-      species,
+      `taxon code` = species,
       `growth habit` = habit,
       morphotype,
       gemmae, 

### 2019-08-15 ms/manuscript.Rmd
Continue using image folder for now

@@
 ```{r traits-used-image, fig.height = 9, fig.width = 6, fig.cap = "Examples of traits used in this study."}
-knitr::include_graphics(here::here("ms/Fig_1.pdf"))
+knitr::include_graphics(here::here("images/Figure_3_1_traits_lowres.pdf"))
 ```

### 2019-08-28 ms/manuscript.Rmd
Use format_pval

@@
 ```{r corr-numbers}
-stipe_pval <- pic_results %>% filter(trait == "stipe") %>% pull(p.value) %>% signif(2)
-width_pval <- pic_results %>% filter(trait == "width") %>% pull(p.value) %>% signif(2)
-dissection_pval <- pic_results %>% filter(trait == "dissection") %>% pull(p.value) %>% signif(2)
-morphotype_pval <- correlated_evo_test %>% filter(trait == "morphotype") %>% pull(pval) %>% signif(2)
+stipe_pval <- pic_results %>% filter(trait == "stipe") %>% pull(p.value) %>% format_pval(., equals_sign = TRUE)
+width_pval <- pic_results %>% filter(trait == "width") %>% pull(p.value) %>% format_pval(., equals_sign = TRUE)
+dissection_pval <- pic_results %>% filter(trait == "dissection") %>% pull(p.value) %>% format_pval(., equals_sign = TRUE)
+morphotype_pval <- correlated_evo_test %>% filter(trait == "morphotype") %>% pull(pval) %>% format_pval(., equals_sign = TRUE)
 ```
@@
 Epiphytic species tend to have smaller values for all traits (\tab \@ref(tab:pics); contrasts were made by subtracting epiphytic from terrestrial values).
-Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval = `r stipe_pval`), as well as frond width (\pval = `r width_pval`) and degree of laminae dissection (\pval = `r dissection_pval`).
+Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval `r stipe_pval`), as well as frond width (\pval `r width_pval`) and degree of laminae dissection (\pval `r dissection_pval`).
 For the gametophyte traits, only morphotype (non-cordate vs.
-cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval = `r morphotype_pval`; \tab \@ref(tab:corr-evo)).
+cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval `r morphotype_pval`; \tab \@ref(tab:corr-evo)).
 Similar results were obtained using the "strict" dataset (\tab S5).

### 2019-08-28 ms/SI.rmd
Fix names of references in SI table 1

@@
 nocite: | 
-  @NMNS2008, @Copeland1932, @FloraNSW2019, @Brownsey1987, @Murdock2003, @Nayar1971, @Lloyd1980, @Zhang2008, @Bierhorst1967, @Martin2006, @Tigerschiold1989, @Tigerschiold1990, @Atkinson1975, @Chen2014a
+  @NMNS2008, @Copeland1932, @FloraNSW2019, @Brownsey1987, @Murdock2003, @Nayar1971, @Lloyd1980, @Zhang2008, @Bierhorst1967, @Martin2006, @Tigerschiold1989, @Tigerschiold1990, @Atkinson1975, @Palmer2003
 ---
@@
 
-Article title: Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns
+Article title: Constrained, yet diverse: Phylogenetic and morphological filtering in epiphytic ferns
 
@@
 
+Continuous measurements shown as mean, standard deviation (sample size).
 'taxon code' is the name used for that taxon in analyses (see Table S1).

### 2019-08-28 code/functions.R

@@ -372,7 +372,7 @@ process_trait_data_for_si <- function (sla_raw, morph_cont_raw, morph_qual_raw) 
     # Add column for mean +/- sd (n) formatted for printing
     mutate(
       print = case_when(
-        n > 1 ~ glue("{mean} ± {sd} ({n})"),
+        n > 1 ~ glue("{mean}, {sd} ({n})"),
         n == 1 ~ glue("{mean} ({n})")
       )
     ) %>%
@@ -384,18 +384,26 @@ process_trait_data_for_si <- function (sla_raw, morph_cont_raw, morph_qual_raw) 
   morph_table <- full_join(morph_mean, morph_qual_raw) %>%
     mutate(gameto_source = jntools::paste3(source_1, source_2) %>%
              str_replace(" ", ", ")) %>%
-    select(-source_1, -source_2)
+    select(-source_1, -source_2) %>%
+    mutate(gameto_source = str_replace_all(gameto_source, "Lloyd1980", "Lloyd (1980)")) %>%
+    mutate(gameto_source = str_replace_all(gameto_source, "Atkinson1975", "Atkinson (1975)")) %>%
+    mutate(gameto_source = str_replace_all(gameto_source, "Nayar1971", "Nayar and Kaur (1971)")) %>%
+    mutate(gameto_source = str_replace_all(gameto_source, "Tigerschiold1990", "Tigerschiold (1990)")) %>%
+    mutate(gameto_source = str_replace_all(gameto_source, "Tigerschiold1989", "Tigerschiold (1989)")) %>%
+    mutate(gameto_source = str_replace_all(gameto_source, "Martin2006", "Martin et al (2006)")) %>%
+    mutate(gameto_source = str_replace_all(gameto_source, "Zhang2008", "Zhang et al (2008)")) %>%
+    mutate(gameto_source = str_replace_all(gameto_source, "Bierhorst1967", "Bierhorst (1967)"))
   
   # Add in sources for measurements
   meas_sources <- morph_cont_raw %>% 
     # Rename sources according to abbreviations
     mutate(
       source = case_when(
-        source == "Ferns & Fern-Allies of South Pacific Islands" ~ "NMNS2008", 
-        source == "Pteridophytes of the Society Islands" ~ "Copeland1932",
-        source == "Hawaii's Ferns and Fern Allies" ~ "Palmer2003",
-        source == "Flora of New South Wales" ~ "FloraNSW2019",
[…]

### 2019-08-28 code/process_refs.R

@@ -1,7 +1,27 @@
 # Make clean bib files for each Rmd file that only includes cited references
 
 library(jntools)
+library(tidyverse)
 
-make_ref_list("ms/manuscript.Rmd", "ms/references_raw.bib", "ms/references.bib")
+make_ref_list(
+  rmd_file = "ms/manuscript.Rmd", 
+  raw_bib = "ms/references_raw.bib",
+  final_bib = "ms/references.bib")
 
-make_ref_list("ms/SI.Rmd", "ms/references_raw.bib", "ms/si_references.bib")
+make_ref_list(
+  rmd_file = "ms/SI.Rmd", 
+  raw_bib = "ms/references_raw.bib", 
+  final_bib = "ms/si_references.bib"
+)
+
+# Make some manual fixes to authors in SI bibliography
+# (these are institutions, so need double brackets to
+# avoid latex thinking they have first and last names)
+read_lines("ms/si_references.bib") %>%
+  str_replace(
+    "National Museum of Nature and Science",
+    "\\{National Museum of Nature and Science\\}") %>%
+  str_replace(
+    "Royal Botanic Gardens and Domain Trust",
+    "\\{Royal Botanic Gardens and Domain Trust\\}") %>%
+  write_lines("ms/si_references.bib")

### 2019-09-30 ms/SI.rmd
Use explicit cache
called canopy_cache

@@
 # Load analysis results
-loadd(trait_data_for_si)
-loadd(climate_ancova_results)
-loadd(phylosig_binary_traits_strict)
-loadd(correlated_evo_test_strict)
-loadd(best_fit_div_models_moran)
-loadd(func_comm_struc_combined)
+loadd(trait_data_for_si, cache = canopy_cache)
+loadd(climate_ancova_results, cache = canopy_cache)
+loadd(phylosig_binary_traits_strict, cache = canopy_cache)
+loadd(correlated_evo_test_strict, cache = canopy_cache)
+loadd(best_fit_div_models_moran, cache = canopy_cache)
+loadd(func_comm_struc_combined, cache = canopy_cache)
 

### 2019-09-30 ms/manuscript.Rmd
Use explicit cache
called canopy_cache

@@
 # Load analysis results
-loadd(best_fit_div_models)
-loadd(climate_ancova_results)
-loadd(climate_plot)
-loadd(climate_select)
-loadd(combined_comm_div_plots)
-loadd(combined_cwm_plots)
-loadd(comm)
-loadd(correlated_evo_test)
-loadd(correlated_evo_test_gemmae)
-loadd(daily_mean_climate)
-loadd(div_el_model_summaries)
-loadd(div_t_test_results)
-loadd(div_metrics_all)
-loadd(fern_traits)
-loadd(importance_heatmap)
-loadd(pca_plot)
-loadd(pca_results)
-loadd(phy)
-loadd(phylosig_binary_traits)
-loadd(phylosig_cont_traits)
-loadd(pic_results)
-loadd(species_list)
-loadd(traits_with_tree)
+loadd(list = c(
+  "best_fit_div_models",
+  "climate_ancova_results",
+  "climate_plot",
+  "climate_select",
+  "combined_comm_div_plots",
+  "combined_cwm_plots",
+  "comm",
+  "correlated_evo_test",
+  "correlated_evo_test_gemmae",
+  "daily_mean_climate",
+  "div_el_model_summaries",
+  "div_t_test_results",
+  "div_metrics_all",
+  "fern_traits",
+  "importance_heatmap",
+  "pca_plot",
+  "pca_results",
+  "phy",
+  "phylosig_binary_traits",
+  "phylosig_cont_traits",
+  "pic_results",
+  "species_list",
+  "traits_with_tree"),
+  cache = canopy_cache)
 ```

### 2019-09-30 make.R

@@ -19,8 +19,11 @@ pkgconfig::set_config("drake::strings_in_dots" = "literals")
 source("code/functions.R")
 source("code/plan.R")
 
+# Load cache
+canopy_cache <- new_cache("canopy_cache")
+
 # Set seed for reproducibility.
 set.seed(1954)
 
 # Make plan.
-make(plan)
+make(plan, cache = canopy_cache)

### 2019-10-18 ms/manuscript.Rmd
Revise methods
Fix plot vs. site, spelling errors, remove some redundant wording ("datalogger")

@@
 Tohiea \SI{393}{\meter} plot malfunctioned and was missing data for a substantial part of the final survey period (March 12--July 5, 2014), so we used data for the same dates from the preliminary survey period for this datalogger instead.
-Three other dataloggers also failed early during the final survey period, but lacked data from the preliminary period, so were excluded from analysis.
-Other days missing data for any datalogger during the final survey period were excluded from the dataset.
-The final dataset included `r total_days` days of data for `r total_sites %>% filter(habit == "epiphytic") %>% pull(n)` epiphytic dataloggers and `r total_sites %>% filter(habit == "terrestrial") %>% pull(n)` terrestrial dataloggers.
+Three others also failed early during the final survey period, but lacked data from the preliminary period, so were excluded from analysis.
+Other days missing data during the final survey period were excluded from the dataset.
+The final dataset included `r total_days` days of data for `r total_sites %>% filter(habit == "epiphytic") %>% pull(n)` dataloggers and `r total_sites %>% filter(habit == "terrestrial") %>% pull(n)` terrestrial dataloggers.
 We calculated vapor pressure deficit (VPD) from relative humidity and temperature using the \package{plantecophys} R package [@Duursma2015].
@@
 
-Morphological traits of sporophytes were measured from herbarium specimens for most species, or obtained from the literature when plant material was unavailable.
+Morphological traits of sporophytes were measured from herbarium specimens, or obtained from the literature when plant material was unavailable.
 Measurements were made directly on plants in the field for a few species with large fronds.
@@
 For species with \textless \SI{2}{\milli\meter} between primary veins, leaf fragments excluding primary veins were obtained by dissection, dried at \SI{60}{\celsius} for \SI{24}{\hour}, weighed, and scanned.
-ImageJ [@Abramoff2004] software was used to measure area per fragment, and SLA calculated in the same way as leaf punches.
+ImageJ [@Abramoff2004] software was used to measure area per fragment.
 
@@
 
-We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each site, then overall means from the daily values.
+We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each datalogger, then overall means from the daily values.
 We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (hereafter, "climatic variables").
@@
 We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone and their interactions, then selecting the one with the lowest corrected Akaike Information Criterion (AIC~c~) [@Akaike1973].
-A single outlier site with unusually high temperature and low humidity (Mt.
-Rotui, 830 m) site was excluded from ANCOVA and linear models.
+A single outlier plot with unusually high temperature and low humidity (Mt.
+Rotui, 830 m plot) was excluded from ANCOVA and linear models.
 Mt.
@@
 For phylogenetic distances, we used the tree of @Nitta2017.
-For functional distances, we log-transformed and scaled traits as in the PCA, excluded frond width and length (both frond size traits highly correlated with stipe length), then calculated the distance matrix using Gower's dissimilarity coeffecient [@Gower1971] with the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
+For functional distances, we log-transformed and scaled traits as in the PCA, excluded frond width and length (both frond size traits highly correlated with stipe length), then calculated the distance matrix using Gower's dissimilarity coefficient [@Gower1971] with the \function{gowdis} function in the R package \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
 To test for the effect of different trait selection schemes, we compared the results of running the same analyses for various sets of traits with and without log-transformation (Supporting Information).
@@
 We tested the effects of climate and growth habit on diversity metrics using full-subsets multiple regression with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
-For each diversity metric, we built a set of models (Gaussian family, with site as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
+For each diversity metric, we built a set of models (Gaussian family, with plot as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
 Our dataset included `r nrow(div_metrics_all) - nrow(climate_select)` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.

### 2019-10-18 ms/manuscript.Rmd
Update MS
Restructure discussion

@@
 
-Here, we show for the first time using a broadly sampled phylogenetic dataset that multiple morphological traits related to epiphytic growth are evolutionarily conserved in ferns.
-We find that epiphytic ferns are constrained by size, having consistently smaller fronds and rhizomes than terrerstial species.
-However, contrary to our hypothesis, epiphytic fern communities appear to be more diverse than terrestrial communities.
-This suggests that environmental filtering plays a strong role in limiting plant size, while allowing for diversification across other trait axes in epiphytic ferns.
+Here, we provide the first study to our knowledge to combine comparative methods with phylogenetic community ecology to analyze the evolutionary implications of epiphytic growth in ferns.
+We find that epiphytic ferns are constrained in size, having consistently smaller fronds and rhizomes than terrestrial species.
+In contrast, epiphytic ferns vary greatly in other traits, leading epiphytic fern communities to be more functionally diverse overall than terrestrial communities.
+Taken together, our results suggest that environmental filtering plays a strong role in limiting plant size, while simultaneously allowing for diversification across other trait axes in epiphytic ferns.
 
-### Effect of Phylogeny on Analysis of Epiphytic Traits
-
-Although several other studies have addressed the relationship of functional traits with growth habit in ferns [e.g., @Hietz1998; @Kluge2007; @WatkinsJr.2010], few have accounted for evolutionary history.
-We find that growth habit shows phylogenetic signal, as do several of the traits associated with it.
-In one of the few previous studies to analyze phylogenetic signal in functional traits of ferns, @Zhang2014b found that only one of 16 traits showed significant phylogenetic signal.
-The only trait in common between the current study and that of @Zhang2014b is SLA, which did not show phylogenetic signal in either study.
-Given the high degree of phylogenetic signal we observe in most traits and that at least some of the traits between our two studies are likely to be correlated, it is surprising that @Zhang2014b did not detect phylogenetic signal in more traits.
-However, @Zhang2014b only used \K, which tended to indicate less phylogenetic signal than $\lambda$ in our study.
-In simulation studies, $\lambda$ has lower error (both Type I and Type II) than \K, which can be inconsistent unless the strength of BM is great [@Munkemuller2012].
-Our results indicate the importance of including a phylogenetic component, and including multiple measures of phylogenetic signal, when analyzing traits related to epiphytic growth in ferns.
-
-### Sporophyte Traits
+### Convergence in small size and non-cordate morphotype
 
@@
 Our results support previous non-phylogenetically controlled studies that found a prevalence of short stipes in epiphytic ferns [@WatkinsJr.2010; @Creese2011].
-However, given that these were all correlated with each other in our study, and similar traits have been shown to scale together previously [@Arcand2008; @Creese2011], careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
+However, given that stipe and frond length, frond width, and rhizome diameter were correlated, and similar traits have been shown to scale together previously [@Arcand2008; @Creese2011], careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
 
-```{r sla-numbers, results = "hide"}
-# First smallest SLA: Humata_anderssonii
-smallest_sla_1 <- fern_traits %>% arrange(sla) %>% slice(1) %>% pull(sla) %>% signif(3)
-assert_that(fern_traits %>% arrange(sla) %>% slice(1) %>% pull(species) == "Humata_anderssonii")
+Another clear pattern to emerge from our study is the association between epiphytism and non-cordate (\eg, elongate or filamentous) morphotype.
+Morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived ($< \SI{1}{\year}$), cordate gametophytes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; @Farrar2008].
+The complex, three-dimensional structure of non-cordate gametophytes slows their drying rates, and may be an adaptation to drought-prone epiphytic habitats [@WatkinsJr.2007c; @Pittermann2013], and promote out-crossing [@Farrar2008].
+Our study is the first to our knowledge to demonstrate that the correlation between non-cordate morphology and epiphytism in ferns is significant while controlling for phylogeny.
+@Sundue2015 found that non-cordate morphology is significantly correlated with increased diversification rates within Polypodiaceae, consistent with the hypothesis that this trait promoted diversification within the epiphytic niche.
 
-# Second smallest SLA: Ctenopterella_blechnoides
-smallest_sla_2 <- fern_traits %>% arrange(sla) %>% slice(2) %>% pull(sla) %>% signif(3)
-assert_that(fern_traits %>% arrange(sla) %>% slice(2) %>% pull(species) == "Ctenopterella_blechnoides")
-```
+### Remarkable diversity in other traits
 
-In contrast to their clearly smaller size, epiphytes were remarkably diverse in other traits.
-Epiphytes tended to have less dissected fronds with fewer pinna than terrestrial species (\tab \@ref(tab:pics)), but the range of values for these traits was quite wide at the community level (error bars in \fig \@ref(fig:cwm-div)).
+In contrast to their clearly smaller size and non-cordate morphotypes, epiphytes were remarkably diverse in other traits.
+Epiphytes tended to have less dissected fronds with fewer pinnae than terrestrial species (\tab \@ref(tab:pics)), but the range of values for these traits was quite wide at the community level (error bars in \fig \@ref(fig:cwm-div)).
+It is likely that the high diversity of non-size related traits we observed reflects different physiological adaptations to various canopy niches.
+For example, filmy ferns (family Hymenophyllaceae), which depend on desiccation tolerance to survive dry conditions, have vastly different morphology from bird's nest ferns (*Asplenium nidus* and others), which capture detritus in their rosette leaves.
+
 Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
@@
 We expected that epiphytes would have lower SLA, given their more extreme growing conditions. 
-However, the parallel relationship in CWM values of SLA with elevation in both epiphytes and terrestrial species (\fig \@ref(fig:cwm-div)) indicates that SLA may be responding more strongly to temperature in both epiphytes and terrestrial species, while only being weakly modified by humidity.
+The lack of a clear difference in SLA between epiphytes and terrestrial species observed here may be an effect of scale: the LES was defined at regional to global scales, but tends to break down at local scales encompassing less environmental variation [@Messier2017].
 
-### Gametophyte Traits
-
-Only gametophyte morphotype was found to have a significant association with epiphytic growth.
-Morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived ($< \SI{1}{\year}$), cordate gametophytes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate (\eg, elongate or filamentous) gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; @Farrar2008].
-The complex, three-dimensional structure of non-cordate gametophytes slows their drying rates, and may be an adaptation to drought-prone epiphytic habitats [@WatkinsJr.2007c; @Pittermann2013], and promote out-crossing [@Farrar2008].
-Our study is the first to our knowledge to demonstrate that the correlation between non-cordate morphology and epiphytism in ferns is significant while controlling for phylogeny.
-@Sundue2015 found that non-cordate morphology is significantly correlated with increased diversification rates within Polypodiaceae, consistent with the hypothesis that this trait promoted diversification within the epiphytic niche.
-
-We did not find that gemmae production was linked with epiphytism.
+We did not find that gemmae production in gametophytes was linked with epiphytism. 
 Gemmae are asexual propagules that allow gametophytes to persist over multiple growing seasons and attain large population sizes [@Farrar2008].
@@
 
-Presence of hairs and glands has strong phylogenetic signal (\tab \@ref(tab:phylosig-binary)), and these traits are correlated with certain taxonomic groups [@Stokey1951; @Nayar1971].
+Presence of hairs and glands in gametophytes has very strong phylogenetic signal (\tab \@ref(tab:phylosig-binary)), and these traits are associated with certain taxonomic groups [@Stokey1951; @Nayar1971].
 Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], we did not detect a correlation between hair production and epiphytic growth.
@@
 
-### Environmental Filtering
+### The role of environmental filtering in the canopy
 
@@
 
-We obtained partial support for our hypothesis that epiphytic fern communities would be more phylogenetically clustered than terrestrial ones; this was observed for MNTD, but not MPD (\fig \@ref(fig:comm-div)).
+We obtained partial support for the hypothesis that epiphytic fern communities would be more phylogenetically clustered than terrestrial ones; this was observed for MNTD, but not MPD (\fig \@ref(fig:comm-div)).
 This makes sense given that most epiphytes have radiated within five major clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely-related species from two or more epiphytic clades, but has a similar overall level of phylogentic diversity as terrestrial communities.
@@
 However, this result is dependent on the traits used to measure functional diversity; when the same analyses were done using only size-related traits (scaled, untransformed stipe length, frond length and width, and rhizome diameter), the opposite result was observed (\fig S1).
-Futhermore, terrestrial species show a much wider range of values for size traits (\fig \@ref(fig:cwm-div)), and neither functional group appears to be restricted to a smaller area in trait space (\fig \@ref(fig:pca)).
+Furthermore, terrestrial species show a much wider range of values for size traits (\fig \@ref(fig:cwm-div)), and neither functional group appears to be restricted to a smaller area in trait space (\fig \@ref(fig:pca)).
 This suggests that epiphytic ferns are subject to environmental filtering for size-related traits, while simultaneously diversifying along other trait axes.
@@
 The vertical stratification of epiphytic species has been well documented in ferns [\eg, @Parra2009; @WatkinsJr.2006a; @Acebey2017] and broader studies [\eg, @Kromer2007; @Zotz2007].
-The goal of our study was to quantify differences between epiphytic and terrestrial fern communities; future research should focus on bringing comparative phylogenetic and community ecological methods to bear on the large body of work characterizing differences within epiphytic microhabitats.
+Future research should focus on bringing comparative phylogenetic and community ecological methods to bear on the large body of work characterizing differences within epiphytic microhabitats.
 
-### Conclusion and Future Directions
+### Conclusion and future directions
 
@@
 Finally, we observed drier conditions in the canopy relative to terrestrial habitats. 
-Taken together, our results support a scenario of environmental filtering acting to limit size and phylogenetic diversity in epiphytic ferns, while simulataneously enabling diversification along other trait axes.
+Taken together, our results support a scenario of environmental filtering acting to limit size and phylogenetic diversity in epiphytic ferns, while simultaneously enabling diversification along other trait axes.
 The generality of this pattern could be further tested by investigating the functional traits of other groups with large epiphytic radiations such as orchids and bromeliads.
@@
 Recent studies [e.g., @Sundue2015; @Testo2016a; @Lehtonen2017; @Testo2018a] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
-Epiphytism itself is likely a complex trait requiring the evolution of multiple associated traits, and the timing and order of the evolution of such traits may be related to the complex diversification dynamics observed in these studies.
+Epiphytism itself is a complex trait requiring the evolution of multiple associated traits, and the timing and order of the evolution of such traits may be related to the complex diversification dynamics observed in these studies.
 Small size seems to be a prerequisite for epiphytic growth, as indicated both by our results and the global study of @Testo2018a.
-Here, we identify several traits in addition to plant size (\eg, pinna number, frond dissection, gametophyte morphotype) that may also belong to the suite of epiphytic traits.
+Here, we have shown that size appears to be the exception, not the rule, when it comes to trait diversity of epiphytes: although epiphytes are universally small, our data suggest they are diverse in most other traits.
+This emphasizes that fern epiphytes should not be treated as a monolithic group, but rather encompass a wide range of adaptive forms.
 We hope future studies will build on our results by analyzing traits involved in epiphytic growth other than body size across densely sampled phylogenies to dissect how each may relate to diversification and growth habit.
-Additionally, future studies should focus on determining in more detail how niche space is partitioned within epiphytic vs.
-terrestrial habitats, which is clearly a major dividing line for ferns.
+Additionally, future studies should focus on determining in more detail how niche space is partitioned within epiphytic vs. terrestrial habitats, which is clearly a major dividing line for ferns.
 

### 2019-10-23 ms/SI.rmd
Add Fig S2 showing community-wide SD values

@@
     pandoc_args: [ "--csl", "new-phytologist.csl"]
+    includes:
+      in_header: newphyt.sty
 nocite: | 
@@
 
-```{r fig-s1}
+```{r fig-s2}
 library(corrr)
@@
 
+**Fig.
+S2** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
+Epiphytic communities in green, terrestrial communities in brown.
+Points randomly jittered along x-axis to reduce overlap.
+Boxplots show median values (bold lines) by growth habit.
+Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 $\times$ the interquartile range.
+Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+
+**(a)**, traits related to plant size **(b)**, traits not related to plant size.
+
+```{r fig-s1}
+loadd(cwsd_scatterplot, cache = canopy_cache)
+ggsave(plot = cwsd_scatterplot, filename = here::here("ms/Fig_S2.pdf"), height = 6, width = 7, units = "in")
+```
+
 **References**

### 2019-10-23 code/functions.R

@@ -1973,7 +1973,8 @@ make_div_scatterplot <- function (yval, xval = "el", ylab = yval, xlab = "Elevat
 #' )
 make_boxplot <- function (yval, ylab = yval, xlab = "Growth habit",
                           data, summaries, 
-                          habit_colors) {
+                          habit_colors,
+                          include_points = FALSE) {
   
   yval_sym <- sym(yval)
   
@@ -1994,8 +1995,15 @@ make_boxplot <- function (yval, ylab = yval, xlab = "Growth habit",
   # Make basic boxplot by habit
   # Add asterisk for t-test result between the bars only if they
   # differ by habit (will print nothing if not signif.)
-  ggplot(data, aes(x = habit)) +
-    geom_boxplot(aes(y = !!yval_sym, color = habit)) +
+  plot <- ggplot(data, aes(x = habit))
+  
+  if(isTRUE(include_points)) {
+    plot <- plot +
+      geom_jitter(aes(y = !!yval_sym, color = habit))
+  }
+  
+  plot +
+    geom_boxplot(aes(y = !!yval_sym, color = habit), fill = "transparent") +
     annotate(
       "text", label = asterisk, size = 5, fontface = "bold",
       x = 1.5, 
@@ -2851,6 +2859,97 @@ combine_comm_div_plots <- function (scatterplots, boxplots) {
     # Tweak margins to remove whitespace between plots
     plot.margin = margin(t = -0.2, r = -0.1, b = 0, l = -0.1, unit = "in")
   )
+}
+
+
+#' Make a composite plot of community-wide
+#' standard deviations of quantitative trait values
+#' 
+#' Includes results of t-test comparing epiphytic vs. terrestrial
+#' SD values.
+#'
+#' @param cwm_long Community-wide mean and standard deviation
+#' in trait values. Output of calculate_cwm()
+#' @param habit_colors Vector of colors to use for growth habit
+#'
+#' @return Plot
+#' 
+make_cwsd_scatterplot <- function(cwm_long, habit_colors) {
+  
+  # Make vector of quantitative traits in the order they will be plotted
+  quant_traits <- c("stipe", "length", "width", "rhizome", 
+                    "sla", "pinna", "dissection") %>%
[…]

### 2019-10-23 code/plan.R

@@ -485,6 +485,10 @@ plan <- drake_plan(
 
   # Make heatmap of importance scores.
   importance_heatmap = make_heatmap(important_div_vars),
+  
+  # Make jitter plus box plot of community-wide SD values
+  # for SI
+  cwsd_scatterplot = make_cwsd_scatterplot(cwm_long, habit_colors),
 
   # Manuscript ----
   ms = rmarkdown::render(

### 2019-10-24 ms/manuscript.Rmd
Update references

@@
 nocite: | 
-  @Vogel1968, @Wollenweber2000
+  @Vogel1968
 ---
@@
 Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife.
-Raw and processed data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED [@Nitta2019].
+Raw and processed data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED [@Nitta2019a].
 
@@
 Presence of hairs and glands in gametophytes has very strong phylogenetic signal (\tab \@ref(tab:phylosig-binary)), and these traits are associated with certain taxonomic groups [@Stokey1951; @Nayar1971].
-Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], we did not detect a correlation between hair production and epiphytic growth.
-Similarly, production of glands does not appear to be correlated with epiphytic growth.
+Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], and glands may affect rates of osmosis by modifying lipid concentrations [@Crow2011], we did not detect a correlation between hair or gland production and epiphytic growth.
 Wax-secreting glands are prevalent in both sporophytes and gametophytes of many notholaenid ferns, a xeric-adapted clade, but their functional significance has not been established [@Johnson2012].

### 2019-10-24 code/process_refs.R

@@ -25,3 +25,12 @@ read_lines("ms/si_references.bib") %>%
     "Royal Botanic Gardens and Domain Trust",
     "\\{Royal Botanic Gardens and Domain Trust\\}") %>%
   write_lines("ms/si_references.bib")
+
+# Make some manual fixes to authors in SI bibliography
+# (these are institutions, so need double brackets to
+# avoid latex thinking they have first and last names)
+read_lines("ms/references.bib") %>%
+  str_replace(
+    "Pteridophyte Phylogeny Group I",
+    "\\{Pteridophyte Phylogeny Group I\\}") %>%
+  write_lines("ms/references.bib")

### 2019-10-27 ms/manuscript.Rmd
Update MS
Finalize title and authors, summary, add Fig S2

@@
 ---
-title: "Constrained, yet diverse: Phylogenetic and morphological filtering in epiphytic ferns"
+title: "Life in the canopy: community trait assessments reveal substantial functional diversity among fern epiphytes"
 bibliography: "references.bib"
@@
 
-Joel H. Nitta^1,4^, James E. Watkins, Jr.^2^, Ravahere Taputuarai^3^, and Charles C. Davis^1^
+Joel H. Nitta^1,3^, James E. Watkins, Jr.^2^, and Charles C. Davis^1^
 
@@
 
-^3^Association Te Rau Ati Ati a Tau a Hiti Noa Tu, B.P. 11553, Mahina, Tahiti, French Polynesia
+^3^Current affiliation: Department of Botany, National Museum of Natural History, Smithsonian Institution, Washington, D.C., 20013, U.S.A.
 
-^4^Current affiliation: Department of Botany, National Museum of Natural History, Smithsonian Institute, Washington, D.C., 20013, U.S.A.
+Authors for correspondence:  
+*Joel H. Nitta*  
+*Tel: +1 2026331000*  
+*Email: joelnitta@gmail.com*
 
-Author for correspondence:  
-*Joel H. Nitta*  
+*Charles C. Davis*  
 *Tel: +1 6174960515*  
-*Email: joelnitta@gmail.com*
+*Email: cdavis@oeb.harvard.edu*
 
@@
 
-- The appearance of angiosperm-dominated forests in the Cretaceous and early Cenozoic had a profound effect on terrestrial biota by creating novel ecological niches.
-The diversity of modern fern lineages is thought to be a response to angiosperm dominance, radiating both in the canopy and on the forest floor.
-Despite the significance of epiphytic growth for fern evolution, however, few studies have investigated traits related to epiphytism in ferns while accounting for phylogeny.
+- The expansion of angiosperm-dominated forests in the Cretaceous and early Cenozoic had a profound effect on terrestrial biota by creating novel ecological niches.
+The majority of modern fern lineages are hypothesized to have arisen in response to this expansion, particularly fern epiphytes that radiated into the canopy.
+Recent evidence, however, suggests that epiphytism does not correlate with increased diversification rates in ferns, calling into question the role of the canopy habitat in their diversification.
 
-- We investigated evolutionary patterns in traits related to epiphytic growth at the species and community level using a phylogeny including representatives of multiple epiphytic fern radiations on the island of Moorea, French Polynesia.
+- To bridge this impasse, we investigated functional traits of fern sporophytes and gametophytes at the species and community level across a broad phylogenetic sampling on the island of Moorea, French Polynesia, including > 120 species and representatives of five major epiphytic radiations.
 
-- We observed phylogenetic conservatism in multiple traits related to epiphytic growth in fern sporophytes and gametophytes.
-Epiphytes had significantly shorter fronds, smaller rhizomes, and a higher frequency of non-cordate gametophytes.
-Epiphytic communities were clustered phylogenetically but diverse functionally, indicating that the canopy acts as a filter due to its low humidity to limit size while simultaneously enabling diversification into finely divided niches.
+- While epiphytes showed convergence in small size and a higher frequency of non-cordate gametophytes, they showed greater functional diversity at the community level relative to terrestrial ferns.
 
-- Our results provide an important phylogenetic perspective on the role of functional traits in ferns, and pave the way for more detailed studies examining how niche space is filled within terrestrial and epiphytic habitats.
+- These results suggest previously overlooked functional diversity among fern epiphytes, and raise the hypothesis that while the angiosperm canopy acted as a complex filter that restricted plant size, it also facilitated diversification into finely partitioned niches.
+Characterizing these niche axes and adaptations of epiphytic ferns occupying them should be a priority for future pteridological research.
 
@@
 
+**Fig. S2** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
+
 \clearpage

### 2019-10-27 ms/manuscript.Rmd
Split intro into one sentence per line
Update MS
Revise intro

@@
 
-The appearance of angiosperm-dominated forests in the late Cretaceous and early Cenozoic had an immense effect on terrestrial ecosystems [@Niklas1983; @Kenrick1997a; @Lloyd2008].
+The appearance of angiosperm-dominated forests in the late Cretaceous and early Cenozoic fundamentally changed climates and had an immense effect on terrestrial ecosystems, including increased global precipitation and the creation of novel niches [@Niklas1983; @Kenrick1997a; @Lloyd2008; @Boyce2010].
 The heterogeneous niche space of the angiosperm canopy comprising stratified light and humidity levels and a range of substrates for growth became rapidly filled by a variety of epiphytic plants, including mosses [@Fiz-Palacios2011], leafy liverworts [@Feldberg2014], orchids [@Givnish2015b], bromeliads [@Givnish2014a], and ferns [@Schneider2004d].
-Diversification rates in some of these groups may have further increased during the Paleocene / Eocene thermal maximum, a period of warmer and wetter climates that may have enabled movement into the canopy [@Schuettpelz2009; @Testo2016a].
-The transition to epiphytic growth has been particularly important for fern diversity, with ca.
-29% of all species exhibiting this growth form relative to ca.
-10% across all vascular plants [@Kress1986].
-Major transitions to epiphytism occurred at least five times during the evolution of ferns, each followed by diversification during the late Cretaceous or early Cenozoic [@Schuettpelz2009; @WatkinsJr.2012].
-Recent studies [@Sundue2015; @Testo2016a; @Lehtonen2017] have failed to find evidence of rapid diversification characteristic of adaptive radiations associated with epiphytic growth in ferns, but have confirmed gradually increasing rates of diversification in Polypodiaceae [@Sundue2015] and generally support the significance of epiphytic growth as a key innovation for extant fern diversity (@Testo2016a; @WatkinsJr.2012; but see @Lehtonen2017).
+The transition to epiphytic growth was particularly important for modern fern diversification: ca. 29% of all fern species are epiphytes vs. ca. 10% across all vascular plants [@Kress1986].
+Moreover, major transitions to epiphytism occurred at least five times within ferns, each followed by diversification during the late Cretaceous or early Cenozoic [@Schuettpelz2009; @WatkinsJr.2012].
+This has led to the hypothesis that epiphytism was a key factor contributing to extant fern diversification [@Schneider2004d; @Schuettpelz2009; @WatkinsJr.2012].
+Recent studies exploring this idea [@Sundue2015; @Testo2016a; @Lehtonen2017], however, have largely failed to identify evidence of rapid diversification associated with epiphytism in ferns. 
+This raises the possibility that modern fern diversification was not shaped by the evolution of epiphytism *per se*, but rather by the exploitation of various niches within the canopy and on the forest floor.
 
-Despite the ecological opportunities afforded by life in the canopy, including escape from disturbance [@Page2002] and abundant light for photosynthesis [@Benzing2000], epiphytes must contend with greater stress in the form of increased evaporation and extreme oscillations in temperature relative to terrestrial plants [@Zotz2001; @Lowman2012].
-Associated with this change, epiphytic ferns have evolved various morphological adaptations that minimize water loss from their leaves, including shorter stipes [@WatkinsJr.2010], thicker leaves [@Kluge2007; @WatkinsJr.2007b], and protection with scales or hairs [@WatkinsJr.2006b].
-Some lineages have specialized adaptations to epiphytic growth, such as humus-collection (e.g., *Asplenium*; @Benzing1990, *Drynaria*; @Janssen2005), vegetative desiccation tolerance (DT) (e.g., Hymenophyllaceae; @Shreve1911, *Pleopeltis*; @Pessin1925), CAM photosynthesis (e.g., *Pyrrosia*; @Wong1976), and symbiotic relationships with ants (e.g., *Antrophyum*; @WatkinsJr.2008, *Lecanopteris*; @Gay1993a).
-Many of these same adaptations are found in epiphytic seed plants [@Benzing1990].
-However, the evolution of epiphytism in ferns was more complex given that ferns rely on a free-living gametophyte [@Tsusumi2013].
-The fern gametophyte is anatomically simple, and differs from the sporophyte in anatomy, morphology, and physiology.
-Such differences create a system of 'dual ecology' where selective pressures acting on one stage are different than those acting on another.
-For example, DT is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is rare in sporophytes [@WatkinsJr.2007c].
-Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no studies to our knowledge have analyzed the evolution of such traits across the fern phylogeny including both life stages and representatives of all five Cretaceous epiphytic radiations *sensu* [@Schuettpelz2009].
+To better understand the role of the canopy in fern diversification, one obvious direction is to explore putatively adaptive fern epiphyte traits in a phylogenetic context. 
+In contrast to most terrestrial plants, epiphytes must contend with greater stress in the form of nutrient limitation, increased evaporative demand, drought, and extreme oscillations in temperature  [@Zotz2001; @WatkinsJr.2009; @Lowman2012].
+Previous studies have identified multiple traits as likely adaptations to these conditions, albeit many without consideration of phylogeny. 
+Epiphytic ferns tend to have morphological features that minimize water loss from their leaves, including shorter stipes [@WatkinsJr.2010], thicker leaves [@Kluge2007; @WatkinsJr.2007b], and protection with scales or hairs [@WatkinsJr.2006b].
+Additional specialized adaptations to epiphytic growth include humus collection (e.g., *Asplenium*; @Benzing1990, *Drynaria*; @Janssen2005), vegetative desiccation tolerance (DT) (e.g., Hymenophyllaceae; @Shreve1911, *Pleopeltis*; @Pessin1925), CAM photosynthesis (e.g., *Pyrrosia*; @Wong1976), and symbiotic relationships with ants (e.g., *Antrophyum*; @WatkinsJr.2008, *Lecanopteris*; @Gay1993a).
 
-Most studies investigating adaptations to epiphytism have utilized a univariate, cross-species approach, analyzing traits individually between terrestrial and epiphytic species [e.g., @WatkinsJr.2007b; @WatkinsJr.2010].
-However, given multiple radiations into the epiphytic niche [@Schuettpelz2009], it is likely that traits related to epiphytism carry at least some degree of phylogenetic signal (i.e., the tendency for closely related species to resemble each other).
-Therefore, a comparative approach incorporating phylogenetic relationships (and if possible, branch lengths) is appropriate [@Felsenstein1985].
-However, such approaches have only recently been applied to studies of epiphytism in ferns, and at relatively limited phylogenetic scales [e.g., @Zhang2014b; @Sundue2015].
+While many similar adaptations are found in epiphytic seed plants and serve as key reference points  [@Benzing1990], the evolution of epiphytism in ferns is more complex owing to their unique life-cycle.
+Ferns produce a free-living gametophyte that is typically smaller than the sporophyte and relies on aquatic motile sperm.
+Unlike the sporophyte, the fern gametophyte lacks vascular tissue, stomata, or cuticle, placing it in constant equilibrium with its external environment. 
+In parallel with some of the adaptations in the sporophytes, the gametophytes of some epiphytic ferns also exhibit morphological and physiological adaptations to minimize water loss and tolerate water scarcity.
+For example, extreme desiccation tolerance is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is rare in sporophytes  [@WatkinsJr.2007c; Pittermann2013].
+In addition, epiphytic fern gametophytes tend to exhibit more complex three-dimensional morphologies that may function to retain water and slow its loss [WatkinsJr.2007c].
+Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no previous studies to our knowledge have analyzed the functional ecology of fern epiphytes in a phylogenetic context including both life stages and representatives of all five Cretaceous epiphytic radiations *sensu* [@Schuettpelz2009].
 
-It should be possible to gain greater insight into the selective strength exerted by the epiphytic habitat by utilizing a community-based approach.
-If the harsh abiotic conditions of the canopy act as a filter, we expect to find lower functional richness at a community level, which should result in decreased phylogenetic diversity under a scenario of phylogenetic trait conservatism [@Webb2002].
+To gain insight into the role of the canopy in promoting or restricting diversity, a functional and community-based approach can be applied to study this question. 
+If the canopy is uniformly harsher than the terrestrial environment, thus acting as a "filter", we would expect functional convergence of epiphytes at the community level, which in turn may result in decreased phylogenetic diversity of epiphytic communities if traits are phylogenetically conserved [@Webb2002; @Cavender-Bares2009].
 However, the relative strength of this environmental filter may depend on overall abiotic conditions; for example, relative humidity (RH) and temperature also vary with elevation [@Cardelus2006]. 
-Elevation is a strong predictor of fern species richness, with maximal richness occurring at mid-elevations, especially in the tropics [@Kessler2011a].
-Functional diversity is also correlated with species richness (few species tend to have exactly the same functional traits; @Schleuter2010).
+Conversely, the stratification of the canopy may stimulate functional diversification by niche partitioning, leading to increased functional diversity of epiphytic communities and reduced phylogenetic signal of relevant traits.
+Furthermore, the role of the canopy as a promoter or restrictor of diversity may be modulated by overall abiotic conditions, including most notably relative humidity (RH) and temperature. 
+For instance, elevation is a strong predictor of fern species richness, with maximal richness occurring at mid-elevations, especially in the tropics [@Kessler2011a]. 
+Functional diversity is also correlated with species richness [@Schleuter2010]. 
 Thus, environmental variation along elevational gradients is expected to impact the functional, phylogenetic, and species diversity of epiphytic and terrestrial fern communities.
 
-Here, we investigate the evolution and role in community assembly of functional traits related to epiphytic growth in fern gametophytes and sporophytes on Moorea, French Polynesia.
-Moorea is an isolated tropical oceanic island with a high phylogenetic diversity of ferns distributed over a steep elevational gradient, making it ideal for a community phylogenetic analysis of functional traits.
+Here, we seek to understand the role of the canopy in fern diversification by investigating traits related to epiphytic growth in fern gametophytes and sporophytes on Moorea, French Polynesia.
+Moorea is an isolated tropical oceanic island with high phylogenetic diversity of ferns distributed across a steep elevational gradient, making it ideal for a community phylogenetic analysis of functional traits.
 Specifically, we test the following hypotheses related to epiphytic growth in ferns:
 
-1. The epiphytic environment experiences greater extremes in temperature and humidity than the terrestrial environment, thus acting as a filter on the traits of plants that grow there.
+1. Epiphytic ferns have converged upon similar morphologies in both sporophyte and gametophyte phases.
 
-2. Epiphytic ferns have converged upon similar morphologies multiple times in both sporophyte and gametophyte phases.
-
-3.
-Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities due to environmental filtering, and the strength of this filtering varies with changes in environment.
+2. Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities.
 

### 2019-10-27 ms/manuscript.Rmd
Update MS
to gemmae production in discussion

@@
 In parallel with some of the adaptations in the sporophytes, the gametophytes of some epiphytic ferns also exhibit morphological and physiological adaptations to minimize water loss and tolerate water scarcity.
-For example, extreme desiccation tolerance is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is rare in sporophytes  [@WatkinsJr.2007c; Pittermann2013].
-In addition, epiphytic fern gametophytes tend to exhibit more complex three-dimensional morphologies that may function to retain water and slow its loss [WatkinsJr.2007c].
+For example, extreme desiccation tolerance is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is rare in sporophytes  [@WatkinsJr.2007c; @Pittermann2013].
+In addition, epiphytic fern gametophytes tend to exhibit more complex three-dimensional morphologies that may function to retain water and slow its loss [@WatkinsJr.2007c].
 Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no previous studies to our knowledge have analyzed the functional ecology of fern epiphytes in a phylogenetic context including both life stages and representatives of all five Cretaceous epiphytic radiations *sensu* [@Schuettpelz2009].
@@
 Moorea, French Polynesia (\ang{17;30;}S, \ang{149;50;}W) is a small (\SI{135}{\kilo\meter\squared}) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
-This distance acts as a strong barrier to dispersal [@Carson1995;@Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders \sensu @PteridophytePhylogenyGroupI2016) (\tab S1).
+This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders \sensu @PteridophytePhylogenyGroupI2016) (\tab S1).
 Ferns of Moorea include representatives of all five of the Cretaceous epiphytic radiations described by @Schuettpelz2009 (\ie, polygrammoids, elaphoglossoids, asplenioids, vittarioids, and filmy ferns) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
@@
 We used distance-based methods to analyze community phylogenetic and functional diversity.
-Mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) describe the mean distance between all species in a community (MPD) or the mean distance between the closest pairs of species in a community (MNTD) [@Webb2000; Webb2002].
+Mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) describe the mean distance between all species in a community (MPD) or the mean distance between the closest pairs of species in a community (MNTD) [@Webb2000; @Webb2002].
 Although these metrics were originally formulated for phylogenetic distances, they can be applied in the same way to trait-based distances, enabling a direct comparison between the two [@Sessa2018].
@@
 
-Here, we provide the first study to our knowledge to combine comparative methods with phylogenetic community ecology to analyze the evolutionary implications of epiphytic growth in ferns.
-We find that epiphytic ferns are constrained in size, having consistently smaller fronds and rhizomes than terrestrial species.
-In contrast, epiphytic ferns vary greatly in other traits, leading epiphytic fern communities to be more functionally diverse overall than terrestrial communities.
-Taken together, our results suggest that environmental filtering plays a strong role in limiting plant size, while simultaneously allowing for diversification across other trait axes in epiphytic ferns.
+Our study reveals a striking duality in the phylogenetic and functional diversity of fern epiphytes. 
+On one hand, we find that fern epiphytes are consistently smaller than terrestrial species at both the species and community level (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)), and that epiphytic communities are phylogenetically clustered (\fig \@ref(fig:comm-div)).
+At the same time, we find that epiphytic fern communities are significantly more diverse functionally compared to terrestrial communities (\fig \@ref(fig:comm-div)).
+Collectively, these results suggest that environmental filtering plays a strong role in limiting plant size, while simultaneously promoting diversification across other trait axes in epiphytic ferns.
 
-### Convergence in small size and non-cordate morphotype
+### Convergence in sporophyte size and gametophyte morphotype in fern epiphytes
 
-Epiphytes were markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
-Our results at the local scale are corroborated by @Testo2018a, who also observed abrupt size reduction associated with epiphytic growth in a globally sampled fern phylogeny.
-Stipe length in particular may be a critical trait controlling whole leaf hydraulic conductance; since ferns have resistive stipes in general, species growing in water-limited habitats are expected to decrease stipe length to minimize resistance [@WatkinsJr.2010].
-Our results support previous non-phylogenetically controlled studies that found a prevalence of short stipes in epiphytic ferns [@WatkinsJr.2010; @Creese2011].
-However, given that stipe and frond length, frond width, and rhizome diameter were correlated, and similar traits have been shown to scale together previously [@Arcand2008; @Creese2011], careful experimental manipulations are needed to distinguish the role (or lack thereof) of each in epiphytic growth.
+Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
+These results are corroborated by @Testo2018a, who similarly observed abrupt size reduction associated with epiphytic growth in a globally sampled fern phylogeny. 
+What might explain such consistent and convergent reduction in fern sporophyte size when transitioning into the canopy?
+An obvious answer is that water limitation in the canopy exerts strong selective pressure to maintain small size. 
+In spite of similar and predictable leaf-level economic tradeoffs between angiosperms and ferns [@Karst2007], fern leaves are limited in their hydraulic flexibility as they rely wholly on primary xylem that feeds into laminar tissues with reduced leaf venation.
+Stipe length and structure in particular are critical for regulating whole leaf hydraulic conductance in ferns [@Pittermann2013].
+Species growing in water-limited habitats like those in the canopy are expected by biomechanical necessity to decrease stipe length to minimize resistance [@WatkinsJr.2010].
+Our results support previous studies that identified a prevalence of short stipes among epiphytic ferns [@WatkinsJr.2010; @Creese2011].
+However, given that stipe and frond length, frond width, and rhizome diameter were correlated in our study, and similar traits have previously been shown to co-vary [@Arcand2008; @Creese2011], careful experimental manipulations are required to distinguish the role of each trait in facilitating epiphytism.
 
-Another clear pattern to emerge from our study is the association between epiphytism and non-cordate (\eg, elongate or filamentous) morphotype.
-Morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived ($< \SI{1}{\year}$), cordate gametophytes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; @Farrar2008].
-The complex, three-dimensional structure of non-cordate gametophytes slows their drying rates, and may be an adaptation to drought-prone epiphytic habitats [@WatkinsJr.2007c; @Pittermann2013], and promote out-crossing [@Farrar2008].
-Our study is the first to our knowledge to demonstrate that the correlation between non-cordate morphology and epiphytism in ferns is significant while controlling for phylogeny.
-@Sundue2015 found that non-cordate morphology is significantly correlated with increased diversification rates within Polypodiaceae, consistent with the hypothesis that this trait promoted diversification within the epiphytic niche.
+Another clear and convergent pattern to emerge from our study supporting the hypothesis of convergence is the association between epiphytism and the prevalence of non-cordate gametophyte morphotypes (\eg, elongate or filamentous).
+Gametophyte morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived ($< \SI{1}{\year}$), cordate morphotypes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; @Farrar2008].
+The complex, three-dimensional structure of non-cordate gametophytes may serve two important functions. 
+Complex thallus morphology can slow drying rates, and slow dehydration is correlated with the ability to recover from drought in gametophytes---likely a critical adaptation to drought-prone epiphytic habitats [@WatkinsJr.2007c; @Pittermann2013].
+Non-cordate thallus morphology may also promote outcrossing through the formation of long-lived gametophyte banks that can persist in epiphytic but not in more disturbance-prone terrestrial habitats [@WatkinsJr.2007a; @Farrar2008].
+Our study is the first to our knowledge to demonstrate that the correlation between non-cordate morphology and epiphytism in ferns is significant across broad phylogenetic scales.
 
-### Remarkable diversity in other traits
+### Unexpected diversity in non-size related traits of epiphytes
 
-In contrast to their clearly smaller size and non-cordate morphotypes, epiphytes were remarkably diverse in other traits.
-Epiphytes tended to have less dissected fronds with fewer pinnae than terrestrial species (\tab \@ref(tab:pics)), but the range of values for these traits was quite wide at the community level (error bars in \fig \@ref(fig:cwm-div)).
-It is likely that the high diversity of non-size related traits we observed reflects different physiological adaptations to various canopy niches.
-For example, filmy ferns (family Hymenophyllaceae), which depend on desiccation tolerance to survive dry conditions, have vastly different morphology from bird's nest ferns (*Asplenium nidus* and others), which capture detritus in their rosette leaves.
+Despite the apparent constraints on size and the recurrence of non-cordate gametophytes exhibited by epiphytes, we observed greater morphological diversity among epiphytes versus terrestrial species for several traits.
+These results do not support the hypothesis that fern epiphytes exhibit reduced morphological diversity.
+For example, although epiphytes demonstrated less dissected fronds at the species level (\tab \@ref(tab:pics)), they exhibited much greater variation in this trait relative to terrestrial ferns at the community level (Fig.
+S2).
+It is likely that the high diversity of this trait reflects different physiological adaptations to various canopy niches.
+For instance, filmy ferns (Hymenophyllaceae), which depend on desiccation tolerance to survive dry conditions, have extremely thin, highly dissected fronds.
+Bird’s nest ferns (e.g., *Asplenium nidus*), in contrast, avoid desiccation by capturing detritus and water in the bases of their simple, rosette leaves. 
+At low-elevation sites on Moorea where these ferns co-occur, filmy ferns frequently grow in more humid portions of trees including near the main trunk, whereas bird’s nest ferns often grow higher in the canopy and on more exposed, drought-prone branches (J.
+Nitta, pers.
+obs.).
+Similar predictable patterns of niche partitioning at the local scale has been reported in neotropical epiphtyic ferns as well [@Tausz2001; @Hietz2008]. 
+This distribution may reflect alternative water-usage strategies aligned to specific microhabitats in the canopy [desiccation tolerance vs. water capture; @Hietz2008]. 
+Such hypotheses need to be carefully tested with additional quantitative investigation, including more finely examining microclimatic variation and species composition across the canopy.
 
-Surprisingly, SLA was only slightly lower in epiphytes relative to terrestrial species (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
-Specific leaf area is an important part of the leaf economic spectrum, which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
-slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a].
-We expected that epiphytes would have lower SLA, given their more extreme growing conditions. 
-The lack of a clear difference in SLA between epiphytes and terrestrial species observed here may be an effect of scale: the LES was defined at regional to global scales, but tends to break down at local scales encompassing less environmental variation [@Messier2017].
+Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA) (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
+Specific leaf area is part of the leaf economic spectrum (LES), which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
+slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a; Diaz2015].
+We expected that epiphytes would exhibit lower SLA, given their more extreme growing conditions.
+In contrast, we observed overlapping, wide ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes (\fig \@ref(fig:cwm-div)).
+This may reflect adaptations to specific niches that are not captured in the binary categorization of habitat commonly applied in ferns. 
+For example, the following species ranked among those with the five lowest SLA values (i.e., thick, well-protected leaves) on Moorea: *Humata anderssonii* (a twig epiphyte), *Acrostichum aureum* (a terrestrial mangrove specialist), and *Paesia divaticatissima* (a terrestrial species occurring on high-elevation ridges) (Table S2).
+Each of these species is likely to experience prolonged periods of drought-stress. 
+In-depth characterization of such microhabitats was beyond the scope of our study and should be investigated further in the future.
 

### 2019-10-28 code/functions.R
Change 'sla' in PCA Fig to 'SLA'

@@ -2136,6 +2136,7 @@ make_pca_plot <- function (pca_results, habit_colors, traits) {
   a <- 
     pca_results$traits_locs %>%
     filter(analysis_type == "standard") %>%
+    mutate(trait = str_replace_all(trait, "sla", "SLA")) %>%
     ggplot(aes(x = PC1, y = PC2)) +
     geom_vline(xintercept = 0, linetype = "longdash", color = "dark gray") +
     geom_hline(yintercept = 0, linetype = "longdash", color = "dark gray") +
@@ -2157,6 +2158,7 @@ make_pca_plot <- function (pca_results, habit_colors, traits) {
   b <- 
     pca_results$traits_locs %>%
     filter(analysis_type == "phylogenetic") %>%
+    mutate(trait = str_replace_all(trait, "sla", "SLA")) %>%
     ggplot(aes(x = PC1, y = PC2)) +
     geom_vline(xintercept = 0, linetype = "longdash", color = "dark gray") +
     geom_hline(yintercept = 0, linetype = "longdash", color = "dark gray") +

### 2019-10-28 ms/manuscript.Rmd
Update MS

@@
 The heterogeneous niche space of the angiosperm canopy comprising stratified light and humidity levels and a range of substrates for growth became rapidly filled by a variety of epiphytic plants, including mosses [@Fiz-Palacios2011], leafy liverworts [@Feldberg2014], orchids [@Givnish2015b], bromeliads [@Givnish2014a], and ferns [@Schneider2004d].
-The transition to epiphytic growth was particularly important for modern fern diversification: ca. 29% of all fern species are epiphytes vs. ca. 10% across all vascular plants [@Kress1986].
+The transition to epiphytic growth was particularly important for modern fern diversification: \circa 29% of all fern species are epiphytes vs. \circa 10% across all vascular plants [@Kress1986].
 Moreover, major transitions to epiphytism occurred at least five times within ferns, each followed by diversification during the late Cretaceous or early Cenozoic [@Schuettpelz2009; @WatkinsJr.2012].
@@
 If the canopy is uniformly harsher than the terrestrial environment, thus acting as a "filter", we would expect functional convergence of epiphytes at the community level, which in turn may result in decreased phylogenetic diversity of epiphytic communities if traits are phylogenetically conserved [@Webb2002; @Cavender-Bares2009].
-However, the relative strength of this environmental filter may depend on overall abiotic conditions; for example, relative humidity (RH) and temperature also vary with elevation [@Cardelus2006]. 
 Conversely, the stratification of the canopy may stimulate functional diversification by niche partitioning, leading to increased functional diversity of epiphytic communities and reduced phylogenetic signal of relevant traits.
@@
 Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and pinna number (\tval-test, all \pval < 0.05; \fig \@ref(fig:cwm-div)).
-Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval = `r get_mod_result(model_summ_print, "sla", "p")`).
+Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval `r get_mod_result(model_summ_print, "sla", "p")`).
 Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval = `r get_mod_result(model_summ_print, "dissection", "p")`).
@@
 Species richness increased with elevation for both epiphytes and terrestrial communities (\fig \@ref(fig:comm-div)).
-Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "mpd.obs.z.func", "p")` and `r get_mod_result(t_test_print, "mntd.obs.z.func", "p")`, respectively).
-Functional diversity increased increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval = `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; \fig \@ref(fig:comm-div)).
+Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, \pval `r get_mod_result(t_test_print, "mpd.obs.z.func", "p")` and `r get_mod_result(t_test_print, "mntd.obs.z.func", "p")`, respectively).
+Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval `r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval `r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval = `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; \fig \@ref(fig:comm-div)).
 No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (\fig \@ref(fig:comm-div)).
@@
 Despite the apparent constraints on size and the recurrence of non-cordate gametophytes exhibited by epiphytes, we observed greater morphological diversity among epiphytes versus terrestrial species for several traits.
-These results do not support the hypothesis that fern epiphytes exhibit reduced morphological diversity.
-For example, although epiphytes demonstrated less dissected fronds at the species level (\tab \@ref(tab:pics)), they exhibited much greater variation in this trait relative to terrestrial ferns at the community level (Fig.
-S2).
+These results do not support the hypothesis that fern epiphytes have reduced functional diversity.
+For example, although epiphytes demonstrated less dissected fronds at the species level (\tab \@ref(tab:pics)), they exhibited much greater variation in this trait relative to terrestrial ferns at the community level (\fig S2).
 It is likely that the high diversity of this trait reflects different physiological adaptations to various canopy niches. 
@@
 obs.).
-Similar predictable patterns of niche partitioning at the local scale has been reported in neotropical epiphtyic ferns as well [@Tausz2001; @Hietz2008]. 
-This distribution may reflect alternative water-usage strategies aligned to specific microhabitats in the canopy [desiccation tolerance vs. water capture; @Hietz2008]. 
+Similar predictable patterns of niche partitioning at the local scale has been reported in neotropical epiphtyic ferns as well [@Hietz1998; @Tausz2001]. 
+This distribution may reflect alternative water-usage strategies aligned to specific microhabitats in the canopy [desiccation tolerance vs. water capture; @Hietz1998]. 
 Such hypotheses need to be carefully tested with additional quantitative investigation, including more finely examining microclimatic variation and species composition across the canopy.
@@
 Specific leaf area is part of the leaf economic spectrum (LES), which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
-slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a; Diaz2015].
+slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a; @Diaz2015].
 We expected that epiphytes would exhibit lower SLA, given their more extreme growing conditions.
@@
 We did not find that gemmae production in gametophytes was linked with epiphytism. 
-Gemmae are asexual propagules that allow gametophytes to persist over multiple growing seasons and attain large population sizes [@Farrar2008].
-Gemmae production should be a useful trait in an epiphytic context because it may allow gametophytes to increase their surface area and hedge against population loss due to extreme drying events [@Farrar2008].
-Gemmae can also function sexually by producing antherida in the presence of antheridiogen secreted by other gametophytes [@Emigh1977], which may confer a selective advantage in the epiphytic environment by enabling outcrossing [@Dassler2001].
-It is possible that some, but not all, epiphytic ferns rely on such strategies, and that they are not under similarly strong selection as other traits such as leaf size and gametophyte morphotype. 
-Interestingly, when we classified 'widespread' species as those with gametophytes occurring above or below the minimum or maximum elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with widespread growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution).
-Thus, it seems that gemmae may function in long-term gametophyte persistence, but are not required for epiphytic growth.
+Rather, occurrence of gemmae strongly reflects phylogeny, appearing concentrated in filmy ferns (Hymenophyllaceae) and vittarioid ferns (\fig \@ref(fig:tree)).
+Gemmae are asexual propagules occurring primarily in non-cordate gametophytes that allow gametophytes to persist over multiple growing seasons and attain large population sizes, and have been hypothesized to be associated with harsh environmental conditions [@Farrar2008].
+In particular, gemmae production should be useful in an epiphytic context because they may allow gametophytes to increase their boundary layer and hedge against population loss during extreme drying [@Farrar2008].
+Gemmae can also function sexually by producing antheridia in the presence of antheridiogen secreted by other gametophytes [@Emigh1977], which may confer a selective advantage in the epiphytic environment by enabling outcrossing [@Dassler2001].
+As with our results summarized above, it is possible that the diversity of gemmae and non-gemmae producing epiphytic species reflects undetected niche variation in the canopy.
+Interestingly, when we classified 'widespread' species as those with gametophytes occurring above or below the minimum or maximum elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with widespread growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
 
-Presence of hairs and glands in gametophytes has very strong phylogenetic signal (\tab \@ref(tab:phylosig-binary)), and these traits are associated with certain taxonomic groups [@Stokey1951; @Nayar1971].
+Finally, the presence of hairs and glands in gametophytes also exhibit very high phylogenetic signal (\ie, much greater than expected under Brownian motion; \tab \@ref(tab:phylosig-binary)), and these traits are concentrated in certain taxonomic groups [\eg, hairs, grammitid ferns; glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
 Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], and glands may affect rates of osmosis by modifying lipid concentrations [@Crow2011], we did not detect a correlation between hair or gland production and epiphytic growth.
-Wax-secreting glands are prevalent in both sporophytes and gametophytes of many notholaenid ferns, a xeric-adapted clade, but their functional significance has not been established [@Johnson2012].
-Further tests are needed to determine if such glands serve a functional role in preventing water loss in desert ferns and epiphytes.
+Given their high phylogenetic signal, it is possible that these traits are unrelated to environment but rather reflect phylogenetically conserved developmental pathways [@Johnson2012].
+Alternatively, the possibility remains that some epiphytic species rely on hairs or glands to prevent or slow drying depending on their niche within the canopy (\eg, hair-bearing grammitid gametophytes at mid to high elevation).
 
-### The role of environmental filtering in the canopy
+### The dual roles of the canopy as both a filter and promoter of diversity
 
-Growth habit and temperature are clearly important factors in most of the community diversity metrics we studied, as shown by the full-subsets model analysis (\fig \ref{fig:heatmap}).
-This is also supported by the linear elevation models (\figs \ref{fig:cwm-div}, \ref{fig:comm-div}), given that temperature shows a strong linear relationship with elevation (\fig \ref{fig:climate}).
+Our results demonstrate that the canopy is far more dynamic with regard to fern epiphytism than has been previously appreciated. 
+On the one hand, the canopy appears to have served as a strong barrier to entry, sorting species by size. 
+Furthermore, the results of our phylogenetic community structure analyses partially support the hypothesis of decreased phylogenetic diversity in epiphytic communities due to filtering: we observed that epiphytic fern communities were more phylogenetically clustered for MNTD~phy~, but not MPD~phy~ (\fig \@ref(fig:comm-div)).
+This makes sense given that most epiphytes have radiated within five major clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely related species from two or more epiphytic clades, but exhibits a similar overall degree of phylogenetic diversity as terrestrial communities.
+
+However, while the canopy clearly acts as a filter, we also find that it simultaneously functions as a promoter of diversity. 
+This is evident in our analyses of MPD~func~ and MNTD~func~, which both indicate that epiphytic communities are more functionally diverse compared to terrestrial species (\fig \@ref(fig:comm-div)), thus contradicting our hypothesis of decreased functional diversity in epiphytic communities.
+Additional analyses using different trait combinations support the role of non-size related traits driving this pattern (\fig S1).
+We propose that the high functional diversity of epiphytic fern communities is maintained by previously undetected niche partitioning within epiphytic habitats as follows.
+The canopy is generally drier than terrestrial habitats, exerting strong selection for small size. 
+At the same time, microclimatic conditions in trees vary rapidly over short distances with height and distance from the trunk [@Hietz1995; @Cardelus2005], contributing to greater environmental variation per unit area than terrestrial habitats [@Rojas-Hernandez2018].
+This vertical and horizontal diversity of niches is occupied by different fern species [@WatkinsJr.2006a; @Cardelus2007; @Parra2009; @Acebey2017].
+Thus, when analyzed together, epiphytic fern communities encompass a wider range of morphological variation than terrestrial communities on average.
+Assuming that the traits we measured reflect physiological strategies of epiphytes, this high morphological diversity likely derives from physiological adaptations to a variety of within-canopy niches.
+Clearly, fine-grained characterization of canopy niches and associated adaptations of epiphytic ferns occupying them is an important goal for future research.
+
+### The primacy of growth habit in determining fern community composition as indicated by modelling
+
+The primary role of growth habit, in combination with temperature, in structuring fern communities is supported by the full-subsets model analysis (\fig \ref{fig:heatmap}).
 It is somewhat puzzling that VPD did not contribute significantly to any models, as we hypothesized that this would be important for epiphytes that are frequently subjected to water stress. 
-One reason could be due to conditions at our study site: periods of prolonged rain result in the humidity sensor tip becoming wet, and it does not function normally again (i.e., shows 100% RH) until it dries out.
-The raw data show several extended periods of 100% RH in terrestrial dataloggers at higher elevations.
-Despite these possible inaccuracies however, epiphytic habitats are clearly drier than terrestrial (\fig \ref{fig:climate}).
-
-We obtained partial support for the hypothesis that epiphytic fern communities would be more phylogenetically clustered than terrestrial ones; this was observed for MNTD, but not MPD (\fig \@ref(fig:comm-div)).
-This makes sense given that most epiphytes have radiated within five major clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely-related species from two or more epiphytic clades, but has a similar overall level of phylogentic diversity as terrestrial communities.
-However, we did not detect significant trends in phylogenetic community structure of epiphytic and terrestrial communities along the elevational gradient.
-It is possible that the scale of the current study (from \SIrange{200}{1200}{\meter}) prevented the detection of such trends.
-When the plots in this study were analyzed together with additional plots from Tahiti reaching a maximum altitude of just over \SI{2000}{\meter}, @Nitta2017 detected increasing phylogenetic clustering in fern sporophyte communities (epiphytes and terrestrial species together) with elevation.
-@Kluge2011 also compared phylogenetic and trait diversity between epiphytic and terrestrial ferns over an elevational gradient from \SIrange{100}{3400}{\meter} in Costa Rica, and found trait overdispersion at mid-elevations for epiphytes but no trends in phylogenetic diversity for either group.
-They accounted this to evolutionary plasticity in traits, but did not quantify the degree of phylogenetic signal in the traits they studied.
-
-Results for functional community diversity contradicted our hypotheses; epiphytic communities were more functionally diverse than terrestrial as measured by both MPD and MNTD (\fig \@ref(fig:comm-div)).
-However, this result is dependent on the traits used to measure functional diversity; when the same analyses were done using only size-related traits (scaled, untransformed stipe length, frond length and width, and rhizome diameter), the opposite result was observed (\fig S1).
-Furthermore, terrestrial species show a much wider range of values for size traits (\fig \@ref(fig:cwm-div)), and neither functional group appears to be restricted to a smaller area in trait space (\fig \@ref(fig:pca)).
-This suggests that epiphytic ferns are subject to environmental filtering for size-related traits, while simultaneously diversifying along other trait axes.
-This could be maintained by niche partitioning within epiphytic habitats, which may be more diverse than terrestrial for a given area [@Rojas-Hernandez2018].
-Epiphytic habitats are highly heterogeneous, and microclimatic conditions vary with height along the host and distance from the trunk [@Hietz1995].
-The vertical stratification of epiphytic species has been well documented in ferns [\eg, @Parra2009; @WatkinsJr.2006a; @Acebey2017] and broader studies [\eg, @Kromer2007; @Zotz2007].
-Future research should focus on bringing comparative phylogenetic and community ecological methods to bear on the large body of work characterizing differences within epiphytic microhabitats.
+Previous studies have shown marked variation in the air water potential across epiphytic habitats, with @WatkinsJr.2009 reporting a \circa \SI{40}{\mega\pascal} shift along the trunks of emergent canopy trees.
+Our epiphytic dataloggers only showed slightly greater VPD relative to those on the ground, with this difference largely dissipating at high elevations (\fig \ref{fig:climate}).
+One reason for such low observed variation could be due to environmental conditions at our study site and placement of dataloggers: Moorea is a generally wet site overall (average annual rainfall \SI{325}{\centi\meter}), and we placed the epiphytic dataloggers on tree trunks, not in the higher emergent portions of the canopy.
+Therefore, these measurements should be interpreted as reflecting the minimum amount of difference between terrestrial and canopy environments. 
+Future studies characterizing within-canopy niches of epiphytic ferns should include placement of dataloggers at multiple sites throughout the canopy, which will allow for more detailed tests of the hypothesis that epiphytic fern diversity is maintained by niche partitioning.
 
@@
 
-We have demonstrated that growth habit, and several traits associated with it, show significant phylogenetic signal in ferns.
-Furthermore, we find that epiphytes are constrained by size while being remarkably diverse in other traits.
-Finally, we observed drier conditions in the canopy relative to terrestrial habitats. 
-Taken together, our results support a scenario of environmental filtering acting to limit size and phylogenetic diversity in epiphytic ferns, while simultaneously enabling diversification along other trait axes.
-The generality of this pattern could be further tested by investigating the functional traits of other groups with large epiphytic radiations such as orchids and bromeliads.
+Recent studies [e.g., @Sundue2015; @Testo2016a; @Lehtonen2017; @Testo2018a] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
+Our study provides one possible explanation for this finding. 
+Here, we demonstrate that epiphytism is likely a more complex trait than previously realized, and almost certainly requires the evolution of multiple associated and dependent traits. 
+We propose that the timing and order of the evolution of such traits are related to the complex underlying diversification dynamics observed in these studies. 
+Future research investigating the tempo and mode of the evolution of epiphytism in ferns should seek to better partition this diversity when exploring hypotheses of exceptional diversification.
 
-Recent studies [e.g., @Sundue2015; @Testo2016a; @Lehtonen2017; @Testo2018a] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
-Epiphytism itself is a complex trait requiring the evolution of multiple associated traits, and the timing and order of the evolution of such traits may be related to the complex diversification dynamics observed in these studies.
-Small size seems to be a prerequisite for epiphytic growth, as indicated both by our results and the global study of @Testo2018a.
-Here, we have shown that size appears to be the exception, not the rule, when it comes to trait diversity of epiphytes: although epiphytes are universally small, our data suggest they are diverse in most other traits.
-This emphasizes that fern epiphytes should not be treated as a monolithic group, but rather encompass a wide range of adaptive forms.
-We hope future studies will build on our results by analyzing traits involved in epiphytic growth other than body size across densely sampled phylogenies to dissect how each may relate to diversification and growth habit.
-Additionally, future studies should focus on determining in more detail how niche space is partitioned within epiphytic vs. terrestrial habitats, which is clearly a major dividing line for ferns.
+Small size appears to be a prerequisite for epiphytic growth, as indicated both by our results and by @Testo2018a.
+However, we have shown that size appears to be the exception, not the rule, when it comes to trait diversity of epiphytes: although epiphytes are generally small, our data suggest they are diverse in numerous other traits.
+This emphasizes that fern epiphytes should not be treated as a single, static entity, but rather encompass a wide range of adaptive forms. 
+Our results suggest new targets for future studies of fern ecology and evolution. 
+We hope future global phylogenetic studies will build on our results by analyzing traits involved in epiphytic growth other than plant size (\eg, gametophyte morphotype, frond dissection, SLA, and hair and gland presence) to dissect how each may relate to diversification and growth habit.
+Additionally, future community ecological studies should focus on determining in more detail how niche space is partitioned within epiphytic vs.
+terrestrial habitats, which is clearly a major dividing line for ferns.
 
@@
 Suzanne Vinette, Saad Amer, Tristan Wang, and Tohei Theophilus helped with community surveys and measurement of traits.
+Ravahere Taputuarai helped collect environmental data.
 Staff at the University of California, Berkeley Richard B.
@@
 
-J.H.N. conceived and carried out the study. R.T. collected data. J.H.N. performed the analyses. J.H.N., J.E.W., and C.C.D. wrote the manuscript.
+J.H.N. conceived and carried out the study and analyzed the data. J.H.N. and C.C.D. wrote the draft with editorial input from J.E.W.
 
@@
 Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from July 7, 2013 to July 5, 2014, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).
-'Epiphytic' dataloggers were placed at ca.
-2 m on trees, and 'terrestrial' dataloggers placed at ground level.
+'Epiphytic' dataloggers were placed at \circa 2 m on trees, and 'terrestrial' dataloggers placed at ground level.
 Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown.
@@
 Quantitative traits only, including only species with no missing observations for any traits ($\n = 104$ \si{\species}).
+Trait abbreviations as follows: dissection, degree of frond dissection; length, frond length; width, frond width; pinna, number of pinnae pairs; rhizome, rhizome diameter; stipe, stipe length; SLA, specific leaf area.
 
@@
 Major Cretaceous epiphytic radiations \sensu @Schuettpelz2009 labeled on tree in green: H, hymenophylloids; T, trichomanoids, V, vittarioids; A, asplenioids; E, elaphoglossoids; P, polygrammoids (trichomanoids diversified prior to the Cretaceous but are included as an important extant epiphytic clade).
-Units on timescale for tree in millions of years.
 
@@
 ```{r traits-used-image, fig.height = 9, fig.width = 6, fig.cap = "Examples of traits used in this study."}
-knitr::include_graphics(here::here("images/Figure_3_1_traits_lowres.pdf"))
+knitr::include_graphics(here::here("images/Fig_1_traits.pdf"))
 ```

### 2019-10-28 ms/manuscript.Rmd
Update MS
Output eps image format for MS submission

Standardize 'pinna no.'

@@
 The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
-In the standard PCA, size-related traits tended to load on PC1, with SLA, pinna number, and degree of dissection on PC2.
-In the phylogenetic PCA, SLA was strongly loaded on PC1, while dissection, pinna, and other size-related traits were mostly loaded on PC2.
+In the standard PCA, size-related traits tended to load on PC1, with SLA, number of pinnae pairs, and degree of dissection on PC2.
+In the phylogenetic PCA, SLA was strongly loaded on PC1, while dissection, pinnae, and other size-related traits were mostly loaded on PC2.
 
@@
 Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs \@ref(tab:phylosig-cont), \@ref(tab:phylosig-binary), \fig \@ref(fig:tree)), and for quantitative traits, different results were obtained for $\lambda$ and K.
-When measured with $\lambda$, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of $\lambda$ close to 1; only number of pinna pairs had $\lambda$ close to zero (\tab \@ref(tab:phylosig-cont)).
+When measured with $\lambda$, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of $\lambda$ close to 1; only number of pinnae pairs had $\lambda$ close to zero (\tab \@ref(tab:phylosig-cont)).
 However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of $\K < 1$ (\tab \@ref(tab:phylosig-cont)).
@@
 
-Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and pinna number (\tval-test, all \pval < 0.05; \fig \@ref(fig:cwm-div)).
+Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval < 0.05; \fig \@ref(fig:cwm-div)).
 Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval `r get_mod_result(model_summ_print, "sla", "p")`).
@@
 For instance, filmy ferns (Hymenophyllaceae), which depend on desiccation tolerance to survive dry conditions, have extremely thin, highly dissected fronds.
-Bird’s nest ferns (e.g., *Asplenium nidus*), in contrast, avoid desiccation by capturing detritus and water in the bases of their simple, rosette leaves. 
+Bird’s nest ferns (e.g., *Asplenium nidus* L.), in contrast, avoid desiccation by capturing detritus and water in the bases of their simple, rosette leaves. 
 At low-elevation sites on Moorea where these ferns co-occur, filmy ferns frequently grow in more humid portions of trees including near the main trunk, whereas bird’s nest ferns often grow higher in the canopy and on more exposed, drought-prone branches (J.
@@
 This may reflect adaptations to specific niches that are not captured in the binary categorization of habitat commonly applied in ferns. 
-For example, the following species ranked among those with the five lowest SLA values (i.e., thick, well-protected leaves) on Moorea: *Humata anderssonii* (a twig epiphyte), *Acrostichum aureum* (a terrestrial mangrove specialist), and *Paesia divaticatissima* (a terrestrial species occurring on high-elevation ridges) (Table S2).
+For example, the following species ranked among those with the five lowest SLA values (i.e., thick, well-protected leaves) on Moorea: *Humata anderssonii* (Mett.) C.Chr.
+(a twig epiphyte), *Acrostichum aureum* L.
+(a terrestrial mangrove specialist), and *Paesia divaticatissima* (Dryand.) Copel.
+(a terrestrial species occurring on high-elevation ridges) (Table S2).
 Each of these species is likely to experience prolonged periods of drought-stress. 
@@
 **(a)** Sporophyte traits.
-All sporophyte traits were quantitative, including frond length and width, stipe length, rhizome dia., number of pinna pairs (in this example, nine) and degree of lamina dissection (in this example, once-pinnate).
+All sporophyte traits were quantitative, including frond length and width, stipe length, rhizome dia., number of pinnae pairs (in this example, nine) and degree of lamina dissection (in this example, once-pinnate).
 Specific leaf area (ratio of area to mass of leaf lamina) not depicted.
@@
 climate_plot
-ggsave(plot = climate_plot, filename = here::here("ms/Fig_2.pdf"), height = 4, width = 7, units = "in")
+ggsave(plot = climate_plot, filename = here::here("ms/Fig_2.eps"), height = 4, width = 7, units = "in")
 ```
@@
 pca_plot
-ggsave(plot = pca_plot, filename = here::here("ms/Fig_3.pdf"), height = 6, width = 6, units = "in")
+ggsave(plot = pca_plot, filename = here::here("ms/Fig_3.eps"), height = 6, width = 6, units = "in")
 ```
@@
 traits_with_tree
-ggsave(plot = traits_with_tree, filename = here::here("ms/Fig_4.pdf"), height = 8, width = 6, units = "in")
+ggsave(plot = traits_with_tree, filename = here::here("ms/Fig_4.eps"), height = 8, width = 6, units = "in")
 ```
@@
 importance_heatmap
-ggsave(plot = importance_heatmap, filename = here::here("ms/Fig_5.pdf"), height = 4, width = 6, units = "in")
+ggsave(plot = importance_heatmap, filename = here::here("ms/Fig_5.eps"), height = 4, width = 6, units = "in")
 ```
@@
 combined_cwm_plots
-ggsave(plot = combined_cwm_plots, filename = here::here("ms/Fig_6.pdf"), height = 4, width = 7, units = "in")
+ggsave(plot = combined_cwm_plots, filename = here::here("ms/Fig_6.eps"), height = 4, width = 7, units = "in")
 ```
@@
 combined_comm_div_plots
-ggsave(plot = combined_comm_div_plots, filename = here::here("ms/Fig_7.pdf"), height = 8, width = 4, units = "in")
+ggsave(plot = combined_comm_div_plots, filename = here::here("ms/Fig_7.eps"), height = 8, width = 4, units = "in")
 ```

### 2019-10-28 code/functions.R

@@ -2628,7 +2628,7 @@ plot_traits_on_tree <- function (traits, phy, ppgi) {
     scale_x_discrete(
       position = "top",
       breaks= c("sla", "dissection", "stipe", "rhizome", "pinna"),
-      labels= c("SLA", "Dissection", "Stipe length", "Rhizome diam.", "Pinna num.")) +
+      labels= c("SLA", "Dissection", "Stipe length", "Rhizome diam.", "Pinna no.")) +
     scale_fill_scico(
       palette = "lajolla",
       breaks = c(0, 1),

### 2019-10-28 ms/SI.rmd
Update MS
final changes before initial sumbission to New Phytologist

@@
 
-ggsave(plot = fig_s1, filename = here::here("ms/Fig_S1.pdf"), height = 8, width = 4, units = "in")
+ggsave(plot = fig_s1, filename = here::here("ms/Fig_S1.eps"), height = 8, width = 4, units = "in")
 ```
@@
 loadd(cwsd_scatterplot, cache = canopy_cache)
-ggsave(plot = cwsd_scatterplot, filename = here::here("ms/Fig_S2.pdf"), height = 6, width = 7, units = "in")
+ggsave(plot = cwsd_scatterplot, filename = here::here("ms/Fig_S2.eps"), height = 6, width = 7, units = "in")
 ```

### 2019-10-28 ms/manuscript.Rmd
Update MS
final changes before initial sumbission to New Phytologist

@@
   words = c(
-    "Total word count (excluding summary, references and legends):_n",
+    "Total word count (Introduction, Materials and Methods, Results, Discussion):_n",
     "Summary:_n",
@@
     "No. of tables:_5",
-    "No. of supporting information files:_7 (Fig. S1; Table S1-S6)", "", "", "", "")
+    "No. of supporting information files:_8 (Figs. S1-S2; Tables S1-S6)", "", "", "", "")
 ) %>%
@@
 
-**Key words:** community ecology, functional traits, gametophyte, Moorea, Society Islands, Tahiti.
+**Key words:** community ecology, epiphyte, fern, functional traits, gametophyte, Moorea, Society Islands, Tahiti.
 
@@
 In addition, epiphytic fern gametophytes tend to exhibit more complex three-dimensional morphologies that may function to retain water and slow its loss [@WatkinsJr.2007c].
-Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no previous studies to our knowledge have analyzed the functional ecology of fern epiphytes in a phylogenetic context including both life stages and representatives of all five Cretaceous epiphytic radiations *sensu* [@Schuettpelz2009].
+Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no previous studies to our knowledge have analyzed the functional ecology of fern epiphytes in a phylogenetic context including both life stages and representatives of all five Cretaceous epiphytic radiations *sensu* @Schuettpelz2009.
 
@@
 Moorea is an isolated tropical oceanic island with high phylogenetic diversity of ferns distributed across a steep elevational gradient, making it ideal for a community phylogenetic analysis of functional traits.
-Specifically, we test the following hypotheses related to epiphytic growth in ferns:
-
-1. Epiphytic ferns have converged upon similar morphologies in both sporophyte and gametophyte phases.
-
-2. Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities.
+Specifically, we test the following hypotheses related to epiphytic growth in ferns: 1) Epiphytic ferns have converged upon similar morphologies in both sporophyte and gametophyte phases; 2) Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities.
 
@@
 
-Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (\fig \@ref(fig:pca)(c, d)).
+Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (\fig \@ref(fig:pca)*c*, *d*).
 The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
@@
 In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (\fig \ref{fig:heatmap}).
-Most of the the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (\tab S6).
+Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (\tab S6).
 Spatial autocorrelation was not detected for any of the diversity metrics (\tab S6).
@@
 obs.).
-Similar predictable patterns of niche partitioning at the local scale has been reported in neotropical epiphtyic ferns as well [@Hietz1998; @Tausz2001]. 
+Similar predictable patterns of niche partitioning at the local scale have been reported in neotropical epiphtyic ferns as well [@Hietz1998; @Tausz2001]. 
 This distribution may reflect alternative water-usage strategies aligned to specific microhabitats in the canopy [desiccation tolerance vs. water capture; @Hietz1998]. 
@@
 
-### The primacy of growth habit in determining fern community composition as indicated by modelling
+### The primacy of growth habit in determining fern community composition as indicated by modeling
 

### 2019-12-24 ms/manuscript.Rmd
Revise intro

@@
 
-To better understand the role of the canopy in fern diversification, one obvious direction is to explore putatively adaptive fern epiphyte traits in a phylogenetic context. 
+To better understand the role of the canopy in fern diversification, one obvious direction is to explore putatively adaptive fern epiphyte traits in a phylogenetic context [@Givnish2015]. 
 In contrast to most terrestrial plants, epiphytes must contend with greater stress in the form of nutrient limitation, increased evaporative demand, drought, and extreme oscillations in temperature [@Zotz2001; @WatkinsJr.2009; @Lowman2012].
-Previous studies have identified multiple traits as likely adaptations to these conditions, albeit many without consideration of phylogeny. 
+Previous studies have identified multiple traits as likely adaptations to these conditions, albeit many without consideration of phylogeny [\eg, @Kluge2007]. 
 Epiphytic ferns tend to have morphological features that minimize water loss from their leaves, including shorter stipes [@WatkinsJr.2010], thicker leaves [@Kluge2007; @WatkinsJr.2007b], and protection with scales or hairs [@WatkinsJr.2006b].
@@
 In addition, epiphytic fern gametophytes tend to exhibit more complex three-dimensional morphologies that may function to retain water and slow its loss [@WatkinsJr.2007c].
-Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no previous studies to our knowledge have analyzed the functional ecology of fern epiphytes in a phylogenetic context including both life stages and representatives of all five Cretaceous epiphytic radiations *sensu* @Schuettpelz2009.
+Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no previous studies to our knowledge have analyzed the functional ecology of fern epiphytes in a broad phylogenetic context including both life stages.
 
@@
 If the canopy is uniformly harsher than the terrestrial environment, thus acting as a "filter", we would expect functional convergence of epiphytes at the community level, which in turn may result in decreased phylogenetic diversity of epiphytic communities if traits are phylogenetically conserved [@Webb2002; @Cavender-Bares2009].
-Conversely, the stratification of the canopy may stimulate functional diversification by niche partitioning, leading to increased functional diversity of epiphytic communities and reduced phylogenetic signal of relevant traits.
+Conversely, the stratification of the canopy may stimulate functional diversification by niche partitioning, leading to increased functional diversity of epiphytic communities. 
 Furthermore, the role of the canopy as a promoter or restrictor of diversity may be modulated by overall abiotic conditions, including most notably relative humidity (RH) and temperature. 
-For instance, elevation is a strong predictor of fern species richness, with maximal richness occurring at mid-elevations, especially in the tropics [@Kessler2011a]. 
+For instance, elevation is a strong predictor of fern species richness, with maximal richness occurring at mid-elevations, especially in the tropics [@Kromer2005; @Kessler2011a]. 
 Functional diversity is also correlated with species richness [@Schleuter2010]. 

### 2019-12-24 ms/manuscript.Rmd
Add single line break between sentences for methods and results
so subsequent changes are clear in commits
Revise methods

@@
 Moorea, French Polynesia (\ang{17;30;}S, \ang{149;50;}W) is a small (\SI{135}{\kilo\meter\squared}) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
-This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders \sensu @PteridophytePhylogenyGroupI2016) (\tab S1).
-Ferns of Moorea include representatives of all five of the Cretaceous epiphytic radiations described by @Schuettpelz2009 (\ie, polygrammoids, elaphoglossoids, asplenioids, vittarioids, and filmy ferns) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
+This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders; taxonomy follows @PteridophytePhylogenyGroupI2016; \tab S1).
+Ferns of Moorea include representatives of multiple Cretaceous epiphytic radiations (\eg, *Asplenium*, *Elaphoglossum*, Hymenophylloideae, Polypodiaceae, Vittarioideae; [@Schuettpelz2009]) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
 
@@
 A fern community survey was carried out on Moorea comprising 17, \SI{10}{\meter} $\times$ \SI{10}{\meter} plots spanning an elevational gradient from \SIrange{20}{2000}{\meter}, with plots placed \circa every \SI{200}{\meter} [@Nitta2017].
-Most plots were located along trails leading to the three main peaks on the island, Mt.
-Rotui (\SI{899}{\meter}), Mt.
-Mouaputa (\SI{890}{\meter}), and Mt.
-Tohiea (\SI{1206}{\meter}).
+Most plots were located along trails leading to the three main peaks on the island, Mt. Rotui (\SI{899}{\meter}), Mt. Mouaputa (\SI{890}{\meter}), and Mt. Tohiea (\SI{1206}{\meter}).
 Each plot was divided into 25, \SI{2}{\meter} $\times$ \SI{2}{\meter} subplots.
@@
 Growth habit was coded as a binary trait (epiphytic or terrestrial) based on our field observations.
-Some species do not fall clearly into either category (\eg, hemi-epiphytes; [@Benzing1990; @Dubuisson2003; @Nitta2009]).
+Some species do not fall clearly into either category (\eg, hemi-epiphytes; [@Benzing1990; @Dubuisson2003; @Zotz2013]).
 We used connection to the soil as the criterion for defining binary growth habit; thus, epipetric plants were treated as epiphytic, and hemi-epiphytes and climbing plants were treated as terrestrial. 
@@
 We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each datalogger, then overall means from the daily values.
-We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (hereafter, "climatic variables").
+We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (a single variable was arbitrarily selected from each set of correlated variables; hereafter, "climatic variables").
 Variables relating to VPD, which have no meaning if less than zero, and some other variables that had skewed distributions, were square-root transformed prior to fitting models.
@@
 We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone and their interactions, then selecting the one with the lowest corrected Akaike Information Criterion (AIC~c~) [@Akaike1973].
-A single outlier plot with unusually high temperature and low humidity (Mt.
-Rotui, 830 m plot) was excluded from ANCOVA and linear models.
-Mt.
-Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established (J.
-Nitta, pers.
-obs.).
+A single outlier plot with unusually high temperature and low humidity (Mt. Rotui, 830 m plot) was excluded from ANCOVA and linear models.
+Mt. Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established (J. Nitta, pers. obs.).
 
-For comparative analyses, we used the dated phylogenetic tree of @Nitta2017.
+For comparative analyses, we used the dated plastid phylogenetic tree of @Nitta2017.
+*Microsorum maximum* (Brack.) Copel., a genetically verified hybrid between *M.
+grossum* (Langsd.
+& Fisch.) S.
+B.
+Andrews and *M.
+commutatum* Copel.
+[@Nitta2018] was excluded from analysis because it shares identical plastid sequences with *M.
+grossum*, so the plastid tree does not accurately represent its evolutionary distinctiveness.
+
 We tested for phylogenetic signal in quantitative traits using Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
@@
 
-We detected significant phylogenetic signal in most traits, so we tested for differences between traits of epiphytic vs.
-terrestrial species using phylogenetic comparative methods.
+We detected significant phylogenetic signal in most traits, so we tested for differences between traits of epiphytic vs. terrestrial species using phylogenetic comparative methods.
 For quantitative traits, we used the \function{brunch} function in the \package{caper} package [@Orme2013], which calculates phylogenetically independent contrasts in a quantitative character between alternative states of a binary trait.
@@
 Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife.
-Raw and processed data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED [@Nitta2019a].
+Data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED [@Nitta2019a].
 

### 2019-12-24 ms/manuscript.Rmd
Revise discussion and acknowledgements

@@
 As with our results summarized above, it is possible that the diversity of gemmae and non-gemmae producing epiphytic species reflects undetected niche variation in the canopy.
-Interestingly, when we classified 'widespread' species as those with gametophytes occurring above or below the minimum or maximum elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with widespread growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
+Interestingly, when we classified gametophytes as ‘independent’ if they occurred beyond the range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with independent growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
 
@@
 Our results demonstrate that the canopy is far more dynamic with regard to fern epiphytism than has been previously appreciated. 
-On the one hand, the canopy appears to have served as a strong barrier to entry, sorting species by size. 
+On the one hand, the canopy appears to have served as a potent barrier to entry, strongly selecting for small size. 
 Furthermore, the results of our phylogenetic community structure analyses partially support the hypothesis of decreased phylogenetic diversity in epiphytic communities due to filtering: we observed that epiphytic fern communities were more phylogenetically clustered for MNTD~phy~, but not MPD~phy~ (\fig \@ref(fig:comm-div)).
-This makes sense given that most epiphytes have radiated within five major clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely related species from two or more epiphytic clades, but exhibits a similar overall degree of phylogenetic diversity as terrestrial communities.
+This makes sense given that most epiphytes have radiated within a limited number of clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely related species from two or more epiphytic clades, but exhibits a similar overall degree of phylogenetic diversity as terrestrial communities.
 
@@
 Michele Holbrook, and Jonathan Losos for helpful discussion and comments on drafts.
+Three anonymous reviewers provided comments that improved the manuscript.
 Suzanne Vinette, Saad Amer, Tristan Wang, and Tohei Theophilus helped with community surveys and measurement of traits.
@@
 Gump South Pacific Research Station, Moorea, French Polynesia, in particular Valentine Brotherson, Hinano Teavai-Murphy, Frank Murphy, and Neil Davies, provided support with obtaining research and collection permits, laboratory space, and logistics.
-Funding provided in part by National Science Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to JHN and CCD), Setup Funds from Harvard University to CCD, the American Society of Plant Taxonomists (Research Grant for Graduate Students to JHN), the Garden Club of America (Award in Tropical Botany to JHN), the Harvard University Herbaria (Fernald Fieldwork Fellowship to JHN), the Society of Systematic Biologists (Graduate Student Research Award to JHN), and the Systematics Association (Systematics Research Fund to JHN).
+Funding provided in part by National Science Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to J.H.N.
+and C.C.D.), Setup Funds from Harvard University to C.C.D., the American Society of Plant Taxonomists (Research Grant for Graduate Students to J.H.N.), the Garden Club of America (Award in Tropical Botany to J.H.N.), the Harvard University Herbaria (Fernald Fieldwork Fellowship to J.H.N.), the Society of Systematic Biologists (Graduate Student Research Award to J.H.N.), and the Systematics Association (Systematics Research Fund to J.H.N.).
 

### 2019-12-24 ms/manuscript.Rmd
Revise figure legends

@@
 **\fig \@ref(fig:climate)**: Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter on Moorea, French Polynesia.
-Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from July 7, 2013 to July 5, 2014, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).
+Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from July 7, 2013 to July 5, 2014.
+Vapor pressure deficit (VPD) was calculated from relative humidity and temperature as described in the methods, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).
 'Epiphytic' dataloggers were placed at \circa 2 m on trees, and 'terrestrial' dataloggers placed at ground level.
@@
 **\fig \@ref(fig:tree)**: Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
+Tree obtained from @Nitta2017.
 Relative value of quantitative (sporophyte) traits shown by heat gradient: low values are lighter, high values are darker.
@@
 For a summary of traits, see \tab \@ref(tab:traits-used).
-Major Cretaceous epiphytic radiations \sensu @Schuettpelz2009 labeled on tree in green: H, hymenophylloids; T, trichomanoids, V, vittarioids; A, asplenioids; E, elaphoglossoids; P, polygrammoids (trichomanoids diversified prior to the Cretaceous but are included as an important extant epiphytic clade).
+Major epiphytic radiations identified by @Schuettpelz2009 labeled on tree in green: H, Hymenophylloideae; T, Trichomanoideae; V, Vittarioideae; A, *Asplenium*; E, *Elaphoglossum*; P, Polypodiaceae; taxonomy follows @PteridophytePhylogenyGroupI2016.
 

### 2019-12-24 ms/manuscript.Rmd
Exclude M. x maximum and revise MS with updated phylogenetic PCA results

@@
 In the standard PCA, size-related traits tended to load on PC1, with SLA, number of pinnae pairs, and degree of dissection on PC2.
-In the phylogenetic PCA, SLA was strongly loaded on PC1, while dissection, pinnae, and other size-related traits were mostly loaded on PC2.
+Loadings in the phylogenetic PCA are similar but approximately reversed relative to the standard PCA.
 
@@
 Epiphytes in green, terrestrial species in brown.
-Quantitative traits only, including only species with no missing observations for any traits ($\n = 104$ \si{\species}).
+Quantitative traits only, including only species with no missing observations for any traits (\n = `r pca_results[["species_locs"]] %>% pull(species) %>% n_distinct` \si{\species}).
 Trait abbreviations as follows: dissection, degree of frond dissection; length, frond length; width, frond width; pinna, number of pinnae pairs; rhizome, rhizome diameter; stipe, stipe length; SLA, specific leaf area.

### 2019-12-24 code/plan.R

@@ -23,6 +23,8 @@ plan <- drake_plan(
   species_list = read_csv(file_in("data/nitta_2017/species.csv")) %>%
     clean_names() %>%
     filter(include == 1, tahiti_only == 0) %>%
+    # Exclude Microsorum_xmaximum (hybrid between M. grossum and M. commutatum)
+    filter(genus_sp != "Microsorum_xmaximum") %>%
     pull(genus_sp),
   
   # - Process raw specific leaf area (SLA) measurments
@@ -103,7 +105,9 @@ plan <- drake_plan(
   # Ferns on Moorea and Tahiti (146 species)
   phy = ape::read.tree(
     file_in("data/nitta_2017/treepl_Moorea_Tahiti.tre"
-    )),
+    )) %>%
+    # Exclude Microsorum_xmaximum (hybrid between M. grossum and M. commutatum)
+    drop.tip("Microsorum_xmaximum"),
   
   ### Community data for ferns of Moorea ###
   # Including sporophyte only,

### 2019-12-24 code/functions.R
Add "T" label to tree

@@ -1120,7 +1120,7 @@ analyze_phy_struc_by_habit <- function (comm, phy, traits, null_model = "phyloge
   # Make sure all species are in the phylogeny
   # (but don't subset phylogeny to only species in Moorea
   # communities - we want to use the species pool including
-  # both ferns of Moorea and Tahiti fro mpd, mntd).
+  # both ferns of Moorea and Tahiti for mpd, mntd).
   assert_that(all(comm$species %in% phy$tip.label))
   
   # Split communities into epiphytic / terrestrial taxa, 
@@ -2678,6 +2678,17 @@ plot_traits_on_tree <- function (traits, phy, ppgi) {
       
       H = phy_tax_data %>%
         filter(genus == "Hymenophyllum") %>%
+        pull(species) %>%
+        getMRCA(phy, .),
+      
+      T = phy_tax_data %>%
+        filter(genus %in% c(
+          "Abrodictyum", 
+          "Callistopteris", 
+          "Crepidomanes", 
+          "Didymoglossum", 
+          "Polyphlebium", 
+          "Vandenboschia")) %>%
         pull(species) %>%
         getMRCA(phy, .),
       

### 2019-12-28 ms/manuscript.Rmd
Don't use knitr cache

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
 library(drake)

### 2019-12-28 ms/manuscript.Rmd
Write out tables in html
MS Word can read them fine, and less buggy than rtf

@@
 
-# Write out table in rich text format for submitting
+# Write out table in html for submitting
 # MS as word doc
 traits_used %>%
-  gt %>%
+  gt(rowname_col = "trait") %>%
   fmt_markdown(columns = everything()) %>%
-  gt(rowname_col = "trait") %>%
   tab_row_group(
@@
   ) %>% 
-  as_rtf() %>%
-  write_lines(here::here("ms/Table_1.rtf"))
+  gtsave(here::here("ms/Table_1.html"))
 
@@
 
-# Write out table in rich text format
+# Write out table in html format
 phylosig_cont_traits_table %>%
@@
   )  %>%
-  as_rtf() %>%
-  write_lines(here::here("ms/Table_2.rtf"))
+  gtsave(here::here("ms/Table_2.html"))
 
@@
 
-# Write out table in rich text format
+# Write out table in html
 phylosig_binary_traits_table %>%
@@
   ) %>% 
-  as_rtf() %>%
-  write_lines(here::here("ms/Table_3.rtf"))
+  gtsave(here::here("ms/Table_3.html"))
 
@@
 
-# Write out table in rich text format
+# Write out table in html
 pic_table %>%
@@
   ) %>% 
-  as_rtf() %>%
-  write_lines(here::here("ms/Table_4.rtf"))
+  gtsave(here::here("ms/Table_4.html"))
 
@@
   ) %>% 
-  as_rtf() %>%
-  write_lines(here::here("ms/Table_5.rtf"))
+  gtsave(here::here("ms/Table_5.html"))
 

### 2019-12-28 ms/manuscript.Rmd
Use cairo_ps with high-res fallback for outputting eps figures

@@
 climate_plot
-ggsave(plot = climate_plot, filename = here::here("ms/Fig_2.eps"), height = 4, width = 7, units = "in")
+
+# Tweaked version of cairo_ps device function with higher-resolution
+# for fallback zones
+cairo_ps_high_res <- function(...) {
+  cairo_ps(fallback_resolution = 1200, ...)
+}
+
+ggsave(plot = climate_plot, filename = here::here("ms/Fig_2.eps"), height = 4, width = 7, units = "in", device = cairo_ps_high_res)
 ```
@@
 pca_plot
-ggsave(plot = pca_plot, filename = here::here("ms/Fig_3.eps"), height = 6, width = 6, units = "in")
+ggsave(plot = pca_plot, filename = here::here("ms/Fig_3.eps"), height = 6, width = 6, units = "in", device = cairo_ps_high_res)
 ```
@@
 traits_with_tree
-ggsave(plot = traits_with_tree, filename = here::here("ms/Fig_4.eps"), height = 8, width = 6, units = "in")
+ggsave(plot = traits_with_tree, filename = here::here("ms/Fig_4.eps"), height = 8, width = 6, units = "in", device = cairo_ps_high_res)
 ```
@@
 importance_heatmap
-ggsave(plot = importance_heatmap, filename = here::here("ms/Fig_5.eps"), height = 4, width = 6, units = "in")
+ggsave(plot = importance_heatmap, filename = here::here("ms/Fig_5.eps"), height = 4, width = 6, units = "in", device = cairo_ps_high_res)
 ```
@@
 combined_cwm_plots
-ggsave(plot = combined_cwm_plots, filename = here::here("ms/Fig_6.eps"), height = 4, width = 7, units = "in")
+ggsave(plot = combined_cwm_plots, filename = here::here("ms/Fig_6.eps"), height = 4, width = 7, units = "in", device = cairo_ps_high_res)
 ```
@@
 combined_comm_div_plots
-ggsave(plot = combined_comm_div_plots, filename = here::here("ms/Fig_7.eps"), height = 8, width = 4, units = "in")
+ggsave(plot = combined_comm_div_plots, filename = here::here("ms/Fig_7.eps"), height = 8, width = 4, units = "in", device = cairo_ps_high_res)
 ```

### 2019-12-28 install_latex.R
Add some latex packages

@@ -20,6 +20,11 @@ latex_packages <- c(
   "trimspaces",
   "ulem",
   "makecell",
-  "xcolor")
+  "xcolor",
+  "unicode-math",
+  "ucharcat",
+  "filehook",
+  "mdwtools",
+  "lm-math")
 
 tinytex::tlmgr_install(latex_packages)

### 2019-12-30 ms/SI.rmd
Set margins of all figures to 0.2 in
also fixes cutoff portions of figs

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
 
@@
   scale_fill_scico(palette = "roma", breaks = c(-1, 1)) +
-  labs(x = NULL, y = NULL, title = "a") +
+  labs(x = NULL, y = NULL, title = "(a)") +
   theme_minimal() +
@@
   scale_fill_scico(palette = "roma", breaks = c(-1, 1)) +
-  labs(x = NULL, y = NULL, title = "b") +
+  labs(x = NULL, y = NULL, title = "(b)") +
   theme_minimal() +
@@
 fig_s1 <- mpd_corr_plot + mntd_corr_plot + plot_layout(nrow = 2, ncol = 1)
+
+fig_s1 <- fig_s1 + plot_annotation(theme = theme(plot.margin = margin(
+      0.2, 0.2, 0.2, 0.2, unit = "in")))
 

### 2019-12-30 code/functions.R

@@ -2108,19 +2108,25 @@ make_climate_plot <- function (data, resp_vars,
   }
   
   # Combine plots into single output
-  wrap_plots(plots_df$plot, ncol = 3, nrow = 2) & theme(
+  
+  # Set margins between subplots
+  mar_adj <- c(-0.2, 0.1, 0, 0.1) #top, right, bottom, left
+  
+  p <- wrap_plots(plots_df$plot, ncol = 3, nrow = 2) & theme(
     legend.position = "none",
     # Tweak margins to remove whitespace between plots
-    plot.margin = margin(t = -0.2, r = 0.1, b = 0, l = 0.1, unit = "in")
+    plot.margin = margin(mar_adj[1], mar_adj[2], mar_adj[3], mar_adj[4], unit = "in")
   )
   
-  # Combine plots into single output
-  wrap_plots(plots_df$plot, ncol = 3, nrow = 2) & theme(
-    legend.position = "none",
-    # Tweak margins to remove whitespace between plots
-    plot.margin = margin(t = -0.2, r = 0.1, b = 0, l = 0.1, unit = "in")
-  )
+  # Set outer margins
+  mar_out = .2 + -1*(mar_adj)
   
+  # Set overall margins of combined plot: 0.2 inch margins, accounting for
+  # previous adjustments
+  p +
+    plot_annotation(theme = theme(plot.margin = margin(
+      mar_out[1], mar_out[2], mar_out[3], mar_out[4], unit = "in")))
+
 }
 
 make_pca_plot <- function (pca_results, habit_colors, traits) {
@@ -2217,9 +2223,14 @@ make_pca_plot <- function (pca_results, habit_colors, traits) {
       subtitle = "(d)"
     )
   
-  a + b + c + d + plot_layout(ncol = 2, nrow = 2) &
+  # Combine into final plot with 0.2 in margins
+  p <- a + b + c + d + plot_layout(ncol = 2, nrow = 2) &
     standard_theme2() &
     theme(legend.position = "none")
+  
+  p +
+    plot_annotation(theme = theme(plot.margin = margin(
+      0.2, 0.2, 0.2, 0.2, unit = "in")))
   
 }
 
@@ -2417,7 +2428,10 @@ make_heatmap <- function(important_div_vars, vars_select = c(
     ) +
     scale_y_discrete(labels = parse(text = levels(plot_data$resp_var))) +
     standard_theme2() +
[…]

### 2020-01-08 ms/manuscript.Rmd
Fix climate data processing
- Write out moorea climate data for Dryad
- Use 'rh' as column name in raw data
- Fix measurement interval (15 min not 30 min) in MS, use code for start and end times instead of hard values

@@
   "climate_plot",
+  "climate_raw",
   "climate_select",
@@
   "importance_heatmap",
+  "moorea_climate",
   "pca_plot",
@@
 ```{r climate-numbers, results="hide"}
-# Count total number of days in climate data
-total_days <- daily_mean_climate %>% pull(date) %>% n_distinct()
+# Calculate summary statistics (number of sites, start date, end date) for raw 
+# climate data for preliminary survey period (until 2013-07-07) and final survey
+# period (2013-07-07 and after) separately.
 
-# Check start and end dates
-# - start July 7, 2013
-assert_that(daily_mean_climate %>% pull(date) %>% min == "2013-07-07")
-# - end July 5, 2014
-assert_that(daily_mean_climate %>% pull(date) %>% max == "2014-07-05")
+prelim_climate_survey <-
+  # Subset to Moorea sites used in prelim survey (Mt. Tohiea only)
+  climate_raw %>%
+  filter(str_detect(site, "Aorai", negate = TRUE)) %>%
+  filter(date < "2013-07-07") %>%
+  filter(str_detect(site, "Tohiea")) %>%
+  # Only count by site name, not distinguishing between epi vs. ter dataloggers
+  mutate(site = str_remove_all(site, "_epi|_ter")) %>%
+  summarize(
+    n_sites = n_distinct(site),
+    start = min(date) %>% 
+      format('%B %e, %Y') %>%
+      str_replace_all("  ", " "),
+    end = max(date) %>% 
+      format('%B %e, %Y') %>%
+      str_replace_all("  ", " ")
+  )
 
-# Count total number of sites in climate data
+final_climate_survey <-
+  # Subset to Moorea sites used in final survey
+  climate_raw %>%
+  filter(str_detect(site, "Aorai", negate = TRUE)) %>%
+  filter(date >= "2013-07-07") %>%
+  # Only count by site name, not distinguishing between epi vs. ter dataloggers
+  mutate(site = str_remove_all(site, "_epi|_ter")) %>%
+  summarize(
+    n_sites = n_distinct(site),
+    start = min(date) %>% 
+      format('%B %e, %Y') %>%
+      str_replace_all("  ", " "),
+    end = max(date) %>% 
+      format('%B %e, %Y') %>%
+      str_replace_all("  ", " ")
+  )
+
+# Count total number of days in clean climate data
+total_days <- moorea_climate %>% pull(date) %>% n_distinct()
+
+# Count total number of sites / dataloggers in clean climate data
 total_sites <-
-daily_mean_climate %>%
-  select(site, habit) %>%
-  unique() %>%
+  moorea_climate %>%
+  count(site, habit) %>%
   count(habit)
@@
 Tohiea \SI{1200}{\meter} plot, which only had an epiphytic datalogger.
-Temperature and RH were logged once every \SI{30}{\minute} during two survey periods: a preliminary survey from July 18, 2012 to July 7, 2013 at six plots, and a final survey from July 7, 2013 to July 5, 2014 at 15 plots.
+Temperature and RH were logged once every \SI{15}{\minute} during two survey periods: a preliminary survey from `r prelim_climate_survey$start` to `r prelim_climate_survey$end` at `r prelim_climate_survey$n_sites %>% english2` plots, and a final survey from `r final_climate_survey$start` to `r final_climate_survey$end` at `r final_climate_survey$n_sites %>% english2` plots.
 Two plots were not included in the environmental survey.
 We used the 2013–-2014 data for the analysis, with one exception: the terrestrial datalogger at the Mt.
-Tohiea \SI{393}{\meter} plot malfunctioned and was missing data for a substantial part of the final survey period (March 12--July 5, 2014), so we used data for the same dates from the preliminary survey period for this datalogger instead.
+Tohiea \SI{393}{\meter} plot malfunctioned and was missing data for a substantial part (March 12--July 5, 2014) of the final survey period, so we used data for the same dates from the preliminary survey period for this datalogger instead.
 Three others also failed early during the final survey period, but lacked data from the preliminary period, so were excluded from analysis.
 Other days missing data during the final survey period were excluded from the dataset.
-The final dataset included `r total_days` days of data for `r total_sites %>% filter(habit == "epiphytic") %>% pull(n)` dataloggers and `r total_sites %>% filter(habit == "terrestrial") %>% pull(n)` terrestrial dataloggers.
+The final dataset included `r total_days` days of data for `r total_sites %>% filter(habit == "epiphytic") %>% pull(n)` epiphytic dataloggers and `r total_sites %>% filter(habit == "terrestrial") %>% pull(n)` terrestrial dataloggers.
 We calculated vapor pressure deficit (VPD) from relative humidity and temperature using the \package{plantecophys} R package [@Duursma2015].
@@
 Although trait values can vary within fern species with elevation [@Kessler2007a], our study includes a broad phylogenetic sampling, so we assume intraspecific variation is outweighed by interspecies differences.
-To measure specific leaf area (SLA; \si{\square\meter\per\kilo\gram}), \circa 10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} \diameter punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
+To measure specific leaf area (SLA; \si{\square\meter\per\kilo\gram}), \circa 10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} diam.
+punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
 For species with \textless \SI{2}{\milli\meter} between primary veins, leaf fragments excluding primary veins were obtained by dissection, dried at \SI{60}{\celsius} for \SI{24}{\hour}, weighed, and scanned.
@@
 **\fig \@ref(fig:climate)**: Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter on Moorea, French Polynesia.
-Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from July 7, 2013 to July 5, 2014.
+Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from `r final_climate_survey$start` to `r final_climate_survey$end`.
 Vapor pressure deficit (VPD) was calculated from relative humidity and temperature as described in the methods, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).

### 2020-01-08 code/functions.R

@@ -446,16 +446,12 @@ process_trait_data_for_si <- function (sla_raw, morph_cont_raw, morph_qual_raw) 
 
 #' Filter raw climate data, also calculate VPD
 #'
-#' @param climate_raw_path Path to raw climate data measured with
+#' @param climate_raw Raw climate data measured with
 #' Hobo dataloggers (relative humidity and temperature) every 15 min
 #' on Moorea and Tahiti from 2012-07-18 to 2015-02-06
 #'
 #' @return Tibble
-process_raw_climate <- function (climate_raw_path) {
-  
-  # Read in raw data, includes all dataloggers from Moorea and Tahiti from
-  # 2012-07-18 to 2015-02-06
-  climate_raw <- read_csv(climate_raw_path)
+process_raw_climate <- function (climate_raw) {
   
   # Reformat data, filter out Tahiti sites (on Mt. Aorai)
   # Split apart date into components
@@ -513,7 +509,7 @@ process_raw_climate <- function (climate_raw_path) {
       ),
       habit = fct_relevel(habit, c("epiphytic", "terrestrial")),
       site = str_remove_all(site, "_epi|_ter"),
-      vpd = plantecophys::RHtoVPD(RH, temp)
+      vpd = plantecophys::RHtoVPD(rh, temp)
     )
   
 }
@@ -548,6 +544,7 @@ process_community_matrix <- function (community_matrix_path, species_list, moore
 get_daily_means <- function (climate_raw) {
   climate_raw %>%
     group_by(site, date, habit) %>%
+    rename(RH = rh) %>%
     summarize_at(
       vars(temp, RH, vpd),
       list(

### 2020-01-08 code/plan.R

@@ -91,14 +91,24 @@ plan <- drake_plan(
     filter(str_detect(site, "Aorai", negate = TRUE)),
   
   ### Climate data ###
-  # - Process raw climate data (rel. humidity and temperature every 15 min.)
-  # - Also calculate vapor pressure deficit (VPD) from temp and RH
-  moorea_climate_raw = process_raw_climate(
-    file_in("data/hobo_moorea_aorai_2-24-15.csv"
-    )),
   
-  # - Calculate mean climate variables
-  daily_mean_climate = get_daily_means(moorea_climate_raw),
+  # - Read in raw climate data (rel. humidity and temperature) from all dataloggers 
+  # on Moorea and Tahiti from 2012-07-18 to 2015-02-06.
+  climate_raw = read_csv(file_in("data/hobo_moorea_aorai_2-24-15.csv")),
+  
+  # - Process raw climate data: 
+  # subset to only Moorea and days without missing data,
+  # replace one chunk of missing data with same dates from previous year,
+  # calculate vapor pressure deficit (VPD) from temp and RH.
+  moorea_climate = process_raw_climate(climate_raw),
+  
+  # - Write out moorea_climate as CSV file to include with Dryad data.
+  moorea_climate_out = moorea_climate %>%
+    select(date, time, site, habit, temp, rh, vpd) %>%
+    write_csv(file_out("ms/moorea_climate.csv")),
+  
+  # - Calculate mean climate variables.
+  daily_mean_climate = get_daily_means(moorea_climate),
   grand_mean_climate = get_grand_means(daily_mean_climate),
   
   ### Phylogeny ###
@@ -417,7 +427,7 @@ plan <- drake_plan(
 
   # Set color scheme: epiphytes in green, terrestrial in brown.
   habit_colors = brewer.pal(9, "Set1")[c(3,7)] %>%
-    set_names(levels(moorea_climate_raw$habit)),
+    set_names(levels(moorea_climate$habit)),
 
   # Make climate plot.
   climate_plot = make_climate_plot(

### 2020-01-08 ms/manuscript.Rmd
Update MS

@@
 Epiphytic ferns tend to have morphological features that minimize water loss from their leaves, including shorter stipes [@WatkinsJr.2010], thicker leaves [@Kluge2007; @WatkinsJr.2007b], and protection with scales or hairs [@WatkinsJr.2006b].
-Additional specialized adaptations to epiphytic growth include humus collection (e.g., *Asplenium*; @Benzing1990, *Drynaria*; @Janssen2005), vegetative desiccation tolerance (DT) (e.g., Hymenophyllaceae; @Shreve1911, *Pleopeltis*; @Pessin1925), CAM photosynthesis (e.g., *Pyrrosia*; @Wong1976), and symbiotic relationships with ants (e.g., *Antrophyum*; @WatkinsJr.2008, *Lecanopteris*; @Gay1993a).
+Additional specialized adaptations to epiphytic growth include humus collection (\eg, *Asplenium*; @Benzing1990, *Drynaria*; @Janssen2005), vegetative desiccation tolerance (\eg, Hymenophyllaceae; @Shreve1911, *Pleopeltis*; @Pessin1925), CAM photosynthesis (\eg, *Pyrrosia*; @Wong1976), and symbiotic relationships with ants (\eg, *Antrophyum*; @WatkinsJr.2008, *Lecanopteris*; @Gay1993a).
 
@@
 This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders; taxonomy follows @PteridophytePhylogenyGroupI2016; \tab S1).
-Ferns of Moorea include representatives of multiple Cretaceous epiphytic radiations (\eg, *Asplenium*, *Elaphoglossum*, Hymenophylloideae, Polypodiaceae, Vittarioideae; [@Schuettpelz2009]) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
+Ferns of Moorea include representatives of multiple Cretaceous epiphytic radiations (\eg, *Asplenium*, *Elaphoglossum*, Hymenophylloideae, Polypodiaceae, Vittarioideae; @Schuettpelz2009) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
 
@@
 Species were categorized as epiphytic or terrestrial based on field observations (see Trait Selection and Measurements below).
-The list of species in each plot was split into epiphytic and terrestrial groups and treated separately during data analysis, hereafter referred to as 'epiphytic' and 'terrestrial' communities.
-Vouchers were deposited at GH and UC (abbreviations follow Index Herbariorum [@Thiers2019]).
+The list of species in each plot was split into epiphytic and terrestrial groups and treated separately during data analysis, hereafter referred to as "epiphytic" and "terrestrial" communities.
+Vouchers were deposited at GH and UC (abbreviations follow Index Herbariorum; @Thiers2019).
 
@@
 Temperature and RH of epiphytic and terrestrial habitats were measured using Hobo Pro v2 dataloggers with the RS3 Solar Radiation Shield (Onset Corporation, USA).
-A pair of dataloggers was installed for each plot, one mounted at \circa \SI{2}{\meter} on a tree ('epiphytic datalogger'), and one at ground level ('terrestrial datalogger'), except for the Mt.
+A pair of dataloggers was installed for each plot, one mounted at \circa \SI{2}{\meter} on a tree ("epiphytic datalogger"), and one at ground level ("terrestrial datalogger"), except for the Mt.
 Tohiea \SI{1200}{\meter} plot, which only had an epiphytic datalogger.
@@
 We selected several traits relevant to epiphytic growth (\tab \@ref(tab:traits-used) and \fig \@ref(fig:traits-used-image)).
-We preferred morphological traits over physiological traits for our study because data for morphological traits are easier to gather at the scale of our study (\circa \SI{130}{\species} distributed over an elevational gradient from \SIrange{200}{2000}{\meter}).
+We preferred morphological traits over physiological traits because data for morphological traits are easier to gather at the scale of our study (\circa \SI{130}{\species} distributed over an elevational gradient from \SIrange{200}{2000}{\meter}).
 Plant morphology and physiology are closely linked, and physiological traits such as rates of photosynthesis and water use efficiency are known to be correlated with morphological traits such as stipe length in ferns [@WatkinsJr.2010].
@@
 Growth habit was coded as a binary trait (epiphytic or terrestrial) based on our field observations.
-Some species do not fall clearly into either category (\eg, hemi-epiphytes; [@Benzing1990; @Dubuisson2003; @Zotz2013]).
+Some species do not fall clearly into either category [\eg, hemi-epiphytes; @Benzing1990; @Dubuisson2003; @Zotz2013].
 We used connection to the soil as the criterion for defining binary growth habit; thus, epipetric plants were treated as epiphytic, and hemi-epiphytes and climbing plants were treated as terrestrial. 
@@
 We tested for differences between epiphytic and terrestrial dataloggers while accounting for elevation with ANCOVA, using growth habit as the categorical variable and elevation as the covariate.
-We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone and their interactions, then selecting the one with the lowest corrected Akaike Information Criterion (AIC~c~) [@Akaike1973].
+We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone and their interactions, then selecting the one with the lowest corrected Akaike Information Criterion [AIC~c~; @Akaike1973].
 A single outlier plot with unusually high temperature and low humidity (Mt. Rotui, 830 m plot) was excluded from ANCOVA and linear models.
@@
 
-We tested for phylogenetic signal in quantitative traits using Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
-$\lambda$ is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion [BM]).
+We tested for phylogenetic signal in quantitative traits using Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999], two metrics with slightly different interpretations and sensitivity to different evolutionary scenarios [@Munkemuller2012], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
+$\lambda$ is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion (BM)).
 \K describes the ratio between the amount of observed variance in traits vs.
@@
 We used distance-based methods to analyze community phylogenetic and functional diversity.
-Mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) describe the mean distance between all species in a community (MPD) or the mean distance between the closest pairs of species in a community (MNTD) [@Webb2000; @Webb2002].
+Mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) describe the mean distance between all species in a community (MPD) or the mean distance between the closest pairs of species in a community [MNTD; @Webb2000; @Webb2002].
 Although these metrics were originally formulated for phylogenetic distances, they can be applied in the same way to trait-based distances, enabling a direct comparison between the two [@Sessa2018].
@@
 We constructed null communities using the "independent swap" algorithm [@Gotelli2000].
-We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available (`r Ntip(phy)` \si{\species} total; [@Nitta2017]) for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct`) for functional null communities.
+We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available [`r Ntip(phy)` \si{\species} total; @Nitta2017] for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct`) for functional null communities.
 We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate phylogenetic and functional diversity metrics.
@@
 For each diversity metric, we built a set of models (Gaussian family, with plot as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
-Our dataset included `r nrow(div_metrics_all) - nrow(climate_select)` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
+Our dataset included `r (nrow(div_metrics_all) - nrow(climate_select)) %>% english2` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
 Therefore, we also constructed linear models for each diversity metric related to elevation including all sites using the same model selection procedure as in the climate analysis.
@@
 Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife.
-Data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED [@Nitta2019a].
+Data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED [@Nitta2020].
 
@@
 For instance, filmy ferns (Hymenophyllaceae), which depend on desiccation tolerance to survive dry conditions, have extremely thin, highly dissected fronds.
-Bird’s nest ferns (e.g., *Asplenium nidus* L.), in contrast, avoid desiccation by capturing detritus and water in the bases of their simple, rosette leaves. 
+Bird’s nest ferns (\eg, *Asplenium nidus* L.), in contrast, avoid desiccation by capturing detritus and water in the bases of their simple, rosette leaves. 
 At low-elevation sites on Moorea where these ferns co-occur, filmy ferns frequently grow in more humid portions of trees including near the main trunk, whereas bird’s nest ferns often grow higher in the canopy and on more exposed, drought-prone branches (J.
@@
 
-Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA) (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
+Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA; \fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
 Specific leaf area is part of the leaf economic spectrum (LES), which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
@@
 As with our results summarized above, it is possible that the diversity of gemmae and non-gemmae producing epiphytic species reflects undetected niche variation in the canopy.
-Interestingly, when we classified gametophytes as ‘independent’ if they occurred beyond the range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with independent growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
+Interestingly, when we classified gametophytes as ‘independent’ if they occurred beyond the elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with independent growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
 
@@
 
-Recent studies [e.g., @Sundue2015; @Testo2016a; @Lehtonen2017; @Testo2018a] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
+Recent studies [\eg, @Sundue2015; @Testo2016a; @Lehtonen2017; @Testo2018a] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
 Our study provides one possible explanation for this finding. 
@@
 Vapor pressure deficit (VPD) was calculated from relative humidity and temperature as described in the methods, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).
-'Epiphytic' dataloggers were placed at \circa 2 m on trees, and 'terrestrial' dataloggers placed at ground level.
+"Epiphytic" dataloggers were placed at \circa 2 m on trees, and "terrestrial" dataloggers placed at ground level.
 Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown.

### 2020-01-08 ms/manuscript.Rmd
Update MS
- Use some overly-complicated code to report same vs. different p-value for MPDfunc vs. MNTDfunc

@@
   mutate(stat = glue("{mean_rich} ± {sd_rich}"))
+
+# Do some special formatting for reporting rounded MPDfunc vs MNTDfunc P-val
+# - combine into a single value if same, or report each different value with 
+# "respectively" if different
+if(
+  magrittr::equals(
+    t_test_print %>%
+      filter(var == "mpd.obs.z.func") %>%
+      pull(p),
+    t_test_print %>%
+      filter(var == "mntd.obs.z.func") %>%
+      pull(p)
+  )
+) {
+  p_mpd_func_mntd_func <- get_mod_result(t_test_print, "mpd.obs.z.func", "p")
+  both <- "both "
+} else {
+  p_mpd_func_mntd_func <- paste0(
+    get_mod_result(t_test_print, "mpd.obs.z.func", "p"),
+    " and ",
+    get_mod_result(t_test_print, "mntd.obs.z.func", "p"),
+    ", respectively"
+  )
+  both <- ""
+}
 ```
@@
 We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial species.
-Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \si{\species}per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "ntaxa", "p")`; \fig \@ref(fig:comm-div)).
+Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \si{\species} per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "ntaxa", "p")`; \fig \@ref(fig:comm-div)).
 Species richness increased with elevation for both epiphytes and terrestrial communities (\fig \@ref(fig:comm-div)).
-Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, \pval `r get_mod_result(t_test_print, "mpd.obs.z.func", "p")` and `r get_mod_result(t_test_print, "mntd.obs.z.func", "p")`, respectively).
+Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, `r both`\pval `r p_mpd_func_mntd_func`).
 Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval `r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.

### 2020-01-08 ms/manuscript.Rmd
Use double quotes for R functions and packages
- Make all quotes throughout MS double

@@
 However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of $\K < 1$ (\tab \@ref(tab:phylosig-cont)).
-All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under BM (\tab \@ref(tab:phylosig-binary)).
+All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae, \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under BM (\tab \@ref(tab:phylosig-binary)).
 Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; \tab \@ref(tab:phylosig-binary)).
@@
 For instance, filmy ferns (Hymenophyllaceae), which depend on desiccation tolerance to survive dry conditions, have extremely thin, highly dissected fronds.
-Bird’s nest ferns (\eg, *Asplenium nidus* L.), in contrast, avoid desiccation by capturing detritus and water in the bases of their simple, rosette leaves. 
-At low-elevation sites on Moorea where these ferns co-occur, filmy ferns frequently grow in more humid portions of trees including near the main trunk, whereas bird’s nest ferns often grow higher in the canopy and on more exposed, drought-prone branches (J.
+Bird's nest ferns (\eg, *Asplenium nidus* L.), in contrast, avoid desiccation by capturing detritus and water in the bases of their simple, rosette leaves. 
+At low-elevation sites on Moorea where these ferns co-occur, filmy ferns frequently grow in more humid portions of trees including near the main trunk, whereas bird's nest ferns often grow higher in the canopy and on more exposed, drought-prone branches (J.
 Nitta, pers.
@@
 As with our results summarized above, it is possible that the diversity of gemmae and non-gemmae producing epiphytic species reflects undetected niche variation in the canopy.
-Interestingly, when we classified gametophytes as ‘independent’ if they occurred beyond the elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with independent growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
+Interestingly, when we classified gametophytes as "independent" if they occurred beyond the elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with independent growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
 

### 2020-01-09 ms/SI.rmd
Use bold subtitle for figure part labels
Same size as axis labels

@@
   scale_fill_scico(palette = "roma", breaks = c(-1, 1)) +
-  labs(x = NULL, y = NULL, title = "(a)") +
+  labs(x = NULL, y = NULL, subtitle = "(a)") +
   theme_minimal() +
@@
     panel.grid.major = element_blank(),
-    title = element_text(face = "bold"),
+    plot.subtitle = element_text(face = "bold"),
     axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
@@
   scale_fill_scico(palette = "roma", breaks = c(-1, 1)) +
-  labs(x = NULL, y = NULL, title = "(b)") +
+  labs(x = NULL, y = NULL, subtitle = "(b)") +
   theme_minimal() +

### 2020-01-09 code/functions.R

@@ -2161,7 +2161,7 @@ make_pca_plot <- function (pca_results, habit_colors, traits) {
     labs(
       x = get_label(axis_labels, "standard", "x"),
       y = get_label(axis_labels, "standard", "y"),
-      title = "(a)"
+      subtitle = "(a)"
     )
   
   # Phylogenetic trait PCA 
@@ -2183,7 +2183,7 @@ make_pca_plot <- function (pca_results, habit_colors, traits) {
     labs(
       x = get_label(axis_labels, "phylogenetic", "x"),
       y = get_label(axis_labels, "phylogenetic", "y"),
-      title = "(b)"
+      subtitle = "(b)"
     )
   
   c <-
@@ -2200,7 +2200,7 @@ make_pca_plot <- function (pca_results, habit_colors, traits) {
     labs(
       x = get_label(axis_labels, "standard", "x"),
       y = get_label(axis_labels, "standard", "y"),
-      title = "(c)"
+      subtitle = "(c)"
     )
   
   d <-
@@ -2217,7 +2217,7 @@ make_pca_plot <- function (pca_results, habit_colors, traits) {
     labs(
       x = get_label(axis_labels, "phylogenetic", "x"),
       y = get_label(axis_labels, "phylogenetic", "y"),
-      title = "(d)"
+      subtitle = "(d)"
     )
   
   # Combine into final plot with 0.2 in margins
@@ -2225,7 +2225,7 @@ make_pca_plot <- function (pca_results, habit_colors, traits) {
     standard_theme2() &
     theme(
       legend.position = "none",
-      plot.title = element_text(face = "bold"))
+      plot.subtitle = element_text(face = "bold"))
   
   p +
     plot_annotation(theme = theme(plot.margin = margin(
@@ -3006,12 +3006,12 @@ make_cwsd_scatterplot <- function(cwm_long, habit_colors) {
   
   # Add plot part labels
   cwsd_boxplots[["stipe"]] <- cwsd_boxplots[["stipe"]] +
-    labs(title = "(a)") +
-    theme(plot.title = element_text(face = "bold"))
+    labs(subtitle = "(a)") +
+    theme(plot.subtitle = element_text(face = "bold"))
   
   cwsd_boxplots[["dissection"]] <- cwsd_boxplots[["dissection"]] +
-    labs(title = "(b)") +
-    theme(plot.title = element_text(face = "bold"))
+    labs(subtitle = "(b)") +
+    theme(plot.subtitle = element_text(face = "bold"))
   
   # Combine plots into single output
   

### 2020-01-09 ms/manuscript.Rmd
Keep tex and render docx output

@@
     fig_caption: yes
-    keep_tex: no
+    keep_tex: yes
     latex_engine: xelatex

### 2020-01-09 code/functions.R

@@ -3132,6 +3132,42 @@ test_corr_evo_range_gemmae <- function(community_matrix_path, phy, traits, moore
   
 }
 
+# MS rendering
+
+# Dummy function to track arbitary output from rmarkdown::render()
+render_tracked <- function (tracked_output, ...) {
+  rmarkdown::render(...)
+}
+
+#' Convert latex file to docx
+#' 
+#' Requires pandoc to be installed and on command line
+#'
+#' @param latex Path to input latex file
+#' @param docx Path to output docx file
+#' @param wd Working directory to run conversion. Should be same as
+#' directory containing any files needed to render latex to pdf.
+#'
+#' @return List including STDOUT of pandoc; externally, the
+#' docx file will be rendered in `wd`.
+#' 
+latex2docx <- function (latex, docx, wd) {
+  
+  assertthat::assert_that(assertthat::is.readable(latex))
+  
+  assertthat::assert_that(assertthat::is.dir(fs::path_dir(docx)))
+  
+  latex <- fs::path_abs(latex)
+  
+  docx <- fs::path_abs(docx)
+  
+  processx::run(
+    command = "pandoc",
+    args = c("-s", latex, "-o", docx),
+    wd = wd
+  )
+  
+}
 
 # Misc ----
 

### 2020-01-09 code/plan.R

@@ -508,9 +508,20 @@ plan <- drake_plan(
   cwsd_scatterplot = make_cwsd_scatterplot(cwm_long, habit_colors),
 
   # Manuscript ----
-  ms = rmarkdown::render(
+  
+  # First render to PDF, keeping the latex
+  ms_pdf = render_tracked(
     knitr_in("ms/manuscript.Rmd"),
-    quiet = TRUE),
+    quiet = TRUE,
+    tracked_output = file_out(here::here("ms/manuscript.tex"))
+    ),
+  
+  # Next use the latex to convert to docx with word
+  ms_docx = latex2docx(
+    latex = file_in(here::here("ms/manuscript.tex")),
+    docx = file_out(here::here("ms/manuscript.docx")),
+    wd = here::here("ms")
+  ),
 
   # SI ----
   si = rmarkdown::render(

### 2020-01-10 ms/manuscript.Rmd
Add MS Word template file in style of New Phytologist

--- 2020-01-08 ms/manuscript.Rmd
@@
     fig_caption: yes
-    keep_tex: no
+    keep_tex: yes # need for converting tex to docx
     latex_engine: xelatex
@@
     includes:
-      in_header: newphyt.sty
+      in_header: new-phytologist.sty
 header-includes: 
@@
 
-<!-- FOR CONVERSION TO MS WORD ONLY: remove page numbers -->
-\pagenumbering{gobble}
-
-<!-- Set left justification (better conversion to Word than default full justification) -->
+<!-- Uncomment to remove page numbers -->
+<!-- \pagenumbering{gobble}  -->
+
+<!-- Set left justification  -->
 \raggedright

### 2020-01-10 code/functions.R

@@ -3143,15 +3143,16 @@ render_tracked <- function (tracked_output, ...) {
 #' 
 #' Requires pandoc to be installed and on command line
 #'
-#' @param latex Path to input latex file
-#' @param docx Path to output docx file
+#' @param latex Path to input latex file.
+#' @param docx Path to output docx file.
+#' @param template Path to template docx file.
 #' @param wd Working directory to run conversion. Should be same as
 #' directory containing any files needed to render latex to pdf.
 #'
 #' @return List including STDOUT of pandoc; externally, the
 #' docx file will be rendered in `wd`.
 #' 
-latex2docx <- function (latex, docx, wd) {
+latex2docx <- function (latex, docx, template = NULL, wd = getwd()) {
   
   assertthat::assert_that(assertthat::is.readable(latex))
   
@@ -3161,9 +3162,15 @@ latex2docx <- function (latex, docx, wd) {
   
   docx <- fs::path_abs(docx)
   
+  template <- if (!is.null(template)) {
+    glue::glue("--reference-doc={fs::path_abs(template)}")
+  } else {
+    NULL
+  }
+  
   processx::run(
     command = "pandoc",
-    args = c("-s", latex, "-o", docx),
+    args = c("-s", latex, template, "-o", docx),
     wd = wd
   )
   

### 2020-01-10 code/plan.R

@@ -520,6 +520,7 @@ plan <- drake_plan(
   ms_docx = latex2docx(
     latex = file_in(here::here("ms/manuscript.tex")),
     docx = file_out(here::here("ms/manuscript.docx")),
+    template = file_in(here::here("ms/new-phytologist.docx")),
     wd = here::here("ms")
   ),
 

### 2020-01-10 ms/manuscript.Rmd
Use unicode lambda instead of latex

@@
 
-We tested for phylogenetic signal in quantitative traits using Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999], two metrics with slightly different interpretations and sensitivity to different evolutionary scenarios [@Munkemuller2012], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
-$\lambda$ is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion (BM)).
+We tested for phylogenetic signal in quantitative traits using Blomberg's *K* [@Blomberg2003] and Pagel's λ [@Pagel1999], two metrics with slightly different interpretations and sensitivity to different evolutionary scenarios [@Munkemuller2012], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
+λ is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion (BM)).
 \K describes the ratio between the amount of observed variance in traits vs.
-the amount of variance expected under BM: when $\K > 1$, traits are more conserved than expected under BM; when $\K < 1$, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
-We tested for significance by comparing observed \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in $\lambda$ by conducting a likelihood ratio test comparing the log likelihoods of the observed value of $\lambda$ vs.
-$\lambda = 0$ (no phylogenetic signal).
+the amount of variance expected under BM: when \K > 1, traits are more conserved than expected under BM; when \K < 1, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
+We tested for significance by comparing observed \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in λ by conducting a likelihood ratio test comparing the log likelihoods of the observed value of λ vs.
+λ = 0 (no phylogenetic signal).
 We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' *D* [@Fritz2010].
@@
 
-Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs \@ref(tab:phylosig-cont), \@ref(tab:phylosig-binary), \fig \@ref(fig:tree)), and for quantitative traits, different results were obtained for $\lambda$ and K.
-When measured with $\lambda$, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of $\lambda$ close to 1; only number of pinnae pairs had $\lambda$ close to zero (\tab \@ref(tab:phylosig-cont)).
-However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of $\K < 1$ (\tab \@ref(tab:phylosig-cont)).
+Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs \@ref(tab:phylosig-cont), \@ref(tab:phylosig-binary), \fig \@ref(fig:tree)), and for quantitative traits, different results were obtained for λ and K.
+When measured with λ, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of λ close to 1; only number of pinnae pairs had λ close to zero (\tab \@ref(tab:phylosig-cont)).
+However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of \K < 1 (\tab \@ref(tab:phylosig-cont)).
 All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae, \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under BM (\tab \@ref(tab:phylosig-binary)).
@@
   footnote(general = "K, Blomberg's \\\\textit{K} (Blomberg \\\\textit{et al.}, 2003); lambda, Pagel's lambda (Pagel, 1999).
-Values of lambda or \\\\textit{K} greater than 1 indicate traits evolving according to Brownian Motion (BM); values of lambda or K near 0 indicate random distribution of traits.", general_title = "", threeparttable = TRUE, escape = FALSE)
+Values of lambda or \\\\textit{K} greater than 1 indicate traits evolving according to Brownian Motion (BM); values of lambda or \\\\textit{K} near 0 indicate random distribution of traits.", general_title = "", threeparttable = TRUE, escape = FALSE)
 ```

### 2020-01-10 ms/manuscript.Rmd
Use captioner for figure captions

@@
 library(tidyverse)
+library(captioner)
 library(conflicted)
@@
   cache = canopy_cache)
+
+# Set up captions
+fig_nums <- captioner(prefix = "Fig.")
+fig_nums(name = "traits-used-image", caption = "Examples of traits used in this study.")
+fig_nums(name = "climate", caption = "Microclimatic variables.")
+fig_nums(name = "pca", caption = "Trait PCA.")
+fig_nums(name = "tree", caption = "Traits on tree.")
+fig_nums(name = "heatmap", caption = "Variable importance scores.")
+fig_nums(name = "cwm-div", caption = "Community-weighted means.")
+fig_nums(name = "comm-div", caption = "Community diversity metrics.")
+
+# Make short version of citation function for figs
+citef <- pryr::partial(fig_nums, display = "cite")
+
 ```
@@
 
-We selected several traits relevant to epiphytic growth (\tab \@ref(tab:traits-used) and \fig \@ref(fig:traits-used-image)).
+We selected several traits relevant to epiphytic growth (\tab \@ref(tab:traits-used) and `r citef("traits-used-image")`).
 We preferred morphological traits over physiological traits because data for morphological traits are easier to gather at the scale of our study (\circa \SI{130}{\species} distributed over an elevational gradient from \SIrange{200}{2000}{\meter}).
@@
 For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie, heart-shaped) or non-cordate.
-Non-cordate morphotypes include elongate (strap or ribbon \sensu @Farrar2008) and filamentous forms (\fig \@ref(fig:traits-used-image)).
+Non-cordate morphotypes include elongate (strap or ribbon \sensu @Farrar2008) and filamentous forms (`r citef("traits-used-image")`).
 Gametophytes were cultivated on Bold's media [@Bold1957] supplemented with Nitch's micronutrients [@Nitsch1951].
@@
 Mean, maximum, and standard deviation (SD) of temperature and mean, minimum, and SD of vapor pressure deficit (VPD) were retained after excluding variables with correlation coefficients > 0.9.
-Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (\tab S3; \fig \ref{fig:climate}).
+Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (\tab S3; `r citef("climate")`).
 Variables related to VPD differed by growth habit and elevation.
@@
 
-Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (\fig \@ref(fig:pca)*c*, *d*).
+Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (`r citef("pca")`**c**, **d**).
 The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
@@
 
-Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs \@ref(tab:phylosig-cont), \@ref(tab:phylosig-binary), \fig \@ref(fig:tree)), and for quantitative traits, different results were obtained for λ and K.
+Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs \@ref(tab:phylosig-cont), \@ref(tab:phylosig-binary), `r citef("tree")`), and for quantitative traits, different results were obtained for λ and K.
 When measured with λ, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of λ close to 1; only number of pinnae pairs had λ close to zero (\tab \@ref(tab:phylosig-cont)).
@@
 
-In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (\fig \ref{fig:heatmap}).
+In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (`r citef("heatmap")`).
 Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (\tab S6).
@@
 
-Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval < 0.05; \fig \@ref(fig:cwm-div)).
+Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval < 0.05; `r citef("cwm-div")`).
 Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval `r get_mod_result(model_summ_print, "sla", "p")`).
 Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval = `r get_mod_result(model_summ_print, "dissection", "p")`).
-Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (\fig \@ref(fig:cwm-div)).
+Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r citef("cwm-div")`).
 
 We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial species.
-Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \si{\species} per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "ntaxa", "p")`; \fig \@ref(fig:comm-div)).
-Species richness increased with elevation for both epiphytes and terrestrial communities (\fig \@ref(fig:comm-div)).
+Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \si{\species} per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "ntaxa", "p")`; `r citef("comm-div")`).
+Species richness increased with elevation for both epiphytes and terrestrial communities (`r citef("comm-div")`).
 Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, `r both`\pval `r p_mpd_func_mntd_func`).
 Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval `r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval `r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval = `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; \fig \@ref(fig:comm-div)).
-No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (\fig \@ref(fig:comm-div)).
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval `r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval = `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; `r citef("comm-div")`).
+No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (`r citef("comm-div")`).
 
@@
 Our study reveals a striking duality in the phylogenetic and functional diversity of fern epiphytes. 
-On one hand, we find that fern epiphytes are consistently smaller than terrestrial species at both the species and community level (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)), and that epiphytic communities are phylogenetically clustered (\fig \@ref(fig:comm-div)).
-At the same time, we find that epiphytic fern communities are significantly more diverse functionally compared to terrestrial communities (\fig \@ref(fig:comm-div)).
+On one hand, we find that fern epiphytes are consistently smaller than terrestrial species at both the species and community level (`r citef("cwm-div")`, \tab \@ref(tab:pics)), and that epiphytic communities are phylogenetically clustered (`r citef("comm-div")`).
+At the same time, we find that epiphytic fern communities are significantly more diverse functionally compared to terrestrial communities (`r citef("comm-div")`).
 Collectively, these results suggest that environmental filtering plays a strong role in limiting plant size, while simultaneously promoting diversification across other trait axes in epiphytic ferns.
@@
 
-Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (\fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
+Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (`r citef("cwm-div")`, \tab \@ref(tab:pics)), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
 These results are corroborated by @Testo2018a, who similarly observed abrupt size reduction associated with epiphytic growth in a globally sampled fern phylogeny. 
@@
 
-Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA; \fig \@ref(fig:cwm-div), \tab \@ref(tab:pics)).
+Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA; `r citef("cwm-div")`, \tab \@ref(tab:pics)).
 Specific leaf area is part of the leaf economic spectrum (LES), which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
@@
 We expected that epiphytes would exhibit lower SLA, given their more extreme growing conditions.
-In contrast, we observed overlapping, wide ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes (\fig \@ref(fig:cwm-div)).
+In contrast, we observed overlapping, wide ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes (`r citef("cwm-div")`).
 This may reflect adaptations to specific niches that are not captured in the binary categorization of habitat commonly applied in ferns. 
@@
 We did not find that gemmae production in gametophytes was linked with epiphytism. 
-Rather, occurrence of gemmae strongly reflects phylogeny, appearing concentrated in filmy ferns (Hymenophyllaceae) and vittarioid ferns (\fig \@ref(fig:tree)).
+Rather, occurrence of gemmae strongly reflects phylogeny, appearing concentrated in filmy ferns (Hymenophyllaceae) and vittarioid ferns (`r citef("tree")`).
 Gemmae are asexual propagules occurring primarily in non-cordate gametophytes that allow gametophytes to persist over multiple growing seasons and attain large population sizes, and have been hypothesized to be associated with harsh environmental conditions [@Farrar2008].
@@
 On the one hand, the canopy appears to have served as a potent barrier to entry, strongly selecting for small size. 
-Furthermore, the results of our phylogenetic community structure analyses partially support the hypothesis of decreased phylogenetic diversity in epiphytic communities due to filtering: we observed that epiphytic fern communities were more phylogenetically clustered for MNTD~phy~, but not MPD~phy~ (\fig \@ref(fig:comm-div)).
+Furthermore, the results of our phylogenetic community structure analyses partially support the hypothesis of decreased phylogenetic diversity in epiphytic communities due to filtering: we observed that epiphytic fern communities were more phylogenetically clustered for MNTD~phy~, but not MPD~phy~ (`r citef("comm-div")`).
 This makes sense given that most epiphytes have radiated within a limited number of clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely related species from two or more epiphytic clades, but exhibits a similar overall degree of phylogenetic diversity as terrestrial communities.
@@
 However, while the canopy clearly acts as a filter, we also find that it simultaneously functions as a promoter of diversity. 
-This is evident in our analyses of MPD~func~ and MNTD~func~, which both indicate that epiphytic communities are more functionally diverse compared to terrestrial species (\fig \@ref(fig:comm-div)), thus contradicting our hypothesis of decreased functional diversity in epiphytic communities.
+This is evident in our analyses of MPD~func~ and MNTD~func~, which both indicate that epiphytic communities are more functionally diverse compared to terrestrial species (`r citef("comm-div")`), thus contradicting our hypothesis of decreased functional diversity in epiphytic communities.
 Additional analyses using different trait combinations support the role of non-size related traits driving this pattern (\fig S1).
@@
 
-The primary role of growth habit, in combination with temperature, in structuring fern communities is supported by the full-subsets model analysis (\fig \ref{fig:heatmap}).
+The primary role of growth habit, in combination with temperature, in structuring fern communities is supported by the full-subsets model analysis (`r citef("heatmap")`).
 It is somewhat puzzling that VPD did not contribute significantly to any models, as we hypothesized that this would be important for epiphytes that are frequently subjected to water stress. 
 Previous studies have shown marked variation in the air water potential across epiphytic habitats, with @WatkinsJr.2009 reporting a \circa \SI{40}{\mega\pascal} shift along the trunks of emergent canopy trees.
-Our epiphytic dataloggers only showed slightly greater VPD relative to those on the ground, with this difference largely dissipating at high elevations (\fig \ref{fig:climate}).
+Our epiphytic dataloggers only showed slightly greater VPD relative to those on the ground, with this difference largely dissipating at high elevations (`r citef("climate")`).
 One reason for such low observed variation could be due to environmental conditions at our study site and placement of dataloggers: Moorea is a generally wet site overall (average annual rainfall \SI{325}{\centi\meter}), and we placed the epiphytic dataloggers on tree trunks, not in the higher emergent portions of the canopy.
@@
 
-**\fig \@ref(fig:traits-used-image)**: Examples of traits used in this study.
+**`r citef("traits-used-image")`**: Examples of traits used in this study.
 **(a)** Sporophyte traits.
@@
 
-**\fig \@ref(fig:climate)**: Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter on Moorea, French Polynesia.
+**`r citef("climate")`**: Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter on Moorea, French Polynesia.
 Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from `r final_climate_survey$start` to `r final_climate_survey$end`.
@@
 
-**\fig \@ref(fig:pca)**: Principal components analysis (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
+**`r citef("pca")`**: Principal components analysis (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
 **(a, b)** PC loadings.
@@
 
-**\fig \@ref(fig:tree)**: Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
+**`r citef("tree")`**: Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
 Tree obtained from @Nitta2017.
@@
 
-**\fig \@ref(fig:heatmap)**: Variable importance scores from a full-subsets analysis using general additive models exploring the effect of climate and growth habit (terrestrial vs.
+**`r citef("heatmap")`**: Variable importance scores from a full-subsets analysis using general additive models exploring the effect of climate and growth habit (terrestrial vs.
 epiphytic growth) on community diversity of ferns on Moorea, French Polynesia.
@@
 
-**\fig \@ref(fig:cwm-div)**: Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
+**`r citef("cwm-div")`**: Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
 Epiphytic communities in green, terrestrial communities in brown.
@@
 
-**\fig \@ref(fig:comm-div)**: Functional and phylogenetic diversity of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
+**`r citef("comm-div")`**: Functional and phylogenetic diversity of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
 Epiphytic communities in green, terrestrial communities in brown.
@@
 
-```{r traits-used-image, fig.height = 9, fig.width = 6, fig.cap = "Examples of traits used in this study."}
+```{r traits-used-image, fig.height = 9, fig.width = 6}
 knitr::include_graphics(here::here("images/Fig_1_traits.pdf"))
@@
 
-\clearpage
-
-```{r climate, fig.height = 4, fig.width = 7, fig.cap = "Microclimatic variables."}
+`r fig_nums("traits-used-image")`
+
+\clearpage
+
+```{r climate, fig.height = 4, fig.width = 7}
 climate_plot
@@
 
-\clearpage
-
-```{r pca, fig.height = 6, fig.width = 6, fig.cap = "Trait PCA."}
+`r fig_nums("climate")`
+
+\clearpage
+
+```{r pca, fig.height = 6, fig.width = 6}
 pca_plot
@@
 
-\clearpage
-
-```{r tree, fig.height = 8, fig.width = 6, fig.cap = "Traits on tree."}
+`r fig_nums("pca")`
+
+\clearpage
+
+```{r tree, fig.height = 8, fig.width = 6}
 traits_with_tree
@@
 
-\clearpage
-
-```{r heatmap, fig.height = 4, fig.width = 6, fig.cap = "Variable importance scores."}
+`r fig_nums("tree")`
+
+\clearpage
+
+```{r heatmap, fig.height = 4, fig.width = 6}
 importance_heatmap
@@
 
-\clearpage
-
-```{r cwm-div, fig.height = 4, fig.width = 7, fig.cap = "Community-weighted means."}
+`r fig_nums("heatmap")`
+
+\clearpage
+
+```{r cwm-div, fig.height = 4, fig.width = 7}
 combined_cwm_plots
@@
 
-\clearpage
-
-```{r comm-div, fig.height = 8, fig.width = 4, fig.cap = "Community diversity metrics."}
+`r fig_nums("cwm-div")`
+
+\clearpage
+
+```{r comm-div, fig.height = 8, fig.width = 4}
 combined_comm_div_plots
@@
 ```
+
+`r fig_nums("comm-div")`

### 2020-01-10 ms/manuscript.Rmd
Use captioner for table and SI captions

@@
 header-includes: 
-  \usepackage{float}
-  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
+  - \usepackage{float}
+  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # Fix position of figs
+  - \makeatletter\renewcommand*{\fps@table}{H}\makeatother # Fix position of tables
+  - \usepackage{caption} # Needed to turn off automatic numbering of table captions
 editor_options: 
@@
 # Set up captions
+# - Figures
 fig_nums <- captioner(prefix = "Fig.")
@@
 
-# Make short version of citation function for figs
+# - Tables
+table_nums <- captioner(prefix = "Table")
+table_nums(name = "traits-used", caption = "Traits used in this study")
+table_nums(name = "phylosig-cont", caption = "Phylogenetic signal in continuous traits")
+table_nums(name = "phylosig-binary", caption = "Phylogenetic signal in binary traits")
+table_nums(name = "pics", caption = "Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth")
+table_nums(name = "corr-evo", caption = "Pagel's (1994) test of correlated evolution between binary traits")
+
+# - SI figures
+s_fig_nums <- captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
+s_fig_nums(name = "corr-plot", "Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.")
+s_fig_nums(name = "cwsd", "Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.")
+
+# - SI tables
+s_table_nums <- captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")
+s_table_nums(name = "species-list", "List of ferns from Moorea, French Polynesia included in this study")
+s_table_nums(name = "trait-values", "Trait values of ferns on Moorea, French Polynesia")
+s_table_nums(name = "ancova", "Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate")
+s_table_nums(name = "phylosig-binary-strict", "Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy")
+s_table_nums(name = "corr-evo-strict", "Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy")
+s_table_nums(name = "model-fit", "Best-fitting models explaining community diversity metrics by climate and growth habit")
+
+# SI figures
+
+# Make short versions of citation functions
 citef <- pryr::partial(fig_nums, display = "cite")
-
-```
+citet <- pryr::partial(table_nums, display = "cite")
+citesf <- pryr::partial(s_fig_nums, display = "cite")
+citest <- pryr::partial(s_table_nums, display = "cite")
+
+```
+
+<!-- Turn off automatic numbering of table captions -->
+\captionsetup[table]{labelformat=empty}
 
@@
 Moorea, French Polynesia (\ang{17;30;}S, \ang{149;50;}W) is a small (\SI{135}{\kilo\meter\squared}) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
-This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders; taxonomy follows @PteridophytePhylogenyGroupI2016; \tab S1).
+This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders; taxonomy follows @PteridophytePhylogenyGroupI2016; `r citest("species-list")`).
 Ferns of Moorea include representatives of multiple Cretaceous epiphytic radiations (\eg, *Asplenium*, *Elaphoglossum*, Hymenophylloideae, Polypodiaceae, Vittarioideae; @Schuettpelz2009) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
@@
 
-We selected several traits relevant to epiphytic growth (\tab \@ref(tab:traits-used) and `r citef("traits-used-image")`).
+We selected several traits relevant to epiphytic growth (`r citet("traits-used")` and `r citef("traits-used-image")`).
 We preferred morphological traits over physiological traits because data for morphological traits are easier to gather at the scale of our study (\circa \SI{130}{\species} distributed over an elevational gradient from \SIrange{200}{2000}{\meter}).
@@
 
-Morphological traits of gametophytes were all qualitative (either categorical or binary; \tab \@ref(tab:traits-used)).
+Morphological traits of gametophytes were all qualitative (either categorical or binary; `r citet("traits-used")`).
 Traits were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017] or grown from spores of known species in the lab, taken from the literature, or inferred from taxonomic group (i.e., genus) if documented as a stable taxonomic character.
@@
 
-Trait values and data sources are summarized in Table S2.
+Trait values and data sources are summarized in `r citest("trait-values")`.
 
@@
 Mean, maximum, and standard deviation (SD) of temperature and mean, minimum, and SD of vapor pressure deficit (VPD) were retained after excluding variables with correlation coefficients > 0.9.
-Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (\tab S3; `r citef("climate")`).
+Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (`r citest("ancova")`; `r citef("climate")`).
 Variables related to VPD differed by growth habit and elevation.
@@
 
-Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs \@ref(tab:phylosig-cont), \@ref(tab:phylosig-binary), `r citef("tree")`), and for quantitative traits, different results were obtained for λ and K.
-When measured with λ, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of λ close to 1; only number of pinnae pairs had λ close to zero (\tab \@ref(tab:phylosig-cont)).
-However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of \K < 1 (\tab \@ref(tab:phylosig-cont)).
-All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae, \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under BM (\tab \@ref(tab:phylosig-binary)).
-Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; \tab \@ref(tab:phylosig-binary)).
-Similar results were obtained using the "strict" dataset (\tab S4).
+Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs `r table_nums("phylosig-cont", display = "n")`, `r table_nums("phylosig-binary", display = "n")`, `r citef("tree")`), and for quantitative traits, different results were obtained for λ and K.
+When measured with λ, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of λ close to 1; only number of pinnae pairs had λ close to zero (`r citet("phylosig-cont")`).
+However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of \K < 1 (`r citet("phylosig-cont")`).
+All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae, \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under BM (`r citet("phylosig-binary")`).
+Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; `r citet("phylosig-binary")`).
+Similar results were obtained using the "strict" dataset (`r citest("phylosig-binary-strict")`).
 
@@
 
-Most sporophyte traits, including all those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (\tab \@ref(tab:pics)).
-Epiphytic species tend to have smaller values for all traits (\tab \@ref(tab:pics); contrasts were made by subtracting epiphytic from terrestrial values).
+Most sporophyte traits, including all those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (`r citet("pics")`).
+Epiphytic species tend to have smaller values for all traits (`r citet("pics")`; contrasts were made by subtracting epiphytic from terrestrial values).
 Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval `r stipe_pval`), as well as frond width (\pval `r width_pval`) and degree of laminae dissection (\pval `r dissection_pval`).
 For the gametophyte traits, only morphotype (non-cordate vs.
-cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval `r morphotype_pval`; \tab \@ref(tab:corr-evo)).
-Similar results were obtained using the "strict" dataset (\tab S5).
+cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval `r morphotype_pval`; `r citet("corr-evo")`).
+Similar results were obtained using the "strict" dataset (`r citest("corr-evo-strict")`).
 
@@
 In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (`r citef("heatmap")`).
-Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (\tab S6).
-Spatial autocorrelation was not detected for any of the diversity metrics (\tab S6).
+Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (`r citest("model-fit")`).
+Spatial autocorrelation was not detected for any of the diversity metrics (`r citest("model-fit")`).
 
@@
 Our study reveals a striking duality in the phylogenetic and functional diversity of fern epiphytes. 
-On one hand, we find that fern epiphytes are consistently smaller than terrestrial species at both the species and community level (`r citef("cwm-div")`, \tab \@ref(tab:pics)), and that epiphytic communities are phylogenetically clustered (`r citef("comm-div")`).
+On one hand, we find that fern epiphytes are consistently smaller than terrestrial species at both the species and community level (`r citef("cwm-div")`, `r citet("pics")`), and that epiphytic communities are phylogenetically clustered (`r citef("comm-div")`).
 At the same time, we find that epiphytic fern communities are significantly more diverse functionally compared to terrestrial communities (`r citef("comm-div")`).
@@
 
-Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (`r citef("cwm-div")`, \tab \@ref(tab:pics)), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
+Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (`r citef("cwm-div")`, `r citet("pics")`), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
 These results are corroborated by @Testo2018a, who similarly observed abrupt size reduction associated with epiphytic growth in a globally sampled fern phylogeny. 
@@
 These results do not support the hypothesis that fern epiphytes have reduced functional diversity.
-For example, although epiphytes demonstrated less dissected fronds at the species level (\tab \@ref(tab:pics)), they exhibited much greater variation in this trait relative to terrestrial ferns at the community level (\fig S2).
+For example, although epiphytes demonstrated less dissected fronds at the species level (`r citet("pics")`), they exhibited much greater variation in this trait relative to terrestrial ferns at the community level (`r citesf("cwsd")`).
 It is likely that the high diversity of this trait reflects different physiological adaptations to various canopy niches. 
@@
 
-Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA; `r citef("cwm-div")`, \tab \@ref(tab:pics)).
+Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA; `r citef("cwm-div")`, `r citet("pics")`).
 Specific leaf area is part of the leaf economic spectrum (LES), which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
@@
 (a terrestrial mangrove specialist), and *Paesia divaticatissima* (Dryand.) Copel.
-(a terrestrial species occurring on high-elevation ridges) (Table S2).
+(a terrestrial species occurring on high-elevation ridges) (`r citest("trait-values")`).
 Each of these species is likely to experience prolonged periods of drought-stress. 
@@
 
-Finally, the presence of hairs and glands in gametophytes also exhibit very high phylogenetic signal (\ie, much greater than expected under Brownian motion; \tab \@ref(tab:phylosig-binary)), and these traits are concentrated in certain taxonomic groups [\eg, hairs, grammitid ferns; glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
+Finally, the presence of hairs and glands in gametophytes also exhibit very high phylogenetic signal (\ie, much greater than expected under Brownian motion; `r citet("phylosig-binary")`), and these traits are concentrated in certain taxonomic groups [\eg, hairs, grammitid ferns; glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
 Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], and glands may affect rates of osmosis by modifying lipid concentrations [@Crow2011], we did not detect a correlation between hair or gland production and epiphytic growth.
@@
 This is evident in our analyses of MPD~func~ and MNTD~func~, which both indicate that epiphytic communities are more functionally diverse compared to terrestrial species (`r citef("comm-div")`), thus contradicting our hypothesis of decreased functional diversity in epiphytic communities.
-Additional analyses using different trait combinations support the role of non-size related traits driving this pattern (\fig S1).
+Additional analyses using different trait combinations support the role of non-size related traits driving this pattern (`r citesf("corr-plot")`).
 We propose that the high functional diversity of epiphytic fern communities is maintained by previously undetected niche partitioning within epiphytic habitats as follows.
@@
 
-**Table S1** List of ferns from Moorea, French Polynesia included in this study.
-
-**Table S2** Trait values of ferns on Moorea, French Polynesia.
-
-**Table S3** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate.
-
-**Table S4** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
-
-**Table S5** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
-
-**Table S6** Best-fitting models explaining community diversity metrics by climate and growth habit.
-
-**Fig. S1** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
-
-**Fig. S2** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
+**`r citest("species-list")`** List of ferns from Moorea, French Polynesia included in this study.
+
+**`r citest("trait-values")`** Trait values of ferns on Moorea, French Polynesia.
+
+**`r citest("ancova")`** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate.
+
+**`r citest("phylosig-binary-strict")`** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+
+**`r citest("corr-evo-strict")`** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+
+**`r citest("model-fit")`** Best-fitting models explaining community diversity metrics by climate and growth habit.
+
+**`r citesf("corr-plot")`** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
+
+**`r citesf("cwsd")`** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
 
@@
 Species missing data for six or more traits not shown.
-For a summary of traits, see \tab \@ref(tab:traits-used).
+For a summary of traits, see `r citet("traits-used")`.
 Major epiphytic radiations identified by @Schuettpelz2009 labeled on tree in green: H, Hymenophylloideae; T, Trichomanoideae; V, Vittarioideae; A, *Asplenium*; E, *Elaphoglossum*; P, Polypodiaceae; taxonomy follows @PteridophytePhylogenyGroupI2016.
@@
 \clearpage
+
+**Table `r table_nums("traits-used", display = "n")`** Traits used in this study
 
@@
   column_to_rownames("trait") %>%
-  knitr::kable(caption = "Traits used in this study", format = "latex", booktabs = TRUE, escape = TRUE) %>%
+  knitr::kable(caption = "", format = "latex", booktabs = TRUE, escape = TRUE) %>%
   kable_styling(full_width = FALSE) %>%
@@
 \clearpage
+
+**Table `r table_nums("phylosig-cont", display = "n")`** Phylogenetic signal in continuous traits
 
@@
   column_to_rownames("trait") %>%
-  kable(caption = "Phylogenetic signal in continuous traits", format = "latex", booktabs = TRUE) %>%
+  kable(caption = "", format = "latex", booktabs = TRUE) %>%
   footnote(general = "K, Blomberg's \\\\textit{K} (Blomberg \\\\textit{et al.}, 2003); lambda, Pagel's lambda (Pagel, 1999).
@@
 \clearpage
+
+**Table `r table_nums("phylosig-binary", display = "n")`** Phylogenetic signal in binary traits
 
@@
 column_to_rownames("trait") %>%
-  kable(caption = "Phylogenetic signal in binary traits", format = "latex", booktabs = TRUE) %>%
+  kable(caption = "", format = "latex", booktabs = TRUE) %>%
   footnote(general = "D, Fritz and Purvis’ \\\\textit{D} (Fritz \\\\& Purvis, 2010); P(rnd), probability of random distribution of traits; P(bm), probability of traits evolving under Brownian motion (BM).
@@
 \clearpage
+
+**Table `r table_nums("pics", display = "n")`** Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth
 
@@
   column_to_rownames("trait") %>%
-  kable(caption = "Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth", format = "latex", booktabs = TRUE) %>%
+  kable(caption = "", format = "latex", booktabs = TRUE) %>%
   footnote(general = "All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades.", general_title = "", threeparttable = TRUE)
@@
 \clearpage
+
+**Table `r table_nums("corr-evo", display = "n")`** Pagel's (1994) test of correlated evolution between binary traits
 
@@
   column_to_rownames("trait") %>%
-  kable(caption = "Pagel's (1994) test of correlated evolution between binary traits", format = "latex", booktabs = TRUE) %>%
+  kable(caption = "", format = "latex", booktabs = TRUE) %>%
   footnote(general = "LogL, log-likelihood.

### 2020-01-11 ms/manuscript.Rmd
Use latex (mostly via gt) for all tables

@@
 
-# Read in csv version of the table
-traits_used <- read_csv(here::here("data/table_1.csv"))
-
-# Write out table in html for submitting
-# MS as word doc
-traits_used %>%
-  gt(rowname_col = "trait") %>%
-  fmt_markdown(columns = everything()) %>%
-  tab_row_group(
-    group = "Sporophyte traits",
-    rows = matches("Stipe|Frond|Rhizome|Pinna|Leaf")
-  ) %>%
-  tab_row_group(
-    group = "Gametophyte traits",
-    rows = matches("Morphotype|Gemmae|Glands|Hairs")
-  ) %>%
-  cols_label(
-    data_type = "Data type (unit)",
-    ecological_significance = "Ecological Significance",
-    reference = "Reference(s)"
-  ) %>% 
-  gtsave(here::here("ms/Table_1.html"))
-
-# Include in PDF using kableExtra to convert to latex format
-traits_used %>% 
+# Use kable instead of gt because gt can't control table width,
+# and the "ecological significance" descriptions would run off the page in PDF.
+
+# FIXME: need to manually add "Sporophyte traits" and "Gametophyte traits" to
+# table in docx (this happens with gt or kable).
+read_csv(here::here("data/table_1.csv")) %>%
   rename(`Data type (unit)` = data_type, `Functional significance` = ecological_significance, `Reference(s)` = reference) %>%
   column_to_rownames("trait") %>%
-  knitr::kable(caption = "", format = "latex", booktabs = TRUE, escape = TRUE) %>%
+  knitr::kable(format = "latex", booktabs = TRUE, escape = TRUE) %>%
   kable_styling(full_width = FALSE) %>%
@@
   column_spec(3, width = "12em") %>%
-  column_spec(4, width = "8em") %>%
-  footnote(general = "Noncordate gametophyte morphotypes include ribbon, strap and filamentous.", general_title = "", threeparttable = TRUE, escape = TRUE)
-```
+  column_spec(4, width = "8em")
+```
+
+Noncordate gametophyte morphotypes include ribbon, strap and filamentous.
 
@@
 
-# Format the raw results for table presentation
-phylosig_cont_traits_table <-
 phylosig_cont_traits %>%
@@
                    "SLA", "Pinna no.", "Frond dissection"))) %>%
-  arrange(trait)
-
-# Write out table in html format
-phylosig_cont_traits_table %>%
+  arrange(trait) %>%
   gt(rowname_col = "trait") %>%
@@
   )  %>%
-  gtsave(here::here("ms/Table_2.html"))
-
-# Include in PDF in latex format
-phylosig_cont_traits_table %>%
-  rename(K = kval, `P(K)` = k_pval, `P(lambda)` = lambda_pval) %>%
-  column_to_rownames("trait") %>%
-  kable(caption = "", format = "latex", booktabs = TRUE) %>%
-  footnote(general = "K, Blomberg's \\\\textit{K} (Blomberg \\\\textit{et al.}, 2003); lambda, Pagel's lambda (Pagel, 1999).
-Values of lambda or \\\\textit{K} greater than 1 indicate traits evolving according to Brownian Motion (BM); values of lambda or \\\\textit{K} near 0 indicate random distribution of traits.", general_title = "", threeparttable = TRUE, escape = FALSE)
-```
+  as_latex
+
+```
+
+\K, Blomberg's *K* [@Blomberg2003]; lambda, Pagel's lambda [@Pagel1999].
+Values of lambda or \K greater than 1 indicate traits evolving according to Brownian Motion (BM); values of lambda or \K near 0 indicate random distribution of traits.
 
@@
 
-# Format the raw results for table presentation
-phylosig_binary_traits_table <-
 phylosig_binary_traits %>%
@@
     new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
-  arrange(trait)
-
-# Write out table in html
-phylosig_binary_traits_table %>%
+  arrange(trait) %>%
   gt(rowname_col = "trait") %>%
@@
     prob_random = md("*P*(rnd)"),
-    prob_brownian = md("*P*(bm)")
+    prob_brownian = md("*P*(BM)")
   ) %>% 
-  gtsave(here::here("ms/Table_3.html"))
-
-# Include in PDF in latex format
-phylosig_binary_traits_table %>%
-  select(trait, `Number of presences` = num_present, `Number of absences` = num_absent,
-         D, `P(rnd)` = prob_random, `P(bm)` = prob_brownian) %>%
-column_to_rownames("trait") %>%
-  kable(caption = "", format = "latex", booktabs = TRUE) %>%
-  footnote(general = "D, Fritz and Purvis’ \\\\textit{D} (Fritz \\\\& Purvis, 2010); P(rnd), probability of random distribution of traits; P(bm), probability of traits evolving under Brownian motion (BM).
-Values of \\\\textit{D} 0 or less indicate evolution according to BM; values near 1 indicate random trait distribution.", general_title = "", threeparttable = TRUE, escape = FALSE)
-```
+  as_latex
+
+```
+
+\D, Fritz and Purvis' *D* [@Fritz2010]; *P*(rnd), probability of random distribution of traits; *P*(BM), probability of traits evolving under Brownian motion (BM).
+Values of \D 0 or less indicate evolution according to BM; values near 1 indicate random trait distribution.
 
@@
 
-# Format the raw results for table presentation
-pic_table <- 
 pic_results %>%
@@
                    "SLA", "Pinna no.", "Frond dissection"))) %>%
-  arrange(trait)
-
-# Write out table in html
-pic_table %>%
+  arrange(trait) %>%
   gt(rowname_col = "trait") %>%
@@
   ) %>% 
-  gtsave(here::here("ms/Table_4.html"))
-
-# Include in PDF in latex format
-pic_table %>%
-  rename(t = statistic, P = p.value, `Num. constrasts` = n_contrasts, `Num. pos. constrasts` = n_pos_con) %>%
-  column_to_rownames("trait") %>%
-  kable(caption = "", format = "latex", booktabs = TRUE) %>%
-  footnote(general = "All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades.", general_title = "", threeparttable = TRUE)
-```
+  as_latex
+```
+
+All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades.
 
@@
 
-# Format the raw results for table presentation
-correlated_evo_test_table <-
 correlated_evo_test %>%
@@
     new_levels = c("Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
-  arrange(trait)
-
-# Write out table in rich text format
-correlated_evo_test_table %>%
+  arrange(trait) %>%
   gt(rowname_col = "trait") %>%
@@
   ) %>% 
-  gtsave(here::here("ms/Table_5.html"))
-
-# Include in PDF in latex format
-correlated_evo_test_table %>%
-  rename(`LogL(independent)` = logL_indep, `LogL(dependent)` = logL_dep, `Likelihood ratio` = likelihood_ratio, P = pval) %>%
-  column_to_rownames("trait") %>%
-  kable(caption = "", format = "latex", booktabs = TRUE) %>%
-  footnote(general = "LogL, log-likelihood.
+  as_latex
+```
+
+LogL, log-likelihood.
 Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
 terrestrial growth).
-Traits with a significantly lower log likelihood for the dependent model are significantly correlated with growth habit.", general_title = "", threeparttable = TRUE)
-```
+Traits with a significantly lower log likelihood for the dependent model are significantly correlated with growth habit.
 

### 2020-01-14 ms/manuscript.Rmd
Update MS
- Fix spacing when gobbled after xspace
- Use single quotes throughout (New Phyt style)
- Don't use equations
- Only use siunitx functions supported by pandoc

@@
 
+<!-- Work-around to get pandoc to recognize custom siunitx units -->
+\renewcommand{\hour}{h}
+\renewcommand{\second}{s}
+\renewcommand{\ma}{MA}
+\renewcommand{\year}{yr}
+\renewcommand{\month}{mo}
+\renewcommand{\week}{wk}
+\renewcommand{\day}{d}
+\renewcommand{\basepair}{bp}
+
 <!-- Turn off automatic numbering of table captions -->
@@
   ) %>%
-  as_rtf() %>%
-  write_lines(here::here("ms/word_count.rtf"))
+  as_latex
 ```
@@
 The heterogeneous niche space of the angiosperm canopy comprising stratified light and humidity levels and a range of substrates for growth became rapidly filled by a variety of epiphytic plants, including mosses [@Fiz-Palacios2011], leafy liverworts [@Feldberg2014], orchids [@Givnish2015b], bromeliads [@Givnish2014a], and ferns [@Schneider2004d].
-The transition to epiphytic growth was particularly important for modern fern diversification: \circa 29% of all fern species are epiphytes vs. \circa 10% across all vascular plants [@Kress1986].
+The transition to epiphytic growth was particularly important for modern fern diversification: \circa{ }29% of all fern species are epiphytes vs. \circa{ }10% across all vascular plants [@Kress1986].
 Moreover, major transitions to epiphytism occurred at least five times within ferns, each followed by diversification during the late Cretaceous or early Cenozoic [@Schuettpelz2009; @WatkinsJr.2012].
@@
 To gain insight into the role of the canopy in promoting or restricting diversity, a functional and community-based approach can be applied to study this question. 
-If the canopy is uniformly harsher than the terrestrial environment, thus acting as a "filter", we would expect functional convergence of epiphytes at the community level, which in turn may result in decreased phylogenetic diversity of epiphytic communities if traits are phylogenetically conserved [@Webb2002; @Cavender-Bares2009].
+If the canopy is uniformly harsher than the terrestrial environment, thus acting as a 'filter', we would expect functional convergence of epiphytes at the community level, which in turn may result in decreased phylogenetic diversity of epiphytic communities if traits are phylogenetically conserved [@Webb2002; @Cavender-Bares2009].
 Conversely, the stratification of the canopy may stimulate functional diversification by niche partitioning, leading to increased functional diversity of epiphytic communities. 
@@
 
-Moorea, French Polynesia (\ang{17;30;}S, \ang{149;50;}W) is a small (\SI{135}{\kilo\meter\squared}) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
-This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa \SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa \SI{130}{\species}, 8/11 orders; taxonomy follows @PteridophytePhylogenyGroupI2016; `r citest("species-list")`).
+Moorea, French Polynesia (17°30'S, 149°50'W) is a small (\SI{135}{\kilo\meter}^2^) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
+This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa{ }\SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa{ }130 \species, 8/11 orders; taxonomy follows @PteridophytePhylogenyGroupI2016; `r citest("species-list")`).
 Ferns of Moorea include representatives of multiple Cretaceous epiphytic radiations (\eg, *Asplenium*, *Elaphoglossum*, Hymenophylloideae, Polypodiaceae, Vittarioideae; @Schuettpelz2009) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
@@
 
-A fern community survey was carried out on Moorea comprising 17, \SI{10}{\meter} $\times$ \SI{10}{\meter} plots spanning an elevational gradient from \SIrange{20}{2000}{\meter}, with plots placed \circa every \SI{200}{\meter} [@Nitta2017].
+A fern community survey was carried out on Moorea comprising 17, \SI{10}{\meter} × \SI{10}{\meter} plots spanning an elevational gradient from \SI{20}{\meter} to \SI{2000}{\meter}, with plots placed \circa every \SI{200}{\meter} [@Nitta2017].
 Most plots were located along trails leading to the three main peaks on the island, Mt. Rotui (\SI{899}{\meter}), Mt. Mouaputa (\SI{890}{\meter}), and Mt. Tohiea (\SI{1206}{\meter}).
-Each plot was divided into 25, \SI{2}{\meter} $\times$ \SI{2}{\meter} subplots.
+Each plot was divided into 25, \SI{2}{\meter} × \SI{2}{\meter} subplots.
 Presence/absence of all sporophytes in each subplot was scored and summed to produce an abundance ranking for each species per plot from zero (not observed) to 25 (observed in all subplots).
-Epiphytic ferns were sampled by hand to \circa \SI{2}{\meter} on tree trunks or confirmed visually from the ground.
+Epiphytic ferns were sampled by hand to \circa{ }\SI{2}{\meter} on tree trunks or confirmed visually from the ground.
 Species were categorized as epiphytic or terrestrial based on field observations (see Trait Selection and Measurements below).
-The list of species in each plot was split into epiphytic and terrestrial groups and treated separately during data analysis, hereafter referred to as "epiphytic" and "terrestrial" communities.
+The list of species in each plot was split into epiphytic and terrestrial groups and treated separately during data analysis, hereafter referred to as 'epiphytic' and 'terrestrial' communities.
 Vouchers were deposited at GH and UC (abbreviations follow Index Herbariorum; @Thiers2019).
@@
 Temperature and RH of epiphytic and terrestrial habitats were measured using Hobo Pro v2 dataloggers with the RS3 Solar Radiation Shield (Onset Corporation, USA).
-A pair of dataloggers was installed for each plot, one mounted at \circa \SI{2}{\meter} on a tree ("epiphytic datalogger"), and one at ground level ("terrestrial datalogger"), except for the Mt.
+A pair of dataloggers was installed for each plot, one mounted at \circa{ }\SI{2}{\meter} on a tree ('epiphytic datalogger'), and one at ground level ('terrestrial datalogger'), except for the Mt.
 Tohiea \SI{1200}{\meter} plot, which only had an epiphytic datalogger.
@@
 Two plots were not included in the environmental survey.
-We used the 2013–-2014 data for the analysis, with one exception: the terrestrial datalogger at the Mt.
+We used the 2013--2014 data for the analysis, with one exception: the terrestrial datalogger at the Mt.
 Tohiea \SI{393}{\meter} plot malfunctioned and was missing data for a substantial part (March 12--July 5, 2014) of the final survey period, so we used data for the same dates from the preliminary survey period for this datalogger instead.
@@
 We selected several traits relevant to epiphytic growth (`r citet("traits-used")` and `r citef("traits-used-image")`).
-We preferred morphological traits over physiological traits because data for morphological traits are easier to gather at the scale of our study (\circa \SI{130}{\species} distributed over an elevational gradient from \SIrange{200}{2000}{\meter}).
+We preferred morphological traits over physiological traits because data for morphological traits are easier to gather at the scale of our study (\circa{ }130 \species distributed over an elevational gradient from \SI{200}{\meter} to \SI{2000}{\meter}).
 Plant morphology and physiology are closely linked, and physiological traits such as rates of photosynthesis and water use efficiency are known to be correlated with morphological traits such as stipe length in ferns [@WatkinsJr.2010].
@@
 Although trait values can vary within fern species with elevation [@Kessler2007a], our study includes a broad phylogenetic sampling, so we assume intraspecific variation is outweighed by interspecies differences.
-To measure specific leaf area (SLA; \si{\square\meter\per\kilo\gram}), \circa 10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} diam.
-punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
-For species with \textless \SI{2}{\milli\meter} between primary veins, leaf fragments excluding primary veins were obtained by dissection, dried at \SI{60}{\celsius} for \SI{24}{\hour}, weighed, and scanned.
+To measure specific leaf area (SLA; m^2^ kg^-1^), \circa{ }10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} \diameter punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
+For species with < \SI{2}{\milli\meter} between primary veins, leaf fragments excluding primary veins were obtained by dissection, dried at \SI{60}{\celsius} for \SI{24}{\hour}, weighed, and scanned.
 ImageJ [@Abramoff2004] software was used to measure area per fragment.
@@
 Morphological traits of gametophytes were all qualitative (either categorical or binary; `r citet("traits-used")`).
-Traits were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017] or grown from spores of known species in the lab, taken from the literature, or inferred from taxonomic group (i.e., genus) if documented as a stable taxonomic character.
-For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie, heart-shaped) or non-cordate.
-Non-cordate morphotypes include elongate (strap or ribbon \sensu @Farrar2008) and filamentous forms (`r citef("traits-used-image")`).
+Traits were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017] or grown from spores of known species in the lab, taken from the literature, or inferred from taxonomic group (\ie, genus) if documented as a stable taxonomic character.
+For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie, heart-shaped) or non-cordate [@Ebihara2013].
+Non-cordate morphotypes include elongate (strap or ribbon \sensu{ }@Farrar2008) and filamentous forms (`r citef("traits-used-image")`).
 Gametophytes were cultivated on Bold's media [@Bold1957] supplemented with Nitch's micronutrients [@Nitsch1951].
@@
 We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each datalogger, then overall means from the daily values.
-We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (a single variable was arbitrarily selected from each set of correlated variables; hereafter, "climatic variables").
+We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (a single variable was arbitrarily selected from each set of correlated variables; hereafter, 'climatic variables').
 Variables relating to VPD, which have no meaning if less than zero, and some other variables that had skewed distributions, were square-root transformed prior to fitting models.
@@
 
-We tested for phylogenetic signal in quantitative traits using Blomberg's *K* [@Blomberg2003] and Pagel's λ [@Pagel1999], two metrics with slightly different interpretations and sensitivity to different evolutionary scenarios [@Munkemuller2012], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
-λ is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion (BM)).
+We tested for phylogenetic signal in quantitative traits using Blomberg's \K{ }[@Blomberg2003] and Pagel's \plam{ }[@Pagel1999], two metrics with slightly different interpretations and sensitivity to different evolutionary scenarios [@Munkemuller2012], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
+\plam is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion (BM)).
 \K describes the ratio between the amount of observed variance in traits vs.
-the amount of variance expected under BM: when \K > 1, traits are more conserved than expected under BM; when \K < 1, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
-We tested for significance by comparing observed \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in λ by conducting a likelihood ratio test comparing the log likelihoods of the observed value of λ vs.
-λ = 0 (no phylogenetic signal).
-We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' *D* [@Fritz2010].
+the amount of variance expected under BM: when \K{ }> 1, traits are more conserved than expected under BM; when \K{ }< 1, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
+We tested for significance by comparing observed \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in \plam by conducting a likelihood ratio test comparing the log likelihoods of the observed value of \plam vs.
+\plam{ }= 0 (no phylogenetic signal).
+We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' \D{ }[@Fritz2010].
 Values of \D range from 0 under a model of trait evolution by BM, to 1 for random distribution of traits.
 We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], and conducted significance test by comparing observed \D to distributions of simulated values produced under two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM model.
-To account for possible circularity in measuring the phylogenetic signal of traits that were scored based on taxonomy, we conducted the same analyses on a dataset excluding species that had been scored this way ("strict" dataset).
+To account for possible circularity in measuring the phylogenetic signal of traits that were scored based on taxonomy, we conducted the same analyses on a dataset excluding species that had been scored this way ('strict' dataset).
 
@@
 A significantly better fit of the second model indicates that evolution of the two traits is correlated.
-The same analyses were also done on the "strict" dataset.
+The same analyses were also done on the 'strict' dataset.
 
 We investigated differences in multivariate trait space between epiphytes and terrestrial species using principal components analysis (PCA).
-We used only quantitative traits, and included only species that had no missing data (*n* = `r quant_traits_nsp` \si{\species}).
+We used only quantitative traits, and included only species that had no missing data (\n{ }= `r quant_traits_nsp` \species).
 We first log-transformed traits as necessary to reduce skew, then scaled each trait by its standard deviation.
@@
 Raw values of MPD and MNTD are statistically related to species richness [@Swenson2014]; to obtain metrics that are uncorrelated with species richness, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{999} randomized communities.
-We constructed null communities using the "independent swap" algorithm [@Gotelli2000].
-We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available [`r Ntip(phy)` \si{\species} total; @Nitta2017] for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct`) for functional null communities.
+We constructed null communities using the 'independent swap' algorithm [@Gotelli2000].
+We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available [`r Ntip(phy)` \species total; @Nitta2017] for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct`) for functional null communities.
 We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate phylogenetic and functional diversity metrics.
-Hereafter, "MPD~phy~" and "MNTD~phy~", and "MPD~func~" and "MNTD~func~", refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.
+Hereafter, 'MPD~phy~' and 'MNTD~phy~', and 'MPD~func~' and 'MNTD~func~', refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.
 
@@
 For each diversity metric, we built a set of models (Gaussian family, with plot as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
-Our dataset included `r (nrow(div_metrics_all) - nrow(climate_select)) %>% english2` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
+Our dataset included `r (nrow(div_metrics_all) - nrow(climate_select)) %>% jntools::english2` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
 Therefore, we also constructed linear models for each diversity metric related to elevation including all sites using the same model selection procedure as in the climate analysis.
@@
 
-Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs `r table_nums("phylosig-cont", display = "n")`, `r table_nums("phylosig-binary", display = "n")`, `r citef("tree")`), and for quantitative traits, different results were obtained for λ and K.
-When measured with λ, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of λ close to 1; only number of pinnae pairs had λ close to zero (`r citet("phylosig-cont")`).
-However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of \K < 1 (`r citet("phylosig-cont")`).
-All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae, \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under BM (`r citet("phylosig-binary")`).
-Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; `r citet("phylosig-binary")`).
-Similar results were obtained using the "strict" dataset (`r citest("phylosig-binary-strict")`).
+Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs{ }`r table_nums("phylosig-cont", display = "n")`, `r table_nums("phylosig-binary", display = "n")`, `r citef("tree")`), and for quantitative traits, different results were obtained for \plam and \K.
+When measured with \plam, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of \plam close to 1; only number of pinnae pairs had \plam close to zero (`r citet("phylosig-cont")`).
+However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of \K{ }< 1 (`r citet("phylosig-cont")`).
+All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D{ }= `r hairs_d`) or more conserved (gemmae, \D{ }= `r gemmae_d`; glands, \D{ }= `r glands_d`; morphotype, \D{ }= `r morphotype_d`) than expected under BM (`r citet("phylosig-binary")`).
+Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D{ }= `r habit_d`; `r citet("phylosig-binary")`).
+Similar results were obtained using the 'strict' dataset (`r citest("phylosig-binary-strict")`).
 
@@
 Epiphytic species tend to have smaller values for all traits (`r citet("pics")`; contrasts were made by subtracting epiphytic from terrestrial values).
-Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval `r stipe_pval`), as well as frond width (\pval `r width_pval`) and degree of laminae dissection (\pval `r dissection_pval`).
+Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval{ }`r stipe_pval`), as well as frond width (\pval{ }`r width_pval`) and degree of laminae dissection (\pval{ }`r dissection_pval`).
 For the gametophyte traits, only morphotype (non-cordate vs.
-cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval `r morphotype_pval`; `r citet("corr-evo")`).
-Similar results were obtained using the "strict" dataset (`r citest("corr-evo-strict")`).
+cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval{ }`r morphotype_pval`; `r citet("corr-evo")`).
+Similar results were obtained using the 'strict' dataset (`r citest("corr-evo-strict")`).
 
@@
 
-Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval < 0.05; `r citef("cwm-div")`).
-Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval `r get_mod_result(model_summ_print, "sla", "p")`).
-Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval = `r get_mod_result(model_summ_print, "dissection", "p")`).
+Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r citef("cwm-div")`).
+Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval{ }`r get_mod_result(model_summ_print, "sla", "p")`).
+Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "dissection", "p")`).
 Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r citef("cwm-div")`).
 
-We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial species.
-Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \si{\species} per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \si{\species} per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval = `r get_mod_result(t_test_print, "ntaxa", "p")`; `r citef("comm-div")`).
+We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
+Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r citef("comm-div")`).
 Species richness increased with elevation for both epiphytes and terrestrial communities (`r citef("comm-div")`).
-Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, `r both`\pval `r p_mpd_func_mntd_func`).
-Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval `r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval `r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval = `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; `r citef("comm-div")`).
+Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, `r both`\pval{ }`r p_mpd_func_mntd_func`).
+Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval{ }`r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval{ }`r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval{ }= `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; `r citef("comm-div")`).
 No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (`r citef("comm-div")`).
@@
 Another clear and convergent pattern to emerge from our study supporting the hypothesis of convergence is the association between epiphytism and the prevalence of non-cordate gametophyte morphotypes (\eg, elongate or filamentous).
-Gametophyte morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived ($< \SI{1}{\year}$), cordate morphotypes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; @Farrar2008].
+Gametophyte morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived (< \SI{1}{\year}), cordate morphotypes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; @Farrar2008].
 The complex, three-dimensional structure of non-cordate gametophytes may serve two important functions. 
@@
 This may reflect adaptations to specific niches that are not captured in the binary categorization of habitat commonly applied in ferns. 
-For example, the following species ranked among those with the five lowest SLA values (i.e., thick, well-protected leaves) on Moorea: *Humata anderssonii* (Mett.) C.Chr.
+For example, the following species ranked among those with the five lowest SLA values (\ie, thick, well-protected leaves) on Moorea: *Humata anderssonii* (Mett.) C.Chr.
 (a twig epiphyte), *Acrostichum aureum* L.
@@
 As with our results summarized above, it is possible that the diversity of gemmae and non-gemmae producing epiphytic species reflects undetected niche variation in the canopy.
-Interestingly, when we classified gametophytes as "independent" if they occurred beyond the elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with independent growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
+Interestingly, when we classified gametophytes as 'independent' if they occurred beyond the elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with independent growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval{ }= `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
 
@@
 It is somewhat puzzling that VPD did not contribute significantly to any models, as we hypothesized that this would be important for epiphytes that are frequently subjected to water stress. 
-Previous studies have shown marked variation in the air water potential across epiphytic habitats, with @WatkinsJr.2009 reporting a \circa \SI{40}{\mega\pascal} shift along the trunks of emergent canopy trees.
+Previous studies have shown marked variation in the air water potential across epiphytic habitats, with @WatkinsJr.2009 reporting a \circa{ }\SI{40}{\mega\pascal} shift along the trunks of emergent canopy trees.
 Our epiphytic dataloggers only showed slightly greater VPD relative to those on the ground, with this difference largely dissipating at high elevations (`r citef("climate")`).
@@
 Arrows point out instances of each binary trait.
-Scalebars 1 mm except for glands, which is 0.1 mm.
+Scalebars \SI{1}{\milli\meter} except for glands, which is \SI{.1}{\milli\meter}.
 Photographs by J.
@@
 
-**`r citef("climate")`**: Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter on Moorea, French Polynesia.
+**`r citef("climate")`**: Selected microclimatic variables of study plots along an elevational gradient from \SI{200}{\meter} to \SI{1200}\meter on Moorea, French Polynesia.
 Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from `r final_climate_survey$start` to `r final_climate_survey$end`.
 Vapor pressure deficit (VPD) was calculated from relative humidity and temperature as described in the methods, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).
-"Epiphytic" dataloggers were placed at \circa 2 m on trees, and "terrestrial" dataloggers placed at ground level.
+'Epiphytic' dataloggers were placed at \circa{ }\SI{2}{\meter} on trees, and 'terrestrial' dataloggers placed at ground level.
 Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown.
-Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval \textless \num{.05}
+Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval{ }< 0.05.
 
@@
 Epiphytes in green, terrestrial species in brown.
-Quantitative traits only, including only species with no missing observations for any traits (\n = `r pca_results[["species_locs"]] %>% pull(species) %>% n_distinct` \si{\species}).
+Quantitative traits only, including only species with no missing observations for any traits (\n{ }= `r pca_results[["species_locs"]] %>% pull(species) %>% n_distinct` \species).
 Trait abbreviations as follows: dissection, degree of frond dissection; length, frond length; width, frond width; pinna, number of pinnae pairs; rhizome, rhizome diameter; stipe, stipe length; SLA, specific leaf area.
@@
 Error bars are each one standard deviation (values below zero not shown).
-Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at $\pval<0.05$.
-\tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at $\pval<0.05$; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at \pval{ }< 0.05.
+\tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01; \* = \pval{ }< 0.05.
 Results for frond width were very similar to frond length and not shown.
@@
 Response variable abbreviations as follows: MPD~phy~, standard effect size (SES) of mean phylogenetic distance; MNTD~phy~, SES of mean nearest taxon distance; MPD~func~, SES of mean functional distance; MNTD~func~, SES of mean nearest functional distance.
-Trendlines indicate significant relationships as determined by a linear model at $\pval<0.05$.
+Trendlines indicate significant relationships as determined by a linear model at \pval{ }< 0.05.
 Boxplots show median values (bold lines) by growth habit.
-Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 $\times$ the interquartile range.
-Asterisks indicate statistical significance of two-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
+Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range.
+Asterisks indicate statistical significance of two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01; \* = \pval{ }< 0.05.
 
@@
 
-\K, Blomberg's *K* [@Blomberg2003]; lambda, Pagel's lambda [@Pagel1999].
-Values of lambda or \K greater than 1 indicate traits evolving according to Brownian Motion (BM); values of lambda or \K near 0 indicate random distribution of traits.
+\K, Blomberg's \K{ }[@Blomberg2003]; \plam, Pagel's \plam{ }[@Pagel1999].
+Values of \plam or \K greater than 1 indicate traits evolving according to Brownian Motion (BM); values of \plam or \K near 0 indicate random distribution of traits.
 
@@
 
-\D, Fritz and Purvis' *D* [@Fritz2010]; *P*(rnd), probability of random distribution of traits; *P*(BM), probability of traits evolving under Brownian motion (BM).
+\D, Fritz and Purvis' \D{ }[@Fritz2010]; *P*(rnd), probability of random distribution of traits; *P*(BM), probability of traits evolving under Brownian motion (BM).
 Values of \D 0 or less indicate evolution according to BM; values near 1 indicate random trait distribution.

### 2020-01-14 ms/manuscript.Rmd
Update MS
- Format word count table

@@
 library(captioner)
+library(jntools)
 library(conflicted)
@@
 
-<!-- Uncomment to remove page numbers -->
-<!-- \pagenumbering{gobble}  -->
-
 <!-- Set left justification  -->
@@
 
-<!-- Don't indent authors -->
+<!-- Don't indent -->
 \setlength{\parindent}{0ex}
@@
 *Email: cdavis@oeb.harvard.edu*
-
-<!-- Set indent to about 3 characters hereafter -->
-\setlength{\parindent}{3ex}
 
@@
 ) %>%
-  separate(words, c("words", "count"), sep = "_") %>%
-  separate(figs_tabs, c("figs_tabs", "n"), sep = "_", fill = "right") %>%
-  mutate(n = replace_na("")) %>%
-  gt() %>%
-  cols_label(
-    words = "",
-    count = "",
-    figs_tabs = "",
-    n = "",
-  ) %>%
-  as_latex
+  separate(words, c("words", "word_count"), sep = "_") %>%
+  separate(figs_tabs, c("figs", "fig_count"), sep = "_", fill = "right") %>%
+  mutate_all(~replace_na(., "")) %>%
+  knitr::kable(format = "latex", booktabs = TRUE, escape = TRUE, col.names = rep("", ncol(.))) %>%
+  kable_styling(full_width = FALSE) %>%
+  column_spec(1, width = "12em") %>%
+  column_spec(2, width = "2em") %>%
+  column_spec(3, width = "12em") %>%
+  column_spec(4, width = "8em")
 ```
@@
 Tohiea \SI{1200}{\meter} plot, which only had an epiphytic datalogger.
-Temperature and RH were logged once every \SI{15}{\minute} during two survey periods: a preliminary survey from `r prelim_climate_survey$start` to `r prelim_climate_survey$end` at `r prelim_climate_survey$n_sites %>% english2` plots, and a final survey from `r final_climate_survey$start` to `r final_climate_survey$end` at `r final_climate_survey$n_sites %>% english2` plots.
+Temperature and RH were logged once every \SI{15}{\minute} during two survey periods: a preliminary survey from `r prelim_climate_survey$start` to `r prelim_climate_survey$end` at `r prelim_climate_survey$n_sites %>% english2()` plots, and a final survey from `r final_climate_survey$start` to `r final_climate_survey$end` at `r final_climate_survey$n_sites %>% english2()` plots.
 Two plots were not included in the environmental survey.
@@
 Although trait values can vary within fern species with elevation [@Kessler2007a], our study includes a broad phylogenetic sampling, so we assume intraspecific variation is outweighed by interspecies differences.
-To measure specific leaf area (SLA; m^2^ kg^-1^), \circa{ }10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} \diameter punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
+To measure specific leaf area (SLA; m^2^ kg^-1^), \circa{ }10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} diam.
+punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
 For species with < \SI{2}{\milli\meter} between primary veins, leaf fragments excluding primary veins were obtained by dissection, dried at \SI{60}{\celsius} for \SI{24}{\hour}, weighed, and scanned.
@@
 For each diversity metric, we built a set of models (Gaussian family, with plot as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
-Our dataset included `r (nrow(div_metrics_all) - nrow(climate_select)) %>% jntools::english2` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
+Our dataset included `r (nrow(div_metrics_all) - nrow(climate_select)) %>% english2()` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
 Therefore, we also constructed linear models for each diversity metric related to elevation including all sites using the same model selection procedure as in the climate analysis.
@@
 
-**`r citef("traits-used-image")`**: Examples of traits used in this study.
+**`r citef("traits-used-image")`** Examples of traits used in this study.
 **(a)** Sporophyte traits.
@@
 
-**`r citef("climate")`**: Selected microclimatic variables of study plots along an elevational gradient from \SI{200}{\meter} to \SI{1200}\meter on Moorea, French Polynesia.
+**`r citef("climate")`** Selected microclimatic variables of study plots along an elevational gradient from \SI{200}{\meter} to \SI{1200}\meter on Moorea, French Polynesia.
 Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from `r final_climate_survey$start` to `r final_climate_survey$end`.
@@
 
-**`r citef("pca")`**: Principal components analysis (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
+**`r citef("pca")`** Principal components analysis (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
 **(a, b)** PC loadings.
@@
 
-**`r citef("tree")`**: Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
+**`r citef("tree")`** Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
 Tree obtained from @Nitta2017.
@@
 
-**`r citef("heatmap")`**: Variable importance scores from a full-subsets analysis using general additive models exploring the effect of climate and growth habit (terrestrial vs.
+**`r citef("heatmap")`** Variable importance scores from a full-subsets analysis using general additive models exploring the effect of climate and growth habit (terrestrial vs.
 epiphytic growth) on community diversity of ferns on Moorea, French Polynesia.
@@
 
-**`r citef("cwm-div")`**: Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
+**`r citef("cwm-div")`** Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
 Epiphytic communities in green, terrestrial communities in brown.
@@
 
-**`r citef("comm-div")`**: Functional and phylogenetic diversity of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
+**`r citef("comm-div")`** Functional and phylogenetic diversity of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
 Epiphytic communities in green, terrestrial communities in brown.

### 2020-01-14 ms/manuscript.Rmd
Change names for custom citation functions

@@
 # Make short versions of citation functions
-citef <- pryr::partial(fig_nums, display = "cite")
-citet <- pryr::partial(table_nums, display = "cite")
-citesf <- pryr::partial(s_fig_nums, display = "cite")
-citest <- pryr::partial(s_table_nums, display = "cite")
+figure <- pryr::partial(fig_nums, display = "cite")
+table <- pryr::partial(table_nums, display = "cite")
+s_figure <- pryr::partial(s_fig_nums, display = "cite")
+s_table <- pryr::partial(s_table_nums, display = "cite")
 
@@
 Moorea, French Polynesia (17°30'S, 149°50'W) is a small (\SI{135}{\kilo\meter}^2^) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
-This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa{ }\SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa{ }130 \species, 8/11 orders; taxonomy follows @PteridophytePhylogenyGroupI2016; `r citest("species-list")`).
+This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa{ }\SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa{ }130 \species, 8/11 orders; taxonomy follows @PteridophytePhylogenyGroupI2016; `r s_table("species-list")`).
 Ferns of Moorea include representatives of multiple Cretaceous epiphytic radiations (\eg, *Asplenium*, *Elaphoglossum*, Hymenophylloideae, Polypodiaceae, Vittarioideae; @Schuettpelz2009) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
@@
 
-We selected several traits relevant to epiphytic growth (`r citet("traits-used")` and `r citef("traits-used-image")`).
+We selected several traits relevant to epiphytic growth (`r table("traits-used")` and `r figure("traits-used-image")`).
 We preferred morphological traits over physiological traits because data for morphological traits are easier to gather at the scale of our study (\circa{ }130 \species distributed over an elevational gradient from \SI{200}{\meter} to \SI{2000}{\meter}).
@@
 
-Morphological traits of gametophytes were all qualitative (either categorical or binary; `r citet("traits-used")`).
+Morphological traits of gametophytes were all qualitative (either categorical or binary; `r table("traits-used")`).
 Traits were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017] or grown from spores of known species in the lab, taken from the literature, or inferred from taxonomic group (\ie, genus) if documented as a stable taxonomic character.
 For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie, heart-shaped) or non-cordate [@Ebihara2013].
-Non-cordate morphotypes include elongate (strap or ribbon \sensu{ }@Farrar2008) and filamentous forms (`r citef("traits-used-image")`).
+Non-cordate morphotypes include elongate (strap or ribbon \sensu{ }@Farrar2008) and filamentous forms (`r figure("traits-used-image")`).
 Gametophytes were cultivated on Bold's media [@Bold1957] supplemented with Nitch's micronutrients [@Nitsch1951].
@@
 
-Trait values and data sources are summarized in `r citest("trait-values")`.
+Trait values and data sources are summarized in `r s_table("trait-values")`.
 
@@
 Mean, maximum, and standard deviation (SD) of temperature and mean, minimum, and SD of vapor pressure deficit (VPD) were retained after excluding variables with correlation coefficients > 0.9.
-Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (`r citest("ancova")`; `r citef("climate")`).
+Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (`r s_table("ancova")`; `r figure("climate")`).
 Variables related to VPD differed by growth habit and elevation.
@@
 
-Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (`r citef("pca")`**c**, **d**).
+Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (`r figure("pca")`**c**, **d**).
 The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
@@
 
-Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs{ }`r table_nums("phylosig-cont", display = "n")`, `r table_nums("phylosig-binary", display = "n")`, `r citef("tree")`), and for quantitative traits, different results were obtained for \plam and \K.
-When measured with \plam, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of \plam close to 1; only number of pinnae pairs had \plam close to zero (`r citet("phylosig-cont")`).
-However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of \K{ }< 1 (`r citet("phylosig-cont")`).
-All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D{ }= `r hairs_d`) or more conserved (gemmae, \D{ }= `r gemmae_d`; glands, \D{ }= `r glands_d`; morphotype, \D{ }= `r morphotype_d`) than expected under BM (`r citet("phylosig-binary")`).
-Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D{ }= `r habit_d`; `r citet("phylosig-binary")`).
-Similar results were obtained using the 'strict' dataset (`r citest("phylosig-binary-strict")`).
+Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs{ }`r table_nums("phylosig-cont", display = "n")`, `r table_nums("phylosig-binary", display = "n")`, `r figure("tree")`), and for quantitative traits, different results were obtained for \plam and \K.
+When measured with \plam, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of \plam close to 1; only number of pinnae pairs had \plam close to zero (`r table("phylosig-cont")`).
+However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of \K{ }< 1 (`r table("phylosig-cont")`).
+All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D{ }= `r hairs_d`) or more conserved (gemmae, \D{ }= `r gemmae_d`; glands, \D{ }= `r glands_d`; morphotype, \D{ }= `r morphotype_d`) than expected under BM (`r table("phylosig-binary")`).
+Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D{ }= `r habit_d`; `r table("phylosig-binary")`).
+Similar results were obtained using the 'strict' dataset (`r s_table("phylosig-binary-strict")`).
 
@@
 
-Most sporophyte traits, including all those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (`r citet("pics")`).
-Epiphytic species tend to have smaller values for all traits (`r citet("pics")`; contrasts were made by subtracting epiphytic from terrestrial values).
+Most sporophyte traits, including all those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (`r table("pics")`).
+Epiphytic species tend to have smaller values for all traits (`r table("pics")`; contrasts were made by subtracting epiphytic from terrestrial values).
 Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval{ }`r stipe_pval`), as well as frond width (\pval{ }`r width_pval`) and degree of laminae dissection (\pval{ }`r dissection_pval`).
 For the gametophyte traits, only morphotype (non-cordate vs.
-cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval{ }`r morphotype_pval`; `r citet("corr-evo")`).
-Similar results were obtained using the 'strict' dataset (`r citest("corr-evo-strict")`).
+cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval{ }`r morphotype_pval`; `r table("corr-evo")`).
+Similar results were obtained using the 'strict' dataset (`r s_table("corr-evo-strict")`).
 
@@
 
-In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (`r citef("heatmap")`).
-Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (`r citest("model-fit")`).
-Spatial autocorrelation was not detected for any of the diversity metrics (`r citest("model-fit")`).
-
-Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r citef("cwm-div")`).
+In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (`r figure("heatmap")`).
+Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (`r s_table("model-fit")`).
+Spatial autocorrelation was not detected for any of the diversity metrics (`r s_table("model-fit")`).
+
+Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
 Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval{ }`r get_mod_result(model_summ_print, "sla", "p")`).
 Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "dissection", "p")`).
-Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r citef("cwm-div")`).
+Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r figure("cwm-div")`).
 
 We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
-Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r citef("comm-div")`).
-Species richness increased with elevation for both epiphytes and terrestrial communities (`r citef("comm-div")`).
+Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r figure("comm-div")`).
+Species richness increased with elevation for both epiphytes and terrestrial communities (`r figure("comm-div")`).
 Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, `r both`\pval{ }`r p_mpd_func_mntd_func`).
 Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval{ }`r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval{ }`r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval{ }= `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; `r citef("comm-div")`).
-No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (`r citef("comm-div")`).
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval{ }`r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval{ }= `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; `r figure("comm-div")`).
+No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (`r figure("comm-div")`).
 
@@
 Our study reveals a striking duality in the phylogenetic and functional diversity of fern epiphytes. 
-On one hand, we find that fern epiphytes are consistently smaller than terrestrial species at both the species and community level (`r citef("cwm-div")`, `r citet("pics")`), and that epiphytic communities are phylogenetically clustered (`r citef("comm-div")`).
-At the same time, we find that epiphytic fern communities are significantly more diverse functionally compared to terrestrial communities (`r citef("comm-div")`).
+On one hand, we find that fern epiphytes are consistently smaller than terrestrial species at both the species and community level (`r figure("cwm-div")`, `r table("pics")`), and that epiphytic communities are phylogenetically clustered (`r figure("comm-div")`).
+At the same time, we find that epiphytic fern communities are significantly more diverse functionally compared to terrestrial communities (`r figure("comm-div")`).
 Collectively, these results suggest that environmental filtering plays a strong role in limiting plant size, while simultaneously promoting diversification across other trait axes in epiphytic ferns.
@@
 
-Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (`r citef("cwm-div")`, `r citet("pics")`), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
+Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (`r figure("cwm-div")`, `r table("pics")`), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
 These results are corroborated by @Testo2018a, who similarly observed abrupt size reduction associated with epiphytic growth in a globally sampled fern phylogeny. 
@@
 These results do not support the hypothesis that fern epiphytes have reduced functional diversity.
-For example, although epiphytes demonstrated less dissected fronds at the species level (`r citet("pics")`), they exhibited much greater variation in this trait relative to terrestrial ferns at the community level (`r citesf("cwsd")`).
+For example, although epiphytes demonstrated less dissected fronds at the species level (`r table("pics")`), they exhibited much greater variation in this trait relative to terrestrial ferns at the community level (`r s_figure("cwsd")`).
 It is likely that the high diversity of this trait reflects different physiological adaptations to various canopy niches. 
@@
 
-Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA; `r citef("cwm-div")`, `r citet("pics")`).
+Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA; `r figure("cwm-div")`, `r table("pics")`).
 Specific leaf area is part of the leaf economic spectrum (LES), which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
@@
 We expected that epiphytes would exhibit lower SLA, given their more extreme growing conditions.
-In contrast, we observed overlapping, wide ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes (`r citef("cwm-div")`).
+In contrast, we observed overlapping, wide ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes (`r figure("cwm-div")`).
 This may reflect adaptations to specific niches that are not captured in the binary categorization of habitat commonly applied in ferns. 
@@
 (a terrestrial mangrove specialist), and *Paesia divaticatissima* (Dryand.) Copel.
-(a terrestrial species occurring on high-elevation ridges) (`r citest("trait-values")`).
+(a terrestrial species occurring on high-elevation ridges) (`r s_table("trait-values")`).
 Each of these species is likely to experience prolonged periods of drought-stress. 
@@
 We did not find that gemmae production in gametophytes was linked with epiphytism. 
-Rather, occurrence of gemmae strongly reflects phylogeny, appearing concentrated in filmy ferns (Hymenophyllaceae) and vittarioid ferns (`r citef("tree")`).
+Rather, occurrence of gemmae strongly reflects phylogeny, appearing concentrated in filmy ferns (Hymenophyllaceae) and vittarioid ferns (`r figure("tree")`).
 Gemmae are asexual propagules occurring primarily in non-cordate gametophytes that allow gametophytes to persist over multiple growing seasons and attain large population sizes, and have been hypothesized to be associated with harsh environmental conditions [@Farrar2008].
@@
 
-Finally, the presence of hairs and glands in gametophytes also exhibit very high phylogenetic signal (\ie, much greater than expected under Brownian motion; `r citet("phylosig-binary")`), and these traits are concentrated in certain taxonomic groups [\eg, hairs, grammitid ferns; glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
+Finally, the presence of hairs and glands in gametophytes also exhibit very high phylogenetic signal (\ie, much greater than expected under Brownian motion; `r table("phylosig-binary")`), and these traits are concentrated in certain taxonomic groups [\eg, hairs, grammitid ferns; glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
 Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], and glands may affect rates of osmosis by modifying lipid concentrations [@Crow2011], we did not detect a correlation between hair or gland production and epiphytic growth.
@@
 On the one hand, the canopy appears to have served as a potent barrier to entry, strongly selecting for small size. 
-Furthermore, the results of our phylogenetic community structure analyses partially support the hypothesis of decreased phylogenetic diversity in epiphytic communities due to filtering: we observed that epiphytic fern communities were more phylogenetically clustered for MNTD~phy~, but not MPD~phy~ (`r citef("comm-div")`).
+Furthermore, the results of our phylogenetic community structure analyses partially support the hypothesis of decreased phylogenetic diversity in epiphytic communities due to filtering: we observed that epiphytic fern communities were more phylogenetically clustered for MNTD~phy~, but not MPD~phy~ (`r figure("comm-div")`).
 This makes sense given that most epiphytes have radiated within a limited number of clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely related species from two or more epiphytic clades, but exhibits a similar overall degree of phylogenetic diversity as terrestrial communities.
@@
 However, while the canopy clearly acts as a filter, we also find that it simultaneously functions as a promoter of diversity. 
-This is evident in our analyses of MPD~func~ and MNTD~func~, which both indicate that epiphytic communities are more functionally diverse compared to terrestrial species (`r citef("comm-div")`), thus contradicting our hypothesis of decreased functional diversity in epiphytic communities.
-Additional analyses using different trait combinations support the role of non-size related traits driving this pattern (`r citesf("corr-plot")`).
+This is evident in our analyses of MPD~func~ and MNTD~func~, which both indicate that epiphytic communities are more functionally diverse compared to terrestrial species (`r figure("comm-div")`), thus contradicting our hypothesis of decreased functional diversity in epiphytic communities.
+Additional analyses using different trait combinations support the role of non-size related traits driving this pattern (`r s_figure("corr-plot")`).
 We propose that the high functional diversity of epiphytic fern communities is maintained by previously undetected niche partitioning within epiphytic habitats as follows.
@@
 
-The primary role of growth habit, in combination with temperature, in structuring fern communities is supported by the full-subsets model analysis (`r citef("heatmap")`).
+The primary role of growth habit, in combination with temperature, in structuring fern communities is supported by the full-subsets model analysis (`r figure("heatmap")`).
 It is somewhat puzzling that VPD did not contribute significantly to any models, as we hypothesized that this would be important for epiphytes that are frequently subjected to water stress. 
 Previous studies have shown marked variation in the air water potential across epiphytic habitats, with @WatkinsJr.2009 reporting a \circa{ }\SI{40}{\mega\pascal} shift along the trunks of emergent canopy trees.
-Our epiphytic dataloggers only showed slightly greater VPD relative to those on the ground, with this difference largely dissipating at high elevations (`r citef("climate")`).
+Our epiphytic dataloggers only showed slightly greater VPD relative to those on the ground, with this difference largely dissipating at high elevations (`r figure("climate")`).
 One reason for such low observed variation could be due to environmental conditions at our study site and placement of dataloggers: Moorea is a generally wet site overall (average annual rainfall \SI{325}{\centi\meter}), and we placed the epiphytic dataloggers on tree trunks, not in the higher emergent portions of the canopy.
@@
 
-**`r citest("species-list")`** List of ferns from Moorea, French Polynesia included in this study.
-
-**`r citest("trait-values")`** Trait values of ferns on Moorea, French Polynesia.
-
-**`r citest("ancova")`** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate.
-
-**`r citest("phylosig-binary-strict")`** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
-
-**`r citest("corr-evo-strict")`** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
-
-**`r citest("model-fit")`** Best-fitting models explaining community diversity metrics by climate and growth habit.
-
-**`r citesf("corr-plot")`** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
-
-**`r citesf("cwsd")`** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
-
-\clearpage
-
-**`r citef("traits-used-image")`** Examples of traits used in this study.
+**`r s_table("species-list")`** List of ferns from Moorea, French Polynesia included in this study.
+
+**`r s_table("trait-values")`** Trait values of ferns on Moorea, French Polynesia.
+
+**`r s_table("ancova")`** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate.
+
+**`r s_table("phylosig-binary-strict")`** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+
+**`r s_table("corr-evo-strict")`** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+
+**`r s_table("model-fit")`** Best-fitting models explaining community diversity metrics by climate and growth habit.
+
+**`r s_figure("corr-plot")`** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
+
+**`r s_figure("cwsd")`** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
+
+\clearpage
+
+**`r figure("traits-used-image")`** Examples of traits used in this study.
 **(a)** Sporophyte traits.
@@
 
-**`r citef("climate")`** Selected microclimatic variables of study plots along an elevational gradient from \SI{200}{\meter} to \SI{1200}\meter on Moorea, French Polynesia.
+**`r figure("climate")`** Selected microclimatic variables of study plots along an elevational gradient from \SI{200}{\meter} to \SI{1200}\meter on Moorea, French Polynesia.
 Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from `r final_climate_survey$start` to `r final_climate_survey$end`.
@@
 
-**`r citef("pca")`** Principal components analysis (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
+**`r figure("pca")`** Principal components analysis (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
 **(a, b)** PC loadings.
@@
 
-**`r citef("tree")`** Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
+**`r figure("tree")`** Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
 Tree obtained from @Nitta2017.
@@
 Species missing data for six or more traits not shown.
-For a summary of traits, see `r citet("traits-used")`.
+For a summary of traits, see `r table("traits-used")`.
 Major epiphytic radiations identified by @Schuettpelz2009 labeled on tree in green: H, Hymenophylloideae; T, Trichomanoideae; V, Vittarioideae; A, *Asplenium*; E, *Elaphoglossum*; P, Polypodiaceae; taxonomy follows @PteridophytePhylogenyGroupI2016.
 
-**`r citef("heatmap")`** Variable importance scores from a full-subsets analysis using general additive models exploring the effect of climate and growth habit (terrestrial vs.
+**`r figure("heatmap")`** Variable importance scores from a full-subsets analysis using general additive models exploring the effect of climate and growth habit (terrestrial vs.
 epiphytic growth) on community diversity of ferns on Moorea, French Polynesia.
@@
 
-**`r citef("cwm-div")`** Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
+**`r figure("cwm-div")`** Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
 Epiphytic communities in green, terrestrial communities in brown.
@@
 
-**`r citef("comm-div")`** Functional and phylogenetic diversity of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
+**`r figure("comm-div")`** Functional and phylogenetic diversity of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
 Epiphytic communities in green, terrestrial communities in brown.

### 2020-01-14 ms/manuscript.Rmd
Update MS
- Fix pvalue formatting in Table 3
- Use unicode lambda in Table 2

@@
 ## References
+
+<!-- FIXME: for some reason, a single space is inserted before each reference when
+coverting to docx from tex by pandoc. Remove these from docx file. -->
+
+<!-- FIXME: manually remove "Jr" from inline-citations in docx file. -->
+
 <!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
@@
     k_pval = md("*P*(*K*)"),
-    lambda_pval = md("*P*(lambda)"),
-    lambda = "lambda"
+    lambda_pval = md("*P*(λ)"),
+    lambda = "λ"
   )  %>%
@@
   mutate(D = round_t(D, 2)) %>%
+  mutate_at(vars(prob_random, prob_brownian), format_pval) %>%
   mutate(trait = factor(

### 2020-01-14 ms/SI.rmd
Change location of ms output to results/
Add docx conversion for SI

@@
     includes:
-      in_header: newphyt.sty
+      in_header: new-phytologist.sty
 nocite: | 
@@
 
-Article title: Constrained, yet diverse: Phylogenetic and morphological filtering in epiphytic ferns
+Article title: Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes
 
@@
 
-**Table S1** List of ferns from Moorea, French Polynesia included in this study.
+**Table S1** List of ferns from Moorea, French Polynesia included in this study
 
@@
 
-**Table S2** Trait values of ferns on Moorea, French Polynesia.
+**Table S2** Trait values of ferns on Moorea, French Polynesia
 
 SLA, specific leaf area.
-
 Continuous measurements shown as mean, standard deviation (sample size).
@@
 trait_data_for_si %>%
-  write_csv(here::here("ms/Table_S2.csv"))
+  write_csv(here::here("results/Table_S2.csv"))
 ```
 
-**Table S3** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate.
+**Table S3** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate
 
@@
   select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value) %>%
-  write_csv(here::here("ms/Table_S3.csv"))
+  write_csv(here::here("results/Table_S3.csv"))
 
@@
 
-**Table S4** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+**Table S4** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
 
-P_rnd, probability of obtaining the observed value of *D* under random trait evolution; P_bm, probability of obtaining the observed value of *D* under simulation of trait evolution by Brownian motion.
+P_rnd, probability of obtaining the observed value of \D under random trait evolution; P_bm, probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.
 
@@
          P_rnd = prob_random, P_bm = prob_brownian) %>%
-  write_csv(here::here("ms/Table_S4.csv"))
+  write_csv(here::here("results/Table_S4.csv"))
 ```
 
-**Table S5** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
+**Table S5** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
 
 LL, log likelihood.
-
 Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
@@
   ) %>%
-  write_csv(here::here("ms/Table_S5.csv"))
+  write_csv(here::here("results/Table_S5.csv"))
 ```
 
-**Table S6** Best-fitting models explaining community diversity metrics by climate and growth habit.
+**Table S6** Best-fitting models explaining community diversity metrics by climate and growth habit
 
 MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, *P* value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's *I*.
-
 Models within delta AICc 3 of the best-fitting model shown.
@@
   ) %>%
-  write_csv(here::here("ms/Table_S6.csv"))
+  write_csv(here::here("results/Table_S6.csv"))
 ```
 
-**Fig. S1** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
-
+**Fig.
+S1** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
 **(a)**, standard effect size of mean phylogenetic distance (MPD~func~) **(b)**, standard effect size of mean nearest taxon distance (MNTD~func~).
 For details on calculations of MPD~func~ and MNTD~func~, see Methods.
-Correlations (values in each tile) calculated with Pearson's correlation coefficient using the R package "corrr".
+Correlations (values in each tile) calculated with Pearson's correlation coefficient using the R package 'corrr'.
 All trait sets were scaled to standard deviation of each trait.
-The following trait sets were used: "all" (all traits), "size" (sporophyte traits related to size; frond length and width, stipe length, rhizome width, pinna number, SLA, degree of lamina dissection), "other" (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), "gameto" (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), "reduced" (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter).
+The following trait sets were used: 'all' (all traits), 'size' (sporophyte traits related to size; frond length and width, stipe length, rhizome width, pinna number, SLA, degree of lamina dissection), 'other' (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), 'gameto' (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), 'reduced' (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter).
 For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), both log-transformed and untransformed sets were included.
-The log-transformed "reduced" dataset was used for the main results.
+The log-transformed 'reduced' dataset was used for the main results.
 
@@
 
-ggsave(plot = fig_s1, filename = here::here("ms/Fig_S1.eps"), height = 8, width = 4, units = "in")
+ggsave(plot = fig_s1, filename = here::here("results/Fig_S1.eps"), height = 8, width = 4, units = "in")
 ```
@@
 Boxplots show median values (bold lines) by growth habit.
-Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 $\times$ the interquartile range.
-Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.
-
-**(a)**, traits related to plant size **(b)**, traits not related to plant size.
+Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range.
+Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01; \* = \pval{ }< 0.05 **(a)**, traits related to plant size **(b)**, traits not related to plant size.
 
@@
 loadd(cwsd_scatterplot, cache = canopy_cache)
-ggsave(plot = cwsd_scatterplot, filename = here::here("ms/Fig_S2.eps"), height = 6, width = 7, units = "in")
+ggsave(plot = cwsd_scatterplot, filename = here::here("results/Fig_S2.eps"), height = 6, width = 7, units = "in")
 ```

### 2020-01-14 ms/manuscript.Rmd
Change location of ms output to results/
Add docx conversion for SI

@@
 ---
-title: "Life in the canopy: community trait assessments reveal substantial functional diversity among fern epiphytes"
+title: "Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes"
 bibliography: "references.bib"
@@
 
-ggsave(plot = climate_plot, filename = here::here("ms/Fig_2.eps"), height = 4, width = 7, units = "in", device = cairo_ps_high_res)
+ggsave(plot = climate_plot, filename = here::here("results/Fig_2.eps"), height = 4, width = 7, units = "in", device = cairo_ps_high_res)
 ```
@@
 pca_plot
-ggsave(plot = pca_plot, filename = here::here("ms/Fig_3.eps"), height = 6, width = 6, units = "in", device = cairo_ps_high_res)
+ggsave(plot = pca_plot, filename = here::here("results/Fig_3.eps"), height = 6, width = 6, units = "in", device = cairo_ps_high_res)
 ```
@@
 traits_with_tree
-ggsave(plot = traits_with_tree, filename = here::here("ms/Fig_4.eps"), height = 8, width = 6, units = "in", device = cairo_ps_high_res)
+ggsave(plot = traits_with_tree, filename = here::here("results/Fig_4.eps"), height = 8, width = 6, units = "in", device = cairo_ps_high_res)
 ```
@@
 importance_heatmap
-ggsave(plot = importance_heatmap, filename = here::here("ms/Fig_5.eps"), height = 4, width = 6, units = "in", device = cairo_ps_high_res)
+ggsave(plot = importance_heatmap, filename = here::here("results/Fig_5.eps"), height = 4, width = 6, units = "in", device = cairo_ps_high_res)
 ```
@@
 combined_cwm_plots
-ggsave(plot = combined_cwm_plots, filename = here::here("ms/Fig_6.eps"), height = 4, width = 7, units = "in", device = cairo_ps_high_res)
+ggsave(plot = combined_cwm_plots, filename = here::here("results/Fig_6.eps"), height = 4, width = 7, units = "in", device = cairo_ps_high_res)
 ```
@@
 combined_comm_div_plots
-ggsave(plot = combined_comm_div_plots, filename = here::here("ms/Fig_7.eps"), height = 8, width = 4, units = "in", device = cairo_ps_high_res)
+ggsave(plot = combined_comm_div_plots, filename = here::here("results/Fig_7.eps"), height = 8, width = 4, units = "in", device = cairo_ps_high_res)
 ```

### 2020-01-14 code/plan.R

@@ -105,7 +105,7 @@ plan <- drake_plan(
   # - Write out moorea_climate as CSV file to include with Dryad data.
   moorea_climate_out = moorea_climate %>%
     select(date, time, site, habit, temp, rh, vpd) %>%
-    write_csv(file_out("ms/moorea_climate.csv")),
+    write_csv(file_out("results/moorea_climate.csv")),
   
   # - Calculate mean climate variables.
   daily_mean_climate = get_daily_means(moorea_climate),
@@ -306,7 +306,7 @@ plan <- drake_plan(
       traits_select_list = c(all, size, other, gameto, reduced),
     )
   ),
-
+  
   func_comm_struc_combined = target(
     bind_data(func_comm_struc_test),
     transform = combine(func_comm_struc_test)
@@ -331,11 +331,11 @@ plan <- drake_plan(
   cwm_long = calculate_cwm(fern_traits, comm, moorea_sites),
   # - and by site
   cwm_by_site = cwm_long %>% select(-sd, -el) %>% spread(trait, cwm),
-
+  
   # Modeling ----
-
+  
   ### Prepare diversity data (response variables) ###
-
+  
   # - Combine diversity metrics into single dataframe
   div_metrics_all = left_join(
     select(phy_comm_struc, 
@@ -348,18 +348,18 @@ plan <- drake_plan(
            mntd.obs.z.func = mntd.obs.z)) %>%
     left_join(cwm_by_site) %>%
     left_join(moorea_sites),
-
+  
   ### Run full-subset analysis using GAMs ###
-
+  
   # - Merge all diversity metrics and climate data for GAMs
   # (transformed version of climate, but including outlier)
   div_climate_trans = inner_join(div_metrics_all, climate_trans),
-
+  
   # - Make named vector of response variables to test
   resp_vars = div_metrics_all %>%
     select(-site, -habit, -el, -long, -lat) %>%
     colnames %>% rlang::set_names(),
-
+  
   # - Run full-subsets model analysis.
   fss_div_results = purrr::map(
     resp_vars,
@@ -368,13 +368,13 @@ plan <- drake_plan(
       indep_vars = climate_vars,
       resp_var = .)
   ),
-
+  
   # - Extract table of importance of each environmental variable.
   important_div_vars = get_important_vars(fss_div_results),
-
+  
   # - Extract table of best-fit models.
   best_fit_div_models = get_best_fss_mods(fss_div_results),
-
+  
[…]

### 2020-01-14 code/resolve_names.R

@@ -84,4 +84,4 @@ taxa_names_with_authors <- gnr_resolve(
   )) %>%
   rename_all(~str_replace(., "_", " "))
   
-write_csv(taxa_names_with_authors, here::here("ms/Table_S1.csv"))
+write_csv(taxa_names_with_authors, here::here("results/Table_S1.csv"))

### 2020-01-14 ms/SI.rmd
Use default latex font for pdf

@@
 bibliography: "si_references.bib"
-mainfont: Times New Roman.ttf
 fontsize: 12pt
-mainfontoptions: 
-- BoldFont=Times New Roman Bold.ttf
-- ItalicFont=Times New Roman Italic.ttf
-- BoldItalicFont=Times New Roman Bold Italic.ttf
 output: 

### 2020-01-14 ms/manuscript.Rmd
Use default latex font for pdf

@@
 bibliography: "references.bib"
-mainfont: Times New Roman.ttf
 fontsize: 12pt
-mainfontoptions: 
-- BoldFont=Times New Roman Bold.ttf
-- ItalicFont=Times New Roman Italic.ttf
-- BoldItalicFont=Times New Roman Bold Italic.ttf
 output: 

### 2020-01-14 ms/manuscript.Rmd
Put all library() calls in setup.R
and arrange alphabetically

@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
-library(drake)
-library(gt)
-library(patchwork)
-library(knitr)
-library(kableExtra)
-library(magrittr)
-library(scales)
-library(tidyverse)
-library(captioner)
-library(jntools)
-library(conflicted)
-
-conflict_prefer("expand", "tidyr")
-conflict_prefer("select", "dplyr")
-conflict_prefer("filter", "dplyr")
 

### 2020-01-14 code/setup.R

@@ -1,36 +1,43 @@
-library(conflicted)
-library(FD)
-library(FSSgam)
-library(FactoMineR)
-library(RColorBrewer)
+# Load packages
 library(assertr)
 library(assertthat)
 library(broom)
 library(caper)
+library(captioner)
+library(conflicted)
 library(corrr)
 library(cowplot)
 library(drake)
+library(FactoMineR)
+library(FD)
+library(FSSgam)
 library(ggrepel)
 library(ggtree)
 library(glue)
+library(gt)
 library(janitor)
 library(jntools)
+library(kableExtra)
+library(knitr)
 library(magrittr)
 library(mgcv)
 library(patchwork)
 library(phytools)
 library(picante)
 library(plantecophys)
+library(RColorBrewer)
 library(scales)
 library(scico)
 library(sme)
 library(sp)
 library(spdep)
-library(viridis)
 library(tidyverse)
+library(viridis)
 
+# Resolve conflicts
+conflict_prefer("expand", "tidyr")
 conflict_prefer("filter", "dplyr")
 conflict_prefer("gather", "tidyr")
+conflict_prefer("ggsave", "ggplot2")
 conflict_prefer("map", "purrr")
 conflict_prefer("select", "dplyr")
-conflict_prefer("ggsave", "ggplot2")

### 2020-01-15 ms/manuscript.Rmd
Use separate bib and yaml for normal and website refs

@@
 title: "Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes"
-bibliography: "references.bib"
+bibliography: 
+  - "references.bib"
+  - "references_other.yaml"
 fontsize: 12pt

### 2020-01-15 code/process_refs.R

@@ -6,31 +6,35 @@ library(tidyverse)
 make_ref_list(
   rmd_file = "ms/manuscript.Rmd", 
   raw_bib = "ms/references_raw.bib",
-  final_bib = "ms/references.bib")
+  final_bib = "ms/references.bib",
+  strip_fields = "abstract",
+  exclude = c("Laliberte2011", "Orme2013", "Thiers2019"))
 
 make_ref_list(
   rmd_file = "ms/SI.Rmd", 
   raw_bib = "ms/references_raw.bib", 
-  final_bib = "ms/si_references.bib"
+  final_bib = "ms/references_si.bib",
+  exclude = "FloraNSW2019"
 )
 
 # Make some manual fixes to authors in SI bibliography
 # (these are institutions, so need double brackets to
 # avoid latex thinking they have first and last names)
-read_lines("ms/si_references.bib") %>%
+read_lines("ms/references_si.bib") %>%
   str_replace(
     "National Museum of Nature and Science",
     "\\{National Museum of Nature and Science\\}") %>%
-  str_replace(
-    "Royal Botanic Gardens and Domain Trust",
-    "\\{Royal Botanic Gardens and Domain Trust\\}") %>%
-  write_lines("ms/si_references.bib")
+  write_lines("ms/references_si.bib")
 
-# Make some manual fixes to authors in SI bibliography
-# (these are institutions, so need double brackets to
-# avoid latex thinking they have first and last names)
+# Make some manual fixes to main bibliography
 read_lines("ms/references.bib") %>%
   str_replace(
     "Pteridophyte Phylogeny Group I",
     "\\{Pteridophyte Phylogeny Group I\\}") %>%
+  str_replace(
+    "Adaptive radiation versus 'radiation' and 'explosive diversification'",
+    "Adaptive radiation versus `radiation' and `explosive diversification'") %>%
+  str_replace(
+    "'Sun leaves' and 'shade leaves'",
+    "`Sun leaves' and `shade leaves'") %>%
   write_lines("ms/references.bib")

### 2020-01-15 ms/SI.rmd
Update MS
various fixes to output docx correctly

@@
 title: ""
-bibliography: "si_references.bib"
+bibliography: 
+  - "references_si.bib"
+  - "references_other.yaml"
 fontsize: 12pt
@@
 nocite: | 
-  @NMNS2008, @Copeland1932, @FloraNSW2019, @Brownsey1987, @Murdock2003, @Nayar1971, @Lloyd1980, @Zhang2008, @Bierhorst1967, @Martin2006, @Tigerschiold1989, @Tigerschiold1990, @Atkinson1975, @Palmer2003
+  @NMNS2008, @Copeland1932, @FloraNSW2019, @Brownsey1987, @Nayar1971, @Lloyd1980, @Zhang2008, @Bierhorst1967, @Martin2006, @Tigerschiold1989, @Tigerschiold1990, @Atkinson1975, @Palmer2003
 ---
@@
 # Load analysis results
-loadd(trait_data_for_si, cache = canopy_cache)
-loadd(climate_ancova_results, cache = canopy_cache)
-loadd(phylosig_binary_traits_strict, cache = canopy_cache)
-loadd(correlated_evo_test_strict, cache = canopy_cache)
-loadd(best_fit_div_models_moran, cache = canopy_cache)
-loadd(func_comm_struc_combined, cache = canopy_cache)
+loadd(list = c(
+  "trait_data_for_si",
+  "climate_ancova_results",
+  "phylosig_binary_traits_strict",
+  "correlated_evo_test_strict",
+  "best_fit_div_models_moran",
+  "func_comm_struc_combined"), 
+  cache = canopy_cache)
 
+# Set up captions
+# - SI figures
+s_fig_nums <- captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
+s_fig_nums(name = "corr-plot", "Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.")
+s_fig_nums(name = "cwsd", "Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.")
+
+# - SI tables
+s_table_nums <- captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")
+s_table_nums(name = "species-list", "List of ferns from Moorea, French Polynesia included in this study")
+s_table_nums(name = "trait-values", "Trait values of ferns on Moorea, French Polynesia")
+s_table_nums(name = "ancova", "Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate")
+s_table_nums(name = "phylosig-binary-strict", "Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy")
+s_table_nums(name = "corr-evo-strict", "Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy")
+s_table_nums(name = "model-fit", "Best-fitting models explaining community diversity metrics by climate and growth habit")
+
+# Make short versions of citation functions
+s_figure <- pryr::partial(s_fig_nums, display = "cite")
+s_table <- pryr::partial(s_table_nums, display = "cite")
 ```
@@
 
-Authors: Joel H. Nitta, James E. Watkins, Jr., Ravahere Taputuarai, and Charles C. Davis
+Authors: Joel H. Nitta, James E. Watkins, Jr., and Charles C. Davis
 
@@
 
-**Table S1** List of ferns from Moorea, French Polynesia included in this study
+**`r s_table("species-list")`** List of ferns from Moorea, French Polynesia included in this study
 
@@
 'informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them.
-'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith, 2003.
+'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = @Murdock2003.
 
-**Table S2** Trait values of ferns on Moorea, French Polynesia
+**`r s_table("trait-values")`** Trait values of ferns on Moorea, French Polynesia
 
@@
 
-**Table S3** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate
+**`r s_table("ancova")`** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate
 
@@
 
-**Table S4** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
+**`r s_table("phylosig-binary-strict")`** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
 
@@
 
-**Table S5** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
+**`r s_table("corr-evo-strict")`** Pagel's [-@Pagel1994] test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
 
@@
 
-**Table S6** Best-fitting models explaining community diversity metrics by climate and growth habit
+**`r s_table("model-fit")`** Best-fitting models explaining community diversity metrics by climate and growth habit
 
-MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, *P* value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's *I*.
+MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, \pval\ value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's *I*.
 Models within delta AICc 3 of the best-fitting model shown.
@@
 
-**Fig.
-S1** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
+**`r s_figure("corr-plot")`** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
 **(a)**, standard effect size of mean phylogenetic distance (MPD~func~) **(b)**, standard effect size of mean nearest taxon distance (MNTD~func~).
@@
 
-**Fig.
-S2** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
+**`r s_figure("cwsd")`** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
 Epiphytic communities in green, terrestrial communities in brown.

### 2020-01-15 ms/manuscript.Rmd
Update MS
various fixes to output docx correctly

@@
 # Set up captions
+
 # - Figures
@@
 
-# SI figures
-
 # Make short versions of citation functions
@@
 s_table <- pryr::partial(s_table_nums, display = "cite")
-
 ```
@@
 <!-- Work-around to get pandoc to recognize custom siunitx units -->
+<!-- see https://github.com/jgm/pandoc/issues/4773 -->
 \renewcommand{\hour}{h}
@@
 
-^1^Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, Cambridge, MA 02138, U.S.A.
-
-^2^Department of Biology, Colgate University, 13 Oak Drive, Hamilton, New York 13346, U.S.A.
-
-^3^Current affiliation: Department of Botany, National Museum of Natural History, Smithsonian Institution, Washington, D.C., 20013, U.S.A.
+^1^Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, Cambridge, MA 02138, U.S.A.; ^2^Department of Biology, Colgate University, 13 Oak Drive, Hamilton, New York 13346, U.S.A.; ^3^Current address: Department of Botany, National Museum of Natural History, Smithsonian Institution, Washington, D.C., 20013, U.S.A.
 
@@
 *Tel: +1 2026331000*  
-*Email: joelnitta@gmail.com*
+*Email: \textnormal{joelnitta@gmail.com}*
 
@@
 *Tel: +1 6174960515*  
-*Email: cdavis@oeb.harvard.edu*
+*Email: \textnormal{cdavis@oeb.harvard.edu}*
+
+<!-- FIXME: add borders to table in docx -->
 
@@
   words = c(
-    "Total word count (Introduction, Materials and Methods, Results, Discussion):_n",
-    "Summary:_n",
-    "Introduction:_n",
-    "Materials and Methods:_n",
-    "Results:_n",
-    "Discussion:_n",
-    "Acknowledgements:_n"),
+    "Total word count (Introduction, Materials and Methods, Results, Discussion):_6520",
+    "Summary:_198",
+    "Introduction:_924",
+    "Materials and Methods:_2409",
+    "Results:_798",
+    "Discussion:_2389",
+    "Acknowledgements:_170"),
   figs_tabs = c(
@@
   mutate_all(~replace_na(., "")) %>%
-  knitr::kable(format = "latex", booktabs = TRUE, escape = TRUE, col.names = rep("", ncol(.))) %>%
+  knitr::kable(format = "latex", escape = TRUE, col.names = NULL) %>%
   kable_styling(full_width = FALSE) %>%
@@
 Functional diversity is also correlated with species richness [@Schleuter2010]. 
-Thus, environmental variation along elevational gradients is expected to impact the functional, phylogenetic, and species diversity of epiphytic and terrestrial fern communities.
+Thus, environmental variation along elevational gradients is expected to impact the functional, phylogenetic, and taxonomic diversity of epiphytic and terrestrial fern communities.
 
@@
 
-### Study Site
-
-Moorea, French Polynesia (17°30'S, 149°50'W) is a small (\SI{135}{\kilo\meter}^2^) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
+### Study site
+
+Moorea, French Polynesia (17°32'S, 149°50'W) is a small (\SI{135}{\kilo\meter}^2^) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
 This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa{ }\SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa{ }130 \species, 8/11 orders; taxonomy follows @PteridophytePhylogenyGroupI2016; `r s_table("species-list")`).
@@
 
-### Community Survey
+### Community survey
 
@@
 Epiphytic ferns were sampled by hand to \circa{ }\SI{2}{\meter} on tree trunks or confirmed visually from the ground.
-Species were categorized as epiphytic or terrestrial based on field observations (see Trait Selection and Measurements below).
+Species were categorized as epiphytic or terrestrial based on field observations (see 'Trait selection and measurements' below).
 The list of species in each plot was split into epiphytic and terrestrial groups and treated separately during data analysis, hereafter referred to as 'epiphytic' and 'terrestrial' communities.
@@
 
-### Environmental Survey
+### Environmental survey
 
@@
 
-### Trait Selection and Measurements
+### Trait selection and measurements
 
@@
 Morphological traits of gametophytes were all qualitative (either categorical or binary; `r table("traits-used")`).
-Traits were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017] or grown from spores of known species in the lab, taken from the literature, or inferred from taxonomic group (\ie, genus) if documented as a stable taxonomic character.
+Traits were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017], grown from spores of known species in the lab, taken from the literature, or inferred from taxonomic group (\ie, genus) if documented as a stable taxonomic character.
 For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie, heart-shaped) or non-cordate [@Ebihara2013].
@@
 the amount of variance expected under BM: when \K{ }> 1, traits are more conserved than expected under BM; when \K{ }< 1, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
-We tested for significance by comparing observed \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in \plam by conducting a likelihood ratio test comparing the log likelihoods of the observed value of \plam vs.
+We tested for significance in \K by comparing observed values of \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in \plam by conducting a likelihood ratio test comparing the log likelihoods of observed values of \plam vs.
 \plam{ }= 0 (no phylogenetic signal).
 We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' \D{ }[@Fritz2010].
-Values of \D range from 0 under a model of trait evolution by BM, to 1 for random distribution of traits.
-We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], and conducted significance test by comparing observed \D to distributions of simulated values produced under two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM model.
+Values of \D range from 0 under a model of trait evolution by BM, to 1 under random distribution of traits.
+We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], and conducted a significance test by comparing observed \D to distributions of simulated values produced under two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM model.
 To account for possible circularity in measuring the phylogenetic signal of traits that were scored based on taxonomy, we conducted the same analyses on a dataset excluding species that had been scored this way ('strict' dataset).
@@
 We constructed null communities using the 'independent swap' algorithm [@Gotelli2000].
-We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available [`r Ntip(phy)` \species total; @Nitta2017] for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct`) for functional null communities.
+We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available [`r Ntip(phy)` \species total; @Nitta2017] for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct` \species) for functional null communities.
 We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate phylogenetic and functional diversity metrics.
@@
 
-### Principal Components Analysis of Trait Data
+### Principal components analysis of trait data
 
@@
 
-### Phylogenetic Signal
+### Phylogenetic signal
 
@@
 
-### Correlation of Traits with Growth Habit
+### Correlation of traits with growth habit
 
@@
 
-### Functional and Phylogenetic Diversity
+### Functional and phylogenetic diversity
 
@@
 
-We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
+We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` \species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
 Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r figure("comm-div")`).
@@
 Our study reveals a striking duality in the phylogenetic and functional diversity of fern epiphytes. 
-On one hand, we find that fern epiphytes are consistently smaller than terrestrial species at both the species and community level (`r figure("cwm-div")`, `r table("pics")`), and that epiphytic communities are phylogenetically clustered (`r figure("comm-div")`).
+On one hand, we find that fern epiphytes are consistently smaller than terrestrial species at both the species and community level (`r table("pics")`, `r figure("cwm-div")`), and that epiphytic communities are phylogenetically clustered (`r figure("comm-div")`).
 At the same time, we find that epiphytic fern communities are significantly more diverse functionally compared to terrestrial communities (`r figure("comm-div")`).
@@
 
-Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (`r figure("cwm-div")`, `r table("pics")`), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
+Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (`r table("pics")`, `r figure("cwm-div")`), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
 These results are corroborated by @Testo2018a, who similarly observed abrupt size reduction associated with epiphytic growth in a globally sampled fern phylogeny. 
@@
 
-Another clear and convergent pattern to emerge from our study supporting the hypothesis of convergence is the association between epiphytism and the prevalence of non-cordate gametophyte morphotypes (\eg, elongate or filamentous).
+Another clear pattern to emerge from our study supporting the hypothesis of convergence is the association between epiphytism and the prevalence of non-cordate (\eg, elongate or filamentous) gametophyte morphotypes.
 Gametophyte morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived (< \SI{1}{\year}), cordate morphotypes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; @Farrar2008].
@@
 
-Despite the apparent constraints on size and the recurrence of non-cordate gametophytes exhibited by epiphytes, we observed greater morphological diversity among epiphytes versus terrestrial species for several traits.
+Despite the apparent constraints on size and the recurrence of non-cordate gametophytes exhibited by epiphytes, we observed greater morphological diversity among epiphytic versus terrestrial communities for several traits.
 These results do not support the hypothesis that fern epiphytes have reduced functional diversity.
@@
 We did not find that gemmae production in gametophytes was linked with epiphytism. 
-Rather, occurrence of gemmae strongly reflects phylogeny, appearing concentrated in filmy ferns (Hymenophyllaceae) and vittarioid ferns (`r figure("tree")`).
+Rather, occurrence of gemmae strongly reflects phylogeny, appearing concentrated in filmy ferns (Hymenophyllaceae) and vittarioid ferns (Vittarioideae) (`r figure("tree")`).
 Gemmae are asexual propagules occurring primarily in non-cordate gametophytes that allow gametophytes to persist over multiple growing seasons and attain large population sizes, and have been hypothesized to be associated with harsh environmental conditions [@Farrar2008].
@@
 
-Finally, the presence of hairs and glands in gametophytes also exhibit very high phylogenetic signal (\ie, much greater than expected under Brownian motion; `r table("phylosig-binary")`), and these traits are concentrated in certain taxonomic groups [\eg, hairs, grammitid ferns; glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
+Finally, the presence of hairs and glands in gametophytes also exhibit very high phylogenetic signal (\ie, much greater than expected under Brownian motion; `r table("phylosig-binary")`), and these traits are concentrated in certain taxonomic groups [\eg, hairs, grammitid ferns (Grammitidoideae); glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
 Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], and glands may affect rates of osmosis by modifying lipid concentrations [@Crow2011], we did not detect a correlation between hair or gland production and epiphytic growth.
@@
 It is somewhat puzzling that VPD did not contribute significantly to any models, as we hypothesized that this would be important for epiphytes that are frequently subjected to water stress. 
-Previous studies have shown marked variation in the air water potential across epiphytic habitats, with @WatkinsJr.2009 reporting a \circa{ }\SI{40}{\mega\pascal} shift along the trunks of emergent canopy trees.
+Previous studies have shown marked variation in air water potential across epiphytic habitats, with @WatkinsJr.2009 reporting a \circa{ }\SI{40}{\mega\pascal} shift along the trunks of emergent canopy trees.
 Our epiphytic dataloggers only showed slightly greater VPD relative to those on the ground, with this difference largely dissipating at high elevations (`r figure("climate")`).
@@
 Small size appears to be a prerequisite for epiphytic growth, as indicated both by our results and by @Testo2018a.
-However, we have shown that size appears to be the exception, not the rule, when it comes to trait diversity of epiphytes: although epiphytes are generally small, our data suggest they are diverse in numerous other traits.
+However, we have shown that size appears to be the exception, not the rule, when it comes to trait diversity of epiphytic communities: although epiphytes are generally small, our data suggest they are diverse in numerous other traits.
 This emphasizes that fern epiphytes should not be treated as a single, static entity, but rather encompass a wide range of adaptive forms. 
@@
 
-<!-- FIXME: manually remove "Jr" from inline-citations in docx file. -->
+<!-- FIXME: The publisher "Taylor and Francis" gets converted to "Taylor; Francis" by pandoc.
+Fix this in the docx file. -->
 
@@
 
-**`r s_table("species-list")`** List of ferns from Moorea, French Polynesia included in this study.
-
-**`r s_table("trait-values")`** Trait values of ferns on Moorea, French Polynesia.
-
-**`r s_table("ancova")`** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate.
-
-**`r s_table("phylosig-binary-strict")`** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
-
-**`r s_table("corr-evo-strict")`** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.
-
-**`r s_table("model-fit")`** Best-fitting models explaining community diversity metrics by climate and growth habit.
-
-**`r s_figure("corr-plot")`** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
-
-**`r s_figure("cwsd")`** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
+**`r s_table("species-list")`** List of ferns from Moorea, French Polynesia included in this study
+
+**`r s_table("trait-values")`** Trait values of ferns on Moorea, French Polynesia
+
+**`r s_table("ancova")`** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate
+
+**`r s_table("phylosig-binary-strict")`** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
+
+**`r s_table("corr-evo-strict")`** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
+
+**`r s_table("model-fit")`** Best-fitting models explaining community diversity metrics by climate and growth habit
+
+**`r s_figure("corr-plot")`** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input
+
+**`r s_figure("cwsd")`** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia
 
@@
 # FIXME: need to manually add "Sporophyte traits" and "Gametophyte traits" to
-# table in docx (this happens with gt or kable).
+# table in docx (this happens with gt or kable), and fix italics, superscript, etc.
 read_csv(here::here("data/table_1.csv")) %>%
@@
 \D, Fritz and Purvis' \D{ }[@Fritz2010]; *P*(rnd), probability of random distribution of traits; *P*(BM), probability of traits evolving under Brownian motion (BM).
-Values of \D 0 or less indicate evolution according to BM; values near 1 indicate random trait distribution.
+Values of \D{ }0 or less indicate evolution according to BM; values near 1 indicate random trait distribution.
 
@@
 
+<!-- FIXME: delete images from docx manually -->
+
 \clearpage

### 2020-01-15 ms/SI.rmd
Submit SI as pdf, not docx

@@
       in_header: new-phytologist.sty
+header-includes: 
+  - \usepackage{float}
+  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # Fix position of figs
+  - \makeatletter\renewcommand*{\fps@table}{H}\makeatother # Fix position of tables
 nocite: | 
@@
 
-Article acceptance date:
+Article acceptance date: ADD_DATE_WHEN_ACCEPTED
 
@@
 
-**`r s_figure("corr-plot")`** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
-**(a)**, standard effect size of mean phylogenetic distance (MPD~func~) **(b)**, standard effect size of mean nearest taxon distance (MNTD~func~).
-For details on calculations of MPD~func~ and MNTD~func~, see Methods.
-Correlations (values in each tile) calculated with Pearson's correlation coefficient using the R package 'corrr'.
-All trait sets were scaled to standard deviation of each trait.
-The following trait sets were used: 'all' (all traits), 'size' (sporophyte traits related to size; frond length and width, stipe length, rhizome width, pinna number, SLA, degree of lamina dissection), 'other' (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), 'gameto' (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), 'reduced' (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter).
-For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), both log-transformed and untransformed sets were included.
-The log-transformed 'reduced' dataset was used for the main results.
+\clearpage
 
-```{r fig-s2}
-library(corrr)
+```{r Fig-S1, fig.height = 8, fig.width = 4, fig.cap = ""}
 
@@
 
-ggsave(plot = fig_s1, filename = here::here("results/Fig_S1.eps"), height = 8, width = 4, units = "in")
+fig_s1
+
+# ggsave(plot = fig_s1, filename = here::here("results/Fig_S1.eps"), height = 8, width = 4, units = "in")
+```
+
+**`r s_figure("corr-plot")`** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
+**(a)**, standard effect size of mean phylogenetic distance (MPD~func~) **(b)**, standard effect size of mean nearest taxon distance (MNTD~func~).
+For details on calculations of MPD~func~ and MNTD~func~, see Methods.
+Correlations (values in each tile) calculated with Pearson's correlation coefficient using the R package 'corrr'.
+All trait sets were scaled to standard deviation of each trait.
+The following trait sets were used: 'all' (all traits), 'size' (sporophyte traits related to size; frond length and width, stipe length, rhizome width, pinna number, SLA, degree of lamina dissection), 'other' (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), 'gameto' (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), 'reduced' (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter).
+For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), both log-transformed and untransformed sets were included.
+The log-transformed 'reduced' dataset was used for the main results.
+
+\clearpage
+
+```{r Fig-S2, fig.height = 6, fig.width = 7, fig.cap = ""}
+loadd(cwsd_scatterplot, cache = canopy_cache)
+
+cwsd_scatterplot
+
+# ggsave(plot = cwsd_scatterplot, filename = here::here("results/Fig_S2.eps"), height = 6, width = 7, units = "in")
 ```
@@
 
-```{r fig-s1}
-loadd(cwsd_scatterplot, cache = canopy_cache)
-ggsave(plot = cwsd_scatterplot, filename = here::here("results/Fig_S2.eps"), height = 6, width = 7, units = "in")
-```
+\clearpage
 

### 2020-01-15 code/plan.R

@@ -531,13 +531,6 @@ plan <- drake_plan(
     quiet = TRUE,
     output_dir = here::here("results"),
     tracked_output = file_out(here::here("results/SI.tex"))
-  ),
-  
-  si_docx = latex2docx(
-    latex = file_in(here::here("results/SI.tex")),
-    docx = file_out(here::here("results/SI.docx")),
-    template = file_in(here::here("ms/new-phytologist.docx")),
-    wd = here::here("results")
   )
   
 )

### 2020-01-15 ms/manuscript.Rmd
Update MS
- Fix typo in SI units

@@
 
-**`r figure("climate")`** Selected microclimatic variables of study plots along an elevational gradient from \SI{200}{\meter} to \SI{1200}\meter on Moorea, French Polynesia.
+**`r figure("climate")`** Selected microclimatic variables of study plots along an elevational gradient from \SI{200}{\meter} to \SI{1200}{\meter} on Moorea, French Polynesia.
 Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from `r final_climate_survey$start` to `r final_climate_survey$end`.

### 2020-01-27 code/functions.R
Make PCA plot geom_text labels smaller, points grey

@@ -2150,8 +2150,13 @@ make_pca_plot <- function (pca_results, habit_colors, traits) {
     ggplot(aes(x = PC1, y = PC2)) +
     geom_vline(xintercept = 0, linetype = "longdash", color = "dark gray") +
     geom_hline(yintercept = 0, linetype = "longdash", color = "dark gray") +
-    geom_point(size=3, shape=16) +
-    geom_text_repel(aes(label = trait), show.legend=FALSE, segment.colour = NA) +
+    geom_point(size=3, shape=16, color = "gray50") +
+    geom_text_repel(
+      aes(label = trait),
+      size = 3,
+      show.legend = FALSE,
+      segment.color = "transparent"
+    ) +
     scale_x_continuous(
       limits = c(-1,1)
     ) +
@@ -2172,8 +2177,13 @@ make_pca_plot <- function (pca_results, habit_colors, traits) {
     ggplot(aes(x = PC1, y = PC2)) +
     geom_vline(xintercept = 0, linetype = "longdash", color = "dark gray") +
     geom_hline(yintercept = 0, linetype = "longdash", color = "dark gray") +
-    geom_point(size=3, shape=16) +
-    geom_text_repel(aes(label = trait), show.legend=FALSE, segment.colour = NA) +
+    geom_point(size=3, shape=16, color = "gray50") +
+    geom_text_repel(
+      aes(label = trait),
+      size = 3,
+      show.legend = FALSE,
+      segment.color = "transparent"
+    ) +
     scale_x_continuous(
       limits = c(-1,1)
     ) +

### 2020-01-27 ms/manuscript.Rmd
Revise MS as per submission V2

@@
   words = c(
-    "Total word count (Introduction, Materials and Methods, Results, Discussion):_6520",
-    "Summary:_198",
-    "Introduction:_924",
-    "Materials and Methods:_2409",
+    "Total word count (Introduction, Materials and Methods, Results, Discussion):_6498",
+    "Summary:_199",
+    "Introduction:_881",
+    "Materials and Methods:_2410",
     "Results:_798",
-    "Discussion:_2389",
+    "Discussion:_2409",
     "Acknowledgements:_170"),
@@
 The majority of modern fern lineages are hypothesized to have arisen in response to this expansion, particularly fern epiphytes that radiated into the canopy.
-Recent evidence, however, suggests that epiphytism does not correlate with increased diversification rates in ferns, calling into question the role of the canopy habitat in their diversification.
-
-- To bridge this impasse, we investigated functional traits of fern sporophytes and gametophytes at the species and community level across a broad phylogenetic sampling on the island of Moorea, French Polynesia, including > 120 species and representatives of five major epiphytic radiations.
+Recent evidence, however, suggests that epiphytism does not correlate with increased diversification rates in ferns, calling into question the role of the canopy habitat in fern evolution.
+
+- To understand the role of the canopy in structuring fern community diversity, we investigated functional traits of fern sporophytes and gametophytes across a broad phylogenetic sampling on the island of Moorea, French Polynesia, including > 120 species and representatives of multiple epiphytic radiations.
 
@@
 The heterogeneous niche space of the angiosperm canopy comprising stratified light and humidity levels and a range of substrates for growth became rapidly filled by a variety of epiphytic plants, including mosses [@Fiz-Palacios2011], leafy liverworts [@Feldberg2014], orchids [@Givnish2015b], bromeliads [@Givnish2014a], and ferns [@Schneider2004d].
-The transition to epiphytic growth was particularly important for modern fern diversification: \circa{ }29% of all fern species are epiphytes vs. \circa{ }10% across all vascular plants [@Kress1986].
+The transition to epiphytic growth was particularly important for modern fern species diversity: \circa{ }29% of all fern species are epiphytes vs.
+\circa{ }10% across all vascular plants [@Kress1986].
 Moreover, major transitions to epiphytism occurred at least five times within ferns, each followed by diversification during the late Cretaceous or early Cenozoic [@Schuettpelz2009; @WatkinsJr.2012].
 This has led to the hypothesis that epiphytism was a key factor contributing to extant fern diversification [@Schneider2004d; @Schuettpelz2009; @WatkinsJr.2012].
-Recent studies exploring this idea [@Sundue2015; @Testo2016a; @Lehtonen2017], however, have largely failed to identify evidence of rapid diversification associated with epiphytism in ferns. 
-This raises the possibility that modern fern diversification was not shaped by the evolution of epiphytism *per se*, but rather by the exploitation of various niches within the canopy and on the forest floor.
-
-To better understand the role of the canopy in fern diversification, one obvious direction is to explore putatively adaptive fern epiphyte traits in a phylogenetic context [@Givnish2015]. 
-In contrast to most terrestrial plants, epiphytes must contend with greater stress in the form of nutrient limitation, increased evaporative demand, drought, and extreme oscillations in temperature [@Zotz2001; @WatkinsJr.2009; @Lowman2012].
-Previous studies have identified multiple traits as likely adaptations to these conditions, albeit many without consideration of phylogeny [\eg, @Kluge2007]. 
-Epiphytic ferns tend to have morphological features that minimize water loss from their leaves, including shorter stipes [@WatkinsJr.2010], thicker leaves [@Kluge2007; @WatkinsJr.2007b], and protection with scales or hairs [@WatkinsJr.2006b].
-Additional specialized adaptations to epiphytic growth include humus collection (\eg, *Asplenium*; @Benzing1990, *Drynaria*; @Janssen2005), vegetative desiccation tolerance (\eg, Hymenophyllaceae; @Shreve1911, *Pleopeltis*; @Pessin1925), CAM photosynthesis (\eg, *Pyrrosia*; @Wong1976), and symbiotic relationships with ants (\eg, *Antrophyum*; @WatkinsJr.2008, *Lecanopteris*; @Gay1993a).
-
-While many similar adaptations are found in epiphytic seed plants and serve as key reference points [@Benzing1990], the evolution of epiphytism in ferns is more complex owing to their unique life-cycle.
-Ferns produce a free-living gametophyte that is typically smaller than the sporophyte and relies on aquatic motile sperm.
-Unlike the sporophyte, the fern gametophyte lacks vascular tissue, stomata, or cuticle, placing it in constant equilibrium with its external environment. 
-In parallel with some of the adaptations in the sporophytes, the gametophytes of some epiphytic ferns also exhibit morphological and physiological adaptations to minimize water loss and tolerate water scarcity.
-For example, extreme desiccation tolerance is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is rare in sporophytes [@WatkinsJr.2007c; @Pittermann2013].
-In addition, epiphytic fern gametophytes tend to exhibit more complex three-dimensional morphologies that may function to retain water and slow its loss [@WatkinsJr.2007c].
-Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no previous studies to our knowledge have analyzed the functional ecology of fern epiphytes in a broad phylogenetic context including both life stages.
-
-To gain insight into the role of the canopy in promoting or restricting diversity, a functional and community-based approach can be applied to study this question. 
-If the canopy is uniformly harsher than the terrestrial environment, thus acting as a 'filter', we would expect functional convergence of epiphytes at the community level, which in turn may result in decreased phylogenetic diversity of epiphytic communities if traits are phylogenetically conserved [@Webb2002; @Cavender-Bares2009].
-Conversely, the stratification of the canopy may stimulate functional diversification by niche partitioning, leading to increased functional diversity of epiphytic communities. 
-Furthermore, the role of the canopy as a promoter or restrictor of diversity may be modulated by overall abiotic conditions, including most notably relative humidity (RH) and temperature. 
+Recent studies exploring this idea [@Sundue2015; @Testo2016a; @Lehtonen2017; @Testo2018a], however, have largely failed to identify evidence of rapid diversification associated with epiphytism in ferns, calling into question the role of the canopy habitat in fern evolution.
+
+While shifts in diversification rate are relevant, they are only one way epiphytic growth may have influenced fern evolution.
+The impact of epiphytic growth could manifest in other ways, such as by shaping phylogenetic and functional diversity at the community level. 
+For example, if the canopy is uniformly harsher than the terrestrial environment, we would expect functional convergence of epiphytes at the community level, which in turn may result in decreased phylogenetic diversity of epiphytic communities if traits are phylogenetically conserved [@Webb2002; @Cavender-Bares2009].
+Conversely, the stratification of the canopy may stimulate functional diversification by niche partitioning, leading to greater functional diversity within epiphytic communities. 
+Furthermore, the role of the canopy as a promoter or inhibitor of community diversity may be modulated by overall abiotic conditions, including most notably relative humidity (RH) and temperature. 
 For instance, elevation is a strong predictor of fern species richness, with maximal richness occurring at mid-elevations, especially in the tropics [@Kromer2005; @Kessler2011a]. 
@@
 
-Here, we seek to understand the role of the canopy in fern diversification by investigating traits related to epiphytic growth in fern gametophytes and sporophytes on Moorea, French Polynesia.
+Fern epiphytes possess numerous adaptations that may contribute to their functional diversity.
+In contrast to most terrestrial plants, epiphytes must contend with greater stress in the form of nutrient limitation, increased evaporative demand, drought, and extreme oscillations in temperature [@Zotz2001; @WatkinsJr.2009; @Lowman2012].
+Epiphytic ferns tend to have morphological features that minimize water loss from their leaves, including shorter stipes [@WatkinsJr.2010], thicker leaves [@Kluge2007; @WatkinsJr.2007b], and protection with scales or hairs [@WatkinsJr.2006b].
+Additional specialized adaptations to epiphytic growth include humus collection (\eg, *Asplenium*; @Benzing1990, *Drynaria*; @Janssen2005), vegetative desiccation tolerance (\eg, Hymenophyllaceae; @Shreve1911, *Pleopeltis*; @Pessin1925), CAM photosynthesis (\eg, *Pyrrosia*; @Wong1976), and symbiotic relationships with ants (\eg, *Antrophyum*; @WatkinsJr.2008, *Lecanopteris*; @Gay1993a).
+
+While many similar adaptations are found in epiphytic seed plants and serve as key reference points [@Benzing1990], the evolution of epiphytism in ferns is more complex owing to their unique life-cycle.
+Ferns produce a free-living gametophyte that is typically smaller than the sporophyte and relies on aquatic, motile sperm.
+Unlike the sporophyte, the fern gametophyte lacks vascular tissue, stomata, or a well developed cuticle, placing it in constant equilibrium with its external environment. 
+In parallel with some of the adaptations of sporophytes, the gametophytes of some epiphytic ferns also exhibit morphological and physiological adaptations to minimize water loss and tolerate water scarcity.
+For example, extreme desiccation tolerance is prevalent in gametophytes of multiple lineages of epiphytic ferns, but is rare in sporophytes [@WatkinsJr.2007c; @Pittermann2013].
+In addition, epiphytic fern gametophytes tend to exhibit complex three-dimensional morphologies that may function to retain water and slow its loss [@WatkinsJr.2007c].
+Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no previous studies to our knowledge have analyzed the functional ecology of fern epiphytes in a broad phylogenetic context including both life stages.
+
+Here, we seek to understand the role of the canopy in structuring fern community diversity by investigating putatively functional traits related to epiphytic growth in fern gametophytes and sporophytes on Moorea, French Polynesia.
 Moorea is an isolated tropical oceanic island with high phylogenetic diversity of ferns distributed across a steep elevational gradient, making it ideal for a community phylogenetic analysis of functional traits.
-Specifically, we test the following hypotheses related to epiphytic growth in ferns: 1) Epiphytic ferns have converged upon similar morphologies in both sporophyte and gametophyte phases; 2) Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities.
+Specifically, we test the following hypotheses related to epiphytic growth in ferns: 1) Epiphytic ferns have convergent morphologies in both sporophyte and gametophyte phases; 2) Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities.
 
@@
 Moorea, French Polynesia (17°32'S, 149°50'W) is a small (\SI{135}{\kilo\meter}^2^) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
-This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island (\circa{ }\SI{1.5}{\ma} [@Duncan1976]) has led to a relatively small, yet phylogenetically diverse fern flora (\circa{ }130 \species, 8/11 orders; taxonomy follows @PteridophytePhylogenyGroupI2016; `r s_table("species-list")`).
-Ferns of Moorea include representatives of multiple Cretaceous epiphytic radiations (\eg, *Asplenium*, *Elaphoglossum*, Hymenophylloideae, Polypodiaceae, Vittarioideae; @Schuettpelz2009) and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
+This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island [\circa{ }\SI{1.5}{\ma}; @Duncan1976] has led to a relatively small, yet phylogenetically diverse fern flora [\circa{ }130 \species, 8/11 orders; `r s_table("species-list")`; taxonomy follows @PteridophytePhylogenyGroupI2016].
+Ferns of Moorea include representatives of multiple Cretaceous epiphytic radiations [\eg, *Asplenium*, *Elaphoglossum*, Hymenophylloideae, Polypodiaceae, Vittarioideae; @Schuettpelz2009] and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
 
@@
 The list of species in each plot was split into epiphytic and terrestrial groups and treated separately during data analysis, hereafter referred to as 'epiphytic' and 'terrestrial' communities.
-Vouchers were deposited at GH and UC (abbreviations follow Index Herbariorum; @Thiers2019).
+Vouchers were deposited at GH and UC [abbreviations follow Index Herbariorum; @Thiers2019].
 
@@
 
-We selected several traits relevant to epiphytic growth (`r table("traits-used")` and `r figure("traits-used-image")`).
-We preferred morphological traits over physiological traits because data for morphological traits are easier to gather at the scale of our study (\circa{ }130 \species distributed over an elevational gradient from \SI{200}{\meter} to \SI{2000}{\meter}).
+We selected several traits putatively relevant to epiphytic growth (`r table("traits-used")` and `r figure("traits-used-image")`).
+We preferred morphological traits over physiological traits because data for morphological traits are easier to gather at the scale of our study (\circa{ }130 \species distributed over an elevational gradient from \SI{200}{\meter} to \SI{1200}{\meter}).
 Plant morphology and physiology are closely linked, and physiological traits such as rates of photosynthesis and water use efficiency are known to be correlated with morphological traits such as stipe length in ferns [@WatkinsJr.2010].
@@
 At the same time, we find that epiphytic fern communities are significantly more diverse functionally compared to terrestrial communities (`r figure("comm-div")`).
-Collectively, these results suggest that environmental filtering plays a strong role in limiting plant size, while simultaneously promoting diversification across other trait axes in epiphytic ferns.
+Collectively, these results suggest that environmental filtering plays a strong role in limiting plant size, while simultaneously supporting functional diversity across other trait axes in epiphytic ferns.
 
@@
 We expected that epiphytes would exhibit lower SLA, given their more extreme growing conditions.
-In contrast, we observed overlapping, wide ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes (`r figure("cwm-div")`).
-This may reflect adaptations to specific niches that are not captured in the binary categorization of habitat commonly applied in ferns. 
+In contrast, we observed overlapping, wide ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes when taking elevation into account (`r figure("cwm-div")`).
+This may reflect adaptations to specific niches that are not captured by a binary categorization of growth habit.
 For example, the following species ranked among those with the five lowest SLA values (\ie, thick, well-protected leaves) on Moorea: *Humata anderssonii* (Mett.) C.Chr.
@@
 
-However, while the canopy clearly acts as a filter, we also find that it simultaneously functions as a promoter of diversity. 
+However, while the canopy clearly acts as a filter, we also find that it simultaneously functions as a promoter of functional diversity. 
 This is evident in our analyses of MPD~func~ and MNTD~func~, which both indicate that epiphytic communities are more functionally diverse compared to terrestrial species (`r figure("comm-div")`), thus contradicting our hypothesis of decreased functional diversity in epiphytic communities.
@@
 Recent studies [\eg, @Sundue2015; @Testo2016a; @Lehtonen2017; @Testo2018a] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
-Our study provides one possible explanation for this finding. 
-Here, we demonstrate that epiphytism is likely a more complex trait than previously realized, and almost certainly requires the evolution of multiple associated and dependent traits. 
-We propose that the timing and order of the evolution of such traits are related to the complex underlying diversification dynamics observed in these studies. 
-Future research investigating the tempo and mode of the evolution of epiphytism in ferns should seek to better partition this diversity when exploring hypotheses of exceptional diversification.
+Our study provides a fresh perspective on this issue. 
+Here, we demonstrate that epiphytism is likely a complex trait, and almost certainly requires the evolution of multiple associated and dependent traits. 
+We propose that the timing and order of the evolution of such traits are related to the conflicting diversification dynamics observed in these studies.
+Another possibility raised by our results is that fern epiphytes have radiated morphologically without corresponding species-level diversification [@Givnish2015].
+Future research investigating the tempo and mode of the evolution of epiphytism in ferns should seek to better partition this functional diversity when exploring hypotheses of exceptional species diversification.
 
@@
 Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval{ }< 0.05.
+Partially transparent colors indicate outliers excluded from the models (see Methods).
 
@@
 \clearpage
-
+<!-- FIXME: Run the code below from Rstudio within the container, or lines along X-Y axes won't print -->
 ```{r pca, fig.height = 6, fig.width = 6}

### 2020-01-28 ms/manuscript.Rmd
Revise MS as per submission V2

@@
   words = c(
-    "Total word count (Introduction, Materials and Methods, Results, Discussion):_6498",
+    "Total word count (Introduction, Materials and Methods, Results, Discussion):_6499",
     "Summary:_199",
@@
     "Results:_798",
-    "Discussion:_2409",
+    "Discussion:_2410",
     "Acknowledgements:_170"),
@@
 
-### The dual roles of the canopy as both a filter and promoter of diversity
+### The dual roles of the canopy as both a filter and promoter of community diversity
 
@@
 Another possibility raised by our results is that fern epiphytes have radiated morphologically without corresponding species-level diversification [@Givnish2015].
-Future research investigating the tempo and mode of the evolution of epiphytism in ferns should seek to better partition this functional diversity when exploring hypotheses of exceptional species diversification.
+Future research investigating the tempo and mode of the evolution of epiphytism in ferns should seek to better partition this morphological diversity when exploring hypotheses of exceptional species diversification.
 

### 2020-02-07 ms/SI.rmd
For SI load captions separately and add header and footer to tables

@@
 
-# Set up captions
-# - SI figures
-s_fig_nums <- captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
-s_fig_nums(name = "corr-plot", "Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.")
-s_fig_nums(name = "cwsd", "Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.")
+# Load captions
+source(here::here("ms/captions.R"))
 
-# - SI tables
-s_table_nums <- captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")
-s_table_nums(name = "species-list", "List of ferns from Moorea, French Polynesia included in this study")
-s_table_nums(name = "trait-values", "Trait values of ferns on Moorea, French Polynesia")
-s_table_nums(name = "ancova", "Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate")
-s_table_nums(name = "phylosig-binary-strict", "Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy")
-s_table_nums(name = "corr-evo-strict", "Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy")
-s_table_nums(name = "model-fit", "Best-fitting models explaining community diversity metrics by climate and growth habit")
+# Set up table headers
+si_table_header <- c(
+  "New Phytologist Supporting Information",
+  "Authors: Joel H. Nitta, James E. Watkins, Jr., and Charles C. Davis" %>% wrap_quotes,
+  "Article acceptance date: 5 February 2020"
+)
+```
 
-# Make short versions of citation functions
-s_figure <- pryr::partial(s_fig_nums, display = "cite")
-s_table <- pryr::partial(s_table_nums, display = "cite")
-```
+<!-- Set left justification  -->
+\raggedright
+
+<!-- Don't indent -->
+\setlength{\parindent}{0ex}
 
@@
 
-Article acceptance date: ADD_DATE_WHEN_ACCEPTED
+Article acceptance date: 5 February 2020
 
@@
 
-**`r s_table("species-list")`** List of ferns from Moorea, French Polynesia included in this study
+**`r s_table("species-list")`** `r s_table_cap("species-list")` (see separate file)
 
@@
 'informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them.
-'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = @Murdock2003.
+'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith (2003).
+'NA' indicates not applicable.
 
-**`r s_table("trait-values")`** Trait values of ferns on Moorea, French Polynesia
+```{r species-list}
+# Set header and footer
+header <- c(
+  si_table_header,
+  s_table_nums("species-list") %>% wrap_quotes
+)
+
+# Wrap footer at ca. 100 characters per line. Each line needs to be an element of the
+# character vector.
+footer <- "'taxon code' is the name used for that taxon in analyses.
+'informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them.
+'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith (2003).
+'NA' indicates not applicable." %>%
+  str_wrap(width = 100) %>%
+  str_split("\\n") %>%
+  unlist %>%
+  purrr::map_chr(wrap_quotes)
+
+# Write out table with metadata
+write_csv_with_meta(
+  taxonomic_data_for_si, 
+  here::here("results/Table_S1.csv"), 
+  header,
+  footer)
+```
+
+**`r s_table("trait-values")`** `r s_table_cap("trait-values")` (see separate file)
 
@@
 For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI.
+'NA' indicates missing data.
 
-```{r}
-trait_data_for_si %>%
-  write_csv(here::here("results/Table_S2.csv"))
+```{r trait_data}
+# Set header and footer
+header <- c(
+  si_table_header,
+  s_table_nums("trait-values") %>% wrap_quotes
+)
+
+footer <- "SLA, specific leaf area.
+Continuous measurements shown as mean, standard deviation (sample size).
+'taxon code' is the name used for that taxon in analyses (see Table S1).
+For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI.
+'NA' indicates missing data." %>%
+  str_wrap(width = 100) %>%
+  str_split("\\n") %>%
+  unlist %>%
+  purrr::map_chr(wrap_quotes)
+
+# Write out table with metadata
+write_csv_with_meta(
+  trait_data_for_si, 
+  here::here("results/Table_S2.csv"), 
+  header,
+  footer)
 ```
 
-**`r s_table("ancova")`** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate
+**`r s_table("ancova")`** `r s_table_cap("ancova")` (see separate file)
 
@@
 ```{r climate-ancova}
-# Format and write out table as CSV file
+# Format table
+climate_ancova_results_formatted <-
 climate_ancova_results %>% 
@@
   ) %>%
-  select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value) %>%
-  write_csv(here::here("results/Table_S3.csv"))
+  select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value)
 
+# Set header and footer
+header <- c(
+  si_table_header,
+  s_table_nums("ancova") %>% wrap_quotes
+)
+
+footer <- "d.f., degrees of freedom; MS, mean squares." %>% wrap_quotes
+
+# Write out table with metadata
+write_csv_with_meta(
+  climate_ancova_results_formatted, 
+  here::here("results/Table_S3.csv"), 
+  header,
+  footer)
 ```
 
-**`r s_table("phylosig-binary-strict")`** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
+**`r s_table("phylosig-binary-strict")`** `r s_table_cap("phylosig-binary-strict")` (see separate file)
 
@@
 ```{r phylosig-binary-strict}
-# Format and write out table as CSV file
+# Format table
+phylosig_binary_traits_strict_formatted <-
   phylosig_binary_traits_strict %>%
@@
   select(Trait = trait, `Number of presences` = num_present, `Number of absences` = num_absent, D,
-         P_rnd = prob_random, P_bm = prob_brownian) %>%
-  write_csv(here::here("results/Table_S4.csv"))
+         P_rnd = prob_random, P_bm = prob_brownian)
+
+# Set header and footer
+header <- c(
+  si_table_header,
+  s_table_nums("phylosig-binary-strict") %>% wrap_quotes
+)
+
+footer <- "P_rnd, probability of obtaining the observed value of D under random trait evolution; P_bm, probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion." %>%
+  str_wrap(width = 100) %>%
+  str_split("\\n") %>%
+  unlist %>%
+  purrr::map_chr(wrap_quotes)
+
+# Write out table with metadata
+write_csv_with_meta(
+  phylosig_binary_traits_strict_formatted, 
+  here::here("results/Table_S4.csv"), 
+  header,
+  footer)
 ```
 
-**`r s_table("corr-evo-strict")`** Pagel's [-@Pagel1994] test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
+**`r s_table("corr-evo-strict")`** `r s_table_cap("corr-evo-strict")` (see separate file)
 
@@
 ```{r corr-evo-strict}
-# Format and write out table as CSV file
+# Format table
+correlated_evo_test_strict_formatted <-
   correlated_evo_test_strict %>%
@@
     `P` = pval
-  ) %>%
-  write_csv(here::here("results/Table_S5.csv"))
+  )
+
+# Set header and footer
+header <- c(
+  si_table_header,
+  s_table_nums("corr-evo-strict") %>% wrap_quotes
+)
+
+footer <- "LL, log likelihood.
+Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
+terrestrial).
+Lower log likelihood for the dependent model indicates correlation with growth habit." %>%
+  str_wrap(width = 100) %>%
+  str_split("\\n") %>%
+  unlist %>%
+  purrr::map_chr(wrap_quotes)
+
+# Write out table with metadata
+write_csv_with_meta(
+  correlated_evo_test_strict_formatted, 
+  here::here("results/Table_S5.csv"), 
+  header,
+  footer)
 ```
 
-**`r s_table("model-fit")`** Best-fitting models explaining community diversity metrics by climate and growth habit
+**`r s_table("model-fit")`** `r s_table_cap("model-fit")` (see separate file)
 
@@
 
-```{r best-fit-mods}
-# Format and write out table as CSV file
+```{r model-fit}
+# Format table
+best_fit_div_models_moran_formatted <-
 best_fit_div_models_moran %>%
@@
     `P_moran` = morans_I_pval
-  ) %>%
-  write_csv(here::here("results/Table_S6.csv"))
+  )
+
+# Set header and footer
+header <- c(
+  si_table_header,
+  s_table_nums("model-fit") %>% wrap_quotes
+)
+
+footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
+Models within delta AICc 3 of the best-fitting model shown.
+Trait names refer to community-weighted mean values (see Methods).
+Frond length and width were highly correlated with rhizome diameter and not shown.
+For details on calculations of MPD and MNTD, see Methods." %>%
+  str_wrap(width = 100) %>%
+  str_split("\\n") %>%
+  unlist %>%
+  purrr::map_chr(wrap_quotes)
+
+# Write out table with metadata
+write_csv_with_meta(
+  best_fit_div_models_moran_formatted, 
+  here::here("results/Table_S6.csv"), 
+  header,
+  footer)
+
 ```
@@
 
-```{r Fig-S1, fig.height = 8, fig.width = 4, fig.cap = ""}
+```{r Fig-S1, fig.height = 6, fig.width = 7, fig.cap = ""}
+loadd(cwsd_scatterplot, cache = canopy_cache)
+
+cwsd_scatterplot
+```
+
+**`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`.
+Epiphytic communities in green, terrestrial communities in brown.
+Points randomly jittered along x-axis to reduce overlap.
+Boxplots show median values (bold lines) by growth habit.
+Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range.
+Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \* = \pval{ }< 0.05 **(a)**, traits related to plant size **(b)**, traits not related to plant size.
+Response variable abbreviations: SLA, specific leaf area.
+
+\clearpage
+
+```{r Fig-S2, fig.height = 8, fig.width = 4, fig.cap = ""}
 

### 2020-02-07 code/functions.R

@@ -3142,7 +3142,32 @@ test_corr_evo_range_gemmae <- function(community_matrix_path, phy, traits, moore
   
 }
 
-# MS rendering
+#' Write out a csv file with metadata appended to start and end
+#' 
+#' Shamelessly copied from SO (and modified)
+#' https://stackoverflow.com/questions/12381117/add-header-to-file-created-by-write-csv
+#' 
+#' @param x Dataframe to write out
+#' @param file Path to write file
+#' @param header Character vector: header to append to CSV file
+#' @param footer Character vector: footer to append to CSV file
+#' @param ... Additional arguments passed to readr::write_csv()
+#'
+#' @return Nothing; externally, `x` will be writted to `file`
+#'
+write_csv_with_meta <- function(x, file, header, footer, ...){
+  # create and open the file connection
+  datafile <- file(file, open = 'wt')
+  # close on exit
+  on.exit(close(datafile))
+  # if a header is defined, write it to the file (@CarlWitthoft's suggestion)
+  if(!missing(header)) writeLines(header,con = datafile)
+  # write the file using the defined function and required addition arguments  
+  readr::write_csv(x = x, path = datafile, ...)
+  # Add footer
+  if(!missing(footer)) writeLines(footer, con = datafile)
+}
+
 
 # Dummy function to track arbitary output from rmarkdown::render()
 render_tracked <- function (tracked_output, ...) {

### 2020-02-07 ms/captions.R

@@ -1,0 +1,45 @@
+# Set up captions
+
+# - Figures
+fig_nums <- captioner::captioner(prefix = "Fig.")
+fig_nums(name = "traits-used-image", caption = "Examples of traits used in this study.")
+fig_nums(name = "climate", caption = "Microclimatic variables.")
+fig_nums(name = "pca", caption = "Trait PCA.")
+fig_nums(name = "tree", caption = "Traits on tree.")
+fig_nums(name = "heatmap", caption = "Variable importance scores.")
+fig_nums(name = "cwm-div", caption = "Community-weighted means.")
+fig_nums(name = "comm-div", caption = "Community diversity metrics.")
+
+# - Tables
+table_nums <- captioner::captioner(prefix = "Table")
+table_nums(name = "traits-used", caption = "Fern traits used in this study")
+table_nums(name = "phylosig-cont", caption = "Phylogenetic signal in continuous traits of ferns from Moorea, French Polynesia")
+table_nums(name = "phylosig-binary", caption = "Phylogenetic signal in binary traits of ferns from Moorea, French Polynesia")
+table_nums(name = "pics", caption = "Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth in ferns from Moorea, French Polynesia")
+table_nums(name = "corr-evo", caption = "Pagel's (1994) test of correlated evolution between binary traits in ferns from Moorea, French Polynesia")
+
+# - SI figures
+s_fig_nums <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
+s_fig_nums(name = "cwsd", "Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia")
+s_fig_nums(name = "corr-plot", "Correlation plot comparing functional diversity of fern communities on Moorea, French Polynesia calculated using different sets of traits as input")
+
+# - SI tables
+s_table_nums <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")
+s_table_nums(name = "species-list", "List of ferns from Moorea, French Polynesia included in this study")
+s_table_nums(name = "trait-values", "Trait values of ferns on Moorea, French Polynesia")
[…]

### 2020-02-07 code/functions.R
Resolve names in plan
- remove Microsorum x-maximum from Table S1

@@ -3168,6 +3168,98 @@ write_csv_with_meta <- function(x, file, header, footer, ...){
   if(!missing(footer)) writeLines(footer, con = datafile)
 }
 
+# Wrap a vector in quotes. Useful for writing out strings to CSV file that you
+# don't want split across columns
+wrap_quotes <- function (x) {paste0("\"", x, "\"")}
+
+#' Reformat a vector of species names into a dataframe
+#' with columns for genus, specific epithet, etc
+#'
+#' @param species_list Vector of species names
+#'
+#' @return Tibble
+#' 
+reformat_species_names <- function (species_list) {
+  
+    tibble(taxon_code = species_list) %>%
+    separate(taxon_code, c("genus", "specific_epithet"), sep = "_", remove = FALSE) %>%
+    mutate(informal_variety = str_match(specific_epithet, "[0-9]") %>% map_chr(1)) %>%
+    mutate(specific_epithet = str_remove_all(specific_epithet, "[0-9]")) %>%
+    mutate(species = paste(genus, specific_epithet)) %>%
+    mutate(informal_variety = case_when(
+      specific_epithet == "sp" ~ NA_character_,
+      TRUE ~ informal_variety
+    )) %>%
+    mutate(infaspecific_epithet = case_when(
+      species == "Lindsaea repens" ~ "marquesensis",
+      species == "Davallia solida" ~ "solida",
+      species == "Ctenitis sciaphila" ~ "sciaphila",
+      species == "Dicranopteris linearis" ~ "linearis",
+      species == "Deparia petersenii" ~ "congrua",
+      TRUE ~ NA_character_
+    )) %>%
+    mutate(taxon = jntools::paste3(genus, specific_epithet, infaspecific_epithet))
+  
+}
+
+#' Lookup taxonomic data (authorship) for a df of taxon names of ferns from Moorea
+#' 
+#' Custom taxonomy added for names not in IPNI or TROPICOS
+#'
+#' @param taxa_names_df Dataframe of taxonomic names of ferns from Moorea
+#'
+#' @return Tibble
+#' 
+lookup_taxonomy <- function (taxa_names_df) {
+  
+  # Use the Global Name Resolver (https://resolver.globalnames.org/) to match the names
+  # to either IPNI or TROPICOS databases.
[…]

### 2020-02-07 code/plan.R

@@ -19,7 +19,7 @@ plan <- drake_plan(
   
   ### Trait data ###
   # - Read in list of accepted species to use for this study
-  # i.e., ferns of Moorea (130 spp)
+  # i.e., ferns of Moorea (129 spp)
   species_list = read_csv(file_in("data/nitta_2017/species.csv")) %>%
     clean_names() %>%
     filter(include == 1, tahiti_only == 0) %>%
@@ -131,11 +131,20 @@ plan <- drake_plan(
   ### PPGI taxonomy ###
   ppgi = read_csv(file_in("data/ppgi_taxonomy.csv")),
   
-  ### Format trait data for SI ###
-  # Include SD and n for all quantitative (continuous) data
+  # Format data for SI ----
+  # - Traits: include SD and n for all quantitative (continuous) data
   trait_data_for_si = process_trait_data_for_si(
     sla_raw, morph_cont_raw, morph_qual_raw
   ),
+  
+  # - Species names
+  # Look up scientific names (specis names with taxonomic author) using
+  # a live search of taxonomic databases via the Global Name Resolver 
+  # (https://resolver.globalnames.org/). Since results have the possibility
+  # of changing, the results are checked-in to the repo at the time of manuscript 
+  # submission.
+  taxonomic_data_for_si = reformat_species_names(species_list) %>%
+    lookup_taxonomy,
   
   # Climate analysis ----
   

### 2020-02-07 code/resolve_names.R

@@ -1,87 +1,0 @@
-# This script looks up scientific names (specis names with taxonomic author) using
-# a live search of taxonomic databases via the Global Name Resolver 
-# (https://resolver.globalnames.org/). Since results have the possibility
-# of changing, it is run once and the results
-# checked-in to the repo at the time of manuscript submission.
-# The script is left here though for reproducibility.
-
-library(taxize)
-library(janitor)
-library(tidyverse)
-
-# Read in list of accepted species names used in this study,
-# extract the genus and specific epithet
-species_list <- read_csv("data/nitta_2017/species.csv") %>%
-  clean_names() %>%
-  filter(include == 1, tahiti_only == 0) %>%
-  pull(genus_sp)
-
-# Reformat the genus_species list of names into a dataframe
-# with the taxon_code used to refer to that taxon in scripts,
-# and other names at various taxonomic ranks.
-taxa_names <-
-  tibble(taxon_code = species_list) %>%
-  separate(taxon_code, c("genus", "specific_epithet"), by = "_", remove = FALSE) %>%
-  mutate(informal_variety = str_match(specific_epithet, "[0-9]") %>% map_chr(1)) %>%
-  mutate(specific_epithet = str_remove_all(specific_epithet, "[0-9]")) %>%
-  mutate(species = paste(genus, specific_epithet)) %>%
-  mutate(informal_variety = case_when(
-    specific_epithet == "sp" ~ NA_character_,
-    TRUE ~ informal_variety
-  )) %>%
-  mutate(infaspecific_epithet = case_when(
-    species == "Lindsaea repens" ~ "marquesensis",
-    species == "Davallia solida" ~ "solida",
-    species == "Ctenitis sciaphila" ~ "sciaphila",
-    species == "Dicranopteris linearis" ~ "linearis",
-    species == "Deparia petersenii" ~ "congrua",
-    TRUE ~ NA_character_
-  )) %>%
-  mutate(taxon = jntools::paste3(genus, specific_epithet, infaspecific_epithet))
-
-# Use the Global Name Resolver (https://resolver.globalnames.org/) to match the names
-# to either IPNI or TROPICOS databases.
-taxa_names_with_authors <- gnr_resolve(
[…]

### 2020-02-07 ms/SI.rmd
Hard-code SI refs, swap order of Fig S1 and S2

@@
 title: ""
-bibliography: 
-  - "references_si.bib"
-  - "references_other.yaml"
-fontsize: 12pt
+mainfont: Roboto-Regular.ttf
+fontsize: 11pt
+mainfontoptions: 
+- BoldFont=Roboto-Bold.ttf
+- ItalicFont=Roboto-Italic.ttf
+- BoldItalicFont=Roboto-BoldItalic.ttf
 output: 
   bookdown::pdf_document2:
-    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
+    citation_package: none
     toc: no
@@
   - \makeatletter\renewcommand*{\fps@table}{H}\makeatother # Fix position of tables
-nocite: | 
-  @NMNS2008, @Copeland1932, @FloraNSW2019, @Brownsey1987, @Nayar1971, @Lloyd1980, @Zhang2008, @Bierhorst1967, @Martin2006, @Tigerschiold1989, @Tigerschiold1990, @Atkinson1975, @Palmer2003
+editor_options: 
+  chunk_output_type: console
 ---
@@
 loadd(list = c(
+  "taxonomic_data_for_si",
   "trait_data_for_si",
@@
 
-fig_s1 <- mpd_corr_plot + mntd_corr_plot + plot_layout(nrow = 2, ncol = 1)
+fig_s2 <- mpd_corr_plot + mntd_corr_plot + plot_layout(nrow = 2, ncol = 1)
 
-fig_s1 <- fig_s1 + plot_annotation(theme = theme(plot.margin = margin(
+fig_s2 <- fig_s2 + plot_annotation(theme = theme(plot.margin = margin(
       0.2, 0.2, 0.2, 0.2, unit = "in")))
 
-fig_s1
-
-# ggsave(plot = fig_s1, filename = here::here("results/Fig_S1.eps"), height = 8, width = 4, units = "in")
+fig_s2
 ```
 
-**`r s_figure("corr-plot")`** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.
+**`r s_figure("corr-plot")`** `r s_figure_cap("corr-plot")`.
 **(a)**, standard effect size of mean phylogenetic distance (MPD~func~) **(b)**, standard effect size of mean nearest taxon distance (MNTD~func~).
 For details on calculations of MPD~func~ and MNTD~func~, see Methods.
-Correlations (values in each tile) calculated with Pearson's correlation coefficient using the R package 'corrr'.
+Correlations between functional diversity as measured with different sets of traits (values in each tile) calculated with Pearson's correlation coefficient using the R package 'corrr'.
 All trait sets were scaled to standard deviation of each trait.
@@
 
-```{r Fig-S2, fig.height = 6, fig.width = 7, fig.cap = ""}
-loadd(cwsd_scatterplot, cache = canopy_cache)
+**References**
 
-cwsd_scatterplot
+<!-- There is a bug somewhere in Rmd -> tex -> pdf using pandoc such that the 
+first bold words until a period in each reference do not appear bold. As a 
+work-around, first render to pdf with `bibliography` and `nocite` in the yaml, 
+then convert Rmd -> tex -> docx with pandoc, edit docx manually to fix formatting, 
+then docx -> md with pandoc, and paste properly formatted refs below -->
 
-# ggsave(plot = cwsd_scatterplot, filename = here::here("results/Fig_S2.eps"), height = 6, width = 7, units = "in")
-```
+**Atkinson L**. **1975**. The gametophyte of five Old World thelypteroid
+ferns. *Phytomorphology* **25**: 38--54.
 
-**`r s_figure("cwsd")`** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.
-Epiphytic communities in green, terrestrial communities in brown.
-Points randomly jittered along x-axis to reduce overlap.
-Boxplots show median values (bold lines) by growth habit.
-Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range.
-Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01; \* = \pval{ }< 0.05 **(a)**, traits related to plant size **(b)**, traits not related to plant size.
+**Bierhorst DW**. **1967**. The gametophyte of *Schizaea dichotoma*.
+*American Journal of Botany* **54**: 538--549.
 
-\clearpage
+**Brownsey PJ**. **1987**. A review of the fern genus *Hypolepis*
+(Dennstaedtiaceae) in the Malesian and Pacific regions. *Blumea* **32**:
+227--276.
 
-**References**
+**Copeland EB**. **1932**. Pteridophytes of the Society Islands.
+*Bernice P. Bishop Museum Bulletin* **93**: 1--86.
+
+**Lloyd RM**. **1980**. Reproductive biology and gametophyte morphology
+of New World populations of *Acrostichum aureum*. *American Fern
+Journal* **70**: 99--110.
+
+**Martin K, Sini S, Zhang C, Slater A, Madhusoodanan P**. **2006**.
+Efficient induction of apospory and apogamy in vitro in silver fern
+(*Pityrogramma calomelanos* L.). *Plant Cell Reports* **25**:
+1300--1307.
+
+**Murdock AG, Smith AR**. **2003**. Pteridophytes of Moorea, French
+Polynesia, with a new species, *Tmesipteris gracilis* (Psilotaceae).
+*Pacific Science* **57**: 253--265.
+
+**National Museum of Nature and Science**. **2008**. *Illustrated Flora
+of Ferns and Fern Allies of South Pacific Islands* (T Nakamura and S
+Matsumoto, Eds.). Tokai University Press.
+
+**Nayar BK, Kaur S**. **1971**. Gametophytes of homosporous ferns. *The
+Botanical Review* **37**: 295--396.
+
+**Pagel M**. **1994**. Detecting correlated evolution on phylogenies: A
+general method for the comparative analysis of discrete characters.
+*Proceedings of the Royal Society B: Biological Sciences* **255**:
+37--45.
+
+**Palmer DD**. **2003**. *Hawaii's Ferns and Fern Allies*. Honolulu:
+University of Hawaii Press.
+
+**Royal Botanic Gardens and Domain Trust**. **2004**. *Flora of New
+South Wales. PlantNET (The NSW Plant Information Network System)* [WWW
+document] URL <http://plantnet.rbgsyd.nsw.gov.au/floraonline.htm>.
+[accessed 1 January 2020].
+
+**Tigerschiöld E**. **1989**. Dehiscence of antheridia in thelypteroid
+ferns. *Nordic Journal of Botany* **9**: 407--412.
+
+**Tigerschiöld E**. **1990**. Gametophytes of some Ceylonese species of
+Thelypteridaceae. *Nordic Journal of Botany* **9**: 657--664.
+
+**Zhang K, Shi L, Zhang X, Jiang C, Tim-Chun W**. **2008**. Gametophyte
+morphology and development of six chinese species of *Pteris*
+(Pteridaceae). *American Fern Journal* **98**: 33--41.

### 2020-02-07 ms/manuscript.Rmd
Load captions separately

@@
 
-# Set up captions
-
-# - Figures
-fig_nums <- captioner(prefix = "Fig.")
-fig_nums(name = "traits-used-image", caption = "Examples of traits used in this study.")
-fig_nums(name = "climate", caption = "Microclimatic variables.")
-fig_nums(name = "pca", caption = "Trait PCA.")
-fig_nums(name = "tree", caption = "Traits on tree.")
-fig_nums(name = "heatmap", caption = "Variable importance scores.")
-fig_nums(name = "cwm-div", caption = "Community-weighted means.")
-fig_nums(name = "comm-div", caption = "Community diversity metrics.")
-
-# - Tables
-table_nums <- captioner(prefix = "Table")
-table_nums(name = "traits-used", caption = "Traits used in this study")
-table_nums(name = "phylosig-cont", caption = "Phylogenetic signal in continuous traits")
-table_nums(name = "phylosig-binary", caption = "Phylogenetic signal in binary traits")
-table_nums(name = "pics", caption = "Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth")
-table_nums(name = "corr-evo", caption = "Pagel's (1994) test of correlated evolution between binary traits")
-
-# - SI figures
-s_fig_nums <- captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
-s_fig_nums(name = "corr-plot", "Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.")
-s_fig_nums(name = "cwsd", "Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.")
-
-# - SI tables
-s_table_nums <- captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")
-s_table_nums(name = "species-list", "List of ferns from Moorea, French Polynesia included in this study")
-s_table_nums(name = "trait-values", "Trait values of ferns on Moorea, French Polynesia")
-s_table_nums(name = "ancova", "Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate")
-s_table_nums(name = "phylosig-binary-strict", "Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy")
-s_table_nums(name = "corr-evo-strict", "Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy")
-s_table_nums(name = "model-fit", "Best-fitting models explaining community diversity metrics by climate and growth habit")
-
-# Make short versions of citation functions
-figure <- pryr::partial(fig_nums, display = "cite")
-table <- pryr::partial(table_nums, display = "cite")
-s_figure <- pryr::partial(s_fig_nums, display = "cite")
-s_table <- pryr::partial(s_table_nums, display = "cite")
+# Load captions
+source(here::here("ms/captions.R"))
 ```

### 2020-02-07 ms/manuscript.Rmd
Use Roboto font

@@
   - "references_other.yaml"
-fontsize: 12pt
+mainfont: Roboto-Regular.ttf
+fontsize: 11pt
+mainfontoptions: 
+- BoldFont=Roboto-Bold.ttf
+- ItalicFont=Roboto-Italic.ttf
+- BoldItalicFont=Roboto-BoldItalic.ttf
 output: 

### 2020-02-07 ms/manuscript.Rmd
Update MS as per editor comments 2020-02-05

@@
 
+Brief heading: "Community trait assessments reveal substantial functional diversity among fern epiphytes"
+
+Social media    
+Twitter: \@joel_nitta  
+Website: \textnormal{https://joelnitta.com}
+
+Orcid IDs   
+Joel H. Nitta: \textnormal{https://orcid.org/0000-0003-4719-7472}  
+Charles C. Davis: \textnormal{https://orcid.org/0000-0001-8747-1101}
+
 <!-- FIXME: add borders to table in docx -->
@@
 
-Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife.
-Data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED [@Nitta2020].
+Code to replicate all analyses, figures, and this manuscript is available at https://github.com/joelnitta/canopylife.
+Data are openly available from the Dryad repository at https://doi.org/10.5061/dryad.fqz612jps [@Nitta2020].
 
@@
 
+<!-- FIXME: add DOI for Dryad Nitta et al. 2020 dataset: https://doi.org/10.5061/dryad.fqz612jps -->
+
 <!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
@@
 
-**`r s_table("species-list")`** List of ferns from Moorea, French Polynesia included in this study
-
-**`r s_table("trait-values")`** Trait values of ferns on Moorea, French Polynesia
-
-**`r s_table("ancova")`** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate
-
-**`r s_table("phylosig-binary-strict")`** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
-
-**`r s_table("corr-evo-strict")`** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy
-
-**`r s_table("model-fit")`** Best-fitting models explaining community diversity metrics by climate and growth habit
-
-**`r s_figure("corr-plot")`** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input
-
-**`r s_figure("cwsd")`** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia
-
-\clearpage
-
-**`r figure("traits-used-image")`** Examples of traits used in this study.
+**`r s_table("species-list")`** `r s_table_cap("species-list")`
+
+**`r s_table("trait-values")`** `r s_table_cap("trait-values")`
+
+**`r s_table("ancova")`** `r s_table_cap("ancova")`
+
+**`r s_table("phylosig-binary-strict")`** `r s_table_cap("phylosig-binary-strict")`
+
+**`r s_table("corr-evo-strict")`** `r s_table_cap("corr-evo-strict")`
+
+**`r s_table("model-fit")`** `r s_table_cap("model-fit")`
+
+**`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`
+
+**`r s_figure("corr-plot")`** `r s_figure_cap("corr-plot")`
+
+\clearpage
+
+**`r figure("traits-used-image")`** Examples of fern traits used in this study.
 **(a)** Sporophyte traits.
@@
 Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval{ }< 0.05.
+Asterisks indicate statistical significance of linear model; \*\*\* = \pval{ }< 0.001; \* = \pval{ }< 0.05.
 Partially transparent colors indicate outliers excluded from the models (see Methods).
@@
 Major epiphytic radiations identified by @Schuettpelz2009 labeled on tree in green: H, Hymenophylloideae; T, Trichomanoideae; V, Vittarioideae; A, *Asplenium*; E, *Elaphoglossum*; P, Polypodiaceae; taxonomy follows @PteridophytePhylogenyGroupI2016.
+Scale in millions of years (my).
 
@@
 Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at \pval{ }< 0.05.
-\tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01; \* = \pval{ }< 0.05.
+\tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01.
 Results for frond width were very similar to frond length and not shown.
+Response variable abbreviations: SLA, specific leaf area.
 
@@
 Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range.
-Asterisks indicate statistical significance of two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01; \* = \pval{ }< 0.05.
-
-\clearpage
-
-**Table `r table_nums("traits-used", display = "n")`** Traits used in this study
+Asterisks indicate statistical significance of two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01.
+
+\clearpage
+
+**Table `r table_nums("traits-used", display = "n")`** Fern traits used in this study
 
@@
 
-**Table `r table_nums("phylosig-cont", display = "n")`** Phylogenetic signal in continuous traits
+**Table `r table_nums("phylosig-cont", display = "n")`** Phylogenetic signal in continuous traits of ferns from Moorea, French Polynesia
 
@@
 
-**Table `r table_nums("phylosig-binary", display = "n")`** Phylogenetic signal in binary traits
+**Table `r table_nums("phylosig-binary", display = "n")`** Phylogenetic signal in binary traits of ferns from Moorea, French Polynesia
 
@@
 
-**Table `r table_nums("pics", display = "n")`** Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth
+**Table `r table_nums("pics", display = "n")`** Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth in ferns from Moorea, French Polynesia
 
@@
 
-**Table `r table_nums("corr-evo", display = "n")`** Pagel's (1994) test of correlated evolution between binary traits
+**Table `r table_nums("corr-evo", display = "n")`** Pagel's (1994) test of correlated evolution between binary traits in ferns from Moorea, French Polynesia
 

### 2020-02-07 make.R
Make plan in local environment

@@ -15,8 +15,13 @@ source("code/setup.R")
 # Update drake settings.
 pkgconfig::set_config("drake::strings_in_dots" = "literals")
 
+# Specify non-global environment
+# to get around taxize and captioner modifying global env
+# (cf https://github.com/ropensci/drake/issues/749)
+envir <- new.env(parent = globalenv())
+
 # Load functions and plans.
-source("code/functions.R")
+source("code/functions.R", local = envir)
 source("code/plan.R")
 
 # Load cache
@@ -26,4 +31,4 @@ canopy_cache <- new_cache("canopy_cache")
 set.seed(1954)
 
 # Make plan.
-make(plan, cache = canopy_cache)
+make(plan, cache = canopy_cache, envir = envir)

### 2020-02-07 ms/captions.R
Don't use any text needing special formatting in caption titles

@@ -27,7 +27,7 @@ s_fig_nums(name = "corr-plot", "Correlation plot comparing functional diversity 
 s_table_nums <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")
 s_table_nums(name = "species-list", "List of ferns from Moorea, French Polynesia included in this study")
 s_table_nums(name = "trait-values", "Trait values of ferns on Moorea, French Polynesia")
-s_table_nums(name = "ancova", "Analysis of covariance (ANCOVA) for differences between climatic variables measured with dataloggers placed on the ground (terrestrial growth habit) vs. mounted at c. 2 m on trees (epiphytic growth habit) on Moorea, French Polynesia, with elevation as a covariate")
+s_table_nums(name = "ancova", "Analysis of covariance (ANCOVA) for differences between climatic variables measured with dataloggers placed on the ground (terrestrial growth habit) vs. mounted at 2 m on trees (epiphytic growth habit) on Moorea, French Polynesia, with elevation as a covariate")
 s_table_nums(name = "phylosig-binary-strict", "Phylogenetic signal in binary traits of ferns on Moorea, French Polynesia, strict dataset excluding any species whose traits were scored based on taxonomy")
 s_table_nums(name = "corr-evo-strict", "Pagel's (1994) test of correlated evolution between binary traits of ferns on Moorea, French Polynesia, strict dataset excluding any species whose traits were scored based on taxonomy")
 s_table_nums(name = "model-fit", "Best-fitting models explaining community diversity metrics of ferns on Moorea, French Polynesia by climate and growth habit")

### 2020-02-07 ms/SI.rmd
Output SI to pdf, no docx

@@
     fig_caption: yes
-    keep_tex: yes
+    keep_tex: no
     latex_engine: xelatex

### 2020-02-07 code/plan.R

@@ -535,11 +535,10 @@ plan <- drake_plan(
   ),
   
   # SI ----
-  si_pdf = render_tracked(
+  si_pdf = rmarkdown::render(
     knitr_in("ms/SI.Rmd"),
     quiet = TRUE,
-    output_dir = here::here("results"),
-    tracked_output = file_out(here::here("results/SI.tex"))
+    output_dir = here::here("results")
   )
   
 )

### 2020-02-07 code/functions.R
Fix year in Flora of NSW ref

@@ -410,7 +410,7 @@ process_trait_data_for_si <- function (sla_raw, morph_cont_raw, morph_qual_raw) 
         source == "Ferns & Fern-Allies of South Pacific Islands" ~ "National Museum of Nature and Science (2008)", 
         source == "Pteridophytes of the Society Islands" ~ "Copeland (1932)",
         source == "Hawaii's Ferns and Fern Allies" ~ "Palmer (2003)",
-        source == "Flora of New South Wales" ~ "Royal Botanic Gardens and Domain Trust (2019)",
+        source == "Flora of New South Wales" ~ "Royal Botanic Gardens and Domain Trust (2004)",
         source == "Brownsey 1987" ~ "Brownsey (1987)",
         source == "measurement" ~ "M"
       )

### 2020-02-07 ms/SI.rmd
Update SI
- add units for SI table 3 legend
- fix definition of size traits in fig 6 legend

@@
 d.f., degrees of freedom; MS, mean squares.
+Response variables include mean, minimum ('min'), maximum ('max'), and standard deviation ('sd') of temperature ('temp', °C) and vapor pressure deficit ('vpd', kPa).
+'NA' indicates not applicable.
 
@@
 
-footer <- "d.f., degrees of freedom; MS, mean squares." %>% wrap_quotes
+footer <- "d.f., degrees of freedom; MS, mean squares.
+Response variables include mean, minimum ('min'), maximum ('max'), and standard deviation ('sd') of temperature ('temp', °C) and vapor pressure deficit ('vpd', kPa).
+'NA' indicates not applicable." %>% wrap_quotes
 
@@
 All trait sets were scaled to standard deviation of each trait.
-The following trait sets were used: 'all' (all traits), 'size' (sporophyte traits related to size; frond length and width, stipe length, rhizome width, pinna number, SLA, degree of lamina dissection), 'other' (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), 'gameto' (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), 'reduced' (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter).
+The following trait sets were used: 'all' (all traits), 'size' (sporophyte traits related to size; frond length and width, stipe length, rhizome width), 'other' (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), 'gameto' (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), 'reduced' (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter).
 For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), both log-transformed and untransformed sets were included.

### 2020-02-12 reports/raw_data_viz.Rmd
Add report visualizing raw data

---
title: "Data plots"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load packages and custom functions
source(here::here("code/setup.R"))
source(here::here("code/functions.R"))
```

Visualize the data and check for any outliers.

## Climate

Load raw data.

```{r load-climate-data}
# Read in raw climate data
climate_raw <- read_csv(here::here("data/hobo_moorea_aorai_2-24-15.csv"))

# Subset to Moorea, format site as 
climate_raw_moorea <- climate_raw %>%
   filter(str_detect(site, "Aorai", negate = TRUE)) %>%
   mutate(date_time = (lubridate::ymd(date) + lubridate::hms(time)) %>% as.POSIXct())
```

Identify time periods with erroneous values to exclude from cleaned dataset. These are the values that will be excluded by `process_raw_climate()`.

Note the band of red from 2013-11-19 to 2014-03-21 is actually OK for most dataloggers, but not for Tohiea 1170m ter + Tohiea 2000m ter, so it is flagged to be excluded from all sites.

```{r format-raw-climate}
# Make dataframe of dates/sites to flag as having weird data.
highlight_chunks <- 
  climate_raw_moorea %>%
  select(date, site) %>%
  unique %>%
  mutate(flag = case_when(
    date >= "2014-03-12" & date < "2014-07-06" & site == "Tohiea_400m_ter" ~ TRUE,
    site == "Rotui_600m_epi" ~ TRUE,
    site == "Mouaputa_800m_epi" ~ TRUE,
    date >= "2013-11-19" & date <= "2014-03-21" ~ TRUE,
    date >= "2013-07-26" & date <= "2013-07-29" & site == "Tohiea_600m_epi"~ TRUE,
    date == "2013-08-08" & site == "Tohiea_600m_ter" ~ TRUE,
    date %in% c("2014-06-15", "2014-06-16", "2014-06-19", "2014-06-20") & site == "Tohiea_600m_ter" ~ TRUE,
    TRUE ~ FALSE
  ))

# Add the "flag" column back into the original data
climate_raw_moorea_highlight <-
  left_join(climate_raw_moorea, highlight_chunks, by = c("site", "date")) %>%
  # Reformat growth habit as factor, separate from site name
   mutate(
      habit = case_when(
        str_detect(site, "epi") ~ "epiphytic",
        str_detect(site,  "ter") ~ "terrestrial"
      ),
      habit = fct_relevel(habit, c("epiphytic", "terrestrial")),
      site = str_remove_all(site, "_epi|_ter")
    )
```

Plot the raw data, highlighting the data that will be excluded from the cleaned dataset in red.

```{r raw-temp-plot, fig.height = 24, fig.width = 8}
ggplot(climate_raw_moorea_highlight, aes(x = date_time, y = temp, color = flag)) +
  geom_line() +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  labs(
    title = "Moorea temperature",
    subtitle = "raw data") +
  theme(legend.position = "bottom") +
  facet_grid(vars(site), vars(habit))
```

```{r raw-rh-plot, fig.height = 24, fig.width = 8}
ggplot(climate_raw_moorea_highlight, aes(x = date_time, y = rh, color = flag)) +
  geom_line() +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  labs(
    title = "Moorea rel. humidity",
    subtitle = "raw data") +
  theme(legend.position = "bottom") +
  facet_grid(vars(site), vars(habit))
```

Process raw data, excluding time periods with erroneous values.

```{r process-climate}
# Process raw climate data: 
# subset to only Moorea and days without missing data,
# replace one chunk of missing data with same dates from previous year,
# calculate vapor pressure deficit (VPD) from temp and RH.
moorea_climate <- process_raw_climate(climate_raw) %>%
  # Add date_time column for plotting
   mutate(date_time = (lubridate::ymd(date) + lubridate::hms(time)) %>% as.POSIXct())
```

Plot the cleaned climate data.

```{r clean-temp-plot, fig.height = 24, fig.width = 8}
ggplot(moorea_climate, aes(x = date_time, y = temp)) +
  geom_line() +
  labs(
    title = "Moorea temperature",
    subtitle = "cleaned data") +
  facet_grid(vars(site), vars(habit))
```

```{r clean-rh-plot, fig.height = 24, fig.width = 8}
ggplot(moorea_climate, aes(x = date_time, y = rh)) +
  geom_line() +
  labs(
    title = "Moorea rel. humidity",
    subtitle = "cleaned data") +
  facet_grid(vars(site), vars(habit))
```

### 2020-02-13 reports/raw_data_viz.Rmd
Update raw data viz report

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
 # Load packages and custom functions
@@
 
-# Subset to Moorea, format site as 
-climate_raw_moorea <- climate_raw %>%
+# Subset to Moorea
+climate <- climate_raw %>%
     filter(str_detect(site, "Aorai", negate = TRUE)) %>%
-   mutate(date_time = (lubridate::ymd(date) + lubridate::hms(time)) %>% as.POSIXct())
+    mutate(
+      day = lubridate::day(date) %>% as.numeric,
+      month = lubridate::month(date) %>% as.numeric,
+      year = lubridate::year(date) %>% as.numeric,
+      # add date_time col for plotting
+      date_time = (lubridate::ymd(date) + lubridate::hms(time)) %>% as.POSIXct())
 ```
 
-Identify time periods with erroneous values to exclude from cleaned dataset. These are the values that will be excluded by `process_raw_climate()`.
+Identify time periods with erroneous values to exclude from cleaned dataset. Do this with code first below and inspect the results.
 
-Note the band of red from 2013-11-19 to 2014-03-21 is actually OK for most dataloggers, but not for Tohiea 1170m ter + Tohiea 2000m ter, so it is flagged to be excluded from all sites.
+```{r flag-extremes}
+# Identify start and end dates of survey including all sites on Moorea
+survey_window <-
+climate %>%
+  group_by(site) %>%
+  summarize(
+    last_measured_day = max(date),
+    first_measured_day = min(date)
+  ) %>%
+  summarize(
+    end_date = min(last_measured_day),
+    start_date = max(first_measured_day)
+  ) %>%
+  mutate_all(as.POSIXct)
+
+# Tally days with extreme (possibly erroneous) data
+extremes_tally <-
+climate %>%
+  # Exclude two sites completely that are mostly missing data
+  filter(site != "Rotui_600m_epi") %>%
+  filter(site != "Mouaputa_800m_epi") %>%
+  # Crop start and end dates
+  filter(date > as.Date(survey_window$start_date) & date < as.Date(survey_window$end_date)) %>%
+  # Flag any days as possible bad data if values are extremely low or high
+  mutate(
+    temp_hi = ifelse(temp > 50, TRUE, FALSE),
+    temp_low = ifelse(temp < 5, TRUE, FALSE),
+    rh_low = ifelse(rh < 10, TRUE, FALSE)) %>%
+  group_by(site, date) %>%
+  summarize(
+    temp_hi = sum(temp_hi),
+    temp_low = sum(temp_low),
+    rh_low = sum(rh_low)
+  ) %>%
+  ungroup %>%
+  filter_at(vars(temp_hi, temp_low, rh_low), any_vars(. > 0)) %>%
+  mutate(flag = TRUE) %>%
+  arrange(site, date)
+
+extremes_tally %>% DT::datatable()
+```
+
+After inspecting raw data plots and the points identified above, flag days for removal by hard-coding.
 
@@
 # Make dataframe of dates/sites to flag as having weird data.
-highlight_chunks <- 
-  climate_raw_moorea %>%
+climate_data_flags <- 
+  climate %>%
   select(date, site) %>%
   unique %>%
-  mutate(flag = case_when(
-    date >= "2014-03-12" & date < "2014-07-06" & site == "Tohiea_400m_ter" ~ TRUE,
-    site == "Rotui_600m_epi" ~ TRUE,
-    site == "Mouaputa_800m_epi" ~ TRUE,
-    date >= "2013-11-19" & date <= "2014-03-21" ~ TRUE,
-    date >= "2013-07-26" & date <= "2013-07-29" & site == "Tohiea_600m_epi"~ TRUE,
+  mutate(flag_rh = case_when(
+    # Flag 2013-11-19 to 2013-12-15 (Tohiea 1170m ter 0% RH)
+    date >= "2013-11-19" & date <= "2013-12-15" & site == "Tohiea_1170m_ter" ~ TRUE,
+    # Flag 2013-12-29 to 2014-03-12 (Tohiea 200m ter unusually low RH)
+    date >= "2013-12-29" & date <= "2014-03-12" & site == "Tohiea_200m_ter" ~ TRUE,
+    # Flag 2014-06-15 to 2014-06-20 (Tohiea_800m_ter 0% RH)
+    date >= "2014-06-15" & date <= "2014-06-20" & site == "Tohiea_800m_ter" ~ TRUE,
+    # Flag 2013-08-08 (Tohiea_600m_ter 1% RH)
     date == "2013-08-08" & site == "Tohiea_600m_ter" ~ TRUE,
-    date %in% c("2014-06-15", "2014-06-16", "2014-06-19", "2014-06-20") & site == "Tohiea_600m_ter" ~ TRUE,
+    # Flag Tohiea sites missing data
+    date >= "2014-03-13" & date <= "2014-03-20" & str_detect(site, "Tohiea") ~ TRUE,
+    TRUE ~ FALSE
+  )) %>%
+  mutate(flag_temp = case_when(
+    # Delete 2013-07-26 to 2013-07-29 (Tohiea_600m epi temperature below 0C)
+    date >= "2013-07-26" & date <= "2013-07-29" & site == "Tohiea_600m_epi" ~ TRUE,
+    # Flag Tohiea sites missing data
+    date >= "2014-03-13" & date <= "2014-03-20" & str_detect(site, "Tohiea") ~ TRUE,
     TRUE ~ FALSE
@@
 # Add the "flag" column back into the original data
-climate_raw_moorea_highlight <-
-  left_join(climate_raw_moorea, highlight_chunks, by = c("site", "date")) %>%
-  # Reformat growth habit as factor, separate from site name
+climate_flagged <- left_join(climate, climate_data_flags, by = c("site", "date"))
+```
+
+Plot the raw data, highlighting the data that will be excluded from the cleaned dataset in red, and lines indicating cropped starting and ending dates in blue.
+
+```{r raw-temp-plot, fig.height = 24, fig.width = 8}
+ggplot(climate_flagged, aes(x = date_time, y = temp, color = flag_temp)) +
+  geom_line() +
+  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
+  annotate("segment", 
+           x = survey_window$start_date,
+           xend = survey_window$start_date, 
+           y = -Inf, yend = Inf, color = "blue") +
+  annotate("segment", 
+           x = survey_window$end_date, 
+           xend = survey_window$end_date, 
+           y = -Inf, yend = Inf, color = "blue") +
+  labs(
+    title = "Moorea temperature",
+    subtitle = "raw data") +
+  theme(legend.position = "bottom") +
+  facet_wrap(vars(site), ncol = 2)
+```
+
+```{r raw-rh-plot, fig.height = 24, fig.width = 8}
+ggplot(climate_flagged, aes(x = date_time, y = rh, color = flag_rh)) +
+  geom_line() +
+  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
+  annotate("segment", 
+           x = survey_window$start_date,
+           xend = survey_window$start_date, 
+           y = -Inf, yend = Inf, color = "blue") +
+  annotate("segment", 
+           x = survey_window$end_date, 
+           xend = survey_window$end_date, 
+           y = -Inf, yend = Inf, color = "blue") +
+  labs(
+    title = "Moorea rel. humidity",
+    subtitle = "raw data") +
+  theme(legend.position = "bottom") +
+  facet_wrap(vars(site), ncol = 2)
+```
+
+Process raw data, excluding time periods with erroneous values. Any day with erroneous values for any site will be excluded from ALL sites.
+
+```{r process-climate}
+# Process climate using same code as function in analysis:
+
+# Subset to Moorea
+climate_moorea <- climate_raw %>%
+    filter(str_detect(site, "Aorai", negate = TRUE)) %>%
+    mutate(
+      day = lubridate::day(date) %>% as.numeric,
+      month = lubridate::month(date) %>% as.numeric,
+      year = lubridate::year(date) %>% as.numeric)
+
+# Split out set of days from 2013 and 2014 to use for filling in 
+# missing 2014 Tohiea 400 m ter data
+tohiea_400m_2013 <- climate_moorea %>%
+  filter(date >= "2013-03-12" & date < "2013-07-06" & site == "Tohiea_400m_ter")
+
+tohiea_400m_2014 <- climate_moorea %>%
+  filter(date >= "2014-03-12" & date < "2014-07-06" & site == "Tohiea_400m_ter")
+
+# Replace the year 2013 with 2014 only for these dates in the Tohiea 400 m ter data
+tohiea_400m_as_2014 <-
+  tohiea_400m_2013 %>%
+  mutate(year = year + 1) %>%
+  mutate(date = paste(year, month, day, sep = "-") %>% as.Date)
+
+# Process data
+climate_filtered <- climate_moorea %>%
+  # Remove bad Tohiea data from 2014
+  anti_join(tohiea_400m_2014, by = c("date", "site")) %>%
+  # Replace bad Tohiea data from 2014 with data from 2013
+  bind_rows(tohiea_400m_as_2014) %>%
+  # Remove two sites that have are (nearly) complete failures
+  filter(site != "Rotui_600m_epi") %>%
+  filter(site != "Mouaputa_800m_epi") %>%
+  # Crop to dates when all dataloggers were set up on Moorea
+  filter(date >= "2013-07-07" & date <= "2014-07-05") %>%
+  # Flag dates to remove:
+  mutate(
+    flag = case_when(
+      # 2013-11-19 to 2013-12-15 (Tohiea 1170m ter 0% RH)
+      date >= "2013-11-19" & date <= "2013-12-15" ~ TRUE,
+      # 2013-12-29 to 2014-03-12 (Tohiea 200m ter unusually low RH)
+      date >= "2013-12-29" & date <= "2014-03-12" ~ TRUE,
+      # 2014-03-13 to 2014-03-20 (Missing data for most Tohiea sites)
+      date >= "2014-03-13" & date <= "2014-03-20" ~ TRUE,
+      # 2014-06-15 to 2014-06-20 (Tohiea_800m_ter 0% RH)
+      date >= "2014-06-15" & date <= "2014-06-20" ~ TRUE,
+      # 2013-07-26 to 2013-07-29 (Tohiea_600m epi temperature below 0C)
+      date >= "2013-07-26" & date <= "2013-07-29" ~ TRUE,
+      # 2013-08-08 (Tohiea_600m_ter 1% RH)
+      date == "2013-08-08" ~ TRUE,
+      TRUE ~ FALSE
+    )) %>%
+  # Instead of filtering out flagged data, set temp and RH for these to NA
+  # this will make them appear as gaps in the plot
+  mutate(
+    rh = case_when(
+      flag == FALSE ~ rh,
+      flag == TRUE ~ NaN
+    )
+  ) %>%
+  mutate(
+    temp = case_when(
+      flag == FALSE ~ temp,
+      flag == TRUE ~ NaN
+    )
+  ) %>%
+  select(-flag) %>%
+  # Check that there are no duplicate measurements for a given time/site/date
+  assert_rows(col_concat, is_uniq, date, time, site) %>%
+  # Check for temp, RH in reasonable range 
+  # (shouldn't be any freezing, expect at least some humidity)
+  assert(within_bounds(0,50, include.lower = FALSE), temp) %>%
+  assert(within_bounds(1,100), rh) %>%
+  # Reformat filtered data, add VPD
   mutate(
@@
     habit = fct_relevel(habit, c("epiphytic", "terrestrial")),
-      site = str_remove_all(site, "_epi|_ter")
+    site = str_remove_all(site, "_epi|_ter"),
+    vpd = plantecophys::RHtoVPD(rh, temp),
+    # add date_time col for plotting
+    date_time = (lubridate::ymd(date) + lubridate::hms(time)) %>% as.POSIXct()
   )
-```
-
-Plot the raw data, highlighting the data that will be excluded from the cleaned dataset in red.
-
-```{r raw-temp-plot, fig.height = 24, fig.width = 8}
-ggplot(climate_raw_moorea_highlight, aes(x = date_time, y = temp, color = flag)) +
-  geom_line() +
-  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
-  labs(
-    title = "Moorea temperature",
-    subtitle = "raw data") +
-  theme(legend.position = "bottom") +
-  facet_grid(vars(site), vars(habit))
-```
-
-```{r raw-rh-plot, fig.height = 24, fig.width = 8}
-ggplot(climate_raw_moorea_highlight, aes(x = date_time, y = rh, color = flag)) +
-  geom_line() +
-  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
-  labs(
-    title = "Moorea rel. humidity",
-    subtitle = "raw data") +
-  theme(legend.position = "bottom") +
-  facet_grid(vars(site), vars(habit))
-```
-
-Process raw data, excluding time periods with erroneous values.
-
-```{r process-climate}
-# Process raw climate data: 
-# subset to only Moorea and days without missing data,
-# replace one chunk of missing data with same dates from previous year,
-# calculate vapor pressure deficit (VPD) from temp and RH.
-moorea_climate <- process_raw_climate(climate_raw) %>%
-  # Add date_time column for plotting
-   mutate(date_time = (lubridate::ymd(date) + lubridate::hms(time)) %>% as.POSIXct())
 ```
@@
 ```{r clean-temp-plot, fig.height = 24, fig.width = 8}
-ggplot(moorea_climate, aes(x = date_time, y = temp)) +
+ggplot(climate_filtered, aes(x = date_time, y = temp)) +
   geom_line() +
@@
 ```{r clean-rh-plot, fig.height = 24, fig.width = 8}
-ggplot(moorea_climate, aes(x = date_time, y = rh)) +
+ggplot(climate_filtered, aes(x = date_time, y = rh)) +
   geom_line() +
@@
 ```
+
+## Traits
+
+Load raw data and prepare for plotting.
+
+```{r load-trait-data}
+# Read in species list
+species_list <- read_csv(here::here("data/nitta_2017/species.csv")) %>%
+    clean_names() %>%
+    filter(include == 1, tahiti_only == 0) %>%
+    # Exclude Microsorum_xmaximum (hybrid between M. grossum and M. commutatum)
+    filter(genus_sp != "Microsorum_xmaximum") %>%
+    pull(genus_sp)
+
+# Process raw continuous trait data
+# (includes one measure per specimen,
+#  multiple specimens per species).
+# All measurements in cm, except number of pinna pairs.
+  morph_cont_raw <- process_raw_cont_morph(
+    here::here("data/morph_measurements_2020-02-10.csv"),
+    species_list
+  )
+
+# Angiopteris evecta is much bigger than other species, obscures
+# points in plot. Exclude this outlier for plotting.
+morph_cont_raw_no_evecta <-
+  morph_cont_raw %>%
+  filter(!str_detect(species, "evecta"))
+
+# Define a function for plotting trait data
+
+plot_raw_quant_morph <- function(morph_cont_raw, trait_select) {
+
+  # Subset to selected trait
+  trait_select_enquo <- enquo(trait_select)
+
+  morph_cont_sel <- select(morph_cont_raw, species, !!trait_select_enquo)
+
+  # Calculate mean trait value to arrange species box plots by mean
+  means <-
+    morph_cont_sel %>%
+    group_by(species) %>%
+    summarize(
+      mean = mean(!!trait_select_enquo, na.rm = TRUE)
+    )
+
+  morph_cont_sel %>%
+    left_join(means) %>%
+    mutate(species = as.factor(species) %>% fct_reorder(mean)) %>%
+    ggplot(aes(x = species, y = !!trait_select_enquo)) +
+    geom_boxplot() +
+    coord_flip()
+
+}
+```
+
+Plot raw quantitative trait data.
+
+```{r plot-stipe, fig.height = 16, fig.width = 6}
+plot_raw_quant_morph(morph_cont_raw_no_evecta, stipe_length) +
+  labs(title = "Stipe length")
+```
+
+```{r plot-frond, fig.height = 16, fig.width = 6}
+plot_raw_quant_morph(morph_cont_raw_no_evecta, frond_length) +
+  labs(title = "Frond length")
+```
+
+```{r plot-lamina-width, fig.height = 16, fig.width = 6}
+plot_raw_quant_morph(morph_cont_raw_no_evecta, lamina_width) +
+  labs(title = "Lamina width")
+```
+
+```{r plot-rhizome, fig.height = 16, fig.width = 6}
+plot_raw_quant_morph(morph_cont_raw_no_evecta, rhizome_dia) +
+  labs(title = "Rhizome diameter")
+```
+
+```{r plot-pinna-pairs, fig.height = 16, fig.width = 6}
+plot_raw_quant_morph(morph_cont_raw_no_evecta, pinna_pairs) +
+  labs(title = "No. pinna pairs")
+```

### 2020-02-18 ms/SI.rmd
Add new SI table for linear model fits
Rename 'model-fit' SI caption to 'gam-fit'

@@
   "best_fit_div_models_moran",
-  "func_comm_struc_combined"), 
+  "func_comm_struc_combined",
+  "climate_model_summaries",
+  "climate_models",
+  "div_el_models",
+  "div_el_model_summaries",
+  "cwsd_scatterplot"), 
   cache = canopy_cache)
@@
 # Write out table with metadata
+filename <- s_table("species-list") %>%
+  str_replace_all(" ", "_") %>%
+  paste0(".csv")
+
 write_csv_with_meta(
   taxonomic_data_for_si, 
-  here::here("results/Table_S1.csv"), 
+  here::here("results", filename), 
   header,
@@
 # Write out table with metadata
+filename <- s_table("trait-values") %>%
+  str_replace_all(" ", "_") %>%
+  paste0(".csv")
+
 write_csv_with_meta(
   trait_data_for_si, 
-  here::here("results/Table_S2.csv"), 
+  here::here("results", filename), 
   header,
@@
 # Write out table with metadata
+filename <- s_table("ancova") %>%
+  str_replace_all(" ", "_") %>%
+  paste0(".csv")
+
 write_csv_with_meta(
   climate_ancova_results_formatted, 
-  here::here("results/Table_S3.csv"), 
+  here::here("results", filename), 
+  header,
+  footer)
+```
+
+**`r s_table("linear-fit")`** `r s_table_cap("linear-fit")` (see separate file)
+
+MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
+Trait names refer to community-weighted mean values (see Methods).
+For details on calculations of MPD and MNTD, see Methods.
+'NA' indicates not applicable.
+
+```{r linear-fit}
+# Combine summaries of models for diversity metrics
+div_summaries <-
+  div_el_model_summaries %>%
+  left_join(
+    select(div_el_models,
+           var, model_type, delta_AICc, morans_I, I_pval), 
+    by = c("var", "model_type")
+  )
+
+# Combine summaries of models for climate metrics
+climate_summaries <-
+  climate_model_summaries %>%
+  left_join(
+    select(climate_models,
+           var, model_type, delta_AICc, morans_I, I_pval), 
+    by = c("var", "model_type")
+  )
+
+# Combine summaries of diversity and climate metrics, reformat model names
+table <-
+  bind_rows(div_summaries, climate_summaries) %>%
+  mutate(
+    model_type = model_type %>% 
+      str_replace_all("el_only", "Elevation") %>%
+      str_replace_all("both", "Elevation + Growth habit") %>%
+      str_replace_all("habit_only", "Growth habit") %>%
+      str_replace_all("interaction", "Elevation + Growth habit + Elevation x Growth habit")
+    )
+
+# Reformat variable names and put in preferred order by joining
+linear_fit_table <-
+tribble(
+  ~var, ~Response,
+  "temp_max", "Max. temp.",
+  "temp_mean", "Mean temp.",
+  "vpd_mean", "Mean VPD",
+  "vpd_min", "Min. VPD",
+  "temp_sd", "SD temp.",
+  "vpd_sd", "SD VPD",
+  "ntaxa", "Richness",
+  "mpd.obs.z.phy", "MPDphy",
+  "mntd.obs.z.phy", "MNTDphy",
+  "mpd.obs.z.func", "MPDfunc",
+  "mntd.obs.z.func", "MNTDfunc",
+  "dissection", "Dissection",
+  "length", "Frond length",
+  "pinna", "Pinna no.",
+  "rhizome", "Rhizome diam.",
+  "stipe", "Stipe length",
+  "sla", "SLA",
+  "width", "Frond width"
+) %>%
+  left_join(table, by = "var") %>%
+  select(
+    Response,
+    Model = model_type,
+    AICc,
+    f = statistic,
+    `Adjusted R2` = adj.r.squared,
+    P = p.value,
+    `d.f.` = df,
+    delta_AICc,
+    `Moran's I` = morans_I,
+    P_moran = I_pval
+  )
+
+# Set header and footer
+header <- c(
+  si_table_header,
+  s_table_nums("linear-fit") %>% wrap_quotes
+)
+
+footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
+Trait names refer to community-weighted mean values (see Methods).
+For details on calculations of MPD and MNTD, see Methods.
+'NA' indicates not applicable." %>% wrap_quotes
+
+filename <- s_table("linear-fit") %>%
+  str_replace_all(" ", "_") %>%
+  paste0(".csv")
+
+# Write out table with metadata
+write_csv_with_meta(
+  linear_fit_table, 
+  here::here("results", filename), 
   header,
@@
 # Write out table with metadata
+filename <- s_table("phylosig-binary-strict") %>%
+  str_replace_all(" ", "_") %>%
+  paste0(".csv")
+
+# Write out table with metadata
 write_csv_with_meta(
   phylosig_binary_traits_strict_formatted, 
-  here::here("results/Table_S4.csv"), 
+  here::here("results", filename), 
   header,
@@
 # Write out table with metadata
+filename <- s_table("corr-evo-strict") %>%
+  str_replace_all(" ", "_") %>%
+  paste0(".csv")
+
+# Write out table with metadata
 write_csv_with_meta(
   correlated_evo_test_strict_formatted, 
-  here::here("results/Table_S5.csv"), 
+  here::here("results", filename), 
   header,
@@
 
-**`r s_table("model-fit")`** `r s_table_cap("model-fit")` (see separate file)
+**`r s_table("gam-fit")`** `r s_table_cap("gam-fit")` (see separate file)
 
@@
 
-```{r model-fit}
+```{r gam-fit}
 # Format table
@@
   si_table_header,
-  s_table_nums("model-fit") %>% wrap_quotes
+  s_table_nums("gam-fit") %>% wrap_quotes
 )
@@
 # Write out table with metadata
+filename <- s_table("gam-fit") %>%
+  str_replace_all(" ", "_") %>%
+  paste0(".csv")
+
+# Write out table with metadata
 write_csv_with_meta(
   best_fit_div_models_moran_formatted, 
-  here::here("results/Table_S6.csv"), 
+  here::here("results", filename), 
   header,
   footer)
-
 ```
@@
 ```{r Fig-S1, fig.height = 6, fig.width = 7, fig.cap = ""}
-loadd(cwsd_scatterplot, cache = canopy_cache)
-
 cwsd_scatterplot

### 2020-02-18 ms/manuscript.Rmd
Add new SI table for linear model fits
Rename 'model-fit' SI caption to 'gam-fit'

@@
 Epiphytic dataloggers had higher values than terrestrial ones for variables related to VPD.
+Spatial autocorrelation was not detected in any of the climate models (`r s_table("linear-fit")`).
 
@@
 In the full-subsets analysis including all climate variables and diversity metrics, growth habit and temperature emerged as the most important variables (`r figure("heatmap")`).
-Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (`r s_table("model-fit")`).
-Spatial autocorrelation was not detected for any of the diversity metrics (`r s_table("model-fit")`).
+Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (`r s_table("gam-fit")`).
+Spatial autocorrelation was not detected in any of the best-fitting full-subsets models (`r s_table("gam-fit")`).
 
 Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
-Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "sla", "r")`, \pval{ }`r get_mod_result(model_summ_print, "sla", "p")`).
+Spatial autocorrelation was detected in the linear model for SLA (`r s_table("linear-fit")`), so a spatial generalized mixed model was used instead (see Methods). 
+Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities.
 Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "dissection", "p")`).
@@
 No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (`r figure("comm-div")`).
+Spatial autocorrelation was not detected in any of the linear models for the community diversity metrics (`r s_table("linear-fit")`).
 
@@
 
+**`r s_table("linear-fit")`** `r s_table_cap("linear-fit")`
+
 **`r s_table("phylosig-binary-strict")`** `r s_table_cap("phylosig-binary-strict")`
@@
 
-**`r s_table("model-fit")`** `r s_table_cap("model-fit")`
+**`r s_table("gam-fit")`** `r s_table_cap("gam-fit")`
 

### 2020-02-18 ms/captions.R

@@ -28,9 +28,10 @@ s_table_nums <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suf
 s_table_nums(name = "species-list", "List of ferns from Moorea, French Polynesia included in this study")
 s_table_nums(name = "trait-values", "Trait values of ferns on Moorea, French Polynesia")
 s_table_nums(name = "ancova", "Analysis of covariance (ANCOVA) for differences between climatic variables measured with dataloggers placed on the ground (terrestrial growth habit) vs. mounted at 2 m on trees (epiphytic growth habit) on Moorea, French Polynesia, with elevation as a covariate")
+s_table_nums(name = "linear-fit", "Best-fitting linear models explaining climate and community diversity metrics of ferns on Moorea, French Polynesia by elevation and growth habit")
 s_table_nums(name = "phylosig-binary-strict", "Phylogenetic signal in binary traits of ferns on Moorea, French Polynesia, strict dataset excluding any species whose traits were scored based on taxonomy")
 s_table_nums(name = "corr-evo-strict", "Pagel's (1994) test of correlated evolution between binary traits of ferns on Moorea, French Polynesia, strict dataset excluding any species whose traits were scored based on taxonomy")
-s_table_nums(name = "model-fit", "Best-fitting models explaining community diversity metrics of ferns on Moorea, French Polynesia by climate and growth habit")
+s_table_nums(name = "gam-fit", "Best-fitting models explaining community diversity metrics of ferns on Moorea, French Polynesia by climate and growth habit")
 
 # Make short versions of citation functions
 # - Just the number

### 2020-02-18 ms/manuscript.Rmd
Update MS

@@
 We tested for differences between epiphytic and terrestrial dataloggers while accounting for elevation with ANCOVA, using growth habit as the categorical variable and elevation as the covariate.
-We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone and their interactions, then selecting the one with the lowest corrected Akaike Information Criterion [AIC~c~; @Akaike1973].
+We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone, in combination, and in combination with their interactions.
+We then calculated the corrected Akaike Information Criterion [AIC~c~; @Akaike1973] and selected the best model with the combination of the fewest non-significant parameters and lowest AIC~c~.
 A single outlier plot with unusually high temperature and low humidity (Mt. Rotui, 830 m plot) was excluded from ANCOVA and linear models.
@@
 In the first model, rates of evolution of the two traits are independent.
-In the second model, the rate of evolution of one trait depends on the other.
+In the second model, the rate of evolution of each trait depends on the other.
 A significantly better fit of the second model indicates that evolution of the two traits is correlated.
@@
 Our dataset included `r (nrow(div_metrics_all) - nrow(climate_select)) %>% english2()` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
-Therefore, we also constructed linear models for each diversity metric related to elevation including all sites using the same model selection procedure as in the climate analysis.
+Therefore, we also constructed linear models for each diversity metric in response to elevation and growth habit including all sites using the same model selection procedure as in the climate analysis.
 We checked for spatial autocorrelation in the model residuals with Moran's *I* [@Moran1950] as implemented in the \package{spdep} R package [@Bivand2013]. 
+In the case the spatial autocorrelation was detected, a generalized linear mixed model was fit with a Mátern spatial correlation structure using the \package{spAMM} R package [@Rousset2014].
 
@@
 In the standard PCA, size-related traits tended to load on PC1, with SLA, number of pinnae pairs, and degree of dissection on PC2.
-Loadings in the phylogenetic PCA are similar but approximately reversed relative to the standard PCA.
+Loadings in the phylogenetic PCA are similar, but size-related traits are approximately reversed relative to the standard PCA.
 
@@
 
-Most sporophyte traits, including all those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (`r table("pics")`).
+Most sporophyte traits, particularly those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (`r table("pics")`).
 Epiphytic species tend to have smaller values for all traits (`r table("pics")`; contrasts were made by subtracting epiphytic from terrestrial values).
@@
 **`r figure("cwm-div")`** Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
+Response variable abbreviations: SLA, specific leaf area.
 Epiphytic communities in green, terrestrial communities in brown.
 Error bars are each one standard deviation (values below zero not shown).
-Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at \pval{ }< 0.05.
+Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at \pval{ }< 0.05 (\rval^2^ not shown for SLA, which was fit with a spatial generalized linear mixed model; see Methods).
 \tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01.
 Results for frond width were very similar to frond length and not shown.
-Response variable abbreviations: SLA, specific leaf area.
 

### 2020-02-19 code/functions.R
Prefer linear models producing non-negative predictions for VPD

@@ -573,7 +573,9 @@ process_raw_climate <- function (climate_raw) {
       ),
       habit = fct_relevel(habit, c("epiphytic", "terrestrial")),
       site = str_remove_all(site, "_epi|_ter"),
-      vpd = plantecophys::RHtoVPD(rh, temp)
+      # plantecophys::RHtoVPD takes RH in percent (1 to 100), temp in C,
+      # outputs VPD in kPA
+      vpd = plantecophys::RHtoVPD(RH = rh, TdegC = temp)
     )
   
   # Final checks:
@@ -1542,9 +1544,12 @@ analyze_cwm_by_habit <- function (traits, comm) {
 #' and "lat" columns.
 #' @param resp_vars Vector of response variables in data
 #' to use. Defaults to column names of data
+#' @param prefer_positive Should models within AICc of 2 that produce
+#' positive response values be preferred? (applies to models where response
+#' var should be positive).
 #'
 #' @return Dataframe.
-choose_habit_elevation_models <- function (data, resp_vars) {
+choose_habit_elevation_models <- function (data, resp_vars, prefer_positive = FALSE) {
   
   # Helper function: check if any terms of a model are non-significant
   any_nonsignif_terms <- function(model) {
@@ -1558,6 +1563,14 @@ choose_habit_elevation_models <- function (data, resp_vars) {
     broom::tidy(model) %>%
       dplyr::summarize(all(p.value > 0.05)) %>%
       dplyr::pull(1)
+  }
+  
+  # Helper function: calculate the sum of all negative fits
+  sum_neg_fits <- function(model) {
+    broom::augment(model) %>%
+      filter(.fitted < 0) %>%
+      summarize(sum(.fitted)) %>%
+      pull(1)
   }
   
   # Reformat data into nested set by response variable
@@ -1588,35 +1601,46 @@ choose_habit_elevation_models <- function (data, resp_vars) {
     data_tibble %>%
     # Construct a linear model for each variable
     mutate(
-      interaction = map(data, ~lm(value ~ el * habit, .)),
-      both = map(data, ~lm(value ~ el + habit, .)),
-      habit_only = map(data, ~lm(value ~ habit, .)),
-      el_only = map(data, ~lm(value ~ el, .)),
+      interaction = map(data, ~lm(value ~ el * habit, data = .)),
+      both = map(data, ~lm(value ~ el + habit, data = .)),
[…]

### 2020-02-19 code/plan.R

@@ -185,9 +185,19 @@ plan <- drake_plan(
   # each climate var by elevation, growth habit, and
   # their interaction, then choose the best model for each.
   # Exclude single outlier site.
-  climate_models = choose_habit_elevation_models(
+  temperature_models = choose_habit_elevation_models(
     data = climate_trans_no_outlier,
-    resp_vars = climate_vars),
+    resp_vars = climate_vars[str_detect(climate_vars, "temp")],
+    prefer_positive = FALSE),
+  
+  # Prefer models within 2 delta AICc that have 
+  # non-negative predicted values for VPD.
+  vpd_models = choose_habit_elevation_models(
+    data = climate_trans_no_outlier,
+    resp_vars = climate_vars[str_detect(climate_vars, "vpd")],
+    prefer_positive = TRUE),
+  
+  climate_models = bind_rows(temperature_models, vpd_models),
   
   # - Check for spatial autocorrelation in models
   check_climate_model_spatial_autocorr = check_moran_pval(climate_models),
@@ -409,7 +419,10 @@ plan <- drake_plan(
   # by elevation, growth habit, and
   # their interaction, then choose the best model for each.
   # Also checks for spatial autocorrelation in the residuals.
-  div_el_models_nonspatial = choose_habit_elevation_models(div_metrics_all, resp_vars),
+  div_el_models_nonspatial = choose_habit_elevation_models(
+    data = div_metrics_all, 
+    resp_vars = resp_vars,
+    prefer_positive = FALSE),
   
   # - Only SLA showed spatial autocorrelation, 
   # so make a spatially-explicit model for this

### 2020-02-19 ms/SI.rmd
Add SI table for spatial model of SLA

@@
   "div_el_model_summaries",
+  "div_el_model_parameter_summaries",
   "cwsd_scatterplot"), 
@@
 
+**`r s_table("spatial-sla")`** `r s_table_cap("spatial-sla")` (see separate file)
+
+Spatial autocorrelation accounted for by including a Mátern spatial correlation structure in model (see Methods).
+
+```{r spatial-sla}
+# Format table
+spatial_sla_formatted <-
+div_el_model_parameter_summaries %>%
+  filter(var == "sla") %>%
+  select(term, Coefficient = estimate, SE = std.error, t = t_value) %>%
+  mutate(term = term %>%
+           str_replace_all("habitterrestrial", "Growth habit") %>%
+           str_replace_all("el", "Elevation")) %>%
+  rename(Term = term)
+
+# Set header and footer
+header <- c(
+  si_table_header,
+  s_table_nums("spatial-sla") %>% wrap_quotes
+)
+
+footer <- "Spatial autocorrelation accounted for by including a Mátern spatial correlation structure in model (see Methods)." %>%
+  str_wrap(width = 100) %>%
+  str_split("\\n") %>%
+  unlist %>%
+  purrr::map_chr(wrap_quotes)
+
+# Write out table with metadata
+filename <- s_table("spatial-sla") %>%
+  str_replace_all(" ", "_") %>%
+  paste0(".csv")
+
+# Write out table with metadata
+write_csv_with_meta(
+  spatial_sla_formatted, 
+  here::here("results", filename), 
+  header,
+  footer)
+```
+
 \clearpage

### 2020-02-19 ms/manuscript.Rmd
Add SI table for spatial model of SLA

@@
   "species_list",
-  "traits_with_tree"),
+  "traits_with_tree",
+  "spatial_sla_model"),
   cache = canopy_cache)
@@
 We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone, in combination, and in combination with their interactions.
-We then calculated the corrected Akaike Information Criterion [AIC~c~; @Akaike1973] and selected the best model with the combination of the fewest non-significant parameters and lowest AIC~c~.
+We then calculated the corrected Akaike Information Criterion [AIC~c~; @Akaike1973] for each model and retained models within 2 AIC~c~ of the lowest (\ie, 2 delta AIC~c~).
+From these selected we the model with the combination of the fewest non-significant parameters and lowest AIC~c~ (for VPD, we the selected the model producing the smallest amount of negative predicted values).
 A single outlier plot with unusually high temperature and low humidity (Mt. Rotui, 830 m plot) was excluded from ANCOVA and linear models.
@@
 }
+
+# Extract model summary values for SLA
+sla_model <-
+spatial_sla_model %>%
+  pull(model) %>%
+  magrittr::extract2(1)
+
+sla_parameters <- spaMM::get_ranPars(sla_model)
+
+sla_nu <- sla_parameters[["corrPars"]][["1"]][["nu"]] %>% round(2)
+sla_rho <- sla_parameters[["corrPars"]][["1"]][["rho"]] %>% round(2)
+sla_AIC <- AIC(sla_model)[[1]] %>% round(2)
 ```
@@
 Spatial autocorrelation was detected in the linear model for SLA (`r s_table("linear-fit")`), so a spatial generalized mixed model was used instead (see Methods). 
-Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities.
+Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (spatial generalized linear mixed model, nu = `r sla_nu`, rho = `r sla_rho`, AIC = `r sla_AIC`; `r s_table("spatial-sla")`).
 Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "dissection", "p")`).

### 2020-02-19 ms/captions.R

@@ -32,6 +32,7 @@ s_table_nums(name = "linear-fit", "Best-fitting linear models explaining climate
 s_table_nums(name = "phylosig-binary-strict", "Phylogenetic signal in binary traits of ferns on Moorea, French Polynesia, strict dataset excluding any species whose traits were scored based on taxonomy")
 s_table_nums(name = "corr-evo-strict", "Pagel's (1994) test of correlated evolution between binary traits of ferns on Moorea, French Polynesia, strict dataset excluding any species whose traits were scored based on taxonomy")
 s_table_nums(name = "gam-fit", "Best-fitting models explaining community diversity metrics of ferns on Moorea, French Polynesia by climate and growth habit")
+s_table_nums(name = "spatial-sla", "Output of spatial generalized linear mixed model for community-weighted mean specific leaf area of ferns on Moorea, French Polynesia predicted by elevation and growth habit")
 
 # Make short versions of citation functions
 # - Just the number

### 2020-02-19 code/functions.R
Use updated species names for Fig 4 and SI Table 1

@@ -2832,10 +2832,35 @@ make_heatmap <- function(important_div_vars, vars_select = c(
 #' @param traits Dataframe with traits of fern sporophytes and gametophytes
 #' including growth habit and others
 #' @param phy Phylogeny
+#' @param ppgi Dataframe with Pteridophyte Phylogeny Group I taxonomy
+#' @param updated_names Dataframe with new synonyms to use to replace original
+#' species names in `traits` and `phy`.
 #'
 #' @return ggplot object
 #' 
-plot_traits_on_tree <- function (traits, phy, ppgi) {
+plot_traits_on_tree <- function (traits, phy, ppgi, updated_names) {
+  
+  # Update species names in traits
+  traits <-
+  traits %>%
+    left_join(updated_names, by = c(species = "original_name")) %>%
+    mutate(species = case_when(
+      !is.na(new_name) ~ new_name,
+      TRUE ~ species
+    )) %>%
+    select(-new_name)
+  
+  # Update species names in tree
+  new_tip_labs <-
+  tibble(species = phy$tip.label) %>%
+    left_join(updated_names, by = c(species = "original_name")) %>%
+    mutate(species = case_when(
+      !is.na(new_name) ~ new_name,
+      TRUE ~ species
+    )) %>%
+    pull(species)
+  
+  phy$tip.label <- new_tip_labs
   
   # Subset and transform traits
   traits <-
@@ -2920,7 +2945,7 @@ plot_traits_on_tree <- function (traits, phy, ppgi) {
     ) %>%
     gather(trait, value, -species) %>%
     # reorder species by phylo order when plotting tree
-    left_join(phy_order) %>%
+    left_join(phy_order, by = "species") %>%
     mutate(
       species = fct_reorder(species, phy_order),
       trait = factor(trait, levels = c("habit", "morphotype", "gemmae", "glands", "hairs")),
@@ -2939,7 +2964,7 @@ plot_traits_on_tree <- function (traits, phy, ppgi) {
     select(species, habit, c("sla", "dissection", "stipe", "rhizome", "pinna")) %>%
     gather(trait, value, -species, -habit) %>%
     # reorder species by phylo order when plotting tree
-    left_join(phy_order) %>%
+    left_join(phy_order, by = "species") %>%
     mutate(species = fct_reorder(species, phy_order))
   
   ### Format data: phylogenetic tree
[…]

### 2020-02-19 code/plan.R

@@ -128,8 +128,13 @@ plan <- drake_plan(
     species_list,
     moorea_sites),
   
-  ### PPGI taxonomy ###
+  ### Taxonomy ###
+  
+  # - Pteridophyte Phylogeny Group 2016 taxonomic system
   ppgi = read_csv(file_in("data/ppgi_taxonomy.csv")),
+  
+  # - Updated synonymy for some species
+  updated_names = read_csv(file_in("data/name_update.csv")),
   
   # Format data for SI ----
   # - Traits: include SD and n for all quantitative (continuous) data
@@ -143,7 +148,7 @@ plan <- drake_plan(
   # (https://resolver.globalnames.org/). Since results have the possibility
   # of changing, the results are checked-in to the repo at the time of manuscript 
   # submission.
-  taxonomic_data_for_si = reformat_species_names(species_list) %>%
+  taxonomic_data_for_si = reformat_species_names(species_list, updated_names) %>%
     lookup_taxonomy,
   
   # Climate analysis ----
@@ -479,7 +484,8 @@ plan <- drake_plan(
   traits_with_tree = plot_traits_on_tree(
     traits = fern_traits,
     phy = phy,
-    ppgi = ppgi
+    ppgi = ppgi,
+    updated_names = updated_names
   ),
   
   # Make CWM scatterplots.

### 2020-02-19 ms/SI.rmd
Use a function to write out SI tables

--- 2020-02-07 ms/SI.rmd
@@
   "best_fit_div_models_moran",
-  "func_comm_struc_combined"), 
+  "func_comm_struc_combined",
+  "climate_model_summaries",
+  "climate_models",
+  "div_el_models",
+  "div_el_model_summaries",
+  "div_el_model_parameter_summaries",
+  "cwsd_scatterplot"), 
   cache = canopy_cache)
@@
 
-# Set up table headers
+# Load functions
+source(here::here("code/functions.R"))
+
+# Set up main table header
 si_table_header <- c(
@@
 )
+
 ```
@@
 
-'taxon code' is the name used for that taxon in analyses.
-'informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them.
-'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith (2003).
-'NA' indicates not applicable.
+```{r species-list}
+species_list_footer <- "'Taxon code' is the name used for that taxon in analyses.
+'Informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them.
+'Name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith (2003).
+The genus in 'Taxon code' may differ from the 'Genus' column for names whose preferred synonymy has changed since data were originally recorded.
+'NA' indicates not applicable."
 
-```{r species-list}
-# Set header and footer
-header <- c(
-  si_table_header,
-  s_table_nums("species-list") %>% wrap_quotes
+write_si_table(
+  table_id = "species-list", 
+  main_header = si_table_header,
+  table = taxonomic_data_for_si,
+  table_footer = species_list_footer
 )
+```
 
-# Wrap footer at ca. 100 characters per line. Each line needs to be an element of the
-# character vector.
-footer <- "'taxon code' is the name used for that taxon in analyses.
-'informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them.
-'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith (2003).
-'NA' indicates not applicable." %>%
-  str_wrap(width = 100) %>%
-  str_split("\\n") %>%
-  unlist %>%
-  purrr::map_chr(wrap_quotes)
-
-# Write out table with metadata
-write_csv_with_meta(
-  taxonomic_data_for_si, 
-  here::here("results/Table_S1.csv"), 
-  header,
-  footer)
-```
+`r species_list_footer`
 
@@
 
-SLA, specific leaf area.
+```{r trait_data}
+trait_values_footer <- "SLA, specific leaf area.
 Continuous measurements shown as mean, standard deviation (sample size).
@@
 For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI.
-'NA' indicates missing data.
+'NA' indicates missing data."
 
-```{r trait_data}
-# Set header and footer
-header <- c(
-  si_table_header,
-  s_table_nums("trait-values") %>% wrap_quotes
+write_si_table(
+  table_id = "trait-values", 
+  table = trait_data_for_si,
+  main_header = si_table_header,
+  table_footer = trait_values_footer
 )
+```
 
-footer <- "SLA, specific leaf area.
-Continuous measurements shown as mean, standard deviation (sample size).
-'taxon code' is the name used for that taxon in analyses (see Table S1).
-For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI.
-'NA' indicates missing data." %>%
-  str_wrap(width = 100) %>%
-  str_split("\\n") %>%
-  unlist %>%
-  purrr::map_chr(wrap_quotes)
-
-# Write out table with metadata
-write_csv_with_meta(
-  trait_data_for_si, 
-  here::here("results/Table_S2.csv"), 
-  header,
-  footer)
-```
+`r trait_values_footer`
 
@@
 
-d.f., degrees of freedom; MS, mean squares.
-Response variables include mean, minimum ('min'), maximum ('max'), and standard deviation ('sd') of temperature ('temp', °C) and vapor pressure deficit ('vpd', kPa).
-'NA' indicates not applicable.
-
-```{r climate-ancova}
+```{r ancova}
 # Format table
-climate_ancova_results_formatted <-
+ancova_table <-
 climate_ancova_results %>% 
@@
 
-# Set header and footer
-header <- c(
-  si_table_header,
-  s_table_nums("ancova") %>% wrap_quotes
+ancova_footer <- "d.f., degrees of freedom; MS, mean squares.
+Response variables include mean, minimum ('min'), maximum ('max'), and standard deviation ('sd') of temperature ('temp', °C) and vapor pressure deficit ('vpd', kPa).
+'NA' indicates not applicable."
+
+write_si_table(
+  table_id = "ancova", 
+  table = ancova_table,
+  main_header = si_table_header,
+  table_footer = ancova_footer
+)
+```
+
+`r ancova_footer`
+
+**`r s_table("linear-fit")`** `r s_table_cap("linear-fit")` (see separate file)
+
+```{r linear-fit}
+# Combine summaries of models for diversity metrics
+div_summaries <-
+  div_el_model_summaries %>%
+  left_join(
+    select(div_el_models,
+           var, model_type, delta_AICc, morans_I, I_pval), 
+    by = c("var", "model_type")
   )
 
-footer <- "d.f., degrees of freedom; MS, mean squares.
-Response variables include mean, minimum ('min'), maximum ('max'), and standard deviation ('sd') of temperature ('temp', °C) and vapor pressure deficit ('vpd', kPa).
-'NA' indicates not applicable." %>% wrap_quotes
+# Combine summaries of models for climate metrics
+climate_summaries <-
+  climate_model_summaries %>%
+  left_join(
+    select(climate_models,
+           var, model_type, delta_AICc, morans_I, I_pval), 
+    by = c("var", "model_type")
+  )
 
-# Write out table with metadata
-write_csv_with_meta(
-  climate_ancova_results_formatted, 
-  here::here("results/Table_S3.csv"), 
-  header,
-  footer)
+# Combine summaries of diversity and climate metrics, reformat model names
+combined_summaries <-
+  bind_rows(div_summaries, climate_summaries) %>%
+  mutate(
+    model_type = model_type %>% 
+      str_replace_all("el_only", "Elevation") %>%
+      str_replace_all("both", "Elevation + Growth habit") %>%
+      str_replace_all("habit_only", "Growth habit") %>%
+      str_replace_all("interaction", "Elevation + Growth habit + Elevation x Growth habit")
+    )
+
+# Reformat variable names and put in preferred order by joining
+linear_fit_table <-
+tribble(
+  ~var, ~Response,
+  "temp_max", "Max. temp.",
+  "temp_mean", "Mean temp.",
+  "vpd_mean", "Mean VPD",
+  "vpd_min", "Min. VPD",
+  "temp_sd", "SD temp.",
+  "vpd_sd", "SD VPD",
+  "ntaxa", "Richness",
+  "mpd.obs.z.phy", "MPDphy",
+  "mntd.obs.z.phy", "MNTDphy",
+  "mpd.obs.z.func", "MPDfunc",
+  "mntd.obs.z.func", "MNTDfunc",
+  "dissection", "Dissection",
+  "length", "Frond length",
+  "pinna", "Pinna no.",
+  "rhizome", "Rhizome diam.",
+  "stipe", "Stipe length",
+  "sla", "SLA",
+  "width", "Frond width"
+) %>%
+  left_join(combined_summaries, by = "var") %>%
+  select(
+    Response,
+    Model = model_type,
+    AICc,
+    f = statistic,
+    `Adjusted R2` = adj.r.squared,
+    P = p.value,
+    `d.f.` = df,
+    delta_AICc,
+    `Moran's I` = morans_I,
+    P_moran = I_pval
+  )
+
+linear_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
+Trait names refer to community-weighted mean values (see Methods).
+For details on calculations of MPD and MNTD, see Methods.
+'NA' indicates not applicable."
+
+write_si_table(
+  table_id = "linear-fit", 
+  table = linear_fit_table,
+  main_header = si_table_header,
+  table_footer = linear_fit_footer
+)
 ```
+
+`r linear_fit_footer`
 
@@
 
-P_rnd, probability of obtaining the observed value of \D under random trait evolution; P_bm, probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.
-
 ```{r phylosig-binary-strict}
 # Format table
-phylosig_binary_traits_strict_formatted <-
+phylosig_binary_strict_table <-
   phylosig_binary_traits_strict %>%
@@
 
-# Set header and footer
-header <- c(
-  si_table_header,
-  s_table_nums("phylosig-binary-strict") %>% wrap_quotes
+phylosig_binary_strict_footer <- "P_rnd, probability of obtaining the observed value of D under random trait evolution; P_bm, probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion."
+
+write_si_table(
+  table_id = "phylosig-binary-strict", 
+  table = phylosig_binary_strict_table,
+  main_header = si_table_header,
+  table_footer = phylosig_binary_strict_footer
 )
+```
 
-footer <- "P_rnd, probability of obtaining the observed value of D under random trait evolution; P_bm, probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion." %>%
-  str_wrap(width = 100) %>%
-  str_split("\\n") %>%
-  unlist %>%
-  purrr::map_chr(wrap_quotes)
-
-# Write out table with metadata
-write_csv_with_meta(
-  phylosig_binary_traits_strict_formatted, 
-  here::here("results/Table_S4.csv"), 
-  header,
-  footer)
-```
+P_rnd, probability of obtaining the observed value of \D under random trait evolution; P_bm, probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.
 
@@
 
-LL, log likelihood.
-Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
-terrestrial).
-Lower log likelihood for the dependent model indicates correlation with growth habit.
-
 ```{r corr-evo-strict}
 # Format table
-correlated_evo_test_strict_formatted <-
+corr_evo_strict_table <-
   correlated_evo_test_strict %>%
@@
 
-# Set header and footer
-header <- c(
-  si_table_header,
-  s_table_nums("corr-evo-strict") %>% wrap_quotes
-)
-
-footer <- "LL, log likelihood.
+corr_evo_strict_footer <- "LL, log likelihood.
 Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
 terrestrial).
-Lower log likelihood for the dependent model indicates correlation with growth habit." %>%
-  str_wrap(width = 100) %>%
-  str_split("\\n") %>%
-  unlist %>%
-  purrr::map_chr(wrap_quotes)
+Lower log likelihood for the dependent model indicates correlation with growth habit."
 
-# Write out table with metadata
-write_csv_with_meta(
-  correlated_evo_test_strict_formatted, 
-  here::here("results/Table_S5.csv"), 
-  header,
-  footer)
+write_si_table(
+  table_id = "corr-evo-strict", 
+  table = corr_evo_strict_table,
+  main_header = si_table_header,
+  table_footer = corr_evo_strict_footer
+)
 ```
 
-**`r s_table("model-fit")`** `r s_table_cap("model-fit")` (see separate file)
+`r corr_evo_strict_footer`
 
-MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, \pval\ value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's *I*.
-Models within delta AICc 3 of the best-fitting model shown.
-Trait names refer to community-weighted mean values (see Methods).
-Frond length and width were highly correlated with rhizome diameter and not shown.
-For details on calculations of MPD and MNTD, see Methods.
+**`r s_table("gam-fit")`** `r s_table_cap("gam-fit")` (see separate file)
 
-```{r model-fit}
+```{r gam-fit}
 # Format table
-best_fit_div_models_moran_formatted <-
+gam_fit_table <-
 best_fit_div_models_moran %>%
@@
 
-# Set header and footer
-header <- c(
-  si_table_header,
-  s_table_nums("model-fit") %>% wrap_quotes
-)
-
-footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
+gam_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
 Models within delta AICc 3 of the best-fitting model shown.
@@
 Frond length and width were highly correlated with rhizome diameter and not shown.
-For details on calculations of MPD and MNTD, see Methods." %>%
-  str_wrap(width = 100) %>%
-  str_split("\\n") %>%
-  unlist %>%
-  purrr::map_chr(wrap_quotes)
+For details on calculations of MPD and MNTD, see Methods."
 
-# Write out table with metadata
-write_csv_with_meta(
-  best_fit_div_models_moran_formatted, 
-  here::here("results/Table_S6.csv"), 
-  header,
-  footer)
+write_si_table(
+  table_id = "gam-fit", 
+  table = gam_fit_table,
+  main_header = si_table_header,
+  table_footer = gam_fit_footer
+)
 
 ```
+
+`r gam_fit_footer`
+
+**`r s_table("spatial-sla")`** `r s_table_cap("spatial-sla")` (see separate file)
+
+```{r spatial-sla}
+# Format table
+spatial_sla_table <-
+div_el_model_parameter_summaries %>%
+  filter(var == "sla") %>%
+  select(term, Coefficient = estimate, SE = std.error, t = t_value) %>%
+  mutate(term = term %>%
+           str_replace_all("habitterrestrial", "Growth habit") %>%
+           str_replace_all("el", "Elevation")) %>%
+  rename(Term = term)
+
+spatial_sla_footer <- "A Mátern spatial correlation structure was included in the model to account for spatial autocorrelation (see Methods)."
+
+write_si_table(
+  table_id = "spatial-sla", 
+  table = spatial_sla_table,
+  main_header = si_table_header,
+  table_footer = spatial_sla_footer
+)
+
+```
+
+`r spatial_sla_footer`
 

### 2020-02-19 code/functions.R

@@ -3670,8 +3670,48 @@ lookup_taxonomy <- function (taxa_names_df) {
       name_source == "The International Plant Names Index" ~ "IPNI",
       TRUE ~ name_source
     )) %>%
-    rename_all(~str_replace(., "_", " "))
+    assert(not_na, taxon_code, genus, specific_epithet, scientific_name, name_source) %>%
+    arrange(taxon_code) %>%
+    rename_all(~str_replace(., "_", " ") %>% str_to_sentence)
   
+}
+
+#' Write out an SI table
+#'
+#' @param table_id Table ID: should match the one used for captioner (see ms/captions.R)
+#' @param table Tibble: the formatted table to write out
+#' @param main_header A vector of main header text to add to the top of each table
+#' @param table_footer A table-specific footer
+#'
+#' @return Nothing. Externally, the table is written out to "results".
+#' 
+write_si_table <- function(table_id, table, main_header, table_footer) {
+  
+  # Combine main header with the header for this table
+  header <- c(
+    main_header,
+    s_table_nums(table_id) %>% wrap_quotes
+  )
+  
+  # Wrap footer at ca. 100 characters per line. Each line needs to be an element of the
+  # character vector.
+  footer <- table_footer %>% 
+    str_wrap(width = 100) %>%
+    str_split("\\n") %>%
+    unlist %>%
+    purrr::map_chr(wrap_quotes)
+  
+  # Set filename: 'Table_S1.csv, etc
+  filename <- s_table(table_id) %>%
+    str_replace_all(" ", "_") %>%
+    paste0(".csv")
+  
+  write_csv_with_meta(
+    table, 
+    here::here("results", filename), 
+    header,
+    footer)
+ 
 }
 
 # MS rendering ----

### 2020-02-19 ms/SI.rmd
Make correlation plot in plan

@@
   "best_fit_div_models_moran",
-  "func_comm_struc_combined",
+  "func_div_corr_plot",
   "climate_model_summaries",
@@
 ```{r Fig-S2, fig.height = 8, fig.width = 4, fig.cap = ""}
-
-# Tweak names for easier viewing
-func_comm_struc_combined_renamed <-
-func_comm_struc_combined %>%
-  mutate(dataset = str_remove_all(dataset, "func_comm_struc_test_fern_traits_")) %>%
-  mutate(dataset = str_remove_all(dataset, "\\.pool")) %>%
-  mutate(dataset = str_replace_all(dataset, "all_reduced", "reduced")) %>%
-  mutate(dataset = str_remove_all(dataset, "scaled_")) %>%
-  mutate(dataset = str_replace_all(dataset, "log_", "log-trans_"))
-
-# Don't need log-transformed sets that don't include log-transformed traits
-func_comm_struc_combined_renamed <- 
-  func_comm_struc_combined_renamed %>%
-  filter(
-    !dataset %in% c("log-trans_gameto", "log-trans_other")
-  )
-
-# Convert to wide format, run corrr: MPD
-mpd_corr <-
-  func_comm_struc_combined_renamed %>%
-  select(dataset, site, habit, mpd.obs.z) %>%
-  spread(dataset, mpd.obs.z) %>%
-  select(-site, -habit) %>%
-  correlate %>% 
-  rearrange(absolute = FALSE) # arrange so most correlated variables are closest
-
-# Make plot 
-# Modified from https://github.com/tidymodels/corrr/issues/35
-order <- mpd_corr %>% 
-  select(-rowname) %>% 
-  colnames()
-
-mpd_corr_plot <-
-mpd_corr %>% 
-  stretch(na.rm = TRUE) %>%
-  mutate_at(c("x", "y"), forcats::fct_relevel, ... = order) %>% 
-  ggplot(aes(x, y, fill = r)) +
-  geom_tile() +
-  geom_text(aes(label = as.character(fashion(r))), color = "white", size = 3) +
-  scale_fill_scico(palette = "roma", breaks = c(-1, 1)) +
-  labs(x = NULL, y = NULL, subtitle = "(a)") +
-  theme_minimal() +
-  theme(
-    panel.grid.minor = element_blank(),
-    panel.grid.major = element_blank(),
-    plot.subtitle = element_text(face = "bold"),
-    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
-  )
-
-# Convert to wide format, run corrr: MNTD
-mntd_corr <-
-  func_comm_struc_combined_renamed %>%
-  select(dataset, site, habit, mntd.obs.z) %>%
-  spread(dataset, mntd.obs.z) %>%
-  select(-site, -habit) %>%
-  correlate %>% 
-  rearrange(absolute = FALSE)
-
-mntd_corr_plot <-
-  mntd_corr %>% 
-  stretch(na.rm = TRUE) %>%
-  mutate_at(c("x", "y"), forcats::fct_relevel, ... = order) %>% # Use same order as MPD
-  ggplot(aes(x, y, fill = r)) +
-  geom_tile() +
-  geom_text(aes(label = as.character(fashion(r))), color = "white", size = 3) +
-  scale_fill_scico(palette = "roma", breaks = c(-1, 1)) +
-  labs(x = NULL, y = NULL, subtitle = "(b)") +
-  theme_minimal() +
-  theme(
-    panel.grid.minor = element_blank(),
-    panel.grid.major = element_blank(),
-    plot.subtitle = element_text(face = "bold"),
-    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
-  )
-
-fig_s2 <- mpd_corr_plot + mntd_corr_plot + plot_layout(nrow = 2, ncol = 1)
-
-fig_s2 <- fig_s2 + plot_annotation(theme = theme(plot.margin = margin(
-      0.2, 0.2, 0.2, 0.2, unit = "in")))
-
-fig_s2
+func_div_corr_plot
 ```

### 2020-02-19 code/functions.R

@@ -3450,6 +3450,91 @@ make_cwsd_scatterplot <- function(cwm_long, habit_colors) {
   
 }
 
+# Make plot showing correlation between results of running functional structure analyses
+# using different sets of traits as input
+make_func_div_corr_plot <- function (func_comm_struc_combined) {
+  
+  # Tweak names for easier viewing
+  func_comm_struc_combined_renamed <-
+    func_comm_struc_combined %>%
+    mutate(dataset = str_remove_all(dataset, "func_comm_struc_test_fern_traits_")) %>%
+    mutate(dataset = str_remove_all(dataset, "\\.pool")) %>%
+    mutate(dataset = str_replace_all(dataset, "all_reduced", "reduced")) %>%
+    mutate(dataset = str_remove_all(dataset, "scaled_")) %>%
+    mutate(dataset = str_replace_all(dataset, "log_", "log-trans_"))
+  
+  # Don't need log-transformed sets that don't include log-transformed traits
+  func_comm_struc_combined_renamed <- 
+    func_comm_struc_combined_renamed %>%
+    filter(
+      !dataset %in% c("log-trans_gameto", "log-trans_other")
+    )
+  
+  # Convert to wide format, run corrr: MPD
+  mpd_corr <-
+    func_comm_struc_combined_renamed %>%
+    select(dataset, site, habit, mpd.obs.z) %>%
+    spread(dataset, mpd.obs.z) %>%
+    select(-site, -habit) %>%
+    correlate %>% 
+    rearrange(absolute = FALSE) # arrange so most correlated variables are closest
+  
+  # Make plot 
+  # Modified from https://github.com/tidymodels/corrr/issues/35
+  order <- mpd_corr %>% 
+    select(-rowname) %>% 
+    colnames()
+  
+  mpd_corr_plot <-
+    mpd_corr %>% 
+    stretch(na.rm = TRUE) %>%
+    mutate_at(c("x", "y"), forcats::fct_relevel, ... = order) %>% 
+    ggplot(aes(x, y, fill = r)) +
+    geom_tile() +
+    geom_text(aes(label = as.character(fashion(r))), color = "white", size = 3) +
+    scale_fill_scico(palette = "roma", breaks = c(-1, 1)) +
+    labs(x = NULL, y = NULL, subtitle = "(a)") +
+    theme_minimal() +
+    theme(
[…]

### 2020-02-19 code/plan.R

@@ -544,6 +544,9 @@ plan <- drake_plan(
   # for SI
   cwsd_scatterplot = make_cwsd_scatterplot(cwm_long, habit_colors),
   
+  # Make correlation plot of testing different sets of traits for calculating functional diversity
+  func_div_corr_plot = make_func_div_corr_plot(func_comm_struc_combined),
+  
   # Manuscript ----
   
   # First render to PDF, keeping the latex

### 2020-02-19 ms/SI.rmd
Improve table S2

@@
 ```{r trait_data}
-trait_values_footer <- "SLA, specific leaf area.
+trait_values_footer <- "SLA, specific leaf area; 'Taxon code', name used for that taxon in analysis (see Table S1); 'Growth habit', growth habit observed in the field; 'Binary growth habit', binary growth habit used in analysis (see Methods).
 Continuous measurements shown as mean, standard deviation (sample size).
-'taxon code' is the name used for that taxon in analyses (see Table S1).
 For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI.
+For frond dissection, 1 = simple, 2 = pinnatifid, 3 = one-pinnate, 4 = one-pinnate-pinnatifid, 5 = two-pinnate, 6 = two-pinnate-pinnatifid, 7 = bipinnatifid_or_tripinnatifid, 8 = three-pinnate, 9 = three-pinnate-pinnatifid.
 'NA' indicates missing data."
+
+# Format table: capitalize column names
+trait_values_table <-
+trait_data_for_si %>% 
+  rename_all(str_to_sentence) %>% 
+  rename(`SLA (sq m per kg)` = `Sla (sq m per kg)`)
 
@@
   table_id = "trait-values", 
-  table = trait_data_for_si,
+  table = trait_values_table,
   main_header = si_table_header,
@@
 ancova_table <-
-climate_ancova_results %>% 
+  tribble(
+  ~variable, ~Response,
+  "temp_max", "Max. temp.",
+  "temp_mean", "Mean temp.",
+  "temp_sd", "SD temp.",
+  "vpd_min", "Min. VPD",
+  "vpd_mean", "Mean VPD",
+  "vpd_sd", "SD VPD",
+) %>%
+  left_join(climate_ancova_results, by = "variable") %>% 
   mutate(
@@
   ) %>%
-  select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value)
+  select(Response, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value)
 
 ancova_footer <- "d.f., degrees of freedom; MS, mean squares.
-Response variables include mean, minimum ('min'), maximum ('max'), and standard deviation ('sd') of temperature ('temp', °C) and vapor pressure deficit ('vpd', kPa).
+Response variables include mean, minimum, maximum, and standard deviation of temperature ('temp', °C) and vapor pressure deficit ('VPD', kPa).
+Vapor pressure deficit was square-root transformed prior to analysis.
 'NA' indicates not applicable."
@@
   "temp_mean", "Mean temp.",
+  "temp_sd", "SD temp.",
+  "vpd_min", "Min. VPD",
   "vpd_mean", "Mean VPD",
-  "vpd_min", "Min. VPD",
-  "temp_sd", "SD temp.",
   "vpd_sd", "SD VPD",
@@
     `d.f.` = df,
-    delta_AICc,
+    `delta AICc` = delta_AICc,
     `Moran's I` = morans_I,
-    P_moran = I_pval
+    `P moran` = I_pval
   )
 
-linear_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
+linear_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; 'P moran', P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
 Trait names refer to community-weighted mean values (see Methods).
@@
   arrange(trait) %>%
-  select(Trait = trait, `Number of presences` = num_present, `Number of absences` = num_absent, D,
-         P_rnd = prob_random, P_bm = prob_brownian)
+  select(Trait = trait, 
+         `Number of presences` = num_present, 
+         `Number of absences` = num_absent, 
+         D,
+         `P rnd` = prob_random, 
+         `P bm` = prob_brownian)
 
-phylosig_binary_strict_footer <- "P_rnd, probability of obtaining the observed value of D under random trait evolution; P_bm, probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion."
+phylosig_binary_strict_footer <- "'P rnd', probability of obtaining the observed value of D under random trait evolution; 'P bm', probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion."
 
@@
 
-P_rnd, probability of obtaining the observed value of \D under random trait evolution; P_bm, probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.
+'P rnd', probability of obtaining the observed value of \D under random trait evolution; 'P bm', probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.
 
@@
 # Format table
+
+# First make join table for converting variable names, in order
+resp_var_table <-
+tribble(
+  ~resp_var, ~Response,
+  "ntaxa", "Richness",
+  "mpd.obs.z.phy", "MPDphy",
+  "mntd.obs.z.phy", "MNTDphy",
+  "mpd.obs.z.func", "MPDfunc",
+  "mntd.obs.z.func", "MNTDfunc",
+  "dissection", "Dissection",
+  "length", "Frond length",
+  "width", "Frond width",
+  "pinna", "Pinna no.",
+  "rhizome", "Rhizome diam.",
+  "stipe", "Stipe length",
+  "sla", "SLA"
+)
+
 gam_fit_table <-
-best_fit_div_models_moran %>%
-  # Drop frond length and width (highly correlated with rhizome dia.)
-  filter(resp_var %in% c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
-                   "mpd.obs.z.func", "mntd.obs.z.func",
-                   "dissection", "sla", "stipe", "rhizome", "pinna")) %>%
-  mutate(resp_var = factor(
-        resp_var, 
-        levels = c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
-                   "mpd.obs.z.func", "mntd.obs.z.func",
-                   "dissection", "sla", "stipe", "rhizome", "pinna"))) %>%
-  mutate(
-    resp_var = lvls_revalue(
-        resp_var, 
-        new_levels = c("Richness", "MPDphy", "MNTDphy", 
-                       "MPDfunc", "MNTDfunc",
-                       "Dissection", "SLA", "Stipe length", "Rhizome diam.", "Pinna no.")) 
-  ) %>%
+resp_var_table %>%
+  left_join(best_fit_div_models_moran, by = "resp_var") %>%
+  mutate(resp_var = fct_relevel(resp_var, resp_var_table$resp_var)) %>%
   arrange(resp_var, delta_AICc) %>%
@@
   select(
-    `Metric` = resp_var,
+    Response,
     `Model` = modname,
@@
     `R2` = r2,
-    delta_AICc,
+    `delta AICc` = delta_AICc,
     `Moran's I` = morans_I,
-    `P_moran` = morans_I_pval
+    `P moran` = morans_I_pval
   )
 
-gam_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
+gam_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; SLA, specific leaf area; VPD, vapor pressure deficit, AICc, corrected Akaike Information Criterion; 'P moran', P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
 Models within delta AICc 3 of the best-fitting model shown.
 Trait names refer to community-weighted mean values (see Methods).
-Frond length and width were highly correlated with rhizome diameter and not shown.
 For details on calculations of MPD and MNTD, see Methods."

### 2020-02-19 code/functions.R

@@ -223,31 +223,14 @@ process_raw_cont_morph <- function (raw_morph_path, species_list) {
 
 #' Process raw morphological qualitative trait data
 #'
-#' @param raw_morph_path Data to raw data file with morphological
-#' trait observations
+#' @param raw_morph Raw data with morphological trait observations
 #' @param species_list Vector of accepted species to include
 #'
 #' @return Tibble, with one trait value per species
 #'
-process_raw_qual_morph <- function (raw_morph_path, species_list) {
+process_raw_qual_morph <- function (raw_morph, species_list) {
   
-  # Format data sources to use names instead of number codes to minimize confusion
-  # original scheme:
-  #1 = lab obs.
-  #2 = Nayar and Kauar 1971 "Gametophytes of homosporous ferns" The Botanical Review [@Nayar1971]
-  #3 = Lloyd 1980 "Reproductive biology and gametophyte morphology of New World populations of Acrostichum aureum" AFJ [@Lloyd1980]
-  #4 = field obs.
-  #5 = genus/family level character (taxonomy)
-  #6 = Zhang et al 2008 "Gametophyte Morphology and Development of Six Chinese Species of Pteris (Pteridaceae)" AFJ [@Zhang2008]
-  #7 = Bierhorst 1967 "The gametophyte of Schizaea dichotoma" AJB [@Bierhorst1967]
-  #8 = Martin et al 2006 "Efficient induction of apospory and apogamy in vitro in silver fern (Pityrogramma calomelanos L.)" Plant Cell Reports [@Martin2006]
-  #9 = Tigerschiöld 1989 "Dehiscence of antheridia in thelypteroid ferns" Nordic Journal of Botany [@Tigerschiold1989]
-  #10 = Tigerschiöld 1990 "Gametophytes of some Ceylonese species of Thelypteridaceae" Nordic Journal of Botany [@Tigerschiold1990]
-  #11 = Atkinson 1975 "The gametophyte of five Old World thelypteroid ferns" Phytomorphology [@Atkinson1975]
-  #12 = Chen et al 2014 "First insights into the evolutionary history of the Davallia repens complex" Blumea [@Chen2014a]
-  
-  # Read in raw trait data
-  read_csv(raw_morph_path, na = c("?", "n/a", "NA", "N/A")) %>%
+  raw_morph %>%
     clean_names() %>%
[…]

### 2020-02-19 code/plan.R

@@ -40,28 +40,31 @@ plan <- drake_plan(
   # (includes one measure per specimen,
   #  multiple specimens per species).
   # All measurements in cm, except number of pinna pairs.
-  morph_cont_raw = process_raw_cont_morph(
+  morph_cont = process_raw_cont_morph(
     file_in("data/morph_measurements.csv"),
     species_list
   ),
   
   # - Process raw qualitative trait data
   # (includes one obsevation per specimen).
-  morph_qual_raw = process_raw_qual_morph(
+  morph_qual_raw = read_csv(
     file_in("data/morph_qual_traits.csv"),
+    na = c("?", "n/a", "NA", "N/A")),
+  
+  morph_qual = process_raw_qual_morph(
+    morph_qual_raw,
     species_list
   ),
   
   # - Combine raw trait data into final trait matrix
-  fern_traits = make_trait_matrix(sla_raw, morph_cont_raw, morph_qual_raw) %>%
-    select(-source_1, -source_2),
+  fern_traits = make_trait_matrix(sla_raw, morph_cont, morph_qual) %>%
+    select(-in_strict),
   
   # - Also make "strict" trait dataset by excluding species with 
   # gameto traits only known from taxonomy
-  fern_traits_strict = make_trait_matrix(sla_raw, morph_cont_raw, morph_qual_raw) %>%
-    mutate_at(vars(source_1, source_2), ~replace_na(., "none")) %>%
-    filter(source_1 != "T", source_2 != "T") %>%
-    select(-source_1, -source_2),
+  fern_traits_strict = make_trait_matrix(sla_raw, morph_cont, morph_qual) %>%
+    filter(in_strict == TRUE) %>%
+    select(-in_strict),
   
   # - Also make trait dataset with continuous traits scaled, but not log-transformed.
   fern_traits_scaled = transform_traits(
@@ -139,7 +142,7 @@ plan <- drake_plan(
   # Format data for SI ----
   # - Traits: include SD and n for all quantitative (continuous) data
   trait_data_for_si = process_trait_data_for_si(
-    sla_raw, morph_cont_raw, morph_qual_raw
+    sla_raw, morph_cont, morph_qual, morph_qual_raw
   ),
   
   # - Species names
@@ -398,7 +401,7 @@ plan <- drake_plan(
   # - Extract table of importance of each environmental variable.
   important_div_vars = get_important_vars(fss_div_results),
   
-  # - Extract table of best-fit models.
[…]

### 2020-02-20 code/functions.R
Put R2 in lower right corner for MPDfunc scatterplot

@@ -2153,6 +2153,12 @@ make_climate_scatterplot <- function (yval, xval = "el", ylab = yval, xlab = "El
   model_type <- pull(summaries, model_type)
   
   # Plot lines only for models with a significant elevation effect
+  
+  # First make sure we can deal with the type of model passed to this function
+  assertthat::assert_that(
+    model_type %in% c("interaction", "el_only", "habit_only", "both"),
+    msg = "model_type not in valid set for plotting") 
+  
   if(str_detect(model_type, "el_only")) {
     plot <- ggplot(data, aes(x = !!xval_sym))
     if (p < 0.05) plot <- plot + geom_line(data = fits, aes(y = .fitted))
@@ -2306,7 +2312,7 @@ make_div_scatterplot <- function (yval, xval = "el", ylab = yval, xlab = "Elevat
   plot <- plot + geom_point(aes(y = !!yval_sym, color = habit))  
   
   # Set location of where to print R-squared
-  if((model_type == "el_only" | model_type == "interaction") & !is.na(p) & p < 0.05) {
+  if((model_type == "el_only" | model_type == "interaction" | model_type == "both") & !is.na(p) & p < 0.05) {
     if (isTRUE(r_upper)) {
       plot <- plot +
         annotate("text", x = Inf, y = Inf, 
@@ -2315,6 +2321,7 @@ make_div_scatterplot <- function (yval, xval = "el", ylab = yval, xlab = "Elevat
                  size = 9 / .pt,
                  parse = TRUE)
     } else {
+      # bottom right
       plot <- plot +
         annotate("text", x = Inf, y = -Inf, 
                  label = glue::glue("italic(R) ^ 2 == {r2}"),

### 2020-02-20 code/plan.R

@@ -522,14 +522,16 @@ plan <- drake_plan(
     scatterplots = cwm_scatterplots),
   
   # Make community diversity scatterplots
-  div_scatterplots = map(
+  div_scatterplots = map2(
     c("ntaxa", "mpd.obs.z.phy",  "mntd.obs.z.phy", 
       "mpd.obs.z.func", "mntd.obs.z.func") %>% set_names(.),
+    c(TRUE, TRUE, TRUE, FALSE, TRUE),
     ~ make_div_scatterplot(
       data = div_metrics_all,
       fits = div_el_model_fits,
       summaries = div_el_model_summaries,
-      yval = .,
+      yval = .x,
+      r_upper = .y,
       habit_colors = habit_colors)
   ),
   

### 2020-02-20 ms/SI.rmd
Update MS

@@
       term == "el" ~ "Elevation",
-      term == "habit:el" ~ "Growth habit × elevation",
+      term == "habit:el" ~ "Growth habit x elevation",
       TRUE ~ term
@@
 
-phylosig_binary_strict_footer <- "'P rnd', probability of obtaining the observed value of D under random trait evolution; 'P bm', probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion."
+phylosig_binary_strict_footer <- "D, Fritz and Purvis' D (Fritz & Purvis, 2010); 'P rnd', probability of obtaining the observed value of D under random trait evolution; 'P bm', probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion."
 
@@
 
-'P rnd', probability of obtaining the observed value of \D under random trait evolution; 'P bm', probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.
+\D, Fritz and Purvis' \D (Fritz & Purvis, 2010); 'P rnd', probability of obtaining the observed value of \D under random trait evolution; 'P bm', probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.
 
@@
 terrestrial).
-Lower log likelihood for the dependent model indicates correlation with growth habit."
+Higher (less negative) log likelihood for the dependent model indicates correlation with growth habit."
 
@@
 
-```{r Fig-S1, fig.height = 6, fig.width = 7, fig.cap = ""}
-cwsd_scatterplot
-```
-
-**`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`.
-Epiphytic communities in green, terrestrial communities in brown.
-Points randomly jittered along x-axis to reduce overlap.
-Boxplots show median values (bold lines) by growth habit.
-Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range.
-Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \* = \pval{ }< 0.05 **(a)**, traits related to plant size **(b)**, traits not related to plant size.
-Response variable abbreviations: SLA, specific leaf area.
-
-\clearpage
-
-```{r Fig-S2, fig.height = 8, fig.width = 4, fig.cap = ""}
+```{r corr-plot, fig.height = 8, fig.width = 4, fig.cap = ""}
 func_div_corr_plot
@@
 The log-transformed 'reduced' dataset was used for the main results.
+
+\clearpage
+
+```{r cwsd, fig.height = 6, fig.width = 7, fig.cap = ""}
+cwsd_scatterplot
+```
+
+**`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`.
+Epiphytic communities in green, terrestrial communities in brown.
+Points randomly jittered along x-axis to reduce overlap.
+Boxplots show median values (bold lines) by growth habit.
+Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range.
+Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; \*\*\* = \pval{ }< 0.001.
+**(a)**, traits related to plant size **(b)**, traits not related to plant size.
+Response variable abbreviations: SLA, specific leaf area.
 
@@
 *Bernice P. Bishop Museum Bulletin* **93**: 1--86.
+
+**Fritz SA, Purvis A**. **2010**. Selectivity in mammalian extinction risk and 
+threat types: A new measure of phylogenetic signal strength in binary traits. 
+*Conservation Biology* **24**: 1042--1051.
 

### 2020-02-20 ms/manuscript.Rmd
Make join explicit
Update MS

@@
 We then calculated the corrected Akaike Information Criterion [AIC~c~; @Akaike1973] for each model and retained models within 2 AIC~c~ of the lowest (\ie, 2 delta AIC~c~).
-From these selected we the model with the combination of the fewest non-significant parameters and lowest AIC~c~ (for VPD, we the selected the model producing the smallest amount of negative predicted values).
+From these, we selected the model with the combination of the fewest non-significant parameters and lowest AIC~c~ (for VPD, we the selected the model producing the smallest amount of negative predicted values).
 A single outlier plot with unusually high temperature and low humidity (Mt. Rotui, 830 m plot) was excluded from ANCOVA and linear models.
@@
 For functional distances, we log-transformed and scaled traits as in the PCA, excluded frond width and length (both frond size traits highly correlated with stipe length), then calculated the distance matrix using Gower's dissimilarity coefficient [@Gower1971] with the \function{gowdis} function in the R package \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
-To test for the effect of different trait selection schemes, we compared the results of running the same analyses for various sets of traits with and without log-transformation (Supporting Information).
+To test for the effect of different trait selection schemes, we compared the results of running the same analyses for various sets of traits with and without log-transformation (`r s_figure("corr-plot")`).
 
@@
 We checked for spatial autocorrelation in the model residuals with Moran's *I* [@Moran1950] as implemented in the \package{spdep} R package [@Bivand2013]. 
-In the case the spatial autocorrelation was detected, a generalized linear mixed model was fit with a Mátern spatial correlation structure using the \package{spAMM} R package [@Rousset2014].
+If spatial autocorrelation was detected, a generalized linear mixed model was fit with a Mátern spatial correlation structure using the \package{spAMM} R package [@Rousset2014].
 
@@
 
-Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (`r figure("pca")`**c**, **d**).
+Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (`r figure("pca")`**c**, **d**).
 The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
@@
 Most sporophyte traits, particularly those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (`r table("pics")`).
-Epiphytic species tend to have smaller values for all traits (`r table("pics")`; contrasts were made by subtracting epiphytic from terrestrial values).
+Epiphytic species tend to have smaller values for all traits (`r table("pics")`).
 Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval{ }`r stipe_pval`), as well as frond width (\pval{ }`r width_pval`) and degree of laminae dissection (\pval{ }`r dissection_pval`).
@@
   comm %>%
-  mutate(abundance = 
-           comm %>%
-           select_if(is.numeric) %>%
-           rowSums()
-  ) %>%
-  select(species, abundance)
+  pivot_longer(cols = -species, values_to = "abundance", names_to = "site") %>%
+  assert(not_na, abundance) %>%
+  group_by(species) %>%
+  summarize(abundance = sum(abundance)) %>%
+  ungroup()
 
@@
 traits_with_abun <- left_join(
-  fern_traits, abundances
+  fern_traits, abundances, by = "species"
 )
@@
 Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (spatial generalized linear mixed model, nu = `r sla_nu`, rho = `r sla_rho`, AIC = `r sla_AIC`; `r s_table("spatial-sla")`).
-Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "dissection", "p")`).
+Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }`r get_mod_result(model_summ_print, "dissection", "p")`).
 Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r figure("cwm-div")`).
 
-We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` \species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
+We observed a total of `r abundances %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` \species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
 Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r figure("comm-div")`).
@@
 This may reflect adaptations to specific niches that are not captured by a binary categorization of growth habit.
-For example, the following species ranked among those with the five lowest SLA values (\ie, thick, well-protected leaves) on Moorea: *Humata anderssonii* (Mett.) C.Chr.
+For example, the following species ranked among those with the five lowest SLA values (\ie, thick, well-protected leaves) on Moorea: *Davallia anderssonii* Mett.
 (a twig epiphyte), *Acrostichum aureum* L.
@@
 
+**`r s_table("spatial-sla")`** `r s_table_cap("spatial-sla")`
+
+**`r s_figure("corr-plot")`** `r s_figure_cap("corr-plot")`
+
 **`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`
-
-**`r s_figure("corr-plot")`** `r s_figure_cap("corr-plot")`
 
@@
 Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval{ }< 0.05.
-Asterisks indicate statistical significance of linear model; \*\*\* = \pval{ }< 0.001; \* = \pval{ }< 0.05.
+Asterisks indicate statistical significance of linear model; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01.
 Partially transparent colors indicate outliers excluded from the models (see Methods).
@@
 Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at \pval{ }< 0.05 (\rval^2^ not shown for SLA, which was fit with a spatial generalized linear mixed model; see Methods).
-\tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01.
+\tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\* = \pval{ }< 0.001.
 Results for frond width were very similar to frond length and not shown.
@@
 # table in docx (this happens with gt or kable), and fix italics, superscript, etc.
-read_csv(here::here("data/table_1.csv")) %>%
+read_csv(here::here("data/table_1.csv"), col_types = "cccc") %>%
   rename(`Data type (unit)` = data_type, `Functional significance` = ecological_significance, `Reference(s)` = reference) %>%
@@
 terrestrial growth).
-Traits with a significantly lower log likelihood for the dependent model are significantly correlated with growth habit.
+Traits with a significantly higher (\ie, less negative) log likelihood for the dependent model are correlated with growth habit.
 

### 2020-02-20 ms/captions.R

@@ -20,8 +20,8 @@ table_nums(name = "corr-evo", caption = "Pagel's (1994) test of correlated evolu
 
 # - SI figures
 s_fig_nums <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
-s_fig_nums(name = "cwsd", "Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia")
 s_fig_nums(name = "corr-plot", "Correlation plot comparing functional diversity of fern communities on Moorea, French Polynesia calculated using different sets of traits as input")
+s_fig_nums(name = "cwsd", "Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia")
 
 # - SI tables
 s_table_nums <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")

### 2020-02-21 code/plan.R
Add check on Moran's I in non-spatial model for SLA

@@ -453,6 +453,8 @@ plan <- drake_plan(
   
   # - Replace SLA model in model list with the spatial version
   div_el_models = div_el_models_nonspatial %>%
+    # Verify that Moran's I is significant for SLA in non-spatial model before swapping
+    verify(filter(., var == "sla") %>% pull(I_pval) < 0.05) %>%
     filter(var != "sla") %>%
     bind_rows(spatial_sla_model),
   

### 2020-02-21 code/plan.R
Render SI to docx for spell checking

@@ -579,6 +579,13 @@ plan <- drake_plan(
     tracked_output = file_out(here::here("results/manuscript.tex"))
   ),
   
+  si_pdf = render_tracked(
+    knitr_in("ms/SI.Rmd"),
+    quiet = TRUE,
+    output_dir = here::here("results"),
+    tracked_output = file_out(here::here("results/SI.tex"))
+  ),
+  
   # Next use the latex to convert to docx with pandoc
   ms_docx = latex2docx(
     latex = file_in(here::here("results/manuscript.tex")),
@@ -587,11 +594,11 @@ plan <- drake_plan(
     wd = here::here("results")
   ),
   
-  # SI ----
-  si_pdf = rmarkdown::render(
-    knitr_in("ms/SI.Rmd"),
-    quiet = TRUE,
-    output_dir = here::here("results")
+  si_docx = latex2docx(
+    latex = file_in(here::here("results/SI.tex")),
+    docx = file_out(here::here("results/SI.docx")),
+    template = file_in(here::here("ms/new-phytologist.docx")),
+    wd = here::here("results")
   )
   
 )

### 2020-02-21 ms/SI.rmd
Update SI

@@
   "climate_models",
+  "spatial_sla_model",
   "div_el_models",
@@
 'Informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them.
-'Name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith (2003).
+'Name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), Murdock and Smith 2003 = Murdock & Smith (2003), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/).
 The genus in 'Taxon code' may differ from the 'Genus' column for names whose preferred synonymy has changed since data were originally recorded.
@@
 ```{r trait_data}
-trait_values_footer <- "SLA, specific leaf area; 'Taxon code', name used for that taxon in analysis (see Table S1); 'Growth habit', growth habit observed in the field; 'Binary growth habit', binary growth habit used in analysis (see Methods).
+trait_values_footer <- "'Taxon code', name used for that taxon in analysis (see Table S1); 'Growth habit', growth habit observed in the field; 'Binary growth habit', binary growth habit used in analysis (see Methods); SLA, specific leaf area.
 Continuous measurements shown as mean, standard deviation (sample size).
-For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI.
-For frond dissection, 1 = simple, 2 = pinnatifid, 3 = one-pinnate, 4 = one-pinnate-pinnatifid, 5 = two-pinnate, 6 = two-pinnate-pinnatifid, 7 = bipinnatifid_or_tripinnatifid, 8 = three-pinnate, 9 = three-pinnate-pinnatifid.
+For data source, F = field observation of gametophyte subsequently identified using DNA barcoding, L = lab observation of gametophyte cultured from spore, M = measurement of a field-collected specimen, T = morphology inferred based on taxonomy; cited references provided at end of SI.
+For frond dissection, 1 = simple, 2 = pinnatifid, 3 = one-pinnate, 4 = one-pinnate-pinnatifid, 5 = two-pinnate, 6 = two-pinnate-pinnatifid, 7 = bipinnatifid or tripinnatifid, 8 = three-pinnate, 9 = three-pinnate-pinnatifid.
 'NA' indicates missing data."
@@
 Vapor pressure deficit was square-root transformed prior to analysis.
+Highly correlated response variables (Pearson's correlation coefficient > 0.9) not shown.
 'NA' indicates not applicable."
@@
 
+# Add AIC for SLA model
+sla_model <-
+  spatial_sla_model %>%
+  pull(model) %>%
+  magrittr::extract2(1)
+
 # Combine summaries of diversity and climate metrics, reformat model names
@@
     model_type = model_type %>% 
+      str_replace_all("both_spatial", "Elevation + Growth habit") %>%
+      str_replace_all("both", "Elevation + Growth habit") %>%
       str_replace_all("el_only", "Elevation") %>%
-      str_replace_all("both", "Elevation + Growth habit") %>%
       str_replace_all("habit_only", "Growth habit") %>%
       str_replace_all("interaction", "Elevation + Growth habit + Elevation x Growth habit")
+    ) %>%
+  mutate(
+    AIC = case_when(
+      var == "sla" ~ AIC(sla_model)[[1]],
+      TRUE ~ AIC
+    )
   )
@@
   "length", "Frond length",
+  "width", "Frond width",
   "pinna", "Pinna no.",
   "rhizome", "Rhizome diam.",
-  "stipe", "Stipe length",
   "sla", "SLA",
-  "width", "Frond width"
+  "stipe", "Stipe length"
 ) %>%
@@
     Model = model_type,
+    AIC,
     AICc,
@@
 
-linear_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; 'P moran', P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
+linear_fit_footer <- "VPD, vapor pressure deficit; MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; SLA, specific leaf area; AIC, Akaike Information Criterion; AICc, corrected AIC; d.f., degrees of freedom; 'P moran', P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I.
 Trait names refer to community-weighted mean values (see Methods).
+Specific leaf area fit with a generalized linear mixed model including a Mátern spatial correlation structure; all others fit with linear models (see Methods).
 For details on calculations of MPD and MNTD, see Methods.
@@
   "rhizome", "Rhizome diam.",
-  "stipe", "Stipe length",
-  "sla", "SLA"
+  "sla", "SLA",
+  "stipe", "Stipe length"
 )
@@
 
-spatial_sla_footer <- "A Mátern spatial correlation structure was included in the model to account for spatial autocorrelation (see Methods)."
+spatial_sla_footer <- "The effect of growth habit is relative to epiphytic growth.
+A Mátern spatial correlation structure was included in the model to account for spatial autocorrelation (see Methods)."
 
@@
 For details on calculations of MPD~func~ and MNTD~func~, see Methods.
-Correlations between functional diversity as measured with different sets of traits (values in each tile) calculated with Pearson's correlation coefficient using the R package 'corrr'.
+Correlations between functional diversity as measured with different sets of traits (values in each tile) calculated with Pearson's correlation coefficient using the R package 'corrr' (Kuhn *et al.*, 2019).
 All trait sets were scaled to standard deviation of each trait.
 The following trait sets were used: 'all' (all traits), 'size' (sporophyte traits related to size; frond length and width, stipe length, rhizome width), 'other' (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), 'gameto' (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), 'reduced' (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter).
-For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), both log-transformed and untransformed sets were included.
+For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), log-transformed sets ('log-trans_all' and 'log-trans_reduced') were also included.
 The log-transformed 'reduced' dataset was used for the main results.
@@
 
+**Kuhn M, Jackson S, Cimentada J**. **2019**. *Corrr: Correlations in r*. R package 
+version 0.4.0 [WWW document] URL http://CRAN.R-project.org/package=corrr. 
+[accessed 1 January 2020].
+
 **Lloyd RM**. **1980**. Reproductive biology and gametophyte morphology
@@
 **Zhang K, Shi L, Zhang X, Jiang C, Tim-Chun W**. **2008**. Gametophyte
-morphology and development of six chinese species of *Pteris*
+morphology and development of six Chinese species of *Pteris*
 (Pteridaceae). *American Fern Journal* **98**: 33--41.

### 2020-02-21 ms/manuscript.Rmd
Update MS

@@
   words = c(
-    "Total word count (Introduction, Materials and Methods, Results, Discussion):_6499",
+    "Total word count (Introduction, Materials and Methods, Results, Discussion):_6601",
     "Summary:_199",
     "Introduction:_881",
-    "Materials and Methods:_2410",
-    "Results:_798",
-    "Discussion:_2410",
+    "Materials and Methods:_2491",
+    "Results:_822",
+    "Discussion:_2407",
     "Acknowledgements:_170"),
@@
     "No. of tables:_5",
-    "No. of supporting information files:_8 (Figs. S1-S2; Tables S1-S6)", "", "", "", "")
+    "No. of supporting information files:_10 (Figs. S1-S2; Tables S1-S8)", "", "", "", "")
 ) %>%
@@
 
-```{r physig-numbers}
+```{r physig-numbers, include=FALSE}
+# Extract D-values
 habit_d <- phylosig_binary_traits %>% filter(trait == "habit") %>% pull(D) %>% round_t(2)
@@
 morphotype_d <- phylosig_binary_traits %>% filter(trait == "morphotype") %>% pull(D) %>% round_t(2)
+
+# Verify D is negative for all binary traits
+assert(phylosig_binary_traits, within_bounds(-Inf, 0), D, success_fun = success_logical)
 ```
@@
 However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of \K{ }< 1 (`r table("phylosig-cont")`).
-All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D{ }= `r hairs_d`) or more conserved (gemmae, \D{ }= `r gemmae_d`; glands, \D{ }= `r glands_d`; morphotype, \D{ }= `r morphotype_d`) than expected under BM (`r table("phylosig-binary")`).
-Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D{ }= `r habit_d`; `r table("phylosig-binary")`).
+All binary gametophyte traits showed phylogenetic signal more conserved than expected under BM (`r table("phylosig-binary")`).
+Growth habit showed phylogenetic signal similar to that expected under BM (\D{ }= `r habit_d`; `r table("phylosig-binary")`).
 Similar results were obtained using the 'strict' dataset (`r s_table("phylosig-binary-strict")`).
@@
 Most of the best-fitting models included growth habit or the interaction of growth habit and a temperature-related variable (`r s_table("gam-fit")`).
-Spatial autocorrelation was not detected in any of the best-fitting full-subsets models (`r s_table("gam-fit")`).
+Spatial autocorrelation was not detected in any of the best-fitting general additive models (`r s_table("gam-fit")`).
 
 Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
-Spatial autocorrelation was detected in the linear model for SLA (`r s_table("linear-fit")`), so a spatial generalized mixed model was used instead (see Methods). 
+Spatial autocorrelation was detected in the linear model for SLA, so a spatial generalized mixed model was used instead (see Methods).
 Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (spatial generalized linear mixed model, nu = `r sla_nu`, rho = `r sla_rho`, AIC = `r sla_AIC`; `r s_table("spatial-sla")`).
 Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }`r get_mod_result(model_summ_print, "dissection", "p")`).
-Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r figure("cwm-div")`).
+Rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r figure("cwm-div")`).
 
@@
 
-Finally, the presence of hairs and glands in gametophytes also exhibit very high phylogenetic signal (\ie, much greater than expected under Brownian motion; `r table("phylosig-binary")`), and these traits are concentrated in certain taxonomic groups [\eg, hairs, grammitid ferns (Grammitidoideae); glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
+Finally, the presence of hairs and glands in gametophytes also exhibit high phylogenetic signal (\ie, greater than expected under Brownian motion; `r table("phylosig-binary")`), and these traits are concentrated in certain taxonomic groups [\eg, hairs, grammitid ferns (Grammitidoideae); glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
 Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], and glands may affect rates of osmosis by modifying lipid concentrations [@Crow2011], we did not detect a correlation between hair or gland production and epiphytic growth.
@@
 Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown.
-Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval{ }< 0.05.
+Trendlines fit using linear models (see Methods); all trendlines significant at \pval{ }< 0.05.
 Asterisks indicate statistical significance of linear model; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01.
@@
 Error bars are each one standard deviation (values below zero not shown).
-Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at \pval{ }< 0.05 (\rval^2^ not shown for SLA, which was fit with a spatial generalized linear mixed model; see Methods).
+Trendlines and \rval^2^ shown for significant relationships as determined by a linear model at \pval{ }< 0.05 (\rval^2^ not shown for SLA, which was fit with a spatial generalized linear mixed model; see Methods).
 \tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\* = \pval{ }< 0.001.
@@
 Response variable abbreviations as follows: MPD~phy~, standard effect size (SES) of mean phylogenetic distance; MNTD~phy~, SES of mean nearest taxon distance; MPD~func~, SES of mean functional distance; MNTD~func~, SES of mean nearest functional distance.
-Trendlines indicate significant relationships as determined by a linear model at \pval{ }< 0.05.
+Trendlines and \rval^2^ shown for significant relationships as determined by a linear model at \pval{ }< 0.05.
 Boxplots show median values (bold lines) by growth habit.

### 2020-02-21 ms/captions.R

@@ -31,7 +31,7 @@ s_table_nums(name = "ancova", "Analysis of covariance (ANCOVA) for differences b
 s_table_nums(name = "linear-fit", "Best-fitting linear models explaining climate and community diversity metrics of ferns on Moorea, French Polynesia by elevation and growth habit")
 s_table_nums(name = "phylosig-binary-strict", "Phylogenetic signal in binary traits of ferns on Moorea, French Polynesia, strict dataset excluding any species whose traits were scored based on taxonomy")
 s_table_nums(name = "corr-evo-strict", "Pagel's (1994) test of correlated evolution between binary traits of ferns on Moorea, French Polynesia, strict dataset excluding any species whose traits were scored based on taxonomy")
-s_table_nums(name = "gam-fit", "Best-fitting models explaining community diversity metrics of ferns on Moorea, French Polynesia by climate and growth habit")
+s_table_nums(name = "gam-fit", "Best-fitting general additive models explaining community diversity metrics of ferns on Moorea, French Polynesia by climate and growth habit selected using full-subsets analysis")
 s_table_nums(name = "spatial-sla", "Output of spatial generalized linear mixed model for community-weighted mean specific leaf area of ferns on Moorea, French Polynesia predicted by elevation and growth habit")
 
 # Make short versions of citation functions

### 2020-02-21 ms/manuscript.Rmd
Update MS

@@
   words = c(
-    "Total word count (Introduction, Materials and Methods, Results, Discussion):_6601",
+    "Total word count (Introduction, Materials and Methods, Results, Discussion):_6603",
     "Summary:_199",
@@
     "Materials and Methods:_2491",
-    "Results:_822",
+    "Results:_824",
     "Discussion:_2407",
@@
 
-Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
+Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter), frond dissection, and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
 Spatial autocorrelation was detected in the linear model for SLA, so a spatial generalized mixed model was used instead (see Methods).

### 2020-05-06 ms/manuscript.Rmd
Update MS to match corrections to proof

@@
 ---
-title: "Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes"
+title: "Life in the canopy: community trait assessments reveal substantial functional diversity among fern epiphytes"
 bibliography: 
@@
 
-^1^Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, Cambridge, MA 02138, U.S.A.; ^2^Department of Biology, Colgate University, 13 Oak Drive, Hamilton, New York 13346, U.S.A.; ^3^Current address: Department of Botany, National Museum of Natural History, Smithsonian Institution, Washington, D.C., 20013, U.S.A.
+^1^Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, 22 Divinity Avenue, Cambridge, MA 02138, U.S.A.; ^2^Department of Biology, Colgate University, 13 Oak Drive, Hamilton, NY 13346, U.S.A.; ^3^Current address: Department of Botany, National Museum of Natural History, Smithsonian Institution, PO Box 37012, Washington, D.C.
+20013, U.S.A.
 
@@
 <!-- FIXME: add borders to table in docx -->
+<!-- NOTE: the word count matches when the final, revised MS was submitted. It does not match changes made to the proof. -->
 
@@
 
-- While epiphytes showed convergence in small size and a higher frequency of non-cordate gametophytes, they showed greater functional diversity at the community level relative to terrestrial ferns.
+- While epiphytes showed convergence in small size and a higher frequency of noncordate gametophytes, they showed greater functional diversity at the community level relative to terrestrial ferns.
 
@@
 The appearance of angiosperm-dominated forests in the late Cretaceous and early Cenozoic fundamentally changed climates and had an immense effect on terrestrial ecosystems, including increased global precipitation and the creation of novel niches [@Niklas1983; @Kenrick1997a; @Lloyd2008; @Boyce2010].
-The heterogeneous niche space of the angiosperm canopy comprising stratified light and humidity levels and a range of substrates for growth became rapidly filled by a variety of epiphytic plants, including mosses [@Fiz-Palacios2011], leafy liverworts [@Feldberg2014], orchids [@Givnish2015b], bromeliads [@Givnish2014a], and ferns [@Schneider2004d].
-The transition to epiphytic growth was particularly important for modern fern species diversity: \circa{ }29% of all fern species are epiphytes vs.
-\circa{ }10% across all vascular plants [@Kress1986].
+The heterogeneous niche space of the angiosperm canopy comprising stratified light intensities and degrees of humidity and a range of substrates for growth became rapidly filled by a variety of epiphytic plants, including mosses [@Fiz-Palacios2011], leafy liverworts [@Feldberg2014], orchids [@Givnish2015b], bromeliads [@Givnish2014a], and ferns [@Schneider2004d].
+The transition to epiphytic growth was particularly important for modern fern species diversity: \circa{ }29% of all fern species are epiphytes vs \circa{ }10% across all vascular plants [@Kress1986].
 Moreover, major transitions to epiphytism occurred at least five times within ferns, each followed by diversification during the late Cretaceous or early Cenozoic [@Schuettpelz2009; @WatkinsJr.2012].
@@
 Conversely, the stratification of the canopy may stimulate functional diversification by niche partitioning, leading to greater functional diversity within epiphytic communities. 
-Furthermore, the role of the canopy as a promoter or inhibitor of community diversity may be modulated by overall abiotic conditions, including most notably relative humidity (RH) and temperature. 
+Furthermore, the role of the canopy as a promoter or inhibitor of community diversity may be modulated by overall abiotic conditions, including relative humidity (RH) and temperature. 
 For instance, elevation is a strong predictor of fern species richness, with maximal richness occurring at mid-elevations, especially in the tropics [@Kromer2005; @Kessler2011a]. 
@@
 Epiphytic ferns tend to have morphological features that minimize water loss from their leaves, including shorter stipes [@WatkinsJr.2010], thicker leaves [@Kluge2007; @WatkinsJr.2007b], and protection with scales or hairs [@WatkinsJr.2006b].
-Additional specialized adaptations to epiphytic growth include humus collection (\eg, *Asplenium*; @Benzing1990, *Drynaria*; @Janssen2005), vegetative desiccation tolerance (\eg, Hymenophyllaceae; @Shreve1911, *Pleopeltis*; @Pessin1925), CAM photosynthesis (\eg, *Pyrrosia*; @Wong1976), and symbiotic relationships with ants (\eg, *Antrophyum*; @WatkinsJr.2008, *Lecanopteris*; @Gay1993a).
+Additional specialized adaptations to epiphytic growth include humus collection (\eg{ }*Asplenium*; @Benzing1990, *Drynaria*; @Janssen2005), vegetative desiccation tolerance (\eg Hymenophyllaceae; @Shreve1911, *Pleopeltis*; @Pessin1925), Crassulacean acid metabolism (\eg{ }*Pyrrosia*; @Wong1976), and symbiotic relationships with ants (\eg{ }*Antrophyum*; @WatkinsJr.2008, *Lecanopteris*; @Gay1993a).
 
@@
 In addition, epiphytic fern gametophytes tend to exhibit complex three-dimensional morphologies that may function to retain water and slow its loss [@WatkinsJr.2007c].
-Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no previous studies to our knowledge have analyzed the functional ecology of fern epiphytes in a broad phylogenetic context including both life stages.
+Despite evidence for adaptations to epiphytism in multiple lineages in both sporophytes and gametophytes, no previous studies have, to our knowledge, analyzed the functional ecology of fern epiphytes in a broad phylogenetic context including both life stages.
 
@@
 Moorea is an isolated tropical oceanic island with high phylogenetic diversity of ferns distributed across a steep elevational gradient, making it ideal for a community phylogenetic analysis of functional traits.
-Specifically, we test the following hypotheses related to epiphytic growth in ferns: 1) Epiphytic ferns have convergent morphologies in both sporophyte and gametophyte phases; 2) Epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities.
+Specifically, we test the following hypotheses related to epiphytic growth in ferns: 1) epiphytic ferns have convergent morphologies in both sporophyte and gametophyte phases; 2) epiphytic fern communities have lower functional and phylogenetic diversity relative to terrestrial communities.
 
@@
 
-Moorea, French Polynesia (17°32'S, 149°50'W) is a small (\SI{135}{\kilo\meter}^2^) tropical oceanic island more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
+Moorea, French Polynesia (17°32'S, 149°50'W) is a small (\SI{135}{\kilo\meter}^2^) tropical oceanic island located more than \SI{5000}{\kilo\meter} from the nearest continental landmass.
 This distance acts as a strong barrier to dispersal [@Carson1995; @Dassler2001], which combined with the young age of the island [\circa{ }\SI{1.5}{\ma}; @Duncan1976] has led to a relatively small, yet phylogenetically diverse fern flora [\circa{ }130 \species, 8/11 orders; `r s_table("species-list")`; taxonomy follows @PteridophytePhylogenyGroupI2016].
-Ferns of Moorea include representatives of multiple Cretaceous epiphytic radiations [\eg, *Asplenium*, *Elaphoglossum*, Hymenophylloideae, Polypodiaceae, Vittarioideae; @Schuettpelz2009] and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
+Ferns of Moorea include representatives of multiple Cretaceous epiphytic radiations [\eg{ }*Asplenium*, *Elaphoglossum*, Hymenophylloideae, Polypodiaceae, Vittarioideae; @Schuettpelz2009] and their terrestrial relatives, making them suitable for a comparative study of traits related to the evolution of epiphytic growth in ferns.
 
@@
 
-A fern community survey was carried out on Moorea comprising 17, \SI{10}{\meter} × \SI{10}{\meter} plots spanning an elevational gradient from \SI{20}{\meter} to \SI{2000}{\meter}, with plots placed \circa every \SI{200}{\meter} [@Nitta2017].
-Most plots were located along trails leading to the three main peaks on the island, Mt. Rotui (\SI{899}{\meter}), Mt. Mouaputa (\SI{890}{\meter}), and Mt. Tohiea (\SI{1206}{\meter}).
-Each plot was divided into 25, \SI{2}{\meter} × \SI{2}{\meter} subplots.
+A fern community survey was carried out on Moorea comprising 17 plots (\SI{10}{\meter} × \SI{10}{\meter}) spanning an elevational gradient from \circa{ }200 to \SI{1200}{\meter}, with plots placed \circa every \SI{200}{\meter} [@Nitta2017].
+Most plots were located along trails leading to the three main peaks on the island, Mt. Rotui (\SI{899}{\meter}), Mt. Mouaputa (\SI{830}{\meter}), and Mt. Tohiea (\SI{1206}{\meter}).
+Each plot was divided into 25 subplots (\SI{2}{\meter} × \SI{2}{\meter}).
 Presence/absence of all sporophytes in each subplot was scored and summed to produce an abundance ranking for each species per plot from zero (not observed) to 25 (observed in all subplots).
 Epiphytic ferns were sampled by hand to \circa{ }\SI{2}{\meter} on tree trunks or confirmed visually from the ground.
-Species were categorized as epiphytic or terrestrial based on field observations (see 'Trait selection and measurements' below).
+Species were categorized as epiphytic or terrestrial based on field observations (see 'Trait selection and measurements' later).
 The list of species in each plot was split into epiphytic and terrestrial groups and treated separately during data analysis, hereafter referred to as 'epiphytic' and 'terrestrial' communities.
-Vouchers were deposited at GH and UC [abbreviations follow Index Herbariorum; @Thiers2019].
+Vouchers were deposited at UC, with duplicates at GH and PAP [abbreviations follow Index Herbariorum; @Thiers2019].
+All fieldwork was conducted under permits issued by the French Polynesian Government (Délégation à la Recherche) and the Haut-commissariat de la République en Polynésie Francaise (Protocole d'Accueil 2012–2014).
 
@@
     start = min(date) %>% 
-      format('%B %e, %Y') %>%
-      str_replace_all("  ", " "),
+      format("%e %B %Y") %>%
+      str_trim(),
     end = max(date) %>% 
-      format('%B %e, %Y') %>%
-      str_replace_all("  ", " ")
+      format("%e %B %Y") %>%
+      str_trim()
   )
@@
 
-Temperature and RH of epiphytic and terrestrial habitats were measured using Hobo Pro v2 dataloggers with the RS3 Solar Radiation Shield (Onset Corporation, USA).
+Temperature and RH of epiphytic and terrestrial habitats were measured using \textsc{Hobo Pro} v.2 dataloggers with the RS3 Solar Radiation Shield (Onset Corporation, Bourne, MA, USA).
 A pair of dataloggers was installed for each plot, one mounted at \circa{ }\SI{2}{\meter} on a tree ('epiphytic datalogger'), and one at ground level ('terrestrial datalogger'), except for the Mt.
@@
 We used the 2013--2014 data for the analysis, with one exception: the terrestrial datalogger at the Mt.
-Tohiea \SI{393}{\meter} plot malfunctioned and was missing data for a substantial part (March 12--July 5, 2014) of the final survey period, so we used data for the same dates from the preliminary survey period for this datalogger instead.
+Tohiea \SI{393}{\meter} plot malfunctioned and was missing data for a substantial part (12 March 2014 to 5 July 2014) of the final survey period, so we used data for the same dates from the preliminary survey period for this datalogger instead.
 Three others also failed early during the final survey period, but lacked data from the preliminary period, so were excluded from analysis.
 Other days missing data during the final survey period were excluded from the dataset.
-The final dataset included `r total_days` days of data for `r total_sites %>% filter(habit == "epiphytic") %>% pull(n)` epiphytic dataloggers and `r total_sites %>% filter(habit == "terrestrial") %>% pull(n)` terrestrial dataloggers.
-We calculated vapor pressure deficit (VPD) from relative humidity and temperature using the \package{plantecophys} R package [@Duursma2015].
+The final dataset included \SI{`r total_days`}{\day} of data for `r total_sites %>% filter(habit == "epiphytic") %>% pull(n)` epiphytic dataloggers and `r total_sites %>% filter(habit == "terrestrial") %>% pull(n)` terrestrial dataloggers.
+We calculated vapor pressure deficit (VPD) from RH and temperature using the \package{plantecophys} package in R [@Duursma2015].
 
@@
 
-We selected several traits putatively relevant to epiphytic growth (`r table("traits-used")` and `r figure("traits-used-image")`).
-We preferred morphological traits over physiological traits because data for morphological traits are easier to gather at the scale of our study (\circa{ }130 \species distributed over an elevational gradient from \SI{200}{\meter} to \SI{1200}{\meter}).
+We selected several traits putatively relevant to epiphytic growth (`r table("traits-used")`; `r figure("traits-used-image")`).
+We preferred morphological traits over physiological traits because data for morphological traits are easier to gather at the scale of our study (\circa{ }130 \species distributed over an elevational gradient from 200 to \SI{1200}{\meter}).
 Plant morphology and physiology are closely linked, and physiological traits such as rates of photosynthesis and water use efficiency are known to be correlated with morphological traits such as stipe length in ferns [@WatkinsJr.2010].
@@
 Although trait values can vary within fern species with elevation [@Kessler2007a], our study includes a broad phylogenetic sampling, so we assume intraspecific variation is outweighed by interspecies differences.
-To measure specific leaf area (SLA; m^2^ kg^-1^), \circa{ }10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} diam.
-punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
+To measure specific leaf area (SLA; m^2^ kg^-1^), \circa{ }10--12, 2- or 4-mm-diameter punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
 For species with < \SI{2}{\milli\meter} between primary veins, leaf fragments excluding primary veins were obtained by dissection, dried at \SI{60}{\celsius} for \SI{24}{\hour}, weighed, and scanned.
-ImageJ [@Abramoff2004] software was used to measure area per fragment.
+\package{ImageJ} [@Abramoff2004] software was used to measure area per fragment.
 
 Morphological traits of gametophytes were all qualitative (either categorical or binary; `r table("traits-used")`).
-Traits were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017], grown from spores of known species in the lab, taken from the literature, or inferred from taxonomic group (\ie, genus) if documented as a stable taxonomic character.
-For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie, heart-shaped) or non-cordate [@Ebihara2013].
-Non-cordate morphotypes include elongate (strap or ribbon \sensu{ }@Farrar2008) and filamentous forms (`r figure("traits-used-image")`).
+Traits were scored by observing gametophytes that were either collected in the field and identified using DNA barcoding [@Nitta2017] or grown from spores of known species in the laboratory, taken from the literature, or inferred from taxonomic group (\ie genus) if documented as a stable taxonomic character.
+For comparative analysis, gametophyte morphotypes were aggregated into two categories: either cordate (\ie heart-shaped) or noncordate [@Ebihara2013].
+Noncordate morphotypes include elongate (strap or ribbon \sensu{ }@Farrar2008) and filamentous forms (`r figure("traits-used-image")`).
 Gametophytes were cultivated on Bold's media [@Bold1957] supplemented with Nitch's micronutrients [@Nitsch1951].
@@
 Growth habit was coded as a binary trait (epiphytic or terrestrial) based on our field observations.
-Some species do not fall clearly into either category [\eg, hemi-epiphytes; @Benzing1990; @Dubuisson2003; @Zotz2013].
+Some species do not fall clearly into either category [\eg hemi-epiphytes; @Benzing1990; @Dubuisson2003a; @Zotz2013].
 We used connection to the soil as the criterion for defining binary growth habit; thus, epipetric plants were treated as epiphytic, and hemi-epiphytes and climbing plants were treated as terrestrial. 
@@
 
-We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each datalogger, then overall means from the daily values.
-We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (a single variable was arbitrarily selected from each set of correlated variables; hereafter, 'climatic variables').
+We calculated daily maximum, mean, minimum, and standard deviation (SD) of temperature, RH, and VPD for each datalogger, then overall means from the daily values.
+We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained those with correlation coefficients < 0.9 (a single variable was arbitrarily selected from each set of correlated variables; hereafter, 'climatic variables').
 Variables relating to VPD, which have no meaning if less than zero, and some other variables that had skewed distributions, were square-root transformed prior to fitting models.
@@
 We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone, in combination, and in combination with their interactions.
-We then calculated the corrected Akaike Information Criterion [AIC~c~; @Akaike1973] for each model and retained models within 2 AIC~c~ of the lowest (\ie, 2 delta AIC~c~).
-From these, we selected the model with the combination of the fewest non-significant parameters and lowest AIC~c~ (for VPD, we the selected the model producing the smallest amount of negative predicted values).
+We then calculated the corrected Akaike Information Criterion [AICc; @Akaike1973] for each model and retained models within 2 AICc of the lowest (\ie 2ΔAICc). 
+From these, we selected the model with the combination of the fewest non-significant parameters and lowest AICc (for VPD, we selected the model producing the fewest negative predicted values).
 A single outlier plot with unusually high temperature and low humidity (Mt. Rotui, 830 m plot) was excluded from ANCOVA and linear models.
@@
 For comparative analyses, we used the dated plastid phylogenetic tree of @Nitta2017.
-*Microsorum maximum* (Brack.) Copel., a genetically verified hybrid between *M.
+*Microsorum* × *maximum* (Brack.) Copel., a genetically verified hybrid between *M.
 grossum* (Langsd.
@@
 Andrews and *M.
-commutatum* Copel.
-[@Nitta2018] was excluded from analysis because it shares identical plastid sequences with *M.
-grossum*, so the plastid tree does not accurately represent its evolutionary distinctiveness.
-
-We tested for phylogenetic signal in quantitative traits using Blomberg's \K{ }[@Blomberg2003] and Pagel's \plam{ }[@Pagel1999], two metrics with slightly different interpretations and sensitivity to different evolutionary scenarios [@Munkemuller2012], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
+commutatum* (Blume) Copel.
+was excluded from analysis because it shares identical plastid sequences with *M.
+grossum* [@Nitta2018], so the plastid tree does not accurately represent its evolutionary distinctiveness.
+
+We tested for phylogenetic signal in quantitative traits using Blomberg's \K{ }[@Blomberg2003] and Pagel's \plam{ }[@Pagel1999], two metrics with slightly different interpretations and sensitivity to different evolutionary scenarios [@Munkemuller2012], with the \function{phylosig} function in the R package \package{phytools} [@Revell2012].
 \plam is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion (BM)).
-\K describes the ratio between the amount of observed variance in traits vs.
-the amount of variance expected under BM: when \K{ }> 1, traits are more conserved than expected under BM; when \K{ }< 1, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
-We tested for significance in \K by comparing observed values of \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in \plam by conducting a likelihood ratio test comparing the log likelihoods of observed values of \plam vs.
-\plam{ }= 0 (no phylogenetic signal).
-We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' \D{ }[@Fritz2010].
-Values of \D range from 0 under a model of trait evolution by BM, to 1 under random distribution of traits.
-We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], and conducted a significance test by comparing observed \D to distributions of simulated values produced under two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM model.
+\K describes the ratio between the amount of observed variance in traits vs the amount of variance expected under BM: when \K{ }> 1, traits are more conserved than expected under BM; when \K{ }< 1, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
+We tested for significance in \K by comparing observed values of \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in \plam by conducting a likelihood ratio test comparing the log likelihoods of observed values of \plam vs \plam{ }= 0 (no phylogenetic signal).
+We tested for phylogenetic signal in qualitative (binary) traits using Fritz and Purvis' \D{ }[@Fritz2010].
+Values of \D range from 0 under trait evolution by BM to 1 under random distribution of traits, and can exceed these values in cases of extreme clumping (\D{ }< 0) or overdispersion (\D{ }> 1).
+We analyzed \D using the \function{phylo.d} function in the R package \package{caper} [@Orme2013], and conducted a significance test by comparing observed \D to distributions of simulated values produced under two scenarios: random shuffling of traits on the tree, or simulation of a binary trait under a BM model.
 To account for possible circularity in measuring the phylogenetic signal of traits that were scored based on taxonomy, we conducted the same analyses on a dataset excluding species that had been scored this way ('strict' dataset).
@@
 
-We detected significant phylogenetic signal in most traits, so we tested for differences between traits of epiphytic vs. terrestrial species using phylogenetic comparative methods.
+We detected significant phylogenetic signal in most traits, so we tested for differences between traits of epiphytic vs terrestrial species using phylogenetic comparative methods.
 For quantitative traits, we used the \function{brunch} function in the \package{caper} package [@Orme2013], which calculates phylogenetically independent contrasts in a quantitative character between alternative states of a binary trait.
@@
 We used only quantitative traits, and included only species that had no missing data (\n{ }= `r quant_traits_nsp` \species).
-We first log-transformed traits as necessary to reduce skew, then scaled each trait by its standard deviation.
-We conducted standard and phylogenetically corrected PCA, using the \function{PCA} and \function{phyl.pca} functions in the \package{FactoMineR} [@Le2008] and \package{phytools} [@Revell2012] R packages, respectively.
-
-We calculated community-weighted means (CWMs) and standard deviations of quantitative traits by using the overall mean trait value of each species weighted by their abundances in each community.
+We first log-transformed traits as necessary to reduce skew, then scaled each trait by its SD.
+We conducted standard and phylogenetically corrected PCA, using the \function{PCA} and \function{phyl.pca} functions in the R packages \package{FactoMineR} [@Le2008] and \package{phytools} [@Revell2012], respectively.
+
+We calculated community-weighted means and SDs of quantitative traits using the mean trait value of each species weighted by their abundances in each community.
 
@@
 
-Raw values of MPD and MNTD are statistically related to species richness [@Swenson2014]; to obtain metrics that are uncorrelated with species richness, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{999} randomized communities.
+Raw values of MPD and MNTD are statistically related to species richness [@Swenson2014]; to obtain metrics that are uncorrelated with species richness, we calculated their standard effect sizes (SES) by comparing observed values with a distribution of 999 randomized communities.
 We constructed null communities using the 'independent swap' algorithm [@Gotelli2000].
@@
 We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate phylogenetic and functional diversity metrics.
-Hereafter, 'MPD~phy~' and 'MNTD~phy~', and 'MPD~func~' and 'MNTD~func~', refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.
-
-We tested the effects of climate and growth habit on diversity metrics using full-subsets multiple regression with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
-For each diversity metric, we built a set of models (Gaussian family, with plot as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
+Hereafter, MPD~phy~ and MNTD~phy~, and MPD~func~ and MNTD~func~, refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.
+
+We tested the effects of climate and growth habit on diversity metrics using full-subsets multiple regression with general additive models (GAMs) as implemented in the R package \package{FSSgamm} [@Fisher2018].
+For each diversity metric, we built a set of models (Gaussian family, with plot as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AICc to find the models that best explain the data.
 Our dataset included `r (nrow(div_metrics_all) - nrow(climate_select)) %>% english2()` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
-Therefore, we also constructed linear models for each diversity metric in response to elevation and growth habit including all sites using the same model selection procedure as in the climate analysis.
-We checked for spatial autocorrelation in the model residuals with Moran's *I* [@Moran1950] as implemented in the \package{spdep} R package [@Bivand2013]. 
-If spatial autocorrelation was detected, a generalized linear mixed model was fit with a Mátern spatial correlation structure using the \package{spAMM} R package [@Rousset2014].
+Therefore, we also constructed linear models for each diversity metric in response to elevation and growth habit, including all sites using the same model selection procedure as in the climate analysis.
+We checked for spatial autocorrelation in the model residuals with Moran's *I* [@Moran1950] as implemented in the R package \package{spdep} [@Bivand2013]. 
+If spatial autocorrelation was detected, a generalized linear mixed model was fit with a Mátern spatial correlation structure using the R package \package{spAMM} [@Rousset2014].
 
@@
 
-Mean, maximum, and standard deviation (SD) of temperature and mean, minimum, and SD of vapor pressure deficit (VPD) were retained after excluding variables with correlation coefficients > 0.9.
+Mean, maximum, and SD of temperature and mean, minimum, and SD of VPD were retained after excluding variables with correlation coefficients > 0.9.
 Variables related to temperature changed with elevation but did not differ significantly between epiphytic and terrestrial dataloggers (`r s_table("ancova")`; `r figure("climate")`).
@@
 
-Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (`r figure("pca")`**c**, **d**).
+Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (`r figure("pca")`c,d).
 The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
@@
 
-Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs{ }`r table_nums("phylosig-cont", display = "n")`, `r table_nums("phylosig-binary", display = "n")`, `r figure("tree")`), and for quantitative traits, different results were obtained for \plam and \K.
+Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs{ }`r table_nums("phylosig-cont", display = "n")`, `r table_nums("phylosig-binary", display = "n")`; `r figure("tree")`), and for quantitative traits, different results were obtained for \plam and \K.
 When measured with \plam, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of \plam close to 1; only number of pinnae pairs had \plam close to zero (`r table("phylosig-cont")`).
@@
 Epiphytic species tend to have smaller values for all traits (`r table("pics")`).
-Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval{ }`r stipe_pval`), as well as frond width (\pval{ }`r width_pval`) and degree of laminae dissection (\pval{ }`r dissection_pval`).
-For the gametophyte traits, only morphotype (non-cordate vs.
-cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval{ }`r morphotype_pval`; `r table("corr-evo")`).
+Short stipe length, in particular, appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval{ }`r stipe_pval`), as well as frond width (\pval{ }`r width_pval`) and degree of laminae dissection (\pval{ }`r dissection_pval`).
+For the gametophyte traits, only morphotype (noncordate vs cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval{ }`r morphotype_pval`; `r table("corr-evo")`).
 Similar results were obtained using the 'strict' dataset (`r s_table("corr-evo-strict")`).
@@
 
-Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter), frond dissection, and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
-Spatial autocorrelation was detected in the linear model for SLA, so a spatial generalized mixed model was used instead (see Methods).
-Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (spatial generalized linear mixed model, nu = `r sla_nu`, rho = `r sla_rho`, AIC = `r sla_AIC`; `r s_table("spatial-sla")`).
+Community-weighted mean values were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter), frond dissection, and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
+Spatial autocorrelation was detected in the linear model for SLA, so a spatial generalized mixed model was used instead (see Materials and Methods).
+SLA decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (spatial generalized linear mixed model, nu = `r sla_nu`, rho = `r sla_rho`, AIC = `r sla_AIC`; `r s_table("spatial-sla")`).
 Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }`r get_mod_result(model_summ_print, "dissection", "p")`).
@@
 We observed a total of `r abundances %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` \species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
-Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r figure("comm-div")`).
-Species richness increased with elevation for both epiphytes and terrestrial communities (`r figure("comm-div")`).
+Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytic and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial communities (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r figure("comm-div")`).
+Species richness increased with elevation for both epiphytic and terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "ntaxa", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "ntaxa", "p")`).
 Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, `r both`\pval{ }`r p_mpd_func_mntd_func`).
 Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval{ }`r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
-Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval{ }`r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval{ }= `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; `r figure("comm-div")`).
+Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval{ }`r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval{ }= `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`).
 No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (`r figure("comm-div")`).
@@
 Our study reveals a striking duality in the phylogenetic and functional diversity of fern epiphytes.
-On one hand, we find that fern epiphytes are consistently smaller than terrestrial species at both the species and community level (`r table("pics")`, `r figure("cwm-div")`), and that epiphytic communities are phylogenetically clustered (`r figure("comm-div")`).
+On one hand, we find that fern epiphytes are consistently smaller than their terrestrial counterparts at both the species and community levels (`r table("pics")`; `r figure("cwm-div")`), and that epiphytic communities are phylogenetically clustered (`r figure("comm-div")`).
 At the same time, we find that epiphytic fern communities are significantly more diverse functionally compared to terrestrial communities (`r figure("comm-div")`).
@@
 
-Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community level (`r table("pics")`, `r figure("cwm-div")`), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
+Epiphytes are markedly smaller than terrestrial species as measured by stipe and frond length, frond width, and rhizome diameter at both the species and community levels (`r table("pics")`; `r figure("cwm-div")`), supporting both hypotheses (convergence in epiphytic traits and reduced functional diversity of epiphytic communities).
 These results are corroborated by @Testo2018a, who similarly observed abrupt size reduction associated with epiphytic growth in a globally sampled fern phylogeny. 
@@
 In spite of similar and predictable leaf-level economic tradeoffs between angiosperms and ferns [@Karst2007], fern leaves are limited in their hydraulic flexibility as they rely wholly on primary xylem that feeds into laminar tissues with reduced leaf venation.
-Stipe length and structure in particular are critical for regulating whole leaf hydraulic conductance in ferns [@Pittermann2013].
+Stipe length and structure, in particular, are critical for regulating whole leaf hydraulic conductance in ferns [@Pittermann2013].
 Species growing in water-limited habitats like those in the canopy are expected by biomechanical necessity to decrease stipe length to minimize resistance [@WatkinsJr.2010].
@@
 
-Another clear pattern to emerge from our study supporting the hypothesis of convergence is the association between epiphytism and the prevalence of non-cordate (\eg, elongate or filamentous) gametophyte morphotypes.
-Gametophyte morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived (< \SI{1}{\year}), cordate morphotypes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have non-cordate gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; @Farrar2008].
-The complex, three-dimensional structure of non-cordate gametophytes may serve two important functions. 
-Complex thallus morphology can slow drying rates, and slow dehydration is correlated with the ability to recover from drought in gametophytes---likely a critical adaptation to drought-prone epiphytic habitats [@WatkinsJr.2007c; @Pittermann2013].
-Non-cordate thallus morphology may also promote outcrossing through the formation of long-lived gametophyte banks that can persist in epiphytic but not in more disturbance-prone terrestrial habitats [@WatkinsJr.2007a; @Farrar2008].
-Our study is the first to our knowledge to demonstrate that the correlation between non-cordate morphology and epiphytism in ferns is significant across broad phylogenetic scales.
-
-### Unexpected diversity in non-size related traits of epiphytes
-
-Despite the apparent constraints on size and the recurrence of non-cordate gametophytes exhibited by epiphytes, we observed greater morphological diversity among epiphytic versus terrestrial communities for several traits.
+Another clear pattern to emerge from our study supporting the hypothesis of convergence is the association between epiphytism and the prevalence of noncordate (\eg elongate or filamentous) gametophyte morphotypes.
+Gametophyte morphotype was previously observed to correlate with life history and habitat in ferns: terrestrial species tend to have short-lived (< \SI{1}{\year}), cordate morphotypes that establish following disturbance and rapidly produce sporophytes, whereas epiphytic species tend to have noncordate gametophytes that persist over multiple growing seasons and produce sporophytes more slowly [@WatkinsJr.2007a; @Farrar2008].
+The complex, three-dimensional structure of noncordate gametophytes may serve two important functions. 
+Complex thallus morphology can slow drying rates, and slow dehydration is correlated with the ability to recover from drought in gametophytes---probably a critical adaptation to drought-prone epiphytic habitats [@WatkinsJr.2007c; @Pittermann2013].
+Noncordate thallus morphology may also promote outcrossing through the formation of long-lived gametophyte banks that can persist in epiphytic but not in more disturbance-prone terrestrial habitats [@WatkinsJr.2007a; @Farrar2008].
+To our knowledge, our study is the first to demonstrate that the correlation between noncordate morphology and epiphytism in ferns is significant across broad phylogenetic scales.
+
+### Unexpected diversity in traits of epiphytes other than size
+
+Despite the apparent constraints on size and the recurrence of noncordate gametophytes exhibited by epiphytes, we observed greater morphological diversity among epiphytic vs terrestrial communities for several traits.
 These results do not support the hypothesis that fern epiphytes have reduced functional diversity.
@@
 For instance, filmy ferns (Hymenophyllaceae), which depend on desiccation tolerance to survive dry conditions, have extremely thin, highly dissected fronds.
-Bird's nest ferns (\eg, *Asplenium nidus* L.), in contrast, avoid desiccation by capturing detritus and water in the bases of their simple, rosette leaves. 
+Bird's nest ferns (\eg{ }*Asplenium nidus* L.), in contrast, avoid desiccation by capturing detritus and water in the bases of their simple, rosette leaves. 
 At low-elevation sites on Moorea where these ferns co-occur, filmy ferns frequently grow in more humid portions of trees including near the main trunk, whereas bird's nest ferns often grow higher in the canopy and on more exposed, drought-prone branches (J.
@@
 Similar predictable patterns of niche partitioning at the local scale have been reported in neotropical epiphtyic ferns as well [@Hietz1998; @Tausz2001]. 
-This distribution may reflect alternative water-usage strategies aligned to specific microhabitats in the canopy [desiccation tolerance vs. water capture; @Hietz1998]. 
+This distribution may reflect alternative water-usage strategies aligned to specific microhabitats in the canopy [desiccation tolerance vs water capture; @Hietz1998]. 
 Such hypotheses need to be carefully tested with additional quantitative investigation, including more finely examining microclimatic variation and species composition across the canopy.
 
-Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA; `r figure("cwm-div")`, `r table("pics")`).
-Specific leaf area is part of the leaf economic spectrum (LES), which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs.
-slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a; @Diaz2015].
+Another trait that defied simple classification into epiphytic or terrestrial groups was SLA (`r figure("cwm-div")`; `r table("pics")`).
+SLA is part of the leaf economic spectrum, which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans; and slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a; @Diaz2015].
 We expected that epiphytes would exhibit lower SLA, given their more extreme growing conditions.
-In contrast, we observed overlapping, wide ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes when taking elevation into account (`r figure("cwm-div")`).
+By contrast, we observed overlapping, wide-ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes when taking elevation into account (`r figure("cwm-div")`).
 This may reflect adaptations to specific niches that are not captured by a binary categorization of growth habit.
-For example, the following species ranked among those with the five lowest SLA values (\ie, thick, well-protected leaves) on Moorea: *Davallia anderssonii* Mett.
+For example, the following species ranked among those with the five lowest SLA values (\ie thick, well-protected leaves) on Moorea: *Davallia anderssonii* Mett.
 (a twig epiphyte), *Acrostichum aureum* L.
-(a terrestrial mangrove specialist), and *Paesia divaticatissima* (Dryand.) Copel.
+(a terrestrial mangrove specialist), and *Paesia divaricatissima* (Dryand.) Copel.
 (a terrestrial species occurring on high-elevation ridges) (`r s_table("trait-values")`).
@@
 Rather, occurrence of gemmae strongly reflects phylogeny, appearing concentrated in filmy ferns (Hymenophyllaceae) and vittarioid ferns (Vittarioideae) (`r figure("tree")`).
-Gemmae are asexual propagules occurring primarily in non-cordate gametophytes that allow gametophytes to persist over multiple growing seasons and attain large population sizes, and have been hypothesized to be associated with harsh environmental conditions [@Farrar2008].
+Gemmae are asexual propagules occurring primarily in noncordate gametophytes that allow gametophytes to persist over multiple growing seasons and attain large population sizes, and have been hypothesized to be associated with harsh environmental conditions [@Farrar2008].
 In particular, gemmae production should be useful in an epiphytic context because they may allow gametophytes to increase their boundary layer and hedge against population loss during extreme drying [@Farrar2008].
 Gemmae can also function sexually by producing antheridia in the presence of antheridiogen secreted by other gametophytes [@Emigh1977], which may confer a selective advantage in the epiphytic environment by enabling outcrossing [@Dassler2001].
-As with our results summarized above, it is possible that the diversity of gemmae and non-gemmae producing epiphytic species reflects undetected niche variation in the canopy.
-Interestingly, when we classified gametophytes as 'independent' if they occurred beyond the elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with independent growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval{ }= `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
-
-Finally, the presence of hairs and glands in gametophytes also exhibit high phylogenetic signal (\ie, greater than expected under Brownian motion; `r table("phylosig-binary")`), and these traits are concentrated in certain taxonomic groups [\eg, hairs, grammitid ferns (Grammitidoideae); glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
+As discussed earlier, it is possible that the diversity of gemmae- and nongemmae-producing epiphytic species reflects undetected niche variation in the canopy.
+Interestingly, when we classified gametophytes as 'independent' if they occurred beyond the elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with independent growth (Pagel's test of correlated evolution, log-likelihood of dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log-likelihood of independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval{ }= `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`), echoing a pattern observed in Japan of noncordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].
+
+Finally, the presence of hairs and glands in gametophytes also exhibit high phylogenetic signal (\ie greater than expected under BM; `r table("phylosig-binary")`), and these traits are concentrated in certain taxonomic groups [\eg hairs, grammitid ferns (Grammitidoideae); glands, Thelypteridaceae; @Stokey1951; @Nayar1971].
 Although hairs have been hypothesized to reduce rates of drying and increase water uptake after desiccation [@Kappen2007], and glands may affect rates of osmosis by modifying lipid concentrations [@Crow2011], we did not detect a correlation between hair or gland production and epiphytic growth.
 Given their high phylogenetic signal, it is possible that these traits are unrelated to environment but rather reflect phylogenetically conserved developmental pathways [@Johnson2012].
-Alternatively, the possibility remains that some epiphytic species rely on hairs or glands to prevent or slow drying depending on their niche within the canopy (\eg, hair-bearing grammitid gametophytes at mid to high elevation).
+Alternatively, the possibility remains that some epiphytic species rely on hairs or glands to prevent or slow drying depending on their niche within the canopy (\eg hair-bearing grammitid gametophytes at mid to high elevation).
 
@@
 However, while the canopy clearly acts as a filter, we also find that it simultaneously functions as a promoter of functional diversity. 
-This is evident in our analyses of MPD~func~ and MNTD~func~, which both indicate that epiphytic communities are more functionally diverse compared to terrestrial species (`r figure("comm-div")`), thus contradicting our hypothesis of decreased functional diversity in epiphytic communities.
-Additional analyses using different trait combinations support the role of non-size related traits driving this pattern (`r s_figure("corr-plot")`).
+This is evident in our analyses of MPD~func~ and MNTD~func~, which both indicate that epiphytic communities are more functionally diverse compared to terrestrial ones (`r figure("comm-div")`), thus contradicting our hypothesis of decreased functional diversity in epiphytic communities.
+Additional analyses using different trait combinations support the role of nonsize-related traits driving this pattern (`r s_figure("corr-plot")`).
 We propose that the high functional diversity of epiphytic fern communities is maintained by previously undetected niche partitioning within epiphytic habitats as follows.
@@
 Thus, when analyzed together, epiphytic fern communities encompass a wider range of morphological variation than terrestrial communities on average.
-Assuming that the traits we measured reflect physiological strategies of epiphytes, this high morphological diversity likely derives from physiological adaptations to a variety of within-canopy niches.
+Assuming that the traits we measured reflect physiological strategies of epiphytes, this high morphological diversity probably derives from physiological adaptations to a variety of within-canopy niches.
 Clearly, fine-grained characterization of canopy niches and associated adaptations of epiphytic ferns occupying them is an important goal for future research.
@@
 The primary role of growth habit, in combination with temperature, in structuring fern communities is supported by the full-subsets model analysis (`r figure("heatmap")`).
-It is somewhat puzzling that VPD did not contribute significantly to any models, as we hypothesized that this would be important for epiphytes that are frequently subjected to water stress. 
+It is somewhat puzzling that VPD did not contribute to hardly any of the best-fitting models (`r s_table("gam-fit")`), as we hypothesized that this would be important for epiphytes that are frequently subjected to water stress.
 Previous studies have shown marked variation in air water potential across epiphytic habitats, with @WatkinsJr.2009 reporting a \circa{ }\SI{40}{\mega\pascal} shift along the trunks of emergent canopy trees.
 Our epiphytic dataloggers only showed slightly greater VPD relative to those on the ground, with this difference largely dissipating at high elevations (`r figure("climate")`).
-One reason for such low observed variation could be due to environmental conditions at our study site and placement of dataloggers: Moorea is a generally wet site overall (average annual rainfall \SI{325}{\centi\meter}), and we placed the epiphytic dataloggers on tree trunks, not in the higher emergent portions of the canopy.
+Possible reasons for such low observed variation may include environmental conditions at our study site and placement of dataloggers: Moorea is a generally wet site overall (average annual rainfall \SI{325}{\centi\meter}), and we placed the epiphytic dataloggers on tree trunks, not in the higher emergent portions of the canopy.
 Therefore, these measurements should be interpreted as reflecting the minimum amount of difference between terrestrial and canopy environments. 
@@
 
-Recent studies [\eg, @Sundue2015; @Testo2016a; @Lehtonen2017; @Testo2018a] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
+Recent studies [\eg{ }@Sundue2015; @Testo2016a; @Lehtonen2017; @Testo2018a] have called into question the scenario of rapid Cretaceous radiations of epiphytic ferns posited by @Schuettpelz2009 by failing to detect increased rates of diversification associated with epiphytic growth.
 Our study provides a fresh perspective on this issue. 
@@
 Our results suggest new targets for future studies of fern ecology and evolution. 
-We hope future global phylogenetic studies will build on our results by analyzing traits involved in epiphytic growth other than plant size (\eg, gametophyte morphotype, frond dissection, SLA, and hair and gland presence) to dissect how each may relate to diversification and growth habit.
-Additionally, future community ecological studies should focus on determining in more detail how niche space is partitioned within epiphytic vs.
-terrestrial habitats, which is clearly a major dividing line for ferns.
+We hope future global phylogenetic studies will build on our results by analyzing traits involved in epiphytic growth other than plant size (\eg gametophyte morphotype, frond dissection, SLA, and hair and gland presence) to dissect how each may relate to diversification and growth habit.
+Additionally, future community ecological studies should focus on determining in more detail how niche space is partitioned within epiphytic vs terrestrial habitats, which is clearly a major dividing line for ferns.
 
@@
 Suzanne Vinette, Saad Amer, Tristan Wang, and Tohei Theophilus helped with community surveys and measurement of traits.
-Ravahere Taputuarai helped collect environmental data.
+Ravahere Taputuarai helped to collect environmental data.
 Staff at the University of California, Berkeley Richard B.
 Gump South Pacific Research Station, Moorea, French Polynesia, in particular Valentine Brotherson, Hinano Teavai-Murphy, Frank Murphy, and Neil Davies, provided support with obtaining research and collection permits, laboratory space, and logistics.
-Funding provided in part by National Science Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to J.H.N.
-and C.C.D.), Setup Funds from Harvard University to C.C.D., the American Society of Plant Taxonomists (Research Grant for Graduate Students to J.H.N.), the Garden Club of America (Award in Tropical Botany to J.H.N.), the Harvard University Herbaria (Fernald Fieldwork Fellowship to J.H.N.), the Society of Systematic Biologists (Graduate Student Research Award to J.H.N.), and the Systematics Association (Systematics Research Fund to J.H.N.).
+Funding provided in part by National Science Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to JHN and CCD), Setup Funds from Harvard University to CCD, the American Society of Plant Taxonomists (Research Grant for Graduate Students to JHN), the Garden Club of America (Award in Tropical Botany to JHN), the Harvard University Herbaria (Fernald Fieldwork Fellowship to JHN), the Society of Systematic Biologists (Graduate Student Research Award to JHN), and the Systematics Association (Systematics Research Fund to JHN).
 
@@
 
-J.H.N. conceived and carried out the study and analyzed the data. J.H.N. and C.C.D. wrote the draft with editorial input from J.E.W.
+JHN conceived and carried out the study and analyzed the data. JHN and CCD wrote the draft with editorial input from JEW.
 
@@
 
-Additional Supporting Information may be found online in the Supporting Information section at the end of the article:
-
-**`r s_table("species-list")`** `r s_table_cap("species-list")`
-
-**`r s_table("trait-values")`** `r s_table_cap("trait-values")`
-
-**`r s_table("ancova")`** `r s_table_cap("ancova")`
-
-**`r s_table("linear-fit")`** `r s_table_cap("linear-fit")`
-
-**`r s_table("phylosig-binary-strict")`** `r s_table_cap("phylosig-binary-strict")`
-
-**`r s_table("corr-evo-strict")`** `r s_table_cap("corr-evo-strict")`
-
-**`r s_table("gam-fit")`** `r s_table_cap("gam-fit")`
-
-**`r s_table("spatial-sla")`** `r s_table_cap("spatial-sla")`
-
-**`r s_figure("corr-plot")`** `r s_figure_cap("corr-plot")`
-
-**`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`
+Additional Supporting Information may be found online in the Supporting Information section at the end of the article.
+
+**`r s_figure("corr-plot")`** `r s_figure_cap("corr-plot")`.
+
+**`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`.
+
+**`r s_table("species-list")`** `r s_table_cap("species-list")`.
+
+**`r s_table("trait-values")`** `r s_table_cap("trait-values")`.
+
+**`r s_table("ancova")`** `r s_table_cap("ancova")`.
+
+**`r s_table("linear-fit")`** `r s_table_cap("linear-fit")`.
+
+**`r s_table("phylosig-binary-strict")`** `r s_table_cap("phylosig-binary-strict")`.
+
+**`r s_table("corr-evo-strict")`** `r s_table_cap("corr-evo-strict")`.
+
+**`r s_table("gam-fit")`** `r s_table_cap("gam-fit")`.
+
+**`r s_table("spatial-sla")`** `r s_table_cap("spatial-sla")`.
 
@@
 **`r figure("traits-used-image")`** Examples of fern traits used in this study.
-**(a)** Sporophyte traits.
-All sporophyte traits were quantitative, including frond length and width, stipe length, rhizome dia., number of pinnae pairs (in this example, nine) and degree of lamina dissection (in this example, once-pinnate).
-Specific leaf area (ratio of area to mass of leaf lamina) not depicted.
-**(b)** Gametophyte traits.
-Morphotype was defined as a binary trait, either cordate or non-cordate.
-Examples of non-cordate morphotypes include ribbon (thallus elongate and two-dimensional) and filamentous (thallus single lines of cells).
-Other gametophyte traits (gemmae, hairs, and glands) were scored as present or absent.
+(a) Sporophyte traits.
+All sporophyte traits were quantitative, including frond length and width, stipe length, rhizome dia., number of pinnae pairs (in this example, nine), and degree of lamina dissection (in this example, once-pinnate).
+Specific leaf area (ratio of area : mass of leaf lamina) not depicted.
+(b) Gametophyte traits.
+Morphotype was defined as a binary trait, either cordate or noncordate.
+Examples of noncordate morphotypes include ribbon (thallus elongate and two-dimensional; left) and filamentous (thallus single lines of cells; right).
+Other gametophyte traits (hairs, gemmae, and glands) were scored as present or absent.
 Arrows point out instances of each binary trait.
-Scalebars \SI{1}{\milli\meter} except for glands, which is \SI{.1}{\milli\meter}.
+Bars, \SI{1}{\milli\meter} (except for glands, \SI{.1}{\milli\meter}).
 Photographs by J.
@@
 
-**`r figure("climate")`** Selected microclimatic variables of study plots along an elevational gradient from \SI{200}{\meter} to \SI{1200}{\meter} on Moorea, French Polynesia.
+**`r figure("climate")`** Selected microclimatic variables of study plots along an elevational gradient from 200 to \SI{1200}{\meter} on Moorea, French Polynesia.
 Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from `r final_climate_survey$start` to `r final_climate_survey$end`.
-Vapor pressure deficit (VPD) was calculated from relative humidity and temperature as described in the methods, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).
+Vapor pressure deficit (VPD) was calculated from RH and temperature as described in the Materials and Methods section, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD).
 'Epiphytic' dataloggers were placed at \circa{ }\SI{2}{\meter} on trees, and 'terrestrial' dataloggers placed at ground level.
 Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown.
-Trendlines fit using linear models (see Methods); all trendlines significant at \pval{ }< 0.05.
-Asterisks indicate statistical significance of linear model; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01.
-Partially transparent colors indicate outliers excluded from the models (see Methods).
+Trendlines were fitted using linear models (see the Materials and Methods section); all trendlines significant at \pval{ }< 0.05.
+Asterisks indicate statistical significance of linear model; \*\*\*, \pval{ }< 0.001; \*\* = \pval{ }< 0.01.
+Partially transparent colors indicate outliers excluded from the models (see the Materials and Methods section).
 
 **`r figure("pca")`** Principal components analysis (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia.
-**(a, b)** PC loadings.
-**(c, d)** species scores.
-**(a)** and **(c)** are standard PCA, **(b)** and **(d)** are phylogenetic PCA.
-Epiphytes in green, terrestrial species in brown.
+(a, b) PC loadings; (c, d) species scores.
+(a) and (c) are standard PCA, and (b) and (d) are phylogenetic PCA.
+Epiphytes are in green, and terrestrial species in brown.
 Quantitative traits only, including only species with no missing observations for any traits (\n{ }= `r pca_results[["species_locs"]] %>% pull(species) %>% n_distinct` \species).
-Trait abbreviations as follows: dissection, degree of frond dissection; length, frond length; width, frond width; pinna, number of pinnae pairs; rhizome, rhizome diameter; stipe, stipe length; SLA, specific leaf area.
-
-**`r figure("tree")`** Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia with growth habit and associated traits mapped on the tips.
+Trait abbreviations: dissection, degree of frond dissection; length, frond length; width, frond width; pinna, number of pinnae pairs; rhizome, rhizome diameter; stipe, stipe length; SLA, specific leaf area.
+
+**`r figure("tree")`** Time-calibrated phylogenetic tree of ferns from Moorea, French Polynesia, with growth habit and associated traits mapped on the tips.
 Tree obtained from @Nitta2017.
-Relative value of quantitative (sporophyte) traits shown by heat gradient: low values are lighter, high values are darker.
+Relative value of quantitative (sporophyte) traits shown by heat gradient: lower values are lighter, higher values are darker.
 Stipe length and rhizome diameter were log-transformed prior to scaling.
-States of qualitative (gametophyte) states and growth habit indicated by colors in key.
-Grey indicates missing data or non-applicable trait states.
-Values for three leaf size traits (stipe length, frond length, and frond with) were correlated, so of these, we only present stipe length.
+States of qualitative (gametophyte) traits and growth habit are indicated by colors in the legend.
+Gray indicates missing data or nonapplicable trait states.
+Values for three leaf size traits (stipe length, frond length, and frond width) were correlated, so of these, we only present stipe length.
 Species missing data for six or more traits not shown.
@@
 Scale in millions of years (my).
-
-**`r figure("heatmap")`** Variable importance scores from a full-subsets analysis using general additive models exploring the effect of climate and growth habit (terrestrial vs.
-epiphytic growth) on community diversity of ferns on Moorea, French Polynesia.
+SLA, specific leaf area.
+
+**`r figure("heatmap")`** Variable importance scores from a full-subsets analysis using general additive models exploring the effect of climate and growth habit (terrestrial vs epiphytic growth) on community diversity of ferns on Moorea, French Polynesia.
 Predictor variable abbreviations: VPD, vapor pressure deficit; SD, standard deviation.
@@
 Response variable abbreviations: SLA, specific leaf area.
-Epiphytic communities in green, terrestrial communities in brown.
-Error bars are each one standard deviation (values below zero not shown).
-Trendlines and \rval^2^ shown for significant relationships as determined by a linear model at \pval{ }< 0.05 (\rval^2^ not shown for SLA, which was fit with a spatial generalized linear mixed model; see Methods).
-\tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\* = \pval{ }< 0.001.
+Epiphytic communities are in green, and terrestrial communities in brown.
+Error bars are each one standard deviation (values < 0 are not shown).
+Trendlines and \rval^2^ shown for significant relationships as determined by a linear model at \pval{ }< 0.05 (\rval^2^ not shown for SLA, which was fit with a spatial generalized linear mixed model; see the Materials and Methods section).
+\tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\*, \pval{ }< 0.001.
 Results for frond width were very similar to frond length and not shown.
@@
 **`r figure("comm-div")`** Functional and phylogenetic diversity of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia.
-Epiphytic communities in green, terrestrial communities in brown.
-Response variable abbreviations as follows: MPD~phy~, standard effect size (SES) of mean phylogenetic distance; MNTD~phy~, SES of mean nearest taxon distance; MPD~func~, SES of mean functional distance; MNTD~func~, SES of mean nearest functional distance.
+Epiphytic communities are in green, and terrestrial communities in brown.
+Response variable abbreviations: MPD~phy~, standard effect size (SES) of mean phylogenetic distance; MNTD~phy~, SES of mean nearest taxon distance; MPD~func~, SES of mean functional distance; MNTD~func~, SES of mean nearest functional distance.
 Trendlines and \rval^2^ shown for significant relationships as determined by a linear model at \pval{ }< 0.05.
@@
 Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range.
-Asterisks indicate statistical significance of two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01.
-
-\clearpage
-
-**Table `r table_nums("traits-used", display = "n")`** Fern traits used in this study
+Asterisks indicate statistical significance of two-sided \tval-test; \*\*\*, \pval{ }< 0.001; \*\*, \pval{ }< 0.01.
+
+\clearpage
+
+**Table `r table_nums("traits-used", display = "n")`** Fern traits used in this study.
 
@@
 read_csv(here::here("data/table_1.csv"), col_types = "cccc") %>%
+  mutate(trait = str_to_sentence(trait)) %>%
   rename(`Data type (unit)` = data_type, `Functional significance` = ecological_significance, `Reference(s)` = reference) %>%
@@
 
-**Table `r table_nums("phylosig-cont", display = "n")`** Phylogenetic signal in continuous traits of ferns from Moorea, French Polynesia
+**Table `r table_nums("phylosig-cont", display = "n")`** Phylogenetic signal in quantitative (sporophyte) traits of ferns from Moorea, French Polynesia.
 
@@
     trait,
-    new_levels = c("Stipe length", "Frond length", "Frond width", "Rhizome diam.", 
+    new_levels = c("Stipe length", "Frond length", "Frond width", "Rhizome diameter", 
                    "SLA", "Pinna no.", "Frond dissection"))) %>%
@@
 \K, Blomberg's \K{ }[@Blomberg2003]; \plam, Pagel's \plam{ }[@Pagel1999].
-Values of \plam or \K greater than 1 indicate traits evolving according to Brownian Motion (BM); values of \plam or \K near 0 indicate random distribution of traits.
-
-\clearpage
-
-**Table `r table_nums("phylosig-binary", display = "n")`** Phylogenetic signal in binary traits of ferns from Moorea, French Polynesia
+Values of \plam or \K near 1 indicate traits evolving according to Brownian Motion (BM); values of \plam or \K near 0 indicate random distribution of traits.
+SLA, specific leaf area.
+
+\clearpage
+
+**Table `r table_nums("phylosig-binary", display = "n")`** Phylogenetic signal in binary traits of ferns from Moorea, French Polynesia.
 
@@
 
-\D, Fritz and Purvis' \D{ }[@Fritz2010]; *P*(rnd), probability of random distribution of traits; *P*(BM), probability of traits evolving under Brownian motion (BM).
-Values of \D{ }0 or less indicate evolution according to BM; values near 1 indicate random trait distribution.
-
-\clearpage
-
-**Table `r table_nums("pics", display = "n")`** Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth in ferns from Moorea, French Polynesia
+\D, Fritz and Purvis's \D{ }[@Fritz2010]; *P*(rnd), probability of random distribution of traits; *P*(BM), probability of traits evolving under Brownian motion (BM).
+Values of \D{ }may range from may range from < 0 (clumped), to 0 (evolving according to BM), to 1 (random), to > 1 (overdispersed).
+
+\clearpage
+
+**Table `r table_nums("pics", display = "n")`** Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth in ferns from Moorea, French Polynesia.
 
@@
   cols_label(
-    n_contrasts = "Num. constrasts",
-    n_pos_con = "Num. pos. constrasts",
+    n_contrasts = "Number of constrasts",
+    n_pos_con = "Number of positive constrasts",
     statistic = md("*t*"),
@@
 
-All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades.
-
-\clearpage
-
-**Table `r table_nums("corr-evo", display = "n")`** Pagel's (1994) test of correlated evolution between binary traits in ferns from Moorea, French Polynesia
+All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades. SLA, specific leaf area.
+
+\clearpage
+
+**Table `r table_nums("corr-evo", display = "n")`** Pagel's (1994) test of correlated evolution between binary traits in ferns from Moorea, French Polynesia.
 
@@
   cols_label(
-    logL_indep = "LogL(independent)",
-    logL_dep = "LogL(dependent)",
+    logL_indep = "Log L(independent)",
+    logL_dep = "Log L(dependent)",
     likelihood_ratio = "Likelihood ratio",
@@
 
-LogL, log-likelihood.
-Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs.
-terrestrial growth).
-Traits with a significantly higher (\ie, less negative) log likelihood for the dependent model are correlated with growth habit.
+Log *L*, log-likelihood.
+The presence or absence of each gametophyte trait was tested for correlated evolution with growth habit (epiphytic vs terrestrial growth).
+Traits with a significantly higher (\ie less negative) log likelihood for the dependent model are correlated with growth habit.
 
